### 2020-02-02 battery-preprocessing_v3_DB.Rmd
Version of analysis file using functions. Modified outlier method to retain code for outliers, and to combine outlier detection for all data (both sides). This revealed some subjects with odd patterns that look like multiple responses. Added method for excluding these. Added exploratory t statistic for laterality, which looks promising for reliability of RDT.

---
title: "Online Laterality Battery Pre-Processing - short version"
author: "Adam Parker + Dorothy Bishop"
date: "1st Feb 2020"
output: html_document
---

This R Markdown file prcoesses the various components of the Online Laterality Battery which is implemented online via Gorilla.sc.

This version updated on 1st Feb 2020 to make script more reproducible and compact by using functions for repeated operations.

The script pulls in csv files and processes them so that pre-registered analyses can be conducted (pre-registered on the OSF: ). First, questionnaires are analyses followed by visual half field tasks. This is done for time 1 and time 2. A "full" laterality index output is then appended to a file with the demographics. This is then used in subsequent R markdown scripts.

This script analyses the following:

1) Questionnaires
  - Demographics
  - Miles test
  - Porta test
  - Edinburgh Handedness Inventory
  - lexTALE
2) Chimeric face task
3) Rhyme detection task
4) Key Pressing Task (finger dexterity)

# Questionnaires

This R Markdown processes the following questionnaires from the Online battery:

- Demographics
- Miles test
- Porta test
- Edinburgh Handedness Inventory
- lexTALE

Once processed, a file is made.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# libraries 
library(readr)
library(yarrr)
library(dplyr)
library(tidyr)
library(car)
library(ggplot2)
library(knitr)
require(readr)  # for read_csv()
require(dplyr)  # for mutate()
require(tidyr)  # for unnest()
require(purrr)  # for map(), reduce()
library(GGally)
require(stringr) # string matching
```

First, the demographics is processed.This gives us a participant ID, age, gender, footedness, and handedness.

```{r demographics, warning=FALSE}
demo_dat_R <- read_csv("./dat2/demographics_R.csv", na = c("empty", "NA")) # read in data
#demo_dat_L <- read_csv("./dat2/demographics_L.csv", na = c("empty", "NA")) # read in data
demo_dat <- rbind(demo_dat_R)
# relabel 
demo_dat$subject <- demo_dat$`Participant Public ID`
demo_dat$subject <- as.factor(demo_dat$subject)
# count and print subjects
nsub <-length(levels(demo_dat$subject))
print(paste0("subjects: ", nsub))

# create data frame
demographics <- data.frame(matrix(ncol = 5, nrow = nsub))
  colnames(demographics) <- c("subject", "age", "gender", "footedness", "handedness")
myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  subname <- levels(demo_dat$subject)[i] # find subject ID
  myrows <- which(demo_dat$subject==subname) # select rows for this subject
  tmp <- data.frame(demo_dat[myrows,])

    myrow <- myrow+1

    demographics$subject[myrow] <- subname
    demographics$age[myrow] <- tmp[tmp$Question.Key == "response-1",]$Response
    demographics$gender[myrow] <- tmp[tmp$Question.Key == "Gender",]$Response
    demographics$footedness[myrow] <- tmp[tmp$Question.Key == "alcohol",]$Response
    demographics$handedness[myrow] <- tmp[tmp$Question.Key == "categorical_hand",]$Response
  }
```

Now we import and append the Porta and Miles Test to the demographics data frame.

```{r sightednesss, warning=FALSE}
miles_dat_R <- read_csv("./dat2/Miles_R.csv", na = c("empty", "NA")) # read in data
#miles_dat_L <- read_csv("./dat2/Miles_L.csv", na = c("empty", "NA")) # read in data
miles_dat <- rbind(miles_dat_R)
porta_dat_R <- read_csv("./dat2/Porta_R.csv", na = c("empty", "NA")) # read in data
#porta_dat_L <- read_csv("./dat2/Porta_L.csv", na = c("empty", "NA")) # read in data
porta_dat <- rbind(porta_dat_R)
# add test ID
miles_dat$test_ID <- "miles"
porta_dat$test_ID <- "porta"
# merge the two
sight_dat <- rbind(miles_dat, porta_dat)
# relabel 
sight_dat$subject <- sight_dat$`Participant Public ID`
sight_dat$subject <- as.factor(sight_dat$subject)

# count and print subjects
nsub <-length(levels(sight_dat$subject))
print(paste0("subjects: ", nsub))

# create data frame
sightedness <- data.frame(matrix(ncol = 3, nrow = nsub))
  colnames(sightedness) <- c("subject", "miles", "porta")
myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  subname <- levels(sight_dat$subject)[i] # find subject ID
  myrows <- which(sight_dat$subject==subname) # select rows for this subject
  tmp <- data.frame(sight_dat[myrows,])

    myrow <- myrow+1

    sightedness$subject[myrow] <- subname
    sightedness$miles[myrow] <- tmp[tmp$Question.Key == "response-2",]$Response
    sightedness$porta[myrow] <- tmp[tmp$Question.Key == "response-2",]$Response
  }

# now merge sightedness and demographics
demographics <- merge(demographics, sightedness, by= "subject")
```

Now we add the EHI. Here we score and plot the Edinburgh handedness inventory (Oldfield, 1971) in its short form.

```{r EHI, warning=FALSE}
EHI_R <- read_csv("./dat2/EHI_R.csv", na = "empty") # read in data
#EHI_L <- read_csv("./dat2/EHI_L.csv", na = "empty") # read in data
EHI <- rbind(EHI_R)

# select only responses
EHI <- EHI[EHI$`Question Key` == "response-2-quantised" | EHI$`Question Key` == "response-3-quantised" | 
           EHI$`Question Key` == "response-4-quantised" | EHI$`Question Key` == "response-5-quantised" |
           EHI$`Question Key` == "response-6-quantised" | EHI$`Question Key` == "response-7-quantised" | 
           EHI$`Question Key` == "response-8-quantised" | EHI$`Question Key` == "response-9-quantised" |
           EHI$`Question Key` == "response-10-quantised" | EHI$`Question Key` == "response-11-quantised" | 
           EHI$`Question Key` == "response-12-quantised" | EHI$`Question Key` == "response-13-quantised" |
           EHI$`Question Key` == "response-14-quantised" | EHI$`Question Key` == "response-15-quantised" | 
           EHI$`Question Key` == "response-16-quantised" | EHI$`Question Key` == "response-17-quantised" |
           EHI$`Question Key` == "response-18-quantised" | EHI$`Question Key` == "response-19-quantised" | 
           EHI$`Question Key` == "response-20-quantised",]

# recode variabels
EHI$subject <- EHI$`Participant Public ID`
EHI$subject <- as.factor(EHI$subject)

# reduce data
# new data
meaningful <- c("subject", "Question Key", "Response") # select wanted columns
c<-which(names(EHI) %in% meaningful) #find colnumbers of unwanted
EHI <-EHI[,c] #remove unwanted columns
EHI <- na.omit(EHI) # remove NAs

# score responses accoridng to the original scoring method in Oldifeld (1971)
EHI$right_hand <- 0
EHI$left_hand <- 0
for (r in 1:nrow(EHI)) {
  if(EHI$Response[r] == 1){
    EHI$right_hand[r]= 2 
  } else {
    if(EHI$Response[r] == 2){
      EHI$right_hand[r]= 1
    } else {
      if(EHI$Response[r] == 4){
        EHI$left_hand[r]= 1
      } else {
        if(EHI$Response[r] == 5){
          EHI$left_hand[r]= 2 
        } else {
          (EHI$right_hand[r]= 1) & (EHI$left_hand[r]= 1)
        }
      }
    }
  }
}

# re-reduce data
# new data
meaningful <- c("subject", "right_hand", "left_hand") # select wanted columns
c<-which(names(EHI) %in% meaningful) #find colnumbers of unwanted
EHI <-EHI[,c] #remove unwanted columns
EHI <- na.omit(EHI) # remove NAs

# find sum for left and right for each participant 
# create hand data 
nsub <-length(levels(EHI$subject))
EHI2 <- data.frame(matrix(ncol = 3, nrow = nsub))
  colnames(EHI2) <- c("subject", "right_hand", "left_hand")

myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  myrow <- myrow+1
  subname <- levels(EHI$subject)[i] # find subject ID
  myrows <- which(EHI$subject==subname) # select rows for this subject
  tmp <- data.frame(EHI[myrows,])

    EHI2$subject[myrow] <- subname # add row for subject
    EHI2$right_hand[myrow] <- sum(tmp$right_hand) # add row for hand
    EHI2$left_hand[myrow] <- sum(tmp$left_hand) # add row for number of buttons pressed
  }

# This is a summary of the EHI
for_index2 <- 
  EHI2 %>% 
  group_by(subject) %>%
  mutate(index_EHI= ((right_hand-left_hand)/(right_hand + left_hand))*100)

# now create a data frame only for index 
EHI_index <- select(for_index2, "subject", "index_EHI")
# merge which demographics
demographics <- merge(demographics, EHI_index, by= "subject")

#now plot EHI in relation to handedness (modified by DB)
x<-ggplot(demographics, aes(x=index_EHI, y=subject, fill=subject)) +
  geom_dotplot(binaxis='y', stackdir='center') + xlim(-100, 100) +
  ggtitle("Handedness index on the Edinburgh Handedness Inventory")
x + theme_bw() + theme(legend.position = "none")

```

Next, the lexTale is preprocesssed. This is done using the weighted average for the test.

```{r lexTALE, warning=FALSE}
# read data
#lexTALE_L <- read_csv("./dat2/lexTale_L.csv", na = "empty")
lexTALE_R <- read_csv("./dat2/lexTale_R.csv", na = "empty")
lexTALE <- rbind(lexTALE_R)
# filter
lexTALE <- lexTALE[lexTALE$Attempt == 1,]
lexTALE <- lexTALE[lexTALE$display == "Task",]
# rename variables
lexTALE$subject <- lexTALE$`Participant Public ID`
lexTALE$accuracy <- lexTALE$Correct
lexTALE$condition <- lexTALE$ANSWER
# code as factors
lexTALE$subject <- as.factor(lexTALE$subject)
lexTALE$condition <- as.factor(lexTALE$condition)
  levels(lexTALE$condition) <- c("non-word", "word")

# create data frame
lexTale_score <- data.frame(matrix(ncol = 4, nrow = nsub))
  colnames(lexTale_score) <- c("subject", "Word", "nonWord", "lexTALE")
myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  subname <- levels(lexTALE$subject)[i] # find subject ID
  myrows <- which(lexTALE$subject==subname) # select rows for this subject
  tmp <- data.frame(lexTALE[myrows,])

    myrow <- myrow+1

    lexTale_score$subject[myrow] <- subname
    lexTale_score$Word[myrow] <- sum(tmp[tmp$condition == "word",]$accuracy)
    lexTale_score$nonWord[myrow] <- sum(tmp[tmp$condition == "non-word",]$accuracy)
    lexTale_score$lexTALE[myrow] <- ((sum(tmp[tmp$condition == "word",]$accuracy)/40*100) + (sum(tmp[tmp$condition == "non-word",]$accuracy)/20*100)) / 2

}

# plot
x<-ggplot(lexTale_score, aes(x=lexTALE, y=subject, fill=subject)) +
  geom_dotplot(binaxis='y', stackdir='center') + xlim(0, 100) +
  ggtitle("Lextale")
x + theme_bw() + theme(legend.position = "none")

# now create a data frame only for index 
lexTALE <- select(lexTale_score, "subject", "lexTALE")
# now merge with demographics
demographics <- merge(demographics, lexTALE, by= "subject")
# now write the full demographics 
write.csv(demographics, file = "./dat2/demographics_out.csv")

```

# Read in raw data for day 1 and 2

Before doing anything with the data, we now pull in all the data from all nodes and combind them from a sub folder in the directory. Participants, identified by random letter strings, are stacked on top of each other. We start by reading the data and cutting out unwanted information.

```{r readraw,include= FALSE, warning=FALSE}
# Set directory
for (d in 1:2){
  data_file <- "./dat2/day1_combined.csv"

  if (d ==2){
    data_file <-  "./dat2/day2_combined.csv"
  }
  #we will assign raw data to all_dat2 (day2), but then reassign to all_dat1 on first run through loop, so we have correct data for each day
all_dat2 <- read.csv(data_file, na.strings = "NA",stringsAsFactors=F)
# cleaning - do operations that apply for all tasks
all_dat2 %>% filter(Attempt >= 1,display=="e_Task") # only answers 
# write as factors
all_dat2$subject <- as.factor(all_dat2$Participant.Public.ID)
all_dat2$RT <- as.numeric(as.character(all_dat2$Reaction.Time))
all_dat2$Task.Name <- as.factor(all_dat2$Task.Name)
all_dat2$word<-substring(all_dat2$word,66,(nchar(all_dat2$word)-4)) #remove format from written word stimuli
if (d==1)
{all_dat1 <- all_dat2}
}
#check that colnames are the same
colcomp<-data.frame(colnames(all_dat1)[order(colnames(all_dat1))])
colnames(colcomp)[1]<-'name1'
ncol2 <-length(colnames(all_dat2))
colcomp$name2<-NA
colcomp$name2[1:ncol2]<-colnames(all_dat2)[order(colnames(all_dat2))]

#NB the colnames are identical but in different order, and all_dat1 has extra column of word_OG
#It is just conceivable the ordering might affect processing when columns are subsequently selected, so we'll reorder them to be consistent across sessions

all_dat2$word_OG<-NA #create dummy column just to make it easier to align columns
collist <- colnames(all_dat1)
all_dat2 <- all_dat2[,collist]

#Check for overlap in subjects between the two sessions
which(levels(all_dat1$subject) %in% levels(all_dat2$subject))

```

## Chimeric Face Task

This part of the R Markdown file prcoesses the "Chimeric Face Task' implmented online via Gorilla.sc which aims to assess the lateralisation of emotion recognition in a face processing task. In the task, participants are shown a chimeric face. Participants then respond by indicating the emotion that they felt was most strong expressed.

This task is near identical to that reported in Karlsson, Johnstone, and Carey (2019). Stimuli were kindly provided by Dr Michael Burt (https://www.dur.ac.uk/psychology/staff/?id=1942) and were symmetrical average images created from four male and four female faces (see Burt & Perrett, 1997; Innes et al., 2016).

We start by reading combined raw data and cutting out unwanted information.

```{r genericprocessing} 
#This is a function that can apply to all tasks; user specifies the original file
#to start from, the wanted columns, and the specific task to select
# This function just returns a subfile with just the necessary rows/cols
  procdata <- function(origfile,wanted,task){
  nufile <- origfile[origfile$Task.Name==task,] #restrict consideration to this task

  c<-which(names(origfile) %in% wanted) #find colnumbers of wanted
  #NB beware that the columns may not be in the same order as in the wanted list
  nufile <- nufile[,c]#select only wanted columns
  nufile <- na.omit(nufile) # remove NAs
  return(nufile)
}

```

The Hoaglin-Iglewicz function is used to remove outliers. We create an 'outlier' column that codes anticipations as 2, and long RT outliers as 1. All else is zero. The original RT is saved as allRT, and the RT column has NA for outliers, so they will be automatically excluded henceforth.

```{r RToutlierfunction}
remove.outliers <- function(origdat,lowcut,zcut){
# Outlier removal put in separate function, as needed to reuse when considering t-test
# Hoaglin, D. C., & Iglewicz, B. (1987). Fine-tuning some resistant rules for outlier labelling. Journal of the American Statistical Association, 82(400), 1147-1149.

# origdat is the dataframe with data for outlier removal, mycolname is name of variable 
# for outlier removal (usually RT), lowcut is a cutoff value below which data excluded
# (typically v fast responses that are implausible, e.g. 201 ms for RT)
# and zcut is the constant in Hoaglin-Iglewicz method (usually 2.2)

# NB this function returns origdat with outlier rows marked, rather than removed
# This makes it easier to record the N outliers per subject.

# NB FUNCTION ASSUMES THAT THERE IS ARE COLUMNS CALLED subject, Correct and RT IN origfile.
# Outliers are computed by subject - NB contrary to previous version, these are defined
# across all trials, rather than separately for L and R.
origdat$allRT <- origdat$RT #we save a copy of RT, as we will be altering RT so that outliers become NA
origdat$outlier<-0
origdat$outlier[origdat$RT<lowcut] <- 2 #outlier col is coded 2 for anticipations
origdat$RT[origdat$RT<lowcut]<-NA #we exclude anticipations when computing quantiles
#Now extract correct RTs for each subject to compute limits for outliers

  for (i in 1:nsub) { # loop through subjects
  subname <- levels(origdat$subject)[i] # find subject ID
  myrows <- which(origdat$subject==subname) # select rows for this subject
  #NB myrows is NOT a consecutive series, because done in blocks
  firstrow <-min(myrows)
  tmp <- data.frame(origdat[myrows,])

  RTcorr <-tmp$RT[tmp$Correct==1] #correct RTs for this subject as a vector

 # Identify 25th and 75th quartiles, and the difference between them
          lower_quartile <- quantile(RTcorr, probs=0.25, na.rm="TRUE")
          upper_quartile <- quantile(RTcorr, probs=0.75, na.rm="TRUE")
          quartile_diff <- upper_quartile - lower_quartile
        # Outliers are defined as being below or above 2.2 times the quartile difference
          lower_limit <- lower_quartile - zcut*quartile_diff
          upper_limit <- upper_quartile + zcut*quartile_diff
        # create outlier variable
w1 <- which(origdat$RT[myrows]>upper_limit) #rows with outlier, nb RELATIVE to myrows range
origdat$outlier[myrows[w1]]<-1
}
 #For RT we just remove slow outliers, so ignore lower_limit
origdat$RT[origdat$outlier>0]<-NA #RT variable now has outliers excluded as NA
return(origdat)
}
```

```{r read_face_dat, warning=FALSE}
for (d in 1:2){ #we will process day 1 and day 2 together
  origfile <- all_dat1
  if (d==2)
  {origfile <- all_dat2}
#As with earlier reading in, we default to saving CF_dat2 (day2), but at end of the chunk we reassign to CF_day1 in the first run through the d loop
task <- "Chimeric faces (Karlsson, Johnstone, & Carey, 2019)"
wanted <- c("subject", "stimuli", "Response", "RT", "l_hemiface", "r_hemiface") # select wanted columns
CF_dat2 <- procdata(origfile,wanted,task)

# count subjects
nsub <-length(levels(CF_dat2$subject))
print(paste('day',d,task,": subjects: ", nsub))

CF_same2 <- CF_dat2[CF_dat2$l_hemiface == CF_dat2$r_hemiface,] # create data frame for non-chimeras
CF_dat2 <- CF_dat2[CF_dat2$l_hemiface != CF_dat2$r_hemiface,] # remove items where same expression in both hemifaces
CF_dat2 <- remove.outliers(CF_dat2,200,2.2)
outliertable<-table(CF_dat2$subject,CF_dat2$outlier)
print(outliertable)
if (d==1)
{CF_dat1 <- CF_dat2
 CF_same1 <- CF_same2}
}
```

First, we judge participants' ability to correctly identify emotions using the "CF_same" dataframe. Participants are removed if their performance is below 75% due to poor ability to identify emotions that will inevitably influence performance on the Chimeric Face Task.

```{r emotions, warning=FALSE,message=FALSE}
# this will mark correct responses as 1 or 0. It will be importantfor removing participants later on. 
# NB we loop through same process for day1 and day2
for (d in 1:2){
  CF_same <- CF_same1
  if (d==2){
    CF_same <- CF_same2
  }

CF_same <- 
  CF_same %>% 
  mutate(emotion_recog= ifelse(as.character(Response) == as.character(l_hemiface), 1, 0))
# calculate participants average
emotion_mean <- aggregate(FUN= mean, data= CF_same, emotion_recog~ subject)
# add this to demographics 
demographics <- merge(demographics, emotion_mean, by= "subject")
}
mycol <- length(colnames(demographics))
colnames(demographics)[(mycol-1):mycol]<-c('Emot.recog1','Emot.recog2')
```

In this section, the correct sequences are marked as 1, incorrect asnwers are marked as 0. We then mark whether the correct answers were in the left or right hemiface.

```{r mark_face_correct_face, warning=FALSE}
# this will mark correct responses as 1 or 0.
# NB we again loop through same process for day1 and day2
for (d in 1:2){
  CF_dat <- CF_dat1
  if (d==2){
    CF_dat <- CF_dat2
  }
CF_dat <- 
  CF_dat %>% 
  mutate(Correct= ifelse(as.character(Response) == as.character(l_hemiface), 1, 
                         ifelse(as.character(Response) == as.character(r_hemiface), 1, 0)))
# this will mark the side responded to. NB to allow for generic functions for all tasks, we use the label 'side' rather than hemiface
CF_dat <- 
  CF_dat %>% 
  mutate(side= ifelse(as.character(Response) == as.character(l_hemiface) & Correct == 1, "Left",
                       ifelse(as.character(Response) == as.character(r_hemiface) & Correct == 1, "Right", NA)))
# code side as factor
CF_dat$side <- as.factor(CF_dat$side)

  if(d==1){
    CF_dat1 <- CF_dat}
  if (d==2){
    CF_dat2 <- CF_dat
  }
}
```

Now the data is ready, create an empty data frame and populate this. It includes columns for the participant identifier, side, and count of accurate responses for each side.

```{r maketaskdataframe, warning=FALSE}
#This is a generic chunk of code that can be used for Chimeric faces and Rhyme Detection

make.df <- function(origdat,varlist,latlist){
# create compact data frame 
temp_dat <- data.frame(matrix(ncol = 4, nrow = nsub*2))
colnames(temp_dat) <- varlist
nsub<-length(unique(origdat$subject))
myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  subname <- levels(origdat$subject)[i] # find subject ID
  myrows <- which(origdat$subject==subname) # select rows for this subject
  latcol <- which(colnames(origdat) == varlist[2])
  RTcol <- which(colnames(origdat) == varlist[4])
  tmp <- data.frame(origdat[myrows,])

  for (j in 1:2) {
    myrow <- myrow+1
    w<- which(tmp[,latcol]== latlist[j]) # match on laterality
    tmp1 <- tmp[w,]
    tmp2 <- tmp1[tmp1$Correct == 1,]

    temp_dat$subject[myrow] <- subname # add row for subject
    temp_dat$side[myrow] <- latlist[j] # add row for side
    temp_dat$accurate[myrow] <- 100*mean(tmp1$Correct,na.rm=TRUE) # add row for accurate responses - this has been modified to give % correct
    temp_dat$RT[myrow] <- mean(tmp2$RT,na.rm=TRUE) # add row for mean RT
    temp_dat$N[myrow]<-length(tmp1$RT) # add count of rows for this subject
  }
}
return(temp_dat)
}
```

```{r makedataframe, warning=F}
#We create chim_dat files for day1 and day 2
varlist <-   c("subject", "side", "accurate", "RT") #need to be in this order, ie sub, side, acc and RR
origdat <- CF_dat1
latlist <- c("Left","Right") #names of factor levels
chim_dat1 <- make.df(origdat,varlist,latlist)
origdat <- CF_dat2
chim_dat2 <- make.df(origdat,varlist,latlist)

w<-which(is.na(chim_dat1$subject))
if (length(w)>0) {chim_dat1<-chim_dat1[-w,]}
w<-which(is.na(chim_dat2$subject))
if (length(w)>0) {chim_dat2<-chim_dat2[-w,]}
```

The data is visualised using the pirate plot function and t.test conducted.

```{r plotdata_ttest, warning=FALSE}
#Same function can be used for various tasks - we specify relevant data/task in the function call by assigning origdat etc to relevant values
dopirate <- function(origdat,task,measure){
# measure for each side
# plot the number of correct responses in each side
pirateplot(formula = X1 ~ side,
           data = origdat,
           main = measure,
           theme = 0,
           pal = "southpark", # southpark color palette
           bean.f.o = .0, # Bean fill
           point.o = .3, # Points
           inf.f.o = .7, # Inference fill
           inf.b.o = .8, # Inference border
           avg.line.o = 1, # Average line
           bar.f.o = .5, # Bar
           point.pch = 21,
           point.cex = .7,
           inf.method= "ci")
# plot L & R to get slope
q <- ggplot(data=origdat, aes(x=side, y=X1, group=subject, color=subject)) +
    geom_line() +
    geom_point() +
    ggtitle(paste0(task,": ",measure," by side"))
q + theme_bw() + theme(legend.position = "none")
#  t test
print(paste0(task,measure))
myt1<-t.test(origdat$X1~ origdat$side, paired = TRUE, alternative = "two.sided")
print(myt1)

}
```

```{r dosummary}
origdat <- chim_dat1
measure <- 'Accuracy'
origdat$X1 <- chim_dat1$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Chimeric faces day 1'

dopirate(origdat,task,measure)
origdat$X1 <- chim_dat1$RT
measure <- 'RT'
dopirate(origdat,task,measure)

#Now repeat for day 2
origdat <- chim_dat2
origdat$X1 <- chim_dat2$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Chimeric faces day 2'

dopirate(origdat,task,measure)
origdat$X1 <- chim_dat2$RT
measure <- 'RT'
dopirate(origdat,task,measure)

```

Finally, a laterality index is calculated and plotted based on the equation [(R-L)/(R+L)] x 100, where L= left side and R= right side.

If the index is negative, emotion recognition is lateralised in the left hemisphere. If the index is positive, emotion recognition is lateralised in the right hemisphere

Here, a .csv is written for the chimeric face LI.

```{r makeLIs, warning=FALSE}
# Another generic function that can be used for different tasks to make 
# a laterality index and write it to the demographics file
# This function modifies a copy of the demographics file and returns it

# User selects whether accuracy or RT by specifying col X1 in origdat when calling the function

# Polarity is -1 or 1 and can be set to keep all LIs positive if in a given direction
# Cutoff will create NA if level of performance is outside predetermined limits

makeLI <- function(demogs,origdat,task,mycolname,polarity,mycutoff){
# we do this for accuracy first
# reformat data frame
keeps <- c("subject", "side", "X1") #X1 holds either accuracy or RT
LIdf <- origdat[keeps]
LIdf <- spread(LIdf, side, X1)
# calculate each participants' lat index
LIdf <- 
  LIdf %>% 
  group_by(subject) %>%
  mutate(thisLI= ((Right - Left)/(Right + Left))*100)

LIdf$LI <- LIdf$thisLI*polarity #polarity can be set to reverse sign - eg for RT a large score is bad, whereas for acc it is good, so can have opp polarity

# check if accuracy or RT level is outside acceptable cutoff.
# Do this based on average for left and right
# LI will be set to NA for those outside limits
LIdf$avg <- (LIdf$Left+LIdf$Right)/2
w<-which(LIdf$avg<mycutoff)
if(length(w)>0){
LIdf$LI[w] <- NA}

# now create a data frame for lat index 
LIdf <- select(LIdf, "subject", "LI")
# now merge with demographics
demogs <- merge(demogs, LIdf, by= "subject")
lastcol<-length(colnames(demogs))
colnames(demogs)[lastcol]<-mycolname

return(demogs)
}
```

```{r domakeLI}
#We add Chimeric faces LIs to demographics for day 1 and then day2
task<-'Chimeric faces'
origdat<-chim_dat1
origdat$X1 <- origdat$accurate
mycolname<-"Day1_CF_acc_LI"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (-1)#so LI is positive in predicted direction
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

# now do RT
origdat$X1 <- origdat$RT
mycolname<-"Day1_CF_RT_LI"
polarity <-1

#if we don't specify mycutoff, it will remain at level set for accuracy, which means it should not trap any RT values as they will all be much bigger
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

origdat<-chim_dat2
origdat$X1 <- origdat$accurate
mycolname<-"Day2_CF_acc_LI"
polarity<- (-1)#so LI is positive in predicted direction
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
origdat$X1 <- origdat$RT
mycolname<-"Day2_CF_RT_LI"
polarity <-1
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

cor1 <- cor(demographics$Day1_CF_acc_LI,demographics$Day2_CF_acc_LI)
plot(demographics$Day1_CF_acc_LI,demographics$Day2_CF_acc_LI,main='Chimeric LI accuracy')
abline(v=0)
abline(h=0)
text(-10,15,paste0('r = ',round(cor1,3)))
cor2<-cor(demographics$Day1_CF_RT_LI,demographics$Day2_CF_RT_LI)
plot(demographics$Day1_CF_RT_LI,demographics$Day2_CF_RT_LI,main='Chimeric LI RT')
abline(v=0)
abline(h=0)
text(-5,15,paste0('r = ',round(cor2,3)))

```
As we can see, we generally get a left side/right hemisphere advantage in our accuracy data. The RT advantage is not as expected. This may be due to the buttons being clicked on screen and the additional time penalty of moving the mouse.

## Rhyme Detection Task

This portion of the R Markdown file prcoesses the "Rhyme Detection Task' implmented online via Gorilla.sc which aims to assess the lateralisation of language processing. In the task, participants are shown a word in the central visual field and different images in the left and right visual field. An arrow appears pointing to either image. Participants then respond by indicating whether the image and word rhyme.

This reads in a csv file where each row represents a response to a trial. Participants, identified by random letter strings, are stacked on top of each other. We start by reading the data and cutting out unwanted information.

```{r read_dat_RDT, warning=FALSE}
for (d in 1:2){ #same script for days 1 and 2

origfile <- all_dat1
if(d==2)
  {origfile <- all_dat2}
task <- "Rhyme Detection Task"

# nb when creating new file, we default to name 'RDT_dat2', but this is reassigned to 'RDT_dat1' at end of processing for day 1.

# new data
wanted <- c("subject", "Correct", "RT", "ANSWER","Response") # select wanted columns
RDT_dat2 <- procdata(origfile,wanted,task) #Here we reuse the procdata function that we developed for processing Chimeric faces data - just does some generic processing

# do some wrangling on the ANSWER variable to make it compatible with functions
RDT_dat2<-filter(RDT_dat2,ANSWER %in% c('left','right'))
c <- which(colnames(RDT_dat2)=="ANSWER")
colnames(RDT_dat2)[c]<-'side'
RDT_dat2$side<-as.factor(RDT_dat2$side)
levels(RDT_dat2$side)<-c("Left","Right")
RDT_dat2<-remove.outliers(RDT_dat2,200,2.2) #numbers specify min RT and zscore for Hoaglin-Iglewicz respectiely

# count subjects
nsub <-length(levels(RDT_dat2$subject))
print(paste0("Day 1 subjects: ", nsub))
outliertable<-table(RDT_dat2$subject,RDT_dat2$outlier)
print(outliertable)

if(d==1)
{RDT_dat1 <- RDT_dat2}

}
```

Now the data is ready, create an empty data frame and populate this. It includes columns for the participant identifier, rhyme, acuracy, RT (RT only for correct answers).

```{r makedataframeRDT, warning=F}
#We create files for day1 and day 2
# NB warnings created because of different Ns for time 1 and time 2, which means there are blanks in subjects
# We deal with this below by removing rows with no subject code
varlist <-   c("subject", "side", "accurate", "RT") #need to be in this order, ie sub, side, acc and RR - we will use these in the make.df function
origdat <- RDT_dat1
latlist <- c("Left","Right") #names of factor levels
RDT_df1 <- make.df(origdat,varlist,latlist)
origdat <- RDT_dat2
RDT_df2 <- make.df(origdat,varlist,latlist)

# Remove blank rows 
w<-which(is.na(RDT_df1$subject))
if (length(w)>0) {RDT_df1<-RDT_df1[-w,]}
w<-which(is.na(RDT_df2$subject))
if (length(w)>0) {RDT_df2<-RDT_df2[-w,]}

# Exclude cases with low % correct
# NB something weird with these - seem to be ones with too many trials, suggesting some problem with multiple responses

w<-which(RDT_df1$accurate<60)
if (length(w)>0)
{RDT_df1$accurate[w]<-NA
RDT_df1$RT[w]<-NA
}
w<-which(RDT_df2$accurate<60)
if (length(w)>0)
{RDT_df2$accurate[w]<-NA
RDT_df2$RT[w]<-NA
  }
```

```{r checker, warning=FALSE}
for (d in 1:2){
origdat <- RDT_df1
if (d==2)
{origdat <- RDT_df2}
measure <- 'Accuracy'
origdat$X1 <- origdat$accurate
task <- 'Rhyme Detection'

dopirate(origdat,task,measure)
origdat$X1 <- origdat$RT
measure <- 'RT'
dopirate(origdat,task,measure)

#now create LI
origdat$X1 <- origdat$accurate
mycolname<-"RDT_acc_LI"
polarity<- 1 #so LI is positive in predicted direction
mycutoff <- 75
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
origdat$X1 <- origdat$RT
mycolname<-"RDT_RT_LI"
polarity <- (-1)
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

#Cutoff has removed those with low accuracy, but we should also remove their RT LIs
#We do this by finding those with NA in accuracy LI
w<-which(is.na(demographics$RDT_acc_LI))
if(length(w)>0)
{demographics$RDT_RT_LI[w]<-NA}
#t.test to test if laterality index differs from zero
t.test(demographics$RDT_acc_LI)
t.test(demographics$RDT_RT_LI)
#rename column to identify day
myncol <- length(colnames(demographics))
colnames(demographics)[(myncol-1):myncol] <- paste0('Day',d,'.',colnames(demographics)[(myncol-1):myncol])

}
cor1<-cor(demographics$Day1.RDT_acc_LI,demographics$Day2.RDT_acc_LI,use='complete.obs')
cor2<-cor(demographics$Day1.RDT_RT_LI,demographics$Day2.RDT_RT_LI,use='complete.obs')
plot(demographics$Day1.RDT_acc_LI,demographics$Day2.RDT_acc_LI,main='Rhyme detect: LI for accuracy')
abline(v=0)
abline(h=0)
text(-3,3,paste0('r = ',round(cor1,3)))
plot(demographics$Day1.RDT_RT_LI,demographics$Day2.RDT_RT_LI,main='Rhyme detect: LI for RT')
abline(v=0)
abline(h=0)
text(-3,3,paste0('r = ',round(cor2,3)))
```
### T-test as alternative to LI for RDT

This is an exploratory analysis.

```{r do.ttest}
for (d in 1:2){
  myfile<-RDT_dat1
  if(d==2){
    myfile<-RDT_dat2}

  nsubs <- nrow(demographics)
  demographics$tlat_RDT<-NA
   demographics$plat_RDT<-NA
  for (i in 1:nsubs) {
    w<-which(myfile$subject==demographics$subject[i])
    if(length(w)>0)
    {
      myt<-t.test(myfile$RT[w]~myfile$side[w])
      demographics$tlat_RDT[i]<-myt$statistic
      demographics$plat_RDT[i]<-myt$p.value
    }
  }
   if (d==1){
     maxcol <-length(colnames(demographics))
     colnames(demographics)[(maxcol-1):maxcol]<-c('tlat_RDT_1','plat_RDT_1')
   }
   if (d==2){
     maxcol <-length(colnames(demographics))
     colnames(demographics)[(maxcol-1):maxcol]<-c('tlat_RDT_2','plat_RDT_2')
   }
}
colcount<-length(colnames(demographics))
# Remove values for cases that are excluded for low acc: do this for both times by just blitzing last 4 columns of demographics
w<-which(is.na(demographics$Day1.RDT_RT_LI))
demographics[w,(colcount-3):colcount]<-NA
w<-which(is.na(demographics$Day2.RDT_RT_LI))
demographics[w,(colcount-3):colcount]<-NA

#compare t-value with LI : use t = 1.65 as cutoff for sig lateralised
 plot(demographics$tlat_RDT_1,demographics$Day1.RDT_RT_LI)
 abline(v=1.65)
  plot(demographics$tlat_RDT_2,demographics$Day2.RDT_RT_LI)
   abline(v=1.65)
#compare t-value from day 1 and day 2 
     myr <- cor(demographics$tlat_RDT_1,demographics$tlat_RDT_2,use='complete.obs')
    plot(demographics$tlat_RDT_1,demographics$tlat_RDT_2)
   abline(v=1.65) 
   abline(h=1.65)
   text(0,2.5,paste0('r = ',round(myr,3)))
```

## Finger Tapping Task

This portion of the R Markdown file prcoesses the "Finger Tapping Task' implmented in Gorilla.sc which aims to quantify handedness. In the task, participants are required to press a sequence of buttons twice for each hand (Left: E, R, F, D; Right: Y, U, J, H) as many times as possible within a 30 ms. There is a version that requires more precise movements.

A score of 1 is given each time that a participants completes the sequence.

This reads in a csv file where each row represents a button press. Participants, identified by random letter strings, are stacked on top of each other. We start by reading the data and cutting out unwanted information.

```{r readdataFinger, warning=FALSE}
for (d in 1:2){ #we will process day 1 and day 2 together
  origfile <- all_dat1
  if (d==2)
  {origfile <- all_dat2}
#As with earlier reading in, we default to saving finger2 (day2), but at end of the chunk we reassign to finger1 in the first run through the d loop
task <- "Key Sequence Task"
wanted <- c("subject", "Response", "display") # select wanted columns
finger2 <- procdata(origfile,wanted,task)

# count subjects
nsub <-length(levels(finger2$subject))
print(paste('day',d,task,": subjects: ", nsub))

if (d==1)
{finger1 <- finger2
 }
}
```

In this section, the correct sequences are marked (e.g. pressing Y, M, I, B, in order will correspond to a score of 1).

```{r scorefinger}
# We can go directly to creation of a data.frame by subjects, and use string matching to identify N correct seqs
# y, m, i, b- complex
# w, v, r, x- complex
# t, u, m, b- easy
# w, r, v, x- easy

myseq <-c( 'ymib','tumb','wvrx','wrvx') #R complex , R easy, L complex, L easy

subids <- levels(finger1$subject) #just look at data for those doing time1 and
nsub <- length(subids)

finger.df <- data.frame(matrix(NA,nrow=nsub,ncol=9))
colnames(finger.df)<-c('subject','Leasy1','Reasy1','Lhard1','Rhard1',
                       'Leasy2','Reasy2','Lhard2','Rhard2')
for (n in 1:nsub){
  finger.df$subject[n] <- subids[n]
  myc <- 1 #will be incremented so sums written to correct columns
  for (d in 1:2){ #each day data
    myfinger <- finger1
    if(d==2){
      myfinger<- finger2}
      tempfinger <- filter(myfinger,subject==subids[n],Response %in% c('y','m','i','b','t','u','w','v','r','x')) #exclude rows with other stuff in Response
      #(this is not essential as nonsequences will just get ignored)
      wholeseq <-paste0(tempfinger$Response,collapse='') #make a long string of responses
      for (s in 1:4){
        myc<-myc+1 #increment column counter
        target <- myseq[s]
        finger.df[n,myc]<-str_count(wholeseq,target) #count how many target sequences are in the long string that contains all responses
      }
    }
  }
#Need same data in long form for pirate plot, so reorganise
finger.left <- finger.df[,c(1,2,4,6,8)]
finger.right <- finger.df[,c(1,3,5,7,9)]
finger.left$side<-'Left'
finger.right$side <-'Right'
colnames(finger.right)<-c('subject','easy1','hard1','easy2','hard2','side')
colnames(finger.left)<-colnames(finger.right)
finger.long <-rbind(finger.left,finger.right)

```

```{r dosummary, warning = F}
origdat <- finger.long
measure <- 'easy1'
origdat$X1 <- origdat$easy1 #we used a dummy column name X1, so we can vary the data that are used in the pirate plot etc
task <- 'Finger easy day 1'
dopirate(origdat,task,measure)

measure <- 'easy2'
origdat$X1 <- origdat$easy2 #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Finger easy day 2'
dopirate(origdat,task,measure)

measure <- 'hard1'
origdat$X1 <- origdat$hard1 #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Finger hard day 1'
dopirate(origdat,task,measure)

measure <- 'hard2'
origdat$X1 <- origdat$hard2 #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Finger hard day 2'
dopirate(origdat,task,measure)
```

Finally, a handedness index is calculated and plotted. The handedness index is based on the equation [(RH-LH)/(RH+LH)] x 100, where LH= left hand and RH= right hand.

If the index is negative, the participant is categorised as left handed. If the index is positive, the participant is categorised as right handed.

In the next chunk, the LI is also added to demographics

```{r fingerLI}
task<-'Finger'
origdat<-finger.long
origdat$X1 <- origdat$easy1
mycolname<-"Day1_finger.easy_LI"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (1)#so LI is positive in predicted direction
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

origdat$X1 <- origdat$hard1
mycolname<-"Day1_finger.hard_LI"
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

origdat$X1 <- origdat$easy2
mycolname<-"Day2_finger.easy_LI"
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

origdat$X1 <- origdat$hard2
mycolname<-"Day2_finger.hard_LI"
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

cor1 <- cor(demographics$Day1_finger.easy_LI,demographics$Day2_finger.easy_LI)
plot(demographics$Day1_finger.easy_LI,demographics$Day2_finger.easy_LI,main='Finger easy LI')
abline(v=0)
abline(h=0)
text(40,10,paste0('r = ',round(cor1,3)))
cor2 <- cor(demographics$Day1_finger.hard_LI,demographics$Day2_finger.hard_LI)
plot(demographics$Day1_finger.hard_LI,demographics$Day2_finger.hard_LI,main='Finger hard LI')
abline(v=0)
abline(h=0)
text(40,10,paste0('r = ',round(cor2,3)))

```

**References:**

- Burt, D. M., & Perrett, D. I. (1997). Perceptual asymmetries in judgements of facial attractiveness, age, gender, speech and expression. Neuropsychologia, 35(5), 685–693.
- Innes, B. R., Burt, D. M., Birch, Y. K., & Hausmann, M. (2016). A leftward bias however you look at it: Revisiting the emotional chimeric face task as a tool for measuring emotion lateralization. Laterality: Asymmetries of Body, Brain and Cognition, 21(4-6), 643–661.
- Karlsson, E. M., Johnstone, L. T., & Carey, D. P. (2019). The depth and breadth of multiple perceptual asymmetries in right handers and non-right handers. Laterality: Asymmetries of Body, Brain and Cognition, 24(6), 707-739.

### 2020-02-02 battery-preprocessing_v3_DB.Rmd
Found some errors in the finger task analysis so corrected these, and also added an analysis of SDs of RTs on that task. Clarified that for chimeric task, accuracy is not % correct, because both L and R correct, so this is a choice measure. But for RDT, we do have percent correct, and this is now renamed as p.corr to make that more explicit.

@@
     temp_dat$side[myrow] <- latlist[j] # add row for side
-    temp_dat$accurate[myrow] <- 100*mean(tmp1$Correct,na.rm=TRUE) # add row for accurate responses - this has been modified to give % correct
+    temp_dat$accurate[myrow] <- sum(tmp$Correct,na.rm=TRUE) # add row for N accurate response
+    temp_dat$p.corr[myrow] <- 100*mean(tmp$Correct,na.rm=TRUE) # add row for %accurate responses - NB for chimeric only error is wrong emotion - measure is just a measure of choice.
+For RDT, can make error of selecting wrong side, so % correct is meaningful
+
     temp_dat$RT[myrow] <- mean(tmp2$RT,na.rm=TRUE) # add row for mean RT
@@
 task <- 'Chimeric faces day 2'
-
+measure <- 'Accuracy'
 dopirate(origdat,task,measure)
@@
 
-w<-which(RDT_df1$accurate<60)
+w<-which(RDT_df1$p.corr<60)
 if (length(w)>0)
@@
 }
-w<-which(RDT_df2$accurate<60)
+w<-which(RDT_df2$p.corr<60)
 if (length(w)>0)
@@
 #now create LI
-origdat$X1 <- origdat$accurate
-mycolname<-"RDT_acc_LI"
+origdat$X1 <- origdat$p.corr
+mycolname<-"RDT_p.corr_LI"
 polarity<- 1 #so LI is positive in predicted direction
@@
 #t.test to test if laterality index differs from zero
-t.test(demographics$RDT_acc_LI)
+t.test(demographics$RDT_p.corr_LI)
 t.test(demographics$RDT_RT_LI)
@@
 }
-cor1<-cor(demographics$Day1.RDT_acc_LI,demographics$Day2.RDT_acc_LI,use='complete.obs')
+cor1<-cor(demographics$Day1.RDT_p.corr_LI,demographics$Day2.RDT_p.corr_LI,use='complete.obs')
 cor2<-cor(demographics$Day1.RDT_RT_LI,demographics$Day2.RDT_RT_LI,use='complete.obs')
-plot(demographics$Day1.RDT_acc_LI,demographics$Day2.RDT_acc_LI,main='Rhyme detect: LI for accuracy')
+plot(demographics$Day1.RDT_p.corr_LI,demographics$Day2.RDT_p.corr_LI,main='Rhyme detect: LI for accuracy (%correct)')
 abline(v=0)
@@
 task <- "Key Sequence Task"
-wanted <- c("subject", "Response", "display") # select wanted columns
+wanted <- c("subject", "Response", "display","RT") # select wanted columns
+#NB RT is not meaningful, in that we are just counting N sequences in fixed time.
+# But it could be looked at for measure of variability.
 finger2 <- procdata(origfile,wanted,task)
-
+finger2$complex <-'easy'
+task <- "Key Sequence Task (wider keys)"
+finger2a<-procdata(origfile,wanted,task)
+finger2a$complex <-'hard'
+finger2<-rbind(finger2,finger2a)
 # count subjects
@@
 
+finger2$RTbase[2:nrow(finger2)]<-finger2$RT[1:(nrow(finger2)-1)]
+finger2$RTa <- finger2$RT-finger2$RTbase
+#remove negative or v long RTs - can indicate when block changes etc
+w<-c(which(finger2$RTa<0),which(finger2$RTa>1000))
+finger2$RTa[w]<-NA
 if (d==1)
@@
 
+Added exploratory analysis of SDs of RTs for keypress rather than means, but this was not particularly illuminating.
+
 ```{r scorefinger}
 # We can go directly to creation of a data.frame by subjects, and use string matching to identify N correct seqs
-# y, m, i, b- complex
-# w, v, r, x- complex
-# t, u, m, b- easy
-# w, r, v, x- easy
+# y, m, i, b- complex R
+# t, u, m, b- easy R
+# w, v, r, x- complex L
+# w, r, v, x- easy L
 
@@
 
-finger.df <- data.frame(matrix(NA,nrow=nsub,ncol=9))
-colnames(finger.df)<-c('subject','Leasy1','Reasy1','Lhard1','Rhard1',
-                       'Leasy2','Reasy2','Lhard2','Rhard2')
+finger.df <- data.frame(matrix(NA,nrow=nsub,ncol=17))
+colnames(finger.df)<-c('subject','Rhard1','Reasy1','Lhard1','Leasy1',
+                       'Rhard2','Reasy2','Lhard2','Leasy2',
+                       'Rhard1sd','Reasy1sd','Lhard1sd','Leasy1sd',
+                       'Rhard2sd','Reasy2sd','Lhard2sd','Leasy2sd')
 for (n in 1:nsub){
@@
       #(this is not essential as nonsequences will just get ignored)
+
       wholeseq <-paste0(tempfinger$Response,collapse='') #make a long string of responses
@@
         finger.df[n,myc]<-str_count(wholeseq,target) #count how many target sequences are in the long string that contains all responses
-      }
-    }
+
+      }
+      righteasy<-filter(tempfinger,display=='right hand',complex=='easy')
+      righthard<-filter(tempfinger,display=='right hand',complex=='hard')
+      lefteasy<-filter(tempfinger,display=='left hand',complex=='easy')
+      lefthard<-filter(tempfinger,display=='left hand',complex=='hard')
+      startcol<-10
+      if(d==2){startcol<-14}
+      finger.df[n,startcol]<-sd(righthard$RTa,na.rm=T)
+      finger.df[n,(startcol+1)]<-sd(righteasy$RTa,na.rm=T)
+      finger.df[n,(startcol+2)]<-sd(lefthard$RTa,na.rm=T)
+      finger.df[n,(startcol+3)]<-sd(lefteasy$RTa,na.rm=T)
+    }
+}
+# Substitute NA for any totals less than 5 - suggest using wrong keys perhaps
+
+for (c in 2:9){
+  w<-which(finger.df[,c]<5)
+  if (length(w)>0){
+     finger.df[w,2:17]<-NA
+
+  }
+
 }
 #Need same data in long form for pirate plot, so reorganise
-finger.left <- finger.df[,c(1,2,4,6,8)]
-finger.right <- finger.df[,c(1,3,5,7,9)]
+finger.left <- finger.df[,c(1,4,5,8,9,12,13,16,17)]
+finger.right <- finger.df[,c(1,2,3,6,7,10,11,14,15)]
 finger.left$side<-'Left'
 finger.right$side <-'Right'
-colnames(finger.right)<-c('subject','easy1','hard1','easy2','hard2','side')
+colnames(finger.right)<-c('subject','hard1','easy1','hard2','easy2',
+                          'hard1sd','easy1sd','hard2sd','easy2sd','side')
 colnames(finger.left)<-colnames(finger.right)
@@
 origdat$X1 <- origdat$hard2 #we used a dummy column name so we can vary the data that are used in the pirate plot etc
+task <- 'Finger hard day 2'
+dopirate(origdat,task,measure)
+
+measure<-'easy1sd'
+origdat$X1 <- origdat$easy1sd #we used a dummy column name X1, so we can vary the data that are used in the pirate plot etc
+task <- 'Finger easy day 1sd'
+dopirate(origdat,task,measure)
+
+measure <- 'easy2sd'
+origdat$X1 <- origdat$easy2sd #we used a dummy column name so we can vary the data that are used in the pirate plot etc
+task <- 'Finger easy day 2'
+dopirate(origdat,task,measure)
+
+measure <- 'hard1sd'
+origdat$X1 <- origdat$hard1sd #we used a dummy column name so we can vary the data that are used in the pirate plot etc
+task <- 'Finger hard day 1'
+dopirate(origdat,task,measure)
+
+measure <- 'hard2sd'
+origdat$X1 <- origdat$hard2sd #we used a dummy column name so we can vary the data that are used in the pirate plot etc
 task <- 'Finger hard day 2'
@@
 
-cor1 <- cor(demographics$Day1_finger.easy_LI,demographics$Day2_finger.easy_LI)
+cor1 <- cor(demographics$Day1_finger.easy_LI,demographics$Day2_finger.easy_LI,use='complete.obs')
 plot(demographics$Day1_finger.easy_LI,demographics$Day2_finger.easy_LI,main='Finger easy LI')
@@
 text(40,10,paste0('r = ',round(cor1,3)))
-cor2 <- cor(demographics$Day1_finger.hard_LI,demographics$Day2_finger.hard_LI)
+cor2 <- cor(demographics$Day1_finger.hard_LI,demographics$Day2_finger.hard_LI,use='complete.obs')
 plot(demographics$Day1_finger.hard_LI,demographics$Day2_finger.hard_LI,main='Finger hard LI')

### 2020-02-03 battery-preprocessing_v3_DB.Rmd
renamed Trial.Number as Trial and included this when reading RDT. In principle might want to look at how LI changes over blocks, but would need also to code block. Trial refers to trial within block - is there a block code in Gorilla? If not we could compute this. Could check Carey idea that *initial* performance is most informative - seems unlikely but could check!

@@
 all_dat2$RT <- as.numeric(as.character(all_dat2$Reaction.Time))
+all_dat2$Trial <-as.numeric(all_dat2$Trial.Number) #useful if we want to restrict analysis to subset of trials so retain this variable
 all_dat2$Task.Name <- as.factor(all_dat2$Task.Name)
@@
 remove.outliers <- function(origdat,lowcut,zcut){
-# Outlier removal put in separate function, as needed to reuse when considering t-test
-# Hoaglin, D. C., & Iglewicz, B. (1987). Fine-tuning some resistant rules for outlier labelling. Journal of the American Statistical Association, 82(400), 1147-1149.
-
+# Outlier removal put in separate function, which can be applied to different task RTs.
+# 
 # origdat is the dataframe with data for outlier removal, mycolname is name of variable 
@@
 wanted <- c("subject", "stimuli", "Response", "RT", "l_hemiface", "r_hemiface") # select wanted columns
-CF_dat2 <- procdata(origfile,wanted,task)
+CF_dat2 <- procdata(origfile,wanted,task) #use generic function (see above) to read in relevant columns
 
@@
 CF_dat2 <- CF_dat2[CF_dat2$l_hemiface != CF_dat2$r_hemiface,] # remove items where same expression in both hemifaces
-CF_dat2 <- remove.outliers(CF_dat2,200,2.2)
+CF_dat2 <- remove.outliers(CF_dat2,200,2.2) #use generic outlier removal function
 outliertable<-table(CF_dat2$subject,CF_dat2$outlier)
@@
 # new data
-wanted <- c("subject", "Correct", "RT", "ANSWER","Response") # select wanted columns
+wanted <- c("Trial","subject", "Correct", "RT", "ANSWER","Response") # select wanted columns
 RDT_dat2 <- procdata(origfile,wanted,task) #Here we reuse the procdata function that we developed for processing Chimeric faces data - just does some generic processing
@@
 abline(h=0)
-text(40,10,paste0('r = ',round(cor1,3)))
+text(-5,20,paste0('r = ',round(cor1,3)))
 cor2 <- cor(demographics$Day1_finger.hard_LI,demographics$Day2_finger.hard_LI,use='complete.obs')
@@
 abline(h=0)
-text(40,10,paste0('r = ',round(cor2,3)))
+text(-5,20,paste0('r = ',round(cor2,3)))
 
@@
 - Burt, D. M., & Perrett, D. I. (1997). Perceptual asymmetries in judgements of facial attractiveness, age, gender, speech and expression. Neuropsychologia, 35(5), 685–693.
+- Hoaglin, D. C., & Iglewicz, B. (1987). Fine-tuning some resistant rules for outlier labelling. Journal of the American Statistical Association, 82(400), 1147-1149. 
 - Innes, B.

### 2020-02-03 battery-preprocessing_v3_DB.Rmd
pointless change to demonstrate commit

@@
 require(stringr) # string matching
+library(maps)
 ```

### 2020-02-03 battery-preprocessing_v3_DB.Rmd
fixed an error in computing mean accuracy for Chimeric faces and RDT task

--- 2020-02-03 battery-preprocessing_v3_DB.Rmd
@@
 require(stringr) # string matching
+
 ```
@@
     temp_dat$side[myrow] <- latlist[j] # add row for side
-    temp_dat$accurate[myrow] <- sum(tmp$Correct,na.rm=TRUE) # add row for N accurate response
-    temp_dat$p.corr[myrow] <- 100*mean(tmp$Correct,na.rm=TRUE) # add row for %accurate responses - NB for chimeric only error is wrong emotion - measure is just a measure of choice.
+    temp_dat$accurate[myrow] <- sum(tmp2$Correct,na.rm=TRUE) # add row for N accurate response
+    temp_dat$p.corr[myrow] <- 100*mean(tmp2$Correct,na.rm=TRUE) # add row for %accurate responses - NB for chimeric only error is wrong emotion - measure is just a measure of choice.
 For RDT, can make error of selecting wrong side, so % correct is meaningful

### 2020-02-12 battery-preprocessing_v3_DB.Rmd
Modified to correct computation of p.corr for RDT. Also added z-test for difference in proportions for laterality measures based on accuracy

@@
 
-In this section, the correct sequences are marked as 1, incorrect asnwers are marked as 0. We then mark whether the correct answers were in the left or right hemiface.
+In this section, the correct sequences are marked as 1, incorrect answers are marked as 0. We then mark whether the correct answers were in the left or right hemiface.
 
@@
     temp_dat$accurate[myrow] <- sum(tmp2$Correct,na.rm=TRUE) # add row for N accurate response
-    temp_dat$p.corr[myrow] <- 100*mean(tmp2$Correct,na.rm=TRUE) # add row for %accurate responses - NB for chimeric only error is wrong emotion - measure is just a measure of choice.
+    temp_dat$p.corr[myrow] <- 100*mean(tmp1$Correct,na.rm=TRUE) # add row for %accurate responses - NB for chimeric only error is wrong emotion - measure is just a measure of choice.
 For RDT, can make error of selecting wrong side, so % correct is meaningful
+    #need to use tmp1, as this contains errors
 
@@
 
-The data is visualised using the pirate plot function and t.test conducted.
-
+Generic function for pirate plot and group-level t-test:
 ```{r plotdata_ttest, warning=FALSE}
@@
 
+The data is visualised using the pirate plot function and t.test for laterality at group level conducted.
+
 ```{r dosummary}
@@
 
-makeLI <- function(demogs,origdat,task,mycolname,polarity,mycutoff){
-# we do this for accuracy first
+makeLI <- function(demogs,origdat,task,mycolname1,mycolname2,polarity,mycutoff){
+
+# mycolname1 and 2 are names for newly created columns for LI and statistic (z)
 # reformat data frame
@@
 
+#statistic is z score - see http://www.sthda.com/english/wiki/one-proportion-z-test-in-r
+LIdf$statistic <-NA
+for (i in 1:nrow(LIdf)){
+    pN <- (LIdf$Left[i]+LIdf$Right[i])
+  p0 <- LIdf$Left[i]/ pN
+  q <- 1-p0
+  pe <- .5
+  LIdf$statistic[i]<-(p0-pe)/sqrt(p0*q/pN)
+}
+
 # now create a data frame for lat index 
-LIdf <- select(LIdf, "subject", "LI")
+LIdf <- select(LIdf, "subject", "LI","statistic")
 # now merge with demographics
@@
 lastcol<-length(colnames(demogs))
-colnames(demogs)[lastcol]<-mycolname
+colnames(demogs)[(lastcol-1):lastcol]<- c(mycolname1,mycolname2)
 
@@
 origdat$X1 <- origdat$accurate
-mycolname<-"Day1_CF_acc_LI"
+mycolname1<-"Day1_CF_acc_LI"
+mycolname2 <-"Day1_CF_z"
 mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
 polarity<- (-1)#so LI is positive in predicted direction
-demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
-
-# now do RT
-origdat$X1 <- origdat$RT
-mycolname<-"Day1_CF_RT_LI"
-polarity <-1
-
-#if we don't specify mycutoff, it will remain at level set for accuracy, which means it should not trap any RT values as they will all be much bigger
-demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
+demographics <- makeLI(demographics,origdat,task,mycolname1,mycolname2,polarity,mycutoff)
 
@@
 origdat$X1 <- origdat$accurate
-mycolname<-"Day2_CF_acc_LI"
+mycolname1<-"Day2_CF_acc_LI"
+mycolname2 <-"Day2_CF_z"
 polarity<- (-1)#so LI is positive in predicted direction
-demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
-origdat$X1 <- origdat$RT
-mycolname<-"Day2_CF_RT_LI"
-polarity <-1
-demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
-
-cor1 <- cor(demographics$Day1_CF_acc_LI,demographics$Day2_CF_acc_LI)
+demographics <- makeLI(demographics,origdat,task,mycolname1,mycolname2,polarity,mycutoff)
+
+cor1 <- cor(demographics$Day1_CF_acc_LI,demographics$Day2_CF_acc_LI,use="complete.obs")
 plot(demographics$Day1_CF_acc_LI,demographics$Day2_CF_acc_LI,main='Chimeric LI accuracy')
@@
 text(-10,15,paste0('r = ',round(cor1,3)))
-cor2<-cor(demographics$Day1_CF_RT_LI,demographics$Day2_CF_RT_LI)
-plot(demographics$Day1_CF_RT_LI,demographics$Day2_CF_RT_LI,main='Chimeric LI RT')
+cor2<-cor(demographics$Day1_CF_z,demographics$Day2_CF_z,use="complete.obs")
+plot(demographics$Day1_CF_z,demographics$Day2_CF_z,main='Chimeric LI.z')
 abline(v=0)
 abline(h=0)
-text(-5,15,paste0('r = ',round(cor2,3)))
+abline(v=1.65,lty=2,col='red')
+abline(v=-1.65,lty=2,col='red')
+abline(h=1.65,lty=2,col='red')
+abline(h=-1.65,lty=2,col='red')
+text(-1,3,paste0('r = ',round(cor2,3)))
+
+cor3 <- cor(demographics$Day1_CF_acc_LI,demographics$Day1_CF_z,use="complete.obs")
+plot(demographics$Day1_CF_acc_LI,demographics$Day1_CF_z,main='Chimeric LI vs z time1')
+abline(v=0)
+abline(h=0)
 
@@
 As we can see, we generally get a left side/right hemisphere advantage in our accuracy data. 
-The RT advantage is not as expected.
-This may be due to the buttons being clicked on screen and the additional time penalty of moving the mouse.
+Agreement between the z-score and conventional LI is very strong.
+Test-retest seems good.
 
@@
 
-Now the data is ready, create an empty data frame and populate this. It includes columns for the participant identifier, rhyme, acuracy, RT (RT only for correct answers).
+Now the data is ready, create an empty data frame and populate this. It includes columns for the participant identifier, rhyme, accuracy, RT (RT only for correct answers).
 
@@
 RDT_df1$RT[w]<-NA
+RDT_df1$p.corr[w]<-NA
 }
@@
 RDT_df2$RT[w]<-NA
+RDT_df2$p.corr[w]<-NA
   }
@@
 origdat$X1 <- origdat$p.corr
-mycolname<-"RDT_p.corr_LI"
+mycolname1<-"RDT_p.corr_LI"
+mycolname2 <- "RDT_pcorr.z"
 polarity<- 1 #so LI is positive in predicted direction
 mycutoff <- 75
-demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
-origdat$X1 <- origdat$RT
-mycolname<-"RDT_RT_LI"
-polarity <- (-1)
-demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
+
+demographics <- makeLI(demographics,origdat,task,mycolname1,mycolname2,polarity,mycutoff)
 
@@
 {demographics$RDT_RT_LI[w]<-NA}
-#t.test to test if laterality index differs from zero
+#t.test to test if laterality index differs from zero in sample as a whole
 t.test(demographics$RDT_p.corr_LI)
-t.test(demographics$RDT_RT_LI)
+t.test(demographics$RDT_pcorr.z)
 #rename column to identify day
@@
 cor1<-cor(demographics$Day1.RDT_p.corr_LI,demographics$Day2.RDT_p.corr_LI,use='complete.obs')
-cor2<-cor(demographics$Day1.RDT_RT_LI,demographics$Day2.RDT_RT_LI,use='complete.obs')
 plot(demographics$Day1.RDT_p.corr_LI,demographics$Day2.RDT_p.corr_LI,main='Rhyme detect: LI for accuracy (%correct)')
@@
 text(-3,3,paste0('r = ',round(cor1,3)))
-plot(demographics$Day1.RDT_RT_LI,demographics$Day2.RDT_RT_LI,main='Rhyme detect: LI for RT')
-abline(v=0)
-abline(h=0)
-text(-3,3,paste0('r = ',round(cor2,3)))
+
 ```
@@
 
-This is an exploratory analysis.
+Advantage of using a t-statistics to represent LI is that we can also identify people whose L-R difference is significantly different from zero.
 
@@
 
-#compare t-value with LI : use t = 1.65 as cutoff for sig lateralised
- plot(demographics$tlat_RDT_1,demographics$Day1.RDT_RT_LI)
- abline(v=1.65)
-  plot(demographics$tlat_RDT_2,demographics$Day2.RDT_RT_LI)
-   abline(v=1.65)
 #compare t-value from day 1 and day 2 
      myr <- cor(demographics$tlat_RDT_1,demographics$tlat_RDT_2,use='complete.obs')
-    plot(demographics$tlat_RDT_1,demographics$tlat_RDT_2)
-   abline(v=1.65) 
-   abline(h=1.65)
+    plot(demographics$tlat_RDT_1,demographics$tlat_RDT_2,xlim=c(-4,4),ylim=c(-4,4))
+  abline(v=0)
+  abline(h=0)
+   abline(v=-1.65,lty=2,col='red') 
+   abline(h=-1.65,lty=2,col='red')
+    abline(v=1.65,lty=2,col='red') 
+   abline(h=1.65,lty=2,col='red')
    text(0,2.5,paste0('r = ',round(myr,3)))
@@
 origdat$X1 <- origdat$easy1
-mycolname<-"Day1_finger.easy_LI"
+mycolname1<-"Day1_finger.easy_LI"
+mycolname2<-"Day1_finger.easy_z"
 mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
 polarity<- (1)#so LI is positive in predicted direction
-demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
+demographics <- makeLI(demographics,origdat,task,mycolname1,mycolname2,polarity,mycutoff)
 
 origdat$X1 <- origdat$hard1
-mycolname<-"Day1_finger.hard_LI"
-demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
+mycolname1<-"Day1_finger.hard_LI"
+mycolname2<-"Day1_finger.hard_z"
+demographics <- makeLI(demographics,origdat,task,mycolname1,mycolname2,polarity,mycutoff)
 
 origdat$X1 <- origdat$easy2
-mycolname<-"Day2_finger.easy_LI"
-demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
+mycolname1<-"Day2_finger.easy_LI"
+mycolname2<-"Day2_finger.easy_z"
+demographics <- makeLI(demographics,origdat,task,mycolname1,mycolname2,polarity,mycutoff)
 
 origdat$X1 <- origdat$hard2
-mycolname<-"Day2_finger.hard_LI"
-demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
+mycolname1<-"Day2_finger.hard_LI"
+mycolname2<-"Day2_finger.hard_z"
+demographics <- makeLI(demographics,origdat,task,mycolname1,mycolname2,polarity,mycutoff)
 
@@
 text(-5,20,paste0('r = ',round(cor2,3)))
+
+cor3 <- cor(demographics$Day1_finger.easy_z,demographics$Day2_finger.easy_z,use='complete.obs')
+plot(demographics$Day1_finger.easy_z,demographics$Day2_finger.easy_z,main='Finger easy z')
+abline(v=0)
+abline(h=0)
+text(-5,20,paste0('r = ',round(cor1,3)))
 

### 2020-02-12 battery-preprocessing_v3_DB.Rmd
Based t-test on log RT rather than raw RT

@@
     {
-      myt<-t.test(myfile$RT[w]~myfile$side[w])
+      myt<-t.test(log(myfile$RT[w])~myfile$side[w])  #NB USING LOG HERE TO GIVE MORE NORMAL RT DATA
       demographics$tlat_RDT[i]<-myt$statistic

### 2020-04-21 battery-preprocessing_v3_DB.Rmd
saving before modifying for new dataset

@@
 author: "Adam Parker + Dorothy Bishop"
-date: "1st Feb 2020"
+date: "original version from 1st Feb 2020"
 output: html_document

### 2020-04-22 battery-preprocessing_v5_DB.Rmd
based on v3; did new version to avoid conflicts. Analysing new data from 300 people.

---
title: "Online Laterality Battery Pre-Processing - short version"
author: "Adam Parker + Dorothy Bishop"
date: "20th March 2020"
output:
  pdf_document: default
  html_document: default
---

This R Markdown file prcoesses the various components of the Online Laterality Battery which is implemented online via Gorilla.sc.

This version updated on 1st Feb 2020 to make script more reproducible and compact by using functions for repeated operations.

The script pulls in csv files and processes them so that pre-registered analyses can be conducted (pre-registered on the OSF: ). First, questionnaires are analyses followed by visual half field tasks. This is done for time 1 and time 2. A "full" laterality index output is then appended to a file with the demographics. This is then used in subsequent R markdown scripts.

This script analyses the following:

1) Questionnaires
  - Demographics
  - Miles test
  - Porta test
  - Edinburgh Handedness Inventory
  - lexTALE
2) Chimeric face task
3) Rhyme detection task
4) Key Pressing Task (finger dexterity)
5) Summary of demographics for the final sample

# Questionnaires

This R Markdown processes the following questionnaires from the Online battery:

- Demographics
- Miles test
- Porta test
- Edinburgh Handedness Inventory
- lexTALE

Once processed, a file is made.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#NB for present setwd("~/Dropbox/Online LDT/300 Ps")
# libraries 
library(readr)
library(yarrr)
library(dplyr)
library(tidyr)
library(car)
library(ggplot2)
library(knitr)
require(readr)  # for read_csv()
require(dplyr)  # for mutate()
require(tidyr)  # for unnest()
require(purrr)  # for map(), reduce()
library(GGally)
require(stringr) # string matching
library(ggpirate)
library(corrr) #easier correlations!
```

First, the demographics is processed.This gives us a participant ID, age, gender, footedness, and handedness.

```{r demographics, warning=FALSE}
demo_dat_R <- read_csv("./Demographics_R.csv", na = c("empty", "NA")) # read in data
demo_dat_L <- read_csv("./Demographics_L.csv", na = c("empty", "NA")) # read in data
demo_dat <- rbind(demo_dat_R, demo_dat_L)
# relabel 
demo_dat$subject <- demo_dat$`Participant Public ID`
demo_dat$subject <- as.factor(demo_dat$subject)
# count and print subjects
nsub <-length(levels(demo_dat$subject))
print(paste0("subjects: ", nsub))

# create data frame
demographics <- data.frame(matrix(ncol = 7, nrow = nsub))
  colnames(demographics) <- c("subject", "age", "gender", "footedness", "handedness", "bilingual", "strongest.language")
myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  subname <- levels(demo_dat$subject)[i] # find subject ID
  myrows <- which(demo_dat$subject==subname) # select rows for this subject
  tmp <- data.frame(demo_dat[myrows,])

    myrow <- myrow+1

    demographics$subject[myrow] <- subname
    demographics$age[myrow] <- as.numeric(tmp[tmp$Question.Key == "Age",]$Response)
    demographics$gender[myrow] <- tmp[tmp$Question.Key == "Gender",]$Response
    demographics$footedness[myrow] <- tmp[tmp$Question.Key == "footedness",]$Response
    demographics$handedness[myrow] <- tmp[tmp$Question.Key == "categorical_hand",]$Response
    demographics$bilingual[myrow] <- tmp[tmp$Question.Key == "bilingual",]$Response
    demographics$strongest.language[myrow] <- tolower(tmp[tmp$Question.Key == "bilingual-text",]$Response)

    demographics$bilingual <- as.character(demographics$bilingual)
    demographics$bilingual <- gsub('. My strongest language is:', '', demographics$bilingual)
}

# Here we summarise the age broken down by handedness and gender. This will be important for any write up. 
#### NOTE. THIS IS NOT REPRESENTATIVE OF THE FINAL DATASET
# age
demographics %>%
    group_by(handedness, gender) %>%
    summarize(avg = mean(age), n = n(), sd = sd(age))
```

Now we import and append the Porta and Miles Test to the demographics data frame.

```{r sightednesss, warning=FALSE}
miles_dat_R <- read_csv("./Miles_R.csv", na = c("empty", "NA")) # read in data
miles_dat_L <- read_csv("./Miles_L.csv", na = c("empty", "NA")) # read in data
miles_dat <- rbind(miles_dat_R, miles_dat_L)
#porta_dat_R <- read_csv("./dat2/Porta_R.csv", na = c("empty", "NA")) # read in data
porta_dat_L <- read_csv("./Porta_L.csv", na = c("empty", "NA")) # read in data
porta_dat <- rbind(porta_dat_L)
# add test ID
miles_dat$test_ID <- "miles"
porta_dat$test_ID <- "porta"
# merge the two
sight_dat <- rbind(miles_dat, porta_dat)
# relabel 
sight_dat$subject <- sight_dat$`Participant Public ID`
sight_dat$subject <- as.factor(sight_dat$subject)

# count and print subjects
nsub <-length(levels(sight_dat$subject))
print(paste0("subjects: ", nsub))

# create data frame
sightedness <- data.frame(matrix(ncol = 3, nrow = nsub))
  colnames(sightedness) <- c("subject", "miles", "porta")
myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  subname <- levels(sight_dat$subject)[i] # find subject ID
  myrows <- which(sight_dat$subject==subname) # select rows for this subject
  tmp <- data.frame(sight_dat[myrows,])

    myrow <- myrow+1

    sightedness$subject[myrow] <- subname
    sightedness$miles[myrow] <- tmp[tmp$Question.Key == "response-2",]$Response
    sightedness$porta[myrow] <- tmp[tmp$Question.Key == "response-2",]$Response
  }

# now merge sightedness and demographics
demographics <- merge(demographics, sightedness, by= "subject")
```

Now we add the EHI. Here we score and plot the Edinburgh handedness inventory (Oldfield, 1971) in its short form.

```{r EHI, warning=FALSE}
EHI_R <- read_csv("./EHI_R.csv", na = "empty") # read in data
EHI_L <- read_csv("./EHI_L.csv", na = "empty") # read in data
EHI <- rbind(EHI_R, EHI_L)

# select only responses
EHI <- EHI[EHI$`Question Key` == "response-2-quantised" | EHI$`Question Key` == "response-3-quantised" | 
           EHI$`Question Key` == "response-4-quantised" | EHI$`Question Key` == "response-5-quantised" |
           EHI$`Question Key` == "response-6-quantised" | EHI$`Question Key` == "response-7-quantised" | 
           EHI$`Question Key` == "response-8-quantised" | EHI$`Question Key` == "response-9-quantised" |
           EHI$`Question Key` == "response-10-quantised" | EHI$`Question Key` == "response-11-quantised" | 
           EHI$`Question Key` == "response-12-quantised" | EHI$`Question Key` == "response-13-quantised" |
           EHI$`Question Key` == "response-14-quantised" | EHI$`Question Key` == "response-15-quantised" | 
           EHI$`Question Key` == "response-16-quantised" | EHI$`Question Key` == "response-17-quantised" |
           EHI$`Question Key` == "response-18-quantised" | EHI$`Question Key` == "response-19-quantised" | 
           EHI$`Question Key` == "response-20-quantised",]

# recode variabels
EHI$subject <- EHI$`Participant Public ID`
EHI$subject <- as.factor(EHI$subject)

# reduce data
# new data
meaningful <- c("subject", "Question Key", "Response") # select wanted columns
c<-which(names(EHI) %in% meaningful) #find colnumbers of unwanted
EHI <-EHI[,c] #remove unwanted columns
EHI <- na.omit(EHI) # remove NAs

# score responses accoridng to the original scoring method in Oldifeld (1971)
EHI$right_hand <- 0
EHI$left_hand <- 0
for (r in 1:nrow(EHI)) {
  if(EHI$Response[r] == 1){
    EHI$right_hand[r]= 2 
  } else {
    if(EHI$Response[r] == 2){
      EHI$right_hand[r]= 1
    } else {
      if(EHI$Response[r] == 4){
        EHI$left_hand[r]= 1
      } else {
        if(EHI$Response[r] == 5){
          EHI$left_hand[r]= 2 
        } else {
          (EHI$right_hand[r]= 1) & (EHI$left_hand[r]= 1)
        }
      }
    }
  }
}

# re-reduce data
# new data
meaningful <- c("subject", "right_hand", "left_hand") # select wanted columns
c<-which(names(EHI) %in% meaningful) #find colnumbers of unwanted
EHI <-EHI[,c] #remove unwanted columns
EHI <- na.omit(EHI) # remove NAs

# find sum for left and right for each participant 
# create hand data 
nsub <-length(levels(EHI$subject))
EHI2 <- data.frame(matrix(ncol = 3, nrow = nsub))
  colnames(EHI2) <- c("subject", "right_hand", "left_hand")

myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  myrow <- myrow+1
  subname <- levels(EHI$subject)[i] # find subject ID
  myrows <- which(EHI$subject==subname) # select rows for this subject
  tmp <- data.frame(EHI[myrows,])

    EHI2$subject[myrow] <- subname # add row for subject
    EHI2$right_hand[myrow] <- sum(tmp$right_hand) # add row for hand
    EHI2$left_hand[myrow] <- sum(tmp$left_hand) # add row for number of buttons pressed
  }

# This is a summary of the EHI
for_index2 <- 
  EHI2 %>% 
  group_by(subject) %>%
  mutate(index_EHI= ((right_hand-left_hand)/(right_hand + left_hand))*100)

# now create a data frame only for index 
EHI_index <- select(for_index2, "subject", "index_EHI")
# merge which demographics
demographics <- merge(demographics, EHI_index, by= "subject")

#now plot EHI in relation to handedness (modified by DB)
x<-ggplot(demographics, aes(x=index_EHI, y=subject, fill=subject)) +
  geom_dotplot(binaxis='y', stackdir='center') + xlim(-100, 100) +
  ggtitle("Handedness index on the Edinburgh Handedness Inventory")
x + theme_bw() + theme(legend.position = "none") + theme(axis.text.y = element_blank())

hist(demographics$index_EHI)
```

Next, the lexTale is preprocesssed. This is done using the weighted average for the test.

```{r lexTALE, warning=FALSE}
# read data
lexTALE_L <- read_csv("./LexTale_L.csv", na = "empty")
lexTALE_R <- read_csv("./lexTale_R.csv", na = "empty")
lexTALE <- rbind(lexTALE_R, lexTALE_L)
# filter
lexTALE <- lexTALE[lexTALE$Attempt == 1,]
lexTALE <- lexTALE[lexTALE$display == "Task",]
# rename variables
lexTALE$subject <- lexTALE$`Participant Public ID`
lexTALE$accuracy <- lexTALE$Correct
lexTALE$condition <- lexTALE$ANSWER
# code as factors
lexTALE$subject <- as.factor(lexTALE$subject)
lexTALE$condition <- as.factor(lexTALE$condition)
  levels(lexTALE$condition) <- c("non-word", "word")

# create data frame
lexTale_score <- data.frame(matrix(ncol = 4, nrow = nsub))
  colnames(lexTale_score) <- c("subject", "Word", "nonWord", "lexTALE")
myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  subname <- levels(lexTALE$subject)[i] # find subject ID
  myrows <- which(lexTALE$subject==subname) # select rows for this subject
  tmp <- data.frame(lexTALE[myrows,])

    myrow <- myrow+1

    lexTale_score$subject[myrow] <- subname
    lexTale_score$Word[myrow] <- sum(tmp[tmp$condition == "word",]$accuracy)
    lexTale_score$nonWord[myrow] <- sum(tmp[tmp$condition == "non-word",]$accuracy)
    lexTale_score$lexTALE[myrow] <- ((sum(tmp[tmp$condition == "word",]$accuracy)/40*100) + (sum(tmp[tmp$condition == "non-word",]$accuracy)/20*100)) / 2

}

# plot
x<-ggplot(lexTale_score, aes(x=lexTALE, y=subject, fill=subject)) +
  geom_dotplot(binaxis='y', stackdir='center') + xlim(0, 100) +
  ggtitle("Lextale") + geom_vline(xintercept = 80, linetype="dashed", color = "black", size=1.5)
x + theme_bw() + theme(legend.position = "none") + theme(axis.text.y = element_blank())

# now create a data frame only for index 
lexTALE <- select(lexTale_score, "subject", "lexTALE")
# now merge with demographics
demographics <- merge(demographics, lexTALE, by= "subject")
# now write the full demographics 
write.csv(demographics, file = "./demographics_out.csv")

hist(demographics$lexTALE)
abline(v=80,col='red')

paste0("No. of participants below LexTALE cuttoff: ", sum(demographics$lexTALE < 80))

# remove those who didn't pass the LexTALE
demographics <- demographics[demographics$lexTALE >= 80,]
```

# Read in raw data for day 1 and 2

Before doing anything with the data, we now pull in all the data from all nodes and combind them from a sub folder in the directory. Participants, identified by random letter strings, are stacked on top of each other. We start by reading the data and cutting out unwanted information.

```{r readraw,include= FALSE, warning=FALSE}
# Set directory
for (d in 1:2){
  data_file <- "./day1_combined.csv"

  if (d ==2){
    data_file <-  "./day2_combined.csv"
  }
  #we will assign raw data to all_dat2 (day2), but then reassign to all_dat1 on first run through loop, so we have correct data for each day
all_dat2 <- read.csv(data_file, na.strings = "NA",stringsAsFactors=F)
# cleaning - do operations that apply for all tasks
all_dat2 %>% filter(Attempt >= 1) # only answers 
# write as factors
all_dat2$subject <- as.factor(all_dat2$Participant.Public.ID)
all_dat2$RT <- as.numeric(as.character(all_dat2$Reaction.Time))
all_dat2$Trial <-as.numeric(all_dat2$Trial.Number) #useful if we want to restrict analysis to subset of trials so retain this variable
all_dat2$Task.Name <- as.factor(all_dat2$Task.Name)
all_dat2$word<-substring(all_dat2$word,66,(nchar(all_dat2$word)-4)) #remove format from written word stimuli
if (d==1)
{all_dat1 <- all_dat2}
}

# #remove cols from all_dat2 that contain 'branch' or 'randomiser'
# branchcols2 <- c(which(grepl('branch', names(all_dat2))=='TRUE'),which(grepl('randomiser', names(all_dat2))=='TRUE'))
# all_dat2 <- all_dat2[,-branchcols2]
# branchcols1 <- c(which(grepl('branch', names(all_dat1))=='TRUE'),which(grepl('randomiser', names(all_dat1))=='TRUE'))
# all_dat1 <- all_dat1[,-branchcols1]

#check that colnames are the same
# colcomp<-data.frame(colnames(all_dat2)[order(colnames(all_dat2))])
# colnames(colcomp)[1]<-'name1'
# ncol1 <-length(colnames(all_dat1))
# colcomp$name2<-NA
# colcomp$name2[1:ncol1]<-colnames(all_dat1)[order(colnames(all_dat1))]

#NB the colnames are identical but in different order, and all_dat1 has extra column of word_OG
#It is just conceivable the ordering might affect processing when columns are subsequently selected, so we'll reorder them to be consistent across sessions

all_dat2$word_OG<-NA #create dummy column just to make it easier to align columns
collist <- colnames(all_dat1)
all_dat2 <- all_dat2[,collist]

#Check for overlap in subjects between the two sessions
which(levels(all_dat1$subject) %in% levels(all_dat2$subject))
```

# Functions

This is a function that can apply to all tasks; user specifies the original file to start from, the wanted columns, and the specific task to select.

```{r genericprocessing} 
# This function just returns a subfile with just the necessary rows/cols
  procdata <- function(origfile,wanted,task,disp){
  nufile <- origfile[origfile$Task.Name==task,] #restrict consideration to this task
  nufile <- origfile[origfile$display==disp,] #restrict consideration to the correct display

  c<-which(names(origfile) %in% wanted) #find colnumbers of wanted
  #NB beware that the columns may not be in the same order as in the wanted list
  nufile <- nufile[,c]#select only wanted columns
  nufile <- na.omit(nufile) # remove NAs
  return(nufile)
}
```

The Hoaglin-Iglewicz function is used to remove outliers. We create an 'outlier' column that codes anticipations as 2, and long RT outliers as 1. All else is zero. The original RT is saved as allRT, and the RT column has NA for outliers, so they will be automatically excluded henceforth.

```{r RToutlierfunction}
remove.outliers <- function(origdat,lowcut,zcut){
# Outlier removal put in separate function, which can be applied to different task RTs.
# 
# origdat is the dataframe with data for outlier removal, mycolname is name of variable 
# for outlier removal (usually RT), lowcut is a cutoff value below which data excluded
# (typically v fast responses that are implausible, e.g. 201 ms for RT)
# and zcut is the constant in Hoaglin-Iglewicz method (usually 2.2)

# NB this function returns origdat with outlier rows marked, rather than removed
# This makes it easier to record the N outliers per subject.

# NB FUNCTION ASSUMES THAT THERE IS ARE COLUMNS CALLED subject, Correct and RT IN origfile.
# Outliers are computed by subject - NB contrary to previous version, these are defined
# across all trials, rather than separately for L and R.
origdat$allRT <- origdat$RT #we save a copy of RT, as we will be altering RT so that outliers become NA
origdat$outlier<-0
origdat$outlier[origdat$RT<lowcut] <- 2 #outlier col is coded 2 for anticipations
origdat$RT[origdat$RT<lowcut]<-NA #we exclude anticipations when computing quantiles
#Now extract correct RTs for each subject to compute limits for outliers

  for (i in 1:nsub) { # loop through subjects
  subname <- levels(origdat$subject)[i] # find subject ID
  myrows <- which(origdat$subject==subname) # select rows for this subject
  #NB myrows is NOT a consecutive series, because done in blocks
  firstrow <-min(myrows)
  tmp <- data.frame(origdat[myrows,])

  RTcorr <-tmp$RT[tmp$Correct==1] #correct RTs for this subject as a vector

 # Identify 25th and 75th quartiles, and the difference between them
          lower_quartile <- quantile(RTcorr, probs=0.25, na.rm="TRUE")
          upper_quartile <- quantile(RTcorr, probs=0.75, na.rm="TRUE")
          quartile_diff <- upper_quartile - lower_quartile
        # Outliers are defined as being below or above 2.2 times the quartile difference
          lower_limit <- lower_quartile - zcut*quartile_diff
          upper_limit <- upper_quartile + zcut*quartile_diff
        # create outlier variable
w1 <- which(origdat$RT[myrows]>upper_limit) #rows with outlier, nb RELATIVE to myrows range
origdat$outlier[myrows[w1]]<-1
}
 #For RT we just remove slow outliers, so ignore lower_limit
origdat$RT[origdat$outlier>0]<-NA #RT variable now has outliers excluded as NA
return(origdat)
}
```

This is a generic chunk of code that can be used for Chimeric faces and Rhyme Detection.

```{r maketaskdataframe, warning=FALSE}
make.df <- function(origdat,varlist,latlist){
nsub<-length(unique(origdat$subject))
# create compact data frame 
temp_dat <- data.frame(matrix(ncol = 4, nrow = nsub*2))
colnames(temp_dat) <- varlist
myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  subname <- levels(origdat$subject)[i] # find subject ID
  myrows <- which(origdat$subject==subname) # select rows for this subject
  latcol <- which(colnames(origdat) == varlist[2])
  RTcol <- which(colnames(origdat) == varlist[4])
  tmp <- data.frame(origdat[myrows,])

  for (j in 1:2) {
    myrow <- myrow+1
    w<- which(tmp[,latcol]== latlist[j]) # match on laterality
    tmp1 <- tmp[w,]
    tmp2 <- tmp1[tmp1$Correct == 1,]

    temp_dat$subject[myrow] <- subname # add row for subject
    temp_dat$side[myrow] <- latlist[j] # add row for side
    temp_dat$accurate[myrow] <- sum(tmp1$Correct,na.rm=TRUE) # add row for N accurate response
    temp_dat$p.corr[myrow] <- 100*mean(tmp1$Correct,na.rm=TRUE) # add row for %accurate responses - NB for chimeric only error is wrong emotion - measure is just a measure of choice. For RDT, can make error of selecting wrong side, so % correct is meaningful

    temp_dat$RT[myrow] <- mean(as.numeric(tmp2$RT),na.rm=TRUE) # add row for mean RT
    temp_dat$N[myrow] <- length(tmp1$RT) # add count of rows for this subject
  }
}
return(temp_dat)
}
```

This plots data and conducts t-tests for each task.

```{r plotdata_ttest, warning=FALSE}
dopirate <- function(origdat,task,measure){
# measure for each side
# plot the number of correct responses in each side
pirateplot(formula = X1 ~ side,
           data = origdat,
           main = measure,
           theme = 0,
           pal = "southpark", # southpark color palette
           bean.f.o = .0, # Bean fill
           point.o = .3, # Points
           inf.f.o = .7, # Inference fill
           inf.b.o = .8, # Inference border
           avg.line.o = 1, # Average line
           bar.f.o = .5, # Bar
           point.pch = 21,
           point.cex = .7,
           inf.method= "ci")
# plot L & R to get slope
q <- ggplot(data=origdat, aes(x=side, y=X1, group=subject, color=subject)) +
    geom_line() +
    geom_point() +
    ggtitle(paste0(task,": ",measure," by side"))
q + theme_bw() + theme(legend.position = "none")
#  t test
print(paste0(task,measure))
myt1<-t.test(origdat$X1~ origdat$side, paired = TRUE, alternative = "two.sided")
print(myt1)
}
```

Another generic function that can be used for different tasks to make a laterality index and write it to the demographics file

```{r makeLIs, warning=FALSE}
# This function modifies a copy of the demographics file and returns it

# User selects whether accuracy or RT by specifying col X1 in origdat when calling the function

# Polarity is -1 or 1 and can be set to keep all LIs positive if in a given direction
# Cutoff will create NA if level of performance is outside predetermined limits

makeLI <- function(demogs,origdat,task,mycolname,polarity,mycutoff){
# we do this for accuracy first
# reformat data frame
keeps <- c("subject", "side", "X1") #X1 holds either accuracy or RT
LIdf <- origdat[keeps]
LIdf <- spread(LIdf, side, X1)
# calculate each participants' lat index
LIdf <- 
  LIdf %>% 
  group_by(subject) %>%
  mutate(thisLI= ((Right - Left)/(Right + Left))*100)

LIdf$LI <- LIdf$thisLI*polarity #polarity can be set to reverse sign - eg for RT a large score is bad, whereas for acc it is good, so can have opp polarity

# check if accuracy or RT level is outside acceptable cutoff.
# Do this based on average for left and right
# LI will be set to NA for those outside limits
LIdf$avg <- (LIdf$Left+LIdf$Right)/2
w<-which(LIdf$avg<mycutoff)
if(length(w)>0){
LIdf$LI[w] <- NA}

# now create a data frame for lat index 
LIdf <- select(LIdf, "subject", "LI")
# now merge with demographics
demogs <- merge(demogs, LIdf, by= "subject")
lastcol<-length(colnames(demogs))
colnames(demogs)[lastcol]<-mycolname

return(demogs)
}
```

Make Z score LI.
NB this is NOT a conventional z-score based on whole sample, but rather is an individual-based z-score that indicates whether the person's L-R difference is significantly different from chance.
So you either compare their L-sided RTs and their R-sided RTs, or the proportions correct on L vs proportion correct on R, to see if there is bias away from zero in that individual.
Note that this in effect adjusts for any overall differences in accuracy and RT, as the person is compared with themselves.

```{r makeZ, warning=FALSE}
# This function modifies a copy of the demographics file and returns it

# User selects the measure by specifying col X1 in origdat when calling the function

# Polarity is -1 or 1 and can be set to keep all LIs positive if in a given direction
# Cutoff will create NA if level of performance is outside predetermined limits

makeZ <- function(demogs,origdat,task,mycolname,polarity,mycutoff){
# we do this for accuracy first
# reformat data frame
keeps <- c("subject", "side", "X1") #X1 holds either accuracy or RT
Zdf <- origdat[keeps]
Zdf <- spread(Zdf, side, X1)
# calculate each participants' lat index
Zdf <- 
  Zdf %>% 
  group_by(subject) %>%
  mutate(n= Left + Right, 
         pR= Right/n,
         pL= Left/n,
         Z= (pR-.5)/sqrt(pR*pL/n))
# remove infinite results
is.na(Zdf) <- do.call(cbind,lapply(Zdf, is.infinite))
Zdf <- na.omit(Zdf)

Zdf$Z <- Zdf$Z*polarity #polarity can be set to reverse sign - eg for RT a large score is bad, whereas for acc it is good, so can have opp polarity

# now create a data frame for lat index 
Zdf <- select(Zdf, "subject", "Z")
# now merge with demographics
demogs <- merge(demogs, Zdf, by= "subject")
lastcol<-length(colnames(demogs))
colnames(demogs)[lastcol]<-mycolname

return(demogs)
}
```

## Chimeric Face Task

This part of the R Markdown file prcoesses the "Chimeric Face Task' implmented online via Gorilla.sc which aims to assess the lateralisation of emotion recognition in a face processing task. In the task, participants are shown a chimeric face. Participants then respond by indicating the emotion that they felt was most strong expressed.

This task is near identical to that reported in Karlsson, Johnstone, and Carey (2019). Stimuli were kindly provided by Dr Michael Burt (https://www.dur.ac.uk/psychology/staff/?id=1942) and were symmetrical average images created from four male and four female faces (see Burt & Perrett, 1997; Innes et al., 2016).

```{r read_face_dat, warning=FALSE}
for (d in 1:2){ #we will process day 1 and day 2 together
  origfile <- all_dat1
  if (d==2)
  {origfile <- all_dat2}
#As with earlier reading in, we default to saving CF_dat2 (day2), but at end of the chunk we reassign to CF_day1 in the first run through the d loop
task <- "Chimeric faces"
wanted <- c("subject", "stimuli", "Response", "RT", "l_hemiface", "r_hemiface") # select wanted columns
disp <- "e_Task"
CF_dat2 <- procdata(origfile,wanted,task,disp) #use generic function (see above) to read in relevant columns

# count subjects
nsub <-length(levels(CF_dat2$subject))
print(paste('day',d,task,": subjects: ", nsub))

CF_same2 <- CF_dat2[CF_dat2$l_hemiface == CF_dat2$r_hemiface,] # create data frame for non-chimeras
CF_dat2 <- CF_dat2[CF_dat2$l_hemiface != CF_dat2$r_hemiface,] # remove items where same expression in both hemifaces
CF_dat2 <- remove.outliers(CF_dat2,200,1.65) #use generic outlier removal function
outliertable<-table(CF_dat2$subject,CF_dat2$outlier)
#print(outliertable)

if (d==1)
{CF_dat1 <- CF_dat2
 CF_same1 <- CF_same2}
}
```

First, we judge participants' ability to correctly identify emotions using the "CF_same" dataframe. Participants are removed if their performance is below 75% due to poor ability to identify emotions that will inevitably influence performance on the Chimeric Face Task.

```{r emotions, warning=FALSE,message=FALSE}
# this will mark correct responses as 1 or 0. It will be importantfor removing participants later on. 
# NB we loop through same process for day1 and day2
for (d in 1:2){
  CF_same <- CF_same1
  if (d==2){
    CF_same <- CF_same2
  }

CF_same <- 
  CF_same %>% 
  mutate(emotion_recog= ifelse(as.character(Response) == as.character(l_hemiface), 1, 0))
# calculate participants average
emotion_mean <- aggregate(FUN= mean, data= CF_same, emotion_recog~ subject)
# add this to demographics 
demographics <- merge(demographics, emotion_mean, by= "subject")
}
mycol <- length(colnames(demographics))
colnames(demographics)[(mycol-1):mycol]<-c('Emot.recog1','Emot.recog2')

plot(demographics$Emot.recog1,demographics$Emot.recog2)
abline(h=.75)
abline(v=.75)
text(.6,.8,'drop if <75% on either run',cex=.8)

#db modified to mark rather than drop cases with scores < .75
demographics$exclude <- 0 #default is to exclude

# mark the cases where there is less than 75% for identical stimuli
w <- union(which(demographics$Emot.recog1 < .75),which(demographics$Emot.recog2 < .75))
demographics$exclude[w]<-1 #code exclude = 1 if excluded because poor emot recog
```

In this section, the correct sequences are marked as 1, incorrect asnwers are marked as 0. We then mark whether the correct answers were in the left or right hemiface.

```{r mark_face_correct_face, warning=FALSE}
# this will mark correct responses as 1 or 0.
# NB we again loop through same process for day1 and day2
for (d in 1:2){
  CF_dat <- CF_dat1
  if (d==2){
    CF_dat <- CF_dat2
  }
CF_dat <- 
  CF_dat %>% 
  mutate(Correct= ifelse(as.character(Response) == as.character(l_hemiface), 1, 
                         ifelse(as.character(Response) == as.character(r_hemiface), 1, 0)))
# this will mark the side responded to. NB to allow for generic functions for all tasks, we use the label 'side' rather than hemiface
CF_dat <- 
  CF_dat %>% 
  mutate(side= ifelse(as.character(Response) == as.character(l_hemiface) & Correct == 1, "Left",
                       ifelse(as.character(Response) == as.character(r_hemiface) & Correct == 1, "Right", NA)))
# code side as factor
CF_dat$side <- as.factor(CF_dat$side)

  if(d==1){
    CF_dat1 <- CF_dat}
  if (d==2){
    CF_dat2 <- CF_dat
  }
}
```

Now the data is ready, create an empty data frame and populate this. It includes columns for the participant identifier, side, and count of accurate responses for each side.

```{r makedataframe, warning=F}
#We create chim_dat files for day1 and day 2
varlist <-   c("subject", "side", "accurate", "RT") #need to be in this order, ie sub, side, acc and RR
origdat <- CF_dat1
latlist <- c("Left","Right") #names of factor levels

chim_dat1 <- make.df(CF_dat1,varlist,latlist)
chim_dat2 <- make.df(CF_dat2,varlist,latlist)

w<-which(is.na(chim_dat1$subject))
if (length(w)>0) {chim_dat1<-chim_dat1[-w,]}
w<-which(is.na(chim_dat2$subject))
if (length(w)>0) {chim_dat2<-chim_dat2[-w,]}
```

The data is visualised using the pirate plot function and t.test conducted.

```{r dosummary}
origdat <- chim_dat1
measure <- 'Accuracy'
origdat$X1 <- chim_dat1$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Chimeric faces day 1'
dopirate(origdat,task,measure)
origdat$X1 <- chim_dat1$RT
measure <- 'RT'
dopirate(origdat,task,measure)
t.test(origdat$accurate~ origdat$side)
#Now repeat for day 2
origdat <- chim_dat2
origdat$X1 <- chim_dat2$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Chimeric faces day 2'
measure <- 'Accuracy'
dopirate(origdat,task,measure)
origdat$X1 <- chim_dat2$RT
measure <- 'RT'
dopirate(origdat,task,measure)
```

Finally, a laterality index is calculated and plotted based on the equation [(R-L)/(R+L)] x 100, where L= left side and R= right side.

If the index is negative, emotion recognition is lateralised in the left hemisphere. If the index is positive, emotion recognition is lateralised in the right hemisphere

Here, a .csv is written for the chimeric face LI.

```{r domakeLI}
#We add Chimeric faces LIs to demographics for day 1 and then day2
task<-'Chimeric faces'
origdat<-chim_dat1
origdat$X1 <- origdat$accurate
mycolname<-"Day1_CF_acc_LI"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (-1)#so LI is positive in predicted direction
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

# now do RT
origdat$X1 <- origdat$RT
mycolname<-"Day1_CF_RT_LI"
polarity <-1

#if we don't specify mycutoff, it will remain at level set for accuracy, which means it should not trap any RT values as they will all be much bigger
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

origdat<-chim_dat2
origdat$X1 <- origdat$accurate
mycolname<-"Day2_CF_acc_LI"
polarity<- (-1)#so LI is positive in predicted direction
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
origdat$X1 <- origdat$RT
mycolname<-"Day2_CF_RT_LI"
polarity <-1
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

cor1 <- cor(demographics$Day1_CF_acc_LI,demographics$Day2_CF_acc_LI)
plot(demographics$Day1_CF_acc_LI,demographics$Day2_CF_acc_LI,main='Chimeric LI accuracy',col=(1+demographics$exclude))
abline(v=0)
abline(h=0)
text(-50,75,paste0('r = ',round(cor1,3)))
text(-50,65,'red will be excluded',cex=.85)
cor2<-cor(demographics$Day1_CF_RT_LI,demographics$Day2_CF_RT_LI)
plot(demographics$Day1_CF_RT_LI,demographics$Day2_CF_RT_LI,main='Chimeric LI RT',col=(1+demographics$exclude))
abline(v=0)
abline(h=0)
text(-10,20,paste0('r = ',round(cor2,3)))
text(-10,18,'red will be excluded',cex=.85)
```

Here we use the Z score version.

```{r domakeZ}
#We add Chimeric faces LIs to demographics for day 1 and then day2
task<-'Chimeric faces'
origdat<-chim_dat1
origdat$X1 <- origdat$N
mycolname<-"Day1_CF_acc_Z"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (-1)#so LI is positive in predicted direction
demographics <- makeZ(demographics,origdat,task,mycolname,polarity,mycutoff)

origdat<-chim_dat2
origdat$X1 <- origdat$N
mycolname<-"Day2_CF_acc_Z"
polarity<- (-1)#so LI is positive in predicted direction
demographics <- makeZ(demographics,origdat,task,mycolname,polarity,mycutoff)
w<-which(demographics$excluded == 1)
demographics$Day1_CF_acc_Z[w] <- NA #removes zscores for excluded cases
demographics$Day2_CF_acc_Z[w] <- NA

RR<-intersect(which(demographics$Day1_CF_acc_Z>1.96),which(demographics$Day2_CF_acc_Z>1.96))
LL<-intersect(which(demographics$Day1_CF_acc_Z< -1.96),which(demographics$Day2_CF_acc_Z< -1.96))

cor1 <- cor(demographics$Day1_CF_acc_Z,demographics$Day2_CF_acc_Z,use='complete.obs',method='spearman')
plot(demographics$Day1_CF_acc_Z,demographics$Day2_CF_acc_Z,main='Chimeric Z accuracy')
abline(v=0)
abline(h=0)
abline(v = 1.96, col='red')
abline(v = -1.96, col='red')
abline(h = 1.96, col='red')
abline(h = -1.96, col='red')
text(-7,7.5,paste0('r = ',round(cor1,3)),cex=.8)
text(-2,-15,paste('Those in bottom quadrant\n bounded by red are consistently\n L-lateralised, N = ',length(LL)),cex=.8)
text(5,18,paste('Those in top quadrant\n bounded by red are consistently\n R-lateralised, N = ',length(RR)),cex=.8)
text(8,-8,'One outlier excluded',cex=.8)

```

As we can see, we generally get a left side/right hemisphere advantage in our accuracy data. The RT advantage is not as expected. This may be due to the buttons being clicked on screen and the additional time penalty of moving the mouse.

## Bergen Dichotic listening

This part of the R Markdown file prcoesses the "Dichotic listening task' implmented online via Gorilla.sc which aims to assess the lateralisation of receptive language lateralisation. In the task, participants hear a different CV sound in each ear. Participants respond to the sound they heard clearest

This task is near identical to that reported in Karlsson, Johnstone, and Carey (2019) and were original described in Hugdahl, Westerhausen, Alho, Medvedev, Laine, & Hämäläinen (2009). Stimuli were kindly provided by Dr Emma Karlsson (https://research.bangor.ac.uk/portal/en/researchers/emma-karlsson(17f6f06c-8aa9-4d16-93ba-4f98f285297e).html) and permission to use these stimuli was given by Professor Kenneth Hughdahl (https://www.uib.no/en/persons/Kenneth.Hugdahl).

```{r read_dichotic_dat, warning=FALSE}
for (d in 1:2){ #we will process day 1 and day 2 together
  origfile <- all_dat1
  if (d==2)
  {origfile <- all_dat2}
# lowercase for responses
origfile$left_channel <- tolower(origfile$left_channel)
origfile$right_channel <- tolower(origfile$right_channel)
# make meaningful stimuli name
origfile$stimuli <- paste0(origfile$left_channel, "-", origfile$right_channel)
#As with earlier reading in, we default to saving dichotic_dat2 (day2), but at end of the chunk we reassign to CF_day1 in the first run through the d loop
task <- "Dichotic Listening"
wanted <- c("subject", "stimuli", "Response", "RT", "left_channel", "right_channel") # select wanted columns
disp <- "e_task"
dichotic_dat2 <- procdata(origfile,wanted,task,disp) #use generic function (see above) to read in relevant columns

# count subjects
nsub <-length(levels(dichotic_dat2$subject))
print(paste('day',d,task,": subjects: ", nsub))

dichotic_same2 <- dichotic_dat2[dichotic_dat2$left_channel == dichotic_dat2$right_channel,] # create data frame for non-chimeras
dichotic_dat2 <- dichotic_dat2[dichotic_dat2$left_channel != dichotic_dat2$right_channel,] # remove items where same expression in both hemifaces
dichotic_dat2 <- remove.outliers(dichotic_dat2,200,1.65) #use generic outlier removal function
outliertable<-table(dichotic_dat2$subject,dichotic_dat2$outlier)
#print(outliertable)

if (d==1)
{dichotic_dat1 <- dichotic_dat2
 dichotic_same1 <- dichotic_same2}
}
```

First, we judge participants' ability to correctly identify sounds using the "dichotic_same" dataframe. Participants are removed if their performance is below 75%.

```{r soundsability, warning=FALSE,message=FALSE}
# this will mark correct responses as 1 or 0. It will be importantfor removing participants later on. 
# NB we loop through same process for day1 and day2
for (d in 1:2){
  dichotic_same <- dichotic_same1
  if (d==2){
    dichotic_same <- dichotic_same2
  }

dichotic_same <- 
  dichotic_same %>% 
  mutate(sound_recog= ifelse(as.character(Response) == as.character(left_channel), 1, 0))
# calculate participants average
sound_mean <- aggregate(FUN= mean, data= dichotic_same, sound_recog~ subject)
# add this to demographics 
demographics <- merge(demographics, sound_mean, by= "subject")
}
mycol <- length(colnames(demographics))
colnames(demographics)[(mycol-1):mycol]<-c('sound.recog1','sound.recog2')

# once again, mark rather than remove those with poor performance on identical stimuli
w<-union(which(demographics$sound.recog1 < .75), which(demographics$sound.recog2 < .75))
ww<-intersect(which(demographics$exclude==1),w)
#ww contains people who were excluded on both dichotic and faces. 
# code exclude = 2 for the cases where there less than 75% for identical stimuli
demographics$exclude[w]<-2 #code exclude = 2 if excluded because poor dichotic
demographics$exclude[ww]<-12 #code for those excluded on both chimeric and dichotic
table(demographics$exclude) 
```

In this section, the correct sequences are marked as 1, incorrect asnwers are marked as 0. We then mark whether the correct answers were in the left or right ear.

```{r mark_dich_correct_ear, warning=FALSE}
# this will mark correct responses as 1 or 0.
# NB we again loop through same process for day1 and day2
for (d in 1:2){
  dichotic_dat <- dichotic_dat1
  if (d==2){
    dichotic_dat <- dichotic_dat2
  }
dichotic_dat <- 
  dichotic_dat %>% 
  mutate(Correct= ifelse(as.character(Response) == as.character(left_channel), 1, 
                         ifelse(as.character(Response) == as.character(right_channel), 1, 0)))
# this will mark the side responded to. NB to allow for generic functions for all tasks, we use the label 'side' rather than hemiface
dichotic_dat <- 
  dichotic_dat %>% 
  mutate(side= ifelse(as.character(Response) == as.character(left_channel) & Correct == 1, "Left",
                       ifelse(as.character(Response) == as.character(right_channel) & Correct == 1, "Right", NA)))
# code side as factor
dichotic_dat$side <- as.factor(dichotic_dat$side)

  if(d==1){
    dichotic_dat1 <- dichotic_dat}
  if (d==2){
    dichotic_dat2 <- dichotic_dat
  }
}
```

Now the data is ready, create an empty data frame and populate this. It includes columns for the participant identifier, side, and count of accurate responses for each side.

```{r makedataframedich, warning=F}
#We create DL_dat files for day1 and day 2
varlist <-   c("subject", "side", "accurate", "RT") #need to be in this order, ie sub, side, acc and RR
origdat <- dichotic_dat1
latlist <- c("Left","Right") #names of factor levels

DL_dat1 <- make.df(dichotic_dat1,varlist,latlist)
DL_dat2 <- make.df(dichotic_dat2,varlist,latlist)

w<-which(is.na(DL_dat1$subject))
if (length(w)>0) {DL_dat1<-DL_dat1[-w,]}
w<-which(is.na(DL_dat2$subject))
if (length(w)>0) {DL_dat2<-DL_dat2[-w,]}

DL_dat1[is.na(DL_dat1)] <- 0
DL_dat2[is.na(DL_dat2)] <- 0
```

The data is visualised using the pirate plot function and t.test conducted.

```{r dosummarydich}
origdat <- DL_dat1
measure <- 'Accuracy'
origdat$X1 <- DL_dat1$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Dichotic listening day 1'
dopirate(origdat,task,measure)
origdat$X1 <- DL_dat1$RT
measure <- 'RT'
dopirate(origdat,task,measure)
t.test(origdat$accurate~ origdat$side)
#Now repeat for day 2
origdat <- DL_dat2
origdat$X1 <- DL_dat2$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Dichotic listening day 2'
measure <- 'Accuracy'
dopirate(origdat,task,measure)
origdat$X1 <- DL_dat2$RT
measure <- 'RT'
dopirate(origdat,task,measure)
```

Finally, a laterality index is calculated and plotted based on the equation [(R-L)/(R+L)] x 100, where L= left side and R= right side.

If the index is negative, emotion recognition is lateralised in the left hemisphere. If the index is positive, emotion recognition is lateralised in the right hemisphere

Here, a .csv is written for the chimeric face LI.

```{r domakeLIdich}
#We add Dichotic LIs to demographics for day 1 and then day2
task<-'Dichotic listening'
origdat<-DL_dat1
origdat$X1 <- origdat$accurate
mycolname<-"Day1_DL_acc_LI"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (-1)#so LI is positive in predicted direction
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

# now do RT
origdat$X1 <- origdat$RT
mycolname<-"Day1_DL_RT_LI"
polarity <-1

#if we don't specify mycutoff, it will remain at level set for accuracy, which means it should not trap any RT values as they will all be much bigger
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

origdat<-DL_dat2
origdat$X1 <- origdat$accurate
mycolname<-"Day2_DL_acc_LI"
polarity<- (-1)#so LI is positive in predicted direction
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
origdat$X1 <- origdat$RT
mycolname<-"Day2_DL_RT_LI"
polarity <-1
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

cor1 <- cor(demographics$Day1_DL_acc_LI,demographics$Day2_DL_acc_LI)
plot(demographics$Day1_DL_acc_LI,demographics$Day2_DL_acc_LI,main='Dichotic LI accuracy',col=(1+demographics$exclude))
abline(v=0)
abline(h=0)
text(-75,75,paste0('r = ',round(cor1,3)))
cor2<-cor(demographics$Day1_DL_RT_LI,demographics$Day2_DL_RT_LI)
plot(demographics$Day1_DL_RT_LI,demographics$Day2_DL_RT_LI,main='Dichotic LI RT',col=(1+demographics$exclude))
abline(v=0)
abline(h=0)
text(-5,15,paste0('r = ',round(cor2,3)))
```

Here we use the Z score version. This is an individualised score that directly indicates whether or not the person is reliably lateralised on one test occasion.

```{r domakeZ_dich}
#We add dichotic LIs to demographics for day 1 and then day2
task<-'Dichotic listening'
origdat<-DL_dat1
origdat$X1 <- origdat$N
mycolname<-"Day1_Dich_acc_Z"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (-1)#so LI is positive in predicted direction
demographics <- makeZ(demographics,origdat,task,mycolname,polarity,mycutoff)

origdat<-DL_dat2
origdat$X1 <- origdat$N
mycolname<-"Day2_Dich_acc_Z"
polarity<- (-1)#so LI is positive in predicted direction
demographics <- makeZ(demographics,origdat,task,mycolname,polarity,mycutoff)

#convert excluded cases with code 2 or 12 to NA
w<-which(demographics$exclude>1)
demographics$Day1_Dich_acc_Z[w]<-NA
demographics$Day2_Dich_acc_Z[w]<-NA

RR<-intersect(which(demographics$Day1_Dich_acc_Z>1.96),which(demographics$Day2_Dich_acc_Z>1.96))
LL<-intersect(which(demographics$Day1_Dich_acc_Z< -1.96),which(demographics$Day2_Dich_acc_Z< -1.96))

cor1 <- cor(demographics$Day1_Dich_acc_Z,demographics$Day2_Dich_acc_Z,use='complete.obs',method='spearman')
demographics$handpch <-1
demographics$handpch[demographics$handedness=='Left'] <-2
plot(demographics$Day1_Dich_acc_Z,demographics$Day2_Dich_acc_Z,main='Dich Z accuracy',pch=demographics$handpch)
abline(v=0)
abline(h=0)
text(-7.5,15,paste0('Spearman r = ',round(cor1,3)),cex=.8)
abline(v = 1.96, col='red')
abline(v = -1.96, col='red')
abline(h = 1.96, col='red')
abline(h = -1.96, col='red')

text(20,-15,paste('Those in bottom quadrant\n bounded by red are consistently\n L-lateralised, N = ',length(LL)),cex=.8)
text(5,40,paste('Those in top quadrant\n bounded by red are consistently\n R-lateralised, N = ',length(RR)),cex=.8)

```

## Rhyme Detection Task

This portion of the R Markdown file prcoesses the "Rhyme Detection Task' implmented online via Gorilla.sc which aims to assess the lateralisation of language processing. In the task, participants are shown a word in the central visual field and different images in the left and right visual field. An arrow appears pointing to either image. Participants then respond by indicating whether the image and word rhyme.

This reads in a csv file where each row represents a response to a trial. Participants, identified by random letter strings, are stacked on top of each other. We start by reading the data and cutting out unwanted information.

```{r read_dat_RDT, warning=FALSE}
for (d in 1:2){ #same script for days 1 and 2

origfile <- all_dat1
if(d==2)
  {origfile <- all_dat2}
task <- "Rhyme Detection Task"
disp <- "Task"

# nb when creating new file, we default to name 'RDT_dat2', but this is reassigned to 'RDT_dat1' at end of processing for day 1.

# new data
wanted <- c("Trial","subject", "Correct", "RT", "ANSWER","Response") # select wanted columns
RDT_dat2 <- procdata(origfile,wanted,task,disp) #Here we reuse the procdata function that we developed for processing Chimeric faces data - just does some generic processing

# do some wrangling on the ANSWER variable to make it compatible with functions
RDT_dat2<-filter(RDT_dat2,ANSWER %in% c('left','right'))
c <- which(colnames(RDT_dat2)=="ANSWER")
colnames(RDT_dat2)[c]<-'side'
RDT_dat2$side<-as.factor(RDT_dat2$side)
  levels(RDT_dat2$side)<-c("Left","Right")
RDT_dat2<-remove.outliers(RDT_dat2,200,1.65) #numbers specify min RT and zscore for Hoaglin-Iglewicz respectively

# count subjects
nsub <-length(levels(RDT_dat2$subject))
print(paste0("Day 1 subjects: ", nsub))
outliertable<-table(RDT_dat2$subject,RDT_dat2$outlier)
#print(outliertable)

# now remove the cases where there are too many or to few cases
#remove.outlier <- as.data.frame(outliertable)
#remove.outlier <- remove.outlier[remove.outlier$Freq >= 260 & remove.outlier$Freq <= 208,]
#remove.p <- list(unique(remove.outlier$Var1))
#demographics <- demographics[demographics$subject != remove.p,]

if(d==1)
{RDT_dat1 <- RDT_dat2}

}
```

Now the data is ready, create an empty data frame and populate this. It includes columns for the participant identifier, rhyme, acuracy, RT (RT only for correct answers).

```{r makedataframeRDT, warning=F}
#We create files for day1 and day 2
# NB warnings created because of different Ns for time 1 and time 2, which means there are blanks in subjects
# We deal with this below by removing rows with no subject code
varlist <-   c("subject", "side", "accurate", "RT") #need to be in this order, ie sub, side, acc and RR - we will use these in the make.df function
origdat <- RDT_dat1
latlist <- c("Left","Right") #names of factor levels
RDT_df1 <- make.df(origdat,varlist,latlist)
origdat <- RDT_dat2
RDT_df2 <- make.df(origdat,varlist,latlist)

# Remove blank rows 
w<-which(is.na(RDT_df1$subject))
if (length(w)>0) {RDT_df1<-RDT_df1[-w,]}
w<-which(is.na(RDT_df2$subject))
if (length(w)>0) {RDT_df2<-RDT_df2[-w,]}

# Exclude cases with low % correct
# NB something weird with these - seem to be ones with too many trials, suggesting some problem with multiple responses
w<-which(RDT_df1$p.corr<50)
if (length(w)>0)
{RDT_df1$accurate[w]<-NA
RDT_df1$RT[w]<-NA
}
w<-which(RDT_df2$p.corr<50)
if (length(w)>=0)
{RDT_df2$accurate[w]<-NA
RDT_df2$RT[w]<-NA
}

# now remove the cases where there are NA for only one visual field
# this will remove those who responded by pressing only one button.
RDT_df1 <- na.omit(RDT_df1) %>% 
  group_by(subject) %>%
  mutate(no_rows = length(subject))
RDT_df1 <- RDT_df1[RDT_df1$no_rows == 2,]

RDT_df2 <- na.omit(RDT_df2) %>% 
  group_by(subject) %>%
  mutate(no_rows = length(subject))
RDT_df2 <- RDT_df2[RDT_df2$no_rows == 2,]
```

```{r checker, warning=FALSE}
for (d in 1:2){
origdat <-RDT_df1
if (d==2)
{origdat <- RDT_df2}
measure <- 'Accuracy'
origdat$X1 <- origdat$accurate
task <- 'Rhyme Detection'

dopirate(origdat,task,measure)
origdat$X1 <- origdat$RT
measure <- 'RT'
dopirate(origdat,task,measure)

#now create LI
origdat$X1 <- origdat$p.corr
mycolname<-"RDT_p.corr_LI"
polarity<- 1 #so LI is positive in predicted direction
mycutoff <- 75
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
origdat$X1 <- origdat$RT
mycolname<-"RDT_RT_LI"
polarity <- (-1)
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

#Cutoff has removed those with low accuracy, but we should also remove their RT LIs
#We do this by finding those with NA in accuracy LI
w<-which(is.na(demographics$RDT_acc_LI))
if(length(w)>0)
{demographics$RDT_RT_LI[w]<-NA}
#t.test to test if laterality index differs from zero
t.test(demographics$RDT_p.corr_LI)
t.test(demographics$RDT_RT_LI)
#rename column to identify day
myncol <- length(colnames(demographics))
colnames(demographics)[(myncol-1):myncol] <- paste0('Day',d,'.',colnames(demographics)[(myncol-1):myncol])

}
cor1<-cor(demographics$Day1.RDT_p.corr_LI,demographics$Day2.RDT_p.corr_LI,use='complete.obs')
cor2<-cor(demographics$Day1.RDT_RT_LI,demographics$Day2.RDT_RT_LI,use='complete.obs')
plot(demographics$Day1.RDT_p.corr_LI,demographics$Day2.RDT_p.corr_LI,main='Rhyme detect: LI for accuracy (%correct)')
abline(v=0)
abline(h=0)
text(-3,3,paste0('r = ',round(cor1,3)))
plot(demographics$Day1.RDT_RT_LI,demographics$Day2.RDT_RT_LI,main='Rhyme detect: LI for RT')
abline(v=0)
abline(h=0)
text(-3,3,paste0('r = ',round(cor2,3)))
```

### T-test as alternative to LI for RDT

```{r do.ttest}
for (d in 1:2){
  myfile<-RDT_dat1
  if(d==2){
    myfile<-RDT_dat2}

  nsubs <- nrow(demographics)
  demographics$tlat_RDT<-NA
   demographics$plat_RDT<-NA
  for (i in 1:nsubs) {
    w<-which(myfile$subject==demographics$subject[i])
    if(length(w)>0)
    {
      myt<-t.test(myfile$RT[w]~myfile$side[w])
      demographics$tlat_RDT[i]<-myt$statistic
      demographics$plat_RDT[i]<-myt$p.value
    }
  }
   if (d==1){
     maxcol <-length(colnames(demographics))
     colnames(demographics)[(maxcol-1):maxcol]<-c('tlat_RDT_1','plat_RDT_1')
   }
   if (d==2){
     maxcol <-length(colnames(demographics))
     colnames(demographics)[(maxcol-1):maxcol]<-c('tlat_RDT_2','plat_RDT_2')
   }
}
colcount<-length(colnames(demographics))
# Remove values for cases that are excluded for low acc: do this for both times by just blitzing last 4 columns of demographics
w<-which(is.na(demographics$Day1.RDT_RT_LI))
demographics[w,(colcount-3):colcount]<-NA
w<-which(is.na(demographics$Day2.RDT_RT_LI))
demographics[w,(colcount-3):colcount]<-NA

# compare t-value with LI : use t = 1.65 as cutoff for sig lateralised
 plot(demographics$tlat_RDT_1,demographics$Day1.RDT_RT_LI)
 abline(v=1.65)
  plot(demographics$tlat_RDT_2,demographics$Day2.RDT_RT_LI)
   abline(v=1.65)
# compare t-value from day 1 and day 2 
     myr <- cor(demographics$tlat_RDT_1,demographics$tlat_RDT_2,use='complete.obs')
    plot(demographics$tlat_RDT_1,demographics$tlat_RDT_2)
    abline(v=0)
    abline(h=0)
   abline(v=1.65) 
   abline(h=1.65)
      abline(v=-1.65) 
   abline(h=-1.65)
   text(-3,4,paste0('r = ',round(myr,3)))
```

## Finger Tapping Task

This portion of the R Markdown file prcoesses the "Finger Tapping Task' implmented in Gorilla.sc which aims to quantify handedness. In the task, participants are required to press a sequence of buttons twice for each hand (Left: W, R, V, X; Right: T, U, M, B) as many times as possible within a 30 ms. There is a version that requires more precise movements.

A score of 1 is given each time that a participants completes the sequence.

This reads in a csv file where each row represents a button press. Participants, identified by random letter strings, are stacked on top of each other. We start by reading the data and cutting out unwanted information.

NOTE. The create data frame function disrupts order here.

In this section, the correct sequences are marked (e.g. pressing Y, U, J, H, in order will correspond to a score of 1).

```{r readdataFinger, warning=FALSE}

finger <- read.csv("day1_combined.csv", na.strings = "NA")

# cleaning
finger <- finger[finger$Attempt >= 1,] # only answers 
finger <- finger[finger$display == "right hand" | finger$display == "left hand",] # only for task
# write as factors
finger$display <- as.factor(finger$display) 
finger$subject <- as.factor(finger$Participant.Public.ID)
finger$RT <- as.numeric(as.character(finger$Reaction.Time))

# new data
wanted <- c("subject", "display", "Response", "RT") # select wanted columns
c<-which(names(finger) %in% wanted) #find colnumbers of unwanted
finger <-finger[,c] #remove unwanted columns
finger <- na.omit(finger) # remove NAs

all_finger1<- finger

finger <- read.csv("day2_combined.csv", na.strings = "NA")

# cleaning
finger <- finger[finger$Attempt >= 1,] # only answers 
finger <- finger[finger$display == "right hand" | finger$display == "left hand",] # only for task
# write as factors
finger$display <- as.factor(finger$display) 
finger$subject <- as.factor(finger$Participant.Public.ID)
finger$RT <- as.numeric(as.character(finger$Reaction.Time))

# new data
wanted <- c("subject", "display", "Response", "RT") # select wanted columns
c<-which(names(finger) %in% wanted) #find colnumbers of unwanted
finger <-finger[,c] #remove unwanted columns
finger <- na.omit(finger) # remove NAs

all_finger2<- finger

```

```{r readdataFinger2, warning=FALSE}
for (d in 1:2){ #we will process day 1 and day 2 together
  origfile <- all_finger1
  if (d==2)
  {origfile <- all_finger2}
#As with earlier reading in, we default to saving finger2 (day2), but at end of the chunk we reassign to finger1 in the first run through the d loop
finger2 <- origfile

finger2$RTbase[2:nrow(finger2)]<-finger2$RT[1:(nrow(finger2)-1)]
finger2$RTa <- finger2$RT-finger2$RTbase
#remove negative or v long RTs - can indicate when block changes etc
w<-c(which(finger2$RTa<0),which(finger2$RTa>1000))
finger2$RTa[w]<-NA
if (d==1)
{finger1 <- finger2
 }
}
```

In this section, the correct sequences are marked (e.g. pressing Y, M, I, B, in order will correspond to a score of 1).

Added exploratory analysis of SDs of RTs for keypress rather than means, but this was not particularly illuminating.

```{r scorefinger}
# We can go directly to creation of a data.frame by subjects, and use string matching to identify N correct seqs

myseq <-c('tumb','wrvx') # R, # L

subids <- levels(finger1$subject) #just look at data for those doing time1 and
nsub <- length(subids)

finger.df <- data.frame(matrix(NA,nrow=nsub,ncol=9))
colnames(finger.df)<-c('subject', 'R1', 'R2', 'L1', 'L2', 'R1sd', 'R2sd', 'L1sd', 'L2sd')
for (n in 1:nsub){
  finger.df$subject[n] <- subids[n]
  myc <- 1 #will be incremented so sums written to correct columns
  for (d in 1:2){ #each day data
    myfinger <- finger1
    if(d==2){
      myfinger<- finger2}
      tempfinger <- filter(myfinger,subject==subids[n],Response %in% c('t', 'u', 'm', 'b', 'w', 'r', 'v', 'x')) 
      #exclude rows with other stuff in Response
      #(this is not essential as nonsequences will just get ignored)

      wholeseq <-paste0(tempfinger$Response,collapse='') #make a long string of responses
      for (s in 1:2){
        myc<-myc+1 #increment column counter
        target <- myseq[s]
        finger.df[n,myc]<-str_count(wholeseq,target) 
        #count how many target sequences are in the long string that contains all responses

      }
      righth<-filter(tempfinger,display=='right hand')
      lefth<-filter(tempfinger,display=='left hand')
      startcol<-6
      if(d==2){startcol<-8}
      finger.df[n,startcol]<-sd(righth$RTa,na.rm=T)
      finger.df[n,(startcol+1)]<-sd(lefth$RTa,na.rm=T)
    }
}
# Substitute NA for any totals less than 5 - suggest using wrong keys perhaps

for (c in 2:5){
  w<-which(finger.df[,c]<5)
  if (length(w)>0){
     finger.df[w,2:9]<-NA

  }

}
#Need same data in long form for pirate plot, so reorganise
finger.left <- finger.df[,c(1,4,5,8,9)]
finger.right <- finger.df[,c(1,2,3,6,7)]
finger.left$side<-'Left'
finger.right$side <-'Right'
colnames(finger.right)<-c('subject', '1', '2', '1sd', '2sd','side')
colnames(finger.left)<-colnames(finger.right)
finger.long <-rbind(finger.left,finger.right)

```

```{r dosummary2, warning = F}
origdat <- finger.long
measure <- '1'
origdat$X1 <- origdat$`1` #we used a dummy column name X1, so we can vary the data that are used in the pirate plot etc
task <- 'day 1'
dopirate(origdat,task,measure)

origdat <- finger.long
measure <- '2'
origdat$X1 <- origdat$`2` #we used a dummy column name X1, so we can vary the data that are used in the pirate plot etc
task <- 'day 2'
dopirate(origdat,task,measure)

origdat <- finger.long
measure <- '1sd'
origdat$X1 <- origdat$`1sd` #we used a dummy column name X1, so we can vary the data that are used in the pirate plot etc
task <- 'day 1sd'
dopirate(origdat,task,measure)

origdat <- finger.long
measure <- '2sd'
origdat$X1 <- origdat$`2sd` #we used a dummy column name X1, so we can vary the data that are used in the pirate plot etc
task <- 'day 2sd'
dopirate(origdat,task,measure)
```

Finally, a handedness index is calculated and plotted. The handedness index is based on the equation [(RH-LH)/(RH+LH)] x 100, where LH= left hand and RH= right hand.

If the index is negative, the participant is categorised as left handed. If the index is positive, the participant is categorised as right handed.

In the next chunk, the LI is also added to demographics

```{r fingerLI}
task<-'Finger'
origdat<-finger.long
origdat$X1 <- origdat$`1`
mycolname<-"Day1_finger_LI"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (1)#so LI is positive in predicted direction
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

origdat$X1 <- origdat$`2`
mycolname<-"Day2_finger_LI"
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

cor1 <- cor(demographics$Day1_finger_LI,demographics$Day2_finger_LI,use='complete.obs')
plot(demographics$Day1_finger_LI,demographics$Day2_finger_LI,main='Finger Tapping LI')
abline(v=0)
abline(h=0)
text(-20,10,paste0('r = ',round(cor1,3)))
```

Here we use the Z score version.

```{r domakeZ_finger}
#We add Chimeric faces LIs to demographics for day 1 and then day2
task<-'Finger'
origdat<-finger.long
origdat$X1 <- origdat$`1`
mycolname<-"Day1_Finger_acc_Z"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (1)#so LI is positive in predicted direction
demographics <- makeZ(demographics,origdat,task,mycolname,polarity,mycutoff)

origdat<-finger.long
origdat$X1 <- origdat$`2`
mycolname<-"Day2_Finger_acc_Z"
polarity<- (-1)#so LI is positive in predicted direction
demographics <- makeZ(demographics,origdat,task,mycolname,polarity,mycutoff)

cor1 <- cor(demographics$Day1_Finger_acc_Z,demographics$Day2_Finger_acc_Z)
plot(demographics$Day1_Finger_acc_Z,demographics$Day2_Finger_acc_Z,main='Finger Z accuracy')
abline(v=0)
abline(h=0)
text(-2.5,2.5,paste0('r = ',round(cor1,3)))

write.csv(demographics, "LIs.csv")
```

## Final sample summary

Here we report the descriptives for the participants who had full datasets.

```{r Finaldemographics, warning=FALSE}
# first, describe age for the whole sample and plot
demographics %>%
    summarize(avg = mean(age), n = n(), sd = sd(age))
# second, show gender count for the whole population
demographics %>%
  group_by(gender) %>%
    summarize(n = n())
# third, show footedness
demographics %>%
  group_by(footedness) %>%
    summarize(n = n())
# fourth, show handedness
demographics %>%
  group_by(handedness) %>%
    summarize(n = n())
# fifth, show bilingualism
demographics %>%
  group_by(bilingual) %>%
    summarize(n = n())
# sixth, show miles
demographics %>%
  group_by(miles) %>%
    summarize(n = n())
# seventh, show porta
demographics %>%
  group_by(porta) %>%
    summarize(n = n())
# eigth, show LexTALE
demographics %>%
    summarize(avg = mean(lexTALE), n = n(), sd = sd(lexTALE))

# Now that we have the basic scores, let's look at how this looks with regards to handedness
# first, describe age for the whole sample and plot
ggplot(demographics, aes(x = handedness, y = age)) +
  geom_pirate(aes(colour = handedness), bars = FALSE) + theme_classic() + 
  ylim(10, 60) + ggtitle("Age by handedness") + ylab("Age (years)") + xlab("")
# second, show gender count for the whole population
a <- demographics %>%
  group_by(gender, handedness) %>%
    summarize(n = n())
ggplot(data=a, aes(x=handedness, y=n, fill= handedness)) +
  geom_bar(stat="identity")+
  geom_text(aes(label=n), vjust=1.6, color="white", size=3.5)+
  theme_classic() + facet_wrap(.~ gender) + ggtitle("Gender by handedness") + 
  ylab("Count") + xlab("")
# third, show footedness
a <- demographics %>%
  group_by(handedness, footedness) %>%
    summarize(n = n())
ggplot(data=a, aes(x=handedness, y=n, fill= handedness)) +
  geom_bar(stat="identity")+
  geom_text(aes(label=n), vjust=1.6, color="white", size=3.5)+
  theme_classic() + facet_wrap(.~ footedness) + ggtitle("Footedness by handedness") + 
  ylab("Count") + xlab("")
# fourth, show bilingualism
a <- demographics %>%
  group_by(bilingual, handedness) %>%
    summarize(n = n())
ggplot(data=a, aes(x=handedness, y=n, fill= handedness)) +
  geom_bar(stat="identity")+
  geom_text(aes(label=n), vjust=1.6, color="white", size=3.5)+
  theme_classic() + facet_wrap(.~ bilingual) + ggtitle("Bilingualism by handedness") + 
  ylab("Count") + xlab("")
# fifth, show miles
a <- demographics %>%
  group_by(miles, handedness) %>%
    summarize(n = n())
ggplot(data=a, aes(x=handedness, y=n, fill= handedness)) +
  geom_bar(stat="identity")+
  geom_text(aes(label=n), vjust=1.6, color="white", size=3.5)+
  theme_classic() + facet_wrap(.~ miles) + ggtitle("Mile test by handedness") + 
  ylab("Count") + xlab("")
# six, show porta
a <- demographics %>%
  group_by(porta, handedness) %>%
    summarize(n = n())
ggplot(data=a, aes(x=handedness, y=n, fill= handedness)) +
  geom_bar(stat="identity")+
  geom_text(aes(label=n), vjust=1.6, color="white", size=3.5)+
  theme_classic() + facet_wrap(.~ porta) + ggtitle("Porta test by handedness") + 
  ylab("Count") + xlab("")
# seventh, show LexTALE
ggplot(demographics, aes(x = handedness, y = lexTALE)) +
  geom_pirate(aes(colour = handedness), bars = FALSE) + theme_classic() + 
  ylim(70, 110) + ggtitle("LexTALE by handedness") + ylab("Standardised score") + xlab("")
```

```{r corr_between_tasks}
mycols <- c('Day1_CF_acc_Z','Day2_CF_acc_Z','Day1_Dich_acc_Z','Day2_Dich_acc_Z','tlat_RDT_1','tlat_RDT_2','Day1_Finger_acc_Z','Day2_Finger_acc_Z')
colnums<- which(colnames(demographics) %in% mycols)
mycorrs<-correlate(demographics[,colnums],method='spearman')

mycorrsL <-correlate(demographics[demographics$handedness=='Left',colnums],method='spearman')
mycorrsR <-correlate(demographics[demographics$handedness=='Right',colnums],method='spearman')

#compare L and R handers

for (i in colnums){
  print(names(demographics)[i])
 print(t.test(demographics[,i]~demographics$handedness))
}

#check lateralisaiton with single group t vs zero
for (i in colnums){
  print(names(demographics)[i])
 print(t.test(demographics[,i]))
}

```

**References:**

- Burt, D. M., & Perrett, D. I. (1997). Perceptual asymmetries in judgements of facial attractiveness, age, gender, speech and expression. Neuropsychologia, 35(5), 685–693.
- Hoaglin, D. C., & Iglewicz, B. (1987). Fine-tuning some resistant rules for outlier labelling. Journal of the American Statistical Association, 82(400), 1147-1149. 
- Hugdahl, K., Westerhausen, R., Alho, K., Medvedev, S., Laine, M., & Hämäläinen, H. (2009). Attention and cognitive control: unfolding the dichotic listening story. Scandinavian journal of psychology, 50(1), 11-22.
- Innes, B. R., Burt, D. M., Birch, Y. K., & Hausmann, M. (2016). A leftward bias however you look at it: Revisiting the emotional chimeric face task as a tool for measuring emotion lateralization. Laterality: Asymmetries of Body, Brain and Cognition, 21(4-6), 643–661.
- Karlsson, E. M., Johnstone, L. T., & Carey, D. P. (2019). The depth and breadth of multiple perceptual asymmetries in right handers and non-right handers. Laterality: Asymmetries of Body, Brain and Cognition, 24(6), 707-739.

### 2020-04-23 battery-preprocessing_v5_DB.Rmd
Done a bit more with outliers on dichotic, but still grappling with RDT data - why some have NA for RT and what to do with them. Need to look at when fresher.

@@
 author: "Adam Parker + Dorothy Bishop"
-date: "20th March 2020"
+date: "updated April 22nd 2020"
 output:
@@
 ---
+NB. 2 changes Apr 22nd by DB: 
+a) all data now in a Data folder, so when reading need to specify that; 
+b) *demographics* data frame renamed to *allsum*.
 
@@
 This version updated on 1st Feb 2020 to make script more reproducible and compact by using functions for repeated operations.
+Adapted in April 2020 for processing dataset on 300 people
 
 The script pulls in csv files and processes them so that pre-registered analyses can be conducted (pre-registered on the OSF: ).
-First, questionnaires are analyses followed by visual half field tasks.
+First, questionnaires are analyses followed by chimeric, dichotic, visual half field and tapping tasks.
 This is done for time 1 and time 2.
@@
 This script analyses the following: 
-
 1) Questionnaires
-  - Demographics
+  - demographics
   - Miles test
@@
 2) Chimeric face task
-3) Rhyme detection task
-4) Key Pressing Task (finger dexterity)
-5) Summary of demographics for the final sample
+3) Dichotic listening
+4) Rhyme detection task
+5) Key Pressing Task (finger dexterity)
+And then produces summary of demog and results for the final sample
 
@@
 
-- Demographics
+- demographics
 - Miles test
@@
 
-#NB for present setwd("~/Dropbox/Online LDT/300 Ps")
+#NB for original data setwd("~/Dropbox/Online LDT/300 Ps")
 # libraries 
@@
 
-```{r demographics, warning=FALSE}
-demo_dat_R <- read_csv("./Demographics_R.csv", na = c("empty", "NA")) # read in data
-demo_dat_L <- read_csv("./Demographics_L.csv", na = c("empty", "NA")) # read in data
+```{r demograhics, warning=FALSE}
+demo_dat_R <- read_csv("./data/demographics_R.csv", na = c("empty", "NA")) # read in data
+demo_dat_L <- read_csv("./data/demographics_L.csv", na = c("empty", "NA")) # read in data
 demo_dat <- rbind(demo_dat_R, demo_dat_L)
@@
 # create data frame
-demographics <- data.frame(matrix(ncol = 7, nrow = nsub))
-  colnames(demographics) <- c("subject", "age", "gender", "footedness", "handedness", "bilingual", "strongest.language")
+allsum <- data.frame(matrix(ncol = 7, nrow = nsub))
+  colnames(allsum) <- c("subject", "age", "gender", "footedness", "handedness", "bilingual", "strongest.language")
 myrow <- 0 # start row counter
@@
 
-    demographics$subject[myrow] <- subname
-    demographics$age[myrow] <- as.numeric(tmp[tmp$Question.Key == "Age",]$Response)
-    demographics$gender[myrow] <- tmp[tmp$Question.Key == "Gender",]$Response
-    demographics$footedness[myrow] <- tmp[tmp$Question.Key == "footedness",]$Response
-    demographics$handedness[myrow] <- tmp[tmp$Question.Key == "categorical_hand",]$Response
-    demographics$bilingual[myrow] <- tmp[tmp$Question.Key == "bilingual",]$Response
-    demographics$strongest.language[myrow] <- tolower(tmp[tmp$Question.Key == "bilingual-text",]$Response)
-
-    demographics$bilingual <- as.character(demographics$bilingual)
-    demographics$bilingual <- gsub('. My strongest language is:', '', demographics$bilingual)
+    allsum$subject[myrow] <- subname
+    allsum$age[myrow] <- as.numeric(tmp[tmp$Question.Key == "Age",]$Response)
+    allsum$gender[myrow] <- tmp[tmp$Question.Key == "Gender",]$Response
+    allsum$footedness[myrow] <- tmp[tmp$Question.Key == "footedness",]$Response
+    allsum$handedness[myrow] <- tmp[tmp$Question.Key == "categorical_hand",]$Response
+    allsum$bilingual[myrow] <- tmp[tmp$Question.Key == "bilingual",]$Response
+    allsum$strongest.language[myrow] <- tolower(tmp[tmp$Question.Key == "bilingual-text",]$Response)
+
+    allsum$bilingual <- as.character(allsum$bilingual)
+    allsum$bilingual <- gsub('. My strongest language is:', '', allsum$bilingual)
 }
@@
 # age
-demographics %>%
+allsum %>%
     group_by(handedness, gender) %>%
     summarize(avg = mean(age), n = n(), sd = sd(age))
-```
-
-Now we import and append the Porta and Miles Test to the demographics data frame.
+
+#create a numerical code for handedness - can be used in plots to distinguish handedness groups (eg if pch set to handpch, get circles and triangles)
+allsum$handpch <-1
+allsum$handpch[allsum$handedness=='Left'] <-2
+```
+
+Now we import and append the Porta and Miles Test to the allsum data frame.
 
 ```{r sightednesss, warning=FALSE}
-miles_dat_R <- read_csv("./Miles_R.csv", na = c("empty", "NA")) # read in data
-miles_dat_L <- read_csv("./Miles_L.csv", na = c("empty", "NA")) # read in data
+miles_dat_R <- read_csv("./data/Miles_R.csv", na = c("empty", "NA")) # read in data
+miles_dat_L <- read_csv("./data/Miles_L.csv", na = c("empty", "NA")) # read in data
 miles_dat <- rbind(miles_dat_R, miles_dat_L)
 #porta_dat_R <- read_csv("./dat2/Porta_R.csv", na = c("empty", "NA")) # read in data
-porta_dat_L <- read_csv("./Porta_L.csv", na = c("empty", "NA")) # read in data
+porta_dat_L <- read_csv("./data/Porta_L.csv", na = c("empty", "NA")) # read in data
 porta_dat <- rbind(porta_dat_L)
@@
 
-# now merge sightedness and demographics
-demographics <- merge(demographics, sightedness, by= "subject")
+# now merge sightedness and allsum
+allsum <- merge(allsum, sightedness, by= "subject")
 ```
@@
 ```{r EHI, warning=FALSE}
-EHI_R <- read_csv("./EHI_R.csv", na = "empty") # read in data
-EHI_L <- read_csv("./EHI_L.csv", na = "empty") # read in data
+EHI_R <- read_csv("./data/EHI_R.csv", na = "empty") # read in data
+EHI_L <- read_csv("./data/EHI_L.csv", na = "empty") # read in data
 EHI <- rbind(EHI_R, EHI_L)
@@
 
-# score responses accoridng to the original scoring method in Oldifeld (1971)
+# score responses according to the original scoring method in Oldifeld (1971)
 EHI$right_hand <- 0
@@
 EHI_index <- select(for_index2, "subject", "index_EHI")
-# merge which demographics
-demographics <- merge(demographics, EHI_index, by= "subject")
+# merge with allsum
+allsum <- merge(allsum, EHI_index, by= "subject")
 
 #now plot EHI in relation to handedness (modified by DB)
-x<-ggplot(demographics, aes(x=index_EHI, y=subject, fill=subject)) +
-  geom_dotplot(binaxis='y', stackdir='center') + xlim(-100, 100) +
+
+eplot<-ggplot(allsum, aes(handpch,index_EHI, colour=gender)) +
+  geom_jitter(width=.25,height=2)+
   ggtitle("Handedness index on the Edinburgh Handedness Inventory")
-x + theme_bw() + theme(legend.position = "none") + theme(axis.text.y = element_blank())
-
-hist(demographics$index_EHI)
+eplot
+#This is not perfect, but gives and idea to check EHI is valid etc.
+
+ggplot(allsum, aes(x=index_EHI)) + geom_density(aes(group=handpch,fill=handpch))
+
+#NB one supposed right-hander has index_EHI around -50. 
+#I will recode this one, as this seems an error.
+
+oddcase <- intersect(which(allsum$index_EHI<0),which(allsum$handedness=='Right'))
+allsum$handedness[oddcase] <- 'Left'
+allsum$handpch[oddcase] <- 1
 ```
@@
 # read data
-lexTALE_L <- read_csv("./LexTale_L.csv", na = "empty")
-lexTALE_R <- read_csv("./lexTale_R.csv", na = "empty")
+lexTALE_L <- read_csv("./data/LexTale_L.csv", na = "empty")
+lexTALE_R <- read_csv("./data/lexTale_R.csv", na = "empty")
 lexTALE <- rbind(lexTALE_R, lexTALE_L)
@@
 lexTALE <- select(lexTale_score, "subject", "lexTALE")
-# now merge with demographics
-demographics <- merge(demographics, lexTALE, by= "subject")
-# now write the full demographics 
-write.csv(demographics, file = "./demographics_out.csv")
-
-hist(demographics$lexTALE)
+# now merge with allsum
+allsum <- merge(allsum, lexTALE, by= "subject")
+# now write the full allsum 
+write.csv(allsum, file = "./allsum_out.csv")
+
+hist(allsum$lexTALE)
 abline(v=80,col='red')
 
-paste0("No. of participants below LexTALE cuttoff: ", sum(demographics$lexTALE < 80))
+paste0("No. of participants below LexTALE cuttoff: ", sum(allsum$lexTALE < 80))
 
 # remove those who didn't pass the LexTALE
-demographics <- demographics[demographics$lexTALE >= 80,]
+allsum <- allsum[allsum$lexTALE >= 80,]
 ```
@@
 
-Before doing anything with the data, we now pull in all the data from all nodes and combind them from a sub folder in the directory.
+Before doing anything with the data, we now pull in all the data from all nodes and combine them from a sub folder in the directory.
 Participants, identified by random letter strings, are stacked on top of each other.
@@
 for (d in 1:2){
-  data_file <- "./day1_combined.csv"
+  data_file <- "./data/day1_combined.csv"
 
   if (d ==2){
-    data_file <-  "./day2_combined.csv"
+    data_file <-  "./data/day2_combined.csv"
   }
@@
 }
-
-# #remove cols from all_dat2 that contain 'branch' or 'randomiser'
-# branchcols2 <- c(which(grepl('branch', names(all_dat2))=='TRUE'),which(grepl('randomiser', names(all_dat2))=='TRUE'))
-# all_dat2 <- all_dat2[,-branchcols2]
-# branchcols1 <- c(which(grepl('branch', names(all_dat1))=='TRUE'),which(grepl('randomiser', names(all_dat1))=='TRUE'))
-# all_dat1 <- all_dat1[,-branchcols1]
-
-#check that colnames are the same
-# colcomp<-data.frame(colnames(all_dat2)[order(colnames(all_dat2))])
-# colnames(colcomp)[1]<-'name1'
-# ncol1 <-length(colnames(all_dat1))
-# colcomp$name2<-NA
-# colcomp$name2[1:ncol1]<-colnames(all_dat1)[order(colnames(all_dat1))]
-
-#NB the colnames are identical but in different order, and all_dat1 has extra column of word_OG
-#It is just conceivable the ordering might affect processing when columns are subsequently selected, so we'll reorder them to be consistent across sessions
 
@@
 
-Another generic function that can be used for different tasks to make a laterality index and write it to the demographics file
+Another generic function that can be used for different tasks to make a laterality index and write it to the allsum file
 
 ```{r makeLIs, warning=FALSE}
-# This function modifies a copy of the demographics file and returns it
+# This function modifies a copy of the allsum file and returns it
 
@@
 
-makeLI <- function(demogs,origdat,task,mycolname,polarity,mycutoff){
+makeLI <- function(allsums,origdat,task,mycolname,polarity,mycutoff){
 # we do this for accuracy first
@@
 LIdf <- select(LIdf, "subject", "LI")
-# now merge with demographics
-demogs <- merge(demogs, LIdf, by= "subject")
-lastcol<-length(colnames(demogs))
-colnames(demogs)[lastcol]<-mycolname
-
-return(demogs)
+# now merge with allsum
+allsums <- merge(allsums, LIdf, by= "subject")
+lastcol<-length(colnames(allsums))
+colnames(allsums)[lastcol]<-mycolname
+
+return(allsums)
 }
@@
 ```{r makeZ, warning=FALSE}
-# This function modifies a copy of the demographics file and returns it
+# This function modifies a copy of the allsum file and returns it
 
@@
 
-makeZ <- function(demogs,origdat,task,mycolname,polarity,mycutoff){
+makeZ <- function(allsums,origdat,task,mycolname,polarity,mycutoff){
 # we do this for accuracy first
@@
 Zdf <- select(Zdf, "subject", "Z")
-# now merge with demographics
-demogs <- merge(demogs, Zdf, by= "subject")
-lastcol<-length(colnames(demogs))
-colnames(demogs)[lastcol]<-mycolname
-
-return(demogs)
+# now merge with allsum
+allsums <- merge(allsums, Zdf, by= "subject")
+lastcol<-length(colnames(allsums))
+colnames(allsums)[lastcol]<-mycolname
+
+return(allsums)
 }
@@
 emotion_mean <- aggregate(FUN= mean, data= CF_same, emotion_recog~ subject)
-# add this to demographics 
-demographics <- merge(demographics, emotion_mean, by= "subject")
-}
-mycol <- length(colnames(demographics))
-colnames(demographics)[(mycol-1):mycol]<-c('Emot.recog1','Emot.recog2')
-
-plot(demographics$Emot.recog1,demographics$Emot.recog2)
+# add this to allsum 
+allsum <- merge(allsum, emotion_mean, by= "subject")
+}
+mycol <- length(colnames(allsum))
+colnames(allsum)[(mycol-1):mycol]<-c('Emot.recog1','Emot.recog2')
+
+plot(jitter(allsum$Emot.recog1,10),jitter(allsum$Emot.recog2,10),pch=3,cex=.6)
 abline(h=.75)
@@
 #db modified to mark rather than drop cases with scores < .75
-demographics$exclude <- 0 #default is to exclude
+allsum$exclude <- 0 #default is to exclude
 
 # mark the cases where there is less than 75% for identical stimuli
-w <- union(which(demographics$Emot.recog1 < .75),which(demographics$Emot.recog2 < .75))
-demographics$exclude[w]<-1 #code exclude = 1 if excluded because poor emot recog
+w <- union(which(allsum$Emot.recog1 < .75),which(allsum$Emot.recog2 < .75))
+allsum$exclude[w]<-1 #code exclude = 1 if excluded because poor emot recog
 ```
@@
 ```{r domakeLI}
-#We add Chimeric faces LIs to demographics for day 1 and then day2
+#We add Chimeric faces LIs to allsum for day 1 and then day2
 task<-'Chimeric faces'
@@
 polarity<- (-1)#so LI is positive in predicted direction
-demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
+allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
 
@@
 #if we don't specify mycutoff, it will remain at level set for accuracy, which means it should not trap any RT values as they will all be much bigger
-demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
+allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
 
@@
 polarity<- (-1)#so LI is positive in predicted direction
-demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
+allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
 origdat$X1 <- origdat$RT
@@
 polarity <-1
-demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
-
-cor1 <- cor(demographics$Day1_CF_acc_LI,demographics$Day2_CF_acc_LI)
-plot(demographics$Day1_CF_acc_LI,demographics$Day2_CF_acc_LI,main='Chimeric LI accuracy',col=(1+demographics$exclude))
+allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
+
+cor1 <- cor(allsum$Day1_CF_acc_LI,allsum$Day2_CF_acc_LI)
+plot(allsum$Day1_CF_acc_LI,allsum$Day2_CF_acc_LI,main='Chimeric LI accuracy',col=(1+allsum$exclude))
 abline(v=0)
@@
 text(-50,65,'red will be excluded',cex=.85)
-cor2<-cor(demographics$Day1_CF_RT_LI,demographics$Day2_CF_RT_LI)
-plot(demographics$Day1_CF_RT_LI,demographics$Day2_CF_RT_LI,main='Chimeric LI RT',col=(1+demographics$exclude))
+cor2<-cor(allsum$Day1_CF_RT_LI,allsum$Day2_CF_RT_LI)
+plot(allsum$Day1_CF_RT_LI,allsum$Day2_CF_RT_LI,main='Chimeric LI RT',col=(1+allsum$exclude))
 abline(v=0)
@@
 
-Here we use the Z score version.
+Here we use the Z score version. This computes for each subject whether they have a different proportion correct for L and R
 
 ```{r domakeZ}
-#We add Chimeric faces LIs to demographics for day 1 and then day2
+#We add Chimeric faces LIs to allsum for day 1 and then day2
 task<-'Chimeric faces'
@@
 polarity<- (-1)#so LI is positive in predicted direction
-demographics <- makeZ(demographics,origdat,task,mycolname,polarity,mycutoff)
+allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)
 
@@
 polarity<- (-1)#so LI is positive in predicted direction
-demographics <- makeZ(demographics,origdat,task,mycolname,polarity,mycutoff)
-w<-which(demographics$excluded == 1)
-demographics$Day1_CF_acc_Z[w] <- NA #removes zscores for excluded cases
-demographics$Day2_CF_acc_Z[w] <- NA
-
-RR<-intersect(which(demographics$Day1_CF_acc_Z>1.96),which(demographics$Day2_CF_acc_Z>1.96))
-LL<-intersect(which(demographics$Day1_CF_acc_Z< -1.96),which(demographics$Day2_CF_acc_Z< -1.96))
-
-cor1 <- cor(demographics$Day1_CF_acc_Z,demographics$Day2_CF_acc_Z,use='complete.obs',method='spearman')
-plot(demographics$Day1_CF_acc_Z,demographics$Day2_CF_acc_Z,main='Chimeric Z accuracy')
+allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)
+w<-which(allsum$excluded == 1)
+allsum$Day1_CF_acc_Z[w] <- NA #removes zscores for excluded cases
+allsum$Day2_CF_acc_Z[w] <- NA
+
+RR<-intersect(which(allsum$Day1_CF_acc_Z>1.96),which(allsum$Day2_CF_acc_Z>1.96))
+LL<-intersect(which(allsum$Day1_CF_acc_Z< -1.96),which(allsum$Day2_CF_acc_Z< -1.96))
+
+cor1 <- cor(allsum$Day1_CF_acc_Z,allsum$Day2_CF_acc_Z,use='complete.obs',method='spearman')
+plot(allsum$Day1_CF_acc_Z,allsum$Day2_CF_acc_Z,main='Chimeric Z accuracy')
 abline(v=0)
@@
 abline(h = -1.96, col='red')
-text(-7,7.5,paste0('r = ',round(cor1,3)),cex=.8)
+text(-7,7.5,paste0('Spearman r = ',round(cor1,3)),cex=.8)
 text(-2,-15,paste('Those in bottom quadrant\n bounded by red are consistently\n L-lateralised, N = ',length(LL)),cex=.8)
 text(5,18,paste('Those in top quadrant\n bounded by red are consistently\n R-lateralised, N = ',length(RR)),cex=.8)
-text(8,-8,'One outlier excluded',cex=.8)
+#text(8,-8,'One outlier excluded',cex=.8)
+#Removal of the outlier does not have much influence on rs
 
@@
 As we can see, we generally get a left side/right hemisphere advantage in our accuracy data.
-The RT advantage is not as expected.
-This may be due to the buttons being clicked on screen and the additional time penalty of moving the mouse.
+We would not expect a RT advantage for this type of measure, where the response is to a single stimulus with conflicting information from the 2 sides.
 
@@
 
-First, we judge participants' ability to correctly identify sounds using the "dichotic_same" dataframe. Participants are removed if their performance is below 75%.
+First, we judge participants' ability to correctly identify sounds using the "dichotic_same" dataframe. Participants are marked as excluded if their performance is below 75%.
 
@@
 sound_mean <- aggregate(FUN= mean, data= dichotic_same, sound_recog~ subject)
-# add this to demographics 
-demographics <- merge(demographics, sound_mean, by= "subject")
-}
-mycol <- length(colnames(demographics))
-colnames(demographics)[(mycol-1):mycol]<-c('sound.recog1','sound.recog2')
+# add this to allsum 
+allsum <- merge(allsum, sound_mean, by= "subject")
+}
+mycol <- length(colnames(allsum))
+colnames(allsum)[(mycol-1):mycol]<-c('sound.recog1','sound.recog2')
 
 # once again, mark rather than remove those with poor performance on identical stimuli
-w<-union(which(demographics$sound.recog1 < .75), which(demographics$sound.recog2 < .75))
-ww<-intersect(which(demographics$exclude==1),w)
+w<-union(which(allsum$sound.recog1 < .75), which(allsum$sound.recog2 < .75))
+ww<-intersect(which(allsum$exclude==1),w)
 #ww contains people who were excluded on both dichotic and faces. 
 # code exclude = 2 for the cases where there less than 75% for identical stimuli
-demographics$exclude[w]<-2 #code exclude = 2 if excluded because poor dichotic
-demographics$exclude[ww]<-12 #code for those excluded on both chimeric and dichotic
-table(demographics$exclude) 
+allsum$exclude[w]<-2 #code exclude = 2 if excluded because poor dichotic
+allsum$exclude[ww]<-12 #code for those excluded on both chimeric and dichotic
+t <- table(allsum$exclude) 
+rownames(t)<-c('include','exclude CF','exclude dich','exclude CF+dich')
+t
 ```
@@
 ```{r domakeLIdich}
-#We add Dichotic LIs to demographics for day 1 and then day2
+#We add Dichotic LIs to allsum for day 1 and then day2
 task<-'Dichotic listening'
@@
 polarity <-1
-demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
-
-cor1 <- cor(demographics$Day1_DL_acc_LI,demographics$Day2_DL_acc_LI)
-plot(demographics$Day1_DL_acc_LI,demographics$Day2_DL_acc_LI,main='Dichotic LI accuracy',col=(1+demographics$exclude))
+allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
+
+cor1 <- cor(allsum$Day1_DL_acc_LI,allsum$Day2_DL_acc_LI)
+plot(allsum$Day1_DL_acc_LI,allsum$Day2_DL_acc_LI,main='Dichotic LI accuracy',col=(1+allsum$exclude))
 abline(v=0)
@@
 text(-75,75,paste0('r = ',round(cor1,3)))
-cor2<-cor(demographics$Day1_DL_RT_LI,demographics$Day2_DL_RT_LI)
-plot(demographics$Day1_DL_RT_LI,demographics$Day2_DL_RT_LI,main='Dichotic LI RT',col=(1+demographics$exclude))
+cor2<-cor(allsum$Day1_DL_RT_LI,allsum$Day2_DL_RT_LI)
+plot(allsum$Day1_DL_RT_LI,allsum$Day2_DL_RT_LI,main='Dichotic LI RT',col=(1+allsum$exclude))
 abline(v=0)
@@
 ```{r domakeZ_dich}
-#We add dichotic LIs to demographics for day 1 and then day2
+#We add dichotic LIs to allsum for day 1 and then day2
 task<-'Dichotic listening'
@@
 polarity<- (-1)#so LI is positive in predicted direction
-demographics <- makeZ(demographics,origdat,task,mycolname,polarity,mycutoff)
+allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)
 
 #convert excluded cases with code 2 or 12 to NA
-w<-which(demographics$exclude>1)
-demographics$Day1_Dich_acc_Z[w]<-NA
-demographics$Day2_Dich_acc_Z[w]<-NA
-
-RR<-intersect(which(demographics$Day1_Dich_acc_Z>1.96),which(demographics$Day2_Dich_acc_Z>1.96))
-LL<-intersect(which(demographics$Day1_Dich_acc_Z< -1.96),which(demographics$Day2_Dich_acc_Z< -1.96))
-
-cor1 <- cor(demographics$Day1_Dich_acc_Z,demographics$Day2_Dich_acc_Z,use='complete.obs',method='spearman')
-demographics$handpch <-1
-demographics$handpch[demographics$handedness=='Left'] <-2
-plot(demographics$Day1_Dich_acc_Z,demographics$Day2_Dich_acc_Z,main='Dich Z accuracy',pch=demographics$handpch)
+w<-which(allsum$exclude>1)
+allsum$Day1_Dich_acc_Z[w]<-NA
+allsum$Day2_Dich_acc_Z[w]<-NA
+
+RR<-intersect(which(allsum$Day1_Dich_acc_Z>1.96),which(allsum$Day2_Dich_acc_Z>1.96))
+LL<-intersect(which(allsum$Day1_Dich_acc_Z< -1.96),which(allsum$Day2_Dich_acc_Z< -1.96))
+
+cor1 <- cor(allsum$Day1_Dich_acc_Z,allsum$Day2_Dich_acc_Z,use='complete.obs',method='spearman')
+
+plot(allsum$Day1_Dich_acc_Z,allsum$Day2_Dich_acc_Z,main='Dich Z accuracy',pch=allsum$handpch)
 abline(v=0)
@@
 #remove.p <- list(unique(remove.outlier$Var1))
-#demographics <- demographics[demographics$subject != remove.p,]
+#allsum <- allsum[allsum$subject != remove.p,]
 
@@
 mycutoff <- 75
-demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
+allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
 origdat$X1 <- origdat$RT
@@
 polarity <- (-1)
-demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
+allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
 
@@
 #We do this by finding those with NA in accuracy LI
-w<-which(is.na(demographics$RDT_acc_LI))
+w<-which(is.na(allsum$RDT_acc_LI))
 if(length(w)>0)
-{demographics$RDT_RT_LI[w]<-NA}
+{allsum$RDT_RT_LI[w]<-NA}
 #t.test to test if laterality index differs from zero
-t.test(demographics$RDT_p.corr_LI)
-t.test(demographics$RDT_RT_LI)
+t.test(allsum$RDT_p.corr_LI)
+t.test(allsum$RDT_RT_LI)
 #rename column to identify day
-myncol <- length(colnames(demographics))
-colnames(demographics)[(myncol-1):myncol] <- paste0('Day',d,'.',colnames(demographics)[(myncol-1):myncol])
-
-}
-cor1<-cor(demographics$Day1.RDT_p.corr_LI,demographics$Day2.RDT_p.corr_LI,use='complete.obs')
-cor2<-cor(demographics$Day1.RDT_RT_LI,demographics$Day2.RDT_RT_LI,use='complete.obs')
-plot(demographics$Day1.RDT_p.corr_LI,demographics$Day2.RDT_p.corr_LI,main='Rhyme detect: LI for accuracy (%correct)')
+myncol <- length(colnames(allsum))
+colnames(allsum)[(myncol-1):myncol] <- paste0('Day',d,'.',colnames(allsum)[(myncol-1):myncol])
+
+}
+cor1<-cor(allsum$Day1.RDT_p.corr_LI,allsum$Day2.RDT_p.corr_LI,use='complete.obs')
+cor2<-cor(allsum$Day1.RDT_RT_LI,allsum$Day2.RDT_RT_LI,use='complete.obs')
+plot(allsum$Day1.RDT_p.corr_LI,allsum$Day2.RDT_p.corr_LI,main='Rhyme detect: LI for accuracy (%correct)')
 abline(v=0)
@@
 text(-3,3,paste0('r = ',round(cor1,3)))
-plot(demographics$Day1.RDT_RT_LI,demographics$Day2.RDT_RT_LI,main='Rhyme detect: LI for RT')
+plot(allsum$Day1.RDT_RT_LI,allsum$Day2.RDT_RT_LI,main='Rhyme detect: LI for RT')
 abline(v=0)
@@
 
-  nsubs <- nrow(demographics)
-  demographics$tlat_RDT<-NA
-   demographics$plat_RDT<-NA
+  nsubs <- nrow(allsum)
+  allsum$tlat_RDT<-NA
+   allsum$plat_RDT<-NA
   for (i in 1:nsubs) {
-    w<-which(myfile$subject==demographics$subject[i])
+    w<-which(myfile$subject==allsum$subject[i])
     if(length(w)>0)
@@
       myt<-t.test(myfile$RT[w]~myfile$side[w])
-      demographics$tlat_RDT[i]<-myt$statistic
-      demographics$plat_RDT[i]<-myt$p.value
+      allsum$tlat_RDT[i]<-myt$statistic
+      allsum$plat_RDT[i]<-myt$p.value
     }
@@
    if (d==1){
-     maxcol <-length(colnames(demographics))
-     colnames(demographics)[(maxcol-1):maxcol]<-c('tlat_RDT_1','plat_RDT_1')
+     maxcol <-length(colnames(allsum))
+     colnames(allsum)[(maxcol-1):maxcol]<-c('tlat_RDT_1','plat_RDT_1')
    }
    if (d==2){
-     maxcol <-length(colnames(demographics))
-     colnames(demographics)[(maxcol-1):maxcol]<-c('tlat_RDT_2','plat_RDT_2')
-   }
-}
-colcount<-length(colnames(demographics))
-# Remove values for cases that are excluded for low acc: do this for both times by just blitzing last 4 columns of demographics
-w<-which(is.na(demographics$Day1.RDT_RT_LI))
-demographics[w,(colcount-3):colcount]<-NA
-w<-which(is.na(demographics$Day2.RDT_RT_LI))
-demographics[w,(colcount-3):colcount]<-NA
+     maxcol <-length(colnames(allsum))
+     colnames(allsum)[(maxcol-1):maxcol]<-c('tlat_RDT_2','plat_RDT_2')
+   }
+}
+colcount<-length(colnames(allsum))
+# Remove values for cases that are excluded for low acc: do this for both times by just blitzing last 4 columns of allsum
+w<-which(is.na(allsum$Day1.RDT_RT_LI))
+allsum[w,(colcount-3):colcount]<-NA
+w<-which(is.na(allsum$Day2.RDT_RT_LI))
+allsum[w,(colcount-3):colcount]<-NA
 
 # compare t-value with LI : use t = 1.65 as cutoff for sig lateralised
- plot(demographics$tlat_RDT_1,demographics$Day1.RDT_RT_LI)
+ plot(allsum$tlat_RDT_1,allsum$Day1.RDT_RT_LI)
  abline(v=1.65)
-  plot(demographics$tlat_RDT_2,demographics$Day2.RDT_RT_LI)
+  plot(allsum$tlat_RDT_2,allsum$Day2.RDT_RT_LI)
    abline(v=1.65)
 # compare t-value from day 1 and day 2 
-     myr <- cor(demographics$tlat_RDT_1,demographics$tlat_RDT_2,use='complete.obs')
-    plot(demographics$tlat_RDT_1,demographics$tlat_RDT_2)
+     myr <- cor(allsum$tlat_RDT_1,allsum$tlat_RDT_2,use='complete.obs')
+    plot(allsum$tlat_RDT_1,allsum$tlat_RDT_2)
     abline(v=0)
@@
 
-In the next chunk, the LI is also added to demographics
+In the next chunk, the LI is also added to allsum
 
@@
 polarity<- (1)#so LI is positive in predicted direction
-demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
+allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
 
@@
 mycolname<-"Day2_finger_LI"
-demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
-
-cor1 <- cor(demographics$Day1_finger_LI,demographics$Day2_finger_LI,use='complete.obs')
-plot(demographics$Day1_finger_LI,demographics$Day2_finger_LI,main='Finger Tapping LI')
+allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
+
+cor1 <- cor(allsum$Day1_finger_LI,allsum$Day2_finger_LI,use='complete.obs')
+plot(allsum$Day1_finger_LI,allsum$Day2_finger_LI,main='Finger Tapping LI')
 abline(v=0)
@@
 ```{r domakeZ_finger}
-#We add Chimeric faces LIs to demographics for day 1 and then day2
+#We add Chimeric faces LIs to allsum for day 1 and then day2
 task<-'Finger'
@@
 polarity<- (1)#so LI is positive in predicted direction
-demographics <- makeZ(demographics,origdat,task,mycolname,polarity,mycutoff)
+allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)
 
@@
 polarity<- (-1)#so LI is positive in predicted direction
-demographics <- makeZ(demographics,origdat,task,mycolname,polarity,mycutoff)
-
-cor1 <- cor(demographics$Day1_Finger_acc_Z,demographics$Day2_Finger_acc_Z)
-plot(demographics$Day1_Finger_acc_Z,demographics$Day2_Finger_acc_Z,main='Finger Z accuracy')
+allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)
+
+cor1 <- cor(allsum$Day1_Finger_acc_Z,allsum$Day2_Finger_acc_Z)
+plot(allsum$Day1_Finger_acc_Z,allsum$Day2_Finger_acc_Z,main='Finger Z accuracy')
 abline(v=0)
@@
 
-write.csv(demographics, "LIs.csv")
+write.csv(allsum, "LIs.csv")
 ```
@@
 
-```{r Finaldemographics, warning=FALSE}
+```{r Finalallsum, warning=FALSE}
 # first, describe age for the whole sample and plot
-demographics %>%
+allsum %>%
     summarize(avg = mean(age), n = n(), sd = sd(age))
 # second, show gender count for the whole population
-demographics %>%
+allsum %>%
   group_by(gender) %>%
@@
 # third, show footedness
-demographics %>%
+allsum %>%
   group_by(footedness) %>%
@@
 # fourth, show handedness
-demographics %>%
+allsum %>%
   group_by(handedness) %>%
@@
 # fifth, show bilingualism
-demographics %>%
+allsum %>%
   group_by(bilingual) %>%
@@
 # sixth, show miles
-demographics %>%
+allsum %>%
   group_by(miles) %>%
@@
 # seventh, show porta
-demographics %>%
+allsum %>%
   group_by(porta) %>%
@@
 # eigth, show LexTALE
-demographics %>%
+allsum %>%
     summarize(avg = mean(lexTALE), n = n(), sd = sd(lexTALE))
@@
 # first, describe age for the whole sample and plot
-ggplot(demographics, aes(x = handedness, y = age)) +
+ggplot(allsum, aes(x = handedness, y = age)) +
   geom_pirate(aes(colour = handedness), bars = FALSE) + theme_classic() + 
@@
 # second, show gender count for the whole population
-a <- demographics %>%
+a <- allsum %>%
   group_by(gender, handedness) %>%
@@
 # third, show footedness
-a <- demographics %>%
+a <- allsum %>%
   group_by(handedness, footedness) %>%
@@
 # fourth, show bilingualism
-a <- demographics %>%
+a <- allsum %>%
   group_by(bilingual, handedness) %>%
@@
 # fifth, show miles
-a <- demographics %>%
+a <- allsum %>%
   group_by(miles, handedness) %>%
@@
 # six, show porta
-a <- demographics %>%
+a <- allsum %>%
   group_by(porta, handedness) %>%
@@
 # seventh, show LexTALE
-ggplot(demographics, aes(x = handedness, y = lexTALE)) +
+ggplot(allsum, aes(x = handedness, y = lexTALE)) +
   geom_pirate(aes(colour = handedness), bars = FALSE) + theme_classic() + 
@@
 mycols <- c('Day1_CF_acc_Z','Day2_CF_acc_Z','Day1_Dich_acc_Z','Day2_Dich_acc_Z','tlat_RDT_1','tlat_RDT_2','Day1_Finger_acc_Z','Day2_Finger_acc_Z')
-colnums<- which(colnames(demographics) %in% mycols)
-mycorrs<-correlate(demographics[,colnums],method='spearman')
-
-mycorrsL <-correlate(demographics[demographics$handedness=='Left',colnums],method='spearman')
-mycorrsR <-correlate(demographics[demographics$handedness=='Right',colnums],method='spearman')
+colnums<- which(colnames(allsum) %in% mycols)
+mycorrs<-correlate(allsum[,colnums],method='spearman')
+
+mycorrsL <-correlate(allsum[allsum$handedness=='Left',colnums],method='spearman')
+mycorrsR <-correlate(allsum[allsum$handedness=='Right',colnums],method='spearman')
 
@@
 for (i in colnums){
-  print(names(demographics)[i])
- print(t.test(demographics[,i]~demographics$handedness))
+  print(names(allsum)[i])
+ print(t.test(allsum[,i]~allsum$handedness))
 }
@@
 for (i in colnums){
-  print(names(demographics)[i])
- print(t.test(demographics[,i]))
+  print(names(allsum)[i])
+ print(t.test(allsum[,i]))
 }

### 2020-05-21 battery-preprocessing_v5.Rmd
this includes lower number of trails needed for inclusion

--- 2020-04-23 battery-preprocessing_v5_DB.Rmd
@@
 ```{r demograhics, warning=FALSE}
-demo_dat_R <- read_csv("./data/demographics_R.csv", na = c("empty", "NA")) # read in data
-demo_dat_L <- read_csv("./data/demographics_L.csv", na = c("empty", "NA")) # read in data
-demo_dat <- rbind(demo_dat_R, demo_dat_L)
+demo_dat <- read_csv("./data/Demographics.csv", na = c("empty", "NA")) # read in data
 # relabel 
-demo_dat$subject <- demo_dat$`Participant Public ID`
+demo_dat$subject <- demo_dat$Participant.Public.ID
 demo_dat$subject <- as.factor(demo_dat$subject)
@@
 
-# Here we summarise the age broken down by handedness and gender. This will be important for any write up. 
-#### NOTE. THIS IS NOT REPRESENTATIVE OF THE FINAL DATASET
-# age
-allsum %>%
-    group_by(handedness, gender) %>%
-    summarize(avg = mean(age), n = n(), sd = sd(age))
-
 #create a numerical code for handedness - can be used in plots to distinguish handedness groups (eg if pch set to handpch, get circles and triangles)
@@
 ```{r sightednesss, warning=FALSE}
-miles_dat_R <- read_csv("./data/Miles_R.csv", na = c("empty", "NA")) # read in data
-miles_dat_L <- read_csv("./data/Miles_L.csv", na = c("empty", "NA")) # read in data
-miles_dat <- rbind(miles_dat_R, miles_dat_L)
-#porta_dat_R <- read_csv("./dat2/Porta_R.csv", na = c("empty", "NA")) # read in data
-porta_dat_L <- read_csv("./data/Porta_L.csv", na = c("empty", "NA")) # read in data
-porta_dat <- rbind(porta_dat_L)
+miles_dat <- read_csv("./data/Miles.csv", na = c("empty", "NA")) # read in data
+porta_dat <- read_csv("./data/Porta.csv", na = c("empty", "NA")) # read in data
 # add test ID
@@
 # relabel 
-sight_dat$subject <- sight_dat$`Participant Public ID`
+sight_dat$subject <- sight_dat$Participant.Public.ID
 sight_dat$subject <- as.factor(sight_dat$subject)
@@
 ```{r EHI, warning=FALSE}
-EHI_R <- read_csv("./data/EHI_R.csv", na = "empty") # read in data
-EHI_L <- read_csv("./data/EHI_L.csv", na = "empty") # read in data
-EHI <- rbind(EHI_R, EHI_L)
+EHI <- read_csv("./data/EHI.csv", na = "empty") # read in data
 
 # select only responses
-EHI <- EHI[EHI$`Question Key` == "response-2-quantised" | EHI$`Question Key` == "response-3-quantised" | 
-           EHI$`Question Key` == "response-4-quantised" | EHI$`Question Key` == "response-5-quantised" |
-           EHI$`Question Key` == "response-6-quantised" | EHI$`Question Key` == "response-7-quantised" | 
-           EHI$`Question Key` == "response-8-quantised" | EHI$`Question Key` == "response-9-quantised" |
-           EHI$`Question Key` == "response-10-quantised" | EHI$`Question Key` == "response-11-quantised" | 
-           EHI$`Question Key` == "response-12-quantised" | EHI$`Question Key` == "response-13-quantised" |
-           EHI$`Question Key` == "response-14-quantised" | EHI$`Question Key` == "response-15-quantised" | 
-           EHI$`Question Key` == "response-16-quantised" | EHI$`Question Key` == "response-17-quantised" |
-           EHI$`Question Key` == "response-18-quantised" | EHI$`Question Key` == "response-19-quantised" | 
-           EHI$`Question Key` == "response-20-quantised",]
+EHI <- EHI[EHI$Question.Key == "response-2-quantised" | EHI$Question.Key == "response-3-quantised" | 
+           EHI$Question.Key == "response-4-quantised" | EHI$Question.Key == "response-5-quantised" |
+           EHI$Question.Key == "response-6-quantised" | EHI$Question.Key == "response-7-quantised" | 
+           EHI$Question.Key == "response-8-quantised" | EHI$Question.Key == "response-9-quantised" |
+           EHI$Question.Key == "response-10-quantised" | EHI$Question.Key == "response-11-quantised" | 
+           EHI$Question.Key == "response-12-quantised" | EHI$Question.Key == "response-13-quantised" |
+           EHI$Question.Key == "response-14-quantised" | EHI$Question.Key == "response-15-quantised" | 
+           EHI$Question.Key == "response-16-quantised" | EHI$Question.Key == "response-17-quantised" |
+           EHI$Question.Key == "response-18-quantised" | EHI$Question.Key == "response-19-quantised" | 
+           EHI$Question.Key == "response-20-quantised",]
 
 # recode variabels
-EHI$subject <- EHI$`Participant Public ID`
+EHI$subject <- EHI$Participant.Public.ID
 EHI$subject <- as.factor(EHI$subject)
@@
 
-ggplot(allsum, aes(x=index_EHI)) + geom_density(aes(group=handpch,fill=handpch))
+ggplot(allsum, aes(x=index_EHI)) + geom_density(aes(group=handpch,fill=handpch)) + theme_classic()
 
@@
 # read data
-lexTALE_L <- read_csv("./data/LexTale_L.csv", na = "empty")
-lexTALE_R <- read_csv("./data/lexTale_R.csv", na = "empty")
-lexTALE <- rbind(lexTALE_R, lexTALE_L)
+lexTALE <- read_csv("./data/LexTALE.csv", na = "empty")
 # filter
@@
 # rename variables
-lexTALE$subject <- lexTALE$`Participant Public ID`
+lexTALE$subject <- lexTALE$Participant.Public.ID
 lexTALE$accuracy <- lexTALE$Correct
@@
 # create data frame
+nsub <-length(levels(lexTALE$subject))
 lexTale_score <- data.frame(matrix(ncol = 4, nrow = nsub))
@@
 
+ggplot(allsum, aes(x=lexTALE)) + geom_density(aes(group=handpch,fill=handpch)) + theme_classic()
+
 hist(allsum$lexTALE)
@@
 for (d in 1:2){
-  data_file <- "./data/day1_combined.csv"
-
+  data_file <- "./data/combined/day1_combined.csv"
   if (d ==2){
-    data_file <-  "./data/day2_combined.csv"
+    data_file <- "./data/combined/day2_combined.csv"
   }
@@
 # write as factors
-all_dat2$subject <- as.factor(all_dat2$Participant.Public.ID)
+all_dat2$subject <- as.factor(trimws(all_dat2$Participant.Public.ID))
 all_dat2$RT <- as.numeric(as.character(all_dat2$Reaction.Time))
@@
 
+# mark Ps with too many trials indicating issues with the task or cases where less than 80% of data is left
+# we mark this to exclude in the next section
+out2 <- subset(data.frame(outliertable), Var2== 0)
+out2 <- out2[out2$Freq > 96 | out2$Freq < 76,]
+
 if (d==1)
 {CF_dat1 <- CF_dat2
- CF_same1 <- CF_same2}
-}
+ CF_same1 <- CF_same2
+ out1 <- out2}
+}
+# bind one long list of those who have missing trials or too much data
+CF_out <- rbind(out1, out2)
 ```
@@
 emotion_mean <- aggregate(FUN= mean, data= CF_same, emotion_recog~ subject)
+emotion_mean$subject <- as.factor(trimws(as.character(emotion_mean$subject)))
 # add this to allsum 
@@
 allsum$exclude[w]<-1 #code exclude = 1 if excluded because poor emot recog
+# now mark those with too many or too few
+allsum$exclude[allsum$subject %in% CF_out$Var1] <- 100
+
 ```
@@
 if (length(w)>0) {chim_dat2<-chim_dat2[-w,]}
+
+chim_dat1[is.na(chim_dat1)] <- 0
+chim_dat2[is.na(chim_dat2)] <- 0
 ```
@@
 
-cor1 <- cor(allsum$Day1_CF_acc_LI,allsum$Day2_CF_acc_LI)
+cor1 <- cor(allsum$Day1_CF_acc_LI,allsum$Day2_CF_acc_LI, use = "complete.obs")
 plot(allsum$Day1_CF_acc_LI,allsum$Day2_CF_acc_LI,main='Chimeric LI accuracy',col=(1+allsum$exclude))
@@
 text(-50,65,'red will be excluded',cex=.85)
-cor2<-cor(allsum$Day1_CF_RT_LI,allsum$Day2_CF_RT_LI)
+cor2<-cor(allsum$Day1_CF_RT_LI,allsum$Day2_CF_RT_LI, use = "complete.obs")
 plot(allsum$Day1_CF_RT_LI,allsum$Day2_CF_RT_LI,main='Chimeric LI RT',col=(1+allsum$exclude))
@@
 abline(h=0)
-text(-10,20,paste0('r = ',round(cor2,3)))
+text(-10,22,paste0('r = ',round(cor2,3)))
 text(-10,18,'red will be excluded',cex=.85)
@@
 #Removal of the outlier does not have much influence on rs
-
 ```
@@
 
+# remove those with too many trials indicating issues with the task
+# we mark this to exclude in the next section
+out2 <- subset(data.frame(outliertable), Var2== 0)
+out2 <- out2[out2$Freq > 120 | out2$Freq < 96,]
+
 if (d==1)
 {dichotic_dat1 <- dichotic_dat2
- dichotic_same1 <- dichotic_same2}
-}
+ dichotic_same1 <- dichotic_same2
+ out1 <- out2}
+}
+# bind one long list of those who have missing trials or too much data
+DL_out <- rbind(out1, out2)
 ```
@@
 sound_mean <- aggregate(FUN= mean, data= dichotic_same, sound_recog~ subject)
+sound_mean$subject <- as.factor(trimws(as.character(sound_mean$subject)))
 # add this to allsum 
@@
 allsum$exclude[ww]<-12 #code for those excluded on both chimeric and dichotic
+# remove those with too many or missing trials
+allsum$exclude[allsum$subject %in% DL_out$Var1] <- 100
+# table to summarise
 t <- table(allsum$exclude) 
-rownames(t)<-c('include','exclude CF','exclude dich','exclude CF+dich')
+rownames(t)<-c('include','exclude CF','exclude dich','exclude CF+dich','trials')
 t
@@
 
-cor1 <- cor(allsum$Day1_DL_acc_LI,allsum$Day2_DL_acc_LI)
+cor1 <- cor(allsum$Day1_DL_acc_LI,allsum$Day2_DL_acc_LI, use = "complete.obs", method = "pearson")
 plot(allsum$Day1_DL_acc_LI,allsum$Day2_DL_acc_LI,main='Dichotic LI accuracy',col=(1+allsum$exclude))
@@
 text(-75,75,paste0('r = ',round(cor1,3)))
-cor2<-cor(allsum$Day1_DL_RT_LI,allsum$Day2_DL_RT_LI)
+cor2<-cor(allsum$Day1_DL_RT_LI,allsum$Day2_DL_RT_LI, use = "complete.obs")
 plot(allsum$Day1_DL_RT_LI,allsum$Day2_DL_RT_LI,main='Dichotic LI RT',col=(1+allsum$exclude))
@@
 mycolname<-"Day1_Dich_acc_Z"
-mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
+mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.AP: chekced against pre-reg
 polarity<- (-1)#so LI is positive in predicted direction
@@
 
-# now remove the cases where there are too many or to few cases
-#remove.outlier <- as.data.frame(outliertable)
-#remove.outlier <- remove.outlier[remove.outlier$Freq >= 260 & remove.outlier$Freq <= 208,]
-#remove.p <- list(unique(remove.outlier$Var1))
-#allsum <- allsum[allsum$subject != remove.p,]
+# remove those with too many trials indicating issues with the task
+out2 <- subset(data.frame(outliertable), Var2== 0)
+out2 <- out2[out2$Freq > 260 | out2$Freq < 208,]
 
 if(d==1)
-{RDT_dat1 <- RDT_dat2}
-
-}
+{RDT_dat1 <- RDT_dat2
+ out1 <- out2}
+}
+# bind one long list of those who have missing trials or too much data
+RDT_out <- rbind(out1, out2)
+
+# once again, mark rather than remove those with poor performance on identical stimuli
+allsum$exclude[allsum$subject %in% RDT_out$Var1] <- 100
+# table to summarise
+t <- table(allsum$exclude) 
+rownames(t)<-c('include','exclude CF','exclude dich','exclude CF+dich','trials')
+t
 ```
@@
 polarity<- 1 #so LI is positive in predicted direction
-mycutoff <- 75
+mycutoff <- 10 
 allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
@@
 
-finger <- read.csv("day1_combined.csv", na.strings = "NA")
-
+finger <- read.csv("data/combined/day1_combined.csv", na.strings = "NA")
 # cleaning
@@
 
-finger <- read.csv("day2_combined.csv", na.strings = "NA")
-
+finger <- read.csv("data/combined/day2_combined.csv", na.strings = "NA")
 # cleaning
@@
 ```{r Finalallsum, warning=FALSE}
+# comment out to get participants with full data set
+allsum <- allsum[allsum$exclude == 0,]
+
 # first, describe age for the whole sample and plot
@@
 a <- allsum %>%
-  group_by(gender, handedness) %>%
+  group_by(handedness) %>%
     summarize(n = n())
 ggplot(data=a, aes(x=handedness, y=n, fill= handedness)) +
-  geom_bar(stat="identity")+
+  geom_bar(stat="identity", alpha= .5, colour="black")+
   geom_text(aes(label=n), vjust=1.6, color="white", size=3.5)+
-  theme_classic() + facet_wrap(.~ gender) + ggtitle("Gender by handedness") + 
+  theme_minimal() + ggtitle("Handedness") + 
   ylab("Count") + xlab("")
@@
 ```{r corr_between_tasks}
+
 mycols <- c('Day1_CF_acc_Z','Day2_CF_acc_Z','Day1_Dich_acc_Z','Day2_Dich_acc_Z','tlat_RDT_1','tlat_RDT_2','Day1_Finger_acc_Z','Day2_Finger_acc_Z')
@@
 #compare L and R handers
-
 for (i in colnums){
@@
 
+# corrplot the correlation matrix
+mycols <- c('Day1_CF_acc_LI', 'Day1_CF_acc_Z','Day2_CF_acc_LI', 'Day2_CF_acc_Z',
+            'Day1_DL_acc_LI', 'Day1_Dich_acc_Z','Day2_DL_acc_LI', 'Day2_Dich_acc_Z',
+            'Day1.RDT_RT_LI', 'tlat_RDT_1', 'Day2.RDT_RT_LI', 'tlat_RDT_2',
+            'Day1_finger_LI','Day1_Finger_acc_Z', 'Day2_finger_LI','Day2_Finger_acc_Z')
+colnums<- which(colnames(allsum) %in% mycols)
+M <- allsum[,colnums]
+M <- M %>% 
+  rename(
+    "CF (z) 1" = Day1_CF_acc_Z,
+    "CF (z) 2" = Day2_CF_acc_Z,
+    "DL (z) 1" = Day1_Dich_acc_Z,
+    "DL (z) 2" = Day2_Dich_acc_Z,
+    "RD (t) 1" = tlat_RDT_1,
+    "RD (t) 2" = tlat_RDT_2,
+    "FT (z) 1" = Day1_Finger_acc_Z,
+    "FT (z) 2" = Day2_Finger_acc_Z,
+    "CF (LI) 1" = Day1_CF_acc_LI,
+    "CF (LI) 2" = Day2_CF_acc_LI,
+    "DL (LI) 1" = Day1_DL_acc_LI,
+    "DL (LI) 2" = Day2_DL_acc_LI,
+    "RD (LI) 1" = Day1.RDT_RT_LI,
+    "RD (LI) 2" = Day2.RDT_RT_LI,
+    "FT (LI) 1" = Day1_finger_LI,
+    "FT (LI) 2" = Day2_finger_LI)
+# PLOT TO SEE IF DIFFERENT FROM ZERO
+# rearrange df
+N <- M %>% select("CF (LI) 1", "CF (LI) 2", "CF (z) 1", "CF (z) 2",
+             "DL (LI) 1", "DL (LI) 2", "DL (z) 1", "DL (z) 2",
+             "RD (LI) 1", "RD (LI) 2", "RD (t) 1", "RD (t) 2",
+             "FT (LI) 1", "FT (LI) 2", "FT (z) 1", "FT (z) 2")
+# rearrange for pirate plot
+# set participant ID
+N$ID <- 1:nrow(N)
+# format to long
+data_long <- tidyr::gather(N, task, index, `CF (LI) 1`:`FT (z) 2`, factor_key=TRUE)
+# add task identifer, session, etc
+data_long <- data_long %>%
+    mutate(group= if_else(task == "CF (LI) 1" | task == "CF (LI) 2" | task == "DL (LI) 1" | task == "DL (LI) 2" |
+                          task == "RD (LI) 1" | task == "RD (LI) 2" | task == "FT (LI) 1" | task == "FT (LI) 2",
+                          "LI", "t/z"),
+           session= if_else(task == "CF (LI) 1" | task == "CF (z) 1" |
+                            task == "DL (LI) 1" | task == "DL (z) 1" |
+                            task == "RD (LI) 1" | task == "RD (t) 1" |
+                            task == "FT (LI) 1" | task == "FT (z) 1", "1", "2"),
+           task2= if_else(task == "CF (LI) 1" | task == "CF (LI) 2" | task == "CF (z) 1" | task == "CF (z) 2", "Chimeric faces",
+                  if_else(task == "DL (LI) 1" | task == "DL (LI) 2" | task == "DL (z) 1" | task == "DL (z) 2", "Dichotic listening",
+                  if_else(task == "RD (LI) 1" | task == "RD (LI) 2" | task == "RD (t) 1" | task == "RD (t) 2", "Rhyme decision",
+                  "Finger Tapping"))))
+# PLOT CORRELATION BETWEEN DAYS
+# create matrix
+M<-cor(M)
+# removed odds and evens on each axis for plotting
+dat <- M
+toDeleteA <- seq(2, nrow(M), 2)
+toDeleteB <- seq(1, ncol(M), 2)
+dat <- dat[ toDeleteA , toDeleteB ]
+dat=dat[,order(ncol(dat):1)]
+# plot
+ggcorrplot::ggcorrplot(dat, lab = TRUE, 
+                       ggtheme = ggplot2::theme_minimal,
+                       colors = c("#6D9EC1", "white", "#E46726"),
+                       show.legend = FALSE, digits = 2, 
+                       title = "Correlation matrix for LIs on the four behavioural laterality tasks")
+
+# correlate each
+# dichotic
+DL_scatter <- ggplot(allsum, aes(x= Day1_DL_acc_LI, y= Day2_DL_acc_LI)) +
+  geom_hline(yintercept=0, color = "black") +
+  geom_vline(xintercept=0, color = "black") +
+  geom_point(aes(color= handedness)) +
+  geom_rug(aes(color= handedness)) +
+  theme_classic() +
+  ylab("Dichotic listening LI (day 2)") + 
+  xlab("Dichotic listening LI (day 1)") +
+  xlim(-100, 100) + ylim(-100, 100) +
+  annotate("text", x = 90, y = 90, label = "paste(italic(R), \" = .72\")", parse = TRUE) +
+  theme(legend.position = c(.15, .90))
+# Set relative size of marginal plots (main plot 10x bigger than marginals)
+ggExtra::ggMarginal(DL_scatter, groupColour = TRUE, groupFill = TRUE)
+# chimeric
+CF_scatter <- ggplot(allsum, aes(x= Day1_CF_acc_LI, y= Day2_CF_acc_LI)) +
+  geom_hline(yintercept=0, color = "black") +
+  geom_vline(xintercept=0, color = "black") +
+  geom_point(aes(color= handedness)) +
+  geom_rug(aes(color= handedness)) +
+  theme_classic() +
+  ylab("Chimeric faces LI (day 2)") + 
+  xlab("Chimeric faces LI (day 1)") +
+  xlim(-100, 100) + ylim(-100, 100) +
+  annotate("text", x = 90, y = 90, label = "paste(italic(R), \" = .88\")", parse = TRUE) +
+  theme(legend.position = c(.15, .90))
+# Set relative size of marginal plots (main plot 10x bigger than marginals)
+ggExtra::ggMarginal(CF_scatter, groupColour = TRUE, groupFill = TRUE)
+# rhyme
+rhyme_scatter <- ggplot(allsum, aes(x= Day1.RDT_RT_LI, y= Day2.RDT_RT_LI)) +
+  geom_hline(yintercept=0, color = "black") +
+  geom_vline(xintercept=0, color = "black") +
+  geom_point(aes(color= handedness)) + 
+  geom_rug(aes(color= handedness)) +
+  theme_classic() +
+  ylab("Rhyme decision LI (day 2)") + 
+  xlab("Rhyme decision LI (day 1)") +
+  annotate("text", x = 8, y = 8, label = "paste(italic(R), \" = .63\")", parse = TRUE) +
+  theme(legend.position = c(.15, .90))
+# Set relative size of marginal plots (main plot 10x bigger than marginals)
+ggExtra::ggMarginal(rhyme_scatter, groupColour = TRUE, groupFill = TRUE)
+# finger
+finger_scatter <- ggplot(allsum, aes(x= Day1_finger_LI, y= Day2_finger_LI)) +
+  geom_hline(yintercept=0, color = "black") +
+  geom_vline(xintercept=0, color = "black") +
+  geom_point(aes(color= handedness)) +
+  geom_rug(aes(color= handedness)) +
+  theme_classic() +
+  ylab("Finger tapping LI (day 2)") + 
+  xlab("Finger tapping LI (day 1)") +
+  annotate("text", x = 18, y = 18, label = "paste(italic(R), \" = .89\")", parse = TRUE) +
+  theme(legend.position = c(.15, .90))
+# Set relative size of marginal plots (main plot 10x bigger than marginals)
+ggExtra::ggMarginal(finger_scatter, groupColour = TRUE, groupFill = TRUE)
 ```

### 2020-05-23 battery-preprocessing_v5_DB.Rmd
some cosmetic changes; also ensured CF task was scored correctly, which had effect on outlier detection (previously all scored correct); also made subject codes simple numeric to make it easier to track errors. In doing the latter, may have messed up with a search and replace, so needs more checking, which is inprogress!

@@
 author: "Adam Parker + Dorothy Bishop"
-date: "updated April 22nd 2020"
-output:
-  pdf_document: default
-  html_document: default
+date: "updated May 21st 2020 (now notebook format)"
+output: html_notebook
 ---
+22 May 2020: AP made changes to N trials needed for inclusion  
 NB. 2 changes Apr 22nd by DB: 
@@
 
-This R Markdown file prcoesses the various components of the Online Laterality Battery which is implemented online via Gorilla.sc.
+This R Markdown file processes the various components of the Online Laterality Battery which is implemented online via Gorilla.sc.
 
@@
 
+NB. Just also putting here relevant references that need to be looked at!
+https://pubmed.ncbi.nlm.nih.gov/18050003/
+
 # Questionnaires
@@
 require(dplyr)  # for mutate()
-require(tidyr)  # for unnest()
 require(purrr)  # for map(), reduce()
@@
 
-First, the demographics is processed.This gives us a participant ID, age, gender, footedness, and handedness.
+First, the demographics is processed.This gives us a participant subject, age, gender, footedness, and handedness.
 
 ```{r demograhics, warning=FALSE}
-demo_dat_R <- read_csv("./data/demographics_R.csv", na = c("empty", "NA")) # read in data
-demo_dat_L <- read_csv("./data/demographics_L.csv", na = c("empty", "NA")) # read in data
-demo_dat <- rbind(demo_dat_R, demo_dat_L)
+#Reading from OSF - this only works if files are public
+
+demo_dat <- read.csv('https://osf.io/7smj6/download',stringsAsFactor=F)
+
 # relabel 
-demo_dat$subject <- demo_dat$`Participant Public ID`
+demo_dat$subject <- demo_dat$Participant.Public.subject
 demo_dat$subject <- as.factor(demo_dat$subject)
@@
 for (i in 1:nsub) { # loop through subjects
-  subname <- levels(demo_dat$subject)[i] # find subject ID
+  subname <- levels(demo_dat$subject)[i] # find subject subject
   myrows <- which(demo_dat$subject==subname) # select rows for this subject
@@
 
-    myrow <- myrow+1
+    myrow <- i
 
     allsum$subject[myrow] <- subname
-    allsum$age[myrow] <- as.numeric(tmp[tmp$Question.Key == "Age",]$Response)
+    allsum$age[myrow] <- tmp[tmp$Question.Key == "Age",]$Response
     allsum$gender[myrow] <- tmp[tmp$Question.Key == "Gender",]$Response
@@
 
+#Check ages
+range(allsum$age)
 # Here we summarise the age broken down by handedness and gender. This will be important for any write up. 
@@
 #create a numerical code for handedness - can be used in plots to distinguish handedness groups (eg if pch set to handpch, get circles and triangles)
-allsum$handpch <-1
+allsum$handpch <-1  #numeric code for plotting handedness
 allsum$handpch[allsum$handedness=='Left'] <-2
@@
 ```{r sightednesss, warning=FALSE}
-miles_dat_R <- read_csv("./data/Miles_R.csv", na = c("empty", "NA")) # read in data
-miles_dat_L <- read_csv("./data/Miles_L.csv", na = c("empty", "NA")) # read in data
-miles_dat <- rbind(miles_dat_R, miles_dat_L)
-#porta_dat_R <- read_csv("./dat2/Porta_R.csv", na = c("empty", "NA")) # read in data
-porta_dat_L <- read_csv("./data/Porta_L.csv", na = c("empty", "NA")) # read in data
-porta_dat <- rbind(porta_dat_L)
-# add test ID
-miles_dat$test_ID <- "miles"
-porta_dat$test_ID <- "porta"
+porta_dat <- read.csv("https://osf.io/zmqkt/download",stringsAsFactors = F)
+miles_dat <-  read.csv("https://osf.io/ekqsp//download",stringsAsFactors = F)
+
+miles_dat$test_subject <- "miles"
+porta_dat$test_subject <- "porta"
 # merge the two
@@
 # relabel 
-sight_dat$subject <- sight_dat$`Participant Public ID`
+sight_dat$subject <- sight_dat$Participant.Public.subject
 sight_dat$subject <- as.factor(sight_dat$subject)
@@
 for (i in 1:nsub) { # loop through subjects
-  subname <- levels(sight_dat$subject)[i] # find subject ID
+  subname <- levels(sight_dat$subject)[i] # find subject subject
   myrows <- which(sight_dat$subject==subname) # select rows for this subject
@@
 
-    myrow <- myrow+1
+    myrow <- i
 
@@
 ```{r EHI, warning=FALSE}
-EHI_R <- read_csv("./data/EHI_R.csv", na = "empty") # read in data
-EHI_L <- read_csv("./data/EHI_L.csv", na = "empty") # read in data
-EHI <- rbind(EHI_R, EHI_L)
+
+EHI <- read.csv("https://osf.io/kfdma/download",stringsAsFactors = F)
 
 # select only responses
-EHI <- EHI[EHI$`Question Key` == "response-2-quantised" | EHI$`Question Key` == "response-3-quantised" | 
-           EHI$`Question Key` == "response-4-quantised" | EHI$`Question Key` == "response-5-quantised" |
-           EHI$`Question Key` == "response-6-quantised" | EHI$`Question Key` == "response-7-quantised" | 
-           EHI$`Question Key` == "response-8-quantised" | EHI$`Question Key` == "response-9-quantised" |
-           EHI$`Question Key` == "response-10-quantised" | EHI$`Question Key` == "response-11-quantised" | 
-           EHI$`Question Key` == "response-12-quantised" | EHI$`Question Key` == "response-13-quantised" |
-           EHI$`Question Key` == "response-14-quantised" | EHI$`Question Key` == "response-15-quantised" | 
-           EHI$`Question Key` == "response-16-quantised" | EHI$`Question Key` == "response-17-quantised" |
-           EHI$`Question Key` == "response-18-quantised" | EHI$`Question Key` == "response-19-quantised" | 
-           EHI$`Question Key` == "response-20-quantised",]
+EHI <- EHI[EHI$Question.Key == "response-2-quantised" | EHI$Question.Key == "response-3-quantised" | 
+           EHI$Question.Key == "response-4-quantised" | EHI$Question.Key == "response-5-quantised" |
+           EHI$Question.Key == "response-6-quantised" | EHI$Question.Key == "response-7-quantised" | 
+           EHI$Question.Key == "response-8-quantised" | EHI$Question.Key == "response-9-quantised" |
+           EHI$Question.Key == "response-10-quantised" | EHI$Question.Key == "response-11-quantised" | 
+           EHI$Question.Key == "response-12-quantised" | EHI$Question.Key == "response-13-quantised" |
+           EHI$Question.Key == "response-14-quantised" | EHI$Question.Key == "response-15-quantised" | 
+           EHI$Question.Key == "response-16-quantised" | EHI$Question.Key == "response-17-quantised" |
+           EHI$Question.Key == "response-18-quantised" | EHI$Question.Key == "response-19-quantised" | 
+           EHI$Question.Key == "response-20-quantised",]
 
 # recode variabels
-EHI$subject <- EHI$`Participant Public ID`
+EHI$subject <- EHI$Participant.Public.subject
 EHI$subject <- as.factor(EHI$subject)
@@
 
-# score responses according to the original scoring method in Oldifeld (1971)
+# score responses according to the original scoring method in Oldfield (1971)
 EHI$right_hand <- 0
@@
 
-myrow <- 0 # start row counter
 for (i in 1:nsub) { # loop through subjects
-  myrow <- myrow+1
-  subname <- levels(EHI$subject)[i] # find subject ID
+  myrow <- i
+  subname <- levels(EHI$subject)[i] # find subject subject
   myrows <- which(EHI$subject==subname) # select rows for this subject
@@
     EHI2$subject[myrow] <- subname # add row for subject
-    EHI2$right_hand[myrow] <- sum(tmp$right_hand) # add row for hand
-    EHI2$left_hand[myrow] <- sum(tmp$left_hand) # add row for number of buttons pressed
-  }
-
-# This is a summary of the EHI
-for_index2 <- 
-  EHI2 %>% 
-  group_by(subject) %>%
-  mutate(index_EHI= ((right_hand-left_hand)/(right_hand + left_hand))*100)
+    EHI2$right_hand[myrow] <- sum(tmp$right_hand) # sum for R hand
+    EHI2$left_hand[myrow] <- sum(tmp$left_hand) # sum for L hand
+  }
+
+EHI2$index_EHI <- 100*(EHI2$right_hand-EHI2$left_hand)/(EHI2$right_hand+EHI2$left_hand)
 
 # now create a data frame only for index 
-EHI_index <- select(for_index2, "subject", "index_EHI")
 # merge with allsum
-allsum <- merge(allsum, EHI_index, by= "subject")
+allsum$index_EHI <- NA #initialise
+for (i in 1:nrow(allsum)){
+  thissub <- allsum$subject[i]
+  w<-which(EHI2$subject==thissub)
+ allsum$index_EHI[i] <- EHI2$index_EHI[w]
+}
 
@@
 eplot<-ggplot(allsum, aes(handpch,index_EHI, colour=gender)) +
-  geom_jitter(width=.25,height=2)+
+  geom_jitter(wsubjectth=.25,height=2)+
   ggtitle("Handedness index on the Edinburgh Handedness Inventory")
 eplot
-#This is not perfect, but gives and idea to check EHI is valid etc.
+#This is not perfect, but gives and subjectea to check EHI is valid etc.
 
@@
 
-#NB one supposed right-hander has index_EHI around -50. 
-#I will recode this one, as this seems an error.
+#NB two supposed right-handers have index_EHI around -50. 
+#I will recode these, as this seems an error.
 
@@
 # read data
-lexTALE_L <- read_csv("./data/LexTale_L.csv", na = "empty")
-lexTALE_R <- read_csv("./data/lexTale_R.csv", na = "empty")
-lexTALE <- rbind(lexTALE_R, lexTALE_L)
+lexTALE <- read.csv("https://osf.io/h4btp/download",stringsAsFactors = F)
+
 # filter
@@
   colnames(lexTale_score) <- c("subject", "Word", "nonWord", "lexTALE")
-myrow <- 0 # start row counter
+
 for (i in 1:nsub) { # loop through subjects
-  subname <- levels(lexTALE$subject)[i] # find subject ID
+  subname <- levels(lexTALE$subject)[i] # find subject subject
   myrows <- which(lexTALE$subject==subname) # select rows for this subject
@@
 
-# remove those who didn't pass the LexTALE
+# remove those who dIDn't pass the LexTALE
 allsum <- allsum[allsum$lexTALE >= 80,]
@@
 Before doing anything with the data, we now pull in all the data from all nodes and combine them from a sub folder in the directory.
-Participants, identified by random letter strings, are stacked on top of each other.
+Participants, IDentified by random letter strings, are stacked on top of each other.
 We start by reading the data and cutting out unwanted information.
@@
 for (d in 1:2){
-  data_file <- "./data/day1_combined.csv"
+  data_file <- "https://osf.io/mhk29/download" #day 1 data
 
   if (d ==2){
-    data_file <-  "./data/day2_combined.csv"
+    data_file <-  "https://osf.io/ebm6s/download"
   }
@@
 all_dat2 <- all_dat2[,collist]
-
-#Check for overlap in subjects between the two sessions
-which(levels(all_dat1$subject) %in% levels(all_dat2$subject))
+```
+
+Create a simple numerical ID to make life easier if we need to identify datapoints which are odd in some way. Ensure this is consistent across all_dat1 and all_dat2 and allsum.
+We will rename the current subject variable so we can reuse this one  for all future computations.
+
+```{r makesubject}
+#Use allsum as the basis for the check
+#First rename all subject to origID
+allsum <- allsum %>% rename(origID = subject) 
+all_dat1 <- all_dat1 %>% rename(origID = subject) 
+all_dat2 <- all_dat2 %>% rename(origID = subject) 
+for (n in 1:nrow(allsum)){
+  allsum$subject <- n
+  w<-which(all_dat1$origID == allsum$origID[n])
+  all_dat1$subject <-n
+  w<-which(all_dat2$origID == allsum$origID[n])
+  all_dat2$subject <-n
+}
 ```
@@
 
+Outlier removal put in separate function, which can be applied to different task RTs.  
+Origdat is the dataframe with data for outlier removal, mycolname is name of variable  
+For outlier removal (usually RT), lowcut is a cutoff value below which data excluded (typically v fast responses that are implausible, e.g. 201 ms for RT)
+ and zcut is the constant in Hoaglin-Iglewicz method (usually 2.2)
+
 ```{r RToutlierfunction}
-remove.outliers <- function(origdat,lowcut,zcut){
-# Outlier removal put in separate function, which can be applied to different task RTs.
-# 
-# origdat is the dataframe with data for outlier removal, mycolname is name of variable 
-# for outlier removal (usually RT), lowcut is a cutoff value below which data excluded
-# (typically v fast responses that are implausible, e.g. 201 ms for RT)
-# and zcut is the constant in Hoaglin-Iglewicz method (usually 2.2)
+detect.outliers <- function(origdat,lowcut,zcut){
 
@@
   for (i in 1:nsub) { # loop through subjects
-  subname <- levels(origdat$subject)[i] # find subject ID
+  subname <- levels(origdat$subject)[i] # find subject 
   myrows <- which(origdat$subject==subname) # select rows for this subject
@@
 
- # Identify 25th and 75th quartiles, and the difference between them
+ # IDentify 25th and 75th quartiles, and the difference between them
           lower_quartile <- quantile(RTcorr, probs=0.25, na.rm="TRUE")
@@
 for (i in 1:nsub) { # loop through subjects
-  subname <- levels(origdat$subject)[i] # find subject ID
+  subname <- levels(origdat$subject)[i] # find subject subject
   myrows <- which(origdat$subject==subname) # select rows for this subject
@@
     temp_dat$subject[myrow] <- subname # add row for subject
-    temp_dat$side[myrow] <- latlist[j] # add row for side
+    temp_dat$side[myrow] <- latlist[j] # add row for sIDe
     temp_dat$accurate[myrow] <- sum(tmp1$Correct,na.rm=TRUE) # add row for N accurate response
     temp_dat$p.corr[myrow] <- 100*mean(tmp1$Correct,na.rm=TRUE) # add row for %accurate responses - NB for chimeric only error is wrong emotion - measure is just a measure of choice.
-For RDT, can make error of selecting wrong side, so % correct is meaningful
+For RDT, can make error of selecting wrong sIDe, so % correct is meaningful
 
@@
 dopirate <- function(origdat,task,measure){
-# measure for each side
-# plot the number of correct responses in each side
+# measure for each sIDe
+# plot the number of correct responses in each sIDe
+  myhead<-paste0(task,": ",measure)
 pirateplot(formula = X1 ~ side,
            data = origdat,
-           main = measure,
+           main = myhead,
            theme = 0,
@@
 
-makeLI <- function(allsums,origdat,task,mycolname,polarity,mycutoff){
+makeLI <- function(allsum,origdat,task,mycolname,polarity,mycutoff){
 # we do this for accuracy first
@@
 keeps <- c("subject", "side", "X1") #X1 holds either accuracy or RT
-LIdf <- origdat[keeps]
-LIdf <- spread(LIdf, side, X1)
+Lsubjectf <- origdat[keeps]
+Lsubjectf <- spread(Lsubjectf, side, X1)
 # calculate each participants' lat index
-LIdf <- 
-  LIdf %>% 
-  group_by(subject) %>%
-  mutate(thisLI= ((Right - Left)/(Right + Left))*100)
-
-LIdf$LI <- LIdf$thisLI*polarity #polarity can be set to reverse sign - eg for RT a large score is bad, whereas for acc it is good, so can have opp polarity
+
+Lsubjectf$thisLI <- 100*(Lsubjectf$Right - Lsubjectf$Left)/(Lsubjectf$Right + Lsubjectf$Left)
+
+Lsubjectf$LI <- Lsubjectf$thisLI*polarity #polarity can be set to reverse sign - eg for RT a large score is bad, whereas for acc it is good, so can have opp polarity
 
@@
 # LI will be set to NA for those outside limits
-LIdf$avg <- (LIdf$Left+LIdf$Right)/2
-w<-which(LIdf$avg<mycutoff)
+Lsubjectf$avg <- (Lsubjectf$Left+Lsubjectf$Right)/2
+w<-which(Lsubjectf$avg<mycutoff)
 if(length(w)>0){
-LIdf$LI[w] <- NA}
+Lsubjectf$LI[w] <- NA}
 
 # now create a data frame for lat index 
-LIdf <- select(LIdf, "subject", "LI")
+Lsubjectf <- select(Lsubjectf, "subject", "LI")
 # now merge with allsum
-allsums <- merge(allsums, LIdf, by= "subject")
-lastcol<-length(colnames(allsums))
-colnames(allsums)[lastcol]<-mycolname
-
-return(allsums)
+# First check if this column already exists and delete it if so
+w<-which(colnames(allsum)==mycolname)
+allsum<-allsum[,-w]
+#now add this LI
+allsum <- merge(allsum, Lsubjectf, by= "subject")
+lastcol<-length(colnames(allsum))
+colnames(allsum)[lastcol]<-mycolname
+
+return(allsum)
 }
@@
 # now merge with allsum
+# First check if this column already exists and delete it if so
+w<-which(colnames(allsums)==mycolname)
+allsums<-allsums[,-w]
+
 allsums <- merge(allsums, Zdf, by= "subject")
@@
 
-This part of the R Markdown file prcoesses the "Chimeric Face Task' implmented online via Gorilla.sc which aims to assess the lateralisation of emotion recognition in a face processing task.
+This part of the notebook processes the "Chimeric Face Task' implmented online via Gorilla.sc which aims to assess the lateralisation of emotion recognition in a face processing task.
 In the task, participants are shown a chimeric face.
@@
 This task is near identical to that reported in Karlsson, Johnstone, and Carey (2019).
-Stimuli were kindly provided by Dr Michael Burt (https://www.dur.ac.uk/psychology/staff/?id=1942) and were symmetrical average images created from four male and four female faces (see Burt & Perrett, 1997; Innes et al., 2016).
+Stimuli were kindly provided by Dr Michael Burt (https://www.dur.ac.uk/psychology/staff/?subject=1942) and were symmetrical average images created from four male and four female faces (see Burt & Perrett, 1997; Innes et al., 2016).
+Stimuli for each combination of anger, disgust, happy and sad (8 left-first, 8 right-first) plus 8 of each emotion where both sides the same the same to check overall perception.
+Total of 96 chimeric stimuli presented in each session.
+We will exclude those who are inaccurate when both faces the same (as in preregistration).
+Responses to chimeric faces can be coded as whether they correspond to emotion on L or R.
+Note that if the reported emotion does not match either side, then trial is not counted.
+
+The chunk below also detects outliers, and codes them as 1 if the RT is outside the range specified in Hoaglin-Iglewicz, and as 2 if the RT is < 200.
+Outliertable shows each subject with Ns with different outlier codes.
+Some people have 0-1 trials - these seem to be ones where the task was skipped for some reason - I have checked and they seem to have no data.
+One person has twice the number of trials in one session. We'll deal with that by just taking the first 96 trials.
 
@@
 task <- "Chimeric faces"
-wanted <- c("subject", "stimuli", "Response", "RT", "l_hemiface", "r_hemiface") # select wanted columns
+wanted <- c("subject", "stimuli", "Response", "RT", "l_hemiface", "r_hemiface","Correct") # select wanted columns - nb added Correct here
 disp <- "e_Task"
-CF_dat2 <- procdata(origfile,wanted,task,disp) #use generic function (see above) to read in relevant columns
+CF_dat_all <- procdata(origfile,wanted,task,disp) #use generic function (see above) to read in relevant columns
+
+#Before detecting outliers need to add a column indicating if response was correct. Although we read in 'Correct' it is not accurate!
+
+w<-union(which(CF_dat_all$Response==CF_dat_all$l_hemiface),
+         which(CF_dat_all$Response==CF_dat_all$r_hemiface))
+CF_dat_all$Correct[w]<-1
 
 # count subjects
-nsub <-length(levels(CF_dat2$subject))
+nsub <-length(levels(CF_dat_all$subject))
 print(paste('day',d,task,": subjects: ", nsub))
 
-CF_same2 <- CF_dat2[CF_dat2$l_hemiface == CF_dat2$r_hemiface,] # create data frame for non-chimeras
-CF_dat2 <- CF_dat2[CF_dat2$l_hemiface != CF_dat2$r_hemiface,] # remove items where same expression in both hemifaces
-CF_dat2 <- remove.outliers(CF_dat2,200,1.65) #use generic outlier removal function
+CF_same2 <- CF_dat_all[CF_dat_all$l_hemiface == CF_dat_all$r_hemiface,] # create data frame for non-chimeras
+CF_dat2 <- CF_dat_all[CF_dat_all$l_hemiface != CF_dat_all$r_hemiface,] # create df for chimeras: remove items where same expression in both hemifaces
+
+CF_dat2 <- detect.outliers(CF_dat2,200,1.65) #use generic outlier detection function. The parameters are lowcut (200 ms cutoff for anticipations) and then the value used for Hoaglin Iglewicz (1.65)
 outliertable<-table(CF_dat2$subject,CF_dat2$outlier)
-#print(outliertable)
+#Most have 96 accepted trials in CF_dat2. A few have zero - these seem to be ones that had RTs < 200, and may have been responding rapidly at random.
+#One person has 192 trials - seems to have had whole thing presented twice.
 
@@
 First, we judge participants' ability to correctly identify emotions using the "CF_same" dataframe.
-Participants are removed if their performance is below 75% due to poor ability to identify emotions that will inevitably influence performance on the Chimeric Face Task.
+Participants are removed if their performance is below 75% as this indicates poor ability to identify emotions, which will inevitably influence performance on the Chimeric Face Task.
 
@@
 
-CF_same <- 
-  CF_same %>% 
-  mutate(emotion_recog= ifelse(as.character(Response) == as.character(l_hemiface), 1, 0))
+# removed code to compute correct, as already done above
 # calculate participants average
-emotion_mean <- aggregate(FUN= mean, data= CF_same, emotion_recog~ subject)
+emotion_mean <- aggregate(FUN= mean, data= CF_same, Correct~ subject)
 # add this to allsum 
@@
 }
+
+w<-which(colnames(allsum) %in% c("Emot.recog1","Emot.recog2"))
+if (length(w)>0){
+allsum<-allsum[,-w] #remove any columns already created with these names
+}
 mycol <- length(colnames(allsum))
+
 colnames(allsum)[(mycol-1):mycol]<-c('Emot.recog1','Emot.recog2')
 
-plot(jitter(allsum$Emot.recog1,10),jitter(allsum$Emot.recog2,10),pch=3,cex=.6)
+plot(jitter(allsum$Emot.recog1,10),jitter(allsum$Emot.recog2,10),pch=3,cex=.6,xlab='Nonchimeric emotion recog 1',ylab='Nonchimeric emotion recog 2')
 abline(h=.75)
@@
 allsum$exclude[w]<-1 #code exclude = 1 if excluded because poor emot recog
-```
-
-In this section, the correct sequences are marked as 1, incorrect asnwers are marked as 0. We then mark whether the correct answers were in the left or right hemiface.
+print(paste('N excluded because < 75% on at least one trial:',length(w)))
+```
+
+In this section, the correct sequences are marked as 1, incorrect answers are marked as 0. We then mark whether the correct answers were in the left or right hemiface.
 
@@
                        ifelse(as.character(Response) == as.character(r_hemiface) & Correct == 1, "Right", NA)))
+
+#Added coding of 'X' for responses of emotion not shown in either side
+w<-which(CF_dat$Correct==0)
+CF_dat$side[w] <- 'X'
 # code side as factor
@@
 if (length(w)>0) {chim_dat2<-chim_dat2[-w,]}
+
+#find included cases
+w<-which(allsum$exclude==0)
+inclsubs <- allsum$subject[w]
+
+chim_dat1 <- chim_dat1[(chim_dat1$subject %in% inclsubs),]
+chim_dat2 <- chim_dat2[(chim_dat2$subject %in% inclsubs),]
+
 ```
@@
 measure <- 'Accuracy'
-origdat$X1 <- chim_dat1$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
+origdat$X1 <- chim_dat1$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc with the dopirate function
 task <- 'Chimeric faces day 1'
@@
 text(-50,75,paste0('r = ',round(cor1,3)))
-text(-50,65,'red will be excluded',cex=.85)
+myn <- nrow(allsum)
+text(-50,85,paste0('N = ',myn))
 cor2<-cor(allsum$Day1_CF_RT_LI,allsum$Day2_CF_RT_LI)
-plot(allsum$Day1_CF_RT_LI,allsum$Day2_CF_RT_LI,main='Chimeric LI RT',col=(1+allsum$exclude))
+plot(allsum$Day1_CF_RT_LI,allsum$Day2_CF_RT_LI,main='Chimeric LI RT',col=(1+allsum$exclude)) #this would plot excluded as red, but they are now excluded at earlier stage so don't show up
 abline(v=0)
@@
 text(-10,20,paste0('r = ',round(cor2,3)))
-text(-10,18,'red will be excluded',cex=.85)
+
+text(-10,23,paste0('N = ',myn))
+#text(-10,18,'red will be excluded',cex=.85)
 ```
@@
 dichotic_dat2 <- dichotic_dat2[dichotic_dat2$left_channel != dichotic_dat2$right_channel,] # remove items where same expression in both hemifaces
-dichotic_dat2 <- remove.outliers(dichotic_dat2,200,1.65) #use generic outlier removal function
+dichotic_dat2 <- detect.outliers(dichotic_dat2,200,1.65) #use generic outlier removal function
 outliertable<-table(dichotic_dat2$subject,dichotic_dat2$outlier)
@@
 
-If the index is negative, emotion recognition is lateralised in the left hemisphere. If the index is positive, emotion recognition is lateralised in the right hemisphere
-
-Here, a .csv is written for the chimeric face LI.
-
-```{r domakeLIdich}
+If the index is negative, speech recognition is lateralised in the left hemisphere. If the index is positive, speech recognition is lateralised in the right hemisphere
+
+Here, a .csv is written for the dichotic LI.
+
+```{r domakeLidich}
 #We add Dichotic LIs to allsum for day 1 and then day2
@@
 
-cor1 <- cor(allsum$Day1_DL_acc_LI,allsum$Day2_DL_acc_LI)
+w<-which(allsum$exclude==2) #substitute NA for those who were excluded
+ncol<-length(colnames(allsum))
+allsum[w,(ncol-3):ncol]<-NA
+cor1 <- cor(allsum$Day1_DL_acc_LI,allsum$Day2_DL_acc_LI,use='complete.obs',method='spearman')
 plot(allsum$Day1_DL_acc_LI,allsum$Day2_DL_acc_LI,main='Dichotic LI accuracy',col=(1+allsum$exclude))
@@
 text(-75,75,paste0('r = ',round(cor1,3)))
-cor2<-cor(allsum$Day1_DL_RT_LI,allsum$Day2_DL_RT_LI)
+cor2<-cor(allsum$Day1_DL_RT_LI,allsum$Day2_DL_RT_LI,use='complete.obs',method='spearman')
 plot(allsum$Day1_DL_RT_LI,allsum$Day2_DL_RT_LI,main='Dichotic LI RT',col=(1+allsum$exclude))
@@
 text(-5,15,paste0('r = ',round(cor2,3)))
+
+#Plot of LI shows that two participants reported only from one ear (acc LI = 100).
+#Although we did not pre-register this, it would make sense to exclude these, who are clear outliers, doing the task quite differently from others.
+
+w<-which(allsum$Day1_DL_RT_LI== (-100))
+allsum$Day1_DL_RT_LI[w] <- NA
+allsum$Day1_DL_acc_LI[w] <- NA
+allsum$Day2_DL_RT_LI[w] <- NA
+allsum$Day2_DL_acc_LI[w] <- NA
+
+allsum$exclude[w]<-2 #mark as excluded for dichotic
+
+#Redo the correlation and plots
+cor1 <- cor(allsum$Day1_DL_acc_LI,allsum$Day2_DL_acc_LI,use='complete.obs',method='spearman')
+plot(allsum$Day1_DL_acc_LI,allsum$Day2_DL_acc_LI,main='Dichotic LI accuracy, after 2 exclusions',col=(1+allsum$exclude))
+abline(v=0)
+abline(h=0)
+text(-75,75,paste0('r = ',round(cor1,3)))
+cor2<-cor(allsum$Day1_DL_RT_LI,allsum$Day2_DL_RT_LI,use='complete.obs',method='spearman')
+plot(allsum$Day1_DL_RT_LI,allsum$Day2_DL_RT_LI,main='Dichotic LI RT, after 2 exclusions',col=(1+allsum$exclude))
+abline(v=0)
+abline(h=0)
+text(-5,15,paste0('r = ',round(cor2,3)))
+
 ```
@@
   levels(RDT_dat2$side)<-c("Left","Right")
-RDT_dat2<-remove.outliers(RDT_dat2,200,1.65) #numbers specify min RT and zscore for Hoaglin-Iglewicz respectively
+RDT_dat2<-detect.outliers(RDT_dat2,200,1.65) #numbers specify min RT and zscore for Hoaglin-Iglewicz respectively
 
@@
 varlist <-   c("subject", "side", "accurate", "RT") #need to be in this order, ie sub, side, acc and RR - we will use these in the make.df function
+#We first need to remove rows where RT is coded NA - these seem cases where false recording led to excess trials with no RT
+w<-which(is.na(RDT_dat1$RT))
+RDT_dat1<-RDT_dat1[-w,]
 origdat <- RDT_dat1
+
 latlist <- c("Left","Right") #names of factor levels
 RDT_df1 <- make.df(origdat,varlist,latlist)
+w<-which(is.na(RDT_dat2$RT))
+RDT_dat1<-RDT_dat2[-w,]
 origdat <- RDT_dat2

### 2020-05-23 battery-preprocessing_v5_DB.Rmd
Double-checked CF task analysis - excluded cases with 2 or less responses on one side.

@@
 ---
+23 May 2020: Created subject number for those included in final analysis.
+CF outlier detection needed updating to ensure only correct trials used for outliers - Correct had not been coded at that point in prior version
 22 May 2020: AP made changes to N trials needed for inclusion  
@@
 # relabel 
-demo_dat$subject <- demo_dat$Participant.Public.subject
+demo_dat$subject <- demo_dat$Participant.Public.ID
 demo_dat$subject <- as.factor(demo_dat$subject)
@@
 # relabel 
-sight_dat$subject <- sight_dat$Participant.Public.subject
+sight_dat$subject <- sight_dat$Participant.Public.ID
 sight_dat$subject <- as.factor(sight_dat$subject)
@@
 allsum <- merge(allsum, sightedness, by= "subject")
+head(allsum) #just to check all looking OK
 ```
@@
 # recode variabels
-EHI$subject <- EHI$Participant.Public.subject
+EHI$subject <- EHI$Participant.Public.ID
 EHI$subject <- as.factor(EHI$subject)
@@
 
-Before doing anything with the data, we now pull in all the data from all nodes and combine them from a sub folder in the directory.
-Participants, IDentified by random letter strings, are stacked on top of each other.
+Before doing anything with the data, we now pull in all the data from OSF and combine them.
+Participants, identified by random letter strings, are stacked on top of each other.
 We start by reading the data and cutting out unwanted information.
+NB This step is quite slow (slower if reading from OSF than from drive).
 
@@
 
-Create a simple numerical ID to make life easier if we need to identify datapoints which are odd in some way. Ensure this is consistent across all_dat1 and all_dat2 and allsum.
-We will rename the current subject variable so we can reuse this one  for all future computations.
+Create a simple numerical ID to make life easier when screening data if we need to identify subjects whose data is odd in some way. Ensure this is consistent across all_dat1 and all_dat2 and allsum.
+We first rename the current subject variable so we can reuse this name with the numbered ID for all future computations.
+Again, this is rather slow because done in a loop, which is probably not very efficient!
 
@@
 all_dat2 <- all_dat2 %>% rename(origID = subject) 
+all_dat1$subject <- NA
+all_dat2$subject <- NA
 for (n in 1:nrow(allsum)){
-  allsum$subject <- n
+  allsum$subject[n] <- n
   w<-which(all_dat1$origID == allsum$origID[n])
-  all_dat1$subject <-n
+  all_dat1$subject[w] <-n
   w<-which(all_dat2$origID == allsum$origID[n])
-  all_dat2$subject <-n
-}
+  all_dat2$subject[w] <-n
+}
+head(allsum)
 ```
@@
 For outlier removal (usually RT), lowcut is a cutoff value below which data excluded (typically v fast responses that are implausible, e.g. 201 ms for RT)
- and zcut is the constant in Hoaglin-Iglewicz method (usually 2.2)
+ and zcut is the constant in Hoaglin-Iglewicz method (usually 2.2).
+ Here the focus is on just one tail of the distribution (slow RTs) so we use 1.65 as zcut.
 
@@
   for (i in 1:nsub) { # loop through subjects
-  subname <- levels(origdat$subject)[i] # find subject 
+  subname <- origdat$subject[i] # find subject 
   myrows <- which(origdat$subject==subname) # select rows for this subject
@@
 nsub<-length(unique(origdat$subject))
-# create compact data frame 
+# create compact data frame ; one row for each subject, subjects stacked, L above R
 temp_dat <- data.frame(matrix(ncol = 4, nrow = nsub*2))
@@
 for (i in 1:nsub) { # loop through subjects
-  subname <- levels(origdat$subject)[i] # find subject subject
+  subname <- i # find subject 
   myrows <- which(origdat$subject==subname) # select rows for this subject
@@
   RTcol <- which(colnames(origdat) == varlist[4])
-  tmp <- data.frame(origdat[myrows,])
-
-  for (j in 1:2) {
+  tmp <- data.frame(origdat[myrows,]) #all trials from this subject
+
+  for (j in 1:2) { #make row for means for each side for this subject
     myrow <- myrow+1
@@
   }
+  #If there are no selections/trials for one side, then exclude L and R
+  # if(temp_dat$N[myrow==0]||temp_dat$N[(myrow-1)==0]){
+  #   temp_dat$RT[(myrow-1):myrow]<-NA
+  # }
 }
@@
 # plot L & R to get slope
-q <- ggplot(data=origdat, aes(x=side, y=X1, group=subject, color=subject)) +
+q <- ggplot(data=origdat, aes(x=side, y=X1, group=as.factor(subject), color=subject)) +
     geom_line() +
@@
 We will exclude those who are inaccurate when both faces the same (as in preregistration).
-Responses to chimeric faces can be coded as whether they correspond to emotion on L or R.
-Note that if the reported emotion does not match either side, then trial is not counted.
-
-The chunk below also detects outliers, and codes them as 1 if the RT is outside the range specified in Hoaglin-Iglewicz, and as 2 if the RT is < 200.
-Outliertable shows each subject with Ns with different outlier codes.
-Some people have 0-1 trials - these seem to be ones where the task was skipped for some reason - I have checked and they seem to have no data.
-One person has twice the number of trials in one session. We'll deal with that by just taking the first 96 trials.
+Responses to chimeric faces are later coded as whether they correspond to emotion on L or R.
+Note that if the reported emotion does not match either side, then trial will not be counted in LI
 
@@
 # count subjects
-nsub <-length(levels(CF_dat_all$subject))
+nsub <-length(unique(CF_dat_all$subject)) #previously we used the ID code and had to count levels of the factor; now with numbers we just use unique
 print(paste('day',d,task,": subjects: ", nsub))
@@
 CF_dat2 <- CF_dat_all[CF_dat_all$l_hemiface != CF_dat_all$r_hemiface,] # create df for chimeras: remove items where same expression in both hemifaces
-
-CF_dat2 <- detect.outliers(CF_dat2,200,1.65) #use generic outlier detection function. The parameters are lowcut (200 ms cutoff for anticipations) and then the value used for Hoaglin Iglewicz (1.65)
-outliertable<-table(CF_dat2$subject,CF_dat2$outlier)
-#Most have 96 accepted trials in CF_dat2. A few have zero - these seem to be ones that had RTs < 200, and may have been responding rapidly at random.
-#One person has 192 trials - seems to have had whole thing presented twice.
-
 if (d==1)
@@
 }
+```
+
+The chunk below detects outliers, and codes them as 1 if the RT is outside the range specified in Hoaglin-Iglewicz, and as 2 if the RT is < 200.
+Outliertable shows each subject with Ns with different outlier codes.
+Some people have 0-1 trials - these seem to be ones where the task was skipped for some reason - I have checked and they seem to have no data.
+
+```{r CFoutliers}
+CF_dat1 <- detect.outliers(CF_dat1,200,1.65) #use generic outlier detection function. The parameters are lowcut (200 ms cutoff for anticipations) and then the value used for Hoaglin Iglewicz (1.65)
+outliertable1<-table(CF_dat1$subject,CF_dat1$outlier)
+CF_dat2 <- detect.outliers(CF_dat2,200,1.65) #use generic outlier detection function. The parameters are lowcut (200 ms cutoff for anticipations) and then the value used for Hoaglin Iglewicz (1.65)
+outliertable2<-table(CF_dat2$subject,CF_dat2$outlier)
+#Most have 96 accepted trials in CF_dat2. A few have 1-2 - these seem to be ones that had RTs < 200, and may have been responding rapidly at random.
+
 ```
@@
   CF_same <- CF_same1
+  CF_diff <- CF_dat1
   if (d==2){
     CF_same <- CF_same2
+    CF_diff <- CF_dat2
   }
@@
 # calculate participants average
-emotion_mean <- aggregate(FUN= mean, data= CF_same, Correct~ subject)
+  #update: do this for chimeric as well as Same trials
+emot_corr_Same <- aggregate(FUN= mean, data= CF_same, Correct~ subject)
 # add this to allsum 
-allsum <- merge(allsum, emotion_mean, by= "subject")
-}
-
-w<-which(colnames(allsum) %in% c("Emot.recog1","Emot.recog2"))
+allsum <- merge(allsum, emot_corr_Same, by= "subject")
+emot_corr_Diff <- aggregate(FUN= mean, data= CF_diff, Correct~ subject)
+# add this to allsum 
+allsum <- merge(allsum, emot_corr_Diff, by= "subject")
+}
+
+w<-which(colnames(allsum) %in% c("Emot.corr.same1","Emot.corr.diff1","Emot.corr.same2","Emot.corr.diff2"))
 if (length(w)>0){
@@
 
-colnames(allsum)[(mycol-1):mycol]<-c('Emot.recog1','Emot.recog2')
-
-plot(jitter(allsum$Emot.recog1,10),jitter(allsum$Emot.recog2,10),pch=3,cex=.6,xlab='Nonchimeric emotion recog 1',ylab='Nonchimeric emotion recog 2')
+colnames(allsum)[(mycol-3):mycol]<-c("Emot.corr.same1","Emot.corr.diff1","Emot.corr.same2","Emot.corr.diff2")
+
+plot(jitter(allsum$Emot.corr.same1,10),jitter(allsum$Emot.corr.same2,10),pch=3,cex=.6,xlab='Nonchimeric emotion recog 1',ylab='Nonchimeric emotion recog 2')
 abline(h=.75)
@@
 # mark the cases where there is less than 75% for identical stimuli
-w <- union(which(allsum$Emot.recog1 < .75),which(allsum$Emot.recog2 < .75))
+w <- union(which(allsum$Emot.corr.same1 < .75),which(allsum$Emot.corr.same2 < .75))
 allsum$exclude[w]<-1 #code exclude = 1 if excluded because poor emot recog
 print(paste('N excluded because < 75% on at least one trial:',length(w)))
-```
-
-In this section, the correct sequences are marked as 1, incorrect answers are marked as 0. We then mark whether the correct answers were in the left or right hemiface.
+
+# Just for interest look also at % correct on chimeric.
+# Presumably the conflicting information makes this harder so we don't use this for exclusion
+plot(jitter(allsum$Emot.corr.diff1,10),jitter(allsum$Emot.corr.diff2,10),pch=3,cex=.6,xlab='Chimeric emotion recog 1',ylab='Chimeric emotion recog 2',col=(1+allsum$exclude),main='Proportion correct for chimeric (not used for exclusion)')
+abline(h=.75)
+abline(v=.75)
+text(.6,.95,'Red excluded, <75% on same')
+
+```
+
+In this section, we mark whether the correct answers were in the left or right hemiface.
 
@@
   }
-CF_dat <- 
-  CF_dat %>% 
-  mutate(Correct= ifelse(as.character(Response) == as.character(l_hemiface), 1, 
-                         ifelse(as.character(Response) == as.character(r_hemiface), 1, 0)))
+
 # this will mark the side responded to. NB to allow for generic functions for all tasks, we use the label 'side' rather than hemiface
@@
   mutate(side= ifelse(as.character(Response) == as.character(l_hemiface) & Correct == 1, "Left",
-                       ifelse(as.character(Response) == as.character(r_hemiface) & Correct == 1, "Right", NA)))
+                       ifelse(as.character(Response) == as.character(r_hemiface) & Correct == 1, "Right", "X")))
 
 #Added coding of 'X' for responses of emotion not shown in either side
-w<-which(CF_dat$Correct==0)
-CF_dat$side[w] <- 'X'
+
 # code side as factor
@@
 }
+#sanity check: looks fine. X means response was emotion of neither side
+ table(CF_dat1$side)
+ table(CF_dat2$side)
 ```
@@
 varlist <-   c("subject", "side", "accurate", "RT") #need to be in this order, ie sub, side, acc and RR
-origdat <- CF_dat1
 latlist <- c("Left","Right") #names of factor levels
@@
 
-w<-which(is.na(chim_dat1$subject))
-if (length(w)>0) {chim_dat1<-chim_dat1[-w,]}
-w<-which(is.na(chim_dat2$subject))
-if (length(w)>0) {chim_dat2<-chim_dat2[-w,]}
+# w<-which(is.na(chim_dat1$subject))
+# if (length(w)>0) {chim_dat1<-chim_dat1[-w,]}
+# w<-which(is.na(chim_dat2$subject))
+# if (length(w)>0) {chim_dat2<-chim_dat2[-w,]}
 
@@
 ```{r dosummary}
+#First look at accuracy, Day 1
 origdat <- chim_dat1
@@
 dopirate(origdat,task,measure)
+#Same function will do RT next
 origdat$X1 <- chim_dat1$RT
@@
 #Now repeat for day 2
+#-------------------------
+#We need to add a step to exclude accuracy and RT if fewer than 3 trials in one condition. This only affects 2nd session data, so this code is specific to that file
+w<-which(chim_dat2$N<2)
+dropsub <- chim_dat2$subject[w]
+chim_dat2$accurate[chim_dat2$subject %in% dropsub]<-NA
+chim_dat2$RT[chim_dat2$subject %in% dropsub]<-NA
+#-------------------------
 origdat <- chim_dat2
@@
 dopirate(origdat,task,measure)
+
 origdat$X1 <- chim_dat2$RT

### 2020-05-23 battery-preprocessing_v5_DB.Rmd
Started dichotic analysis check.

@@
 NB. Just also putting here relevant references that need to be looked at!
-https://pubmed.ncbi.nlm.nih.gov/18050003/
+Bourne on RT in chimeric faces: https://pubmed.ncbi.nlm.nih.gov/18050003/
 
@@
 library(corrr) #easier correlations!
+library(beeswarm) #just a kind of plot Dorothy likes
 ```
@@
 # we do this for accuracy first
-# reformat data frame
+# reformat data frame so left and right are side by side
 keeps <- c("subject", "side", "X1") #X1 holds either accuracy or RT
@@
 w<-which(colnames(allsum)==mycolname)
-allsum<-allsum[,-w]
+if(length(w)>0){
+allsum<-allsum[,-w]}
 #now add this LI
@@
 
-makeZ <- function(allsums,origdat,task,mycolname,polarity,mycutoff){
+makeZ <- function(allsum,origdat,task,mycolname,polarity,mycutoff){
 # we do this for accuracy first
@@
 Zdf <- spread(Zdf, side, X1)
-# calculate each participants' lat index
+# calculate each participants' z score for laterality 
 Zdf <- 
@@
 # First check if this column already exists and delete it if so
-w<-which(colnames(allsums)==mycolname)
-allsums<-allsums[,-w]
-
-allsums <- merge(allsums, Zdf, by= "subject")
-lastcol<-length(colnames(allsums))
-colnames(allsums)[lastcol]<-mycolname
-
-return(allsums)
+w<-which(colnames(allsum)==mycolname)
+if(length(w)>0){
+allsum<-allsum[,-w]} #delete cols that already exist with that name
+
+allsum <- merge(allsum, Zdf, by= "subject")
+lastcol<-length(colnames(allsum))
+colnames(allsum)[lastcol]<-mycolname
+
+return(allsum)
 }
@@
 First, we judge participants' ability to correctly identify emotions using the "CF_same" dataframe.
-Participants are removed if their performance is below 75% as this indicates poor ability to identify emotions, which will inevitably influence performance on the Chimeric Face Task.
+Participants are flagged if their performance is below 75% as this indicates poor ability to identify emotions, which will inevitably influence performance on the Chimeric Face Task.
 
@@
   # add this to allsum 
+  colnames(emot_corr_Same)[2]<-'Emot_corr_same'
   allsum <- merge(allsum, emot_corr_Same, by= "subject")
@@
   # add this to allsum 
+  colnames(emot_corr_Diff)[2]<-'Emot_corr_diff'
   allsum <- merge(allsum, emot_corr_Diff, by= "subject")
@@
 colnames(allsum)[(mycol-3):mycol]<-c("Emot.corr.same1","Emot.corr.diff1","Emot.corr.same2","Emot.corr.diff2")
-
-plot(jitter(allsum$Emot.corr.same1,10),jitter(allsum$Emot.corr.same2,10),pch=3,cex=.6,xlab='Nonchimeric emotion recog 1',ylab='Nonchimeric emotion recog 2')
-abline(h=.75)
-abline(v=.75)
-text(.6,.8,'drop if <75% on either run',cex=.8)
-
+```
+
+```{r excludelowCFacc}
 #db modified to mark rather than drop cases with scores < .75
@@
 print(paste('N excluded because < 75% on at least one trial:',length(w)))
+
+```
+
+```{r plotCFacc}
+plot(jitter(allsum$Emot.corr.same1,10),jitter(allsum$Emot.corr.same2,10),pch=3,cex=.6,xlab='Nonchimeric emotion recog 1',ylab='Nonchimeric emotion recog 2',col=(1+allsum$exclude))
+abline(h=.75)
+abline(v=.75)
+text(.6,.8,'red: exclude if <75% on either run',cex=.8)
 
@@
 
-Finally, a laterality index is calculated and plotted based on the equation [(R-L)/(R+L)] x 100, where L= left side and R= right side.
+Next, a laterality index is calculated and plotted based on the equation [(R-L)/(R+L)] x 100, where L= left side and R= right side.
 
@@
 
-cor1 <- cor(allsum$Day1_CF_acc_LI,allsum$Day2_CF_acc_LI)
+cor1 <- cor(allsum$Day1_CF_acc_LI,allsum$Day2_CF_acc_LI,use='complete.obs')
 plot(allsum$Day1_CF_acc_LI,allsum$Day2_CF_acc_LI,main='Chimeric LI accuracy',col=(1+allsum$exclude))
@@
 text(-50,85,paste0('N = ',myn))
-cor2<-cor(allsum$Day1_CF_RT_LI,allsum$Day2_CF_RT_LI)
+cor2<-cor(allsum$Day1_CF_RT_LI,allsum$Day2_CF_RT_LI,use='complete.obs')
 plot(allsum$Day1_CF_RT_LI,allsum$Day2_CF_RT_LI,main='Chimeric LI RT',col=(1+allsum$exclude)) #this would plot excluded as red, but they are now excluded at earlier stage so don't show up
@@
 allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)
-w<-which(allsum$excluded == 1)
+w<-which(allsum$exclude == 1)
 allsum$Day1_CF_acc_Z[w] <- NA #removes zscores for excluded cases
@@
 
+#Can also do sanity check: the Chimeric Z accuracy should be very highly correlated with the laterality index, but on a different scale
+
+plot(allsum$Day1_CF_acc_LI,allsum$Day1_CF_acc_Z,main='Day 1: LI vs zlat')
+plot(allsum$Day2_CF_acc_LI,allsum$Day2_CF_acc_Z,main='Da7 2: LI vs zlat')
 ```
@@
 We would not expect a RT advantage for this type of measure, where the response is to a single stimulus with conflicting information from the 2 sides.
+Nevertheless, there is a RT advantage - this is consistent with Bourne:Laterality 2008 Jan;13(1):92-103.
+doi: 10.1080/13576500701754315.
+(just read abstract so far).
+
+## Next Question: Does CF laterality relate to handedness?
+Chunk below indicates there is a significant difference, though plots show much overlap.
+
+```{r CFhanded}
+
+t.test(allsum$Day1_CF_acc_Z~allsum$handpch)
+t.test(allsum$Day2_CF_acc_Z~allsum$handpch)
+beeswarm(Day1_CF_acc_Z~handedness, data = allsum, 
+  pch = 16, cex=.5,col = rainbow(8),
+  main = 'Day 1')
+beeswarm(Day2_CF_acc_Z~handedness, data = allsum, 
+  pch = 16, cex=.5,col = rainbow(8),
+  main = 'Day 2')
+```
+## Next Question: Does CF laterality relate to task accuracy
+Just looks at correlations and plot scattergram.
+
+The scattergram shapes are a bit unusual - not sure if this is just because high levels of accuracy are more common.
+Does suggest high levels of lateralisation in either direction may be more accurate?
+```{r CFlat_acc}
+cor1 <- round(cor(allsum$Day1_CF_acc_Z,allsum$Emot.corr.diff1,use="complete.obs"),3)
+plot(allsum$Day1_CF_acc_Z,allsum$Emot.corr.diff1)
+text(15,.6,paste0('r = ',cor1))
+cor2 <- round(cor(allsum$Day2_CF_acc_Z,allsum$Emot.corr.diff2,use="complete.obs"),3)
+plot(allsum$Day2_CF_acc_Z,allsum$Emot.corr.diff2)
+text(15,.6,paste0('r = ',cor2))
+
+#These are correlations with unsigned z-scores: significant in large sample, but not much going on here
+cor1a <- round(cor(abs(allsum$Day1_CF_acc_Z),allsum$Emot.corr.diff1,use="complete.obs"),3)
+cor2a <- round(cor(abs(allsum$Day2_CF_acc_Z),allsum$Emot.corr.diff1,use="complete.obs"),3)
+```
 
@@
 
-This part of the R Markdown file prcoesses the "Dichotic listening task' implmented online via Gorilla.sc which aims to assess the lateralisation of receptive language lateralisation.
-In the task, participants hear a different CV sound in each ear.
-Participants respond to the sound they heard clearest
-
-This task is near identical to that reported in Karlsson, Johnstone, and Carey (2019) and were original described in Hugdahl, Westerhausen, Alho, Medvedev, Laine, & Hämäläinen (2009).
-Stimuli were kindly provided by Dr Emma Karlsson (https://research.bangor.ac.uk/portal/en/researchers/emma-karlsson(17f6f06c-8aa9-4d16-93ba-4f98f285297e).html) and permission to use these stimuli was given by Professor Kenneth Hughdahl (https://www.uib.no/en/persons/Kenneth.Hugdahl).
+This part of the notebook prcoesses the "Dichotic listening task' implmented online via Gorilla.sc which assesses the lateralisation of receptive language.
+In the task, participants hear a different CV syllable in each ear.
+Participants report the syllable they heard clearest
+
+This task is near identical to that reported in Karlsson, Johnstone, and Carey (2019) and was original described in Hugdahl, Westerhausen, Alho, Medvedev, Laine, & Hämäläinen (2009).
+Stimuli were kindly provided by Dr Emma Karlsson (https://research.bangor.ac.uk/portal/en/researchers/emma-karlsson(17f6f06c-8aa9-4d16-93ba-4f98f285297e).html) and permission to use these stimuli was given by Professor Kenneth Hugdahl (https://www.uib.no/en/persons/Kenneth.Hugdahl).
+
+Much of the processing parallels that of chimeric faces.
 
@@
 disp <- "e_task"
-dichotic_dat2 <- procdata(origfile,wanted,task,disp) #use generic function (see above) to read in relevant columns
-
+dichotic_all <- procdata(origfile,wanted,task,disp) #use generic function (see above) to read in relevant columns
+
+dichotic_all$Correct <- 0 #default is error
+w<-union(which(dichotic_all$Response==dichotic_all$left_channel),which(dichotic_all$Response==dichotic_all$right_channel))
+dichotic_all$Correct[w] <-1
+print(paste0('Day ',d,': total Error (0) and Correct(1)'))
+print(table(dichotic_all$Correct))
 # count subjects
-nsub <-length(levels(dichotic_dat2$subject))
+nsub <-length(unique(dichotic_all$subject))
 print(paste('day',d,task,": subjects: ", nsub))
 
-dichotic_same2 <- dichotic_dat2[dichotic_dat2$left_channel == dichotic_dat2$right_channel,] # create data frame for non-chimeras
-dichotic_dat2 <- dichotic_dat2[dichotic_dat2$left_channel != dichotic_dat2$right_channel,] # remove items where same expression in both hemifaces
-dichotic_dat2 <- detect.outliers(dichotic_dat2,200,1.65) #use generic outlier removal function
-outliertable<-table(dichotic_dat2$subject,dichotic_dat2$outlier)
-#print(outliertable)
-
+dichotic_same2 <- dichotic_all[dichotic_all$left_channel == dichotic_all$right_channel,] # create data frame for trials with both ears same
+dichotic_dat2 <- dichotic_all[dichotic_all$left_channel != dichotic_all$right_channel,] # remove items where same expression in both hemifaces
 if (d==1)
@@
 
-First, we judge participants' ability to correctly identify sounds using the "dichotic_same" dataframe. Participants are marked as excluded if their performance is below 75%.
+Now remove outliers, using same function as for CF.
+```{r dichotic.outliers}
+dichotic_dat1 <- detect.outliers(dichotic_dat1,200,1.65) #use generic outlier removal function
+outliertable1<-table(dichotic_dat1$subject,dichotic_dat1$outlier)
+dichotic_dat2 <- detect.outliers(dichotic_dat2,200,1.65) #use generic outlier removal function
+outliertable2<-table(dichotic_dat2$subject,dichotic_dat2$outlier)
+#print(outliertable)
+
+```
+Inspection of outliertable1 and outliertable2 shows most are data as expected, with 120 trials. Outliers not v numerous. Occasional cases where test does not seem to be done, just one trial recorded.
+
+Now, we judge participants' ability to correctly identify sounds using the "dichotic_same" dataframe. Participants are marked as excluded if their performance is below 75%.
 
 ```{r soundsability, warning=FALSE,message=FALSE}
-# this will mark correct responses as 1 or 0. It will be importantfor removing participants later on. 
+# We consider how many errors there are when both ears have same sound. 
 # NB we loop through same process for day1 and day2
@@
 
-dichotic_same <- 
-  dichotic_same %>% 
-  mutate(sound_recog= ifelse(as.character(Response) == as.character(left_channel), 1, 0))
 # calculate participants average
-sound_mean <- aggregate(FUN= mean, data= dichotic_same, sound_recog~ subject)
+sound_mean <- aggregate(FUN= mean, data= dichotic_same, Correct~ subject)
 # add this to allsum
+
+# Check if sound.recog columns already exist (ie if this chunk was run previously)
+w <- which(colnames(allsum) %in% c('sound.recog1','sound.recog2'))
+if (length(w)>0){
+  allsum<-allsum[,-w]
+}
+
 allsum <- merge(allsum, sound_mean, by= "subject")

### 2020-05-23 battery-preprocessing_v5_DB.Rmd
Got as far as checking LI for dichotic.

@@
 allsum<-allsum[,-w]}
+
 #now add this LI
-allsum <- merge(allsum, Lsubjectf, by= "subject")
+allsum <- merge(allsum, Lsubjectf, by= "subject",all.x=T)
+#Need all.x = T to retain all original subjects from allsum, even if no match
 lastcol<-length(colnames(allsum))
@@
 
-allsum <- merge(allsum, Zdf, by= "subject")
+allsum <- merge(allsum, Zdf, by= "subject",all.x=T)
 lastcol<-length(colnames(allsum))
@@
   colnames(emot_corr_Diff)[2]<-'Emot_corr_diff'
-  allsum <- merge(allsum, emot_corr_Diff, by= "subject")
+  allsum <- merge(allsum, emot_corr_Diff, by= "subject",all.x=T)
 }
@@
 allsum$exclude[w]<-1 #code exclude = 1 if excluded because poor emot recog
-print(paste('N excluded because < 75% on at least one trial:',length(w)))
+print(paste('N flagged for exclusion because < 75% on at least one trial:',length(w)))
 
@@
 
+```
+
+The data is visualised using the pirate plot function and t.test conducted.
+
+```{r dosummary}
 #find included cases
@@
 
-chim_dat1 <- chim_dat1[(chim_dat1$subject %in% inclsubs),]
-chim_dat2 <- chim_dat2[(chim_dat2$subject %in% inclsubs),]
-
-```
-
-The data is visualised using the pirate plot function and t.test conducted.
-
-```{r dosummary}
 #First look at accuracy, Day 1
-origdat <- chim_dat1
+origdat <- chim_dat1[(chim_dat1$subject %in% inclsubs),]
 measure <- 'Accuracy'
-origdat$X1 <- chim_dat1$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc with the dopirate function
+origdat$X1 <- origdat$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc with the dopirate function
 task <- 'Chimeric faces day 1'
@@
 #Same function will do RT next
-origdat$X1 <- chim_dat1$RT
+origdat$X1 <- origdat$RT
 measure <- 'RT'
@@
 #-------------------------
-origdat <- chim_dat2
-origdat$X1 <- chim_dat2$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
+origdat <- chim_dat2[(chim_dat2$subject %in% inclsubs),]
+origdat$X1 <- origdat$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
 task <- 'Chimeric faces day 2'
@@
 
-origdat$X1 <- chim_dat2$RT
+origdat$X1 <- origdat$RT
 measure <- 'RT'
@@
 
-allsum <- merge(allsum, sound_mean, by= "subject")
-}
+  allsum <- merge(allsum, sound_mean, by= "subject",all.x=T)
+}
+
 mycol <- length(colnames(allsum))
@@
   }
-dichotic_dat <- 
-  dichotic_dat %>% 
-  mutate(Correct= ifelse(as.character(Response) == as.character(left_channel), 1, 
-                         ifelse(as.character(Response) == as.character(right_channel), 1, 0)))
+
 # this will mark the side responded to. NB to allow for generic functions for all tasks, we use the label 'side' rather than hemiface
@@
 ```{r dosummarydich}
-origdat <- DL_dat1
+includesubs <- which(allsum$exclude<2) #only plot/t-test included cases
+origdat <- DL_dat1[DL_dat1$subject %in% includesubs,]
 measure <- 'Accuracy'
-origdat$X1 <- DL_dat1$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
+origdat$X1 <- origdat$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
 task <- 'Dichotic listening day 1'
 dopirate(origdat,task,measure)
-origdat$X1 <- DL_dat1$RT
+origdat$X1 <- origdat$RT
 measure <- 'RT'
@@
 #Now repeat for day 2
-origdat <- DL_dat2
-origdat$X1 <- DL_dat2$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
+origdat <-  DL_dat2[DL_dat2$subject %in% includesubs,]
+origdat$X1 <- origdat$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
 task <- 'Dichotic listening day 2'
@@
 dopirate(origdat,task,measure)
-origdat$X1 <- DL_dat2$RT
+origdat$X1 <- origdat$RT
 measure <- 'RT'

### 2020-05-23 battery-preprocessing_v5_DB.Rmd
completed CF and dichotic. Started on RDT.

@@
 
-## Next Question: Does CF laterality relate to handedness?
+## Does CF laterality relate to handedness?
 Chunk below indicates there is a significant difference, though plots show much overlap.
@@
   dichotic_same <- dichotic_same1
+   dichotic_diff<- dichotic_dat1
   if (d==2){
     dichotic_same <- dichotic_same2
+     dichotic_diff <- dichotic_dat2
   }
@@
   sound_mean <- aggregate(FUN= mean, data= dichotic_same, Correct~ subject)
+  dichotic_acc <- aggregate(FUN= mean, data= dichotic_diff, Correct~ subject)
   # add this to allsum
@@
   # Check if sound.recog columns already exist (ie if this chunk was run previously)
-  w <- which(colnames(allsum) %in% c('sound.recog1','sound.recog2'))
+  w <- which(colnames(allsum) %in% c('sound.recog1','sound.recog2','dich_acc1','dich_acc2'))
   if (length(w)>0){
@@
   allsum <- merge(allsum, sound_mean, by= "subject",all.x=T)
+  allsum <- merge(allsum, dichotic_acc, by= "subject",all.x=T)
 }
@@
 mycol <- length(colnames(allsum))
-colnames(allsum)[(mycol-1):mycol]<-c('sound.recog1','sound.recog2')
+colnames(allsum)[(mycol-3):mycol]<-c('sound.recog1','sound.recog2','dichotic.acc1','dichotic.acc2')
 
@@
 ```
+## Does dichotic laterality relate to handedness?
+Chunk below indicates there is a significant difference, but v small effect size.
+
+```{r dichotic_handed}
+partsum <- allsum[allsum$subject %in% includesubs,]
+
+t.test(partsum$Day1_Dich_acc_Z~partsum$handpch)
+t.test(partsum$Day2_Dich_acc_Z~partsum$handpch)
+beeswarm(Day1_Dich_acc_Z~handedness, data = partsum, 
+  pch = 16, cex=.5,col = rainbow(8),
+  main = 'Day 1')
+beeswarm(Day2_Dich_acc_Z~handedness, data = partsum, 
+  pch = 16, cex=.5,col = rainbow(8),
+  main = 'Day 2')
+```
+## Next Question: Does Dichotic laterality relate to task accuracy
+Just looks at correlations and plot scattergram.
+No hint of any relationship.
+
+```{r Cdichotic_acc}
+cor1 <- round(cor(partsum$Day1_Dich_acc_Z,partsum$dichotic.acc1,use="complete.obs"),3)
+plot(partsum$Day1_Dich_acc_Z,partsum$dichotic.acc1)
+text(15,.6,paste0('r = ',cor1))
+cor2 <- round(cor(partsum$Day2_Dich_acc_Z,partsum$dichotic.acc2,use="complete.obs"),3)
+plot(partsum$Day2_Dich_acc_Z,partsum$dichotic.acc2)
+text(15,.6,paste0('r = ',cor2))
+
+```
 
@@
 
-This portion of the R Markdown file prcoesses the "Rhyme Detection Task' implmented online via Gorilla.sc which aims to assess the lateralisation of language processing.
-In the task, participants are shown a word in the central visual field and different images in the left and right visual field.
-An arrow appears pointing to either image.
-Participants then respond by indicating whether the image and word rhyme.
-
-This reads in a csv file where each row represents a response to a trial.
-Participants, identified by random letter strings, are stacked on top of each other.
+This portion of the notebook prcoesses the "Rhyme Detection Task' implemented online via Gorilla.sc, which aims to assess the lateralisation of language processing.
+This is a new task devised for this project.
+In the task, participants are shown a word in the central visual field and two images in the left and right visual field.
+The task is to judge which of the two images has a name that rhymes with the central word.
+
  We start by reading the data and cutting out unwanted information.
@@
   levels(RDT_dat2$side)<-c("Left","Right")
-RDT_dat2<-detect.outliers(RDT_dat2,200,1.65) #numbers specify min RT and zscore for Hoaglin-Iglewicz respectively
 
   # count subjects
-nsub <-length(levels(RDT_dat2$subject))
-print(paste0("Day 1 subjects: ", nsub))
-outliertable<-table(RDT_dat2$subject,RDT_dat2$outlier)
-#print(outliertable)
-
-# now remove the cases where there are too many or to few cases
-#remove.outlier <- as.data.frame(outliertable)
-#remove.outlier <- remove.outlier[remove.outlier$Freq >= 260 & remove.outlier$Freq <= 208,]
-#remove.p <- list(unique(remove.outlier$Var1))
-#allsum <- allsum[allsum$subject != remove.p,]
-
+nsub <-length(unique(RDT_dat2$subject))
+print(paste0("Day ",d," subjects: ", nsub))
 if(d==1)
   {RDT_dat1 <- RDT_dat2}
-
-}
-```
+}
+```
+Now use generic function to remove outliers.
+
+```{r rdt_outliers}
+RDT_dat1<-detect.outliers(RDT_dat1,200,1.65) #numbers specify min RT and 
+RDT_dat2<-detect.outliers(RDT_dat2,200,1.65) #numbers specify min RT and zscore for Hoaglin-Iglewicz respectively
+
+outliertable1<-table(RDT_dat1$subject,RDT_dat1$outlier)
+outliertable2<-table(RDT_dat2$subject,RDT_dat2$outlier)
+
+```
+
+Scrutiny of outliertable indicates a few problematic cases with far fewer than 260 trials in total. Seems related to glitches in Gorilla.
+Will need to exclude these
 
@@
 
-```{r makedataframeRDT, warning=F}
-#We create files for day1 and day 2
-# NB warnings created because of different Ns for time 1 and time 2, which means there are blanks in subjects
-# We deal with this below by removing rows with no subject code
-varlist <-   c("subject", "side", "accurate", "RT") #need to be in this order, ie sub, side, acc and RR - we will use these in the make.df function
-#We first need to remove rows where RT is coded NA - these seem cases where false recording led to excess trials with no RT
-w<-which(is.na(RDT_dat1$RT))
-RDT_dat1<-RDT_dat1[-w,]
-origdat <- RDT_dat1
-
+```{r makedf_rhyme}
+#We create chim_dat files for day1 and day 2
+varlist <-   c("subject", "side", "accurate", "RT") #need to be in this order, ie sub, side, acc and RR
 latlist <- c("Left","Right") #names of factor levels
-RDT_df1 <- make.df(origdat,varlist,latlist)
-w<-which(is.na(RDT_dat2$RT))
-RDT_dat1<-RDT_dat2[-w,]
-origdat <- RDT_dat2
-RDT_df2 <- make.df(origdat,varlist,latlist)
-
-# Remove blank rows 
-w<-which(is.na(RDT_df1$subject))
-if (length(w)>0) {RDT_df1<-RDT_df1[-w,]}
-w<-which(is.na(RDT_df2$subject))
-if (length(w)>0) {RDT_df2<-RDT_df2[-w,]}
-
-# Exclude cases with low % correct
-# NB something weird with these - seem to be ones with too many trials, suggesting some problem with multiple responses
-w<-which(RDT_df1$p.corr<50)
-if (length(w)>0)
-{RDT_df1$accurate[w]<-NA
-RDT_df1$RT[w]<-NA
-}
-w<-which(RDT_df2$p.corr<50)
-if (length(w)>=0)
-{RDT_df2$accurate[w]<-NA
-RDT_df2$RT[w]<-NA
-}
-
-# now remove the cases where there are NA for only one visual field
-# this will remove those who responded by pressing only one button.
-RDT_df1 <- na.omit(RDT_df1) %>% 
-  group_by(subject) %>%
-  mutate(no_rows = length(subject))
-RDT_df1 <- RDT_df1[RDT_df1$no_rows == 2,]
-
-RDT_df2 <- na.omit(RDT_df2) %>% 
-  group_by(subject) %>%
-  mutate(no_rows = length(subject))
-RDT_df2 <- RDT_df2[RDT_df2$no_rows == 2,]
-```
-
-```{r checker, warning=FALSE}
+
+RDT_df1 <- make.df(RDT_dat1,varlist,latlist)
+RDT_df2 <- make.df(RDT_dat2,varlist,latlist)
+
+```
+
+Next exclude those with less than 75% correct responses on either side
+
+```{r RDT_exclude}
+w1<-which(RDT_df1$p.corr < 75)
+w2<-which(RDT_df2$p.corr < 75)
+exRDT<-unique(c(RDT_df1$subject[w1],RDT_df2$subject[w2]))
+
+allsum$exRDT <- 0
+allsum$exRDT[allsum$subject %in% exRDT]<-1 #decided best to have separate column for exclusion for each task
+w<-which(allsum$exRDT==0)  
+incRDT <- allsum$subject[w]
+
+```
+
+Comparison shows that there is significant difference in RT.
+Accuracy is close to ceiling and does not show side difference (but it would be a bit odd for this task, - unlike CF and dichotic, there is no competing information and only one answer is correct).
+```{r rdtpirate}
 for (d in 1:2){
-origdat <-RDT_df1
+origdat <-RDT_df1[RDT_df1$subject %in% incRDT,]
 if (d==2)
-{origdat <- RDT_df2}
+{origdat <-RDT_df2[RDT_df2$subject %in% incRDT,]}
 measure <- 'Accuracy'
-origdat$X1 <- origdat$accurate
+origdat$X1 <- origdat$p.corr
 task <- 'Rhyme Detection'
@@
 dopirate(origdat,task,measure)
-
+}
+```
+
+```{r RDT.LI}
 #now create LI
+for (d in 1:2){
+origdat <-RDT_df1[RDT_df1$subject %in% incRDT,]
+if (d==2)
+{origdat <-RDT_df2[RDT_df2$subject %in% incRDT,]}
 origdat$X1 <- origdat$p.corr
@@
 abline(h=0)
-text(-3,3,paste0('r = ',round(cor2,3)))
+text(-3,10,paste0('r = ',round(cor2,3)))
 ```

### 2020-05-24 battery-preprocessing_v5_DB.Rmd
started looking at item-level aspects of RDT task. Reading in the words and images for each item to identify specific item combinations

@@
 
-  c<-which(names(origfile) %in% wanted) #find colnumbers of wanted
+  c<-which(names(nufile) %in% wanted) #find colnumbers of wanted
   #NB beware that the columns may not be in the same order as in the wanted list
@@
 
- We start by reading the data and cutting out unwanted information.
+ We start by reading the data and cutting out unwanted information. Here we are including item-specific information as it may be used for later analysis.
 
@@
   # new data
-  wanted <- c("Trial","subject", "Correct", "RT", "ANSWER","Response") # select wanted columns
+  wanted <- c("Trial","subject", "Correct", "RT", "ANSWER","Response","word","Left_image","Right_image") # select wanted columns
   RDT_dat2 <- procdata(origfile,wanted,task,disp) #Here we reuse the procdata function that we developed for processing Chimeric faces data - just does some generic processing
@@
   # do some wrangling on the ANSWER variable to make it compatible with functions
-  RDT_dat2<-filter(RDT_dat2,ANSWER %in% c('left','right'))
+  RDT_dat2<-filter(RDT_dat2,ANSWER %in% c('left','right','no rhyme'))
   c <- which(colnames(RDT_dat2)=="ANSWER")
@@
   RDT_dat2$side<-as.factor(RDT_dat2$side)
-  levels(RDT_dat2$side)<-c("Left","Right")
+  levels(RDT_dat2$side)<-c("left","no rhyme","right")
+
+  #get rid of .png extensions
+  RDT_dat2$Left_image<-tools::file_path_sans_ext(RDT_dat2$Left_image)
+   RDT_dat2$Right_image<-tools::file_path_sans_ext(RDT_dat2$Right_image)
 
@@
 #t.test to test if laterality index differs from zero
-t.test(allsum$RDT_p.corr_LI)
-t.test(allsum$RDT_RT_LI)
+partsum <- allsum[allsum$exRDT==0,]
+print('Testing for difference of single mean from zero')
+t.test(partsum$Day1.RDT_p.corr_LI)
+t.test(partsum$Day1.RDT_RT_LI)
+t.test(partsum$Day2.RDT_p.corr_LI)
+t.test(partsum$Day2.RDT_RT_LI)
 #rename column to identify day
@@
 }
-cor1<-cor(allsum$Day1.RDT_p.corr_LI,allsum$Day2.RDT_p.corr_LI,use='complete.obs')
-cor2<-cor(allsum$Day1.RDT_RT_LI,allsum$Day2.RDT_RT_LI,use='complete.obs')
-plot(allsum$Day1.RDT_p.corr_LI,allsum$Day2.RDT_p.corr_LI,main='Rhyme detect: LI for accuracy (%correct)')
+cor1<-cor(partsum$Day1.RDT_p.corr_LI,partsum$Day2.RDT_p.corr_LI,use='complete.obs')
+cor2<-cor(partsum$Day1.RDT_RT_LI,partsum$Day2.RDT_RT_LI,use='complete.obs')
+plot(partsum$Day1.RDT_p.corr_LI,partsum$Day2.RDT_p.corr_LI,main='Rhyme detect: LI for accuracy (%correct)')
 abline(v=0)
@@
 text(-3,3,paste0('r = ',round(cor1,3)))
-plot(allsum$Day1.RDT_RT_LI,allsum$Day2.RDT_RT_LI,main='Rhyme detect: LI for RT')
+plot(partsum$Day1.RDT_RT_LI,partsum$Day2.RDT_RT_LI,main='Rhyme detect: LI for RT')
 abline(v=0)
@@
 ### T-test as alternative to LI for RDT
+
+The LI results are not encouraging - reliability well below acceptable level.
+Previously I also did a t-test that was analogous to the z-test for CF/dichotic, i.e.
+showed if the individual had a significant RT difference between L and R stimuli.
+But that is v highly correlated with LI.
+The only way to make it more sensitive would be to try a pairwise t-test that compares RTs for the same items presented on L and R.
+So the next bit of code is designed to compute that.
 

### 2020-05-24 battery-preprocessing_v5_DB.Rmd
Full analysis of RDT task. It is looking bad on all fronts - not very lateralised, not very reliable, no relation to handedness. Pairwise analysis of laterality by items looks even worse. Medians rather than mean RT checked; doesn't improve things.

@@
 ```{r maketaskdataframe, warning=FALSE}
-make.df <- function(origdat,varlist,latlist){
+make.df <- function(origdat,varlist,latlist,nloop){
 nsub<-length(unique(origdat$subject))
 # create compact data frame ; one row for each subject, subjects stacked, L above R
-temp_dat <- data.frame(matrix(ncol = 4, nrow = nsub*2))
+temp_dat <- data.frame(matrix(ncol = 4, nrow = nsub*nloop))
 colnames(temp_dat) <- varlist
@@
 
-  for (j in 1:2) { #make row for means for each side for this subject
+  for (j in 1:nloop) { #make row for means for each side for this subject
     myrow <- myrow+1
@@
     temp_dat$RT[myrow] <- mean(as.numeric(tmp2$RT),na.rm=TRUE) # add row for mean RT
+    #NB for RDT task, also looked at medians, but it did not make much difference, presumably because v slow responses already omitted via outlier routine
     temp_dat$N[myrow] <- length(tmp1$RT) # add count of rows for this subject
@@
 latlist <- c("Left","Right") #names of factor levels
-
-chim_dat1 <- make.df(CF_dat1,varlist,latlist)
-chim_dat2 <- make.df(CF_dat2,varlist,latlist)
+nloop=2 #same as levels of latlist
+chim_dat1 <- make.df(CF_dat1,varlist,latlist,nloop)
+chim_dat2 <- make.df(CF_dat2,varlist,latlist,nloop)
 
@@
 latlist <- c("Left","Right") #names of factor levels
-
-DL_dat1 <- make.df(dichotic_dat1,varlist,latlist)
-DL_dat2 <- make.df(dichotic_dat2,varlist,latlist)
+nloop=2 #same as levels of latlist
+DL_dat1 <- make.df(dichotic_dat1,varlist,latlist,nloop)
+DL_dat2 <- make.df(dichotic_dat2,varlist,latlist,nloop)
 
@@
 
+   RDT_dat2$picpair <- paste0(RDT_dat2$Left_image,'_',RDT_dat2$Right_image)
+   w<-which(RDT_dat2$side=='right')
+   RDT_dat2$picpair[w]<- paste0(RDT_dat2$Right_image[w],'_',RDT_dat2$Left_image[w])
+
+   RDT_dat2$item <- paste0(RDT_dat2$word,'_',RDT_dat2$picpair) 
+   #Check all OK
+   t<-table(RDT_dat2$item,RDT_dat2$side)
+   t
+   write.csv(t,'items.csv')
+
+   #Reveals one item miscoded, so we will correct it here.
+   w<-which(RDT_dat2$item=='bite_Door_Kite')
+   temp<-RDT_dat2[w,]
+   #This shows that 'no response' was treated as correct, when it should have been 'right'
+
+   w2 <- intersect(w,which(RDT_dat2$Response=='right'))
+   RDT_dat2$Correct[w2]<-1
+   RDT_dat2$item[w2]<-'bite_Kite_Door'
+   RDT_dat2$side[w2]<-'right'
+
+   #Also reveals occasional trials where Response shows 'Could not set screen ...' These were coded as errors, so need removing
+   s <- "Could not set image to requested size.  Either the participants screen or the containing zone is too small."
+   w<-which (RDT_dat2$Response==s)
+   RDT_dat2<-RDT_dat2[-w,]
+
   # count subjects
@@
 varlist <-   c("subject", "side", "accurate", "RT") #need to be in this order, ie sub, side, acc and RR
-latlist <- c("Left","Right") #names of factor levels
-
-RDT_df1 <- make.df(RDT_dat1,varlist,latlist)
-RDT_df2 <- make.df(RDT_dat2,varlist,latlist)
-
-```
-
-Next exclude those with less than 75% correct responses on either side
+latlist <- c("left","no rhyme","right") #names of factor levels
+nloop<-length(latlist)
+RDT_df1 <- make.df(RDT_dat1,varlist,latlist,nloop)
+RDT_df2 <- make.df(RDT_dat2,varlist,latlist,nloop)
+
+#Scrutiny of these reveal a few cases who missed all 'no rhyme' cases.
+#Qu of whether to remove these: think so - this was included to ensure thoughtful responding.
+```
+
+Next exclude those who never get 'no rhyme' items correct, and those with less than 75% correct responses on either side
 
 ```{r RDT_exclude}
+wn <-unique(which(RDT_df1$accurate ==0),which(RDT_df2$accurate ==0)) #this catches those who were always wrong on 'no rhyme' trials, and possibly others where presentation messed up for technical reasons
+
 w1<-which(RDT_df1$p.corr < 75)
 w2<-which(RDT_df2$p.corr < 75)
-exRDT<-unique(c(RDT_df1$subject[w1],RDT_df2$subject[w2]))
+
+exRDT<-unique(c(RDT_df1$subject[w1],RDT_df2$subject[w2],RDT_df2$subject[wn]))
 
@@
 incRDT <- allsum$subject[w]
-
+ninc <- length(incRDT)
+print(paste0('RDT: number remaining after exclusions = ',ninc))
 ```
@@
 {origdat <-RDT_df2[RDT_df2$subject %in% incRDT,]}
+origdat <-origdat[origdat$side %in% c('left','right'),]
 measure <- 'Accuracy'
@@
 origdat$X1 <- origdat$p.corr
+origdat$side <- as.factor(origdat$side)
+levels(origdat$side) <-c('Left','No rhyme','Right') #function assumes capitalised
 mycolname<-"RDT_p.corr_LI"
@@
 abline(h=0)
-text(-3,10,paste0('r = ',round(cor2,3)))
-```
-
+text(-3,9,paste0('r = ',round(cor2,3)))
+```
+## Does RDT relate to handedness?
+Answer is convincing nope!
+
+```{r RDThand}
+t.test(Day1.RDT_RT_LI~handpch,data=partsum)
+t.test(Day2.RDT_RT_LI~handpch,data=partsum)
+```
 ### T-test as alternative to LI for RDT
 
-The LI results are not encouraging - reliability well below acceptable level.
+The LI results are not encouraging - reliability well below acceptable level, and although RT shows lateralised bias, it is tiny. Also, many subjects dropped.
 Previously I also did a t-test that was analogous to the z-test for CF/dichotic, i.e.
@@
 
-```{r do.ttest}
+```{r RDTaggregate}
+#remove 'no rhyme' trials
 for (d in 1:2){
-  myfile<-RDT_dat1
+  thisdat<-RDT_dat2[RDT_dat2$side %in% c('left','right'),]
+  if(d==1){
+    thisdat<-RDT_dat1[RDT_dat1$side %in% c('left','right'),]
+  }
+
+  thisdat <- thisdat[thisdat$outlier==0,] #remove outliers
+  aggdata <- aggregate(thisdat$RT,by=list(thisdat$side,thisdat$item,thisdat$subject),
+                       FUN=mean,na.rm=TRUE)
+  colnames(aggdata)<-c('side','item','subject','RT')
+  mysubs<-unique(aggdata$subject)
+  if(d==1){
+  pairedt <- data.frame(matrix(NA,nrow=length(mysubs),ncol=5))
+  colnames(pairedt) <- c('subject','t1','p1','t2','p2')
+  }
+  nsubs<-length(mysubs)
+  for (i in 1:nsubs){
+    myt<-NA
+    myp<-NA
+    thisdat <- aggdata[aggdata$subject==mysubs[i],]
+    pairedt$subject[i]<- mysubs[i]
+    #check there are data for both left and right on all items
+    checkdat <-table(thisdat$item)
+    w<-which(checkdat<2)
+    if(length(w)==0){
+      t<-t.test(thisdat$RT[thisdat$side=='left'],thisdat$RT[thisdat$side=='right'],paired=T)
+      myt<-t$statistic
+      myp<-t$p.value
+    }
+    if(d==1){
+      pairedt$t1[i] <- myt
+      pairedt$p1[i] <-myp
+    }
     if(d==2){
-    myfile<-RDT_dat2}
-
-  nsubs <- nrow(allsum)
-  allsum$tlat_RDT<-NA
-   allsum$plat_RDT<-NA
-  for (i in 1:nsubs) {
-    w<-which(myfile$subject==allsum$subject[i])
-    if(length(w)>0)
-    {
-      myt<-t.test(myfile$RT[w]~myfile$side[w])
-      allsum$tlat_RDT[i]<-myt$statistic
-      allsum$plat_RDT[i]<-myt$p.value
-    }
-  }
-   if (d==1){
-     maxcol <-length(colnames(allsum))
-     colnames(allsum)[(maxcol-1):maxcol]<-c('tlat_RDT_1','plat_RDT_1')
-   }
-   if (d==2){
-     maxcol <-length(colnames(allsum))
-     colnames(allsum)[(maxcol-1):maxcol]<-c('tlat_RDT_2','plat_RDT_2')
-   }
-}
-colcount<-length(colnames(allsum))
-# Remove values for cases that are excluded for low acc: do this for both times by just blitzing last 4 columns of allsum
-w<-which(is.na(allsum$Day1.RDT_RT_LI))
-allsum[w,(colcount-3):colcount]<-NA
-w<-which(is.na(allsum$Day2.RDT_RT_LI))
-allsum[w,(colcount-3):colcount]<-NA
-
-# compare t-value with LI : use t = 1.65 as cutoff for sig lateralised
- plot(allsum$tlat_RDT_1,allsum$Day1.RDT_RT_LI)
- abline(v=1.65)
-  plot(allsum$tlat_RDT_2,allsum$Day2.RDT_RT_LI)
-   abline(v=1.65)
-# compare t-value from day 1 and day 2 
-     myr <- cor(allsum$tlat_RDT_1,allsum$tlat_RDT_2,use='complete.obs')
-    plot(allsum$tlat_RDT_1,allsum$tlat_RDT_2)
-    abline(v=0)
-    abline(h=0)
-   abline(v=1.65) 
-   abline(h=1.65)
-      abline(v=-1.65) 
-   abline(h=-1.65)
-   text(-3,4,paste0('r = ',round(myr,3)))
-```
+      pairedt$t2[i] <-myt
+      pairedt$p2[i] <-myp
+    }
+  }
+}
+cor(pairedt$t1,pairedt$t2,use='complete.obs')
+```
+
+So the paired items analysis gives a totally nonsignificant correlation between time 1 and time 2. Alas, the RDT task has to be ditched!
 

### 2020-05-24 battery-preprocessing_v5_DB.Rmd
Finalised RDT analysis - confirms this is no good! But finger tapping is much better than we thought: I had confused the labels of columns for this, so L/R was mixed up with day1/day2. Now it is fixed, this gives a nice laterality index which is strongly related to handedness and which has reliability just above .7

@@
 The task is to judge which of the two images has a name that rhymes with the central word.
+The logic of the task is that to perform one must generate the phonology of a pictured word's name, which is a classic left-hemisphere task - it is assumed this will be easier for pictures the are projected to the left hemisphere (ie RVF).
+A small proportion of trials are 'no rhyme' cases, which are designed to assess attention to the task.
 
@@
 levels(origdat$side) <-c('Left','No rhyme','Right') #function assumes capitalised
-mycolname<-"RDT_p.corr_LI"
+mycolname<-"Day1.RDT_p.corr_LI"
+if(d==2){mycolname<-"Day2.RDT_p.corr_LI"}
 polarity<- 1 #so LI is positive in predicted direction
@@
 origdat$X1 <- origdat$RT
-mycolname<-"RDT_RT_LI"
+mycolname<-"Day1.RDT_RT_LI"
+if(d==2){mycolname<-"Day2.RDT_RT_LI"}
 polarity <- (-1)
 allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
-
-#Cutoff has removed those with low accuracy, but we should also remove their RT LIs
-#We do this by finding those with NA in accuracy LI
-w<-which(is.na(allsum$RDT_acc_LI))
-if(length(w)>0)
-{allsum$RDT_RT_LI[w]<-NA}
+}
+
 #t.test to test if laterality index differs from zero
@@
 t.test(partsum$Day2.RDT_RT_LI)
-#rename column to identify day
-myncol <- length(colnames(allsum))
-colnames(allsum)[(myncol-1):myncol] <- paste0('Day',d,'.',colnames(allsum)[(myncol-1):myncol])
-
-}
+
 cor1<-cor(partsum$Day1.RDT_p.corr_LI,partsum$Day2.RDT_p.corr_LI,use='complete.obs')
@@
 
-So the paired items analysis gives a totally nonsignificant correlation between time 1 and time 2. Alas, the RDT task has to be ditched!
+So the paired items analysis gives a totally nonsignificant correlation between time 1 and time 2. This is much worse than the overall LI reliability.
+
+The failure of this task led to further thoughts about why it is barely lateralised, given that it does implicate language-specific word generation. 
+A noteworthy feature of the other two tasks (CF and dichotic) is that they measure side preference rather than accuracy - i.e.
+a response from either L or R is correct, but two potential responses are put in competition.
+That contrasts with this task where only one side has the correct answer.
+We used pictures because with written words there is a concern that  preference may be confounded by the fact we read words from L to R.
+That could potentially dilute the RVF advantage that is expected from L hem language.
 
@@
 
-This portion of the R Markdown file prcoesses the "Finger Tapping Task' implmented in Gorilla.sc which aims to quantify handedness.
-In the task, participants are required to press a sequence of buttons twice for each hand (Left: W, R, V, X; Right: T, U, M, B) as many times as possible within a 30 ms.
-There is a version that requires more precise movements.
-
-A score of 1 is given each time that a participants completes the sequence.
-
-This reads in a csv file where each row represents a button press.
-Participants, identified by random letter strings, are stacked on top of each other.
-We start by reading the data and cutting out unwanted information.
+This portion of the notebook processes the "Finger Tapping Task' implmented in Gorilla.sc, which aims to provide a quantified measure of hand skill.
+In the task, participants are required to press a sequence of buttons for each hand (Left: W, R, V, X; Right: T, U, M, B) as many times as possible within 30 ms.
+We have also devised is a version that requires more complex movements, but early data suggested that does not achieve our aim of emphasising lateral differences.
+
+A score of 1 is given for each sequence completed within the time interval.
+
+In the raw data file, farticipants, identified by random letter strings, are stacked on top of each other. We start by reading the data and cutting out unwanted information.
 
@@
 ```{r readdataFinger, warning=FALSE}
-
-finger <- read.csv("day1_combined.csv", na.strings = "NA")
-
+for (d in 1:2){
+finger <- all_dat2[all_dat2$Task.Name=='Finger Tapping',]
+if(d==1){
+finger <- all_dat1[all_dat1$Task.Name=='Finger Tapping',]}
 # cleaning
@@
 finger <- finger[finger$display == "right hand" | finger$display == "left hand",] # only for task
-# write as factors
-finger$display <- as.factor(finger$display) 
-finger$subject <- as.factor(finger$Participant.Public.ID)
-finger$RT <- as.numeric(as.character(finger$Reaction.Time))
 
@@
 
-all_finger1<- finger
-
-finger <- read.csv("day2_combined.csv", na.strings = "NA")
-
-# cleaning
-finger <- finger[finger$Attempt >= 1,] # only answers 
-finger <- finger[finger$display == "right hand" | finger$display == "left hand",] # only for task
-# write as factors
-finger$display <- as.factor(finger$display) 
-finger$subject <- as.factor(finger$Participant.Public.ID)
-finger$RT <- as.numeric(as.character(finger$Reaction.Time))
-
-# new data
-wanted <- c("subject", "display", "Response", "RT") # select wanted columns
-c<-which(names(finger) %in% wanted) #find colnumbers of unwanted
-finger <-finger[,c] #remove unwanted columns
-finger <- na.omit(finger) # remove NAs
-
 all_finger2<- finger
 
+if(d==1){all_finger1 <- finger}
+}
 ```
@@
 
-subids <- levels(finger1$subject) #just look at data for those doing time1 and
+subids <- unique(finger1$subject) #just look at data for those doing time1 and
 nsub <- length(subids)
@@
   finger.df$subject[n] <- subids[n]
-  myc <- 1 #will be incremented so sums written to correct columns
+  myc <- 1 #column index: will be incremented so sums written to correct columns
+  mycols<-matrix(c(2,4,
+                 3,5),nrow=2,byrow=T)
+  #makes a matrix for looking up column to write to : rows are days, cols are R then L
+
   for (d in 1:2){ #each day data
@@
       myfinger<- finger2}
-      tempfinger <- filter(myfinger,subject==subids[n],Response %in% c('t', 'u', 'm', 'b', 'w', 'r', 'v', 'x')) 
-      #exclude rows with other stuff in Response
-      #(this is not essential as nonsequences will just get ignored)
+    tempfinger <- filter(myfinger,subject==subids[n],
+                         Response %in% c('t', 'u', 'm', 'b', 'w', 'r', 'v', 'x'))  
+    #all rows with these keypresses for this subject
 
@@
     for (s in 1:2){
-        myc<-myc+1 #increment column counter
-        target <- myseq[s]
+      myc<-mycols[d,s] #select column
+      target <- myseq[s] #tumb then wrvx, ie right then left
       finger.df[n,myc]<-str_count(wholeseq,target) 
@@
     }
+    #We then add the SD of the RT - this indicates variability in pace of keypressing
     righth<-filter(tempfinger,display=='right hand')
     lefth<-filter(tempfinger,display=='left hand')
-      startcol<-6
-      if(d==2){startcol<-8}
-      finger.df[n,startcol]<-sd(righth$RTa,na.rm=T)
-      finger.df[n,(startcol+1)]<-sd(lefth$RTa,na.rm=T)
+    myc<-mycols[d,1]+4
+
+    finger.df[n,myc]<-sd(righth$RTa,na.rm=T)
+    finger.df[n,(myc+2)]<-sd(lefth$RTa,na.rm=T)
   }
@@
 ```{r domakeZ_finger}
-#We add Chimeric faces LIs to allsum for day 1 and then day2
+
 task<-'Finger'
@@
 mycolname<-"Day2_Finger_acc_Z"
-polarity<- (-1)#so LI is positive in predicted direction
+polarity<- (1)#so LI is positive in predicted direction
 allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)
 
-cor1 <- cor(allsum$Day1_Finger_acc_Z,allsum$Day2_Finger_acc_Z)
-plot(allsum$Day1_Finger_acc_Z,allsum$Day2_Finger_acc_Z,main='Finger Z accuracy')
+cor1 <- cor(allsum$Day1_Finger_acc_Z,allsum$Day2_Finger_acc_Z,use='complete.obs')
+plot(allsum$Day1_Finger_acc_Z,allsum$Day2_Finger_acc_Z,main='Finger Z accuracy',col=allsum$handpch,pch=15)
 abline(v=0)
 abline(h=0)
-text(-2.5,2.5,paste0('r = ',round(cor1,3)))
+text(-2.5,4,paste0('r = ',round(cor1,3)))
 
@@
 ```{r corr_between_tasks}
-mycols <- c('Day1_CF_acc_Z','Day2_CF_acc_Z','Day1_Dich_acc_Z','Day2_Dich_acc_Z','tlat_RDT_1','tlat_RDT_2','Day1_Finger_acc_Z','Day2_Finger_acc_Z')
+mycols <- c('Day1_CF_acc_Z','Day2_CF_acc_Z','Day1_Dich_acc_Z','Day2_Dich_acc_Z','Day1_Finger_acc_Z','Day2_Finger_acc_Z')
 colnums<- which(colnames(allsum) %in% mycols)

### 2020-05-25 battery-preprocessing_v5_DB.Rmd
A bit of tidying up of code for final plots etc.

@@
     allsum$subject[myrow] <- subname
-    allsum$age[myrow] <- tmp[tmp$Question.Key == "Age",]$Response
+    allsum$age[myrow] <- as.numeric( tmp[tmp$Question.Key == "Age",]$Response)
     allsum$gender[myrow] <- tmp[tmp$Question.Key == "Gender",]$Response
@@
 
+# Previously found one person aged over 200, who should be 21
+# Correct that here
+w <- which (allsum$age>200)
+allsum$age[w]<-21
 #Check ages
@@
 allsum$handpch[allsum$handedness=='Left'] <-2
+
+#numerical code for gender - as we have only one 'prefer not to say' we will assign to NA
+
+allsum$genderpch <-2
+w<-which(allsum$gender=='Male')
+allsum$genderpch[w] <-1
+w<-which(allsum$gender=='Prefer not to say')
+allsum$genderpch[w]<-NA
+
 ```
@@
 Here we report the descriptives for the participants who had full datasets. 
+(Some of these tests etc have been incorporated in analyses above, so can be pruned from script)
 
@@
     summarize(avg = mean(lexTALE), n = n(), sd = sd(lexTALE))
-
+```
+
+```{r finalplots}
 # Now that we have the basic scores, let's look at how this looks with regards to handedness
 # first, describe age for the whole sample and plot
-ggplot(allsum, aes(x = handedness, y = age)) +
-  geom_pirate(aes(colour = handedness), bars = FALSE) + theme_classic() + 
-  ylim(10, 60) + ggtitle("Age by handedness") + ylab("Age (years)") + xlab("")
-# second, show gender count for the whole population
+plot(data=allsum, index_EHI ~ age,col=handpch,xlim=c(18,50),pch=(14+allsum$genderpch))
+
+allsum$handfac <- as.factor(allsum$handedness)
+allsum$genderfac <- as.factor(allsum$genderpch)
+levels(allsum$genderfac)<-c('male','female')
+
 a <- allsum %>%
-  group_by(gender, handedness) %>%
+  group_by(genderfac, handfac) %>%
     summarize(n = n())
-ggplot(data=a, aes(x=handedness, y=n, fill= handedness)) +
+
+ggplot(data=a, aes(x=handfac, y=n, fill= handfac)) +
   geom_bar(stat="identity")+
   geom_text(aes(label=n), vjust=1.6, color="white", size=3.5)+
-  theme_classic() + facet_wrap(.~ gender) + ggtitle("Gender by handedness") + 
-  ylab("Count") + xlab("")
+  theme_classic() + facet_wrap(.~ genderfac) + ggtitle("Gender by handedness") +   ylab("Count") + xlab("")
+
 # third, show footedness
 a <- allsum %>%
-  group_by(handedness, footedness) %>%
+  group_by(handfac, footedness) %>%
     summarize(n = n())
-ggplot(data=a, aes(x=handedness, y=n, fill= handedness)) +
+ggplot(data=a, aes(x=handfac, y=n, fill= handfac)) +
   geom_bar(stat="identity")+
@@
   ylab("Count") + xlab("")
+
 # fourth, show bilingualism
 a <- allsum %>%
-  group_by(bilingual, handedness) %>%
+  group_by(bilingual,handfac) %>%
     summarize(n = n())
-ggplot(data=a, aes(x=handedness, y=n, fill= handedness)) +
-  geom_bar(stat="identity")+
-  geom_text(aes(label=n), vjust=1.6, color="white", size=3.5)+
-  theme_classic() + facet_wrap(.~ bilingual) + ggtitle("Bilingualism by handedness") + 
-  ylab("Count") + xlab("")
+
 # fifth, show miles
 a <- allsum %>%
-  group_by(miles, handedness) %>%
+  group_by(miles, handfac) %>%
     summarize(n = n())
-ggplot(data=a, aes(x=handedness, y=n, fill= handedness)) +
+ggplot(data=a, aes(x=handfac, y=n, fill= handfac)) +
   geom_bar(stat="identity")+
@@
 a <- allsum %>%
-  group_by(porta, handedness) %>%
+  group_by(porta, handfac) %>%
     summarize(n = n())
-ggplot(data=a, aes(x=handedness, y=n, fill= handedness)) +
+ggplot(data=a, aes(x=handfac, y=n, fill= handfac)) +
   geom_bar(stat="identity")+
@@
 # seventh, show LexTALE
-ggplot(allsum, aes(x = handedness, y = lexTALE)) +
-  geom_pirate(aes(colour = handedness), bars = FALSE) + theme_classic() + 
+ggplot(allsum, aes(x = handfac, y = lexTALE)) +
+  geom_pirate(aes(colour = handfac), bars = FALSE) + theme_classic() + 
   ylim(70, 110) + ggtitle("LexTALE by handedness") + ylab("Standardised score") + xlab("")
@@
 
-#check lateralisaiton with single group t vs zero
+#check lateralisation with single group t vs zero
 for (i in colnums){

### 2020-05-28 pre-registered_analyses.Rmd
First go at addressing the hypotheses outlined in the pre-reg

---
title: "Pre-registered Analyses"
author: "Adam Parker"
date: "28/05/2020"
output:
  pdf_document: default
  html_document: default
---

The Markdown script addresses the research questions outlined in the pre-registration for "Assessing reliability of an online behavioural laterality battery"  (https://osf.io/fktmh).

Below are the research questions that we sought to address.

Research Question

- Is it possible to achieve test-retest reliability of .65 or more with laterality indices derived from the four laterality tests? If so, could this be achieved more efficiently by using fewer trials in each test?

- Laterality indices can be calculated in multiple different ways (e.g. using RT, accuracy or a combination of the two; as a raw difference score or a proportion). We will explore these options to determine which metric shows the strongest test-retest reliability.

- Do extreme left-handers show a higher rate of atypical laterality than right-handers on the Rhyme Decision, Dichotic listening, or Chimeric Faces tasks?

-	Do the laterality indices from the language and faces laterality tasks correlate? This question will be considered separately for left- and right-handers, since our previous research on language laterality indicated a single factor provided a good fit to data from right-handers but not left-handers.

To answer these questions, we derived four hypotheses. Each are tested below after the reliability assessment.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
library(MASS)
library(lavaan)
library(qpcR)
library(tidyverse)
library(semPlot)
library(RVAideMemoire)
```

# Reliability

For the reliability assessment, we will first plot the distributions of laterality indices and check them for normality using the Shapiro-Wilk test. For normally-distributed scores, we will compute Pearson correlation coefficients for time 1 vs time 2, after excluding outliers. If scores are not normally distributed, we will consider transforming data to achieve normality (see ‘Transformations’ below), or alternatively use a Spearman correlation if, for instance, data are bimodally distributed.

First, we read in the LI data which was generated by running the the "battery-preprocessing_v5_DB.Rmd" script.

```{r read_dat}
# read in data from csv
LIs <- read.csv("LIs.csv", na.strings = c("NA", ""))
# then we exclude those who we do not have full datasets for 
LIs <- subset(LIs, exclude== 0)
```

Now we plot the distribution of the correlations.

### Rhyme Detection

**LI**

Here, we assess the LI reliability for the Rhyme detection task (RT). First this is done for raw LI.

```{r RDT LI reliability, include=TRUE}
# first, we consider the LIs for RDT accuracy
# day 1
# plot density
ggplot(LIs, aes(x=Day1.RDT_RT_LI))+geom_density(linetype="dashed")+theme_classic()+ggtitle("RDT RT LI (day 1)")
# normality
shapiro.test(LIs$Day1.RDT_RT_LI)
# day 2
# plot denisty
ggplot(LIs, aes(x=Day2.RDT_RT_LI))+geom_density(linetype="dashed")+theme_classic()+ggtitle("RDT RT LI (day 2)")
# normality
shapiro.test(LIs$Day2.RDT_RT_LI)
# as neither are of normal distribution, conduct Spearman's
# plot
rhyme_scatter <- ggplot(LIs, aes(x= Day1.RDT_RT_LI, y= Day2.RDT_RT_LI)) +
  geom_hline(yintercept=0, color = "black") +
  geom_vline(xintercept=0, color = "black") +
  geom_point(aes(color= handedness)) + 
  geom_rug(aes(color= handedness)) +
  theme_classic() +
  ylab("Rhyme decision LI (day 2)") + 
  xlab("Rhyme decision LI (day 1)") +
  annotate("text", x = 8, y = 8, label = "paste(italic(R), \" = .47\")", parse = TRUE) +
  theme(legend.position = c(.15, .90))
# Set relative size of marginal plots (main plot 10x bigger than marginals)
ggExtra::ggMarginal(rhyme_scatter, groupColour = TRUE, groupFill = TRUE)
# spearman's CIs
spearman.ci(LIs$Day1.RDT_RT_LI, LIs$Day2.RDT_RT_LI, nrep = 10000, conf.level = 0.95)
# spearman's test
cor.test(LIs$Day1.RDT_RT_LI, LIs$Day2.RDT_RT_LI, method = "spearman", alternative = "greater")
```

**t-score**

Now consider the t score.

```{r RDT t reliability, include=TRUE}
# first, we consider the LIs for RDT accuracy
# day 1
# plot density
ggplot(LIs, aes(x=t1))+geom_density(linetype="dashed")+theme_classic()+ggtitle("RDT t-score LI (day 1)")
# normality
shapiro.test(LIs$t1)
# day 2
# plot denisty
ggplot(LIs, aes(x=t2))+geom_density(linetype="dashed")+theme_classic()+ggtitle("RDT t-score LI (day 2)")
# normality
shapiro.test(LIs$t2)
# as neither are of normal distribution, conduct Spearman's
# plot
rhyme_scatter <- ggplot(LIs, aes(x= t1, y= t2)) +
  geom_hline(yintercept=0, color = "black") +
  geom_vline(xintercept=0, color = "black") +
  geom_point(aes(color= handedness)) + 
  geom_rug(aes(color= handedness)) +
  theme_classic() +
  ylab("Rhyme decision t-score (day 2)") + 
  xlab("Rhyme decision t-score (day 1)") +
  annotate("text", x = 8, y = 8, label = "paste(italic(R), \" = -.03\")", parse = TRUE) +
  theme(legend.position = c(.15, .90))
# Set relative size of marginal plots (main plot 10x bigger than marginals)
ggExtra::ggMarginal(rhyme_scatter, groupColour = TRUE, groupFill = TRUE)
# spearman's CIs
spearman.ci(LIs$t1, LIs$t2, nrep = 10000, conf.level = 0.95)
# spearman's test
cor.test(LIs$t1, LIs$t2, method = "spearman", alternative = "greater")
```

Analysis indicates this is below the satisfactory r= .65, so the utility of this task doesn't seem ideal.

### Dichotic listening

**LI**

```{r DL reliability, include=TRUE}
# day 1
# plot density
ggplot(LIs, aes(x=Day1_DL_acc_LI))+geom_density(linetype="dashed")+theme_classic()+ggtitle("DL ACC LI (day 1)")
# normality
shapiro.test(LIs$Day1_DL_acc_LI)
# day 2
# plot denisty
ggplot(LIs, aes(x=Day2_DL_acc_LI))+geom_density(linetype="dashed")+theme_classic()+ggtitle("DL ACC LI (day 2)")
# normality
shapiro.test(LIs$Day1_DL_acc_LI)
# as neither are of normal distribution, conduct Spearman's
# plot
DL_scatter <- ggplot(LIs, aes(x= Day1_DL_acc_LI, y= Day2_DL_acc_LI)) +
  geom_hline(yintercept=0, color = "black") +
  geom_vline(xintercept=0, color = "black") +
  geom_point(aes(color= handedness)) +
  geom_rug(aes(color= handedness)) +
  theme_classic() +
  ylab("Dichotic listening LI (day 2)") + 
  xlab("Dichotic listening LI (day 1)") +
  xlim(-100, 100) + ylim(-100, 100) +
  annotate("text", x = 90, y = 90, label = "paste(italic(R), \" = .78\")", parse = TRUE) +
  theme(legend.position = c(.15, .90))
# Set relative size of marginal plots (main plot 10x bigger than marginals)
ggExtra::ggMarginal(DL_scatter, groupColour = TRUE, groupFill = TRUE)
# spearman's CIs
spearman.ci(LIs$Day1_DL_acc_LI, LIs$Day2_DL_acc_LI, nrep = 10000, conf.level = 0.95)
# spearman's test
cor.test(LIs$Day1_DL_acc_LI, LIs$Day2_DL_acc_LI, method = "spearman", alternative = "greater")
```

**z-score**

```{r DL z reliability, include=TRUE}
# day 1
# plot density
ggplot(LIs, aes(x=Day1_Dich_acc_Z))+geom_density(linetype="dashed")+theme_classic()+ggtitle("DL ACC z-score (day 1)")
# normality
shapiro.test(LIs$Day1_Dich_acc_Z)
# day 2
# plot denisty
ggplot(LIs, aes(x=Day2_Dich_acc_Z))+geom_density(linetype="dashed")+theme_classic()+ggtitle("DL ACC z-score (day 2)")
# normality
shapiro.test(LIs$Day2_Dich_acc_Z)
# as neither are of normal distribution, conduct Spearman's
# plot
DL_scatter <- ggplot(LIs, aes(x= Day1_Dich_acc_Z, y= Day2_Dich_acc_Z)) +
  geom_hline(yintercept=0, color = "black") +
  geom_vline(xintercept=0, color = "black") +
  geom_point(aes(color= handedness)) +
  geom_rug(aes(color= handedness)) +
  theme_classic() +
  ylab("Dichotic listening z-score (day 2)") + 
  xlab("Dichotic listening z-score (day 1)") +
  xlim(-100, 100) + ylim(-100, 100) +
  annotate("text", x = 90, y = 90, label = "paste(italic(R), \" = .78\")", parse = TRUE) +
  theme(legend.position = c(.15, .90))
# Set relative size of marginal plots (main plot 10x bigger than marginals)
ggExtra::ggMarginal(DL_scatter, groupColour = TRUE, groupFill = TRUE)
# spearman's CIs
spearman.ci(LIs$Day1_Dich_acc_Z, LIs$Day2_Dich_acc_Z, nrep = 10000, conf.level = 0.95)
# spearman's test
cor.test(LIs$Day1_Dich_acc_Z, LIs$Day2_Dich_acc_Z, method = "spearman", alternative = "greater")
```

The dichotic listening task comes out pretty good and exceeds our threshold for reliabiliity.

### Chimeric Faces

**LI**

```{r CF reliability, include=TRUE}
# day 1
# plot density
ggplot(LIs, aes(x=Day1_CF_acc_LI))+geom_density(linetype="dashed")+theme_classic()+ggtitle("CF ACC LI (day 1)")
# normality
shapiro.test(LIs$Day1_CF_acc_LI)
# day 2
# plot denisty
ggplot(LIs, aes(x=Day2_CF_acc_LI))+geom_density(linetype="dashed")+theme_classic()+ggtitle("CF ACC LI (day 2)")
# normality
shapiro.test(LIs$Day2_CF_acc_LI)
# as neither are of normal distribution, conduct Spearman's
# plot
CF_scatter <- ggplot(LIs, aes(x= Day1_CF_acc_LI, y= Day2_CF_acc_LI)) +
  geom_hline(yintercept=0, color = "black") +
  geom_vline(xintercept=0, color = "black") +
  geom_point(aes(color= handedness)) +
  geom_rug(aes(color= handedness)) +
  theme_classic() +
  ylab("Chimeric faces LI (day 2)") + 
  xlab("Chimeric faces LI (day 1)") +
  xlim(-100, 100) + ylim(-100, 100) +
  annotate("text", x = 90, y = 90, label = "paste(italic(R), \" = .87\")", parse = TRUE) +
  theme(legend.position = c(.15, .90))
# Set relative size of marginal plots (main plot 10x bigger than marginals)
ggExtra::ggMarginal(CF_scatter, groupColour = TRUE, groupFill = TRUE)
# spearman's CIs
spearman.ci(LIs$Day1_CF_acc_LI, LIs$Day2_CF_acc_LI, nrep = 10000, conf.level = 0.95)
# spearman's test
cor.test(LIs$Day1_CF_acc_LI, LIs$Day2_CF_acc_LI, method = "spearman", alternative = "greater")
```

**z-score**

```{r CF z reliability, include=TRUE}
# day 1
# plot density
ggplot(LIs, aes(x=Day1_CF_acc_Z))+geom_density(linetype="dashed")+theme_classic()+ggtitle("CF ACC z-score (day 1)")
# normality
shapiro.test(LIs$Day1_CF_acc_Z)
# day 2
# plot denisty
ggplot(LIs, aes(x=Day2_CF_acc_Z))+geom_density(linetype="dashed")+theme_classic()+ggtitle("CF ACC z-score (day 2)")
# normality
shapiro.test(LIs$Day2_CF_acc_Z)
# as neither are of normal distribution, conduct Spearman's
# plot
CF_scatter <- ggplot(LIs, aes(x= Day1_CF_acc_Z, y= Day2_CF_acc_Z)) +
  geom_hline(yintercept=0, color = "black") +
  geom_vline(xintercept=0, color = "black") +
  geom_point(aes(color= handedness)) +
  geom_rug(aes(color= handedness)) +
  theme_classic() +
  ylab("Chimeric faces z-score (day 2)") + 
  xlab("Chimeric faces z-score (day 1)") +
  xlim(-100, 100) + ylim(-100, 100) +
  annotate("text", x = 90, y = 90, label = "paste(italic(R), \" = .88\")", parse = TRUE) +
  theme(legend.position = c(.15, .90))
# Set relative size of marginal plots (main plot 10x bigger than marginals)
ggExtra::ggMarginal(CF_scatter, groupColour = TRUE, groupFill = TRUE)
# spearman's CIs
spearman.ci(LIs$Day1_CF_acc_Z, LIs$Day2_CF_acc_Z, nrep = 10000, conf.level = 0.95)
# spearman's test
cor.test(LIs$Day1_CF_acc_Z, LIs$Day2_CF_acc_Z, method = "spearman", alternative = "greater")
```

The chimeric faces task does well and seems very reliable.

### Finger tapping

**LI**

```{r FT reliability, include=TRUE}
# day 1
# plot density
ggplot(LIs, aes(x=Day1_finger_LI))+geom_density(linetype="dashed")+theme_classic()+ggtitle("FT LI (day 1)")
# normality
shapiro.test(LIs$Day1_finger_LI)
# day 2
# plot denisty
ggplot(LIs, aes(x=Day2_finger_LI))+geom_density(linetype="dashed")+theme_classic()+ggtitle("FT LI (day 2)")
# normality
shapiro.test(LIs$Day2_finger_LI)
# as neither are of normal distribution, conduct Spearman's
# plot
finger_scatter <- ggplot(LIs, aes(x= Day1_finger_LI, y= Day2_finger_LI)) +
  geom_hline(yintercept=0, color = "black") +
  geom_vline(xintercept=0, color = "black") +
  geom_point(aes(color= handedness)) +
  geom_rug(aes(color= handedness)) +
  theme_classic() +
  ylab("Finger tapping LI (day 2)") + 
  xlab("Finger tapping LI (day 1)") +
  annotate("text", x = 18, y = 18, label = "paste(italic(R), \" = .78\")", parse = TRUE) +
  theme(legend.position = c(.15, .90))
# Set relative size of marginal plots (main plot 10x bigger than marginals)
ggExtra::ggMarginal(finger_scatter, groupColour = TRUE, groupFill = TRUE)
# spearman's CIs
spearman.ci(LIs$Day1_finger_LI, LIs$Day2_finger_LI, nrep = 10000, conf.level = 0.95)
# spearman's test
cor.test(LIs$Day1_finger_LI, LIs$Day2_finger_LI, method = "spearman", alternative = "greater")
```

**z-score**

```{r FT z reliability, include=TRUE}
# day 1
# plot density
ggplot(LIs, aes(x=Day1_Finger_acc_Z))+geom_density(linetype="dashed")+theme_classic()+ggtitle("FT z-score (day 1)")
# normality
shapiro.test(LIs$Day1_Finger_acc_Z)
# day 2
# plot denisty
ggplot(LIs, aes(x=Day2_Finger_acc_Z))+geom_density(linetype="dashed")+theme_classic()+ggtitle("FT z-score (day 2)")
# normality
shapiro.test(LIs$Day2_Finger_acc_Z)
# as neither are of normal distribution, conduct Spearman's
# plot
finger_scatter <- ggplot(LIs, aes(x= Day1_Finger_acc_Z, y= Day2_Finger_acc_Z)) +
  geom_hline(yintercept=0, color = "black") +
  geom_vline(xintercept=0, color = "black") +
  geom_point(aes(color= handedness)) +
  geom_rug(aes(color= handedness)) +
  theme_classic() +
  ylab("Finger tapping z-score (day 2)") + 
  xlab("Finger tapping z-score (day 1)") +
  annotate("text", x = 18, y = 18, label = "paste(italic(R), \" = .78\")", parse = TRUE) +
  theme(legend.position = c(.15, .90))
# Set relative size of marginal plots (main plot 10x bigger than marginals)
ggExtra::ggMarginal(finger_scatter, groupColour = TRUE, groupFill = TRUE)
# spearman's CIs
spearman.ci(LIs$Day1_Finger_acc_Z, LIs$Day2_Finger_acc_Z, nrep = 10000, conf.level = 0.95)
# spearman's test
cor.test(LIs$Day1_Finger_acc_Z, LIs$Day2_Finger_acc_Z, method = "spearman", alternative = "greater")
```

The finger tapping tasks yields a reliable correlation, and seems to relate to categorical handedness.

**Finger tapping and EHI**

```{r fingerEHI}
# plot
finger_scatter <- ggplot(LIs, aes(x= index_EHI, y= Day1_finger_LI)) +
  geom_hline(yintercept=0, color = "black") +
  geom_vline(xintercept=0, color = "black") +
  geom_point(aes(color= handedness)) +
  geom_rug(aes(color= handedness)) +
  theme_classic() +
  ylab("Finger tapping LI (day 1)") + 
  xlab("EHI") +
  annotate("text", x = 90, y = 25, label = "paste(italic(R), \" = .58\")", parse = TRUE) +
  theme(legend.position = c(.15, .90))
# Set relative size of marginal plots (main plot 10x bigger than marginals)
ggExtra::ggMarginal(finger_scatter, groupColour = TRUE, groupFill = TRUE)
# spearman's CIs
spearman.ci(LIs$Day1_finger_LI, LIs$index_EHI, nrep = 10000, conf.level = 0.95)
# spearman's test
cor.test(LIs$Day1_finger_LI, LIs$index_EHI, method = "spearman", alternative = "greater")
# LEFT HANDERS
left <- subset(LIs, handedness== "Left")
# spearman's test
cor.test(left$Day1_finger_LI, left$index_EHI, method = "spearman", alternative = "greater")
# Right HANDERS
right <- subset(LIs, handedness== "Right")
# spearman's test
cor.test(right$Day1_finger_LI, right$index_EHI, method = "spearman", alternative = "greater")
```

Upon an initial look, there is a reasonable correlation between the EHI and day 1 on the finger tapping. However, it does look like there are two clusters. So, we take a loo at the left and right handers seperately. Doing so yields dissappointing results.

## Hypothesis 1

At the population level, the laterality indices for Rhyme Decision will show a right visual field advantage (indicating left hemisphere dominance for language processing); the laterality indices for the Dichotic Listening task will show a right ear advantage (indicating left hemisphere dominance for language); the laterality indices for Chimeric Faces will show a left visual field advantage (indicating right hemisphere dominance for face or emotion processing); and the Finger Tapping task will show a right hand advantage.

To assess hypothesis 1, we pre-registered that the predictions would be tested using a one-sample t-test for each task using session 1 data.

All LIs were calculated as: [(R - L)/(R + L)] × 100, resulting in values between -100 and 100

Now we run a one-sampled t-test against zero for each test using day 1 data. We consider the accuracy LI for chimeric faces, dichotic listening, and finger tapping. The RT LI is used for the rhyme task.

```{r H1}
# corrplot the correlation matrix
mycols <- c('Day1_CF_acc_LI', 'Day1_DL_acc_LI', 'Day1.RDT_RT_LI', 'Day1_finger_LI', "handedness")
colnums<- which(colnames(LIs) %in% mycols)
M <- LIs[,colnums]
N <- M %>% 
  rename(
    "CF" = Day1_CF_acc_LI,
    "DL" = Day1_DL_acc_LI,
    "RD" = Day1.RDT_RT_LI,
    "FT" = Day1_finger_LI)
# PLOT TO SEE IF DIFFERENT FROM ZERO
N$ID <- 1:nrow(N)
# format to long
data_long <- tidyr::gather(N, task, index, `CF`:`FT`, factor_key=TRUE)
# plot data
a <- ggplot(subset(data_long, task== "RD"), aes(x = task, y = index)) +
  geom_hline(yintercept=0, linetype="dashed", color = "red") +
  ggpirate::geom_pirate(aes(colour = "black"), bars = FALSE, 
                        points_params = list(shape = 19, alpha = 0.5), lines_params = list(size = 0.8)) +
  ylab("Laterality Index") + xlab("") + ylim(-30, 30) +
  scale_colour_manual(values = "darkgray") + theme_classic() 
b <- ggplot(subset(data_long, task== "DL"), aes(x = task, y = index)) +
  geom_hline(yintercept=0, linetype="dashed", color = "red") +
  ggpirate::geom_pirate(aes(colour = "black"), bars = FALSE, 
                        points_params = list(shape = 19, alpha = 0.5), lines_params = list(size = 0.8)) +
  ylab("Laterality Index") + xlab("") + ylim(-100, 100) +
  scale_colour_manual(values = "darkgray") + theme_classic() 
c <- ggplot(subset(data_long, task== "CF"), aes(x = task, y = index)) +
  geom_hline(yintercept=0, linetype="dashed", color = "red") +
  ggpirate::geom_pirate(aes(colour = "black"), bars = FALSE, 
                        points_params = list(shape = 19, alpha = 0.5), lines_params = list(size = 0.8)) +
  ylab("Laterality Index") + xlab("") + ylim(-100, 100) +
  scale_colour_manual(values = "darkgray") + theme_classic() 
d <- ggplot(subset(data_long, task== "FT"), aes(x = task, y = index)) +
  geom_hline(yintercept=0, linetype="dashed", color = "red") +
  ggpirate::geom_pirate(aes(colour = "black"), bars = FALSE, 
                        points_params = list(shape = 19, alpha = 0.5), lines_params = list(size = 0.8)) +
  ylab("Laterality Index") + xlab("") + ylim(-30, 30) +
  scale_colour_manual(values = "darkgray") + theme_classic() 
ggpubr::ggarrange(a, b, c, d)
t.test(LIs$Day1_CF_acc_LI, mu= 0)
t.test(LIs$Day1_DL_acc_LI, mu= 0)
t.test(LIs$Day1_finger_LI, mu= 0)
t.test(LIs$Day1.RDT_RT_LI, mu= 0)
```

For hypothesis 1,

- rhyme task is different from zero but indicates a LVF advantage.

- dichotic listening is different from zero and shows a right ear advantage.

- chimeric faces is different from zero and shows a left visual field advantage.

- finger tapping is different from zero and shows a right hand advanatge.

## Hypothesis 2

Strong left-handers (i.e. those with laterality quotient of -90 or less on the Edinburgh Handedness Inventory) will be more likely to show atypical laterality (i.e. laterality in the opposite direction to our population level predictions) than right-handers. This prediction follows from observations by Mazoyer et al (2014) who studied a large group of left- and right-handers using fMRI to measure language laterality. They concluded "concordance of hemispheric dominance for hand and for language occurs barely above the chance level, except in a group of rare individuals (less than 1% in the general population) who exhibit strong right hemisphere dominance for both language and their preferred hand."

The test of this hypothesis will be complicated by having to decide how to define 'atypical'. Ideally, these would be individuals who have significant lateralisation in the opposite direction to the majority, as indicated by p < .05 on the individual z-score or t-score. We anticipate, however, that this criterion may select very few cases on some of our tasks, in which case we will adjust the cutoff on that task until we have at least 20 atypical cases. Once typicality is defined, we will first analyse atypicality on the Rhyme Decision and Chimeric Faces task. Participants will be categorised into four groups: typical on both; atypical on Rhyme Decision only; atypical on Chimeric Faces only, and atypical on both. A chi square test will be used to test for association with the number of people in each of the four groups and the handedness classification from the Edinburgh Handedness Inventory. This will then be repeated for the Dichotic Listening and the Chimeric Faces tasks.

As our rhyme task was not relibale, we drop this task here and focus only on the dichotic and faces.

```{r H2, include=TRUE}
# create new variable for atypicality 
LIs <- 
  LIs %>% 
  mutate(atypicality= ifelse(Day1_CF_acc_Z > 1.96 & Day1_Dich_acc_Z < -1.96, "typical",
                          ifelse(Day1_CF_acc_Z > 1.96 & Day1_Dich_acc_Z > 1.96, "atypical dichotic",
                                 ifelse(Day1_CF_acc_Z < -1.96 & Day1_Dich_acc_Z < -1.96, "atypical face",
                                        ifelse(Day1_CF_acc_Z < -1.96 & Day1_Dich_acc_Z > 1.96, "atypical both", NA)))))
# create handedness for handedness on EHI 
LIs <- 
  LIs %>% 
  mutate(hand_90= ifelse(index_EHI >= 90, "right handed",
                         ifelse(index_EHI <= -9, "left handed", NA)))
# create a table for chi-squared
tbl <- table(LIs$atypicality, LIs$hand_90)
tbl
# run chi squared
chisq.test(tbl)
# use balloon plot to visualise
gplots::balloonplot(t(tbl), main ="Association of EHI handeness and atypical laterality", xlab ="", ylab="",
            label = FALSE, show.margins = FALSE)
```

Using a t score of 1.96 as our cutoff (equaiting to .05 on a two-tailed test) results in more than the 20 atypicals we wanted. Left handers do appear to depart from typcality more.

## Hypothesis 3

For right-handers, a model

(a) where the two language tasks are correlated with each other but not with the chimeric faces will provide better fit to the data than a model

(b) in which laterality indices are correlated between the language and chimeric face tasks, or a base model

(c) where each task is independently lateralised.

For left-handers, we anticipate that that model (c) may be a better fit to the data, indicating that laterality is dissociated across the different tasks.

Using a structural equation modelling framework, we will compare the covariances of the laterality statistics between the two language tasks and chimeric faces to test between three models. This will be done separately for left- and right-handers. Before doing this we will flip the polarity of chimeric faces, so that all predictions are for positive correlations.

- Model 1: the only significant correlations will be within-task. I.e. each task will be reliable in the sense that it correlates with itself, but it will not correlate with other tasks.

- Model 2: there will be significant within-task correlations, and significant between-task correlations for the language tasks only. The chimeric faces will not be correlated with either language task.

- Model 3: There are significant correlations between all tasks, indicating a single laterality factor affecting performance.

We anticipate that model 2 may fit right-handers, whereas model 1 seems more likely for left-handers.

```{r prep_dat}
# now create subset of data frame for analysis
mycols <- c('Day1_CF_acc_LI', 'Day2_CF_acc_LI',
            'Day1_DL_acc_LI', 'Day2_DL_acc_LI',
            'Day1.RDT_RT_LI', 'Day2.RDT_RT_LI',
            'handedness')
colnums<- which(colnames(LIs) %in% mycols)
M <- LIs[,colnums]
M <- M %>% 
  rename(
    "faces1" = Day1_CF_acc_LI,
    "faces2" = Day2_CF_acc_LI,
    "dichotic1" = Day1_DL_acc_LI,
    "dichotic2" = Day2_DL_acc_LI,
    "rhyme1" = Day1.RDT_RT_LI,
    "rhyme2" = Day2.RDT_RT_LI)
# subset into right and left handers
left_dat <- subset(M, handedness== "Left")
right_dat <- subset(M, handedness== "Right")
```

### Left handers:

Here we model left hander data with three prespecified covariance structures, then compare using AIC weights.

This is a subset of SEM that dos not include any latent variables or directional paths. It only looks to constrain particular covariance patterns and report 'best' model according AIC weights.

```{r Roger_mod_left}
model1<-"
dichotic1~~0*rhyme1
dichotic1~~0*faces1 
rhyme1~~0*faces1 
"

model2<-"
dichotic1~~rhyme1 
dichotic1~~0*faces1
rhyme1~~0*faces1 
"

model3<-"
dichotic1~~rhyme1 
dichotic1~~faces1 
rhyme1~~faces1
"

fit1A <- cfa(model1, data=left_dat)
fit2A <- cfa(model2, data=left_dat)
fit3A <- cfa(model3, data=left_dat)

aic_vector1 <- c(fitMeasures(fit1A, "aic"),fitMeasures(fit2A, "aic"),fitMeasures(fit3A, "aic"))

summary(fit1A, fit.measures=TRUE)
summary(fit2A, fit.measures=TRUE)
summary(fit3A, fit.measures=TRUE)

AICweights1<-akaike.weights(aic_vector1)$weights

IC_objectweights1<-data.frame(Model=factor(rep(c('Model 1','Model 2','Model 3'))), IC = factor(rep('AIC',each=3)),values=AICweights1)

ggplot(IC_objectweights1,aes(Model,values,fill=Model))+
  geom_bar(stat='identity',position="dodge")+
  ylab('Akaike weights') + theme_classic()
```

We now repeat this for right handers.

```{r Roger_mod_right}
fit1B <- cfa(model1, data=right_dat)
fit2B <- cfa(model2, data=right_dat)
fit3B <- cfa(model3, data=right_dat)

aic_vector2 <- c(fitMeasures(fit1B, "aic"),fitMeasures(fit2B, "aic"),fitMeasures(fit3B, "aic"))

summary(fit1B, fit.measures=TRUE)
summary(fit2B, fit.measures=TRUE)
summary(fit3B, fit.measures=TRUE)

AICweights2<-akaike.weights(aic_vector2)$weights

IC_objectweights2<-data.frame(Model=factor(rep(c('Model 1','Model 2','Model 3'))), IC = factor(rep('AIC',each=3)),values=AICweights2)

ggplot(IC_objectweights2,aes(Model,values,fill=Model)) +
  geom_bar(stat='identity',position="dodge")+
  ylab('Akaike weights') + theme_classic()
```

### 2020-05-28 pre-registered_analyses.Rmd
bonferroni corrections applied

@@
 
-```{r H2, include=TRUE}
+**Rhyme**
+
+```{r H2 dichotic, include=TRUE}
 # create new variable for atypicality 
@@
   LIs %>% 
-  mutate(atypicality= ifelse(Day1_CF_acc_Z > 1.96 & Day1_Dich_acc_Z < -1.96, "typical",
+  mutate(atypicality_rhyme= ifelse(Day1_CF_acc_Z > 1 & t1 < -1, "typical",
+                          ifelse(Day1_CF_acc_Z > 1 & t1 > 1, "atypical rhyme",
+                                 ifelse(Day1_CF_acc_Z < -1 & t1 < -1, "atypical face",
+                                        ifelse(Day1_CF_acc_Z < -1 & t1 > 1, "atypical both", NA)))))
+# create handedness for handedness on EHI 
+LIs <- 
+  LIs %>% 
+  mutate(hand_90= ifelse(index_EHI >= 90, "right handed",
+                         ifelse(index_EHI <= -9, "left handed", NA)))
+# create a table for chi-squared
+tbl <- table(LIs$atypicality_rhyme, LIs$hand_90)
+tbl
+# run chi squared
+chisq.test(tbl)
+p <- chisq.test(tbl)[3]
+p.adjust(p, "bonferroni", n=2)
+# use balloon plot to visualise
+gplots::balloonplot(t(tbl), main ="Association of EHI handeness and atypical laterality (faces and rhyme)", xlab ="", ylab="",
+            label = FALSE, show.margins = FALSE)
+```
+
+With the rhyme task, the reliability is low.
+Because of this we have to use a t value that is less than 1.96 (i.e.
+larger than 0.05).
+This produces no association.
+But, this isn't a big deal as we can't be sure that the task is measuring language lateralisation.
+
+**Dichotic**
+
+```{r H2 dichotic, include=TRUE}
+# create new variable for atypicality 
+LIs <- 
+  LIs %>% 
+  mutate(atypicality_dichotic= ifelse(Day1_CF_acc_Z > 1.96 & Day1_Dich_acc_Z < -1.96, "typical",
                           ifelse(Day1_CF_acc_Z > 1.96 & Day1_Dich_acc_Z > 1.96, "atypical dichotic",
@@
 # create a table for chi-squared
-tbl <- table(LIs$atypicality, LIs$hand_90)
+tbl <- table(LIs$atypicality_dichotic, LIs$hand_90)
 tbl
@@
 chisq.test(tbl)
+p <- chisq.test(tbl)[3]
+p.adjust(p, "bonferroni", n=2)
 # use balloon plot to visualise
-gplots::balloonplot(t(tbl), main ="Association of EHI handeness and atypical laterality", xlab ="", ylab="",
+gplots::balloonplot(t(tbl), main ="Association of EHI handeness and atypical laterality (faces and dichotic)", xlab ="", ylab="",
             label = FALSE, show.margins = FALSE)

### 2020-06-08 draft_notebook_writeup.Rmd
Notebook merging original preregistration with Adam's analyses, which have been condensed into shortened form. Can't get 'include' and 'echo' right, so the tables aren't showing in html!

---
title: "Assessing reliability of an online behavioural laterality battery: a pre-registered study (draft notebook)"
output: html_notebook
---

## Authors
Adam J Parker, Zoe Woodhead, Paul Andrew Thompson, and Dorothy Bishop

We report here the results of analyses that were preregistered on Open Science Framework. The analyses here are based on the preregistration that was revised after piloting of the battery. On that basis, we made some modifications to the tasks and added a consonant-vowel Dichotic Listening task. We also adapted the analysis plan to include a structural equation modeling to look at dissociations between laterality on different tasks. The preregistration can be found <a href="https://osf.io/qe6b4"> here </a> .

## Introduction
Hemispheric specialisation (lateralisation) is a ubiquitous characteristic of the human brain, and yet many questions remain unanswered as to its causes and consequences. Much of the ambiguity in the literature may arise from unreliable measures of lateralisation and underpowered studies. This study aims to develop a reliable online battery to assess lateralisation, which could be used to test large numbers of participants.

The new battery consists of four behavioural laterality tests, which will be described in more detail in the Methods section:   
1.	Rhyme Decision: A visual half-field task in which participants decide whether a central written word rhymes with the name of a laterally-presented picture. This measures laterality of language processing.  
2.	Dichotic Listening: A dichotic listening task in which participants hear two consonant-vowel (CV) auditory stimuli simultaneously, one to each ear. They are required to report which CV stimulus out of six options (ba, da, ga, ka, pa, ta) provided the best match to the stimulus they heard . This measures laterality of language processing.  
3.	Chimeric Faces: A visual half-field task in which participants view stimuli that are vertically split at the midline with each hemiface displaying one of four emotions (anger, disgust, happiness, and sadness), They are required to report the overall emotion that they perceive in the face. This measures laterality of face and emotion processing.  
4.	Finger Tapping: A speeded sequential key-pressing task performed using a computer keyboard with left and right hands. This provides a quantitative measure of relative skill of the two hands.    
In addition, participants completed online versions of:

•	The Lexical Test for Advanced Learners of English (LexTALE; Lemhöfer & Broersma, 2012), to assess vocabulary as an indicator of language proficiency.  
•	The Edinburgh Handedness Inventory (Oldfield, 1971), to assess manual preference.  
•	Modified versions of the Miles Test (Miles, 1929) and the Porta Test (Porac & Coren, 1976) which are suitable for online testing, to assess ocular dominance.   
•	A basic demographic questionnaire. This included questions about age, gender, footedness, years in education, and bilingualism.

The study was primarily exploratory. Our principal goal was to refine the four laterality tests to maximise test-retest reliability. The whole battery was administered at an initial testing session, then the four laterality tests were repeated at a second session 3-10 days later. In addition, we aimed to test some predictions about the relationships between laterality on the four tests.

The primary outcome measure from each of the four tests is a laterality index; depending on the task, this is based on difference between left and right sides in accuracy (dichotic and chimeric faces tasks), reaction time (rhyme decision), or completed sequences (finger tapping) (see ‘Indices’ section below for details).

As well as establishing reliability of measures, we consider the inter-relationships between them. With online testing, there is the opportunity to test this using an adequately powered sample using measures of known reliability.  Prior studies have been inconclusive because it has been difficult gather large amounts of data to establish test-retest reliability. For instance, Van der Haegen and Brysbaert (2018), found that language laterality did not correlate with face laterality in visual half-field tasks. However, their study was restricted to left-handers, who may be atypical in this regard, and they noted that lack of association was hard to interpret because it may be due to unreliability of measurement.

Our research questions were as follows:

1.	Is it possible to achieve test-retest reliability of .65 or more with laterality indices derived from the four laterality tests?  If so, could this be achieved more efficiently by using fewer trials in each test?  
2.	Laterality indices can be calculated in multiple different ways (e.g. using RT, accuracy or a combination of the two; as a raw difference score or a proportion). We will explore these options (described in more detail in the ‘Indices’ section) to determine which metric shows the strongest test-retest reliability.  
3.	Do extreme left-handers show a higher rate of atypical laterality than right-handers on the Rhyme Decision, Dichotic listening, or Chimeric Faces tasks?  
4.	Do the laterality indices from the language and faces laterality tasks correlate? We aimed to consider this questiron separately for left- and right-handers, since our previous research on language laterality indicated a single factor provided a good fit to data from right-handers but not left-handers.

## Hypotheses

As noted above, our main goal was to devise a battery of laterality assessments with adequate test-retest reliability. In addition we had specific a priori hypotheses relating to questions 3 and 4 as follows:

### Hypothesis 1
At a population level, the laterality indices for Rhyme Decision will show a right visual field advantage (indicating left hemisphere dominance for language processing); the laterality indices for the Dichotic Listening task will show a right ear advantage (indicating left hemisphere dominance for language); the laterality indices for Chimeric Faces will show a left visual field advantage (indicating right hemisphere dominance for face or emotion processing); and the Finger Tapping task will show a right hand advantage.

### Hypothesis 2
Strong left-handers (i.e. those with laterality quotient of -90 or less on the Edinburgh Handedness Inventory) will be more likely to show atypical laterality (i.e. laterality in the opposite direction to our population level predictions) than right-handers. This prediction follows from observations by Mazoyer et al (2014) who studied a large group of left- and right-handers using fMRI to measure language laterality. They concluded "concordance of hemispheric dominance for hand and for language occurs barely above the chance level, except in a group of rare individuals (less than 1% in the general population) who exhibit strong right hemisphere dominance for both language and their preferred hand."

### Hypothesis 3
For right-handers, a model (a) where the two language tasks are correlated with each other but not with the chimeric faces will provide better fit to the data than a model (b) in which laterality indices are correlated between the language and chimeric face tasks, or a base model (c) where each task is independently lateralised. For left-handers, we anticipate that that model (c) may be a better fit to the data, indicating that laterality is dissociated across the different tasks.

## Design Plan
We note here specific study characteristics that are specified for a preregistration on OSF. We have added additional information to specify Methods in more detail.

### Study type
This is an observational Study in which data are collected from study subjects who are not randomly assigned to a treatment.

### Blinding
No blinding is involved

### Study design
We used a within-subjects design whereby participants performed each laterality task twice (repeated-measures) in order to establish test-retest reliability.

The battery of tasks was broken into four blocks. Within each block, participants completed 70 trials of the Rhyme Detection task; 36 trials of the Dichotic Listening task; 32 trials of the Chimeric Face task; and 2 trials of the Finger Tapping task for both their dominant and non-dominant hand. Trial order for blocks of Rhyme Detection, Dichotic Listening, and Chimeric Face task was randomised. For the Finger Tapping task, all participants were asked to respond first with their dominant hand followed by their non-dominant hand.

The order of tasks was balanced by using four possible orders as follows ABCD, BCDA, CDAB, DABC. Each successive participant entering the study was assigned the next task order in the sequence above. Order of tasks remained consistent across blocks for that participant.

### Randomization
No randomization was used in this study.

### Sampling Plan
The study was registered prior to collection of data

## Participants  
### Data collection procedures
We aimed to recruit participants via three methods:

1)	Given that we are also interested in genetically related samples, a portion of participants will be recruited via social media adverts, with encouragement for twins or first-degree relatives (siblings or parent-child pairs) to volunteer.

2)	Undergraduate students from the University of Oxford will be recruited via the Department of Experimental Psychology’s Research Participation Scheme.

3)	Other participants will be recruited from the Prolific web-based population (https://www.prolific.co/).

In practice, the first route was ineffective, with very few related individuals volunteering. 
<!--Adam: can we report how many came from Prolific and how many from student scheme?-->

### Inclusion criteria
Participants were required to be at least 16 years old, native English speakers, with no reported history of neurological disease (e.g. head injury, brain tumour, stroke), no history of significant hearing loss/impairment, and with normal or corrected-to-normal vision. We intentionally oversampled left-handers (see below).

### Sample size
We specified we would continue data collection until we had complete and useable datasets from 300 participants. As explained below, we had fewer participants than this for one task, because of many exclusions. However, as this task performed poorly, we decided not to test further participants.

### Sample size rationale
As our ultimate goal is to use these measures in individual differences research, we aim to optimise the task to achieve observed test-retest reliability of at least .65. To achieve sufficient precision of the estimate of reliability, we estimated we needed 300 participants. With this sample size, if the true correlation is .58, then the 95% confidence interval around the estimate will be .499-.65.  Thus with this sample size, if we observe a correlation of .65, then the true correlation is unlikely to be lower than .58.

We aimed to collect data for approximately 80 left-handers and 220 right-handers. We examined the power that a sample of 80 participants would provide to make the model comparisons outlined under hypothesis 3. By comparing model fit using a likelihood test, 80 participants would provide 100% power for model (a) and 97% power for model (b). Thus, we would have sufficient power to assess the fit of each model using Akaike weights, which are more conservative than likelihood tests.

### Stopping rule
We planned to cease data collection when we had collected 300 full datasets.

## Methods  
### Manipulated variables
In this test-retest design, the within-subject independent variable of interest was Session (1 or 2).  
In the four laterality tasks, stimuli were presented in either the left or right visual fields (Rhyme Decision and Chimeric Faces), to the left or right ear (Dichotic Listening), or using the left or right hand (Finger Tapping).

### Tasks 
The test battery was implemented online via Gorilla (https://www.gorilla.sc/).
<!--Adam: could you add information about the checks for screensize-->

## Rhyme Decision
This was a novel visual-half-field task devised for this study. The goal was to devise a task that engaged phonological processing by requiring the participant to generate a name from a pictured object and judge if it rhymed with a centrally-presented written word. Two pictures were presented to the left and right of fixation, and the prediction was that judgements should be faster when the rhyming picture occurred in the right visual field, which projects directly to the left hemisphere. The stimulus trios (two pictures and a word) are shown in Appendix A. On xx trials, stimuli were presented for which there was no rhyme match; this was used as a means of identifying and excluding participants who were not attending carefully.

## Dichotic Listening 
Materials were taken from a dichotic listening task that has been developed as an app (Hugdahl et al., 2009). Participants wear stereo headphones. On each trial they hear two consonant-vowel (CV) auditory stimuli simultaneously, one to each ear. They are required to report which CV stimulus out of six options (ba, da, ga, ka, pa, ta) provided the best match to the stimulus they heard. This measures laterality of language processing. The original test has xx dichotic stimulus pair, xx of which present the same syllable in both ears. These latter stimuli could be used to identify individuals who had difficulty in discriminating the syllables. 
<!--Adam: could you add information about the checks for headphones-->

##	Chimeric Faces
Materials were taken from Chimeric Faces task developed by Burt & Perrett (1997) and used by Innes, Burt, Birch, & Hausmann (2016). Participants view stimuli that are vertically split at the midline with each hemiface displaying one of four emotions (anger, disgust, happiness, and sadness). They are required to report the overall emotion that they perceive in the face. This measures laterality of face and emotion processing.  The task also includes xx trials in which both sides of the face show the same emotion. This is used as a check to identify whether the participant can reliably identify the intended emotions when there is no conflicting information.

## Finger Tapping
This was a simple task devised for this study. Our goal was to see whether we could obtain a quantification of extent of relative motor skill of the two hands using the individual's computer keyboard. The participant was asked to use the index finger to tap a specified sequence of keys continuously for a 20 second <!--Adam: is that right?--> interval, and the number of correct sequences was counted. The task was performed first with the preferred hand and then with the nonpreferred hand.

## Other tasks  
The Lexical Test for Advanced Learners of English (LexTALE; Lemhöfer & Broersma, 2012) was used to assess vocabulary as an indicator of language proficiency.
The Edinburgh Handedness Inventory (Oldfield, 1971) was used to assess manual preference.  
Modified versions of the Miles Test (Miles, 1929) and the Porta Test (Porac & Coren, 1976) were used to assess ocular dominance, and a question about foot preference.
<!--Adam Could you give more information about Lextale, Miles/Porta, and what was the foot preference question-->

## Measured variables  
For Rhyme Detection, Dichotic Listening, and Chimeric Faces, the measured variables were reaction time (RT) and accuracy. These were converted into laterality indices, as described below. For the Finger Tapping task, the measured variable was correct number of sequences produced with the right and left hands. This was also be converted into a laterality index.

For the LexTALE task, the measured variables was the number of words with correct responses (out of 40) and the number of nonwords with correct responses (out of 20). An adjusted score was computed by averaging the percentage of correct words and the percentage of correct nonwords (e.g., ((number of words correct/40*100) + (number of nonwords correct/20*100)) / 2).

The Edinburgh Handedness Inventory was scored in the standard way, which involves combining information about direction and exclusiveness of hand preference across ten tasks. For simple contrast between left- and right-handers, we treated indices greater than 0 as right handed, and indices less than 0 as left handed. For analysis of extreme left-handers, we focussed on left-handers with an index of -90 or less, which corresponds to near-exclusive use of the left hand.

Adapted versions of the Miles Test and Porta Test will classify participants as being either left or right eye dominant.<!---Adam: not sure if we need this here, as we did not specify how they would be used in analysis?--->

## Laterality Indices
The classic way of calculating a laterality index is to take the difference between left and right sides and divide it between the total for the two sides. This gives a difference measure that is scaled by the individual's overall speed or accuracy.

We also experimented with an alternative approach that allows one to identify whether an individual participant is reliably lateralised. For Reaction Time measures, this involved taking all the RTs for left and right-sided responses in a single participant, and conducting a t-test. One can then set a cutoff for reliable lateralisation, e.g. by using a t-value of 1.96, which would occur in around 5% of individuals if there was no real bias to one side. Based on considerations from pilot data, we planned to transform RT data from the RDT task by  log-transformation before computing t-values.

Similar logic can be used for accuracy measures, where a z-score can be computed and interpreted in relation to the normal distribution, using the formula:
z = (pR-.5)/sqrt(pR*pL/n) 
where pR is the proportion of R responses, pL is proportion of L responses, and n is total L and R responses.

These t- and z-measures amount to rescaling of the conventional laterality quotient and so are not expected to give different values for reliability, but they may be useful if one wanted to distinguish between individuals who showed robust lateralisation versus those who showed less convincing bias to one side.

### Data exclusion

In the laterality tasks, individual trials will be excluded from analysis if the reaction time was faster than 200 ms, as this suggests an anticipatory response. We did not preregister an upper RT boundary, but planned to apply Hoaglin-Iglewicz criteria (Hoaglin & Iglewicz 1987) to minimise the influence of outliers in the rightwards tail of the distribution. As such, outliers are defined as RTs that are 1.65 times the difference between the first and third quantiles (1.65*(Q3-Q1)) below or above the first and third quantile values. Our cut-off criterion was set to 1.65 instead of the typical 2.2. This is because we expected to obtain asymmetric distributions for reaction time, since reaction time is always non-negative, and so most outliers would be in one tail of the distribution.

Participants will be excluded if:
•	Responses were excluded or missing for more than 20% of trials in any condition in either of the language laterality tasks or chimeric faces.  
•	Performance for stimuli presented in their dominant visual-half field is not significantly above chance level. Moreover, participants whose performance for stimuli presented in their non-dominant visual half field is significantly below chance will also be excluded.   
•	LexTALE adjusted score is below 80, indicating that a participant’s level of vocabulary knowledge is below the cut off for lower advanced proficiency. We anticipate that native English speakers will have high proficiency and therefore will score highly on this task.  
•	For the Chimeric Face Task, participants’ data will be excluded if their accuracy on trials with identical emotions in each hemiface is below 75%. This is to ensure that participants can sufficiently judge emotion expression.  
•	For the Dichotic Listening task, participants’ data will be excluded if their accuracy on trials with identical sounds played to each ear is below 75%.  
•	In the dichotic listening task we will screen for headphone use using an adapted version of a task reported by Woods, Siegle, Traer, and McDermott (2017) where participants judge which of three tones is the softest. If accuracy is lower than 5 out of 6, we will remove participants as recommended by Woods et al.

### Missing data

Excluded participants will be replaced, so that a total of 300 complete datasets are collected.

## Data Analysis
### Statistical models

For the reliability assessment, we plotted the distributions of laterality indices and checked them for normality using the Shapiro-Wilk test. For normally-distributed scores, we used Pearson correlation coefficients for time 1 vs time 2, after excluding outliers. In the pre-registration we suggestd we might transform data to achieve normality if necessary, or alternatively use a Spearman correlation.

Optimising the test battery: If reliability turns out to be very high (above .75) with the existing test length, we will reanalyse the data using just the first N trials, reducing N until reliability becomes unsatisfactory (below r=0.65). This will allow us to identify the most efficient way to administer these tasks in future, while retaining good reliability. However, the full data (including all trials) will be used for all subsequent analyses.
<!---Adam: we have not done this yet, but it should not be too hard--->

To test Hypothesis 1 (laterality bias at a population level) one-sample t-tests will be applied for each task and session.

Hypothesis 2 predicts that left-handers will be more likely to show atypical laterality than right-handers. The test of this hypothesis will be complicated by having to decide how to define 'atypical'. Ideally, these would be individuals who have significant lateralisation in the opposite direction to the majority, as indicated by p < .05 on the individual z-score or t-score. We anticipated, however, that this criterion may select very few cases on some of our tasks, in which case we will adjust the cutoff on that task until we have at least 20 atypical cases. Once typicality is defined, we will first analyse atypicality on the Rhyme Decision and Chimeric Faces task. Participants will be categorised into four groups: typical on both; atypical on Rhyme Decision only; atypical on Chimeric Faces only, and atypical on both. A chi square test will be used to test for association with the number of people in each of the four groups and the handedness classification from the Edinburgh Handedness Inventory. This will then be repeated for the Dichotic Listening and the Chimeric Faces tasks. <!---to do--->

Hypothesis 3 concerns the relationships between the laterality measures. We planned to use  structural equation modelling (SEM) to compare the covariances of the laterality statistics between the two language tasks and chimeric faces to test between three models. This would be done separately for left- and right-handers. Before doing this we would flip the polarity of chimeric faces, so that all predictions are for positive correlations.

There were no specific predictions for the finger tapping task; this was included as an exploratory measure, which may allow us to better characterise handedness on a dimension.

## Inference criteria

Assessment of test-retest reliability will use the 95% confidence interval around the estimate, as documented above.  
Assessment of whether a function is lateralised will be tested for a difference from zero using a conventional .05 significance level.
The association between handedness and atypical laterality will be tested using a Bonferroni corrected .025 significance level.
Model fit will be evaluated using AIC model selection using Akaike weights (Wagenmakers & Farrell, 2004). Akaike weights are transformations of the AIC values, which can be directly interpreted as conditional probabilities for each model. The best fitting model to the data for right- and left-handers will be selected as the final model.

## Exploratory analysis

As noted above, we will use the z- or t-statistics to index laterality, but we will have a large dataset that might be used subsequently to explore different approaches to measuring a laterality index, and to consider how choice of measure affects test-retest reliability.

While the role of the LexTALE in the current research is largely for the purpose of data exclusion, we plan to explore correlations between LexTALE score and the laterality index on the Rhyme Decision task. This will enable us to examine whether vocabulary knowledge is related to rhyme judgement when information is displayed in each hemifield. We will similarly explore the influence of sight dominance, handedness/footedness, and the Finger Tapping Task on the laterality index for rhyme decision.

Finally, we will investigate whether the method for measuring handedness has an impact on the associations between handedness and laterality on the Rhyme Decision and Chimeric Faces tests. In our main planned analyses, handedness will be coded using a binary categorisation of the Edinburgh Handedness Inventory scores, where > 0 indicates right handedness and < 0 indicates left handedness. There are many alternative ways of coding handedness: e.g. using three categories (right handed, ambidextrous and left handed); as a continuous measure; using self-report handedness; or using performance on the Finger Tapping test as an objective, continuous measure. We will explore which of these coding systems gives the strongest associations with laterality indices on the Rhyme Decision and Chimeric Faces. To do this we will select half the data at random to explore associations, and then aim to replicate on the second half.

## Results

```{r setup, echo=FALSE,include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
library(MASS)
library(lavaan)
library(qpcR) #?? why do we need this?!
library(tidyverse)
library(semPlot)
library(RVAideMemoire) #for spearman.ci
library(ggplot2)
library(kableExtra)
library(corrr)
options(knitr.kable.NA = '') #should suppress NA in tables

```
<!---Notes to Adam from Dorothy (7th June 2020).
I've deposited the LIs file on OSF and now am reading direct from that.
I want to double check polarity of LIs - not sure they are all consistently in same direction.  
I turned the basic plotting/correlation etc into a function so it can be applied to all tasks.  
I added a vertical line to density plots at zero.
Have added a new column handed3 which distinguishes between extreme left handers and others. This may need tweaking in terms of the colours used to code and the terms,
Added a one-sample t-test to check if significantly lateralised - but then found you had done that later!
No longer removing those without complete datasets
I will try to condense results so they can be reported more succinctly--->

```{r read_dat,echo=FALSE,include=TRUE}
# read in data from csv
LIs <- read.csv('https://osf.io/x5jd2/download',stringsAsFactors = F)
#LIs <- read.csv("LIs.csv", na.strings = c("NA", ""))
# then we exclude those who we do not have full datasets for 
# This step currently omitted: use all available data for each task
# LIs <- subset(LIs, exclude== 0)

#Add a category of handedness that distinguishes extreme left from other left-handers
#In preregistration, extreme left defined as EHI laterality -90 or less.
LIs$handed3 <- LIs$handedness
w<-which(LIs$index_EHI<(-89))
LIs$handed3[w]<-'Extreme.left'

```

```{r make.combotable,echo=FALSE,include=TRUE}
#Make a table to summarise results for all 4 tasks
#This will gradually get populated as we go through each task

#Cells for each task are filled in course of plots.function chunk

combotable <- data.frame(matrix(NA,nrow=7,ncol=9))
colnames(combotable)<-c('.','RhymeRT1','RhymeRT2','DichoticAcc1','DichoticAcc2','ChimericAcc1','ChimericAcc2','FingerAcc1','FingerAcc2')
combotable[,1]<-c('N','mean','SD','p-lateralised','p-normal','test-retest r','CI r')
```

```{r plots.function, echo=FALSE,include=TRUE}
# Adam's code here turned into generic function
# While we have the relevant columns specified here, I have added steps to populate combotable as we go.
# Also saving the scatterplots as png
densplots <- function(t1,t2,name1,name2,scattername,combotable,combocolumn){
  dens1<-ggplot(LIs, aes(x=t1))+geom_density(linetype="dashed")+theme_classic()+ggtitle(name1)
  dens1 <- dens1+ geom_vline(xintercept = 0)
# normality
shap1<-shapiro.test(t1)
ttest1 <- t.test(t1) #Check if significantly lateralised
Nt1 <- length(t1)-length(which(is.na(t1)))
combotable[1,combocolumn]<-Nt1
combotable[2,combocolumn]<-round(mean(t1,na.rm=T),2)
combotable[3,combocolumn]<-round(sd(t1,na.rm=T),2)
combotable[4,combocolumn]<-round(ttest1$p.value,3)
combotable[5,combocolumn]<-round(shap1$p.value,3)

# day 2
# plot denisty
dens2<-ggplot(LIs, aes(x=t2))+geom_density(linetype="dashed")+theme_classic()+ggtitle(name2)
dens2 <- dens2+ geom_vline(xintercept = 0)
# normality
shap2<-shapiro.test(t2)
ttest2 <- t.test(t2)
Nt2 <- length(t2)-length(which(is.na(t2)))
myN<-min(Nt1,Nt2)
combotable[2,(combocolumn+1)]<-round(mean(t2,na.rm=T),2)
combotable[3,(combocolumn+1)]<-round(sd(t2,na.rm=T),2)
combotable[4,(combocolumn+1)]<-round(ttest2$p.value,3)
combotable[5,(combocolumn+1)]<-round(shap2$p.value,3)
# conduct Spearman's wth bootstrapped CI
myspear<-spearman.ci(t1, t2, nrep = 10000, conf.level = 0.95)

combotable[6,combocolumn]<-round(myspear$estimate,2)
myci <- paste0(round(myspear$conf.int[1],2),' - ',round(myspear$conf.int[2],2))
combotable[7,combocolumn]<-myci
# scatterplot time1 x time2
# position for label set to 50%x and 90%y
xpos <- max(t1,na.rm=T)*.5
ypos <- max(t2,na.rm=T)*.9

mycor<-cor.test(t1, t2, method = "spearman", alternative = "greater")$estimate

#Make a label to show the correlation with 95%CI and N on the plot
mycorlabel<-paste0('r = ',round(mycor,2),'\n(95% CI: ',round(myspear$conf.int[1],2),'-',round(myspear$conf.int[2],2),')\nN = ',myN)
#Find position in top L corner for legend
legposx <- .2 #not working - needs tweaking!
legposy <- .85
myscatter <- ggplot(LIs, aes(x= t1, y= t2)) +
  geom_hline(yintercept=0, color = "black") +
  geom_vline(xintercept=0, color = "black") +
  geom_point(aes(color= handed3)) + 
  geom_rug(aes(color= handed3)) +
  theme_classic() +
  ylab(name2) + 
  xlab(name1) +
  annotate("text", x = xpos, y = ypos, label = mycorlabel) +
  theme(legend.position = c(legposx,legposy))
# Set relative size of marginal plots (main plot 10x bigger than marginals)
bigplot<-ggExtra::ggMarginal(myscatter, groupColour = TRUE, groupFill = TRUE)
scattername<-paste0(scattername,".png")
ggsave(scattername,plot=bigplot,dpi = 300)

return(combotable)
}
```
<!--- We start by completing combotable for the 4 main measures. NB need to check polarity of tasks - also oddity re RDT distributions needs checking, and we said we'd do log transform , not sure we did--->
```{r doplots.RDT_RT,echo=FALSE,include=TRUE}
t1<-LIs$Day1.RDT_RT_LI
t2<-LIs$Day2.RDT_RT_LI
name1<-"Rhyme decision LI (day 1)"
name2<-"Rhyme decision LI (day 2)"
scattername<-"RhymeDecisionScatter"
combocolumn<-2 #column to write day 1 summary for this task in combotable
combotable<-densplots(t1,t2,name1,name2,scattername, combotable,combocolumn)

```

```{r doplots.dichotic_Accscore,echo=FALSE,include=TRUE}
t1<-LIs$Day1_DL_acc_LI
t2<-LIs$Day2_DL_acc_LI
name1<-"Dichotic LI acc (day 1)"
name2<-"Dichotic LI acc (day 2)"
scattername<-"DichoticScatter"
combocolumn<-4 #column to write day 1 summary for this task in combotable
combotable<-densplots(t1,t2,name1,name2,scattername, combotable,combocolumn)

```

```{r doplots.chimeric_Acc,echo=FALSE,include=TRUE}
t1<-LIs$Day1_CF_acc_LI
t2<-LIs$Day2_CF_acc_LI
name1<-"Chimeric LI acc (day 1)"
name2<-"Chimeric LI acc (day 2)"
scattername<-"ChimericScatter"
combocolumn<-6 #column to write day 1 summary for this task in combotable
combotable<-densplots(t1,t2,name1,name2,scattername, combotable,combocolumn)

```

```{r doplots.fingerLI,echo=FALSE,include=TRUE}
t1<-LIs$Day1_finger_LI
t2<-LIs$Day2_finger_LI
name1<-"Finger LI (day 1)"
name2<-"Finger LI (day 2)"
scattername<-"FingerScatter"
combocolumn<-8 #column to write day 1 summary for this task in combotable
combotable<-densplots(t1,t2,name1,name2,scattername, combotable,combocolumn)

```
### Table 1

```{r Table1,echo=FALSE,include=TRUE}
#Needs more formatting. For word output may need to use flextable. See https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html
combotable%>%
  kable() %>%
  kable_styling("condensed")

```
### Question 1
We first consider question 1: Is it possible to achieve test-retest reliability of .65 or more with laterality indices derived from the four laterality tests?  If so, could this be achieved more efficiently by using fewer trials in each test?

Table 1 shows results for the Laterality Indices (LIs) for the four tasks at time 1 and time 2 in terms of significance of lateralisation ('p-lateralised', i.e. comparison with distribution with mean of zero), significance of departure from normality on Shapiro-Wilks test ('p-normal'), plus the Spearman correlation between time 1 and time 2 (test-retest reliability).

Figures 1-4 show the scatterplots for time 1 and time 2 LIs for all four tasks, with left-handers colour-coded, and extreme left-handers marked in black.

![Figure 1](RhymeDecisionScatter.png)

![Figure 2](DichoticScatter.png)
![Figure 3](ChimericScatter.png)
![Figure 4](FingerScatter.png)

These analyses show good reliability for Dichotic Listening, Chimeric Faces and Finger Tapping. These tasks are also significantly lateralised in the predicted direction.  The Rhyme Detection Task is unsatisfactory: the test-retest correlation is below our specified cutoff r= .65 for both RT and accuracy (not shown here). For RT, there is significant lateralisation, but in absolute terms it is small, with a high proportion LVF-lateralised.  <!---Adam: you mentioned laterality was opposite to prediction - I think it is in the right direction, but we need to check: for RT a big score means slower performance and I am not sure if I adjusted adequately for that. Need anyhow to do with log RTs since that is what we preregistered - I don't think it will make much difference--->

Note also that we fell short of 300 participants for this task. This is because many participants were dropped because they did not meet our inclusion criteria - e.g. low accuracy.  
We had previously modified this task in an attempt to improve its reliability, but these analyses question whether it is useful, so we decided not to test further participants.

*Still to be done: consider reliabilities with fewer trials. How short a test is feasible without loss of reliability?*

### Question 2
Laterality indices can be calculated in multiple different ways (e.g. using RT, accuracy or a combination of the two; as a raw difference score or a proportion). We will explore these options (described in more detail in the ‘Indices’ section) to determine which metric shows the strongest test-retest reliability.

*Still to be done - we have computed the z-score and t-test versions discussed in introduction *

### Question 3
Do extreme left-handers show a higher rate of atypical laterality than right-handers on the Rhyme Decision, Dichotic listening, or Chimeric Faces tasks?

For our preliminary analysis, we define 'atypical laterality' as laterality that is in the opposite direction to the majority, and which has an absolute z-score or t-score value greater than 1.96. These z and t-scores indicate whether there is significant bias to the left or right within an individual.

According to the pre-registration, we omit non-extreme left-handers, but we show their data here for completeness. We also specified in the preregistration that we would analyse performance for pairs of tasks, but the very weak lateralisation of the Rhyme Decision task made that a questionable analysis. Tables 2 to 4 show associations between handedness categories and laterality classification for individual tasks first. We did not specify how we would handle data from the two sessions: these provide an opportunity for exploratory analyses from session 1 to then be confirmed using session 2 data, so for each task we start just with session 1.

```{r makecontingencies,echo=FALSE,include=TRUE}
#We start by using z-scores to make a 3-way categorisation into:
#1 sig bias in typical direction; 2: NS bias 3: in atypical direction with sig bias.
#
LIs$rhyme1.cut <- cut(LIs$Day1.RDT_p.corr_LI, 
                       breaks = c(-Inf, -1.96, 1.96, Inf), 
                       labels = c("TypLat", "Nonlat", "AtypLat"), 
                       right = FALSE)
LIs$dichotic1.cut <- cut(LIs$Day1_Dich_acc_Z, 
                       breaks = c(-Inf, -1.96, 1.96, Inf), 
                       labels = c("TypLat", "Nonlat", "AtypLat"), 
                       right = FALSE)
LIs$chimeric1.cut <- cut(LIs$Day1_CF_acc_Z, 
                       breaks =  c(-Inf, -1.96, 1.96, Inf), 
                       labels = c("AtypLat", "Nonlat", "TypLat"), 
                       right = FALSE)
mytabr <- table(LIs$rhyme1.cut)
mytabd <- table(LIs$dichotic1.cut)
mytabc <- table(LIs$chimeric1.cut)
mytabc<-mytabc[c(3,2,1)]
mytab2 <-rbind(mytabr,mytabd,mytabc)
row.names(mytab2)<-c('Rhyme decision','Dichotic','Chimeric Faces')
mytab2<-data.frame(mytab2)

mytab2%>%
  kable() %>%
  kable_styling("condensed")

tabr <- table(LIs$handed3,LIs$rhyme1.cut)
chisq.test(tabr)

tabd <- table(LIs$handed3,LIs$dichotic1.cut)
chisq.test(tabd)

tabc <- table(LIs$handed3,LIs$chimeric1.cut)
chisq.test(tabc)

#Planned test omitted group 'Left' and column Nonlat.
chi.r<-chisq.test(tabr[c(1,3),c(1,3)])

chi.d<-chisq.test(tabd[c(1,3),c(1,3)])
chi.c<-chisq.test(tabc[c(1,3),c(1,3)])
```

Table 2: Rhyme judgement
```{r showtabr,echo=FALSE,include=TRUE}
tabr%>%
  kable() %>%
  kable_styling("condensed")
```
Table 3: Dichotic listening
```{r showtabd,echo=FALSE,include=TRUE}
tabd%>%
  kable() %>%
  kable_styling("condensed")
```
Table 4: Chimeric faces
```{r showtabc,echo=FALSE,include=TRUE}
tabc%>%
  kable() %>%
  kable_styling("condensed")
```
It is evident that the Rhyme Judgement task has few individuals who are significantly lateralised in either direction. We will therefore not consider it further, as our analyses are only justified if the task is an indicator of cerebral lateralisation. Both Dichotic Listening and Chimeric Faces, on the other hand, are strongly lateralised.  The planned analyses, just comparing typically lateralised with atypically lateralised individuals, and extreme left vs right-handed individuals, show significant association between extreme left-handedness and atypical laterality on chi square test. For dichotic listening, chi squared = `r round(chi.d$statistic,2)`, d.f. = 1, p = `r round(chi.d$p.value,3)`; for chimeric faces, chi squared = `r round(chi.c$statistic,2)`, d.f. = 1, p = `r round(chi.c$p.value,3)`.

With all participants included, on the dichotic listening task, `r round(100*tabd[1,3]/sum(tabd[,3]),1)`% of those with atypical laterality were extreme left-handers, `r round(100*tabd[3,3]/sum(tabd[,3]),1)`% were right-handers, and the remainder less extreme left-handers. For those with typical lateralisation, the corresponding figures were `r round(100*tabd[1,1]/sum(tabd[,1]),1)`% extreme left-handers, `r round(100*tabd[3,1]/sum(tabd[,1]),1)`% right-handers, and the remainder less extreme left-handers.

<!-- NB v confusing because the table gives the Atypical/Typical in different order for the chimeric faces-->
For chimeric faces, the association was even more extreme.  `r round(100*tabc[1,1]/sum(tabc[,1]),1)`% of those with Atypical laterality were extreme left-handers, `r round(100*tabc[3,1]/sum(tabc[,1]),1)`% were right-handers, and the remainder less extreme left-handers. For those with typical lateralisation, `r round(100*tabc[1,3]/sum(tabc[,3]),1)`% were extreme left-handers, `r round(100*tabc[3,3]/sum(tabc[,3]),1)`% were right-handers, and the remainder less extreme left-handers.

```{r contingency2tasks,echo=FALSE,include=TRUE}
tab2task <- table(LIs$dichotic1.cut,LIs$chimeric1.cut)
chi2task<-chisq.test(tab2task)
w1<-which(LIs$chimeric1.cut=='AtypLat')
w2<-which(LIs$dichotic1.cut=='AtypLat')
w3<-intersect(w1,w2)
LIs$handed3[w3]
```

The planned analysis looking at cases who were atypical on both dichotic listening and chimeric faces was hampered by the fact that the two tasks were not associated (see below), and consequently there were only four cases in this category. Nevertheless, it was noteworthy that two of these were extreme left-handers and the other two were less extreme left-handers.

This brings us to the final set of predictions regarding associations between tests.

## Question 4
Our final question was whether the laterality indices from the language and faces laterality tasks correlate? Specifically we had predicted that in right-handers the two language tasks would correlate together, but the chimeric faces would form a separate factor, but in left-handers, the language tasks might fractionate.

This analysis was handicapped by the fact that one of the language indices, the rhyme judgement task, lacked both reliability and validity. Therefore we present the planned analysis in supplementary material, just for completeness. Here we consider the intercorrelations between the reliable tasks, including finger tapping instead of rhyme judgement. This is motivated by the prior observations that on both dichotic listening and chimeric faces, there was some association between handedness and lateralised task performance.

We had planned to use a structural equation modelling framework to compare the covariances of the laterality statistics between the two language tasks and chimeric faces to test between three models, separately for left- and right-handers. However, in neither handedness group were there any meaningful correlations between laterality indices.

```{r prep_dat,echo=FALSE,include=TRUE}
# now create subset of data frame for analysis
mycols <- c('Day1.RDT_RT_LI', 'Day2.RDT_RT_LI',
            'Day1_CF_acc_LI', 'Day2_CF_acc_LI',
            'Day1_DL_acc_LI', 'Day2_DL_acc_LI',
            'Day1_finger_LI', 'Day2_finger_LI',
            'handedness')
colnums<- which(colnames(LIs) %in% mycols)
M <- LIs[,colnums]
M <- M %>% 
  rename(
    "rhyme1" = Day1.RDT_RT_LI,
    "rhyme2" = Day2.RDT_RT_LI,
    "chimeric1" = Day1_CF_acc_LI,
    "chimeric2" = Day2_CF_acc_LI,
    "dichotic1" = Day1_DL_acc_LI,
    "dichotic2" = Day2_DL_acc_LI,
    "finger1" = Day1_finger_LI,
    "finger2" = Day2_finger_LI)
# subset into right and left handers
left_dat <- subset(M, handedness== "Left")
right_dat <- subset(M, handedness== "Right")
```

```{r showcorrs,echo=FALSE,include=TRUE}
#Will focus just on Day1
#Had considered showing as correlation heatmap - but values so small it wouldn't be worth doing!

#http://www.sthda.com/english/wiki/ggplot2-quick-correlation-matrix-heatmap-r-software-and-data-visualization

corrL<-round(cor(left_dat[,c(2,4,6,8)],use='complete.obs'),2)
corrR<-round(cor(right_dat[,c(2,4,6,8)],use='complete.obs'),2)

```

### Table 5: Intratask correlations for Left handers
```{r dotab5,echo=FALSE,include=TRUE}
corrL%>%
  kable() %>%
  kable_styling("condensed")
```

### Table 6: Intratask correlations for Right handers
```{r dotab6,echo=FALSE,include=TRUE}
corrR%>%
  kable() %>%
  kable_styling("condensed")
```
<!---I have run out of steam at this point. I think some of the correlations in *left-handers* might be statistically significant, though very small. But the only ones that look at all nonzero are with rhyme! Will just redo for time 2 to see if this is robust-->

```{r corrtime2,echo=FALSE,include=TRUE}
corrL2<-round(cor(left_dat[,c(3,5,7,9)],use='complete.obs'),2)
corrR2<-round(cor(right_dat[,c(3,5,7,9)],use='complete.obs'),2)

```

The correlations between tasks are all very low. To check if this pattern is robust, this analysis was repeated for time 2 data.

### Table 7: Intratask correlations time 2, for Left handers
```{r dotab7,echo=FALSE,include=TRUE}
corrL%>%
  kable() %>%
  kable_styling("condensed")
```

### Table 8: Intratask correlations time 2 ,for Right handers
```{r dotab8,echo=FALSE,include=TRUE}
corrR%>%
  kable() %>%
  kable_styling("condensed")

## SEM model 
Here we model left hander data with three prespecified covariance structures, then compare using AIC weights.

This is a subset of SEM that dos not include any latent variables or directional paths. It only looks to constrain particular covariance patterns and report 'best' model according AIC weights.

```{r Roger_mod_left,echo=FALSE,include=TRUE}
model1<-"
dichotic1~~0*rhyme1
dichotic1~~0*chimeric1 
rhyme1~~0*chimeric1 
"

model2<-"
dichotic1~~rhyme1 
dichotic1~~0*chimeric1
rhyme1~~0*chimeric1 
"

model3<-"
dichotic1~~rhyme1 
dichotic1~~chimeric1 
rhyme1~~chimeric1
"

fit1A <- cfa(model1, data=left_dat)
fit2A <- cfa(model2, data=left_dat)
fit3A <- cfa(model3, data=left_dat)

aic_vector1 <- c(fitMeasures(fit1A, "aic"),fitMeasures(fit2A, "aic"),fitMeasures(fit3A, "aic"))

summary(fit1A, fit.measures=TRUE)
summary(fit2A, fit.measures=TRUE)
summary(fit3A, fit.measures=TRUE)

AICweights1<-akaike.weights(aic_vector1)$weights

IC_objectweights1<-data.frame(Model=factor(rep(c('Model 1','Model 2','Model 3'))), IC = factor(rep('AIC',each=3)),values=AICweights1)

ggplot(IC_objectweights1,aes(Model,values,fill=Model))+
  geom_bar(stat='identity',position="dodge")+
  ylab('Akaike weights') + theme_classic()
```

```{r Roger_mod_right,echo=FALSE,include=TRUE}
#We now repeat this for right handers. 
fit1B <- cfa(model1, data=right_dat)
fit2B <- cfa(model2, data=right_dat)
fit3B <- cfa(model3, data=right_dat)

aic_vector2 <- c(fitMeasures(fit1B, "aic"),fitMeasures(fit2B, "aic"),fitMeasures(fit3B, "aic"))

summary(fit1B, fit.measures=TRUE)
summary(fit2B, fit.measures=TRUE)
summary(fit3B, fit.measures=TRUE)

AICweights2<-akaike.weights(aic_vector2)$weights

IC_objectweights2<-data.frame(Model=factor(rep(c('Model 1','Model 2','Model 3'))), IC = factor(rep('AIC',each=3)),values=AICweights2)

ggplot(IC_objectweights2,aes(Model,values,fill=Model)) +
  geom_bar(stat='identity',position="dodge")+
  ylab('Akaike weights') + theme_classic()
```

## Other analyses
<!---Original text: Upon an initial look, there is a reasonable correlation between the EHI and day 1 on the finger tapping. However, it does look like there are two clusters. So, we take a look at the left and right handers seperately. Doing so yields dissappointing results.-->
<!---DB: Not sure I understood this - is the point that strength of handedness on EHI is not related to finger tapping within handedness groups? --->

```{r H1,echo=FALSE,include=TRUE }
# Nice plots of distributions for each LI but I don't think we need these if we have the scatterplots, as they show the same information. T-tests done above.
#I've preserved the script as we may want to use it, but put it in a loop so it is optional for now.
doH1 <-0
if(doH1 ==1){
mycols <- c('Day1.RDT_RT_LI','Day1_DL_acc_LI','Day1_CF_acc_LI',  'Day1_finger_LI', "handedness")
colnums<- which(colnames(LIs) %in% mycols)
M <- LIs[,colnums]
N <- M %>% 
  rename(
    "CF" = Day1_CF_acc_LI,
    "DL" = Day1_DL_acc_LI,
    "RD" = Day1.RDT_RT_LI,
    "FT" = Day1_finger_LI)
# PLOT TO SEE IF DIFFERENT FROM ZERO
N$ID <- 1:nrow(N)
# format to long
data_long <- tidyr::gather(N, task, index, `CF`:`FT`, factor_key=TRUE)
# plot data
a <- ggplot(subset(data_long, task== "RD"), aes(x = task, y = index)) +
  geom_hline(yintercept=0, linetype="dashed", color = "red") +
  ggpirate::geom_pirate(aes(colour = "black"), bars = FALSE, 
                        points_params = list(shape = 19, alpha = 0.5), lines_params = list(size = 0.8)) +
  ylab("Laterality Index") + xlab("") + ylim(-30, 30) +
  scale_colour_manual(values = "darkgray") + theme_classic() 
b <- ggplot(subset(data_long, task== "DL"), aes(x = task, y = index)) +
  geom_hline(yintercept=0, linetype="dashed", color = "red") +
  ggpirate::geom_pirate(aes(colour = "black"), bars = FALSE, 
                        points_params = list(shape = 19, alpha = 0.5), lines_params = list(size = 0.8)) +
  ylab("Laterality Index") + xlab("") + ylim(-100, 100) +
  scale_colour_manual(values = "darkgray") + theme_classic() 
c <- ggplot(subset(data_long, task== "CF"), aes(x = task, y = index)) +
  geom_hline(yintercept=0, linetype="dashed", color = "red") +
  ggpirate::geom_pirate(aes(colour = "black"), bars = FALSE, 
                        points_params = list(shape = 19, alpha = 0.5), lines_params = list(size = 0.8)) +
  ylab("Laterality Index") + xlab("") + ylim(-100, 100) +
  scale_colour_manual(values = "darkgray") + theme_classic() 
d <- ggplot(subset(data_long, task== "FT"), aes(x = task, y = index)) +
  geom_hline(yintercept=0, linetype="dashed", color = "red") +
  ggpirate::geom_pirate(aes(colour = "black"), bars = FALSE, 
                        points_params = list(shape = 19, alpha = 0.5), lines_params = list(size = 0.8)) +
  ylab("Laterality Index") + xlab("") + ylim(-30, 30) +
  scale_colour_manual(values = "darkgray") + theme_classic() 
ggpubr::ggarrange(a, b, c, d)
t.test(LIs$Day1_CF_acc_LI, mu= 0)
t.test(LIs$Day1_DL_acc_LI, mu= 0)
t.test(LIs$Day1_finger_LI, mu= 0)
t.test(LIs$Day1.RDT_RT_LI, mu= 0)
}
```

<!---**Finger tapping and EHI** - commented out for now as focus is on planned analyses-->

```{r fingerEHI,echo=FALSE,include=TRUE }
# plot
FingersEHI <-0
if(FingersEHI==1){
finger_scatter <- ggplot(LIs, aes(x= index_EHI, y= Day1_finger_LI)) +
  geom_hline(yintercept=0, color = "black") +
  geom_vline(xintercept=0, color = "black") +
  geom_point(aes(color= handedness)) +
  geom_rug(aes(color= handedness)) +
  theme_classic() +
  ylab("Finger tapping LI (day 1)") + 
  xlab("EHI") +
  annotate("text", x = 90, y = 25, label = "paste(italic(R), \" = .58\")", parse = TRUE) +
  theme(legend.position = c(.15, .90))
# Set relative size of marginal plots (main plot 10x bigger than marginals)
ggExtra::ggMarginal(finger_scatter, groupColour = TRUE, groupFill = TRUE)
# spearman's CIs
spearman.ci(LIs$Day1_finger_LI, LIs$index_EHI, nrep = 10000, conf.level = 0.95)
# spearman's test
cor.test(LIs$Day1_finger_LI, LIs$index_EHI, method = "spearman", alternative = "greater")
# LEFT HANDERS
left <- subset(LIs, handedness== "Left")
# spearman's test
cor.test(left$Day1_finger_LI, left$index_EHI, method = "spearman", alternative = "greater")
# Right HANDERS
right <- subset(LIs, handedness== "Right")
# spearman's test
cor.test(right$Day1_finger_LI, right$index_EHI, method = "spearman", alternative = "greater")
}
```

### 2020-06-08 draft_notebook_writeup.Rmd
minor changes to echo and include commands so we can see tables in html version

@@
 
-```{r Table1,echo=FALSE,include=TRUE}
+```{r Table1,echo=FALSE}
 #Needs more formatting. For word output may need to use flextable. See https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html
@@
 
-```{r makecontingencies,echo=FALSE,include=TRUE}
+```{r makecontingencies,echo=FALSE}
 #We start by using z-scores to make a 3-way categorisation into:
@@
 Table 2: Rhyme judgement
-```{r showtabr,echo=FALSE,include=TRUE}
+```{r showtabr,echo=FALSE}
 tabr%>%
@@
 Table 3: Dichotic listening
-```{r showtabd,echo=FALSE,include=TRUE}
+```{r showtabd}
 tabd%>%
@@
   kable_styling("condensed")
+
 ```
+
 Table 4: Chimeric faces
-```{r showtabc,echo=FALSE,include=TRUE}
+```{r showtabc,echo=FALSE}
 tabc%>%
@@
   kable_styling("condensed")
+
 ```
@@
 
-```{r prep_dat,echo=FALSE,include=TRUE}
+```{r prep_dat,echo=FALSE}
 # now create subset of data frame for analysis
@@
 
-```{r showcorrs,echo=FALSE,include=TRUE}
+```{r showcorrs,echo=FALSE}
 #Will focus just on Day1
@@
 ### Table 5: Intratask correlations for Left handers
-```{r dotab5,echo=FALSE,include=TRUE}
+```{r dotab5,echo=FALSE}
 corrL%>%
@@
 ### Table 6: Intratask correlations for Right handers
-```{r dotab6,echo=FALSE,include=TRUE}
+```{r dotab6,echo=FALSE}
 corrR%>%
@@
 ### Table 7: Intratask correlations time 2, for Left handers
-```{r dotab7,echo=FALSE,include=TRUE}
-corrL%>%
+```{r dotab7,echo=FALSE}
+corrL2%>%
   kable() %>%
@@
 ### Table 8: Intratask correlations time 2 ,for Right handers
-```{r dotab8,echo=FALSE,include=TRUE}
-corrR%>%
+```{r dotab8,echo=FALSE}
+corrR2%>%
   kable() %>%
   kable_styling("condensed")
+```
 
 ## SEM model
+
 Here we model left hander data with three prespecified covariance structures, then compare using AIC weights.
@@
 
-```{r Roger_mod_left,echo=FALSE,include=TRUE}
+```{r Roger_mod_left,echo=FALSE}
 model1<-"
@@
 
-```{r Roger_mod_right,echo=FALSE,include=TRUE}
+```{r Roger_mod_right,echo=FALSE}
 #We now repeat this for right handers. 

### 2020-06-09 battery-preprocessing_v5_DB.Rmd
updates for logRT

@@
 date: "updated May 21st 2020 (now notebook format)"
-output: html_notebook
+output:
+  pdf_document: default
+  html_notebook: default
 ---
+
+## Updates
+
+09 June 2020: AP made changes to calcultaions of rhyme
+(1) Added t scores to output
+(2) computed LIs for log-RT
+(3) changed polariies so that a negative value was indicative of left lateralisation in the brain.
+
 23 May 2020: Created subject number for those included in final analysis.
 CF outlier detection needed updating to ensure only correct trials used for outliers - Correct had not been coded at that point in prior version
+
 22 May 2020: AP made changes to N trials needed for inclusion  
@@
 
-demo_dat <- read.csv('https://osf.io/7smj6/download',stringsAsFactor=F)
+demo_dat <- read.csv('./data/Demographics.csv',stringsAsFactor=F)
 
@@
 ```{r sightednesss, warning=FALSE}
-porta_dat <- read.csv("https://osf.io/zmqkt/download",stringsAsFactors = F)
-miles_dat <-  read.csv("https://osf.io/ekqsp//download",stringsAsFactors = F)
+porta_dat <- read.csv("./data/Porta.csv",stringsAsFactors = F)
+miles_dat <-  read.csv("./data/Miles.csv",stringsAsFactors = F)
 
@@
 
-EHI <- read.csv("https://osf.io/kfdma/download",stringsAsFactors = F)
+EHI <- read.csv("./data/EHI.csv",stringsAsFactors = F)
 
@@
 # read data
-lexTALE <- read.csv("https://osf.io/h4btp/download",stringsAsFactors = F)
+lexTALE <- read.csv("./data/LexTALE.csv",stringsAsFactors = F)
 
@@
 for (d in 1:2){
-  data_file <- "https://osf.io/mhk29/download" #day 1 data
+  data_file <- "./data/combined/day1_combined.csv" #day 1 data
 
   if (d ==2){
-    data_file <-  "https://osf.io/ebm6s/download"
+    data_file <-  "./data/combined/day2_combined.csv"
   }
@@
 # if (length(w)>0) {chim_dat2<-chim_dat2[-w,]}
-
 ```
@@
 outliertable2<-table(RDT_dat2$subject,RDT_dat2$outlier)
-
 ```
@@
 RDT_df2 <- make.df(RDT_dat2,varlist,latlist,nloop)
+
+# remove NaN
+is.nan.data.frame <- function(x)
+  do.call(cbind, lapply(x, is.nan))
+RDT_df1[is.nan(RDT_df1)] <- NA
+RDT_df2[is.nan(RDT_df2)] <- NA
+
+# add log RT
+RDT_df1$logRT <- log10(RDT_df1$RT)
+RDT_df2$logRT <- log10(RDT_df2$RT)
 
@@
 origdat <-origdat[origdat$side %in% c('left','right'),]
+
+# accuracy
 measure <- 'Accuracy'
@@
 task <- 'Rhyme Detection'
-
 dopirate(origdat,task,measure)
+
+# RT
 origdat$X1 <- origdat$RT
 measure <- 'RT'
+dopirate(origdat,task,measure)
+
+# log RT
+origdat$X1 <- origdat$logRT
+measure <- 'logRT'
 dopirate(origdat,task,measure)
@@
 {origdat <-RDT_df2[RDT_df2$subject %in% incRDT,]}
+
+# accuracy
 origdat$X1 <- origdat$p.corr
@@
 allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
+
+# RT
 origdat$X1 <- origdat$RT
@@
 allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
+
+# log-RT
+origdat$X1 <- origdat$logRT
+mycolname<-"Day1.RDT_logRT_LI"
+if(d==2){mycolname<-"Day2.RDT_logRT_LI"}
+polarity <- (-1)
+mycutoff <- 0
+allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
 }
@@
 t.test(partsum$Day1.RDT_RT_LI)
+t.test(partsum$Day1.RDT_logRT_LI)
 t.test(partsum$Day2.RDT_p.corr_LI)
 t.test(partsum$Day2.RDT_RT_LI)
-
+t.test(partsum$Day2.RDT_logRT_LI)
+
+# correlate
 cor1<-cor(partsum$Day1.RDT_p.corr_LI,partsum$Day2.RDT_p.corr_LI,use='complete.obs')
 cor2<-cor(partsum$Day1.RDT_RT_LI,partsum$Day2.RDT_RT_LI,use='complete.obs')
+cor3<-cor(partsum$Day1.RDT_logRT_LI,partsum$Day2.RDT_logRT_LI,use='complete.obs')
+# acc plot
 plot(partsum$Day1.RDT_p.corr_LI,partsum$Day2.RDT_p.corr_LI,main='Rhyme detect: LI for accuracy (%correct)')
@@
 text(-3,3,paste0('r = ',round(cor1,3)))
+# raw plot
 plot(partsum$Day1.RDT_RT_LI,partsum$Day2.RDT_RT_LI,main='Rhyme detect: LI for RT')
@@
 text(-3,9,paste0('r = ',round(cor2,3)))
+# log plot
+plot(partsum$Day1.RDT_logRT_LI,partsum$Day2.RDT_logRT_LI,main='Rhyme detect: LI for log-RT')
+abline(v=0)
+abline(h=0)
+text(2,1,paste0('r = ',round(cor3,3)))
 ```
@@
 ```
+
 ### T-test as alternative to LI for RDT
@@
 }
+
+# add t score to the data
+t_dat <- select(pairedt, -3, -5)
+# merge to allsum for analysis
+allsum <- merge(allsum, t_dat)
+
 cor(pairedt$t1,pairedt$t2,use='complete.obs')
@@
 mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
-polarity<- (1)#so LI is positive in predicted direction
+polarity<- (-1)#so LI is positive in predicted direction
 allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
@@
 mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
-polarity<- (1)#so LI is positive in predicted direction
+polarity<- (-1)#so LI is positive in predicted direction
 allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)

### 2020-06-11 battery-preprocessing_v5_DB.Rmd
osf links updated

@@
   # new data
-  wanted <- c("Trial","subject", "Correct", "RT", "ANSWER","Response","word","Left_image","Right_image") # select wanted columns
+  wanted <- c("Trial","subject", "Participant.Public.ID", "Correct", "RT", "ANSWER","Response","word","Left_image","Right_image") # select wanted columns
   RDT_dat2 <- procdata(origfile,wanted,task,disp) #Here we reuse the procdata function that we developed for processing Chimeric faces data - just does some generic processing
@@
   thisdat <- thisdat[thisdat$outlier==0,] #remove outliers
-  aggdata <- aggregate(thisdat$RT,by=list(thisdat$side,thisdat$item,thisdat$subject),
+  aggdata <- aggregate(thisdat$RT,by=list(thisdat$side,thisdat$item,thisdat$Participant.Public.ID),
                        FUN=mean,na.rm=TRUE)
+  aggdata$subject <- aggdata$Participant.Public.ID
   colnames(aggdata)<-c('side','item','subject','RT')
@@
 t_dat <- select(pairedt, -3, -5)
+t_dat <- na.omit(t_dat)
 # merge to allsum for analysis
-allsum <- merge(allsum, t_dat)
+allsum <- merge(allsum, t_dat, by.x = "origID", by.y = "subject", all.x=T)
 

### 2020-06-25 draft_notebook_writeup.Rmd
intro and method included. Next to write results.

@@
 ---
-title: "Assessing reliability of an online behavioural laterality battery: a pre-registered study (draft notebook)"
-output: html_notebook
+title: 'Assessing reliability of an online behavioural laterality battery: a pre-registered
+  study.'
+output:
+  html_notebook: default
+  pdf_document: default
+  word_document: default
 ---
@@
 ## Authors
-Adam J Parker, Zoe Woodhead, Paul Andrew Thompson, and Dorothy Bishop
-
-We report here the results of analyses that were preregistered on Open Science Framework.
-The analyses here are based on the preregistration that was revised after piloting of the battery.
-On that basis, we made some modifications to the tasks and added a consonant-vowel Dichotic Listening task.
-We also adapted the analysis plan to include a structural equation modeling to look at dissociations between laterality on different tasks.
-The preregistration can be found <a href="https://osf.io/qe6b4"> here </a> .
+Adam J. Parker, Zoe V. J., Woodhead, Paul A. Thompson, and Dorothy V. M. Bishop
+
+Department of Experimental Psychology, University of Oxford, Anna Watts Building, Radcliffe Observatory Quarter, Woodstock Rd, Oxford OX2 6GG
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+## Abstract
+
+**Needs filling in**
 
 ## Introduction
-Hemispheric specialisation (lateralisation) is a ubiquitous characteristic of the human brain, and yet many questions remain unanswered as to its causes and consequences.
-Much of the ambiguity in the literature may arise from unreliable measures of lateralisation and underpowered studies.
-This study aims to develop a reliable online battery to assess lateralisation, which could be used to test large numbers of participants.
-
-The new battery consists of four behavioural laterality tests, which will be described in more detail in the Methods section:   
-1.	Rhyme Decision: A visual half-field task in which participants decide whether a central written word rhymes with the name of a laterally-presented picture.
-This measures laterality of language processing.
-2.	Dichotic Listening: A dichotic listening task in which participants hear two consonant-vowel (CV) auditory stimuli simultaneously, one to each ear.
-They are required to report which CV stimulus out of six options (ba, da, ga, ka, pa, ta) provided the best match to the stimulus they heard .
-This measures laterality of language processing.
-3.	Chimeric Faces: A visual half-field task in which participants view stimuli that are vertically split at the midline with each hemiface displaying one of four emotions (anger, disgust, happiness, and sadness), They are required to report the overall emotion that they perceive in the face.
-This measures laterality of face and emotion processing.
-4.	Finger Tapping: A speeded sequential key-pressing task performed using a computer keyboard with left and right hands. This provides a quantitative measure of relative skill of the two hands.    
-In addition, participants completed online versions of:
-
-•	The Lexical Test for Advanced Learners of English (LexTALE; Lemhöfer & Broersma, 2012), to assess vocabulary as an indicator of language proficiency.  
-•	The Edinburgh Handedness Inventory (Oldfield, 1971), to assess manual preference.  
-•	Modified versions of the Miles Test (Miles, 1929) and the Porta Test (Porac & Coren, 1976) which are suitable for online testing, to assess ocular dominance.   
-•	A basic demographic questionnaire. This included questions about age, gender, footedness, years in education, and bilingualism.
-
-The study was primarily exploratory.
-Our principal goal was to refine the four laterality tests to maximise test-retest reliability.
-The whole battery was administered at an initial testing session, then the four laterality tests were repeated at a second session 3-10 days later.
-In addition, we aimed to test some predictions about the relationships between laterality on the four tests.
-
-The primary outcome measure from each of the four tests is a laterality index; depending on the task, this is based on difference between left and right sides in accuracy (dichotic and chimeric faces tasks), reaction time (rhyme decision), or completed sequences (finger tapping) (see ‘Indices’ section below for details).
-
-As well as establishing reliability of measures, we consider the inter-relationships between them.
-With online testing, there is the opportunity to test this using an adequately powered sample using measures of known reliability.
+
+Hemispheric specialisation, or lateralisation, is a ubiquitous characteristic of the human brain, yet many questions remain unanswered as to its causes and consequences.
+Much of the ambiguity in the published literature may arise from unreliable measures of lateralisation and underpowered studies.
+Without well powered experiments that use reliable tasks, it becomes difficult to address many research questions regarding the relationship between laterality indices and their stability over time.
+With respect to the relationship between indices, unless a task is reliable, the absence of a correlation cannot be interpreted as evidence for independence of function (Bishop, 2017).
+The best way to determine a task’s reliability is to test it twice.
+Here, the variation between participants should be larger than the variation within subjects.
+Taking this into consideration, the current work introduces an online behavioural laterality battery for which we established the test-retest reliability using a sample of 390 participants.
+The behavioural tasks included were: a novel rhyme decision task, a dichotic listening task (Hugdahl et al., 2009), a chimeric faces task (Burt & Perret, 1997), and a novel finger tapping task.
+We subsequently use this battery to we examine two pre-registered questions: (1) do extreme left-handers show a higher rate of atypical laterality than right handers on each behavioural laterality task; (2) do the laterality indices produced on each measure correlate.
+
+### Behavioural laterality tasks and their reliability.
+
+Behavioural laterality tasks have long been used to establish hemispheric dominance for certain cognitive functions.
+Rather consistently, visual half-field and dichotic listening paradigms agree that the left hemisphere plays the dominant role in language processing  (Hugdhal, 2000; Ocklenburg & Güntürkün, 2018).
+Here we review how each paradigm has been implemented to study language lateralisation.
+
+Evidence from the visual half-field paradigm (see Bourne, 2006, for a review), strongly points towards advantageous processing for letter strings presented in the right, relative to the left, visual field.
+As information in the right visual field projects contralaterally and arrives directly at the language dominant left hemisphere, it is thought to reflect an absence in callosal transfer that is required for the processing of information initially arriving at the right hemisphere.
+
+Despite wide reporting of a right visual field advantage at the population level when using visual half-field tasks, the paradigm is not without critique.
+In particular, the validity of visual half-field tasks has been questioned.
+Krach, Chen, and Hartje (2006) compared the laterality indices (LIs) produced during a visual half-field experiment with those produced using functional Transcranial Doppler Sonography (fTCD).
+The visual half-field experiment itself consisted of lexical decisions made towards abstract nouns presented in the left and right visual-fields.
+During the experiment, the blood flow to left and right hemispheres of 58 participants was measured.
+In addition, blood flow was measured during a word generation task—the gold standard task for assessing hemispheric dominance for language.
+Despite showing the expected right visual field advantage on the visual half-field task, the behavioural LIs produced on the visual half-field paradigm did not correlate with the fTCD LIs produced during word generation (r= .18) or the visual half-field task itself (r -.004).
+The
+lack of validity here likely reflects the poor reliability of the visual field task as Krach et al.
+report the split-half reliability for the correct response rates and reaction times during the behavioural visual half-field paradigm were r= .14 and r= .20 respectively.
+Thus, with such low split-half reliability it is unclear whether the task was able to accurately measure language lateralisation to start with.
+
+In recent years it has become clear that the reliability and validity of visual half-field tasks to identify the language-dominant hemisphere is largely dependent on several methodological aspects.
+For example, Hunter and Brysbaert (2008) noted stronger and more stable visual-field differences when stimuli were presented bilaterally relative to those where only one stimulus was presented either in the LVF or in the RVF (Boles, 1987, 1990, 1994; see also Iacoboni & Zaidel, 1996).
+Brysbaert and Hunter argued that this is the result of stimuli competing during bilateral presentation, which is easier when the stimulus is presented in the language dominant hemisphere.
+Reliability and validity also depend on there being a sufficient number of observations and brief stimulus presentation.
+When accounting for these methodological aspects, the validity of visual half-field tasks seems promising.
+For example, Hunter and Brysbaert (2008) reported that LIs from visual half-field paradigms correlated positively with those from functional Magnetic Resonance Imaging (fMRI): word naming (r= .63), picture naming (r= .77).
+Here 192 trials were used during word naming and 160 trials during picture naming while stimuli were bilaterally presented and masked at offset.
+Furthermore, a promising visual half-field paradigm developed by Hausmann and colleagues (Hausmann et al., 2019; Willemin et al., 2016) which adhered to several of the methodological aspects outline by Brysbaert and Hunter has reported a right visual-field advantage at the population level with little variation occurring between languages.
+The authors, however, have not yet reported reliability measures for this task.
+
+While the visual half-field technique measures lateralisation for visual stimuli, dichotic listening tasks examine hemispheric dominance for the processing of auditory stimuli.
+Like the visual half-field technique, the dichotic listening task makes use of the crossing of fibres at the corpus callosum.
+In this task, different streams of auditory linguistic stimuli (i.e.
+words, syllables) are presented to each ear and participants are either asked simply to report as many words as they can or to attend to one ear selectively.
+In either case, a right ear advantage has been demonstrated; typically, participants report more verbal stimuli presented to the right ear (Hugdahl, 2011; Kimura, 1961).
+Again, this advantage has been attributed to stimuli being projected to the language dominant hemisphere without callosal transfer.
+
+Over the years, various dichotic listening paradigms have been used (Bryden, 1988; Westerhausen, 2019).
+The test-retest reliability of these paradigms have been far from optimal (Voyer, 1998; Kelley & Littenberg, 2019).
+However, recent data from Westerhausen and Samuelsen (2020) indicated that when designed well, the dichotic listening paradigm can be reliable.
+First, based on a review of the published literature, Westerhausen and Samuelsen outlined eight design features that should be used in an optimal dichotic listening paradigm and explain their justification for each.
+The criteria are as follows: (1) stop consonant-vowel (CV) syllables as stimulus materials, (2) pair only CV stimuli from the same voicing category, (3) Alternating trials of voiced and unvoiced stimulus pairs, (4) free-recall instruction, (5) single-stimulus pair per trial, (6) paradigm length of 120 dichotic trials, (7) all stimulus pairs are presented in both orientations in identical frequency, and (8) includes binaural (homonymic) trials.
+Putting these criteria into practice, Westerhausen and Samuelsen administered a modified dichotic listening paradigm using both a verbal and manual response format to 50 right-handers.
+Interestingly, the full paradigm yielded a test-retest reliability of r= .93 for verbal responses and r= .91 for manual responses.
+Thus, the dichotic listening task provides a promising paradigm for those wanting to investigate language lateralisation without reliance on visual stimuli.
+
+So far, the studies reviewed have been interested in language lateralisation.
+That is not to say that behavioural tasks are limited to linguistic materials.
+For instance, the visual half-field paradigm has been used to measure lateralisation for the processing of facial similarity (e.g.
+Gilbert & Bakan, 1973), face emotionality (e.g.
+Levy, Heller, Banich, & Burton, 1983), and local information (Sergent, 1982).
+In a recent review of the reproducibility of visual-field asymmetries, Brederoo, Nieuwenstein, Cornelissen, and Lorist (2019) reported that they were able to replicate five out nine perceptual asymmetries using the visual half-field paradigm.
+
+Of importance to the current work is the use of the visual half-field paradigm to measure the lateralisation of face emotionality, in particular the chimeric faces task.
+In this task participants are presented with vertically split chimeric faces, where each half displays a different emotion.
+Depending on how the task is implemented, participants will either view a single chimeric face and report the emotion being expressed (e.g.
+Drebing, Federman, Edington, & Terzian, 1997; Kucharska-Pieture, & David, 2003) or view two faces, one above the other, and decide which of the two faces looks more emotive (e.g., Bourne, 2005; Levy, Heller, Banich, & Burton, 1983).
+Here the images will be identical, mirror images and typically include a neutral and emotive hemiface (i.e.
+Happiness-Neutral vs Neutral-Happiness).
+Regardless of implementation, participants’ responses point towards a bias for the processing of emotion in the left visual field.
+In the single face version, participants will more often report the emotion shown in the left hemiface.
+In the two-face version, participants show a bias in selecting the face with an emotion depicted in the left hemiface.
+
+In terms of its reliability, the chimeric faces tasks seem promising.
+In Levy et al.’s (1983) early study, they reported good split-half reliability for the two faces version of the chimeric faces task, where the left and right handers both showed a split-half reliability of .93.
+The validity of the chimeric faces task has been demonstrated using behavioural and fMRI data.
+In Yovel, Tambini, and Brandman’s (2008) study, participants performed a behavioural experiment where they had to recognize centrally presented chimeric faces, which presented different identities in the right- and left-visual-field.
+Interestingly, the behavioural measure significantly correlated with the LI calculated based on activation in the fusiform face area (r= .49).
+However, it is of importance to note the sample size used here was small, but at face value the results seem promising.
+
+### Behavioural laterality tasks and (a)typical lateralisation
+
+The studies reviewed so far have all focused on lateralisation at the population level.
+That is not to say that lateralisation does not vary between individuals.
+One factor that has been associated with the degree of lateralisation is handedness.
+Neuroimaging data strongly indicates that these differences in lateralisation are the result of a group of left-handers who are right language dominant rather than inidciateing that this relationship operates on a continuum (Mazoyer et al., 2014This increased departure from typical lateralisation in left-handers is not only present in neuroimaging studies, it is also reflected in behavioural tasks.
+Thus it will be important for our battery to replicate the findings regarding handedness and lateralisation summarised below.
+
+With regards to visual half-field techniques for studying language lateralisation, there is evidence to suggest that the visual field advantage varies with handedness (Elias, Bulman-Fleming, &McManus, 1999; Nicholls, 1994, 1998; Weekes, Capetillo-Cuncliffe, Rayman, Iacoboni, & Zaidel, 1999).
+However, a relatively well powered study, involving 100 participants conducted by Willemin et al.
+(2016) reported no effect of handedness (when coded as left and right) on the visual field advantaged during lexical decision task.
+At first, this finding may seem odd.
+However, it may reflect underlying issues with the task rather than represent strict evidence for a null effect of handedness on visual half-field task.
+Here, Willemin et al.
+asked participants to respond via manual response.
+This contrasts several studies that have reported increased variability when responding aloud.
+Thus, a task that lacks some active speech production component may not yield such strong visual field advantages that enable provide enough variation to detect subtle differences in handedness (Hunter & Brysbaert, 2008).
+This, of course, is something that we kept in mind when developing the visual half-field paradigm used in the current study.
+
+When dichotic listening paradigms to study language lateralisation, the evidence is clear that non-right-handers differ in terms of their laterality.
+While a clear right ear (i.e.
+left hemisphere) advantage is noted at the population level (see Hugdahl et al., 2009, and Westerhausen, 2019, for a review), there is evidence to suggest that this is reduced in left- relative to right-handers (Bethmann, Tempelmann, De Bleser, Scheich, & Brechmann, 2007; Bryden, 1970, 1988; Hugdahl, 2003); although handedness alone does not predict this difference (Van der Haegen, Westerhausen, Hugdahl, & Brysbaert, 2013).
+Across two studies Karlsson, Johnstone, and Carey (2019) provided a clear, yet comprehensive explanation of how performance on the dichotic listening task differed as a function of hand preference.
+Here, Karlsson et al.
+employed a consonant-vowel version of the dichotic listening task where participants reported which one of six syllables that they heard the clearest.
+For study 1, a clear right ear advantage was reported at the population level.
+Furthermore, right-handers were on average more lateralised than non-right-handers—87% of right-handers and 78% of non-right-handers showed a right ear advantage.
+For study 2, a similar pattern of results emerged.
+Again, right-handers were more lateralised with 85% of right-handers and 79% of non-right-handers showing an advantage.
+Together these results indicate that non-right-handers show a larger departure from the ‘typical’ lateralisation during dichotic listening tasks.
+
+Recall that, for the chimeric faces, there is typically a left visual field advantage at the population level— indicating right hemisphere dominance.
+As with other laterality tasks, there appears to be a greater departure from this typical pattern of laterality for left-handers (e.g.
+Gilbert & Bakan, 1973; Heller & Levy, 1983; Karlsson et al., 2019; Levy et al., 1983; Rozkowski & Snelbecker, 1982).
+Again, this departure is well described by Karlsson et al.
+(2019).
+This departure was examined for two version of the chimeric faces task.
+The first involve participants viewing a single face with a different emotion shown in either hemiface.
+Participants then reported the emotion that they felt was most strongly displayed.
+The second was a two faces version where participants viewed two a face with an emotion on one side and the mirror version.
+Here participants had to decide which of the two faces showed a target emotion.
+For the first task, right-handers showed a bias for reporting information presented in the left visual field (i.e.
+right hemisphere); however, left-handers did not show a population bias.
+Specifically, 75% of right-handers and 60% of non-right-handers showed a leftwards bias.
+For the second chimeric faces task, a similar pattern emerged.
+73% of right-handers and 57% of non-right-handers here showed a leftwards bias.
+Based on the above finding, a good laterality task must not only be reliable, but should be able to replicate a departure from typical lateralisation in those who are left-handed.
+
+From the above reviewed studies, it appears that handedness is related to laterality across a range of tasks.
+This relationship points towards a a general laterality factor which governs lateralisation across tasks and behaviours.
+This has been argued in several hypotheses (Behrmann & Plaut, 2015; Cook, 1984; Hellige, 1983; Kosslyn, 1987).
+Common to all hypotheses is the assumption that there would be a correlation between the degree of laterality across complementary tasks.
+While studies do report significant correlations between measures of lateralisation, these are often small (Bradzakova-Trajkov, Haberling, Roberts, & Corballis, 2010).
+Furthermore, there are a large number of studies which report no correlation between laterality indices (Bryden, Hecaen, & Deagostini, 1983; Rosch, Bishop, & Badcock, 2012; Whitehouse & Bishop, 2009; van der Haegen & Brysbaert, 2018).
+This has led to the conclusion that there are multiple independent biases across tasks which lead to hemispheric lateralisation of different cognitive domains.
+
+### The current study
+
+Our principle goal for the current study was to maximise the test-retest reliability of an online laterality battery.
+Specifically, we examined whether it is possible to achieve a test-retest reliability of .65 or above using online behaviour research methods, and, if so, whether this reliability could be achieved more efficiently by using fewer trials in each test.
+The tasks included in the battery were a novel rhyme decision visual half-field tasks, a consonant-vowel dichotic listening tasks, a chimeric faces visual half-field task, and a novel finger tapping tasks.
+As we were primarily interested in establishing test-retest reliability, the behavioural tasks included in the battery were completed twice on two different occasions.
+In addition to establishing the test-retest reliability for the battery, we tested several predictions about the relationship between laterality on the four behavioural tasks.
+We describe these predictions below.
+
+Our first set of predictions related to participants’ performance at a population level on each task. Based on previous work, we predicted that:
+(1)	the laterality indices for Rhyme Decision will show a right visual field advantage (indicating left hemisphere dominance for language processing); 
+(2)	the laterality indices for the Dichotic Listening task will show a right ear advantage (indicating left hemisphere dominance for language); 
+(3)	the laterality indices for Chimeric Faces will show a left visual field advantage (indicating right hemisphere dominance for face or emotion processing); and 
+(4)	the Finger Tapping task will show a right-hand advantage.
+
+Our second prediction related to the inter-relationships between laterality indices.
+With online testing, there is the opportunity to test inter-relationships between laterality indices with an adequately powered sample using measures of known reliability.
 Prior studies have been inconclusive because it has been difficult gather large amounts of data to establish test-retest reliability.
@@
 However, their study was restricted to left-handers, who may be atypical in this regard, and they noted that lack of association was hard to interpret because it may be due to unreliability of measurement.
-
-Our research questions were as follows:
-
-1.	Is it possible to achieve test-retest reliability of .65 or more with laterality indices derived from the four laterality tests?
-If so, could this be achieved more efficiently by using fewer trials in each test?
-2.	Laterality indices can be calculated in multiple different ways (e.g.
-using RT, accuracy or a combination of the two; as a raw difference score or a proportion).
-We will explore these options (described in more detail in the ‘Indices’ section) to determine which metric shows the strongest test-retest reliability.
-3.	Do extreme left-handers show a higher rate of atypical laterality than right-handers on the Rhyme Decision, Dichotic listening, or Chimeric Faces tasks?  
-4.	Do the laterality indices from the language and faces laterality tasks correlate?
-We aimed to consider this questiron separately for left- and right-handers, since our previous research on language laterality indicated a single factor provided a good fit to data from right-handers but not left-handers.
-
-## Hypotheses
-
-As noted above, our main goal was to devise a battery of laterality assessments with adequate test-retest reliability.
-In addition we had specific a priori hypotheses relating to questions 3 and 4 as follows:
-
-### Hypothesis 1
-At a population level, the laterality indices for Rhyme Decision will show a right visual field advantage (indicating left hemisphere dominance for language processing); the laterality indices for the Dichotic Listening task will show a right ear advantage (indicating left hemisphere dominance for language); the laterality indices for Chimeric Faces will show a left visual field advantage (indicating right hemisphere dominance for face or emotion processing); and the Finger Tapping task will show a right hand advantage.
-
-### Hypothesis 2
-Strong left-handers (i.e.
+Originally, we predicted that for right-handers, a model (a) where the two language tasks are correlated with each other but not with the chimeric faces will provide better fit to the data than a model (b) in which laterality indices are correlated between the language and chimeric face tasks, or a base model (c) where each task is independently lateralised.
+For left-handers, we anticipated that that model (c) may be a better fit to the data, indicating that laterality is dissociated across the different tasks.
+However, we were not able to test this planned analysis with the current data set as one task had low reliability.
+Thus, we instead examined the relationship between laterality tasks using correlation analysis.
+
+Our final prediction that we examined relates to handedness and atypical lateralisation.
+We predicted that strong left-handers (i.e.
 those with laterality quotient of -90 or less on the Edinburgh Handedness Inventory) will be more likely to show atypical laterality (i.e.
 laterality in the opposite direction to our population level predictions) than right-handers.
-This prediction follows from observations by Mazoyer et al (2014) who studied a large group of left- and right-handers using fMRI to measure language laterality.
-They concluded "concordance of hemispheric dominance for hand and for language occurs barely above the chance level, except in a group of rare individuals (less than 1% in the general population) who exhibit strong right hemisphere dominance for both language and their preferred hand."
-
-### Hypothesis 3
-For right-handers, a model (a) where the two language tasks are correlated with each other but not with the chimeric faces will provide better fit to the data than a model (b) in which laterality indices are correlated between the language and chimeric face tasks, or a base model (c) where each task is independently lateralised.
-For left-handers, we anticipate that that model (c) may be a better fit to the data, indicating that laterality is dissociated across the different tasks.
+This prediction follows from observations by Mazoyer et al (2014, p.
+1) who studied a large group of left- and right-handers using fMRI to measure language laterality.
+They concluded "concordance of hemispheric dominance for hand and for language occurs barely above the chance level, except in a group of rare individuals (less than 1% in the general population) who exhibit strong right hemisphere dominance for both language and their preferred hand”.
+
+## Method
+
+This project was pre-registered on the Open Science Framework (OSF) prior to the commencement of data collection.
+The registration form along with the analysis scripts and anonymised data can be found on the OSF at https://osf.io/qsxcv/.
+
+### Participants
+
+Prior to data collection, we estimated that we would need 300 participants in order to achieve sufficient precision of the estimate of reliability.
+With this sample size, if the true test-retest correlation is .58, then the 95% confidence interval around the estimate will be .50-.65.
+Thus, with this sample size, if we observed a correlation of .65, then the true correlation is unlikely to be lower than .58.
+As will become clear in the Data Analysis section, our final sample exceeded 300 participants.
+In total 441 native English speakers were initially recruited through Prolific Academic.
+Participants had normal or corrected-to-normal vision and reported no history of neurological disease (e.g.
+head injury, brain tumour, stroke) or significant hearing loss/impairment.
+
+This study was approved by the Medical Sciences Inter-Divisional Research Ethics Committee (reference number: R66638/RE002) and conducted in accordance with the principles of the Declaration of Helsinki.
+Participants gave informed consent and received compensation for their participation.
+
+### Online behavioural laterality battery
+
+The battery consisted of eight tasks and was administered via Gorilla (Anwyl-Irvine, Massonnié, Flitton, Kirkham, & Evershed, 2020), a cloud-based tool for collecting data in the behavioural sciences.
+Each task is described below.
+
+#### Basic demographics questionnaire
+
+The basic demographics questionnaire asked participants to report handedness, footedness, years in education, and whether they were bilingual.
+
+#### Test of ocular dominance
+
+Adapted versions of the Miles test (Miles, 1929) and the Porta test (Porac & Coren, 1976) were used to determine each participant’s eye dominance in central gaze (i.e., when looking straight ahead).
+Each test classified participants as being either left or right eye dominant.
+
+#### Edinburgh handedness inventory
+
+The Edinburgh Handedness Inventory (Oldfield, 1971) was administered to quantify handedness on a continuum.
+For 10 everyday activities, participants indicated their preferred hand use on a 5-step scale (right hand strongly preferred, right hand preferred, no preference, left hand preferred, left hand strongly preferred).
+Scores range between -100 and 100 with negative values indicating a left-hand preference and positive values a right hand preference.
+
+#### LexTALE
+
+The Test for Advanced Learners of English (Lemhöfer & Broersma, 2012) was used to assess participants’ vocabulary knowledge.
+Participants saw 60 letter strings and had to indicate which word they knew.
+Of the 60 strings, 40 were actual English words and 20 were non-words.
+LexTALE scores were corrected the unequal proportion of words and non-word by averaging the percentages corrected for these two item types.
+Accordingly, the LexTALE scores ranged from 0 to 100.
+
+#### Rhyme decision task
+
+The rhyme decision task is a novel visual half-field technique which requires participants decide whether a foveally present letter string rhymed with an image shown in either the left or right hemifield.
+
+##### Materials
+
+Twenty-six monosyllabic words were presented in foveal vision.
+Each word was 3-4 letters in length (*mean*= 3.8; *SD*= 0.40) and had an average Zipf frequency [log10(frequency per million words) + 3] of 4.4 (*SD*= 0.72) when using the SUBTLEX-UK database (van Heuven, Mandera, Keuleers, & Brysbaert, 2014).
+Each foveal word was paired with an image that corresponded to one of 26 monosyllabic words which had a mean Zipf frequency of 4.5 (*SD*= 0.48).
+For example, the word “bite” was paired with the image “kite” and the word “more” was paired with the image “door”.
+Each word-image pair was then paired with another word-image pair for which there was no inter-pair rhyme (i.e.
+bite-kite was paired with more-door).
+This resulted in 13 word-image couplets.
+For each of the foveally presented words in a given word-image couplet, the corresponding images could appear in either the LVF or RVF.
+Thus, for each of the 13 couplets, four combinations were possible.
+For the pairing of bite-kite and more-door the following stimuli were created: kite – bite – door, door – bite – kite, door – more – kite, kite – more – door.
+In total, there were 52 unique stimuli.
+On a given trial, each image was displayed at 2.5 deg.
+eccentricity of the foveal word (see Figure 1).
+
+![Figure 1.
+Schematic illustration of a trial on the rhyme decision task.
+In the example, the word “bite” is presented along with the two images “kite” and “door”.
+When making the correct judgement, participants would respond with by pressing ‘S’ to indicate that the rhyme was on the left.](Picture1.png)
+
+##### Procedure
+
+Each of the 52 unique stimuli were shown five times.
+The 260 stimuli were then displayed across four blocks (65 items per block).
+Each of the four blocks also contained five items for which there was no rhyme.
+For example, the word “moon” was displayed with the images for kite and door: kite – moon – door.
+These foil items were included in attempt to prevent participants from preferentially responding to items in either the LVF or RVF.
+These items were then pseudo-randomised such that image pairs could not appear one after the other.
+
+Each trial began with a fixation cross presented centrally for 800 ms.
+Participants were instructed to keep their eyes on the central fixation point.
+Then a foveally presented word would appear for 200 ms.
+Word stimuli were presented in black lowercase Tahoma font on a white background.
+Font was scaled to 250% equating to 28pt font.
+The foveally presented word was replaced with bilaterally presented images at approximately 2.5 of visual angle either side of fixation.
+The eccentricity of images was maintained using Gorilla’s scaling tool.
+The bilaterally presented images were then presented for 150 ms.
+Brief bilateral presentation has guaranteed adequate control for eye movements in previous VHF experimental with verbal naming (Beaumont, 1982).
+Studies that directly monitored eye movements reported failures of fixation on only 0.5% of trials (Geffen, Bradshaw & Nettleton, 1972, but see also Bourne, 2006).
+At stimulus offset, participants indicated whether the image in the LVF or RVF rhymed with the foveally presented word.
+Participants were instructed to press ‘S’ to indicate that the foveally presented word rhyme with the image in the LVF and ‘K’ if it rhymed with the image in the RVF.
+If participants did not detect a rhyme, they were instructed to press the space bar.
+The response triggered the next trial.
+
+#### Dichotic listening task
+
+The consonant-vowel dichotic stimuli were kindly provided by Dr Emma Karlsson, and were used with the permission of Professor Kenneth Hugdahl.
+The stimuli have been used in prior research and reported elsewhere (Hugdahl et al., 2009; Karlsson et al., 2019).
+
+##### Materials
+
+The consonant-vowel stimuli were formed by pairing six stop-consonants (/b/, /d/, /g/, /p/, /t/, /k/) with the vowel /a/.
+Each of the six consonant-vowel stimuli (/ba/, /da/, /ga/, /pa/, /ta/, /ka/) were paired and played in each sound channel (e.g.
+/ba/-/da/), resulting in 36 pairings including homonyms.
+
+##### Procedure
+
+Participants completed four blocks of the dichotic listening task.
+In each block, participants heard each of the 36 stimuli once.
+Across the four blocks, they responded to 144 stimuli.
+Participants were instructed to wear headphones when completing the dichotic listening task.
+We included two headphone checks to ensure that participants were using stereo-headphones, which we describe further in the General Procedure.
+At the start of each trial, participants saw a fixation cross for 250 ms, after which a stimulus was played.
+After hearing each stimulus, participants were instructed to report the syllable they heard or, if it seemed like they heard two different sounds, the one they heard best or most clearly by clicking a button corresponding to one of the six syllables.
+The response triggered the start of the next trial.
+
+#### Chimeric faces task
+
+The chimeric faces stimuli were kindly provided by Dr Michael Burt, and have been reported elsewhere (Burt & Perrett, 1997; Innes et al., 2016; Karlsson et al., 2019).
+
+##### Materials
+
+The faces consisted of symmetrical averaged images created from four male and four female faces.
+Four emotional expressions were used: anger, disgust, happiness, and sadness.
+The faces were split vertically down the middle of the face and paired so that one emotive hemifaces was attached to another and blended at the midline (see Figure 2 for an example).
+These were paired in all possible combinations creating 16 individual stimuli.
+Four stimuli showed identical emotions in each hemiface.
+
+![Figure 2.
+An example stimulus from the chimeric faces task.
+In the example, “happiness” is shown on the left and “sadness” on the right.
+For a participant who is right lateralised, they might be expected to have stronger emotion processing for the left hemiface.
+As such, they will likely report seeing “happiness”.](chim_example.png)
+
+##### Procedure
+
+Participants completed four blocks of the chimeric faces task.
+In each block, participants viewed each stimulus twice.
+Thus, they viewed 32 stimuli per block and 128 in total.
+At the start of each trial, participants fixated a cross for 1000 ms.
+A chimeric face was then displayed for 400 ms.
+Participants then reported the emotion seen in the face by clicking a button corresponding to one of the four emotions.
+The response triggered the start of the next trial.
+
+#### Finger tapping task
+
+The Finger Tapping task is a novel behavioural task to classify handedness.
+Participants were required to either press the ‘W’, ‘R’, ‘V’, and ‘X’ keys with their left hand or the ‘Y’, ‘I’, ‘M’, ‘B’ keys with their right hands in sequence as many times as possible within a 30 second window.
+Participants completed four blocks of the finger tapping task in total.
+Each block contained required participants to complete the task for both their dominant and non-dominant hand.
+
+#### General Procedure
+
+The current study implemented a test-retest, within-subjects design across two sessions.
+During session 1, participants completed the full battery of tasks.
+Self -report measures were completed first, followed by the behavioural laterality tasks.
+During session 2, participants completed only the behavioural laterality tasks.
+Hence, each participant provided data for the rhyme decision, dichotic listening, chimeric faces, and finger tapping tasks twice.
+Each session lasted approximately 45 minutes and 35 minutes respectively.
+
+As the dichotic listening task demands the use of stereo-headphones, participants were screened at the start of each session.
+Screening consisted of two tasks.
+The first involved participants judging which of three pure tones is quietest, and is described by Woods, Siegel, Traer, and McDermott (2017) in full detail.
+Sound discrimination is designed to be easy to achieve when using headphones but difficult over loudspeakers due to phase-cancellation.
+Participants were removed from the study if they failed to correctly discriminate the quietest sound on at least five out of six trials, indicating that they were unlikely using headphones.
+The second involve participants hearing a sound presented in only one of the two channels.
+Participants were removed if they failed to correctly identify which channel the sound was played on five out of six trials, indicating that they were not using stereo-headphones.
+
+When completing the behavioural tasks participants were instructed to sit approximately 50 cm from the screen.
+Gorilla’s scaling tool was utilised such that stimuli size and eccentricity was maintained across tasks and participants.
+Stimuli were from each task were repeated across four blocks.
+Within each block participants completed 70 trials of the rhyme detection task (R); 36 trials of the dichotic listening task (D); 32 trials of the chimeric face task (C); and 2 trials of the finger tapping task (F) for both their dominant and non-dominant hand.
+The order of tasks with each block were determined as follows RDCF, DCFR, CFRD, FRDC, and the order of tasks remained consistent across blocks for each participant.
+For each task, trial order was randomised, except for the finger tapping task where participants responded with their dominant hand followed by their non-dominant hand.
+
+#### Data analysis
+
+```{r read_data,echo=FALSE,include=TRUE}
+# read in data from csv
+LIs <- read.csv('https://osf.io/x5jd2/download',stringsAsFactors = F)
+# cor test for rhyme 
+corR<- cor(LIs$Day2.RDT_RT_LI, LIs$Day1.RDT_RT_LI, use = "complete.obs")
+# table of counts for footed and sigh
+short_dat <- dplyr::select(LIs, "subject", "handedness", "footedness", "miles", "porta")
+long_dat <- gather(short_dat, task, response, footedness:porta, factor_key=TRUE)
+# make table
+tab <- as.data.frame(table(long_dat$handedness, long_dat$task, long_dat$response))
+tab <- tab %>% 
+  rename(
+    Handedness = Var1,
+    Task = Var2,
+    Response= Var3,
+    Count= Freq
+    )
+split <- ggplot(data=tab, aes(x=Response, y=Count, fill=Handedness)) +
+  geom_bar(stat="identity", position=position_dodge()) + facet_grid(.~Task) +
+    geom_text(aes(label=Count), vjust=-0.3, position = position_dodge(0.9), size=3.5)+
+  theme_classic() + theme(legend.position = "top") + ylab("Count") + xlab("") + ylim(0,220)
+ggsave(
+  "handed_sight.png",
+  plot = split,
+  width = 6, height = 4,
+  dpi = 300
+)
+```
+
+Out of the 441 partiipants recuited via Prolific, 421 (`r round(421/441*100, 1)`%) complete both sessions.
+We then removed 21 (`r round(21/421*100, 1)`%) participants from our final sample as their lexTALE score was below 80, indicating that a participant’s level of vocabulary knowledge was below the cut off for lower advanced proficiency.
+A further 10 participants (`r round(10/400*100, 1)`%) were excluded due to issues with long refresh rates or self report that stimuli were displaying on several tasks.
+This left a sample of 390 participants.
+
+For each task we adaopted the following data cleaning procedures.
+Applying each of these meant that the number of participants left in the final analyses could vary by task.
+First, for all tasks, we removed trials in which reaction times were shorter than 200 ms.
+Outlier reaction times were identified using Hoaglin and Iglewicz’s (1987) procedure, whereby outliers are defined as reaction times that are 1.65 times the difference between the first and third quantiles (1.65 * (Q3-Q1)) below or above the first and third quantile values (e.g: lower limit = Q1 – 1.65 x (Q3-Q1); upper limit = Q3 + 1.65 x (Q3-Q1)).
+Our differences criterion was set to 1.65 instead of the typical 2.2 because we expected to obtain asymmetric distributions for reaction time, since reaction time is always non-negative, and so most outliers would be in one tail of the distribution.
+The Hoaglin and Iglewicz procedure was applied to each participants’ data separately.
+After applying each data trimming procedure, participants left with less than 80% of trials were removed.
+
+Next, for the chimeric faces task, we removed partcipants data if they scored lower than 75% on trials in which an identical emotion was shown in each hemisface.
+These data trimming procedures left data for `r 390-nrow(LIs[LIs$exclude== 1 | LIs$exclude== 12,])` participants on the chimeric faces.
+For the dichotic listening task, we removed those who scored less than 75% correct on trials where the same sound was played to each ear.
+These data trimming procedures left data for `r 390-nrow(LIs[LIs$exclude== 2 | LIs$exclude== 12,])` participants on the dichotic listening task.
+
+With regards to the rhyme task, we removed participants who scored less than 50% accuracy in either visual field.
+Furthermore, we removed particpants who scored less than 75% accuracy on the catch trials (trials in which there was no rhyme).
+These trimming procedures left data for `r nrow(LIs[LIs$exRDT== 0,])` participants on the rhyme decision task.
+Seeing as we were removing a relatively high portion of participants, want wanted to check whether this task was yielding reliably results.
+With `r nrow(LIs[LIs$exRDT== 0,])` participants, the test-retest reliability was *r*= `r round(corR,3)`, which is far smaller than the r= 0.65 that we pre-registered as our criteria for a reliable task.
+Thus, rather than collect further participants so that we would have our pre-registered 300 participants for the rhyme, we chose to end data collection here and analyses the 390 participants that we had collected under the assumption that the rhyme decision is not reliable.
+The final group of 390 participants (`r nrow(LIs[LIs$gender== "Female",])` female) had a mean age of `r round(mean(LIs$age, na.rm= TRUE), 1)` years (*SD*= `r round(sd(LIs$age, na.rm= TRUE), 2)` and included `r nrow(LIs[LIs$handedness== "Left",])` left handers (`r nrow(LIs[LIs$handedness== "Left" & LIs$gender== "Female",])` female).
+`r nrow(LIs[LIs$handedness== "Right" & LIs$bilingual== "Yes",])` right-handers were bilingual.
+`r nrow(LIs[LIs$handedness== "Left" & LIs$bilingual== "Yes",])` left-handers were bilingual.
+The division of footedness and ocular dominance in left- and right-handers are shown in Figure 3.
+Below we describe the analysis procedures used to examine our research questions and hypotheses.
+
+![Figure 3. Count of left- and right-handers who report left, right, or don’t know on self-reported footedness, the Miles test, and the Porta test.](handed_sight.png)
+
+##### Calculations of laterality indices
+
+For the rhyme task, we pre-registered that we would calculate LIs based on reaction time data, as accuracy during piloting was at ceiling.
+We specified that we would calculate LIs in two ways: (1) LIs would be calculated in the traditional fashion (100 x (Right – Left)/(Right + Left)); (2) LIs would be calculated in a way that corresponded to a *t*-value.
+This was achieved by conducting a *t*-test for each participant where accuracy response times are the dependent variable and visual field is the independent variable.
+This would estimate the sensitivity to stimuli presented in either visual field and equate to a laterality index.
+One benefit of this approach is that we could identify participants who show a statistically significant reaction time advantage for one side, i.e.
+where the *t*-value on a 2-sided test has *p*< .05.
+
+For the dichotic listening and chimeric faces, we specified that LI calculated would be based on trials in which participants correctly identified a stimulus.
+The count of correct responses that corresponded to each side would be used to generate an accuracy LI in two ways: (1) LIs would be calculated in the traditional fashion (100 x (Right – Left)/(Right + Left)); (2) LIs would be computed as a *z*-statistic for each participant, using z = (pR-.5)/sqrt(pR x pL/n), where pR is the proportion of correct responses in the RVF, pL is proportion of L responses in the LVF, and n is total LVF and RVF responses.
+LIs for the finger tapping task would be calculated in the same manner, but using the number of sequences generated by the left and right hand as opposed to the proportion of correct responses in each visual field.
+
+##### Reliability assessment
+
+For the reliability assessment, we first plotted the distributions of LIs and checked them for normality using a Shapiro-Wilk test.
+As the LIs were non-normal, we used a series of Spearman’s correlations to examine the correlation of coefficients for session 1 versus session 2.
+
+##### Population level analysis
+
+To address hypothesis 1 and examine the laterality of each task at the population level, we conducted a series of one-sample *t*-tests for each task using session 1 data.
+
+##### Atypicality
+
+To address hypothesis 2, that extreme left-handers will be more likely to show atypical laterality than right-handers, we first identified participants as being atypical.
+Here participants were defined as left-handed if they scored less than -90 on the Edinburgh Handedness Inventory and right-handed if they scored more than 90 of the Edinburgh Handedness Inventory.
+Atypicality was first defined in participants where their *t*- or *z*-score was significantly laterality in the opposite direction from that expected.
+If there were less than 20 cases defined as atypical using this method, we adjusted the cut-off on that task until we acquired 20 atypical participants.
+Once typicality was defined, we first analysed atypicality on the Rhyme Decision and Chimeric Faces task.
+Participants were categorised into four groups: typical on both; atypical on Rhyme Decision only; atypical on Chimeric Faces only, and atypical on both.
+A chi-square test was then used to test for association with the number of people in each of the four groups and the handedness classification from the Edinburgh Handedness Inventory.
+This was then be repeated for the Dichotic Listening and the Chimeric Faces tasks.
+A Bonferroni correction was applied to the outcome of each of the two chi-squared tests.
+
+##### Relationship between language and non-language LIs
+
+We had originally intended to examine the relationship between LIs on the Rhyme Decision, Dichotic Listening, and Chimeric faces by modelling covariances.
+This would have allowed us to compare (1) a model where the two language tasks are correlated with each other but not with the chimeric faces, (2) a model where all LIs are correlated, and (3) a base model where the LIs are unrelated.
+This would have been done separately for right- and left-handers.
+However, due to the Rhyme Decision task having poor reliability, we decided that this analysis would not be appropriate.
+Instead, we examine the correlation coefficient for LIs on the Dichotic Listening and Chimeric Faces tasks.
+For completeness, we include the proposed analysis as a supplementary materials but do not draw conclusions based on this.
+
+1
+1
+1
+1
+1
+1
+1
+1
+1
+1
+1
+1
+1
+1
+1
+1
+1
+1
+1
+1
+1
+1
+1
+1
+
+# Pre-reg info
 
@@
 
-```{r setup, echo=FALSE,include=TRUE}
-knitr::opts_chunk$set(echo = TRUE)
-require(tidyverse)
-library(MASS)
-library(lavaan)
-library(qpcR) #?? why do we need this?!
-library(tidyverse)
-library(semPlot)
-library(RVAideMemoire) #for spearman.ci
-library(ggplot2)
-library(kableExtra)
-library(corrr)
-options(knitr.kable.NA = '') #should suppress NA in tables
-
-```
 <!---Notes to Adam from Dorothy (7th June 2020).
@@
 LIs <- read.csv('https://osf.io/x5jd2/download',stringsAsFactors = F)
+
+cor(LIs$Day2.RDT_RT_LI, LIs$Day1.RDT_RT_LI, use = "complete.obs")
 #LIs <- read.csv("LIs.csv", na.strings = c("NA", ""))

### 2020-07-03 draft_notebook_writeup.Rmd
Drafted majority of results. Still requires minimal trials for stable LI which is next to do.

@@
 
-Department of Experimental Psychology, University of Oxford, Anna Watts Building, Radcliffe Observatory Quarter, Woodstock Rd, Oxford OX2 6GG
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
+Department of Experimental Psychology, University of Oxford, Anna Watts Building, Radcliffe Observatory Quarter, Woodstock Road, Oxford, OX2 6GG.
+
+<!--- Leaving blank for now as not sure how corresponding authorship works in OSCCI lab --->
+*Correspondence regarding this article should be addressed to XXXXXX, Department of Experimental Psychology, University of Oxford, Anna Watts Building, Woodstock Road, Oxford, OX2 6GG. Email: XXXXXX
 
@@
 
-Prior to data collection, we estimated that we would need 300 participants in order to achieve sufficient precision of the estimate of reliability.
-With this sample size, if the true test-retest correlation is .58, then the 95% confidence interval around the estimate will be .50-.65.
-Thus, with this sample size, if we observed a correlation of .65, then the true correlation is unlikely to be lower than .58.
-As will become clear in the Data Analysis section, our final sample exceeded 300 participants.
-In total 441 native English speakers were initially recruited through Prolific Academic.
+As our ultimate goal is to use these measures in individual differences research, we aimed to optimise the task to achieve observed test-retest reliability of at least .65.
+To achieve sufficient precision of the estimate of reliability, we estimated we needed 300 participants.
+With this sample size, if the true correlation is .58, then the 95% confidence interval around the estimate will be .50-.65.
+Thus with this sample size, if we observe a correlation of .65, then the true correlation is unlikely to be lower than .58.
+
+Out of the 300 participants, we originally specified that we'd collected data for least 80 left-handers.
+We had examined the power that a sample of 80 participants would provide to make the originally planned model comparisons outlined under hypothesis 3.
+By comparing model fit using a likelihood test, 80 participants would provide 100% power for model (a) and 97% power for model (b).
+Thus, we would have sufficient power to assess the fit of each model using Akaike weights, which are more conservative than likelihood tests.
+
+Despite setting out to recruit 300 participants, our final sample exceeded this target (we say more about this below).
+In total, 441 native English speakers were initially recruited through Prolific Academic.
 Participants had normal or corrected-to-normal vision and reported no history of neurological disease (e.g.
@@
 
+<!--Dorothy, just a note that all Ps came from Prolific in the end-->
+
 ### Online behavioural laterality battery
@@
 
-The basic demographics questionnaire asked participants to report handedness, footedness, years in education, and whether they were bilingual.
+The basic demographics questionnaire asked participants to report handedness, years in education, and whether they were bilingual.
+We also asked participnats a question about footedness: which foot do you normally use to step up on a ladder/step?
+Participants could answer "left", "right", or "don't know".
 
@@
 For 10 everyday activities, participants indicated their preferred hand use on a 5-step scale (right hand strongly preferred, right hand preferred, no preference, left hand preferred, left hand strongly preferred).
-Scores range between -100 and 100 with negative values indicating a left-hand preference and positive values a right hand preference.
+The Edinburgh handedness inventory was scored in the standard way, which involves combining information about direction and exclusiveness of hand preference across ten tasks.
+For simple contrast between left- and right-handers, we treated indices greater than 0 as right handed, and indices less than 0 as left handed.
+For analysis of extreme left-handers, we focussed on left-handers with an index of -90 or less, which corresponds to near-exclusive use of the left hand.
 
@@
 Of the 60 strings, 40 were actual English words and 20 were non-words.
-LexTALE scores were corrected the unequal proportion of words and non-word by averaging the percentages corrected for these two item types.
+LexTALE scores were corrected the unequal proportion of words and non-word by averaging the percentages corrected for these two item types (e.g., ((number of words correct/40 x 100) + (number of nonwords correct/20 x 100))/2).
 Accordingly, the LexTALE scores ranged from 0 to 100.
@@
 
-```{r read_data,echo=FALSE,include=TRUE}
+```{r read_dat, echo- FALSE, include= FALSE}
+library(ggplot2)
+library(tidyr)
+library(dplyr)
+library(knitr)
+library(kableExtra)
+
 # read in data from csv
 LIs <- read.csv('https://osf.io/x5jd2/download',stringsAsFactors = F)
-# cor test for rhyme 
-corR<- cor(LIs$Day2.RDT_RT_LI, LIs$Day1.RDT_RT_LI, use = "complete.obs")
+```
+
+```{r prepdat,echo=FALSE,include=FALSE}
+# remove those without day 2
+# reformat NA
+LIs$day2.complete <- ifelse(is.na(LIs$day2.complete), 0, LIs$day2.complete)
+# remove
+LIs_c <- LIs[LIs$day2.complete == 1,]
+LIs <- LIs[LIs$day2.complete == 1,]
+# lextale out df
+LIs_no <- LIs[LIs$lexOut != "keep",]
+# now remove them from LI
+LIs <- LIs[LIs$lexOut == "keep",]
 # table of counts for footed and sigh
@@
 )
-```
-
-Out of the 441 partiipants recuited via Prolific, 421 (`r round(421/441*100, 1)`%) complete both sessions.
-We then removed 21 (`r round(21/421*100, 1)`%) participants from our final sample as their lexTALE score was below 80, indicating that a participant’s level of vocabulary knowledge was below the cut off for lower advanced proficiency.
-A further 10 participants (`r round(10/400*100, 1)`%) were excluded due to issues with long refresh rates or self report that stimuli were displaying on several tasks.
-This left a sample of 390 participants.
+#Add a category of handedness that distinguishes extreme left from other left-handers
+#In preregistration, extreme left defined as EHI laterality -90 or less.
+LIs$handed3 <- LIs$handedness
+w<-which(LIs$index_EHI<(-89))
+LIs$handed3[w]<-'Extreme.left'
+```
+
+Out of the 441 partiipants recuited via Prolific, `r nrow(LIs_c)` (`r round(nrow(LIs_c)/441*100, 1)`%) complete both sessions.
+We then removed `r nrow(LIs_no)` participants (`r round(nrow(LIs_no)/nrow(LIs_c)*100, 1)`%) from our final sample as their lexTALE score was below 80, indicating that a participant’s level of vocabulary knowledge was below the cut off for lower advanced proficiency.
+This left a final sample of `r nrow(LIs)` participants.
 
 For each task we adaopted the following data cleaning procedures.
-Applying each of these meant that the number of participants left in the final analyses could vary by task.
+Applying each of these procedures meant that the number of participants left in the final analyses could vary by task.
 First, for all tasks, we removed trials in which reaction times were shorter than 200 ms.
@@
 Next, for the chimeric faces task, we removed partcipants data if they scored lower than 75% on trials in which an identical emotion was shown in each hemisface.
-These data trimming procedures left data for `r 390-nrow(LIs[LIs$exclude== 1 | LIs$exclude== 12,])` participants on the chimeric faces.
+These data trimming procedures left data for `r nrow(LIs)-nrow(LIs[LIs$exclude== 1 | LIs$exclude== 12,])` participants on the chimeric faces.
 For the dichotic listening task, we removed those who scored less than 75% correct on trials where the same sound was played to each ear.
-These data trimming procedures left data for `r 390-nrow(LIs[LIs$exclude== 2 | LIs$exclude== 12,])` participants on the dichotic listening task.
+These data trimming procedures left data for `r nrow(LIs)-nrow(LIs[LIs$exclude== 2 | LIs$exclude== 12,])` participants on the dichotic listening task.
 
@@
 Seeing as we were removing a relatively high portion of participants, want wanted to check whether this task was yielding reliably results.
-With `r nrow(LIs[LIs$exRDT== 0,])` participants, the test-retest reliability was *r*= `r round(corR,3)`, which is far smaller than the r= 0.65 that we pre-registered as our criteria for a reliable task.
-Thus, rather than collect further participants so that we would have our pre-registered 300 participants for the rhyme, we chose to end data collection here and analyses the 390 participants that we had collected under the assumption that the rhyme decision is not reliable.
-The final group of 390 participants (`r nrow(LIs[LIs$gender== "Female",])` female) had a mean age of `r round(mean(LIs$age, na.rm= TRUE), 1)` years (*SD*= `r round(sd(LIs$age, na.rm= TRUE), 2)` and included `r nrow(LIs[LIs$handedness== "Left",])` left handers (`r nrow(LIs[LIs$handedness== "Left" & LIs$gender== "Female",])` female).
-`r nrow(LIs[LIs$handedness== "Right" & LIs$bilingual== "Yes",])` right-handers were bilingual.
-`r nrow(LIs[LIs$handedness== "Left" & LIs$bilingual== "Yes",])` left-handers were bilingual.
+With `r nrow(LIs[LIs$exRDT== 0,])` participants, the test-retest reliability was *r*= `r round(cor(LIs[LIs$exRDT== 0,]$Day1.RDT_RT_LI, LIs[LIs$exRDT== 0,]$Day2.RDT_RT_LI, use = "complete.obs"),3)`, which is smaller than the *r*= 0.65 that we pre-registered as our criteria for a reliable task.
+Thus, rather than collect further participants so that we would have our pre-registered 300 participants for the rhyme decision task, we chose to end data collection here and analyse the `r nrow(LIs)` participants that we had collected under the assumption that the rhyme decision is not reliable.
+The final group of `r nrow(LIs)` participants (`r nrow(LIs[LIs$gender== "Female",])` female) had a mean age of `r round(mean(LIs$age, na.rm= TRUE), 1)` years (*SD*= `r round(sd(LIs$age, na.rm= TRUE), 2)` and included `r nrow(LIs[LIs$handedness== "Left",])` left handers (`r nrow(LIs[LIs$handedness== "Left" & LIs$gender== "Female",])` female).
+`r nrow(LIs[LIs$handedness== "Right" & LIs$bilingual== "Yes",])` right-handers were bilingual and `r nrow(LIs[LIs$handedness== "Left" & LIs$bilingual== "Yes",])` left-handers were bilingual.
 The division of footedness and ocular dominance in left- and right-handers are shown in Figure 3.
@@
 where the *t*-value on a 2-sided test has *p*< .05.
+Based on considerations from pilot data, we planned to transform RT data from the RDT task by log-transformation before computing *t*-values.
 
@@
 
-For the reliability assessment, we first plotted the distributions of LIs and checked them for normality using a Shapiro-Wilk test.
+For the reliability assessment, we plotted the distributions of LIs and checked them for normality using a Shapiro-Wilk test.
 As the LIs were non-normal, we used a series of Spearman’s correlations to examine the correlation of coefficients for session 1 versus session 2.
+
+For tasks where the test-retest reliability was above .75 with the existing test length, we reanalysed the data using just the first N trials, reducing N until reliability becomes unsatisfactory (below *r*= .75)^1^.
+This will allow us to identify the most efficient way to administer these tasks in future, while retaining good reliability.
+However, the full data (including all trials) was used for all subsequent analyses.
 
@@
 
-1
-1
-1
-1
-1
-1
-1
-1
-1
-1
-1
-1
-1
-1
-1
-1
-1
-1
-1
-1
-1
-1
-1
-1
-
-# Pre-reg info
-
-## Design Plan
-We note here specific study characteristics that are specified for a preregistration on OSF. We have added additional information to specify Methods in more detail.
-
-### Study type
-This is an observational Study in which data are collected from study subjects who are not randomly assigned to a treatment.
-
-### Blinding
-No blinding is involved
-
-### Study design
-We used a within-subjects design whereby participants performed each laterality task twice (repeated-measures) in order to establish test-retest reliability.
-
-The battery of tasks was broken into four blocks.
-Within each block, participants completed 70 trials of the Rhyme Detection task; 36 trials of the Dichotic Listening task; 32 trials of the Chimeric Face task; and 2 trials of the Finger Tapping task for both their dominant and non-dominant hand.
-Trial order for blocks of Rhyme Detection, Dichotic Listening, and Chimeric Face task was randomised.
-For the Finger Tapping task, all participants were asked to respond first with their dominant hand followed by their non-dominant hand.
-
-The order of tasks was balanced by using four possible orders as follows ABCD, BCDA, CDAB, DABC.
-Each successive participant entering the study was assigned the next task order in the sequence above.
-Order of tasks remained consistent across blocks for that participant.
-
-### Randomization
-No randomization was used in this study.
-
-### Sampling Plan
-The study was registered prior to collection of data
-
-## Participants  
-### Data collection procedures
-We aimed to recruit participants via three methods:
-
-1)	Given that we are also interested in genetically related samples, a portion of participants will be recruited via social media adverts, with encouragement for twins or first-degree relatives (siblings or parent-child pairs) to volunteer.
-
-2)	Undergraduate students from the University of Oxford will be recruited via the Department of Experimental Psychology’s Research Participation Scheme.
-
-3)	Other participants will be recruited from the Prolific web-based population (https://www.prolific.co/).
-
-In practice, the first route was ineffective, with very few related individuals volunteering. 
-<!--Adam: can we report how many came from Prolific and how many from student scheme?-->
-
-### Inclusion criteria
-Participants were required to be at least 16 years old, native English speakers, with no reported history of neurological disease (e.g.
-head injury, brain tumour, stroke), no history of significant hearing loss/impairment, and with normal or corrected-to-normal vision.
-We intentionally oversampled left-handers (see below).
-
-### Sample size
-We specified we would continue data collection until we had complete and useable datasets from 300 participants.
-As explained below, we had fewer participants than this for one task, because of many exclusions.
-However, as this task performed poorly, we decided not to test further participants.
-
-### Sample size rationale
-As our ultimate goal is to use these measures in individual differences research, we aim to optimise the task to achieve observed test-retest reliability of at least .65.
-To achieve sufficient precision of the estimate of reliability, we estimated we needed 300 participants.
-With this sample size, if the true correlation is .58, then the 95% confidence interval around the estimate will be .499-.65.
-Thus with this sample size, if we observe a correlation of .65, then the true correlation is unlikely to be lower than .58.
-
-We aimed to collect data for approximately 80 left-handers and 220 right-handers.
-We examined the power that a sample of 80 participants would provide to make the model comparisons outlined under hypothesis 3.
-By comparing model fit using a likelihood test, 80 participants would provide 100% power for model (a) and 97% power for model (b).
-Thus, we would have sufficient power to assess the fit of each model using Akaike weights, which are more conservative than likelihood tests.
-
-### Stopping rule
-We planned to cease data collection when we had collected 300 full datasets.
-
-## Methods  
-### Manipulated variables
-In this test-retest design, the within-subject independent variable of interest was Session (1 or 2).  
-In the four laterality tasks, stimuli were presented in either the left or right visual fields (Rhyme Decision and Chimeric Faces), to the left or right ear (Dichotic Listening), or using the left or right hand (Finger Tapping).
-
-### Tasks 
-The test battery was implemented online via Gorilla (https://www.gorilla.sc/).
-<!--Adam: could you add information about the checks for screensize-->
-
-## Rhyme Decision
-This was a novel visual-half-field task devised for this study.
-The goal was to devise a task that engaged phonological processing by requiring the participant to generate a name from a pictured object and judge if it rhymed with a centrally-presented written word.
-Two pictures were presented to the left and right of fixation, and the prediction was that judgements should be faster when the rhyming picture occurred in the right visual field, which projects directly to the left hemisphere.
-The stimulus trios (two pictures and a word) are shown in Appendix A.
-On xx trials, stimuli were presented for which there was no rhyme match; this was used as a means of identifying and excluding participants who were not attending carefully.
-
-## Dichotic Listening 
-Materials were taken from a dichotic listening task that has been developed as an app (Hugdahl et al., 2009).
-Participants wear stereo headphones.
-On each trial they hear two consonant-vowel (CV) auditory stimuli simultaneously, one to each ear.
-They are required to report which CV stimulus out of six options (ba, da, ga, ka, pa, ta) provided the best match to the stimulus they heard.
-This measures laterality of language processing.
-The original test has xx dichotic stimulus pair, xx of which present the same syllable in both ears.
-These latter stimuli could be used to identify individuals who had difficulty in discriminating the syllables.
-<!--Adam: could you add information about the checks for headphones-->
-
-##	Chimeric Faces
-Materials were taken from Chimeric Faces task developed by Burt & Perrett (1997) and used by Innes, Burt, Birch, & Hausmann (2016).
-Participants view stimuli that are vertically split at the midline with each hemiface displaying one of four emotions (anger, disgust, happiness, and sadness).
-They are required to report the overall emotion that they perceive in the face.
-This measures laterality of face and emotion processing.
-The task also includes xx trials in which both sides of the face show the same emotion.
-This is used as a check to identify whether the participant can reliably identify the intended emotions when there is no conflicting information.
-
-## Finger Tapping
-This was a simple task devised for this study.
-Our goal was to see whether we could obtain a quantification of extent of relative motor skill of the two hands using the individual's computer keyboard.
-The participant was asked to use the index finger to tap a specified sequence of keys continuously for a 20 second <!--Adam: is that right?--> interval, and the number of correct sequences was counted.
-The task was performed first with the preferred hand and then with the nonpreferred hand.
-
-## Other tasks  
-The Lexical Test for Advanced Learners of English (LexTALE; Lemhöfer & Broersma, 2012) was used to assess vocabulary as an indicator of language proficiency.
-The Edinburgh Handedness Inventory (Oldfield, 1971) was used to assess manual preference.  
-Modified versions of the Miles Test (Miles, 1929) and the Porta Test (Porac & Coren, 1976) were used to assess ocular dominance, and a question about foot preference.
-<!--Adam Could you give more information about Lextale, Miles/Porta, and what was the foot preference question-->
-
-## Measured variables  
-For Rhyme Detection, Dichotic Listening, and Chimeric Faces, the measured variables were reaction time (RT) and accuracy.
-These were converted into laterality indices, as described below.
-For the Finger Tapping task, the measured variable was correct number of sequences produced with the right and left hands.
-This was also be converted into a laterality index.
-
-For the LexTALE task, the measured variables was the number of words with correct responses (out of 40) and the number of nonwords with correct responses (out of 20).
-An adjusted score was computed by averaging the percentage of correct words and the percentage of correct nonwords (e.g., ((number of words correct/40*100) + (number of nonwords correct/20*100)) / 2).
-
-The Edinburgh Handedness Inventory was scored in the standard way, which involves combining information about direction and exclusiveness of hand preference across ten tasks.
-For simple contrast between left- and right-handers, we treated indices greater than 0 as right handed, and indices less than 0 as left handed.
-For analysis of extreme left-handers, we focussed on left-handers with an index of -90 or less, which corresponds to near-exclusive use of the left hand.
-
-Adapted versions of the Miles Test and Porta Test will classify participants as being either left or right eye dominant.<!---Adam: not sure if we need this here, as we did not specify how they would be used in analysis?--->
-
-## Laterality Indices
-The classic way of calculating a laterality index is to take the difference between left and right sides and divide it between the total for the two sides.
-This gives a difference measure that is scaled by the individual's overall speed or accuracy.
-
-We also experimented with an alternative approach that allows one to identify whether an individual participant is reliably lateralised.
-For Reaction Time measures, this involved taking all the RTs for left and right-sided responses in a single participant, and conducting a t-test.
-One can then set a cutoff for reliable lateralisation, e.g.
-by using a t-value of 1.96, which would occur in around 5% of individuals if there was no real bias to one side.
-Based on considerations from pilot data, we planned to transform RT data from the RDT task by  log-transformation before computing t-values.
-
-Similar logic can be used for accuracy measures, where a z-score can be computed and interpreted in relation to the normal distribution, using the formula:
-z = (pR-.5)/sqrt(pR*pL/n) 
-where pR is the proportion of R responses, pL is proportion of L responses, and n is total L and R responses.
-
-These t- and z-measures amount to rescaling of the conventional laterality quotient and so are not expected to give different values for reliability, but they may be useful if one wanted to distinguish between individuals who showed robust lateralisation versus those who showed less convincing bias to one side.
-
-### Data exclusion
-
-In the laterality tasks, individual trials will be excluded from analysis if the reaction time was faster than 200 ms, as this suggests an anticipatory response.
-We did not preregister an upper RT boundary, but planned to apply Hoaglin-Iglewicz criteria (Hoaglin & Iglewicz 1987) to minimise the influence of outliers in the rightwards tail of the distribution.
-As such, outliers are defined as RTs that are 1.65 times the difference between the first and third quantiles (1.65*(Q3-Q1)) below or above the first and third quantile values.
-Our cut-off criterion was set to 1.65 instead of the typical 2.2.
-This is because we expected to obtain asymmetric distributions for reaction time, since reaction time is always non-negative, and so most outliers would be in one tail of the distribution.
-
-Participants will be excluded if:
-•	Responses were excluded or missing for more than 20% of trials in any condition in either of the language laterality tasks or chimeric faces.  
-•	Performance for stimuli presented in their dominant visual-half field is not significantly above chance level.
-Moreover, participants whose performance for stimuli presented in their non-dominant visual half field is significantly below chance will also be excluded.
-•	LexTALE adjusted score is below 80, indicating that a participant’s level of vocabulary knowledge is below the cut off for lower advanced proficiency.
-We anticipate that native English speakers will have high proficiency and therefore will score highly on this task.
-•	For the Chimeric Face Task, participants’ data will be excluded if their accuracy on trials with identical emotions in each hemiface is below 75%.
-This is to ensure that participants can sufficiently judge emotion expression.
-•	For the Dichotic Listening task, participants’ data will be excluded if their accuracy on trials with identical sounds played to each ear is below 75%.  
-•	In the dichotic listening task we will screen for headphone use using an adapted version of a task reported by Woods, Siegle, Traer, and McDermott (2017) where participants judge which of three tones is the softest.
-If accuracy is lower than 5 out of 6, we will remove participants as recommended by Woods et al.
-
-### Missing data
-
-Excluded participants will be replaced, so that a total of 300 complete datasets are collected.
-
-## Data Analysis
-### Statistical models
-
-For the reliability assessment, we plotted the distributions of laterality indices and checked them for normality using the Shapiro-Wilk test.
-For normally-distributed scores, we used Pearson correlation coefficients for time 1 vs time 2, after excluding outliers.
-In the pre-registration we suggestd we might transform data to achieve normality if necessary, or alternatively use a Spearman correlation.
-
-Optimising the test battery: If reliability turns out to be very high (above .75) with the existing test length, we will reanalyse the data using just the first N trials, reducing N until reliability becomes unsatisfactory (below r=0.65).
-This will allow us to identify the most efficient way to administer these tasks in future, while retaining good reliability.
-However, the full data (including all trials) will be used for all subsequent analyses.
-<!---Adam: we have not done this yet, but it should not be too hard--->
-
-To test Hypothesis 1 (laterality bias at a population level) one-sample t-tests will be applied for each task and session.
-
-Hypothesis 2 predicts that left-handers will be more likely to show atypical laterality than right-handers.
-The test of this hypothesis will be complicated by having to decide how to define 'atypical'.
-Ideally, these would be individuals who have significant lateralisation in the opposite direction to the majority, as indicated by p < .05 on the individual z-score or t-score.
-We anticipated, however, that this criterion may select very few cases on some of our tasks, in which case we will adjust the cutoff on that task until we have at least 20 atypical cases.
-Once typicality is defined, we will first analyse atypicality on the Rhyme Decision and Chimeric Faces task.
-Participants will be categorised into four groups: typical on both; atypical on Rhyme Decision only; atypical on Chimeric Faces only, and atypical on both.
-A chi square test will be used to test for association with the number of people in each of the four groups and the handedness classification from the Edinburgh Handedness Inventory.
-This will then be repeated for the Dichotic Listening and the Chimeric Faces tasks.
-<!---to do--->
-
-Hypothesis 3 concerns the relationships between the laterality measures.
-We planned to use  structural equation modelling (SEM) to compare the covariances of the laterality statistics between the two language tasks and chimeric faces to test between three models.
-This would be done separately for left- and right-handers.
-Before doing this we would flip the polarity of chimeric faces, so that all predictions are for positive correlations.
-
-There were no specific predictions for the finger tapping task; this was included as an exploratory measure, which may allow us to better characterise handedness on a dimension.
-
-## Inference criteria
-
-Assessment of test-retest reliability will use the 95% confidence interval around the estimate, as documented above.  
-Assessment of whether a function is lateralised will be tested for a difference from zero using a conventional .05 significance level.
-The association between handedness and atypical laterality will be tested using a Bonferroni corrected .025 significance level.
-Model fit will be evaluated using AIC model selection using Akaike weights (Wagenmakers & Farrell, 2004).
-Akaike weights are transformations of the AIC values, which can be directly interpreted as conditional probabilities for each model.
-The best fitting model to the data for right- and left-handers will be selected as the final model.
-
-## Exploratory analysis
-
-As noted above, we will use the z- or t-statistics to index laterality, but we will have a large dataset that might be used subsequently to explore different approaches to measuring a laterality index, and to consider how choice of measure affects test-retest reliability.
-
-While the role of the LexTALE in the current research is largely for the purpose of data exclusion, we plan to explore correlations between LexTALE score and the laterality index on the Rhyme Decision task.
-This will enable us to examine whether vocabulary knowledge is related to rhyme judgement when information is displayed in each hemifield.
-We will similarly explore the influence of sight dominance, handedness/footedness, and the Finger Tapping Task on the laterality index for rhyme decision.
-
-Finally, we will investigate whether the method for measuring handedness has an impact on the associations between handedness and laterality on the Rhyme Decision and Chimeric Faces tests.
-In our main planned analyses, handedness will be coded using a binary categorisation of the Edinburgh Handedness Inventory scores, where > 0 indicates right handedness and < 0 indicates left handedness.
-There are many alternative ways of coding handedness: e.g.
-using three categories (right handed, ambidextrous and left handed); as a continuous measure; using self-report handedness; or using performance on the Finger Tapping test as an objective, continuous measure.
-We will explore which of these coding systems gives the strongest associations with laterality indices on the Rhyme Decision and Chimeric Faces.
-To do this we will select half the data at random to explore associations, and then aim to replicate on the second half.
-
 ## Results
 
-<!---Notes to Adam from Dorothy (7th June 2020).
-I've deposited the LIs file on OSF and now am reading direct from that.
-I want to double check polarity of LIs - not sure they are all consistently in same direction.  
-I turned the basic plotting/correlation etc into a function so it can be applied to all tasks.  
-I added a vertical line to density plots at zero.
-Have added a new column handed3 which distinguishes between extreme left handers and others. This may need tweaking in terms of the colours used to code and the terms,
-Added a one-sample t-test to check if significantly lateralised - but then found you had done that later!
-No longer removing those without complete datasets
-I will try to condense results so they can be reported more succinctly--->
-
-```{r read_dat,echo=FALSE,include=TRUE}
-# read in data from csv
-LIs <- read.csv('https://osf.io/x5jd2/download',stringsAsFactors = F)
-
-cor(LIs$Day2.RDT_RT_LI, LIs$Day1.RDT_RT_LI, use = "complete.obs")
-#LIs <- read.csv("LIs.csv", na.strings = c("NA", ""))
-# then we exclude those who we do not have full datasets for 
-# This step currently omitted: use all available data for each task
-# LIs <- subset(LIs, exclude== 0)
-
-#Add a category of handedness that distinguishes extreme left from other left-handers
-#In preregistration, extreme left defined as EHI laterality -90 or less.
-LIs$handed3 <- LIs$handedness
-w<-which(LIs$index_EHI<(-89))
-LIs$handed3[w]<-'Extreme.left'
-
-```
+```{r descriptives, echo=FALSE,include=FALSE}
+# here we present the raw count and RT scores for each task
+# read CF data
+for_figs <- read.csv("./data/for_figs.csv", na.strings = "NA")
+# get CF
+CF.short.dat <- for_figs[for_figs$exclude== 0 | for_figs$exclude== 2,]
+  CF.short.dat <-CF.short.dat[CF.short.dat$task== "CF",]
+# read Dichotic data
+DL.short.dat <- for_figs[for_figs$exclude== 0 | for_figs$exclude== 1,]
+  DL.short.dat <- DL.short.dat[DL.short.dat$task== "DL",]
+# read rhyme data
+RD.short.dat <- for_figs[for_figs$exRDT== 0,]
+  RD.short.dat <- RD.short.dat[RD.short.dat$task== "RD",]
+# read finger data
+finger.short.dat <- read.csv("./data/finger.csv", na.strings = "NA")
+
+# plots
+CF <- ggplot(CF.short.dat, aes(x = side, y = accurate)) +
+        ggpirate::geom_pirate(aes(colour = side)) +
+        theme_classic() + ylab("Count of correct responses") + xlab("") + facet_wrap(.~Session)+ 
+        ggtitle("C") + theme(plot.title = element_text(hjust = 0.5))
+DL <- ggplot(DL.short.dat, aes(x = side, y = accurate)) +
+        ggpirate::geom_pirate(aes(colour = side)) +
+        theme_classic() + ylab("Count of correct responses") + xlab("") + facet_wrap(.~Session)+ 
+        ggtitle("B") + theme(plot.title = element_text(hjust = 0.5))
+RD <- ggplot(RD.short.dat, aes(x = side, y = RT)) +
+        ggpirate::geom_pirate(aes(colour = side)) +
+        theme_classic() + ylab("Reaction time (ms)") + xlab("") + facet_wrap(.~Session) + 
+        ggtitle("A") + theme(plot.title = element_text(hjust = 0.5))
+FT <- ggplot(finger.short.dat, aes(x = side, y = sequences)) +
+        ggpirate::geom_pirate(aes(colour = side)) +
+        theme_classic() + ylab("Count of correct sequences") + xlab("") + facet_wrap(.~Session)+ 
+        ggtitle("D") + theme(plot.title = element_text(hjust = 0.5))
+all.plot <- suppressMessages(suppressWarnings(ggpubr::ggarrange(RD, DL, CF, FT)))
+all.plot
+# save
+ggsave(
+  "all.vf.png",
+  plot = all.plot,
+  width = 6, height = 6,
+  dpi = 300
+)
+```
+
+##### Calculations of laterality indices
+
+Participants' average reaction times on the rhyme decision task are shown in Figure 4 broken down by visual field and session.
+While we do not analyse accuracy, accuracy on the rhyme task was high  (*M*= `r round(mean(RD.short.dat$p.corr, na.rm= TRUE),1)`%, *SD*= `r round(sd(RD.short.dat$p.corr, na.rm= TRUE),2)`%).
+The count of correct answers on the dichotic listening and chimeric faces tasks are also shown in Figure 4, along with the number of sequences gerated by the left and right hand.
+LIs were computed in several ways.
+For all tasks, we computed LIs in the typical way: (100 x (Right – Left)/(Right + Left)).
+It is important to note, however, that we present LIs such that they are reported to be consistent with lateralisation at the brain level (i.e.
+by multiplying LIs by -1).
+A negative LI indicates left lateralisation at the brain level while a positive LI indicates right lateralisation at the brain level.
+For the rhyme decision task, we then computed an LI using a *t*-test for inidiviaul participants fit to log-tranformed reaction time data.
+For the reamining tasks (which made use of count data) we computed a *z*-statistic for each participant using count data.
+
+![Figure 4.
+Mean scorees across the two sessions of the behavioural laterality tasks.
+(A) Reaction times in each visual field on the rhyme decision tasks.
+(B) Count of correctly identified sounds played to the left or right ear.
+(C) Count of correctly identified emotions in each hemiface.
+(D) Count of sequences generated by the left and right hand.](all.vf.png)
+
+For each measure, LIs and their test-retest correlation coefficients are reported in Table 1.
+In addition, Table 1 inlcudes the *p*-values for a one-sampled test and Shapiro-Wlk test for each measure  As the distirbution of each LI was non-normal, spearman's rank correlations were used to estimate the test-retest reliability.
+Confidence intervals were estimated using bootstapping with 10,000 iterations.
+From Table 1, it is clear that we are able to obtain good test-retest reliability when calculating LIs in a typical manner.
+For all but the rhyme decision task, we see good test-retest reliability that exceeds our pre-registered thresold of *r*= .65.
+Furthermore, LIs for each task across the two sessions are lateralised at the participant level (all *p*s< .001).
+Specifically, the rhyme decision, dichotic listening, and finger tapping tasks were all lateralised to the left hemisphere (showing a left visual field advantage).
+The chimeric faces task was lateralised to the right hemipshere (showing a left visual field advantage).
+A highly similar pattern of results emerges for the *t*- and *z*-score measures, where each is lateralised at the population level (all *p*s< .001).
+The test-retest correlations are shown visually in Figure 5.
 
@@
 #This will gradually get populated as we go through each task
-
 #Cells for each task are filled in course of plots.function chunk
-
-combotable <- data.frame(matrix(NA,nrow=7,ncol=9))
-colnames(combotable)<-c('.','RhymeRT1','RhymeRT2','DichoticAcc1','DichoticAcc2','ChimericAcc1','ChimericAcc2','FingerAcc1','FingerAcc2')
+combotable <- data.frame(matrix(NA,nrow=7,ncol=17))
+colnames(combotable)<-c('.','RD 1','RD 2','DL 1','DL 2','CF 1','CF 2','FT 1','FT 2', 
+                            'RD t 1','RD t 2','DL z 1','DL z 2','CF z 1','CF z 2','FT z 1','FT z 2')
 combotable[,1]<-c('N','mean','SD','p-lateralised','p-normal','test-retest r','CI r')
@@
 # Also saving the scatterplots as png
-densplots <- function(t1,t2,name1,name2,scattername,combotable,combocolumn){
-  dens1<-ggplot(LIs, aes(x=t1))+geom_density(linetype="dashed")+theme_classic()+ggtitle(name1)
+densplots <- function(dat, time1,time2,name1,name2,scattername,combotable,combocolumn, y1, y2, x1, x2, ax, ay){
+
+# plots
+dens1<-ggplot(dat, aes(x=time1))+geom_density(linetype="dashed")+theme_classic()+ggtitle(name1)
   dens1 <- dens1+ geom_vline(xintercept = 0)
 # normality
-shap1<-shapiro.test(t1)
-ttest1 <- t.test(t1) #Check if significantly lateralised
-Nt1 <- length(t1)-length(which(is.na(t1)))
+shap1<-shapiro.test(time1)
+ttest1 <- t.test(time1) #Check if significantly lateralised
+Nt1 <- length(time1)-length(which(is.na(time1)))
 combotable[1,combocolumn]<-Nt1
-combotable[2,combocolumn]<-round(mean(t1,na.rm=T),2)
-combotable[3,combocolumn]<-round(sd(t1,na.rm=T),2)
+combotable[2,combocolumn]<-round(mean(time1,na.rm=T),2)
+combotable[3,combocolumn]<-round(sd(time1,na.rm=T),2)
 combotable[4,combocolumn]<-round(ttest1$p.value,3)
@@
 # plot denisty
-dens2<-ggplot(LIs, aes(x=t2))+geom_density(linetype="dashed")+theme_classic()+ggtitle(name2)
+dens2<-ggplot(dat, aes(x=time2))+geom_density(linetype="dashed")+theme_classic()+ggtitle(name2)
   dens2 <- dens2+ geom_vline(xintercept = 0)
 # normality
-shap2<-shapiro.test(t2)
-ttest2 <- t.test(t2)
-Nt2 <- length(t2)-length(which(is.na(t2)))
+shap2<-shapiro.test(time2)
+ttest2 <- t.test(time2)
+Nt2 <- length(time2)-length(which(is.na(time2)))
 myN<-min(Nt1,Nt2)
-combotable[2,(combocolumn+1)]<-round(mean(t2,na.rm=T),2)
-combotable[3,(combocolumn+1)]<-round(sd(t2,na.rm=T),2)
+combotable[2,(combocolumn+1)]<-round(mean(time2,na.rm=T),2)
+combotable[3,(combocolumn+1)]<-round(sd(time2,na.rm=T),2)
 combotable[4,(combocolumn+1)]<-round(ttest2$p.value,3)
 combotable[5,(combocolumn+1)]<-round(shap2$p.value,3)
+
 # conduct Spearman's wth bootstrapped CI
-myspear<-spearman.ci(t1, t2, nrep = 10000, conf.level = 0.95)
-
+myspear<-RVAideMemoire::spearman.ci(time1, time2, nrep = 10000, conf.level = 0.95)
   combotable[6,combocolumn]<-round(myspear$estimate,2)
@@
 # position for label set to 50%x and 90%y
-xpos <- max(t1,na.rm=T)*.5
-ypos <- max(t2,na.rm=T)*.9
-
-mycor<-cor.test(t1, t2, method = "spearman", alternative = "greater")$estimate
+xpos <- max(time1,na.rm=T)*.5
+ypos <- max(time2,na.rm=T)*.9
+
+mycor<-cor.test(time1, time2, method = "spearman", alternative = "greater")$estimate
 
@@
 legposy <- .85
-myscatter <- ggplot(LIs, aes(x= t1, y= t2)) +
+myscatter <- ggplot(dat, aes(x= time1, y= time2)) +
   geom_hline(yintercept=0, color = "black") +
@@
   xlab(name1) +
-  annotate("text", x = xpos, y = ypos, label = mycorlabel) +
-  theme(legend.position = c(legposx,legposy))
+  annotate("text", x = ax, y = ay, label = mycorlabel) +
+  theme(legend.position = c(legposx,legposy), legend.title = element_blank()) + ylim(y1, y2) + xlim(x1, x2)
 # Set relative size of marginal plots (main plot 10x bigger than marginals)
@@
 scattername<-paste0(scattername,".png")
-ggsave(scattername,plot=bigplot,dpi = 300)
+ggsave(scattername,plot=bigplot,dpi = 300, width = 6, height = 6,)
 
@@
 NB need to check polarity of tasks - also oddity re RDT distributions needs checking, and we said we'd do log transform , not sure we did--->
-```{r doplots.RDT_RT,echo=FALSE,include=TRUE}
-t1<-LIs$Day1.RDT_RT_LI
-t2<-LIs$Day2.RDT_RT_LI
+```{r do.plots,LI,echo=FALSE,include=TRUE}
+# RHYME DECISION
+# remove bad cases
+RD.dat <- LIs[LIs$exRDT == 0,]
+# now plot
+dat <- RD.dat
+time1<-RD.dat$Day1.RDT_RT_LI
+time2<-RD.dat$Day2.RDT_RT_LI
 name1<-"Rhyme decision LI (day 1)"
@@
 combocolumn<-2 #column to write day 1 summary for this task in combotable
-combotable<-densplots(t1,t2,name1,name2,scattername, combotable,combocolumn)
-
-```
-
-```{r doplots.dichotic_Accscore,echo=FALSE,include=TRUE}
-t1<-LIs$Day1_DL_acc_LI
-t2<-LIs$Day2_DL_acc_LI
-name1<-"Dichotic LI acc (day 1)"
-name2<-"Dichotic LI acc (day 2)"
+y1= -12
+y2= 12
+x1= -12
+x2= 12
+ax= 8
+ay= -8
+combotable<-densplots(RD.dat, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)
+
+# DICHOTIC LISTENING
+# remove bad cases
+DL.dat <- LIs[LIs$exclude == 0 | LIs$exclude == 1,]
+# now plot
+dat <- DL.dat
+time1<-DL.dat$Day1_DL_acc_LI
+time2<-DL.dat$Day2_DL_acc_LI
+name1<-"Dichotic listening LI acc (day 1)"
+name2<-"Dichotic listening LI acc (day 2)"
 scattername<-"DichoticScatter"
 combocolumn<-4 #column to write day 1 summary for this task in combotable
-combotable<-densplots(t1,t2,name1,name2,scattername, combotable,combocolumn)
-
-```
-
-```{r doplots.chimeric_Acc,echo=FALSE,include=TRUE}
-t1<-LIs$Day1_CF_acc_LI
-t2<-LIs$Day2_CF_acc_LI
-name1<-"Chimeric LI acc (day 1)"
-name2<-"Chimeric LI acc (day 2)"
+y1= -100
+y2= 100
+x1= -100
+x2= 100
+ax= 75
+ay= -75
+combotable<-densplots(DL.dat, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)
+
+# CHIMERIC FACES
+# remove bad cases
+CF.dat <- LIs[LIs$exclude == 0 | LIs$exclude == 2,]
+# now plot
+dat <- CF.dat
+time1<-CF.dat$Day1_CF_acc_LI
+time2<-CF.dat$Day2_CF_acc_LI
+name1<-"Chimeric faces LI acc (day 1)"
+name2<-"Chimeric faces LI acc (day 2)"
 scattername<-"ChimericScatter"
-combocolumn<-6 #column to write day 1 summary for this task in combotable
-combotable<-densplots(t1,t2,name1,name2,scattername, combotable,combocolumn)
-
-```
-
-```{r doplots.fingerLI,echo=FALSE,include=TRUE}
-t1<-LIs$Day1_finger_LI
-t2<-LIs$Day2_finger_LI
-name1<-"Finger LI (day 1)"
-name2<-"Finger LI (day 2)"
+combocolumn<-6 #column to write day 1 summary for this task in combotabley1= -100
+y1= -100
+y2= 100
+x1= -100
+x2= 100
+ax= 75
+ay= -75
+combotable<-densplots(CF.dat, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)
+
+# FINGER TAPPING
+dat <- LIs
+time1<-LIs$Day1_finger_LI
+time2<-LIs$Day2_finger_LI
+name1<-"Finger tapping LI (day 1)"
+name2<-"Finger tapping LI (day 2)"
 scattername<-"FingerScatter"
-combocolumn<-8 #column to write day 1 summary for this task in combotable
-combotable<-densplots(t1,t2,name1,name2,scattername, combotable,combocolumn)
-
-```
-### Table 1
-
+combocolumn<-8 #column to write day 1 summary for this task in combotabley2= 100
+y1= -35
+y2= 35
+x1= -35
+x2= 35
+ax= 26.25
+ay= -26.25
+combotable<-densplots(dat, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)
+```
+
+```{r do.plot.t,echo=FALSE,include=TRUE}
+# RHYME T-TEST
+# now plot
+dat <- RD.dat
+time1<-RD.dat$logt1
+time2<-RD.dat$logt2
+name1<-"Rhyme decision t-score (day 1)"
+name2<-"Rhyme decision t-score (day 2)"
+scattername<-"rhyme.t"
+combocolumn<-10 #column to write day 1 summary for this task in combotable
+y1= -8
+y2= 8
+x1= -8
+x2= 8
+ax= 6
+ay= -6
+combotable<-densplots(RD.dat, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)
+
+# DICHOTIC Z-TEST
+# now plot
+dat <- DL.dat
+time1<-DL.dat$Day1_Dich_acc_Z
+time2<-DL.dat$Day2_Dich_acc_Z
+name1<-"Dichotic listening z-score (day 1)"
+name2<-"Dichotic listening z-score (day 2)"
+scattername<-"dichotic.z"
+combocolumn<-12 #column to write day 1 summary for this task in combotable
+y1= -20
+y2= 20
+x1= -20
+x2= 20
+ax= 15
+ay= -15
+combotable<-densplots(DL.dat, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)
+
+# CHIMERIC FACES Z-TEST
+# now plot
+dat <- CF.dat
+time1<-CF.dat$Day1_CF_acc_Z
+time2<-CF.dat$Day2_CF_acc_Z
+name1<-"Chimeric faces z-score (day 1)"
+name2<-"Chimeric faces z-score (day 2)"
+scattername<-"chimeric.z"
+combocolumn<-14 #column to write day 1 summary for this task in combotable
+y1= -20
+y2= 20
+x1= -20
+x2= 20
+ax= 15
+ay= -15
+combotable<-densplots(CF.dat, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)
+
+# FINGER TAPPING Z-TEST
+# now plot
+dat <- LIs
+time1<-LIs$Day1_Finger_acc_Z
+time2<-LIs$Day2_Finger_acc_Z
+name1<-"Finger tapping z-score (day 1)"
+name2<-"Finger tapping z-score (day 2)"
+scattername<-"finger.z"
+combocolumn<-16 #column to write day 1 summary for this task in combotable
+y1= -8
+y2= 8
+x1= -8
+x2= 8
+ax= 6
+ay= -6
+combotable<-densplots(LIs, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)
+```
+
+Table 1. LIs and test-retest correlations for each taks.
 ```{r Table1,echo=FALSE}
 #Needs more formatting. For word output may need to use flextable. See https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html
-combotable%>%
-  kable() %>%
-  kable_styling("condensed")
-
-```
-### Question 1
-We first consider question 1: Is it possible to achieve test-retest reliability of .65 or more with laterality indices derived from the four laterality tests?
-If so, could this be achieved more efficiently by using fewer trials in each test?
-
-Table 1 shows results for the Laterality Indices (LIs) for the four tasks at time 1 and time 2 in terms of significance of lateralisation ('p-lateralised', i.e.
-comparison with distribution with mean of zero), significance of departure from normality on Shapiro-Wilks test ('p-normal'), plus the Spearman correlation between time 1 and time 2 (test-retest reliability).
-
-Figures 1-4 show the scatterplots for time 1 and time 2 LIs for all four tasks, with left-handers colour-coded, and extreme left-handers marked in black.
-
-![Figure 1](RhymeDecisionScatter.png)
-
-![Figure 2](DichoticScatter.png)
-![Figure 3](ChimericScatter.png)
-![Figure 4](FingerScatter.png)
-
-These analyses show good reliability for Dichotic Listening, Chimeric Faces and Finger Tapping.
-These tasks are also significantly lateralised in the predicted direction.
-The Rhyme Detection Task is unsatisfactory: the test-retest correlation is below our specified cutoff r= .65 for both RT and accuracy (not shown here).
-For RT, there is significant lateralisation, but in absolute terms it is small, with a high proportion LVF-lateralised.
-<!---Adam: you mentioned laterality was opposite to prediction - I think it is in the right direction, but we need to check: for RT a big score means slower performance and I am not sure if I adjusted adequately for that.
-Need anyhow to do with log RTs since that is what we preregistered - I don't think it will make much difference--->
-
-Note also that we fell short of 300 participants for this task. This is because many participants were dropped because they did not meet our inclusion criteria - e.g. low accuracy.  
-We had previously modified this task in an attempt to improve its reliability, but these analyses question whether it is useful, so we decided not to test further participants.
-
-*Still to be done: consider reliabilities with fewer trials. How short a test is feasible without loss of reliability?*
-
-### Question 2
-Laterality indices can be calculated in multiple different ways (e.g.
-using RT, accuracy or a combination of the two; as a raw difference score or a proportion).
-We will explore these options (described in more detail in the ‘Indices’ section) to determine which metric shows the strongest test-retest reliability.
-
-*Still to be done - we have computed the z-score and t-test versions discussed in introduction *
-
-### Question 3
-Do extreme left-handers show a higher rate of atypical laterality than right-handers on the Rhyme Decision, Dichotic listening, or Chimeric Faces tasks?
-
-For our preliminary analysis, we define 'atypical laterality' as laterality that is in the opposite direction to the majority, and which has an absolute z-score or t-score value greater than 1.96.
+kable(combotable) %>%
+  kable_styling(c("bordered")) %>%
+  add_header_above(c(" " = 1, "Typical LI" = 8, "t/z-score" = 8))
+```
+Note.
+LIs are transformed such that a negtaive value equates to a left lateralisation at the brain level.
+A positive value equates to right lateralisation at the brain level.
+RD= rhyme decision, DL= dichotic listening, CF: chimeric faces, and FT: finger tapping.
+
+![Figure 5.
+Scatter plots for day 1 and day 2 data on each laterality task.
+The left column shows LIs calculated in the typical way.
+The right column shows LIs calculated using *t*- and *z*-scores.
+Here, distributions are shown for left and right handers, and an additional group of extreme left-handers who had a score below -90 on the Edinburgh Handedness Inventory.](combined_LI_cor.png)
+
+Moving forward, we chose to use LIs calculated in a typical manner to address each remaining research question.
+To dtermine the smallest number of trials needed to achieve a test-retest reliabilty of *r*= .75, we recaclauted LIs by dropping one trial at a time for each participant...
+
+##### Atypicality
+
+We defined 'atypical laterality' as laterality that is in the opposite direction to the majority, and which has an absolute z-score or t-score value greater than 1.96.
 These z and t-scores indicate whether there is significant bias to the left or right within an individual.
-
-According to the pre-registration, we omit non-extreme left-handers, but we show their data here for completeness.
-We also specified in the preregistration that we would analyse performance for pairs of tasks, but the very weak lateralisation of the Rhyme Decision task made that a questionable analysis.
-Tables 2 to 4 show associations between handedness categories and laterality classification for individual tasks first.
-We did not specify how we would handle data from the two sessions: these provide an opportunity for exploratory analyses from session 1 to then be confirmed using session 2 data, so for each task we start just with session 1.
+Based on our pre-registered analysis, we divided left-handers into two groups (1) those with a score of -90 or below on the Edinburgh Handedness Inventory (c.f.
+Mazoyer et al., 2014), (2) the remaining left handers.
+While our analyses inlcude only right handers and extreme left handers, we include the remaining left-handers in Table 2 for completeness.
 
@@
 #1 sig bias in typical direction; 2: NS bias 3: in atypical direction with sig bias.
-#
 LIs$rhyme1.cut <- cut(LIs$Day1.RDT_p.corr_LI, 
                        breaks = c(-Inf, -1.96, 1.96, Inf), 
-                       labels = c("TypLat", "Nonlat", "AtypLat"), 
+                       labels = c("TypLat", "NonLat", "AtypLat"), 
                        right = FALSE)
@@
                        breaks = c(-Inf, -1.96, 1.96, Inf), 
-                       labels = c("TypLat", "Nonlat", "AtypLat"), 
+                       labels = c("TypLat", "NonLat", "AtypLat"), 
                        right = FALSE)
@@
                        breaks =  c(-Inf, -1.96, 1.96, Inf), 
-                       labels = c("AtypLat", "Nonlat", "TypLat"), 
+                       labels = c("AtypLat", "NonLat", "TypLat"), 
                        right = FALSE)
@@
 mytab2<-data.frame(mytab2)
-
-mytab2%>%
-  kable() %>%
-  kable_styling("condensed")
-
+# combine tables for printing
+mytab3 <- cbind(tabr, tabd, tabc)
+# now do chi-squared tests
+# rhyme
 tabr <- table(LIs$handed3,LIs$rhyme1.cut)
-chisq.test(tabr)
-
+rhy.chi <- chisq.test(tabr)
+# dichotic
 tabd <- table(LIs$handed3,LIs$dichotic1.cut)
-chisq.test(tabd)
-
+dic.chi <- chisq.test(tabd)
+# chimeric
 tabc <- table(LIs$handed3,LIs$chimeric1.cut)
-chisq.test(tabc)
-
+chim.chi <- chisq.test(tabc)
+# BUT these don't actually run the planed tests. We run these by removing lefties here
 #Planned test omitted group 'Left' and column Nonlat.
 chi.r<-chisq.test(tabr[c(1,3),c(1,3)])
-
 chi.d<-chisq.test(tabd[c(1,3),c(1,3)])
@@
 
-Table 2: Rhyme judgement
-```{r showtabr,echo=FALSE}
-tabr%>%
+Table 2. Counts of laterality patterns for right-, left-, and extreme geft-handers acorss the Rhyme decision, Dichotic listening, and Chimeric faces tasks.
+```{r Table 2}
+# plot group table
+mytab3%>%
   kable() %>%  
-  kable_styling("condensed")
-```
-Table 3: Dichotic listening
-```{r showtabd}
-tabd%>%
-  kable() %>%
-  kable_styling("condensed")
-
-```
-
-Table 4: Chimeric faces
-```{r showtabc,echo=FALSE}
-tabc%>%
-  kable() %>%
-  kable_styling("condensed")
-
-```
+  kable_styling(c("bordered")) %>%
+  add_header_above(c(" " = 1, "Rhyme decision" = 3, "Dichotic listening" = 3, "Chimeric faces"= 3))
+```
+Note. TypLat: lateralised in the expected direction, Nonlat: no lateralisation, and AtypLat: show lateralisation in the opposite direction.
+
 It is evident that the Rhyme Judgement task has few individuals who are significantly lateralised in either direction.
@@
 Both Dichotic Listening and Chimeric Faces, on the other hand, are strongly lateralised.
-The planned analyses, just comparing typically lateralised with atypically lateralised individuals, and extreme left vs right-handed individuals, show significant association between extreme left-handedness and atypical laterality on chi square test.
+The planned analyses, comparing typically lateralised with atypically lateralised individuals, and extreme left vs right-handed individuals, show significant association between extreme left-handedness and atypical laterality on chi square test.
 For dichotic listening, chi squared = `r round(chi.d$statistic,2)`, d.f.
-= 1, p = `r round(chi.d$p.value,3)`; for chimeric faces, chi squared = `r round(chi.c$statistic,2)`, d.f.
-= 1, p = `r round(chi.c$p.value,3)`.
+= 1, *p* = `r round(p.adjust(chi.d$p.value, method= "bonferroni", n=2),3)`; for chimeric faces, chi squared = `r round(chi.c$statistic,2)`, d.f.
+= 1, *p* = `r round(p.adjust(chi.c$p.value, method= "bonferroni", n=2),3)`.
 
@@
 For those with typical lateralisation, the corresponding figures were `r round(100*tabd[1,1]/sum(tabd[,1]),1)`% extreme left-handers, `r round(100*tabd[3,1]/sum(tabd[,1]),1)`% right-handers, and the remainder less extreme left-handers.
-
-<!-- NB v confusing because the table gives the Atypical/Typical in different order for the chimeric faces-->
 For chimeric faces, the association was even more extreme.
@@
 w3<-intersect(w1,w2)
-LIs$handed3[w3]
+#LIs$handed3[w3]
 ```
@@
 
-This brings us to the final set of predictions regarding associations between tests.
+##### Relationship between language and non-language LIs
+
+```{r LI-link,echo=FALSE,include=TRUE}
+# make sure we have good cases
+interR <- LIs[LIs$exclude == 0,]
+# now run analysis
+DL.CF<-RVAideMemoire::spearman.ci(interR$Day1_CF_acc_LI, interR$Day1_DL_acc_LI, nrep = 10000, conf.level = 0.95)
+DL.CF.est<-cor.test(interR$Day1_CF_acc_LI, interR$Day1_DL_acc_LI, method = "spearman")
+# run Bayes factor
+BF1 <- BayesFactor::correlationBF(y = interR$Day1_CF_acc_LI, x= interR$Day1_DL_acc_LI, iterations = 10000)
+# day 2
+# now run analysis
+DL.CF.2<-RVAideMemoire::spearman.ci(interR$Day2_CF_acc_LI, interR$Day2_DL_acc_LI, nrep = 10000, conf.level = 0.95)
+DL.CF.est.2<-cor.test(interR$Day2_CF_acc_LI, interR$Day2_DL_acc_LI, method = "spearman")
+# run Bayes factor
+BF2 <- BayesFactor::correlationBF(y = interR$Day2_CF_acc_LI, x= interR$Day2_DL_acc_LI, iterations = 10000)
+```
+
+Using data from session 1, we examined the correlation between LIs on the dichotic listening and chimeric faces task to examine whether the perceptual biases on these tasks are governed by a single laterality factor.
+A spearman's rank order correlation indicated a extrenely small, non-singificant relationship, *rs*= `r round(DL.CF.est$estimate, 2)`, *95% CI* [`r round(DL.CF$conf.int[1], 2)`, `r round(DL.CF$conf.int[2], 2)`] *p*= `r round(DL.CF.est$p.value[1], 3)`.Based on this analysis, there is no evidence of a relationship between laterality indicies on the dichotic listening and chimeric faces tasks.
+To infer the extent to which our data reflect a null relationship, as opposed to a Type II error, we computed a Bayes factor (Rouder, Morey, Speckman, & Province, 2012).
+We computed the Bayes factor in R using the BayesFactor package (Version 0.9.12–4.2; Morey & Rouder, 2018) with 10,000 Monte Carlo iterations.
+We assumed the default Cauchy prior for the effect size.
+The Bayes factor for the alterative model, as opposed to the null model, was .150.
+According to Jeffreys' (1961) inference criteria, this provides moderate evidence for the null hypothesis.
+To validate these findings, we repeated these analysis with session 2 data.
+Again the correlation was small and not significantly different from zero *rs*= `r round(DL.CF.est.2$estimate, 2)`, *95% CI* [`r round(DL.CF.2$conf.int[1], 2)`, `r round(DL.CF.2$conf.int[2], 2)`] *p*= `r round(DL.CF.est.2$p.value[1], 3)`.
+Again the Bayes Factor favoured the null hypothesis, *BF*= .138.
+
+##### Exploratory analyses
+
+```{r fingerEHI,echo=FALSE,include=TRUE }
+spearCI <- RVAideMemoire::spearman.ci(LIs$Day1_finger_LI, -LIs$index_EHI, nrep = 10000, conf.level = 0.95)
+# spearman's test
+speartest <- cor.test(LIs$Day1_finger_LI, -LIs$index_EHI, method = "spearman", alternative = "greater")
+```
+
+The finger tapping task produced reliable results, but the extent to which it correlated with existing measures of handedness is not clear.
+Thus, we used a spearman's rank correlation to obtain an estimate of the relationship between LIs produced on the finger tapping task and the Edinburgh Handedness Inventory using session 1 data.
+The relationship between the tasks was moderate,*rs*= `r round(speartest$estimate, 2)`, *95% CI* [`r round(spearCI$conf.int[1], 2)`, `r round(spearCI$conf.int[2], 2)`] *p*= `r round(speartest$p.value[1], 3)`.
+Thus it indicates that the task likely represents a component of hand preference in conjunction with some aspect of motor planning or sequencing.
+
+## Notes
+
+1.
+While we pre-registered that we would reduce the number of trails until *r*= .65 for tasks where the initial test-retest relaibility was *r*= .75, we chose to reduce the trials down until *r*= .75.
+We decided to do this to maintain a higher level of relaibaility when using shorter version of the battery in inidividual differences research.
+
+## Supplemental Material
+
+We had originally intended to examine the relationship between LIs on the Rhyme Decision, Dichotic Listening, and Chimeric faces by modelling covariances.
+This would have allowed us to compare (1) a model where the two language tasks are correlated with each other but not with the chimeric faces, (2) a model where all LIs are correlated, and (3) a base model where the LIs are unrelated.
+This would have been done separately for right- and left-handers.
+However, due to the Rhyme Decision task having poor reliability, we decided that this analysis would not be appropriate to base our conlcusions on.
+Instead, we report them as a supplmental material.
+The models were fit using a subset of SEM that dos not include any latent variables or directional paths.
+It only looks to constrain particular covariance patterns and report 'best' model according AIC weights.
+Model fit was evaluated using AIC model selection using Akaike weights (Wagenmakers & Farrell, 2004).
+Akaike weights are transformations of the AIC values, which can be directly interpreted as conditional probabilities for each model.
+The best fitting model to the data for right- and left-handers will be selected as the final model.
+
+```{r Roger_mod_left,echo=FALSE}
+library(lavaan)
+library(tidyverse)
+library(MASS)
+library(qpcR)
+# created handedness data
+# now create subset of data frame for analysis
+interRR <- dplyr::select(interR, 'Day1.RDT_RT_LI', 'Day1_CF_acc_LI', 'Day1_DL_acc_LI', "handedness")
+# rename
+interRR <- interRR %>% 
+  rename(
+    "rhyme1" = Day1.RDT_RT_LI,
+    "chimeric1" = Day1_CF_acc_LI,
+    "dichotic1" = Day1_DL_acc_LI)
+# subset into right and left handers
+left_dat <- subset(interRR, handedness== "Left")
+right_dat <- subset(interRR, handedness== "Right")
+# set up model 1
+model1<-"
+dichotic1~~0*rhyme1
+dichotic1~~0*chimeric1 
+rhyme1~~0*chimeric1 
+"
+# set up model 2
+model2<-"
+dichotic1~~rhyme1 
+dichotic1~~0*chimeric1
+rhyme1~~0*chimeric1 
+"
+# set up model 3
+model3<-"
+dichotic1~~rhyme1 
+dichotic1~~chimeric1 
+rhyme1~~chimeric1
+"
+# now fit model
+fit1A <- cfa(model1, data=left_dat)
+fit2A <- cfa(model2, data=left_dat)
+fit3A <- cfa(model3, data=left_dat)
+# get the aic
+aic_vector1 <- c(fitMeasures(fit1A, "aic"),fitMeasures(fit2A, "aic"),fitMeasures(fit3A, "aic"))
+#summary(fit1A, fit.measures=TRUE)
+#summary(fit2A, fit.measures=TRUE)
+#summary(fit3A, fit.measures=TRUE)
+# get weights
+AICweights1<-akaike.weights(aic_vector1)$weights
+# populat df
+IC_objectweights1<-data.frame(Model=factor(rep(c('Model 1','Model 2','Model 3'))), IC = factor(rep('AIC',each=3)),values=AICweights1)
+# now print weights
+left.weights <- ggplot(IC_objectweights1,aes(Model,values))+
+  geom_bar(stat='identity',position="dodge")+
+  ylab('Akaike weights') + xlab("") + theme_classic() + ylim(0, .7) + ggtitle("A") + 
+  theme(plot.title = element_text(hjust = 0.5))
+
+# NOW DO RIGHT HANDERS
+#We now repeat this for right handers. 
+fit1B <- cfa(model1, data=right_dat)
+fit2B <- cfa(model2, data=right_dat)
+fit3B <- cfa(model3, data=right_dat)
+
+aic_vector2 <- c(fitMeasures(fit1B, "aic"),fitMeasures(fit2B, "aic"),fitMeasures(fit3B, "aic"))
+
+#summary(fit1B, fit.measures=TRUE)
+#summary(fit2B, fit.measures=TRUE)
+#summary(fit3B, fit.measures=TRUE)
+
+AICweights2<-akaike.weights(aic_vector2)$weights
+
+IC_objectweights2<-data.frame(Model=factor(rep(c('Model 1','Model 2','Model 3'))), IC = factor(rep('AIC',each=3)),values=AICweights2)
+
+right.weights <- ggplot(IC_objectweights2,aes(Model,values)) +
+  geom_bar(stat='identity',position="dodge")+
+  ylab('Akaike weights') + xlab("") + theme_classic() + ylim(0,.7) + ggtitle("B") + 
+  theme(plot.title = element_text(hjust = 0.5))
+
+model.comp <- ggpubr::ggarrange(left.weights, right.weights, ncol = 2)
+# save
+ggsave(
+  "model.comp.png",
+  plot = model.comp,
+  width = 6, height = 3,
+  dpi = 300
+)
+```
+
+![Figure S1. Plot of Akaike weights for each model for (A) left-handers and (B) right-handers.](model.comp.png)
+
+As shown in Figure S1, the akaike weights favoured a different model for left- and right-handers.
+For left-handers model 2, where all LIs were correlated, was preferred.
+For right-handers model 1, were languge tasks were correlated with eachother but not the chimeric faces was preffered.
+The pattern of results for left handers was surprising given that in the main article we reported no correlation between the dichotic listening and chimeric faces tasks.
+Furthermore, for left-handers, the fact that the prefered model is one where language tasks are related seems to contrast previous work reporting that language tasks fractionate in left-handers (Woodhead et al., 2019).
+This is likely to reflect that the poor reliability of the rhyme decision task.
+For right-handers we see that a model where language tasks are related.
+This is as predicted but it is of course likely that this estimate also represents noise from the rhyme decision task.
+
+# Pre-reg info
+
+To do: 
+- reduce numbers of items in the task.
+
+## Exploratory analysis
+
+As noted above, we will use the z- or t-statistics to index laterality, but we will have a large dataset that might be used subsequently to explore different approaches to measuring a laterality index, and to consider how choice of measure affects test-retest reliability.
+
+While the role of the LexTALE in the current research is largely for the purpose of data exclusion, we plan to explore correlations between LexTALE score and the laterality index on the Rhyme Decision task.
+This will enable us to examine whether vocabulary knowledge is related to rhyme judgement when information is displayed in each hemifield.
+We will similarly explore the influence of sight dominance, handedness/footedness, and the Finger Tapping Task on the laterality index for rhyme decision.
+
+Finally, we will investigate whether the method for measuring handedness has an impact on the associations between handedness and laterality on the Rhyme Decision and Chimeric Faces tests.
+In our main planned analyses, handedness will be coded using a binary categorisation of the Edinburgh Handedness Inventory scores, where > 0 indicates right handedness and < 0 indicates left handedness.
+There are many alternative ways of coding handedness: e.g.
+using three categories (right handed, ambidextrous and left handed); as a continuous measure; using self-report handedness; or using performance on the Finger Tapping test as an objective, continuous measure.
+We will explore which of these coding systems gives the strongest associations with laterality indices on the Rhyme Decision and Chimeric Faces.
+To do this we will select half the data at random to explore associations, and then aim to replicate on the second half.
+
+## Results
+
+<!---I'm not sure how to weave this in as it stands, so leaving here for now--->
+
+*Still to be done: consider reliabilities with fewer trials. How short a test is feasible without loss of reliability?*
 
@@
 #Had considered showing as correlation heatmap - but values so small it wouldn't be worth doing!
-
 #http://www.sthda.com/english/wiki/ggplot2-quick-correlation-matrix-heatmap-r-software-and-data-visualization
-
+# day 1
 corrL<-round(cor(left_dat[,c(2,4,6,8)],use='complete.obs'),2)
 corrR<-round(cor(right_dat[,c(2,4,6,8)],use='complete.obs'),2)
-
-```
-
-### Table 5: Intratask correlations for Left handers
-```{r dotab5,echo=FALSE}
-corrL%>%
-  kable() %>%
-  kable_styling("condensed")
-```
-
-### Table 6: Intratask correlations for Right handers
-```{r dotab6,echo=FALSE}
-corrR%>%
-  kable() %>%
-  kable_styling("condensed")
-```
-<!---I have run out of steam at this point.
-I think some of the correlations in *left-handers* might be statistically significant, though very small.
-But the only ones that look at all nonzero are with rhyme!
-Will just redo for time 2 to see if this is robust-->
-
-```{r corrtime2,echo=FALSE,include=TRUE}
+corr1 <- cbind(corrR, corrL)
+# day 2
 corrL2<-round(cor(left_dat[,c(3,5,7,9)],use='complete.obs'),2)
 corrR2<-round(cor(right_dat[,c(3,5,7,9)],use='complete.obs'),2)
-
+corr2 <- cbind(corrR2, corrL2)
+```
+
+### Table 3: Intratask correlations using session 1 data.
+```{r dotab5,echo=FALSE}
+corr1%>%
+  kable() %>%
+  kable_styling(c("bordered")) %>%
+  add_header_above(c(" " = 1, "Session 1" = 4, "Session 2" = 4))
 ```
@@
 
-### Table 7: Intratask correlations time 2, for Left handers
+### Table 4: Intratask correlations using session 2 data.
 ```{r dotab7,echo=FALSE}
-corrL2%>%
+corr2%>%
   kable() %>%
-  kable_styling("condensed")
-```
-
-### Table 8: Intratask correlations time 2 ,for Right handers
-```{r dotab8,echo=FALSE}
-corrR2%>%
-  kable() %>%
-  kable_styling("condensed")
-```
-
-## SEM model
-
-Here we model left hander data with three prespecified covariance structures, then compare using AIC weights.
-
-This is a subset of SEM that dos not include any latent variables or directional paths. It only looks to constrain particular covariance patterns and report 'best' model according AIC weights.
-
-```{r Roger_mod_left,echo=FALSE}
-model1<-"
-dichotic1~~0*rhyme1
-dichotic1~~0*chimeric1 
-rhyme1~~0*chimeric1 
-"
-
-model2<-"
-dichotic1~~rhyme1 
-dichotic1~~0*chimeric1
-rhyme1~~0*chimeric1 
-"
-
-model3<-"
-dichotic1~~rhyme1 
-dichotic1~~chimeric1 
-rhyme1~~chimeric1
-"
-
-fit1A <- cfa(model1, data=left_dat)
-fit2A <- cfa(model2, data=left_dat)
-fit3A <- cfa(model3, data=left_dat)
-
-aic_vector1 <- c(fitMeasures(fit1A, "aic"),fitMeasures(fit2A, "aic"),fitMeasures(fit3A, "aic"))
-
-summary(fit1A, fit.measures=TRUE)
-summary(fit2A, fit.measures=TRUE)
-summary(fit3A, fit.measures=TRUE)
-
-AICweights1<-akaike.weights(aic_vector1)$weights
-
-IC_objectweights1<-data.frame(Model=factor(rep(c('Model 1','Model 2','Model 3'))), IC = factor(rep('AIC',each=3)),values=AICweights1)
-
-ggplot(IC_objectweights1,aes(Model,values,fill=Model))+
-  geom_bar(stat='identity',position="dodge")+
-  ylab('Akaike weights') + theme_classic()
-```
-
-```{r Roger_mod_right,echo=FALSE}
-#We now repeat this for right handers. 
-fit1B <- cfa(model1, data=right_dat)
-fit2B <- cfa(model2, data=right_dat)
-fit3B <- cfa(model3, data=right_dat)
-
-aic_vector2 <- c(fitMeasures(fit1B, "aic"),fitMeasures(fit2B, "aic"),fitMeasures(fit3B, "aic"))
-
-summary(fit1B, fit.measures=TRUE)
-summary(fit2B, fit.measures=TRUE)
-summary(fit3B, fit.measures=TRUE)
-
-AICweights2<-akaike.weights(aic_vector2)$weights
-
-IC_objectweights2<-data.frame(Model=factor(rep(c('Model 1','Model 2','Model 3'))), IC = factor(rep('AIC',each=3)),values=AICweights2)
-
-ggplot(IC_objectweights2,aes(Model,values,fill=Model)) +
-  geom_bar(stat='identity',position="dodge")+
-  ylab('Akaike weights') + theme_classic()
-```
-
-## Other analyses
-<!---Original text: Upon an initial look, there is a reasonable correlation between the EHI and day 1 on the finger tapping.
-However, it does look like there are two clusters.
-So, we take a look at the left and right handers seperately.
-Doing so yields dissappointing results.-->
-<!---DB: Not sure I understood this - is the point that strength of handedness on EHI is not related to finger tapping within handedness groups? --->
-
-```{r H1,echo=FALSE,include=TRUE }
-# Nice plots of distributions for each LI but I don't think we need these if we have the scatterplots, as they show the same information. T-tests done above.
-#I've preserved the script as we may want to use it, but put it in a loop so it is optional for now.
-doH1 <-0
-if(doH1 ==1){
-mycols <- c('Day1.RDT_RT_LI','Day1_DL_acc_LI','Day1_CF_acc_LI',  'Day1_finger_LI', "handedness")
-colnums<- which(colnames(LIs) %in% mycols)
-M <- LIs[,colnums]
-N <- M %>% 
-  rename(
-    "CF" = Day1_CF_acc_LI,
-    "DL" = Day1_DL_acc_LI,
-    "RD" = Day1.RDT_RT_LI,
-    "FT" = Day1_finger_LI)
-# PLOT TO SEE IF DIFFERENT FROM ZERO
-N$ID <- 1:nrow(N)
-# format to long
-data_long <- tidyr::gather(N, task, index, `CF`:`FT`, factor_key=TRUE)
-# plot data
-a <- ggplot(subset(data_long, task== "RD"), aes(x = task, y = index)) +
-  geom_hline(yintercept=0, linetype="dashed", color = "red") +
-  ggpirate::geom_pirate(aes(colour = "black"), bars = FALSE, 
-                        points_params = list(shape = 19, alpha = 0.5), lines_params = list(size = 0.8)) +
-  ylab("Laterality Index") + xlab("") + ylim(-30, 30) +
-  scale_colour_manual(values = "darkgray") + theme_classic() 
-b <- ggplot(subset(data_long, task== "DL"), aes(x = task, y = index)) +
-  geom_hline(yintercept=0, linetype="dashed", color = "red") +
-  ggpirate::geom_pirate(aes(colour = "black"), bars = FALSE, 
-                        points_params = list(shape = 19, alpha = 0.5), lines_params = list(size = 0.8)) +
-  ylab("Laterality Index") + xlab("") + ylim(-100, 100) +
-  scale_colour_manual(values = "darkgray") + theme_classic() 
-c <- ggplot(subset(data_long, task== "CF"), aes(x = task, y = index)) +
-  geom_hline(yintercept=0, linetype="dashed", color = "red") +
-  ggpirate::geom_pirate(aes(colour = "black"), bars = FALSE, 
-                        points_params = list(shape = 19, alpha = 0.5), lines_params = list(size = 0.8)) +
-  ylab("Laterality Index") + xlab("") + ylim(-100, 100) +
-  scale_colour_manual(values = "darkgray") + theme_classic() 
-d <- ggplot(subset(data_long, task== "FT"), aes(x = task, y = index)) +
-  geom_hline(yintercept=0, linetype="dashed", color = "red") +
-  ggpirate::geom_pirate(aes(colour = "black"), bars = FALSE, 
-                        points_params = list(shape = 19, alpha = 0.5), lines_params = list(size = 0.8)) +
-  ylab("Laterality Index") + xlab("") + ylim(-30, 30) +
-  scale_colour_manual(values = "darkgray") + theme_classic() 
-ggpubr::ggarrange(a, b, c, d)
-t.test(LIs$Day1_CF_acc_LI, mu= 0)
-t.test(LIs$Day1_DL_acc_LI, mu= 0)
-t.test(LIs$Day1_finger_LI, mu= 0)
-t.test(LIs$Day1.RDT_RT_LI, mu= 0)
-}
-```
-
-<!---**Finger tapping and EHI** - commented out for now as focus is on planned analyses-->
-
-```{r fingerEHI,echo=FALSE,include=TRUE }
-# plot
-FingersEHI <-0
-if(FingersEHI==1){
-finger_scatter <- ggplot(LIs, aes(x= index_EHI, y= Day1_finger_LI)) +
-  geom_hline(yintercept=0, color = "black") +
-  geom_vline(xintercept=0, color = "black") +
-  geom_point(aes(color= handedness)) +
-  geom_rug(aes(color= handedness)) +
-  theme_classic() +
-  ylab("Finger tapping LI (day 1)") + 
-  xlab("EHI") +
-  annotate("text", x = 90, y = 25, label = "paste(italic(R), \" = .58\")", parse = TRUE) +
-  theme(legend.position = c(.15, .90))
-# Set relative size of marginal plots (main plot 10x bigger than marginals)
-ggExtra::ggMarginal(finger_scatter, groupColour = TRUE, groupFill = TRUE)
-# spearman's CIs
-spearman.ci(LIs$Day1_finger_LI, LIs$index_EHI, nrep = 10000, conf.level = 0.95)
-# spearman's test
-cor.test(LIs$Day1_finger_LI, LIs$index_EHI, method = "spearman", alternative = "greater")
-# LEFT HANDERS
-left <- subset(LIs, handedness== "Left")
-# spearman's test
-cor.test(left$Day1_finger_LI, left$index_EHI, method = "spearman", alternative = "greater")
-# Right HANDERS
-right <- subset(LIs, handedness== "Right")
-# spearman's test
-cor.test(right$Day1_finger_LI, right$index_EHI, method = "spearman", alternative = "greater")
-}
-```
+  kable_styling(c("bordered")) %>%
+  add_header_above(c(" " = 1, "Session 1" = 4, "Session 2" = 4))
+```

### 2020-07-06 battery-preprocessing_v5_DB.Rmd
smallest number of trials for good test-retest

@@
 ## Updates
+
+1st July 2020: AP edits
+(1) no longer remove lexTALE in file. I think this will make things a little easier when descirbing missing data in the MS.
+(2) corrected the issue wrong files being marked for removal where accuracy is less than 75% (comma in the wrong place).
+(3) identifiers for participants who completed both days
+(4) save data frames to plot average performance across tasks in MS.
+(5) Now include minimum number of trials to achieve .75 ttest-retest.
 
@@
 knitr::opts_chunk$set(echo = TRUE)
-
-#NB for original data setwd("~/Dropbox/Online LDT/300 Ps")
 # libraries 
@@
 #Reading from OSF - this only works if files are public
-
-demo_dat <- read.csv('./data/Demographics.csv',stringsAsFactor=F)
+demo_dat <- read.csv('https://osf.io/7smj6/download',stringsAsFactors = F)
 
@@
 allsum$genderpch[w]<-NA
-
 ```
@@
 ```{r sightednesss, warning=FALSE}
-porta_dat <- read.csv("./data/Porta.csv",stringsAsFactors = F)
-miles_dat <-  read.csv("./data/Miles.csv",stringsAsFactors = F)
+porta_dat <- read.csv("https://osf.io/zmqkt/download",stringsAsFactors = F)
+miles_dat <-  read.csv("https://osf.io/ekqsp/download",stringsAsFactors = F)
 
@@
 ```{r EHI, warning=FALSE}
-
-EHI <- read.csv("./data/EHI.csv",stringsAsFactors = F)
+EHI <- read.csv("https://osf.io/kfdma/download",stringsAsFactors = F)
 
@@
 # read data
-lexTALE <- read.csv("./data/LexTALE.csv",stringsAsFactors = F)
+lexTALE <- read.csv("https://osf.io/h4btp/download",stringsAsFactors = F)
 
@@
 
-# remove those who dIDn't pass the LexTALE
-allsum <- allsum[allsum$lexTALE >= 80,]
+# mark those who ID don't pass the LexTALE
+allsum <- allsum %>% mutate(lexOut= if_else(lexTALE < 80, "remove", "keep"))
 ```
@@
 for (d in 1:2){
-  data_file <- "./data/combined/day1_combined.csv" #day 1 data
+  data_file <- "https://osf.io/mhk29/download" #day 1 data
 
   if (d ==2){
-    data_file <-  "./data/combined/day2_combined.csv"
+    data_file <-  "https://osf.io/ebm6s/download"
   }
@@
 }
+
+# lets keep a running check on the participants completing here
+complete.day2.CF <- unique(CF_dat2$subject)
 ```
@@
 #Most have 96 accepted trials in CF_dat2. A few have 1-2 - these seem to be ones that had RTs < 200, and may have been responding rapidly at random.
-
 ```
@@
   colnames(emot_corr_Same)[2]<-'Emot_corr_same'
-  allsum <- merge(allsum, emot_corr_Same, by= "subject")
+  allsum <- merge(allsum, emot_corr_Same, by= "subject", all = TRUE)
   emot_corr_Diff <- aggregate(FUN= mean, data= CF_diff, Correct~ subject)
@@
   colnames(emot_corr_Diff)[2]<-'Emot_corr_diff'
-  allsum <- merge(allsum, emot_corr_Diff, by= "subject",all.x=T)
+  allsum <- merge(allsum, emot_corr_Diff, by= "subject",all=TRUE)
 }
@@
 ```{r CFhanded}
-
 t.test(allsum$Day1_CF_acc_Z~allsum$handpch)
@@
 ```
+
 ## Next Question: Does CF laterality relate to task accuracy
@@
 }
+
+# lets keep a running check on the participants completing here
+complete.day2.DL <- unique(dichotic_dat2$subject)
 ```
@@
 #print(outliertable)
-
-```
+```
+
 Inspection of outliertable1 and outliertable2 shows most are data as expected, with 120 trials. Outliers not v numerous. Occasional cases where test does not seem to be done, just one trial recorded.
@@
 
-  allsum <- merge(allsum, sound_mean, by= "subject",all.x=T)
-  allsum <- merge(allsum, dichotic_acc, by= "subject",all.x=T)
+  allsum <- merge(allsum, sound_mean, by= "subject",all=TRUE)
+  allsum <- merge(allsum, dichotic_acc, by= "subject",all.x=TRUE)
 }
@@
 text(15,.6,paste0('r = ',cor2))
-
 ```
@@
 }
-```
+
+# lets keep a running check on the participants completing here
+complete.day2.RD <- unique(RDT_dat2$subject)
+```
+
 Now use generic function to remove outliers.
@@
 RDT_df2[is.nan(RDT_df2)] <- NA
-
-# add log RT
-RDT_df1$logRT <- log10(RDT_df1$RT)
-RDT_df2$logRT <- log10(RDT_df2$RT)
 
@@
 measure <- 'RT'
-dopirate(origdat,task,measure)
-
-# log RT
-origdat$X1 <- origdat$logRT
-measure <- 'logRT'
 dopirate(origdat,task,measure)
@@
 if(d==2){mycolname<-"Day2.RDT_RT_LI"}
-polarity <- (-1)
-allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
-
-# log-RT
-origdat$X1 <- origdat$logRT
-mycolname<-"Day1.RDT_logRT_LI"
-if(d==2){mycolname<-"Day2.RDT_logRT_LI"}
-polarity <- (-1)
-mycutoff <- 0
+polarity <- (1)
 allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
@@
 t.test(partsum$Day1.RDT_RT_LI)
-t.test(partsum$Day1.RDT_logRT_LI)
 t.test(partsum$Day2.RDT_p.corr_LI)
 t.test(partsum$Day2.RDT_RT_LI)
-t.test(partsum$Day2.RDT_logRT_LI)
 
@@
 cor2<-cor(partsum$Day1.RDT_RT_LI,partsum$Day2.RDT_RT_LI,use='complete.obs')
-cor3<-cor(partsum$Day1.RDT_logRT_LI,partsum$Day2.RDT_logRT_LI,use='complete.obs')
 # acc plot
@@
 text(-3,9,paste0('r = ',round(cor2,3)))
-# log plot
-plot(partsum$Day1.RDT_logRT_LI,partsum$Day2.RDT_logRT_LI,main='Rhyme detect: LI for log-RT')
-abline(v=0)
-abline(h=0)
-text(2,1,paste0('r = ',round(cor3,3)))
 ```
@@
 
+### log RT
+
+In our pre-reg, we also specified that we would use logRT. So we add this in here.
+
+```{r logRT analysis}
+# now add the logRT to original output
+# first relabel df
+RDT_logdat1 <- RDT_dat1
+RDT_logdat2 <- RDT_dat2
+# now log the RT
+RDT_logdat1$RT <- log(RDT_dat1$RT)
+RDT_logdat2$RT <- log(RDT_dat2$RT)
+# Now use generic function to remove outliers.
+# Here we keep the 1.65 of the H&I removal. 
+# BUT we now specify log 200 for lower cutoff to move onto the same scale
+RDT_logdat1<-detect.outliers(RDT_logdat1,log(200),1.65) #numbers specify min RT and 
+RDT_logdat2<-detect.outliers(RDT_logdat2,log(200),1.65) #numbers specify min RT and zscore for Hoaglin-Iglewicz respectiv
+# We create logRT for day1 and day 2
+varlist <-   c("subject", "side", "accurate", "RT") #need to be in this order, ie sub, side, acc and RR
+latlist <- c("left","no rhyme","right") #names of factor levels
+nloop<-length(latlist)
+RDT_log_df1 <- make.df(RDT_logdat1,varlist,latlist,nloop)
+RDT_log_df2 <- make.df(RDT_logdat2,varlist,latlist,nloop)
+# remove NaN
+is.nan.data.frame <- function(x)
+  do.call(cbind, lapply(x, is.nan))
+RDT_log_df1[is.nan(RDT_log_df1)] <- NA
+RDT_log_df2[is.nan(RDT_log_df2)] <- NA
+# now let's plot the log data
+for (d in 1:2){
+origdat <-RDT_log_df1[RDT_log_df1$subject %in% incRDT,]
+if (d==2)
+{origdat <-RDT_log_df2[RDT_log_df2$subject %in% incRDT,]}
+origdat <-origdat[origdat$side %in% c('left','right'),]
+# RT
+origdat$X1 <- origdat$RT
+measure <- 'log-RT'
+dopirate(origdat,task,measure)
+}
+# make LI for log RT
+# Because outliers will already be marked in allsum, we won't bother going through this again. 
+# We will intead add straight into the data frame. 
+#now create LI
+for (d in 1:2){
+origdat <-RDT_log_df1[RDT_log_df1$subject %in% incRDT,]
+if (d==2)
+{origdat <-RDT_log_df2[RDT_log_df2$subject %in% incRDT,]}
+# log RT
+origdat$side <- as.factor(origdat$side)
+  levels(origdat$side) <-c('Left','No rhyme','Right') #function assumes capitalised
+polarity<- 1 #so LI is positive in predicted direction
+mycutoff <- 0
+origdat$X1 <- origdat$RT
+mycolname<-"Day1_RDT_logRT_LI"
+if(d==2){mycolname<-"Day2_RDT_logRT_LI"}
+polarity <- (1)
+allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
+}
+# plot log RT
+cor1<-cor(allsum$Day1_RDT_logRT_LI,allsum$Day2_RDT_logRT_LI,use='complete.obs')
+plot(allsum$Day1_RDT_logRT_LI,allsum$Day2_RDT_logRT_LI,main='Rhyme detect: LI for log RT')
+abline(v=0)
+abline(h=0)
+text(-1,1,paste0('r = ',round(cor1,3)))
+# FINALLY, make t-test for logRT
+#remove 'no rhyme' trials
+for (d in 1:2){
+  thisdat<-RDT_logdat2[RDT_logdat2$side %in% c('left','right'),]
+  if(d==1){
+    thisdat<-RDT_logdat1[RDT_logdat1$side %in% c('left','right'),]
+  }
+
+  thisdat <- thisdat[thisdat$outlier==0,] #remove outliers
+  aggdata <- aggregate(thisdat$RT,by=list(thisdat$side,thisdat$item,thisdat$Participant.Public.ID),
+                       FUN=mean,na.rm=TRUE)
+  aggdata$subject <- aggdata$Participant.Public.ID
+  colnames(aggdata)<-c('side','item','subject','RT')
+  mysubs<-unique(aggdata$subject)
+  if(d==1){
+  pairedt <- data.frame(matrix(NA,nrow=length(mysubs),ncol=5))
+  colnames(pairedt) <- c('subject','t1','p1','t2','p2')
+  }
+  nsubs<-length(mysubs)
+  for (i in 1:nsubs){
+    myt<-NA
+    myp<-NA
+    thisdat <- aggdata[aggdata$subject==mysubs[i],]
+    pairedt$subject[i]<- mysubs[i]
+    #check there are data for both left and right on all items
+    checkdat <-table(thisdat$item)
+    w<-which(checkdat<2)
+    if(length(w)==0){
+      t<-t.test(thisdat$RT[thisdat$side=='left'],thisdat$RT[thisdat$side=='right'],paired=T)
+      myt<-t$statistic
+      myp<-t$p.value
+    }
+    if(d==1){
+      pairedt$t1[i] <- myt
+      pairedt$p1[i] <-myp
+    }
+    if(d==2){
+      pairedt$t2[i] <-myt
+      pairedt$p2[i] <-myp
+    }
+  }
+}
+
+# add t score to the data
+t_dat <- select(pairedt, -3, -5)
+t_dat <- na.omit(t_dat)
+# relabel for merging
+names(t_dat)[names(t_dat) == "t1"] <- "logt1"
+names(t_dat)[names(t_dat) == "t2"] <- "logt2"
+# merge to allsum for analysis
+allsum <- merge(allsum, t_dat, by.x = "origID", by.y = "subject", all.x=T)
+
+cor(pairedt$t1,pairedt$t2,use='complete.obs')
+```
+
+Even when log-transforming, things don't look any better.
+
 ## Finger Tapping Task
@@
 }
+
+# lets keep a running check on the participants completing here
+complete.day2.FT <- unique(all_finger2$subject)
 ```
@@
 
-## Final sample summary
-
-Here we report the descriptives for the participants who had full datasets. 
-(Some of these tests etc have been incorporated in analyses above, so can be pruned from script)
-
-```{r Finalallsum, warning=FALSE}
-# first, describe age for the whole sample and plot
-allsum %>%
-    summarize(avg = mean(age), n = n(), sd = sd(age))
-# second, show gender count for the whole population
-allsum %>%
-  group_by(gender) %>%
-    summarize(n = n())
-# third, show footedness
-allsum %>%
-  group_by(footedness) %>%
-    summarize(n = n())
-# fourth, show handedness
-allsum %>%
-  group_by(handedness) %>%
-    summarize(n = n())
-# fifth, show bilingualism
-allsum %>%
-  group_by(bilingual) %>%
-    summarize(n = n())
-# sixth, show miles
-allsum %>%
-  group_by(miles) %>%
-    summarize(n = n())
-# seventh, show porta
-allsum %>%
-  group_by(porta) %>%
-    summarize(n = n())
-# eigth, show LexTALE
-allsum %>%
-    summarize(avg = mean(lexTALE), n = n(), sd = sd(lexTALE))
-```
-
-```{r finalplots}
-# Now that we have the basic scores, let's look at how this looks with regards to handedness
-# first, describe age for the whole sample and plot
-plot(data=allsum, index_EHI ~ age,col=handpch,xlim=c(18,50),pch=(14+allsum$genderpch))
-
-allsum$handfac <- as.factor(allsum$handedness)
-allsum$genderfac <- as.factor(allsum$genderpch)
-levels(allsum$genderfac)<-c('male','female')
-
-a <- allsum %>%
-  group_by(genderfac, handfac) %>%
-    summarize(n = n())
-
-ggplot(data=a, aes(x=handfac, y=n, fill= handfac)) +
-  geom_bar(stat="identity")+
-  geom_text(aes(label=n), vjust=1.6, color="white", size=3.5)+
-  theme_classic() + facet_wrap(.~ genderfac) + ggtitle("Gender by handedness") +   ylab("Count") + xlab("")
-
-# third, show footedness
-a <- allsum %>%
-  group_by(handfac, footedness) %>%
-    summarize(n = n())
-ggplot(data=a, aes(x=handfac, y=n, fill= handfac)) +
-  geom_bar(stat="identity")+
-  geom_text(aes(label=n), vjust=1.6, color="white", size=3.5)+
-  theme_classic() + facet_wrap(.~ footedness) + ggtitle("Footedness by handedness") + 
-  ylab("Count") + xlab("")
-
-# fourth, show bilingualism
-a <- allsum %>%
-  group_by(bilingual,handfac) %>%
-    summarize(n = n())
-
-# fifth, show miles
-a <- allsum %>%
-  group_by(miles, handfac) %>%
-    summarize(n = n())
-ggplot(data=a, aes(x=handfac, y=n, fill= handfac)) +
-  geom_bar(stat="identity")+
-  geom_text(aes(label=n), vjust=1.6, color="white", size=3.5)+
-  theme_classic() + facet_wrap(.~ miles) + ggtitle("Mile test by handedness") + 
-  ylab("Count") + xlab("")
-# six, show porta
-a <- allsum %>%
-  group_by(porta, handfac) %>%
-    summarize(n = n())
-ggplot(data=a, aes(x=handfac, y=n, fill= handfac)) +
-  geom_bar(stat="identity")+
-  geom_text(aes(label=n), vjust=1.6, color="white", size=3.5)+
-  theme_classic() + facet_wrap(.~ porta) + ggtitle("Porta test by handedness") + 
-  ylab("Count") + xlab("")
-# seventh, show LexTALE
-ggplot(allsum, aes(x = handfac, y = lexTALE)) +
-  geom_pirate(aes(colour = handfac), bars = FALSE) + theme_classic() + 
-  ylim(70, 110) + ggtitle("LexTALE by handedness") + ylab("Standardised score") + xlab("")
-```
-
-```{r corr_between_tasks}
-mycols <- c('Day1_CF_acc_Z','Day2_CF_acc_Z','Day1_Dich_acc_Z','Day2_Dich_acc_Z','Day1_Finger_acc_Z','Day2_Finger_acc_Z')
-colnums<- which(colnames(allsum) %in% mycols)
-mycorrs<-correlate(allsum[,colnums],method='spearman')
-
-mycorrsL <-correlate(allsum[allsum$handedness=='Left',colnums],method='spearman')
-mycorrsR <-correlate(allsum[allsum$handedness=='Right',colnums],method='spearman')
-
-#compare L and R handers
-
-for (i in colnums){
-  print(names(allsum)[i])
- print(t.test(allsum[,i]~allsum$handedness))
-}
-
-#check lateralisation with single group t vs zero
-for (i in colnums){
-  print(names(allsum)[i])
- print(t.test(allsum[,i]))
-}
-
-```
+### list of participants who completed.
+
+We need a list of participants here who completed the both days for later filtering.
+
+````{r lister}
+# CF
+day2.CF <- data.frame(matrix(nrow=length(complete.day2.CF), ncol= 2))
+  colnames(day2.CF) <- c("task", "subject")
+  day2.CF$subject <- complete.day2.CF
+  day2.CF$task <- "CF"
+# DL
+day2.DL <- data.frame(matrix(nrow=length(complete.day2.DL), ncol= 2))
+  colnames(day2.DL) <- c("task", "subject")
+  day2.DL$subject <- complete.day2.DL
+  day2.DL$task <- "DL"
+# RD
+day2.RD <- data.frame(matrix(nrow=length(complete.day2.RD), ncol= 2))
+  colnames(day2.RD) <- c("task", "subject")
+  day2.RD$subject <- complete.day2.RD
+  day2.RD$task <- "CF"
+# Finger
+day2.FT <- data.frame(matrix(nrow=length(complete.day2.FT), ncol= 2))
+  colnames(day2.FT) <- c("task", "subject")
+  day2.FT$subject <- complete.day2.FT
+  day2.FT$task <- "CF"
+# now create a combined df
+complete.day2a <- merge(day2.CF, day2.DL, by= "subject")
+complete.day2b <- merge(day2.RD, day2.FT, by= "subject")
+complete.day2 <- merge(complete.day2a, complete.day2b, by= "subject")
+# add row to code
+complete.day2$day2.complete <- 1
+complete.day2 <- select(complete.day2, -2, -3, -4, -5)
+# now merge into allsum
+allsum <- merge(allsum, complete.day2, by= "subject", all= TRUE)
+
+# now write this for filtering
+write.csv(allsum, "LIs.csv")
+```
+
+### data frames for general plots
+
+In the MS we add a section on raw scores (simple plots). To do this, we require a data frame for each measure.
+
+```{r dataMS}
+# DAY 1
+# add task ID
+chim_dat1$task <- "CF"
+DL_dat1$task <- "DL"
+RDT_df1$task <- "RD"
+# merge and add day
+plot_data_1 <- rbind(chim_dat1, DL_dat1, RDT_df1[RDT_df1$side != "no rhyme",])
+  plot_data_1$Session<- 1
+
+# DAY 2
+chim_dat2$task <- "CF"
+DL_dat2$task <- "DL"
+RDT_df2$task <- "RD"
+# merge and add day
+plot_data_2 <- rbind(chim_dat2, DL_dat2, RDT_df2[RDT_df2$side != "no rhyme",])
+  plot_data_2$Session<- 2
+
+# merge both
+plot.data <- rbind(plot_data_2, plot_data_1)
+
+# remove those who only did day 1
+plot.data <- merge(plot.data, complete.day2, by= "subject")
+
+# make look up for lextale and remove
+lexLOOK <- allsum %>% dplyr::select("subject", "lexTALE")
+lexLOOK <- lexLOOK[lexLOOK$lexTALE >= 80,]
+# remove those wuth poor lextale
+plot.data <- merge(plot.data, lexLOOK, by= "subject")
+
+# now look up exclusions 
+excludeLOOK <- allsum %>% dplyr::select("subject", "exclude", "exRDT")
+plot.data <- merge(plot.data, excludeLOOK, by= "subject")
+
+# make style conistent
+plot.data$side <- tolower(plot.data$side)
+
+# save data frame
+write.csv(plot.data, "./data/for_figs.csv")
+```
+
+### Reduce to .75
+
+In our pre-registration, we specified that we would reduce the number of trials required for tasks yielding greater than .75 test-retest relaibility until w have a reliability of .65.
+However, given that this is actually considerably lovwer than the results obtained.
+We instead decided to do this until the reliability was .75.
+When removing outliers and exlcuding, we achieve this for the dichotic listening and chimeric faces.
+The code below reduces the data frame and then creates LIs.
+
+#### Chimeric faces
+
+```{r reducer}
+{for (d in 1:2){ #we will process day 1 and day 2 together
+  dat <- CF_dat1
+  if (d==2)
+  {dat <- CF_dat2}
+
+trial <- 96
+nsub <-(unique(dat$subject))
+
+myrow <- 0 # start row counter
+temp.dat <- data.frame(matrix(ncol = 4, nrow=length(nsub)*2*trial))
+  colnames(temp.dat) <- c("subject", "side", "accurate", "trial")
+for (z in nsub) {
+
+    subname <- z
+    myrows <- which(dat$subject==subname) # select rows for this subject
+    tmp <- dat[myrows,]
+
+    # loop through subjects
+    for (y in 1:trial) {
+      # limit data here
+      tmp2 <- tmp[1:y,]
+      tmp2 <- tmp2[tmp2$side != "X",]
+      tmp2 <- tmp2 %>% mutate(side= if_else(side== "Left", 1, 2))
+
+      for (s in 1:2) {
+
+        tmp3 <- tmp2[tmp2$side== s,]
+
+        myrow <- myrow+1
+        temp.dat$subject[myrow] <- z
+        temp.dat$accurate[myrow] <- sum(tmp3$Correct,na.rm=TRUE) # add row for N accurate response
+        temp.dat$trial[myrow] <- y
+        temp.dat$side[myrow] <- if_else(s== 1, "Left", "Right")
+      }
+    }
+  }
+  chim.iter2<- temp.dat
+  if (d==1)
+  {chim.iter <- chim.iter2}
+}
+# now remove NAs to avoid any issues later
+chim.iter <- na.omit(chim.iter)
+chim.iter2 <- na.omit(chim.iter2)
+# create another data frame
+myrow.2 <- 0 # start row counter
+CF.corrs <- data.frame(matrix(ncol = 2, nrow=200))
+  colnames(CF.corrs) <- c("trial", "cor")
+# LOOP THROUGH TRIAL SIZE
+for (y in 2:trial) {
+# task
+task<-'Chimeric faces'  
+# now identify the data frame
+origdat <- chim.iter[chim.iter$trial== y,]
+origdat$X1 <- origdat$accurate
+mycolname<-"Day1_CF_moving"
+mycutoff <- 0
+polarity<- (-1)#so LI is positive in predicted direction
+# select df
+newsum <- dplyr::select(allsum, subject, exclude, lexTALE)
+# make LIs for day 1
+move <- makeLI(newsum,origdat,task,mycolname,polarity,mycutoff)
+# make LIs for day 2
+origdat2 <- chim.iter2[chim.iter2$trial== y,]
+origdat2$X1 <- origdat2$accurate
+mycolname<-"Day2_CF_moving"
+move <- makeLI(move,origdat2,task,mycolname,polarity,mycutoff)
+move <- move[move$exclude == 0 | move$exclude == 2,]
+# correlate
+move_corr <- cor(move[move$lexTALE <= 80,]$Day1_CF_moving, move[move$lexTALE <= 80,]$Day2_CF_moving, use= "complete.obs")
+
+  # write to data 
+  myrow.2 <- myrow.2+1
+  CF.corrs$trial[myrow.2] <- y
+  CF.corrs$cor[myrow.2] <- move_corr
+}
+# plot
+CF.corrs <- na.omit(CF.corrs)
+CF.cor.plot <- ggplot(CF.corrs, aes(x=trial, y=cor)) +
+  geom_hline(yintercept=.65, linetype="dashed", color = "blue") +
+  geom_hline(yintercept=.75, linetype="dashed", color = "red") +
+  geom_line(size=1, alpha=0.8) + ylim(0, 1) + theme_classic() +
+  xlab("No. of trials") + ylab("Test-retest reliability") + ggtitle("B") + theme(plot.title = element_text(hjust = 0.5))
+# print number of trials needed
+print(paste0("Number of chimeric faces required for at least .65 test-retest: ", min(CF.corrs[CF.corrs$cor >= .65,]$trial)))
+print(paste0("Number of chimeric faces required for at least .75 test-retest: ", min(CF.corrs[CF.corrs$cor >= .75,]$trial)))
+}
+```
+
+Based on a couple of rounds with this, removing a lower limit on trials correct here is the only way to examine if less than 1 block would be enough to achieve r= .75.
+From this analysis, it appears that 27 trials would be sufficient to run the chimeric faces wirth reasonable reliability.
+Three repetitions of all combinations would lead to 48 trials (36 with different emotions in each hemifield, 12 same-same trials).
+This would ensure full balancing as 27 items would not allow an equal presentation of each of the four emotions across hemifaces.
+
+#### Dichotic listening
+
+```{r reducerDL}
+{for (d in 1:2){ #we will process day 1 and day 2 together
+  dat <- dichotic_dat1
+  if (d==2)
+  {dat <- dichotic_dat2}
+
+# format for conistency    
+dat <- 
+  dat %>% 
+  mutate(side= ifelse(as.character(Response) == as.character(left_channel) & Correct == 1, "Left",
+                       ifelse(as.character(Response) == as.character(right_channel) & Correct == 1, "Right", "X")))
+# code side as factor
+
+trial <- 120
+nsub <-(unique(dat$subject))
+
+myrow <- 0 # start row counter
+temp.dat <- data.frame(matrix(ncol = 4, nrow=200000))
+  colnames(temp.dat) <- c("subject", "side", "accurate", "trial")
+for (z in nsub) {
+
+    subname <- z
+    myrows <- which(dat$subject==subname) # select rows for this subject
+    tmp <- dat[myrows,]
+
+    # loop through subjects
+    for (y in 1:trial) {
+      # limit data here
+      tmp2 <- tmp[1:y,]
+      tmp2 <- tmp2[tmp2$side != "X",]
+      tmp2 <- tmp2 %>% mutate(side= if_else(side== "Left", 1, 2))
+
+      for (s in 1:2) {
+
+        tmp3 <- tmp2[tmp2$side== s,]
+
+        myrow <- myrow+1
+        temp.dat$subject[myrow] <- z
+        temp.dat$accurate[myrow] <- sum(tmp3$Correct,na.rm=TRUE) # add row for N accurate response
+        temp.dat$trial[myrow] <- y
+        temp.dat$side[myrow] <- if_else(s== 1, "Left", "Right")
+      }
+    }
+  }
+  dich.iter2<- temp.dat
+  if (d==1)
+  {dich.iter <- dich.iter2}
+}
+# now remove NAs to avoid any issues later
+dich.iter <- na.omit(dich.iter)
+dich.iter2 <- na.omit(dich.iter2)
+# create another data frame
+myrow.2 <- 0 # start row counter
+DL.corrs <- data.frame(matrix(ncol = 2, nrow=400))
+  colnames(DL.corrs) <- c("trial", "cor")
+# LOOP THROUGH TRIAL SIZE
+for (y in 2:trial) {
+# task
+task<-'Dichotic listening'  
+# now identify the data frame
+origdat <- dich.iter[dich.iter$trial== y,]
+origdat$X1 <- origdat$accurate
+mycolname<-"Day1_DL_moving"
+mycutoff <- 0
+polarity<- (-1)#so LI is positive in predicted direction
+# select df
+newsum <- dplyr::select(allsum, subject, exclude, lexTALE)
+# make LIs for day 1
+move <- makeLI(newsum,origdat,task,mycolname,polarity,mycutoff)
+# make LIs for day 2
+origdat2 <- dich.iter2[dich.iter2$trial== y,]
+origdat2$X1 <- origdat2$accurate
+mycolname<-"Day2_DL_moving"
+move <- makeLI(move,origdat2,task,mycolname,polarity,mycutoff)
+move <- move[move$exclude == 0 | move$exclude == 2,]
+move <- move[move$lexTALE < 80,]
+# correlate
+move_corr <- cor(move$Day1_DL_moving, move$Day2_DL_moving, use= "complete.obs")
+
+  # write to data 
+  myrow.2 <- myrow.2+1
+  DL.corrs$trial[myrow.2] <- y
+  DL.corrs$cor[myrow.2] <- move_corr
+}
+# plot
+DL.corrs <- na.omit(DL.corrs)
+DL.cor.plot <- ggplot(DL.corrs, aes(x=trial, y=cor)) +
+  geom_hline(yintercept=.65, linetype="dashed", color = "blue") +
+  geom_hline(yintercept=.75, linetype="dashed", color = "red") +
+  geom_line(size=1, alpha=0.8) + ylim(0, 1) + theme_classic() +
+  xlab("No. of trials") + ylab("Test-retest reliability") + ggtitle("A") + theme(plot.title = element_text(hjust = 0.5))
+# print number of trials needed
+print(paste0("Number of chimeric faces required for at least .65 test-retest: ", min(DL.corrs[DL.corrs$cor >= .65,]$trial)))
+print(paste0("Number of chimeric faces required for at least .75 test-retest: ", min(DL.corrs[DL.corrs$cor >= .75,]$trial)))
+}
+
+all.cor.plot <- suppressMessages(suppressWarnings(ggpubr::ggarrange(DL.cor.plot, CF.cor.plot)))
+all.cor.plot
+# save
+ggsave(
+  "all.cor.plot.png",
+  plot = all.cor.plot,
+  width = 6, height = 3,
+  dpi = 300
+)
+```
+
+Based on a couple of rounds with this, removing a lower limit on trials correct here is the only way to examine if less than 1 block would be enough to achieve r= .75.
+From this analysis, it appears that 82 trials would be sufficient to run the DL with reasonable reliability.
+Three repetitions of all combinations would lead to 108 trials (90 with different emotions in each hemifield, 18 same-same trials).
 

### 2020-07-10 draft_notebook_writeup.Rmd
full first draft

@@
 
-## Authors
-Adam J. Parker, Zoe V. J., Woodhead, Paul A. Thompson, and Dorothy V. M. Bishop
+Adam J. Parker
+
+Zoe V. J., Woodhead
+
+Paul A. Thompson
+
+Dorothy V. M. Bishop
 
@@
 
-<!--- Leaving blank for now as not sure how corresponding authorship works in OSCCI lab --->
-*Correspondence regarding this article should be addressed to XXXXXX, Department of Experimental Psychology, University of Oxford, Anna Watts Building, Woodstock Road, Oxford, OX2 6GG. Email: XXXXXX
-
+<!--- not sure how corresponding authorship works in OSCCI lab --->
+*Correspondence regarding this article should be addressed to Adam J.
+Parker, Department of Experimental Psychology, University of Oxford, Anna Watts Building, Woodstock Road, Oxford, OX2 6GG.
+Email: adam.parker@psy.ox.ac.uk.
+
+\newpage
 ## Abstract
 
-**Needs filling in**
-
+Studies of cerebral lateralisation often involve participants completing a series of perceptual tasks under laboratory conditions.
+This has constrained the number of participants recruited in such studies.
+The current work, however, introduces an online laterality battery that can be used to test participants on a variety of tasks: a rhyme decision visual half-field task, a dichotic listening task, a chimeric faces task, and a finger tapping task to assess manual preference.
+To validate the battery and assess its test-retest reliability, 392 participants completed the battery twice.
+Around a third of participants were left-handers.
+Not only do we report hemispheric advantages in the expected direction for each task, we additionally report good test-retest for the dichotic listening, chimeric faces, and finger tapping tasks.
+We also examine the association between handedness and atypical lateralisation and report that left-handers are more likely to depart from typical lateralisation.
+Furthermore, we provide data which indicates that lateralisation across tasks likely stems from multiple independent sources.
+The implications of these findings are discussed.
+
+*Keywords*: cerebral lateralisation, hemispheric asymmetry, language, online behaviour research methods, reliability
+
+\newpage
 ## Introduction
@@
 Here, the variation between participants should be larger than the variation within subjects.
-Taking this into consideration, the current work introduces an online behavioural laterality battery for which we established the test-retest reliability using a sample of 390 participants.
+Taking this into consideration, the current work introduces an online behavioural laterality battery for which we established the test-retest reliability using a sample of 392 participants.
 The behavioural tasks included were: a novel rhyme decision task, a dichotic listening task (Hugdahl et al., 2009), a chimeric faces task (Burt & Perret, 1997), and a novel finger tapping task.
@@
 Behavioural laterality tasks have long been used to establish hemispheric dominance for certain cognitive functions.
-Rather consistently, visual half-field and dichotic listening paradigms agree that the left hemisphere plays the dominant role in language processing  (Hugdhal, 2000; Ocklenburg & Güntürkün, 2018).
-Here we review how each paradigm has been implemented to study language lateralisation.
+Rather consistently, results from visual half-field and dichotic listening paradigms suggest that the left hemisphere plays the dominant role in language processing (Hugdhal, 2000; Ocklenburg & Güntürkün, 2018).
+Here we review how each paradigm has been implemented to study language lateralisation and each paradigms’ reliability.
 
 Evidence from the visual half-field paradigm (see Bourne, 2006, for a review), strongly points towards advantageous processing for letter strings presented in the right, relative to the left, visual field.
-As information in the right visual field projects contralaterally and arrives directly at the language dominant left hemisphere, it is thought to reflect an absence in callosal transfer that is required for the processing of information initially arriving at the right hemisphere.
+As information in the right visual field projects contralaterally and arrives directly at the language dominant left hemisphere, it is thought to reflect an absence in callosal transfer that is required for the processing of linguistic information initially arriving at the right hemisphere.
 
@@
 Krach, Chen, and Hartje (2006) compared the laterality indices (LIs) produced during a visual half-field experiment with those produced using functional Transcranial Doppler Sonography (fTCD).
-The visual half-field experiment itself consisted of lexical decisions made towards abstract nouns presented in the left and right visual-fields.
+The visual half-field experiment itself consisted of lexical decisions made towards abstract nouns presented in the left- and right visual-fields.
 During the experiment, the blood flow to left and right hemispheres of 58 participants was measured.
@@
 Despite showing the expected right visual field advantage on the visual half-field task, the behavioural LIs produced on the visual half-field paradigm did not correlate with the fTCD LIs produced during word generation (r= .18) or the visual half-field task itself (r -.004).
-The
-lack of validity here likely reflects the poor reliability of the visual field task as Krach et al.
-report the split-half reliability for the correct response rates and reaction times during the behavioural visual half-field paradigm were r= .14 and r= .20 respectively.
+The lack of validity here likely reflects the poor reliability of the visual field task as split-half reliability was far from optimal (RT spilt-half: r= .14; accuracy split-half: r= .20).
 Thus, with such low split-half reliability it is unclear whether the task was able to accurately measure language lateralisation to start with.
@@
 Brysbaert and Hunter argued that this is the result of stimuli competing during bilateral presentation, which is easier when the stimulus is presented in the language dominant hemisphere.
-Reliability and validity also depend on there being a sufficient number of observations and brief stimulus presentation.
+Reliability and validity also depend on there being a sufficient number of trials and brief stimulus presentation.
 When accounting for these methodological aspects, the validity of visual half-field tasks seems promising.
@@
 words, syllables) are presented to each ear and participants are either asked simply to report as many words as they can or to attend to one ear selectively.
-In either case, a right ear advantage has been demonstrated; typically, participants report more verbal stimuli presented to the right ear (Hugdahl, 2011; Kimura, 1961).
+In either case, a right ear advantage has been demonstrated; participants report more verbal stimuli presented to the right ear (Hugdahl, 2011; Kimura, 1961).
 Again, this advantage has been attributed to stimuli being projected to the language dominant hemisphere without callosal transfer.
@@
 Over the years, various dichotic listening paradigms have been used (Bryden, 1988; Westerhausen, 2019).
-The test-retest reliability of these paradigms have been far from optimal (Voyer, 1998; Kelley & Littenberg, 2019).
+As with the visual half-field paradigm, the test-retest reliability of dichotic listening paradigms has been far from optimal (Voyer, 1998; Kelley & Littenberg, 2019).
 However, recent data from Westerhausen and Samuelsen (2020) indicated that when designed well, the dichotic listening paradigm can be reliable.
@@
 Drebing, Federman, Edington, & Terzian, 1997; Kucharska-Pieture, & David, 2003) or view two faces, one above the other, and decide which of the two faces looks more emotive (e.g., Bourne, 2005; Levy, Heller, Banich, & Burton, 1983).
-Here the images will be identical, mirror images and typically include a neutral and emotive hemiface (i.e.
+In the two face version images will be identical, mirror images and typically include a neutral and emotive hemiface (i.e.
 Happiness-Neutral vs Neutral-Happiness).
@@
 In terms of its reliability, the chimeric faces tasks seem promising.
-In Levy et al.’s (1983) early study, they reported good split-half reliability for the two faces version of the chimeric faces task, where the left and right handers both showed a split-half reliability of .93.
+In Levy et al.’s (1983) early work, they reported good split-half reliability for the two faces version of the chimeric faces task, where the left and right handers both showed a split-half reliability of .93.
 The validity of the chimeric faces task has been demonstrated using behavioural and fMRI data.
@@
 One factor that has been associated with the degree of lateralisation is handedness.
-Neuroimaging data strongly indicates that these differences in lateralisation are the result of a group of left-handers who are right language dominant rather than inidciateing that this relationship operates on a continuum (Mazoyer et al., 2014This increased departure from typical lateralisation in left-handers is not only present in neuroimaging studies, it is also reflected in behavioural tasks.
+Neuroimaging data strongly indicates that these differences in lateralisation are the result of a group of left-handers who are right language dominant rather than indicating that this relationship operates on a continuum (Mazoyer et al., 2014).
+This increased departure from typical lateralisation in left-handers is not only present in neuroimaging studies, it is also reflected in behavioural tasks.
 Thus it will be important for our battery to replicate the findings regarding handedness and lateralisation summarised below.
 
-With regards to visual half-field techniques for studying language lateralisation, there is evidence to suggest that the visual field advantage varies with handedness (Elias, Bulman-Fleming, &McManus, 1999; Nicholls, 1994, 1998; Weekes, Capetillo-Cuncliffe, Rayman, Iacoboni, & Zaidel, 1999).
+With regards to visual half-field techniques for studying language lateralisation, there is evidence to suggest that the visual field advantage varies with handedness (Elias, Bulman-Fleming, & McManus, 1999; Nicholls, 1994, 1998; Weekes, Capetillo-Cuncliffe, Rayman, Iacoboni, & Zaidel, 1999).
 However, a relatively well powered study, involving 100 participants conducted by Willemin et al.
@@
 
-When dichotic listening paradigms to study language lateralisation, the evidence is clear that non-right-handers differ in terms of their laterality.
+When using dichotic listening paradigms to study language lateralisation, the evidence is clear that non-right-handers differ in terms of their laterality.
 While a clear right ear (i.e.
@@
 From the above reviewed studies, it appears that handedness is related to laterality across a range of tasks.
-This relationship points towards a a general laterality factor which governs lateralisation across tasks and behaviours.
+This relationship points towards a general laterality factor which governs lateralisation across tasks and behaviours.
 This has been argued in several hypotheses (Behrmann & Plaut, 2015; Cook, 1984; Hellige, 1983; Kosslyn, 1987).
@@
 
-Our first set of predictions related to participants’ performance at a population level on each task. Based on previous work, we predicted that:
-(1)	the laterality indices for Rhyme Decision will show a right visual field advantage (indicating left hemisphere dominance for language processing); 
-(2)	the laterality indices for the Dichotic Listening task will show a right ear advantage (indicating left hemisphere dominance for language); 
-(3)	the laterality indices for Chimeric Faces will show a left visual field advantage (indicating right hemisphere dominance for face or emotion processing); and 
-(4)	the Finger Tapping task will show a right-hand advantage.
-
-Our second prediction related to the inter-relationships between laterality indices.
+Our first set of predictions related to participants’ performance at a population level on each task.
+Based on previous work, we predicted that: (1) the laterality indices for Rhyme Decision will show a right visual field advantage (indicating left hemisphere dominance for language processing); (2) the laterality indices for the Dichotic Listening task will show a right ear advantage (indicating left hemisphere dominance for language); (3) the laterality indices for Chimeric Faces will show a left visual field advantage (indicating right hemisphere dominance for face or emotion processing); and (4) the Finger Tapping task will show a right-hand advantage.
+
+Our second prediction that we examined relates to handedness and atypical lateralisation.
+We predicted that strong left-handers (i.e.
+those with laterality quotient of -90 or less on the Edinburgh Handedness Inventory) will be more likely to show atypical laterality (i.e.
+laterality in the opposite direction to our population level predictions) than right-handers.
+This prediction follows from observations by Mazoyer et al (2014, p.
+1) who studied a large group of left- and right-handers using fMRI to measure language laterality.
+They concluded “concordance of hemispheric dominance for hand and for language occurs barely above the chance level, except in a group of rare individuals (less than 1% in the general population) who exhibit strong right hemisphere dominance for both language and their preferred hand”.
+
+Our final prediction related to the inter-relationships between laterality indices.
 With online testing, there is the opportunity to test inter-relationships between laterality indices with an adequately powered sample using measures of known reliability.
@@
 
-Our final prediction that we examined relates to handedness and atypical lateralisation.
-We predicted that strong left-handers (i.e.
-those with laterality quotient of -90 or less on the Edinburgh Handedness Inventory) will be more likely to show atypical laterality (i.e.
-laterality in the opposite direction to our population level predictions) than right-handers.
-This prediction follows from observations by Mazoyer et al (2014, p.
-1) who studied a large group of left- and right-handers using fMRI to measure language laterality.
-They concluded "concordance of hemispheric dominance for hand and for language occurs barely above the chance level, except in a group of rare individuals (less than 1% in the general population) who exhibit strong right hemisphere dominance for both language and their preferred hand”.
-
 ## Method
@@
 As our ultimate goal is to use these measures in individual differences research, we aimed to optimise the task to achieve observed test-retest reliability of at least .65.
-To achieve sufficient precision of the estimate of reliability, we estimated we needed 300 participants.
-With this sample size, if the true correlation is .58, then the 95% confidence interval around the estimate will be .50-.65.
-Thus with this sample size, if we observe a correlation of .65, then the true correlation is unlikely to be lower than .58.
-
-Out of the 300 participants, we originally specified that we'd collected data for least 80 left-handers.
-We had examined the power that a sample of 80 participants would provide to make the originally planned model comparisons outlined under hypothesis 3.
+To achieve sufficient precision around the estimate of reliability, we estimated that we would require 300 participants.
+Out of the 300 participants, we originally specified that 80 would be left-handers.
+We had examined the power that a sample of 80 participants would provide to make the originally planned model comparisons outlined under hypothesis 2.
 By comparing model fit using a likelihood test, 80 participants would provide 100% power for model (a) and 97% power for model (b).
@@
 
-<!--Dorothy, just a note that all Ps came from Prolific in the end-->
-
 ### Online behavioural laterality battery
@@
 The basic demographics questionnaire asked participants to report handedness, years in education, and whether they were bilingual.
-We also asked participnats a question about footedness: which foot do you normally use to step up on a ladder/step?
+We also asked participants a question about footedness: which foot do you normally use to step up on a ladder/step?
 Participants could answer "left", "right", or "don't know".
@@
 For 10 everyday activities, participants indicated their preferred hand use on a 5-step scale (right hand strongly preferred, right hand preferred, no preference, left hand preferred, left hand strongly preferred).
-The Edinburgh handedness inventory was scored in the standard way, which involves combining information about direction and exclusiveness of hand preference across ten tasks.
+The Edinburgh handedness inventory was scored in the standard way, which involves combining information about direction and exclusiveness of hand preference.
 For simple contrast between left- and right-handers, we treated indices greater than 0 as right handed, and indices less than 0 as left handed.
@@
 
-The rhyme decision task is a novel visual half-field technique which requires participants decide whether a foveally present letter string rhymed with an image shown in either the left or right hemifield.
+The rhyme decision task is a novel visual half-field technique which requires participants to decide whether a foveally presented letter string rhymed with an image shown in either the left or right hemifield.
 
@@
 Font was scaled to 250% equating to 28pt font.
-The foveally presented word was replaced with bilaterally presented images at approximately 2.5 of visual angle either side of fixation.
+The foveally presented word was replaced with bilaterally presented images at approximately 2.5 deg.
+of visual angle either side of fixation.
 The eccentricity of images was maintained using Gorilla’s scaling tool.
@@
 
-#### General Procedure
+### General Procedure
 
@@
 
-#### Data analysis
+### Data analysis
 
@@
 
-```{r prepdat,echo=FALSE,include=FALSE}
+```{r prepdat,echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
 # remove those without day 2
@@
 short_dat <- dplyr::select(LIs, "subject", "handedness", "footedness", "miles", "porta")
-long_dat <- gather(short_dat, task, response, footedness:porta, factor_key=TRUE)
+long_dat <- tidyr::gather(short_dat, task, response, footedness:porta, factor_key=TRUE)
 # make table
@@
 
-Out of the 441 partiipants recuited via Prolific, `r nrow(LIs_c)` (`r round(nrow(LIs_c)/441*100, 1)`%) complete both sessions.
+Out of the 441 participants recruited via Prolific, `r nrow(LIs_c)` (`r round(nrow(LIs_c)/441*100, 1)`%) complete both sessions.
 We then removed `r nrow(LIs_no)` participants (`r round(nrow(LIs_no)/nrow(LIs_c)*100, 1)`%) from our final sample as their lexTALE score was below 80, indicating that a participant’s level of vocabulary knowledge was below the cut off for lower advanced proficiency.
@@
 
-For each task we adaopted the following data cleaning procedures.
+For each task we adopted the following data cleaning procedures.
 Applying each of these procedures meant that the number of participants left in the final analyses could vary by task.
@@
 
-Next, for the chimeric faces task, we removed partcipants data if they scored lower than 75% on trials in which an identical emotion was shown in each hemisface.
+Next, for the chimeric faces task, we removed participants data if they scored lower than 75% on trials in which an identical emotion was shown in each hemiface.
 These data trimming procedures left data for `r nrow(LIs)-nrow(LIs[LIs$exclude== 1 | LIs$exclude== 12,])` participants on the chimeric faces.
@@
 
-With regards to the rhyme task, we removed participants who scored less than 50% accuracy in either visual field.
-Furthermore, we removed particpants who scored less than 75% accuracy on the catch trials (trials in which there was no rhyme).
+With regards to the rhyme decision task, we removed participants who scored less than 50% accuracy in either visual field.
+Furthermore, we removed participants who scored less than 75% accuracy on the catch trials (trials in which there was no rhyme).
 These trimming procedures left data for `r nrow(LIs[LIs$exRDT== 0,])` participants on the rhyme decision task.
 Seeing as we were removing a relatively high portion of participants, want wanted to check whether this task was yielding reliably results.
-With `r nrow(LIs[LIs$exRDT== 0,])` participants, the test-retest reliability was *r*= `r round(cor(LIs[LIs$exRDT== 0,]$Day1.RDT_RT_LI, LIs[LIs$exRDT== 0,]$Day2.RDT_RT_LI, use = "complete.obs"),3)`, which is smaller than the *r*= 0.65 that we pre-registered as our criteria for a reliable task.
+With `r nrow(LIs[LIs$exRDT== 0,])` participants, the test-retest reliability was *r*= `r round(cor(LIs[LIs$exRDT== 0,]$Day1.RDT_RT_LI, LIs[LIs$exRDT== 0,]$Day2.RDT_RT_LI, use = "complete.obs"),2)`, which is smaller than the *r*= 0.65 that we pre-registered as our criteria for a reliable task.
 Thus, rather than collect further participants so that we would have our pre-registered 300 participants for the rhyme decision task, we chose to end data collection here and analyse the `r nrow(LIs)` participants that we had collected under the assumption that the rhyme decision is not reliable.
@@
 
-##### Calculations of laterality indices
+##### Calculations of laterality indices, reliability, and population level analysis
 
@@
 We specified that we would calculate LIs in two ways: (1) LIs would be calculated in the traditional fashion (100 x (Right – Left)/(Right + Left)); (2) LIs would be calculated in a way that corresponded to a *t*-value.
-This was achieved by conducting a *t*-test for each participant where accuracy response times are the dependent variable and visual field is the independent variable.
+This was achieved by conducting a *t*-test for each participant where accurate response times are the dependent variable and visual field is the independent variable.
 This would estimate the sensitivity to stimuli presented in either visual field and equate to a laterality index.
@@
 
-##### Reliability assessment
-
-For the reliability assessment, we plotted the distributions of LIs and checked them for normality using a Shapiro-Wilk test.
+For the reliability assessment, we checked LIs for normality using a Shapiro-Wilk test.
 As the LIs were non-normal, we used a series of Spearman’s correlations to examine the correlation of coefficients for session 1 versus session 2.
-
-For tasks where the test-retest reliability was above .75 with the existing test length, we reanalysed the data using just the first N trials, reducing N until reliability becomes unsatisfactory (below *r*= .75)^1^.
+For tasks where the test-retest reliability was above .75 with the existing test length, we reanalysed the data using just the first N trials, reducing N until reliability becomes unsatisfactory (below *r*= .65).
 This will allow us to identify the most efficient way to administer these tasks in future, while retaining good reliability.
 However, the full data (including all trials) was used for all subsequent analyses.
-
-##### Population level analysis
-
 To address hypothesis 1 and examine the laterality of each task at the population level, we conducted a series of one-sample *t*-tests for each task using session 1 data.
@@
 
-```{r descriptives, echo=FALSE,include=FALSE}
+```{r descriptives, echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
 # here we present the raw count and RT scores for each task
@@
 
-##### Calculations of laterality indices
-
-Participants' average reaction times on the rhyme decision task are shown in Figure 4 broken down by visual field and session.
+### Calculations of laterality indices, reliability, and population level analysis
+
+Participants' average reaction times on the rhyme decision task are shown in Figure 4 across visual field and session.
 While we do not analyse accuracy, accuracy on the rhyme task was high  (*M*= `r round(mean(RD.short.dat$p.corr, na.rm= TRUE),1)`%, *SD*= `r round(sd(RD.short.dat$p.corr, na.rm= TRUE),2)`%).
-The count of correct answers on the dichotic listening and chimeric faces tasks are also shown in Figure 4, along with the number of sequences gerated by the left and right hand.
+The count of correct answers on the dichotic listening and chimeric faces tasks are also shown in Figure 4, along with the number of sequences generated by the left and right hand.
 LIs were computed in several ways.
@@
 A negative LI indicates left lateralisation at the brain level while a positive LI indicates right lateralisation at the brain level.
-For the rhyme decision task, we then computed an LI using a *t*-test for inidiviaul participants fit to log-tranformed reaction time data.
-For the reamining tasks (which made use of count data) we computed a *z*-statistic for each participant using count data.
+For the rhyme decision task, we then computed an LI using a *t*-test for inidividual participants fit to log-transformed reaction time data.
+For the remaining tasks (which made use of count data) we computed a *z*-statistic for each participant using count data.
 
 ![Figure 4.
-Mean scorees across the two sessions of the behavioural laterality tasks.
+Mean scores across the two sessions of the behavioural laterality tasks.
 (A) Reaction times in each visual field on the rhyme decision tasks.
@@
 For each measure, LIs and their test-retest correlation coefficients are reported in Table 1.
-In addition, Table 1 inlcudes the *p*-values for a one-sampled test and Shapiro-Wlk test for each measure  As the distirbution of each LI was non-normal, spearman's rank correlations were used to estimate the test-retest reliability.
-Confidence intervals were estimated using bootstapping with 10,000 iterations.
-From Table 1, it is clear that we are able to obtain good test-retest reliability when calculating LIs in a typical manner.
-For all but the rhyme decision task, we see good test-retest reliability that exceeds our pre-registered thresold of *r*= .65.
+In addition, Table 1 includes the *p*-values for one-sample *t* and Shapiro-Wilk tests for each measure  As the distribution of each LI was non-normal, spearman's rank correlations were used to estimate the test-retest reliability.
+Confidence intervals were estimated using the *spearman.ci()* function from the RVAideMemoire package (version 0.9-75; Herv, 2020) with 10,000 iterations.
+From Table 1, it is clear that we were able to obtain good test-retest reliability when calculating LIs in a typical manner.
+For all but the rhyme decision task, we see good test-retest reliability that exceeds our pre-registered threshold of *r*= .65.
 Furthermore, LIs for each task across the two sessions are lateralised at the participant level (all *p*s< .001).
 Specifically, the rhyme decision, dichotic listening, and finger tapping tasks were all lateralised to the left hemisphere (showing a left visual field advantage).
-The chimeric faces task was lateralised to the right hemipshere (showing a left visual field advantage).
+The chimeric faces task was lateralised to the right hemisphere (showing a left visual field advantage).
 A highly similar pattern of results emerges for the *t*- and *z*-score measures, where each is lateralised at the population level (all *p*s< .001).
@@
 
-```{r make.combotable,echo=FALSE,include=TRUE}
+```{r make.combotable,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
 #Make a table to summarise results for all 4 tasks
@@
 
-```{r plots.function, echo=FALSE,include=TRUE}
+```{r plots.function, echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
 # Adam's code here turned into generic function
@@
 NB need to check polarity of tasks - also oddity re RDT distributions needs checking, and we said we'd do log transform , not sure we did--->
-```{r do.plots,LI,echo=FALSE,include=TRUE}
+```{r do.plots,LI,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
 # RHYME DECISION
@@
 
-```{r do.plot.t,echo=FALSE,include=TRUE}
+```{r do.plot.t,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
 # RHYME T-TEST
@@
 
-Table 1. LIs and test-retest correlations for each taks.
-```{r Table1,echo=FALSE}
-#Needs more formatting. For word output may need to use flextable. See https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html
-kable(combotable) %>%
-  kable_styling(c("bordered")) %>%
-  add_header_above(c(" " = 1, "Typical LI" = 8, "t/z-score" = 8))
+Table 1. LIs and test-retest correlations for each task.
+```{r Table1,echo=FALSE, comment=FALSE, warning=FALSE, comment=FALSE, warning=FALSE}
+library(flextable)
+library(officer)
+myft <- flextable(
+  head(combotable))
+myft
 ```
 Note.
-LIs are transformed such that a negtaive value equates to a left lateralisation at the brain level.
+LIs are transformed such that a negative value equates to a left lateralisation at the brain level.
 A positive value equates to right lateralisation at the brain level.
@@
 Moving forward, we chose to use LIs calculated in a typical manner to address each remaining research question.
-To dtermine the smallest number of trials needed to achieve a test-retest reliabilty of *r*= .75, we recaclauted LIs by dropping one trial at a time for each participant...
-
-##### Atypicality
-
-We defined 'atypical laterality' as laterality that is in the opposite direction to the majority, and which has an absolute z-score or t-score value greater than 1.96.
-These z and t-scores indicate whether there is significant bias to the left or right within an individual.
+To determine the smallest number of trials needed to achieve a test-retest reliability of *r*= .65, we recalculated LIs for the dichotic listening and chimeric faces tasks using the range of trials in which different emotions/sounds were presented to each hemiface/ear.
+The absolute test-retest correlation coefficients are shown in Figure 6.
+A minimum of 60 trials was required to achieve a test-retest correlation of *r*= .65 using the dichotic listening task.
+However, as indicated in Figure 6, even with 100 trials the reliability estimate still occasionally approaches this 0.65 threshold.
+Taking this lower number of trials, it would be able to present only two blocks to achieve this threshold of reliability for the dichotic listening task.
+For the chimeric faces task, a minimum of 25 trials is necessary to achieve a test-retest reliability of *r*= .65.
+Recall that all possible combinations of the four emotions would result in 12 stimuli in which a different emotion is shown in each hemiface and 4 where the same emotion appears in each hemiface.
+Three sets of all possible combinations would then be sufficient to yield a good test-retest reliability here (equating to 1.5 blocks of the chimeric faces in the current study).
+
+![Figure 6.
+Test-retest correlations for (A) the dichotic listening task and (B) the chimeric faces task when using a variable number of trials to calculate laterality indices.
+The red dashed line marks the test-retest correlation r= .65.
+The correlations coefficients are shown as absolute values.](all.cor.plot.png)
+
+### Atypicality
+
+We defined 'atypical laterality' as laterality that is in the opposite direction to the majority of participant, and which has an absolute *z*/*t*-score value greater than 1.96.
 Based on our pre-registered analysis, we divided left-handers into two groups (1) those with a score of -90 or below on the Edinburgh Handedness Inventory (c.f.
 Mazoyer et al., 2014), (2) the remaining left handers.
-While our analyses inlcude only right handers and extreme left handers, we include the remaining left-handers in Table 2 for completeness.
-
-```{r makecontingencies,echo=FALSE}
+While our analyses include only right handers and extreme left handers, we include the remaining left-handers in Table 2 for completeness.
+
+```{r makecontingencies,echo=FALSE, comment=FALSE, warning=FALSE}
 #We start by using z-scores to make a 3-way categorisation into:
 #1 sig bias in typical direction; 2: NS bias 3: in atypical direction with sig bias.
-LIs$rhyme1.cut <- cut(LIs$Day1.RDT_p.corr_LI, 
+LIs$rhyme1.cut <- cut(LIs$logt1, 
                        breaks = c(-Inf, -1.96, 1.96, Inf), 
-                       labels = c("TypLat", "NonLat", "AtypLat"), 
+                       labels = c("RD TypLat", "RD NonLat", "RD AtypLat"), 
                        right = FALSE)
@@
                        breaks = c(-Inf, -1.96, 1.96, Inf), 
-                       labels = c("TypLat", "NonLat", "AtypLat"), 
+                       labels = c("DL TypLat", "DL NonLat", "DL AtypLat"), 
                        right = FALSE)
@@
                        breaks =  c(-Inf, -1.96, 1.96, Inf), 
-                       labels = c("AtypLat", "NonLat", "TypLat"), 
+                       labels = c("CF AtypLat", "CF NonLat", "CF TypLat"), 
                        right = FALSE)
@@
 mytab2<-data.frame(mytab2)
-# combine tables for printing
-mytab3 <- cbind(tabr, tabd, tabc)
 # now do chi-squared tests
@@
 chim.chi <- chisq.test(tabc)
+# combine tables for printing
+mytab3 <- cbind(tabr, tabd, tabc)
 # BUT these don't actually run the planed tests. We run these by removing lefties here
@@
 
-Table 2. Counts of laterality patterns for right-, left-, and extreme geft-handers acorss the Rhyme decision, Dichotic listening, and Chimeric faces tasks.
-```{r Table 2}
-# plot group table
-mytab3%>%
-  kable() %>%  
-  kable_styling(c("bordered")) %>%
-  add_header_above(c(" " = 1, "Rhyme decision" = 3, "Dichotic listening" = 3, "Chimeric faces"= 3))
+Table 2. Frequencies of laterality typicality for right-, left-, and extreme left-handers across the Rhyme decision, Dichotic listening, and Chimeric faces tasks.
+```{r Table 2,echo=FALSE, comment=FALSE, warning=FALSE}
+mytab4 <- as.data.frame(mytab3)
+mytab4$` ` <- c("Extreme left", "Left", "Right")
+myft2 <- flextable(
+  head(mytab4),
+  col_keys = c(" ", 
+               "RD TypLat", "RD NonLat", "RD AtypLat",
+               "DL TypLat", "DL NonLat", "DL AtypLat",
+               "CF TypLat", "CF NonLat", "CF AtypLat"))
+myft2
 ```
@@
 
-```{r contingency2tasks,echo=FALSE,include=TRUE}
+```{r contingency2tasks,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
 tab2task <- table(LIs$dichotic1.cut,LIs$chimeric1.cut)
@@
 
-##### Relationship between language and non-language LIs
-
-```{r LI-link,echo=FALSE,include=TRUE}
+### Relationship between language and non-language LIs
+
+```{r LI-link,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
 # make sure we have good cases
@@
 Using data from session 1, we examined the correlation between LIs on the dichotic listening and chimeric faces task to examine whether the perceptual biases on these tasks are governed by a single laterality factor.
-A spearman's rank order correlation indicated a extrenely small, non-singificant relationship, *rs*= `r round(DL.CF.est$estimate, 2)`, *95% CI* [`r round(DL.CF$conf.int[1], 2)`, `r round(DL.CF$conf.int[2], 2)`] *p*= `r round(DL.CF.est$p.value[1], 3)`.Based on this analysis, there is no evidence of a relationship between laterality indicies on the dichotic listening and chimeric faces tasks.
+A spearman’s rank order correlation indicated a extremely small, non-significant relationship, *rs*= `r round(DL.CF.est$estimate, 2)`, *95% CI* [`r round(DL.CF$conf.int[1], 2)`, `r round(DL.CF$conf.int[2], 2)`] *p*= `r round(p.adjust(DL.CF.est$p.value[1], n=2, method= "bonferroni"), 3)`.Based on this analysis, there is no evidence of a relationship between laterality indices on the dichotic listening and chimeric faces tasks.
 To infer the extent to which our data reflect a null relationship, as opposed to a Type II error, we computed a Bayes factor (Rouder, Morey, Speckman, & Province, 2012).
-We computed the Bayes factor in R using the BayesFactor package (Version 0.9.12–4.2; Morey & Rouder, 2018) with 10,000 Monte Carlo iterations.
-We assumed the default Cauchy prior for the effect size.
+We computed the Bayes factor in R using the *correlationBF()* functions from the BayesFactor package (Version 0.9.12–4.2; Morey & Rouder, 2018) with 10,000 Monte Carlo iterations.
+We assumed the default prior (*r*= 1/3) for the effect size.
 The Bayes factor for the alterative model, as opposed to the null model, was .150.
 According to Jeffreys' (1961) inference criteria, this provides moderate evidence for the null hypothesis.
-To validate these findings, we repeated these analysis with session 2 data.
-Again the correlation was small and not significantly different from zero *rs*= `r round(DL.CF.est.2$estimate, 2)`, *95% CI* [`r round(DL.CF.2$conf.int[1], 2)`, `r round(DL.CF.2$conf.int[2], 2)`] *p*= `r round(DL.CF.est.2$p.value[1], 3)`.
+To validate these findings, we repeated these analysis with session 2 data and controlled for multiple comparisons.
+Again the correlation was small and not significantly different from zero *rs*= `r round(DL.CF.est.2$estimate, 2)`, *95% CI* [`r round(DL.CF.2$conf.int[1], 2)`, `r round(DL.CF.2$conf.int[2], 2)`] *p*= `r round(p.adjust(DL.CF.est.2$p.value[1], n= 2, method= "bonferroni"), 3)`.
 Again the Bayes Factor favoured the null hypothesis, *BF*= .138.
 
-##### Exploratory analyses
-
-```{r fingerEHI,echo=FALSE,include=TRUE }
+### Exploratory analyses
+
+#### Finger tapping and Edinburgh Handedness Inventory
+
+```{r fingerEHI,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
 spearCI <- RVAideMemoire::spearman.ci(LIs$Day1_finger_LI, -LIs$index_EHI, nrep = 10000, conf.level = 0.95)
@@
 
-The finger tapping task produced reliable results, but the extent to which it correlated with existing measures of handedness is not clear.
-Thus, we used a spearman's rank correlation to obtain an estimate of the relationship between LIs produced on the finger tapping task and the Edinburgh Handedness Inventory using session 1 data.
-The relationship between the tasks was moderate,*rs*= `r round(speartest$estimate, 2)`, *95% CI* [`r round(spearCI$conf.int[1], 2)`, `r round(spearCI$conf.int[2], 2)`] *p*= `r round(speartest$p.value[1], 3)`.
-Thus it indicates that the task likely represents a component of hand preference in conjunction with some aspect of motor planning or sequencing.
-
-## Notes
-
-1.
-While we pre-registered that we would reduce the number of trails until *r*= .65 for tasks where the initial test-retest relaibility was *r*= .75, we chose to reduce the trials down until *r*= .75.
-We decided to do this to maintain a higher level of relaibaility when using shorter version of the battery in inidividual differences research.
-
+To examine the relationship between the finger tapping task with an existing measures of handedness, we used a spearman's rank correlation to obtain an estimate of the relationship between LIs produced on the finger tapping task and the Edinburgh Handedness Inventory using session 1 data.
+The relationship between the tasks was moderate, *rs*= `r round(speartest$estimate, 2)`, *95% CI* [`r round(spearCI$conf.int[1], 2)`, `r round(spearCI$conf.int[2], 2)`],  *p*< 0.001.
+Thus it indicates that the task likely represents captures some variance in hand preference.
+It is, however, also likely that the task goes beyond this and captures motor planning and coordination.
+We say more about this in the discussion.
+
+#### Handedness measurements and their association with laterality
+
+```{r hand_explore, echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
+hand_cors <- data.frame(matrix(ncol = 5, nrow = 1))
+  colnames(hand_cors) <- c(" ", "DL 1", "CF 1", "DL 2", "CF 2")
+# add rows
+# row 1
+hand_cors[1, 1] = "EHI"
+hand_cors[1, 2] = round(cor(LIs$Day1_DL_acc_LI, LIs$index_EHI, use= "complete.obs"), 2)
+hand_cors[1, 3] = round(cor(LIs$Day1_CF_acc_LI, LIs$index_EHI, use= "complete.obs"), 2)
+hand_cors[1, 4] = round(cor(LIs$Day2_DL_acc_LI, LIs$index_EHI, use= "complete.obs"), 2)
+hand_cors[1, 5] = round(cor(LIs$Day2_CF_acc_LI, LIs$index_EHI, use= "complete.obs"), 2)
+# row 2
+hand_cors[2, 1] = "FT"
+hand_cors[2, 2] = round(cor(LIs$Day1_DL_acc_LI, LIs$Day1_finger_LI, use= "complete.obs"), 2)
+hand_cors[2, 3] = round(cor(LIs$Day1_CF_acc_LI, LIs$Day1_finger_LI, use= "complete.obs"), 2)
+hand_cors[2, 4] = round(cor(LIs$Day2_DL_acc_LI, LIs$Day2_finger_LI, use= "complete.obs"), 2)
+hand_cors[2, 5] = round(cor(LIs$Day2_CF_acc_LI, LIs$Day2_finger_LI, use= "complete.obs"), 2)
+# table 
+corrTabs <- flextable(
+  head(hand_cors))
+# same scale
+LIs$negEHI <- -LIs$index_EHI
+# compare
+cocor.result <- cocor(~Day1_DL_acc_LI + negEHI | Day1_DL_acc_LI + Day1_finger_LI, LIs)
+cocor.result2 <- cocor(~Day1_CF_acc_LI + negEHI | Day1_CF_acc_LI + Day1_finger_LI, LIs)
+cocor.result3 <- cocor(~Day2_DL_acc_LI + negEHI | Day2_DL_acc_LI + Day2_finger_LI, LIs)
+cocor.result4 <- cocor(~Day2_CF_acc_LI + negEHI | Day2_CF_acc_LI + Day2_finger_LI, LIs)
+
+# ASSOCIATIONS
+
+# using self report
+# dichotic
+self_DL <- table(LIs$handedness,LIs$dichotic1.cut)
+self_DL.chi <- chisq.test(self_DL)
+# chimeric
+self_CF <- table(LIs$handedness,LIs$chimeric1.cut)
+self_CF.chi <- chisq.test(self_CF)
+# combinw
+self_hand <- as.data.frame(cbind(self_DL, self_CF))
+  self_hand$` ` <- rownames(self_hand)
+# table
+self_hand_Tabs <- flextable(
+  head(self_hand),
+  col_keys = c(" ", "DL TypLat", "DL NonLat", "DL AtypLat", "CF TypLat", "CF NonLat", "CF AtypLat"))
+
+# using finger tapping
+LIs$finger1.cut <- cut(LIs$Day1_finger_LI, 
+                       breaks = c(-Inf, 0, Inf), 
+                       labels = c("Right", "Left"), 
+                       right = FALSE)
+# dichotic
+finger_DL <- table(LIs$finger1.cut,LIs$dichotic1.cut)
+finger_DL.chi <- chisq.test(finger_DL)
+# chimeric
+finger_CF <- table(LIs$finger1.cut,LIs$chimeric1.cut)
+finger_CF.chi <- chisq.test(finger_CF)
+# combinw
+finger_hand <- as.data.frame(cbind(finger_DL, finger_CF))
+    finger_hand$` ` <- rownames(finger_hand)
+# table
+finger_hand_Tabs <- flextable(
+  head(finger_hand),
+  col_keys = c(" ", "DL TypLat", "DL NonLat", "DL AtypLat", "CF TypLat", "CF NonLat", "CF AtypLat"))
+
+# using EHI
+LIs$EHI.cut <- cut(LIs$index_EHI, 
+                       breaks = c(-Inf, -40, 40, Inf), 
+                       labels = c("Left", "Ambidextrous", "Right"), 
+                       right = FALSE)
+# dichotic
+EHI_DL <- table(LIs$EHI.cut,LIs$dichotic1.cut)
+EHI_DL.chi <- chisq.test(EHI_DL)
+# chimeric
+EHI_CF <- table(LIs$EHI.cut,LIs$chimeric1.cut)
+EHI_CF.chi <- chisq.test(EHI_CF)
+# combinw
+EHI_hand <- as.data.frame(cbind(EHI_DL, EHI_CF))
+    EHI_hand$` ` <- rownames(EHI_hand)
+# table
+EHI_hand_Tabs <- flextable(
+  head(EHI_hand),
+  col_keys = c(" ", "DL TypLat", "DL NonLat", "DL AtypLat", "CF TypLat", "CF NonLat", "CF AtypLat"))
+```
+
+Our planned analysis of the association between handedness and (a)typical lateralisation was based solely on using the Edinburgh Handedness Inventory Scores.
+We additionally examined how different strategies of coding handedness would influence the strength of this association on the dichotic listening and chimeric faces tasks using session 1 and session 2 data.
+We assessed these associations in the following ways: (1) used the Edinburgh Handedness Inventory to code left, ambidextrous, and right handedness, (2) using self-reported handedness, and (3) using performance on the finger tapping task to code handedness.
+To examine the influence of the three handedness groups (left, ambidextrous, right), we defined handedness as: scores of +40 or above denoting right-handedness, −40 or below denoting left handedness, and scores in between denoting ambidexterity (follwoing Whitehouse & Bishop, 2009).
+The frequencies of (a)typicality are shown for left-, ambidextrous, and right-handers are shown in Table 3.
+A chi-squared test of association indicated that there was a significant association between handedness and the dichotic listening, chi squared = `r round(EHI_DL.chi$statistic,2)`, d.f.
+= 4, *p* = `r round(p.adjust(EHI_DL.chi$p.value, method= "bonferroni", n=2),3)`; and chimeric faces, chi squared = `r round(EHI_CF.chi$statistic,2)`, d.f.
+= 4, *p* = `r round(p.adjust(EHI_CF.chi$p.value, method= "bonferroni", n=2),3)`.
+
+Table 3. Frequencies of laterality typicality for right-, ambidextrous, and left-handers across the Dichotic listening and Chimeric faces tasks.
+```{r table 3, echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
+EHI_hand_Tabs
+```
+Note. TypLat: lateralised in the expected direction, Nonlat: no lateralisation, and AtypLat: show lateralisation in the opposite direction.
+
+From Table 4, there is evidence of increased atypicality for left-handers on the dichotic listening when using self-reported handedness.
+The association appears even stronger when considering the chimeric faces task.
+Chi-squared tests of association confirmed the association between handedness and atypicality as statistically significant on the dichotic listening, chi squared = `r round(self_DL.chi$statistic,2)`, d.f.
+= 2, *p* = `r round(p.adjust(self_DL.chi$p.value, method= "bonferroni", n=2),3)`; and chimeric faces, chi squared = `r round(self_CF.chi$statistic,2)`, d.f.
+= 1, *p* = `r round(p.adjust(self_CF.chi$p.value, method= "bonferroni", n=2),3)`.
+
+Table 4. Frequencies of laterality typicality for self-reported right- and left-handers across the Dichotic listening and Chimeric faces tasks.
+```{r table 4, echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
+self_hand_Tabs
+```
+Note. TypLat: lateralised in the expected direction, Nonlat: no lateralisation, and AtypLat: show lateralisation in the opposite direction.
+
+We then assessed the association between handedness, defined using the finger tapping task, and (a)typical lateralisation.
+Here a negative value was coded as a preference for right handedness and a positive value as left handedness.
+From Table 5, the association between left-handedness and atypical laterality on the dichotic listening task was not significant, chi squared = `r round(finger_DL.chi$statistic,2)`, d.f.
+= 2, *p* = `r round(p.adjust(finger_DL.chi$p.value, method= "bonferroni", n=2),3)`.
+There was, however, an association between finger tapping defined handedness and atypicality for chimeric faces, chi squared = `r round(finger_CF.chi$statistic,2)`, d.f.
+= 2, *p* = `r round(p.adjust(finger_CF.chi$p.value, method= "bonferroni", n=2),3)`.
+Taken together, these analyses demonstrate a clear association between handedness and atypical lateralisation on both tasks.
+The chi-squared coefficient was largest when using the Edinburgh Handedness Inventory, indicating that this measure is likely the most suited to identifying those who are likely to show (a)typical lateralisation.
+
+Table 5. Frequencies of laterality typicality for finger tapping defined right- and left-handers across the Dichotic listening and Chimeric faces tasks.
+```{r table 5, echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
+finger_hand_Tabs
+```
+Note. TypLat: lateralised in the expected direction, Nonlat: no lateralisation, and AtypLat: show lateralisation in the opposite direction.
+
+We additionally examined the strength of the relationship of LIs on the dichotic listening and chimeric faces tasks with LIs calculated using the Edinburgh Handedness Inventory and the finger tapping task.
+A series of Spearman’s rank correlations indicated that the relationship between LIs on each task were largely small (see Table 3).
+A series on z-tests were then computed using absolute correlation coefficients to examine whether the Edinburgh Handedness Inventory or finger Tapping were better correlated with LIs on the dichotic listening and chimeric faces.
+We controlled for multiple comparisons using a Bonferroni adjustment.
+For session 1 data, the Edinburgh handedness Inventory was more strongly correlated with the chimeric faces, *z*= `r round(as.numeric(as.htest(cocor.result2)$pearson1898[1]), 2)`, *p*= `r round(p.adjust(as.htest(cocor.result2)$pearson1898[3], n= 2, method = "bonferroni"),3)`, but not the dichotic listening, *z*= `r round(as.numeric(as.htest(cocor.result)$pearson1898[1]), 2)`, *p*= `r round(p.adjust(as.htest(cocor.result)$pearson1898[3], n= 2, method = "bonferroni"),3)`, task.
+For session 2 data, an identical pattern of results emerge whereby the Edinburgh handedness Inventory was more strongly correlated with the chimeric faces, *z*= `r round(as.numeric(as.htest(cocor.result4)$pearson1898[1]), 2)`, *p*= `r round(p.adjust(as.htest(cocor.result4)$pearson1898[3], n= 2, method = "bonferroni"),3)`, but not the dichotic listening, the dichotic listening, *z*= `r round(as.numeric(as.htest(cocor.result3)$pearson1898[1]), 2)`, *p*= `r round(p.adjust(as.htest(cocor.result3)$pearson1898[3], n= 2, method = "bonferroni"),3)`, task.
+
+Table 6.
+Correlation coefficients between LIs on the dichotic listening (DL) and chimeric faces (CF) tasks and the Edinburgh Handedness Inventory and finger tapping task using session 1 and session 2 data.
+```{r table 6, echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
+corrTabs
+```
+
+#### Vocabulary knowledge, footedness, sightedness and laterality
+
+```{r exploreA, echo= FALSE, comment=FALSE, warning=FALSE}
+DL.lex<-RVAideMemoire::spearman.ci(interR$Day1_DL_acc_LI, interR$lexTALE, nrep = 10000, conf.level = 0.95)
+DL.lex.est<-cor.test(interR$Day1_DL_acc_LI, interR$lexTALE, method = "spearman")
+DL.lex1<-RVAideMemoire::spearman.ci(interR$Day2_DL_acc_LI, interR$lexTALE, nrep = 10000, conf.level = 0.95)
+DL.lex.est1<-cor.test(interR$Day2_DL_acc_LI, interR$lexTALE, method = "spearman")
+
+# footedness
+foot.dat <- interR[interR$footedness != "Don't know",]
+foot.dat$footedness <- as.factor(foot.dat$footedness)
+#  levels(droplevels(foot.dat$footedness))
+DL.ft.t.1 <- t.test(foot.dat$Day1_DL_acc_LI~ foot.dat$footedness)
+DL.ft.t.2 <- t.test(foot.dat$Day2_DL_acc_LI~ foot.dat$footedness)
+CF.ft.t.1 <- t.test(foot.dat$Day1_CF_acc_LI~ foot.dat$footedness)
+CF.ft.t.2 <- t.test(foot.dat$Day2_CF_acc_LI~ foot.dat$footedness)
+
+# sightedness
+foot.dat <- interR[interR$miles != "Don't know",]
+foot.dat$miles <- as.factor(foot.dat$miles)
+#  levels(droplevels(foot.dat$miles))
+DL.st.t <- t.test(foot.dat$Day1_DL_acc_LI~ foot.dat$miles)
+DL.st.t <- t.test(foot.dat$Day2_DL_acc_LI~ foot.dat$miles)
+CF.st.t <- t.test(foot.dat$Day1_CF_acc_LI~ foot.dat$miles)
+CF.st.t <- t.test(foot.dat$Day2_CF_acc_LI~ foot.dat$miles)
+```
+
+While the role of the LexTALE in the current research was largely for the purpose of data exclusion, we had planned to explore correlations between LexTALE score and the laterality index on the Rhyme Decision task.
+However, as the rhyme task was not overly reliable, we chose to examine the relationship between the LexTALE and the dichotic listening task instead.
+From session 1 data, after correcting for multiple comparisons, the correlation between the LIs on the dichotic listening and LexTALE score was small and non-significant, *rs*= `r round(DL.lex.est$estimate, 2)`, *95% CI* [`r round(DL.lex$conf.int[1], 2)`, `r round(DL.lex$conf.int[2], 2)`] *p*= `r round(p.adjust(DL.lex.est$p.value[1], n= 2, method= "bonferroni"), 3)`.
+An identical pattern of results was obtained using session 2 data, *rs*= `r round(DL.lex.est1$estimate, 2)`, *95% CI* [`r round(DL.lex1$conf.int[1], 2)`, `r round(DL.lex1$conf.int[2], 2)`] *p*= `r round(p.adjust(DL.lex.est1$p.value[1], n= 2, method= "bonferroni"), 3)`.
+
+We then examined whether footedness was related to the laterality indices on both the dichotic listening and chimeric faces tasks.
+Using session 1 data, LIs on the dichotic listening task did not differ between left and right footers, *t*(275.89)= .07, *95% CI* [-3.29, 11.15], *p*= `r p.adjust(0.284, n= 2, method= "bonferroni")`, nor did LIs on the chimeric faces, *t*(219.58)= -1.91, *95% CI* [-12.78, .20], *p*= `r p.adjust(0.058, n= 2, method= "bonferroni")`.
+Analysis of session 2 data yielded a similar pattern of findings.
+LIs on the dichotic listening, *t*(243.09)= 1.87, *95% CI* [-.39, 14.63], *p*= `r p.adjust(0.063, n= 2, method= "bonferroni")`, and the chimeric faces, *t*(197.01)= -2.03, *p*= `r p.adjust(0.044, n= 2, method= "bonferroni")`, tasks did not differ for right and left footers.
+We then completed the same analyses but using sightedness in the place of footedness.
+As participants self-reported sightedness was always the same on the Miles and Porta test, we analysed data using only the Porta test.
+Using session 1 data, LIs on the dichotic listening task did not differ with sightedness, *t*(176.49)= .02, *95% CI* [-7.49, 7.67], *p*= `r p.adjust(0.982, n= 2, method= "bonferroni")`, yet chimeric faces LIs were significantly less lateralised to the right hemisphere in those with left eye dominances, *t*(-3.06)= -3.06, *95% CI* [-18.96, -4.07], *p*= `r p.adjust(0.003, n= 2, method= "bonferroni")`.
+Analysis of session 2 data yielded a similar pattern of findings for the dichotic listening task, *t*(144.51)= .75, *95% CI* [-17.99, -21.11], *p*= `r p.adjust(0.455, n= 2, method= "bonferroni")`.
+However, LIs on the chimeric faces tasks did not differ between left and righ eye dominant groups here, *t*(124.25)= -1.89, *95% CI* [-17.48, .39], *p*= `r p.adjust(0.061, n= 2, method= "bonferroni")`.
+
+## Discussion
+
+With the advancement of online behaviour research methods, it has become possible to test large samples of participants outside of the lab.
+In the current study we introduced an online behavioural laterality battery that was administered to a relatively large sample of right- and left-handers.
+In addition to completing self-report measures of handedness, footedness, and sightedness, participants completed a visual half-field rhyme decision task, a dichotic listening task, a chimeric faces task, and a finger tapping task twice.
+Thus, we were able to examine the test-retest reliability of each task.
+We also examined the association between handedness and atypical lateralisation and examined the intercorrelations between LIs on the dichotic listening and chimeric faces tasks.
+We discuss each of these points in turn.
+
+Each behavioural task yielded a hemispheric advantage in the expected direction.
+However, the extent to which they were reliable differed substantially between tasks.
+The rhyme decision task yielded the poorest test-retest reliability.
+Compared to early studies (e.g.
+Krach et al., 2006) the rhyme decision task used here was far more reliable but, compared to other visual half-field tasks, it was far from optimal.
+van der Haegen and Brysbaert (2018) have reported that under lab conditions that visual half-field paradigms can yield test-retest reliabilities of *r*= .77 when using pictures and *r*= .83 when using words.
+While the two studies both included a sufficiently large number of trials, there are several potential reasons why the rhyme decision task failed to achieve an acceptable level of test-retest reliability in the current study.
+One key difference between van der Haegen and Brysbaert’s and our study was that their language based visual half-field tasks involved naming.
+Speech production is highly lateralised (Kosslyn, 1987) and requiring speech production will no doubt yield a hemispheric advantage.
+As the rhyme task was run online, it would be extremely difficult to use the onset of speech to trigger the end/start of a trial let alone retain and code these response manually.
+Therefore, we opted to use a rhyme component that should theoretically engage subvocal processes associated with phonological processing.
+Yet, this did not yield such reliable results.
+Therefore it may be of use for future implementations to involve some active production component.
+
+While the rhyme decision task was not so reliable, the dichotic listening task yielded surprising results.
+Not only did participants show a right-ear advantage at the population level, the task exceeded our pre-registered reliability cut-off of *r*= .65.
+This was, perhaps, not unexpected though as the paradigm has been implemented via the iDichotic application (Bless et al., 2013).
+What is more astonishing is the fact test-retest correlation of the dichotic listening task is identical to that using the iDichotic application (both test-retest coefficients= .78) despite the current study using a non-dedicated platform.
+Furthermore, the reliability in the current study exceeds that of the dichotic listening task when run under lab conditions (Bless et al., 2013).
+However, this reliability is somewhat less than the estimates of reliability produced when completing a modified version of the task under lab conditions (Westerhausen & Samuelsen, 2020).
+What is particularly interesting is the observation that with as few as 60 trials, the current implementation of the dichotic listening task yields a test-retest correlation that than *r*= .65, meaning that those who wish to use this paradigm in the future could use a smaller number of trials and add additional tasks on that are important to their ongoing research questions.
+Of course, it would be necessary to screen for adequate and appropriate headphone as we did.
+
+Like the dichotic listening task, the chimeric faces task yielded good test-retest reliability.
+While previous endeavours to assess split-half reliability of the task have yielded strong correlations (i.e.
+Levy et al., 1983), we report a strong test-retest correlation, What is particularly novel about this finding is that we were able to report this for online behavioural research methods.
+Furthermore, the reliability of this task can be astonishingly high with very few trials.
+Our analysis indicated that only 25 trials would be necessary to achieve a test-retest reliability of *r*= .65.
+We believe that the reason for the success of the chimeric faces task online is the task’s simplistic nature.
+As implemented, it requires participants to report the emotion they saw and not have to make judgements which involve comparing representations of emotions as the two faces version might.
+
+We also introduced a novel finger tapping task, which we designed to quantify handedness behaviourally.
+Surprisingly, this simple task produced reliable results.
+Not only was there a right hand advantage at the population level, the LIs produced on this task were reliable.
+Furthermore, our exploratory analyses indicated that this correlated reasonably well with the Edinburgh handedness inventory.
+Yet the Edinburgh Handedness Inventory did seem better equipped to categorise atypical lateralisation.
+This suggests that while the finger tapping task does capture some aspect of hand preference it does also capture something beyond this.
+It is likely that  the task additionally measures the planning and execution of motor coordination.
+However, this remains speculative at current and further work will be needed to clarify the aspects that this task measures beyond simple hand preference and to examine the extent to which LIs resemble that from tasks such as the pegboard task (e.g.
+Hodgson, Hirst, & Hudson, 2016).
+Future work should also examine how other factors such as complexity of key combinations and reach through the midline influence the task.
+
+In addition to assessing lateralisation of the task using traditional LIs, we also examined the potential to use individual *z*/*t*-scores to quantify lateralisation on each task.
+While the *t*-test for the rhyme task does not seem warranted for further use, the *z*-scores for the remaining tasks closely resembled that of the traditional LI.
+This method of calculating Lis is perhaps more useful for those interested in looking atypical lateralisation.
+Rather than selecting those who fall either side of an arbitrary cut-off, *z*-scores allow researchers to assess whether participants are consistently and statistically lateralised in either a typical or atypical manner.
+Let us consider an example to emphasise their importance.
+A researcher plans to examine the influence of atypical language lateralisation on education outcomes and would like to compare those who show strong typicality and those who show strong atypicality.
+Before even meeting the participants for testing, the battery described in the current study could be administered and then those who have *z*-scores > 1.96 or <1.96 could be invited into the lab as they appeared to be significantly lateralised on a behavioural task.
+This has the potential to save time and effort screening participants in a lab setting only to find that they do not meet study inclusion criteria.
+
+While the primary purpose of the study was to validate the battery and establish its test-retest reliability, we addressed two theoretically motivated questions related to handedness and atypicality and intercorrelations between the dichotic listening and chimeric faces task.
+In our pre-registered analysis of handedness, we categorised participants into right,-, left-, and extreme left-handers.
+The data are consistent in showing that extreme left-handers showed increased atypicality on both the dichotic listening and chimeric faces task.
+Mazoyer et al.
+(2014) reported a relationship between strength of lateralisation and hand preference., which was largely driven by the existence of a group with strong left hand preference and strong right hemisphere dominance.
+They further claim that this right hemisphere dominance is almost exclusively present in left-handers.
+We, however, did observe a portion of right-handers with right language dominance on the dichotic listening task—regardless of how we define handedness.
+While issues such as sample size and bias between studies likely represent this, it does further question the whether there is a common mechanism that determines handedness and lateralisation.
+
+A common mechanism that codetermines lateralisation on an array of tasks appears a somewhat hotly debated topic in the laterality literature.
+Several hypotheses have argued that LIs on one task should correlate with another (Behrmann & Plaut, 2015; Cook, 1984; Hellige, 1983; Kosslyn, 1987).
+However, several studies have report no such correlation, and if they do these are often small (Bradzakova-Trajkov et al., 2010; Bryden et al.,1983; Rosch et al., 2012; Whitehouse & Bishop, 2009; van der Haegen & Brysbaert, 2018).
+Data from the current work supports a view in which laterality one each task is independent.
+Not only was our study well powered to examine this relationship, it included a large sample of left-handers who show increased variability in their LIs.
+Specifically, we reported that the correlation between LIs on the dichotic listening and chimeric faces was extremely close to zero and Bayes Factor analysis ruled out the possibility that this was due to a lack of statistical power.
+Thus, we argue that lateralisation is driven by a series of independent biases, which, according to this view, specifies that there is a population bias for left hemisphere language dominance.
+By contrast, there is a population bias for right hemisphere dominance for the processing of faces.
+However, these biases reflect probabilities elating to a combination of multiple independent causal sources, which may be genetic, biological or environmental in origin.
+
+In a series of exploratory analyses we also examined whether sightedness or footedness related to the strength of laterality indices.
+The data were rather consistent in showing that these measures were not predictive of differences in laterality indices in those who report being right- or left-footed/eye dominant.
+Again, this echoes the finding that lateralisation on one function does not determine that on another.
+
+Before moving on to our concluding remarks, it is important to consider the future of online research methods to examine lateralisation.
+On the whole, the results pave a way for a promising future of online data collection to answer several questions with regards to cerebral lateralisation.
+It seems that when we rely on count data (as did the dichotic listening, chimeric faces, and finger tapping tasks) we can obtain reliable results.
+The success of the dichotic listening and chimeric faces task is likely to stem from the use of bilateral stimuli presentation, which results in the two hemispheres competing.
+The story, however, for tasks depending on reaction times for the calculation of LIs seems slightly less optimistic.
+This is likely to result from the increased noise associated with different experimental setups across participants when using typical paradigms.
+Furthermore, while we can format images to be consistent across displays, we cannot control participants’ behaviour.
+Of participant interest to laterality studies is that we cannot control viewing distance across participants, which has several implications for visual half-field paradigms.
+Outside of the lab, participants can freely move their head which will result in the size of images changing on the retina.
+Furthermore,  thing makes it incredibly difficult to ensure images are shown in foveal or parafoveal vision.
+One avenue that might be worth exploring is the extent to which participants can judge images presented in peripheral vision as even with head movements it is unlikely that the change in movements will result in the image shifting between distinct regions on the retina.
+This increased difficultly in processing of information in this region may also result in stronger visual field advantages (c.f.
+Hunter and Brysbaert, 2008).
+
+To conclude, the present study introduced an online behaviour laterality battery that can be used to examine laterality for speech perception, face emotionality, and manual preference with good test-retest reliability.
+The battery of tasks was administered to a large sample of participants with relative ease, and enabled us to examine several questions related to lateralisation.
+We find evidence of increased atypical lateralisation in left- relative to right-handers.
+However, we find no indication that the laterality indices on one task are dependent on another and argue that the causes of lateralisation on one task should be seen as stemming from multiple independent sources.
+
+\newpage
+## Data accessibility
+All analysis scripts and anonymised data are avaiable on the Open Science Framework (https://osf.io/qsxcv/).
+
+\newpage
+## Competing Interests
+We declare we have no competing interestes.
+
+\newpage
+## Funding 
+This study was supported by an Advanced Grant from the European Research Council (694189).
+
+\newpage
+## Acknowledgements
+We would like to thank Dr Mike Burt (University of Durham, Uk) for providing stimuli for the chimeric faces task.
+We would also like to thank Professor Kenneth Hugdahl (University of Bergen, Norway) for providing stimuli for the dichotic listening task.
+Portions of this work were presented at the Online Meeting of the Expeirmental Psychology Society in July 2020.
+
+\newpage
 ## Supplemental Material
 
-We had originally intended to examine the relationship between LIs on the Rhyme Decision, Dichotic Listening, and Chimeric faces by modelling covariances.
+We had originally intended to examine the relationship We had originally intended to examine the relationship between LIs on the Rhyme Decision, Dichotic Listening, and Chimeric faces by modelling covariances.
 This would have allowed us to compare (1) a model where the two language tasks are correlated with each other but not with the chimeric faces, (2) a model where all LIs are correlated, and (3) a base model where the LIs are unrelated.
 This would have been done separately for right- and left-handers.
-However, due to the Rhyme Decision task having poor reliability, we decided that this analysis would not be appropriate to base our conlcusions on.
-Instead, we report them as a supplmental material.
-The models were fit using a subset of SEM that dos not include any latent variables or directional paths.
-It only looks to constrain particular covariance patterns and report 'best' model according AIC weights.
+However, due to the Rhyme Decision task having poor reliability, we decided that this analysis would not be appropriate to base our conclusions on.
+Instead, we report them as a supplemental material.
+The models were fit using a subset of SEM that does not include any latent variables or directional paths.
+It only looks to constrain particular covariance patterns and report ‘best’ model according AIC weights.
 Model fit was evaluated using AIC model selection using Akaike weights (Wagenmakers & Farrell, 2004).
@@
 
-```{r Roger_mod_left,echo=FALSE}
+```{r Roger_mod_left,echo=FALSE, comment=FALSE, warning=FALSE}
 library(lavaan)
@@
 
-As shown in Figure S1, the akaike weights favoured a different model for left- and right-handers.
+As shown in Figure S1, the Akaike weights favoured a different model for left- and right-handers.
 For left-handers model 2, where all LIs were correlated, was preferred.
-For right-handers model 1, were languge tasks were correlated with eachother but not the chimeric faces was preffered.
+For right-handers model 1, were language tasks were correlated with each other but not the chimeric faces was preferred.
 The pattern of results for left handers was surprising given that in the main article we reported no correlation between the dichotic listening and chimeric faces tasks.
-Furthermore, for left-handers, the fact that the prefered model is one where language tasks are related seems to contrast previous work reporting that language tasks fractionate in left-handers (Woodhead et al., 2019).
+Furthermore, for left-handers, the fact that the preferred model is one where language tasks are related seems to contrast previous work reporting that language tasks fractionate in left-handers (Woodhead et al., 2019).
 This is likely to reflect that the poor reliability of the rhyme decision task.
@@
 
+<!---
+
+DOROTHY, I'VE LEFT THESE NOTES IN INCASE YOU FIND THEM USEFUL
+
 # Pre-reg info
 
-To do: 
-- reduce numbers of items in the task.
-
 ## Exploratory analysis
@@
 
-While the role of the LexTALE in the current research is largely for the purpose of data exclusion, we plan to explore correlations between LexTALE score and the laterality index on the Rhyme Decision task.
-This will enable us to examine whether vocabulary knowledge is related to rhyme judgement when information is displayed in each hemifield.
-We will similarly explore the influence of sight dominance, handedness/footedness, and the Finger Tapping Task on the laterality index for rhyme decision.
-
-Finally, we will investigate whether the method for measuring handedness has an impact on the associations between handedness and laterality on the Rhyme Decision and Chimeric Faces tests.
-In our main planned analyses, handedness will be coded using a binary categorisation of the Edinburgh Handedness Inventory scores, where > 0 indicates right handedness and < 0 indicates left handedness.
-There are many alternative ways of coding handedness: e.g.
-using three categories (right handed, ambidextrous and left handed); as a continuous measure; using self-report handedness; or using performance on the Finger Tapping test as an objective, continuous measure.
-We will explore which of these coding systems gives the strongest associations with laterality indices on the Rhyme Decision and Chimeric Faces.
-To do this we will select half the data at random to explore associations, and then aim to replicate on the second half.
-
 ## Results
-
-<!---I'm not sure how to weave this in as it stands, so leaving here for now--->
-
-*Still to be done: consider reliabilities with fewer trials. How short a test is feasible without loss of reliability?*
 
@@
 ```
+--->

### 2020-07-17 online_laterality_reliability.Rmd
reliability writeup draft updated by DB. Being actively revised! Do not edit until DB says!

--- 2020-07-10 draft_notebook_writeup.Rmd
@@
 
-Zoe V. J., Woodhead
+Zoe V. J. Woodhead
 
@@
 <!--- not sure how corresponding authorship works in OSCCI lab --->
+<!--- I'm fine with you in this role--->
 *Correspondence regarding this article should be addressed to Adam J.
@@
 This has constrained the number of participants recruited in such studies.
-The current work, however, introduces an online laterality battery that can be used to test participants on a variety of tasks: a rhyme decision visual half-field task, a dichotic listening task, a chimeric faces task, and a finger tapping task to assess manual preference.
-To validate the battery and assess its test-retest reliability, 392 participants completed the battery twice.
-Around a third of participants were left-handers.
-Not only do we report hemispheric advantages in the expected direction for each task, we additionally report good test-retest for the dichotic listening, chimeric faces, and finger tapping tasks.
-We also examine the association between handedness and atypical lateralisation and report that left-handers are more likely to depart from typical lateralisation.
-Furthermore, we provide data which indicates that lateralisation across tasks likely stems from multiple independent sources.
-The implications of these findings are discussed.
+Online testing can potentially allow for much larger sample sizes, but limits the amount of experimental control that is feasible.
+Here we considered whether online testing could give valid and reliable results on four tasks: a rhyme decision visual half-field task, a dichotic listening task, a chimeric faces task, and a finger tapping task to assess manual preference.
+We recruited 392 participants, oversampling left-handers, who completed the battery twice.
+Three of the tasks showed evidence of both validity and reliability, insofar as they showed hemispheric advantages in the expected direction and test-retest reliability of at least .65.
+The rhyme judgement task was less satisfactory.
+We also found that left-handers were more likely to depart from typical lateralisation.
+Contrary to prediction, lateralisation across tasks was not significantly correlated.
+We conclude that three of the tasks, dichotic listening, chimeric faces and finger-tapping, show considerable promise for online evaluation of cerebral lateralisation, but the results do not support the notion of a common underlying cause of lateralisation for different cognitive tasks.
 
@@
 Much of the ambiguity in the published literature may arise from unreliable measures of lateralisation and underpowered studies.
-Without well powered experiments that use reliable tasks, it becomes difficult to address many research questions regarding the relationship between laterality indices and their stability over time.
+Without well-powered experiments that use reliable tasks, it becomes difficult to address many research questions regarding the relationship between laterality indices and their stability over time.
 With respect to the relationship between indices, unless a task is reliable, the absence of a correlation cannot be interpreted as evidence for independence of function (Bishop, 2017).
-The best way to determine a task’s reliability is to test it twice.
-Here, the variation between participants should be larger than the variation within subjects.
-Taking this into consideration, the current work introduces an online behavioural laterality battery for which we established the test-retest reliability using a sample of 392 participants.
+Reliability can either be assessed within a test session (e.g.
+split-half reliability, from correlations between odd and even items) or from repeated assessments on different occasions, test-retest reliability.
+The latter is generally preferred as it includes all test items and will incorporate fluctuations in performance due to changes in state or test conditiions from one occasion to another.
+Assessment of test-retest reliability, however, requires large samples of people who undergo repeated testing, and this is seldom feasible for in-person testing.
+Online testing allows for recruitment of much larger samples, but raises questions about test validity when opportunities for experimental control are reduced.
+Here we introduce an online behavioural laterality battery for which we established the test-retest reliability using a sample of 392 participants.
 The behavioural tasks included were: a novel rhyme decision task, a dichotic listening task (Hugdahl et al., 2009), a chimeric faces task (Burt & Perret, 1997), and a novel finger tapping task.
-We subsequently use this battery to we examine two pre-registered questions: (1) do extreme left-handers show a higher rate of atypical laterality than right handers on each behavioural laterality task; (2) do the laterality indices produced on each measure correlate.
+Validity of the tasks was assessed in terms of whether lateralisation was observed in the predicted direction, based on prior literature with in-person testing on analogous tasks.
+We also examined two pre-registered questions: (1) do extreme left-handers show a higher rate of atypical laterality than right handers on each behavioural laterality task?
+(2) do the laterality indices from different tasks correlate?
 
@@
 Behavioural laterality tasks have long been used to establish hemispheric dominance for certain cognitive functions.
-Rather consistently, results from visual half-field and dichotic listening paradigms suggest that the left hemisphere plays the dominant role in language processing (Hugdhal, 2000; Ocklenburg & Güntürkün, 2018).
-Here we review how each paradigm has been implemented to study language lateralisation and each paradigms’ reliability.
-
-Evidence from the visual half-field paradigm (see Bourne, 2006, for a review), strongly points towards advantageous processing for letter strings presented in the right, relative to the left, visual field.
-As information in the right visual field projects contralaterally and arrives directly at the language dominant left hemisphere, it is thought to reflect an absence in callosal transfer that is required for the processing of linguistic information initially arriving at the right hemisphere.
-
-Despite wide reporting of a right visual field advantage at the population level when using visual half-field tasks, the paradigm is not without critique.
-In particular, the validity of visual half-field tasks has been questioned.
+Rather consistently, results from visual half-field and dichotic listening paradigms support the view that the left hemisphere plays the dominant role in language processing (Hugdahl, 2000; Ocklenburg & Güntürkün, 2018).
+Here we review how each paradigm has been implemented to study language lateralisation and what is known about reliability and validity.
+
+#### Visual half-field paradigm
+In this paradigm, visual stimuli are briefly presented to the left or right of a central fixation stimulus.
+The visual pathways from the left and right sides of space to visual cortex are crossed for non-foveal viewing, and so the right eye projects directly to the left hemisphere, and the left eye to the right hemisphere.
+The corpus callosum allows for transfer between hemispheres, but the rationale of the method is that there will be a processing advantage for language stimuli that are directly projected to the language-dominant hemisphere.
+This method has been used since the 1950s (Mishkin & Forgays, 1952).
+Bourne (2006) reviewed the literature, which consistently shows advantageous processing for letter strings presented in the right, relative to the left, visual field in most people.
+
+Despite wide reporting of a right visual field advantage at the population level,  the validity of visual half-field tasks has been questioned.
 Krach, Chen, and Hartje (2006) compared the laterality indices (LIs) produced during a visual half-field experiment with those produced using functional Transcranial Doppler Sonography (fTCD).
-The visual half-field experiment itself consisted of lexical decisions made towards abstract nouns presented in the left- and right visual-fields.
+The visual half-field task consisted of lexical decisions made towards abstract nouns presented in the left- and right visual-fields.
 During the experiment, the blood flow to left and right hemispheres of 58 participants was measured.
-In addition, blood flow was measured during a word generation task—the gold standard task for assessing hemispheric dominance for language.
+In addition, blood flow was measured during a word generation task — the gold standard task for assessing hemispheric dominance for language.
 Despite showing the expected right visual field advantage on the visual half-field task, the behavioural LIs produced on the visual half-field paradigm did not correlate with the fTCD LIs produced during word generation (r= .18) or the visual half-field task itself (r -.004).
-The lack of validity here likely reflects the poor reliability of the visual field task as split-half reliability was far from optimal (RT spilt-half: r= .14; accuracy split-half: r= .20).
-Thus, with such low split-half reliability it is unclear whether the task was able to accurately measure language lateralisation to start with.
+The lack of validity here likely reflects the poor reliability of the visual field task as split-half reliability was very low (RT spilt-half: r= .14; accuracy split-half: r= .20).
+This illustrates the difficulty of drawing conclusions about associations between laterality indices when reliability of measurement is so low.
 
@@
 For example, Hunter and Brysbaert (2008) noted stronger and more stable visual-field differences when stimuli were presented bilaterally relative to those where only one stimulus was presented either in the LVF or in the RVF (Boles, 1987, 1990, 1994; see also Iacoboni & Zaidel, 1996).
-Brysbaert and Hunter argued that this is the result of stimuli competing during bilateral presentation, which is easier when the stimulus is presented in the language dominant hemisphere.
-Reliability and validity also depend on there being a sufficient number of trials and brief stimulus presentation.
+Brysbaert and Hunter argued that this is the result of stimuli competing during bilateral presentation, enhancing the advantage to the stimulus presented in the language dominant hemisphere.
+Reliability and validity also depend on there being a sufficient number of trials and brief stimulus presentation to prevent eye movements.
 When accounting for these methodological aspects, the validity of visual half-field tasks seems promising.
-For example, Hunter and Brysbaert (2008) reported that LIs from visual half-field paradigms correlated positively with those from functional Magnetic Resonance Imaging (fMRI): word naming (r= .63), picture naming (r= .77).
-Here 192 trials were used during word naming and 160 trials during picture naming while stimuli were bilaterally presented and masked at offset.
-Furthermore, a promising visual half-field paradigm developed by Hausmann and colleagues (Hausmann et al., 2019; Willemin et al., 2016) which adhered to several of the methodological aspects outline by Brysbaert and Hunter has reported a right visual-field advantage at the population level with little variation occurring between languages.
-The authors, however, have not yet reported reliability measures for this task.
-
-While the visual half-field technique measures lateralisation for visual stimuli, dichotic listening tasks examine hemispheric dominance for the processing of auditory stimuli.
-Like the visual half-field technique, the dichotic listening task makes use of the crossing of fibres at the corpus callosum.
-In this task, different streams of auditory linguistic stimuli (i.e.
-words, syllables) are presented to each ear and participants are either asked simply to report as many words as they can or to attend to one ear selectively.
-In either case, a right ear advantage has been demonstrated; participants report more verbal stimuli presented to the right ear (Hugdahl, 2011; Kimura, 1961).
-Again, this advantage has been attributed to stimuli being projected to the language dominant hemisphere without callosal transfer.
+For example, Hunter and Brysbaert (2008) used two visual half-field tasks, word naming (192 trials) and picture naming (160) in which stimuli were bilaterally presented and masked at offset.
+The two tasks correlated .8 (95% CI .598-.956) in a sample of 26 left-handers.
+Encouragingly, the LIs from these visual half-field tasks also correlated positively with laterality from functional Magnetic Resonance Imaging (fMRI) on a word generation task in 10 left-handers.
+However, the small sample size means that the confidence intervals around these estimates are very wide: word naming, r= .63 (95% CI = .001-.902), picture naming, r= .77 (95% CI = .272-.943).
+<!---I removed Hausmann et al in interests of length, as they didn't report reliability--->
+
+#### Dichotic listening
+Simultaneous presentation of different auditory stimuli to left and right ears for immediate recall was a method originally used in studies of attention, but it became evident that there was usually an advantage for verbal stimuli presented to the right ear (Kimura, 1967).
+Both ears project to both hemispheres, but the weaker ipsilateral pathways appear to be suppressed when the two sides are in competition, and so the right ear, with a contralateral projection to the left hemisphere, will have an advantage in tasks that involve language processing.
 
@@
 First, based on a review of the published literature, Westerhausen and Samuelsen outlined eight design features that should be used in an optimal dichotic listening paradigm and explain their justification for each.
-The criteria are as follows: (1) stop consonant-vowel (CV) syllables as stimulus materials, (2) pair only CV stimuli from the same voicing category, (3) Alternating trials of voiced and unvoiced stimulus pairs, (4) free-recall instruction, (5) single-stimulus pair per trial, (6) paradigm length of 120 dichotic trials, (7) all stimulus pairs are presented in both orientations in identical frequency, and (8) includes binaural (homonymic) trials.
+The criteria are as follows: (1) stop consonant-vowel (CV) syllables as stimulus materials, (2) pair only CV stimuli from the same voicing category, (3) Alternating trials of voiced and unvoiced stimulus pairs, (4) free-recall instruction, (5) single-stimulus pair per trial, (6) paradigm length of 120 dichotic trials, (7) all stimulus pairs are presented to the two sides in identical frequency, and (8) includes binaural (homonymic) trials.
 Putting these criteria into practice, Westerhausen and Samuelsen administered a modified dichotic listening paradigm using both a verbal and manual response format to 50 right-handers.
@@
 
-So far, the studies reviewed have been interested in language lateralisation.
-That is not to say that behavioural tasks are limited to linguistic materials.
-For instance, the visual half-field paradigm has been used to measure lateralisation for the processing of facial similarity (e.g.
-Gilbert & Bakan, 1973), face emotionality (e.g.
-Levy, Heller, Banich, & Burton, 1983), and local information (Sergent, 1982).
-In a recent review of the reproducibility of visual-field asymmetries, Brederoo, Nieuwenstein, Cornelissen, and Lorist (2019) reported that they were able to replicate five out nine perceptual asymmetries using the visual half-field paradigm.
-
-Of importance to the current work is the use of the visual half-field paradigm to measure the lateralisation of face emotionality, in particular the chimeric faces task.
-In this task participants are presented with vertically split chimeric faces, where each half displays a different emotion.
-Depending on how the task is implemented, participants will either view a single chimeric face and report the emotion being expressed (e.g.
-Drebing, Federman, Edington, & Terzian, 1997; Kucharska-Pieture, & David, 2003) or view two faces, one above the other, and decide which of the two faces looks more emotive (e.g., Bourne, 2005; Levy, Heller, Banich, & Burton, 1983).
-In the two face version images will be identical, mirror images and typically include a neutral and emotive hemiface (i.e.
-Happiness-Neutral vs Neutral-Happiness).
-Regardless of implementation, participants’ responses point towards a bias for the processing of emotion in the left visual field.
-In the single face version, participants will more often report the emotion shown in the left hemiface.
-In the two-face version, participants show a bias in selecting the face with an emotion depicted in the left hemiface.
-
-In terms of its reliability, the chimeric faces tasks seem promising.
-In Levy et al.’s (1983) early work, they reported good split-half reliability for the two faces version of the chimeric faces task, where the left and right handers both showed a split-half reliability of .93.
-The validity of the chimeric faces task has been demonstrated using behavioural and fMRI data.
+#### Chimeric faces
+
+In studies of face processing, it is possible to use a modification of the visual half-field paradigm in which single stimuli are viewed, composed of vertically split chimeric faces, where each half displays a different emotion or identity.
+Our focus here is on emotion detection in chimeric faces, where a bias is found towards reporting the emotion shown in the left hemiface.
+
+The reliability of the chimeric faces tasks seem promising.
+Early work by Levy et al (1983) using two vertically presented chimeric faces, found split-half reliability of .93.
+The validity of the chimeric faces task has been demonstrated using behavioural and fMRI data, but in a task of recognising facial identity rather than emotion.
 In Yovel, Tambini, and Brandman’s (2008) study, participants performed a behavioural experiment where they had to recognize centrally presented chimeric faces, which presented different identities in the right- and left-visual-field.
-Interestingly, the behavioural measure significantly correlated with the LI calculated based on activation in the fusiform face area (r= .49).
-However, it is of importance to note the sample size used here was small, but at face value the results seem promising.
+It is encouraging to note that the behavioural measure significantly correlated with the LI calculated based on activation in the fusiform face area (r= .49).
+However, the sample size used here was small: with 17 participants, the 95% confidence interval around the observed correlation of .49 ranged from .012 to .786.
 
@@
 
-The studies reviewed so far have all focused on lateralisation at the population level.
-That is not to say that lateralisation does not vary between individuals.
+The studies reviewed so far have indicated consistent population biases in lateralisation on behavioural tasks, but also individual differences.
 One factor that has been associated with the degree of lateralisation is handedness.
-Neuroimaging data strongly indicates that these differences in lateralisation are the result of a group of left-handers who are right language dominant rather than indicating that this relationship operates on a continuum (Mazoyer et al., 2014).
-This increased departure from typical lateralisation in left-handers is not only present in neuroimaging studies, it is also reflected in behavioural tasks.
-Thus it will be important for our battery to replicate the findings regarding handedness and lateralisation summarised below.
+Neuroimaging data indicates that for language tasks these differences in lateralisation reflect the contribution of a subset of strong left-handers who are right language dominant, rather than indicating that this relationship operates on a continuum (Mazoyer et al., 2014).
+This association can be used to validate behavioural measures of lateralisation: we would expect to see departure from typical lateralisation in strong left-handers.
 
@@
 However, a relatively well powered study, involving 100 participants conducted by Willemin et al.
-(2016) reported no effect of handedness (when coded as left and right) on the visual field advantaged during lexical decision task.
-At first, this finding may seem odd.
-However, it may reflect underlying issues with the task rather than represent strict evidence for a null effect of handedness on visual half-field task.
-Here, Willemin et al.
-asked participants to respond via manual response.
-This contrasts several studies that have reported increased variability when responding aloud.
-Thus, a task that lacks some active speech production component may not yield such strong visual field advantages that enable provide enough variation to detect subtle differences in handedness (Hunter & Brysbaert, 2008).
-This, of course, is something that we kept in mind when developing the visual half-field paradigm used in the current study.
-
-When using dichotic listening paradigms to study language lateralisation, the evidence is clear that non-right-handers differ in terms of their laterality.
-While a clear right ear (i.e.
-left hemisphere) advantage is noted at the population level (see Hugdahl et al., 2009, and Westerhausen, 2019, for a review), there is evidence to suggest that this is reduced in left- relative to right-handers (Bethmann, Tempelmann, De Bleser, Scheich, & Brechmann, 2007; Bryden, 1970, 1988; Hugdahl, 2003); although handedness alone does not predict this difference (Van der Haegen, Westerhausen, Hugdahl, & Brysbaert, 2013).
-Across two studies Karlsson, Johnstone, and Carey (2019) provided a clear, yet comprehensive explanation of how performance on the dichotic listening task differed as a function of hand preference.
-Here, Karlsson et al.
-employed a consonant-vowel version of the dichotic listening task where participants reported which one of six syllables that they heard the clearest.
-For study 1, a clear right ear advantage was reported at the population level.
-Furthermore, right-handers were on average more lateralised than non-right-handers—87% of right-handers and 78% of non-right-handers showed a right ear advantage.
-For study 2, a similar pattern of results emerged.
-Again, right-handers were more lateralised with 85% of right-handers and 79% of non-right-handers showing an advantage.
-Together these results indicate that non-right-handers show a larger departure from the ‘typical’ lateralisation during dichotic listening tasks.
-
-Recall that, for the chimeric faces, there is typically a left visual field advantage at the population level— indicating right hemisphere dominance.
-As with other laterality tasks, there appears to be a greater departure from this typical pattern of laterality for left-handers (e.g.
-Gilbert & Bakan, 1973; Heller & Levy, 1983; Karlsson et al., 2019; Levy et al., 1983; Rozkowski & Snelbecker, 1982).
-Again, this departure is well described by Karlsson et al.
+(2016) reported no hint of any effect of handedness, when coded as left and right, or when divided into left (N = 19), mixed (N = 24) and right-handed (N = 57) groups, on the visual field advantage during a lexical decision task.
+This result illustrates the recurring challenge of interpreting null results in this field: do they reflect a true lack of relationship, poor test reliability, inadequate power to detect small effects, or specific methodological factors?
+We may note that Willemin et al.
+asked participants to respond via manual response, whereas other studies have involved an active speech production component, which could influence task lateralisation.
+
+On dichotic listening paradigms, the clear right ear (i.e.
+left hemisphere) advantage that is noted at the population level (see Hugdahl et al., 2009, and Westerhausen, 2019, for a review) is reduced in left-handers (Bethmann, Tempelmann, De Bleser, Scheich, & Brechmann, 2007; Bryden, 1970, 1988; Hugdahl, 2003).
+More recently, Karlsson, Johnstone, and Carey (2019) employed a consonant-vowel version of the dichotic listening task where participants reported which one of six syllables that they heard the clearest.
+In study 1, 87% of right-handers and 78% of non-right-handers showed a right ear advantage.
+For study 2, 85% of right-handers and 79% of non-right-handers showed a right ear advantage.
+
+For chimeric faces, which typically show a left visual field advantage there again appears to be a greater departure from the typical pattern of laterality for left-handers (e.g.
+Gilbert & Bakan, 1973; Heller & Levy, 1983; Levy et al., 1983; Rozkowski & Snelbecker, 1982).
+This was also studied in the study by Karlsson et al.
 (2019).
-This departure was examined for two version of the chimeric faces task.
-The first involve participants viewing a single face with a different emotion shown in either hemiface.
-Participants then reported the emotion that they felt was most strongly displayed.
-The second was a two faces version where participants viewed two a face with an emotion on one side and the mirror version.
-Here participants had to decide which of the two faces showed a target emotion.
-For the first task, right-handers showed a bias for reporting information presented in the left visual field (i.e.
-right hemisphere); however, left-handers did not show a population bias.
-Specifically, 75% of right-handers and 60% of non-right-handers showed a leftwards bias.
-For the second chimeric faces task, a similar pattern emerged.
-73% of right-handers and 57% of non-right-handers here showed a leftwards bias.
-Based on the above finding, a good laterality task must not only be reliable, but should be able to replicate a departure from typical lateralisation in those who are left-handed.
+When viewing a single face with a different emotion shown in either hemiface, 75% of right-handers and 60% of non-right-handers showed a leftwards bias.
+For the second chimeric faces task that involved two chimeric faces arranged vertically, a similar pattern emerged.
+73% of right-handers and 57% of non-right-handers showing a leftwards bias.
 
@@
 Furthermore, there are a large number of studies which report no correlation between laterality indices (Bryden, Hecaen, & Deagostini, 1983; Rosch, Bishop, & Badcock, 2012; Whitehouse & Bishop, 2009; van der Haegen & Brysbaert, 2018).
-This has led to the conclusion that there are multiple independent biases across tasks which lead to hemispheric lateralisation of different cognitive domains.
+However, it has not been possible to clearly unpick whether the null results indicate that there are multiple independent lateral biases across tasks, or whether there is a more prosaic explanation in terms of poor reliability or validity of tasks.
 
@@
 
-Our principle goal for the current study was to maximise the test-retest reliability of an online laterality battery.
-Specifically, we examined whether it is possible to achieve a test-retest reliability of .65 or above using online behaviour research methods, and, if so, whether this reliability could be achieved more efficiently by using fewer trials in each test.
-The tasks included in the battery were a novel rhyme decision visual half-field tasks, a consonant-vowel dichotic listening tasks, a chimeric faces visual half-field task, and a novel finger tapping tasks.
-As we were primarily interested in establishing test-retest reliability, the behavioural tasks included in the battery were completed twice on two different occasions.
-In addition to establishing the test-retest reliability for the battery, we tested several predictions about the relationship between laterality on the four behavioural tasks.
-We describe these predictions below.
-
-Our first set of predictions related to participants’ performance at a population level on each task.
-Based on previous work, we predicted that: (1) the laterality indices for Rhyme Decision will show a right visual field advantage (indicating left hemisphere dominance for language processing); (2) the laterality indices for the Dichotic Listening task will show a right ear advantage (indicating left hemisphere dominance for language); (3) the laterality indices for Chimeric Faces will show a left visual field advantage (indicating right hemisphere dominance for face or emotion processing); and (4) the Finger Tapping task will show a right-hand advantage.
-
-Our second prediction that we examined relates to handedness and atypical lateralisation.
+The principal goal of the current study was to evaluate the reliability and validity of an online behavioural laterality battery.
+Specifically, we examined whether it is possible to observe the usual lateral biases at the population level in tests of language and emotion processing, and whether it is possible to achieve a test-retest reliability of .65 or above using online behaviour research methods.
+In addition, for tasks where satisfactory reliability was achieved, we considered how many trials were needed to achieve this.
+The tasks included in the battery were a novel rhyme decision visual half-field task, a consonant-vowel dichotic listening tasks, a chimeric faces visual half-field task, and a novel finger tapping task.
+These behavioural tasks were completed twice on two different occasions.
+
+Our first set of predictions concerned validity of the tasks as indicators of cerebral lateralisation.
+Based on previous work, we predicted that at the population level: (1) the laterality indices for Rhyme Decision would show a right visual field advantage (indicating left hemisphere dominance for language processing); (2) the laterality indices for the Dichotic Listening task would show a right ear advantage (indicating left hemisphere dominance for language); and (3) the laterality indices for Chimeric Faces would show a left visual field advantage (indicating right hemisphere dominance for face or emotion processing).
+The Finger Tapping task was designed to provide a quantitive measure of degree of handedness, and was expected to show a right-hand advantage in the overall sample.
+
+Our second prediction related to handedness and atypical lateralisation.
 We predicted that strong left-handers (i.e.
-those with laterality quotient of -90 or less on the Edinburgh Handedness Inventory) will be more likely to show atypical laterality (i.e.
+those with laterality quotient of -90 or less on the Edinburgh Handedness Inventory) would be more likely to show atypical laterality (i.e.
 laterality in the opposite direction to our population level predictions) than right-handers.
-This prediction follows from observations by Mazoyer et al (2014, p.
+This prediction followed from observations by Mazoyer et al (2014, p.
 1) who studied a large group of left- and right-handers using fMRI to measure language laterality.
@@
 However, their study was restricted to left-handers, who may be atypical in this regard, and they noted that lack of association was hard to interpret because it may be due to unreliability of measurement.
-Originally, we predicted that for right-handers, a model (a) where the two language tasks are correlated with each other but not with the chimeric faces will provide better fit to the data than a model (b) in which laterality indices are correlated between the language and chimeric face tasks, or a base model (c) where each task is independently lateralised.
+In the preregistration, we predicted that for right-handers, a model (a) where the two language tasks are correlated with each other but not with the chimeric faces will provide better fit to the data than a model (b) in which laterality indices are correlated between the language and chimeric face tasks, or a base model (c) where each task is independently lateralised.
 For left-handers, we anticipated that that model (c) may be a better fit to the data, indicating that laterality is dissociated across the different tasks.
-However, we were not able to test this planned analysis with the current data set as one task had low reliability.
-Thus, we instead examined the relationship between laterality tasks using correlation analysis.
+However, we were not able to test this planned analysis with the current data set as one language task had low reliability.
+Thus, we confined our analysis to presentation of correlations between laterality indices from the four measures.
 
@@
 
-This project was pre-registered on the Open Science Framework (OSF) prior to the commencement of data collection.
+This project was pre-registered on the Open Science Framework (OSF) prior to the start of data collection.
 The registration form along with the analysis scripts and anonymised data can be found on the OSF at https://osf.io/qsxcv/.
@@
 As our ultimate goal is to use these measures in individual differences research, we aimed to optimise the task to achieve observed test-retest reliability of at least .65.
-To achieve sufficient precision around the estimate of reliability, we estimated that we would require 300 participants.
+Our initial power analysis showed that to achieve sufficient precision around the estimate of reliability, we would require 300 participants.
 Out of the 300 participants, we originally specified that 80 would be left-handers.
-We had examined the power that a sample of 80 participants would provide to make the originally planned model comparisons outlined under hypothesis 2.
-By comparing model fit using a likelihood test, 80 participants would provide 100% power for model (a) and 97% power for model (b).
-Thus, we would have sufficient power to assess the fit of each model using Akaike weights, which are more conservative than likelihood tests.
-
-Despite setting out to recruit 300 participants, our final sample exceeded this target (we say more about this below).
+This would give us power to apply our planned model-fitting analysis to left-handers as a separate group.
+
+In practice, our final sample exceeded 300, because we discarded participants from individual tasks for failing to meet prespecified criteria (see below).
 In total, 441 native English speakers were initially recruited through Prolific Academic.
@@
 
-#### Test of ocular dominance
+#### Tests of ocular dominance
 
@@
 Each test classified participants as being either left or right eye dominant.
-
-#### Edinburgh handedness inventory
-
-The Edinburgh Handedness Inventory (Oldfield, 1971) was administered to quantify handedness on a continuum.
+In practice, the two tests agreed perfectly, so we report here just the Miles test.
+
+#### Edinburgh Handedness Inventory
+
+The Edinburgh Handedness Inventory (EHI: Oldfield, 1971) was administered to quantify handedness on a continuum.
 For 10 everyday activities, participants indicated their preferred hand use on a 5-step scale (right hand strongly preferred, right hand preferred, no preference, left hand preferred, left hand strongly preferred).
-The Edinburgh handedness inventory was scored in the standard way, which involves combining information about direction and exclusiveness of hand preference.
+The EHI was scored in the standard way, which involves combining information about direction and exclusiveness of hand preference.
 For simple contrast between left- and right-handers, we treated indices greater than 0 as right handed, and indices less than 0 as left handed.
@@
 
-The rhyme decision task is a novel visual half-field technique which requires participants to decide whether a foveally presented letter string rhymed with an image shown in either the left or right hemifield.
+The rhyme decision task was developed for this project as a novel visual half-field technique which requires participants to decide whether a foveally presented letter string rhymed with the name of an image shown in either the left or right hemifield.
+With online presentation, a particular concern is potential for eye movements away from fixation.
+We relied on brief bilateral presentation, which has guaranteed adequate control for eye movements in previous VHF experiments with verbal naming (Beaumont, 1982).
+Studies that directly monitored eye movements in this setting reported failures of fixation on only 0.5% of trials (Geffen, Bradshaw & Nettleton, 1972, but see also Bourne, 2006).
+The design we used ensured that participants had to start the trial fixating centrally in order to read the target word.
+Pictures of familiar objects were presented for rhyme matching to the written word, with the goal of ensuring that the participant would have to covertly name the picture, thereby engaging left-hemisphere speech processing.
 
@@
 
-Each of the 52 unique stimuli were shown five times.
-The 260 stimuli were then displayed across four blocks (65 items per block).
+The trials were preceded by a familiarisation phase in which each of the stimulus pictures was presented together with its written name.
+Then 260 stimuli (52 unique stimuli each shown 5 times) were displayed across four blocks (65 items per block).
 Each of the four blocks also contained five items for which there was no rhyme.
@@
 These foil items were included in attempt to prevent participants from preferentially responding to items in either the LVF or RVF.
-These items were then pseudo-randomised such that image pairs could not appear one after the other.
+These items were then pseudo-randomised such that identical image pairs could not appear one after the other.
 
@@
 The bilaterally presented images were then presented for 150 ms.
-Brief bilateral presentation has guaranteed adequate control for eye movements in previous VHF experimental with verbal naming (Beaumont, 1982).
-Studies that directly monitored eye movements reported failures of fixation on only 0.5% of trials (Geffen, Bradshaw & Nettleton, 1972, but see also Bourne, 2006).
 At stimulus offset, participants indicated whether the image in the LVF or RVF rhymed with the foveally presented word.
@@
 
-The consonant-vowel dichotic stimuli were kindly provided by Dr Emma Karlsson, and were used with the permission of Professor Kenneth Hugdahl.
-The stimuli have been used in prior research and reported elsewhere (Hugdahl et al., 2009; Karlsson et al., 2019).
+The consonant-vowel dichotic stimuli were taken from a dichotic listening task that was developed to work as a smartphone app, and shown to give high levels of test-retest reliability (Bless et al., 2013).
+These stimuli were the same as those used by Karlsson et al (2019) and were used with the kind permission of Professor Kenneth Hugdahl.
 
@@
 ##### Procedure
+
+As the dichotic listening task demands the use of stereo-headphones, participants were screened at the start of each session.
+Screening consisted of two tasks.
+The first involved participants judging which of three pure tones is quietest, and is described by Woods, Siegel, Traer, and McDermott (2017) in full detail.
+Sound discrimination is designed to be easy to achieve when using headphones but difficult over loudspeakers due to phase-cancellation.
+Participants were removed from the study if they failed to correctly discriminate the quietest sound on at least five out of six trials, indicating that they were unlikely to be using headphones.
+The second check involve participants hearing a sound presented in only one of the two channels.
+Participants were removed if they failed to correctly identify which channel the sound was played on five out of six trials, indicating that they were not using stereo-headphones.
 
@@
 Four emotional expressions were used: anger, disgust, happiness, and sadness.
-The faces were split vertically down the middle of the face and paired so that one emotive hemifaces was attached to another and blended at the midline (see Figure 2 for an example).
+The faces were split vertically down the middle and paired so that one emotive hemiface was attached to another and blended at the midline (see Figure 2 for an example).
 These were paired in all possible combinations creating 16 individual stimuli.
@@
 
-The Finger Tapping task is a novel behavioural task to classify handedness.
+The Finger Tapping task is a novel behavioural task that we designed as a quantitative index of handedness.
 Participants were required to either press the ‘W’, ‘R’, ‘V’, and ‘X’ keys with their left hand or the ‘Y’, ‘I’, ‘M’, ‘B’ keys with their right hands in sequence as many times as possible within a 30 second window.
@@
 
-As the dichotic listening task demands the use of stereo-headphones, participants were screened at the start of each session.
-Screening consisted of two tasks.
-The first involved participants judging which of three pure tones is quietest, and is described by Woods, Siegel, Traer, and McDermott (2017) in full detail.
-Sound discrimination is designed to be easy to achieve when using headphones but difficult over loudspeakers due to phase-cancellation.
-Participants were removed from the study if they failed to correctly discriminate the quietest sound on at least five out of six trials, indicating that they were unlikely using headphones.
-The second involve participants hearing a sound presented in only one of the two channels.
-Participants were removed if they failed to correctly identify which channel the sound was played on five out of six trials, indicating that they were not using stereo-headphones.
-
 When completing the behavioural tasks participants were instructed to sit approximately 50 cm from the screen.
 Gorilla’s scaling tool was utilised such that stimuli size and eccentricity was maintained across tasks and participants.
-Stimuli were from each task were repeated across four blocks.
+Stimuli from each task were repeated across four blocks.
 Within each block participants completed 70 trials of the rhyme detection task (R); 36 trials of the dichotic listening task (D); 32 trials of the chimeric face task (C); and 2 trials of the finger tapping task (F) for both their dominant and non-dominant hand.
@@
 library(kableExtra)
-
+library(ggpubr)
 # read in data from csv
@@
 ```
-
-Out of the 441 participants recruited via Prolific, `r nrow(LIs_c)` (`r round(nrow(LIs_c)/441*100, 1)`%) complete both sessions.
+<!---Figure was not so easy to read : I think table preferable here--->
+```{r simplifiedlatdescriptives}
+#Miles and Porta are identical so only report one.
+short_dat$cat3<-'R'
+short_dat$cat3[short_dat$handedness=='Left']<-'L'
+short_dat$temp<-'R'
+short_dat$temp[short_dat$footedness=='Left']<-'L'
+short_dat$temp[short_dat$footedness=="Don't know"]<-'X'
+short_dat$cat3<-paste0(short_dat$cat3,short_dat$temp)
+short_dat$temp<-'R'
+short_dat$temp[short_dat$miles=='Left']<-'L'
+short_dat$temp[short_dat$miles=="Don't know"]<-'X'
+short_dat$cat3<-paste0(short_dat$cat3,short_dat$temp)
+t<-table(short_dat$cat3)
+lattab<-data.frame(matrix(NA,nrow=6,ncol=5))
+colnames(lattab)<-c('Foot.Eye','R.handers.N','R.handers.%','L.handers.N','L.handers.%')
+lattab$Foot.Eye<-c('RR','RL','LR','LL','Other','Total')
+lattab$R.handers.N[6]<-sum(t[9:16])
+lattab$R.handers.N[1:4]<-as.vector(t[c(13,12,10,9)])
+lattab$R.handers.N[5]<-lattab$R.handers.N[6]-sum(lattab$R.handers.N[1:4])
+lattab$L.handers.N[6]<-sum(t[1:8])
+lattab$L.handers.N[1:4]<-as.vector(t[c(5,4,2,1)])
+lattab$L.handers.N[5]<-lattab$L.handers.N[6]-sum(lattab$L.handers.N[1:4])
+sumR<-lattab$R.handers.N[6]
+sumL<-lattab$L.handers.N[6]
+lattab$`R.handers.%`<-round(100*lattab$R.handers.N/sumR,1)
+lattab$`L.handers.%`<-round(100*lattab$L.handers.N/sumL,1)
+lattab$`R.handers.%`[6]<-NA
+lattab$`L.handers.%`[6]<-NA
+
+```
+Out of 441 participants recruited via Prolific, `r nrow(LIs_c)` (`r round(nrow(LIs_c)/441*100, 1)`%) completed both sessions.
 We then removed `r nrow(LIs_no)` participants (`r round(nrow(LIs_no)/nrow(LIs_c)*100, 1)`%) from our final sample as their lexTALE score was below 80, indicating that a participant’s level of vocabulary knowledge was below the cut off for lower advanced proficiency.
-This left a final sample of `r nrow(LIs)` participants.
-
+This left a sample of `r nrow(LIs)` participants.
+
+##### Data cleaning
 For each task we adopted the following data cleaning procedures.
@@
 
-Next, for the chimeric faces task, we removed participants data if they scored lower than 75% on trials in which an identical emotion was shown in each hemiface.
-These data trimming procedures left data for `r nrow(LIs)-nrow(LIs[LIs$exclude== 1 | LIs$exclude== 12,])` participants on the chimeric faces.
-For the dichotic listening task, we removed those who scored less than 75% correct on trials where the same sound was played to each ear.
-These data trimming procedures left data for `r nrow(LIs)-nrow(LIs[LIs$exclude== 2 | LIs$exclude== 12,])` participants on the dichotic listening task.
-
-With regards to the rhyme decision task, we removed participants who scored less than 50% accuracy in either visual field.
+For the chimeric faces task, we removed participants if they scored lower than 75% on trials in which an identical emotion was shown in each hemiface, and for the dichotic listening task, we removed those who scored less than 75% correct on trials where the same sound was played to each ear.
+These data trimming procedures left data for `r nrow(LIs)-nrow(LIs[LIs$exclude== 1 | LIs$exclude== 12,])` participants on the chimeric faces `r nrow(LIs)-nrow(LIs[LIs$exclude== 2 | LIs$exclude== 12,])` participants on the dichotic listening task.
+
+For the rhyme decision task, we removed participants who scored less than 50% accuracy in either visual field.
 Furthermore, we removed participants who scored less than 75% accuracy on the catch trials (trials in which there was no rhyme).
-These trimming procedures left data for `r nrow(LIs[LIs$exRDT== 0,])` participants on the rhyme decision task.
-Seeing as we were removing a relatively high portion of participants, want wanted to check whether this task was yielding reliably results.
-With `r nrow(LIs[LIs$exRDT== 0,])` participants, the test-retest reliability was *r*= `r round(cor(LIs[LIs$exRDT== 0,]$Day1.RDT_RT_LI, LIs[LIs$exRDT== 0,]$Day2.RDT_RT_LI, use = "complete.obs"),2)`, which is smaller than the *r*= 0.65 that we pre-registered as our criteria for a reliable task.
-Thus, rather than collect further participants so that we would have our pre-registered 300 participants for the rhyme decision task, we chose to end data collection here and analyse the `r nrow(LIs)` participants that we had collected under the assumption that the rhyme decision is not reliable.
+These trimming procedures left data for `r nrow(LIs[LIs$exRDT== 0,])` participants on this task.
+We considered recruiting further participants to make up the numbers, but this seemed uneconomic as we already had more than 300 cases for the other tasks, and we could tell by this point that the test-retest reliability of the rhyme detection task was unsatisfactory; With `r nrow(LIs[LIs$exRDT== 0,])` participants, the test-retest reliability was *r*= `r round(cor(LIs[LIs$exRDT== 0,]$Day1.RDT_RT_LI, LIs[LIs$exRDT== 0,]$Day2.RDT_RT_LI, use = "complete.obs"),2)`, which is smaller than the *r*= 0.65 that we pre-registered as our criterion for a reliable task.
+Accordingly, we stopped data collection at this point and analysed the `r nrow(LIs)` participants with data on other tasks.
+
 The final group of `r nrow(LIs)` participants (`r nrow(LIs[LIs$gender== "Female",])` female) had a mean age of `r round(mean(LIs$age, na.rm= TRUE), 1)` years (*SD*= `r round(sd(LIs$age, na.rm= TRUE), 2)` and included `r nrow(LIs[LIs$handedness== "Left",])` left handers (`r nrow(LIs[LIs$handedness== "Left" & LIs$gender== "Female",])` female).
 `r nrow(LIs[LIs$handedness== "Right" & LIs$bilingual== "Yes",])` right-handers were bilingual and `r nrow(LIs[LIs$handedness== "Left" & LIs$bilingual== "Yes",])` left-handers were bilingual.
-The division of footedness and ocular dominance in left- and right-handers are shown in Figure 3.
-Below we describe the analysis procedures used to examine our research questions and hypotheses.
-
-![Figure 3. Count of left- and right-handers who report left, right, or don’t know on self-reported footedness, the Miles test, and the Porta test.](handed_sight.png)
-
-##### Calculations of laterality indices, reliability, and population level analysis
-
-For the rhyme task, we pre-registered that we would calculate LIs based on reaction time data, as accuracy during piloting was at ceiling.
-We specified that we would calculate LIs in two ways: (1) LIs would be calculated in the traditional fashion (100 x (Right – Left)/(Right + Left)); (2) LIs would be calculated in a way that corresponded to a *t*-value.
-This was achieved by conducting a *t*-test for each participant where accurate response times are the dependent variable and visual field is the independent variable.
-This would estimate the sensitivity to stimuli presented in either visual field and equate to a laterality index.
-One benefit of this approach is that we could identify participants who show a statistically significant reaction time advantage for one side, i.e.
+
+##### Methods for calcuating laterality indices, reliability, and population level analysis
+
+For the rhyme task, we pre-registered that we would calculate LIs based on response time (RT) data, as accuracy during piloting was at ceiling.
+We specified that we would calculate LIs in two ways: (1) in the traditional fashion (100 x (Right – Left)/(Right + Left)); (2) to correspond to a *t*-value.
+This was achieved by conducting a *t*-test for each participant where accurate RTs are the dependent variable and visual field is the independent variable.
+This latter measure is highly correlated with the conventional laterality index but has the advantage that it can be used to identify participants who show a statistically significant RT advantage for one side, i.e.
 where the *t*-value on a 2-sided test has *p*< .05.
-Based on considerations from pilot data, we planned to transform RT data from the RDT task by log-transformation before computing *t*-values.
+Based on considerations from pilot data, we adopted log-transformation of RTs before computing *t*-values.
 
 For the dichotic listening and chimeric faces, we specified that LI calculated would be based on trials in which participants correctly identified a stimulus.
-The count of correct responses that corresponded to each side would be used to generate an accuracy LI in two ways: (1) LIs would be calculated in the traditional fashion (100 x (Right – Left)/(Right + Left)); (2) LIs would be computed as a *z*-statistic for each participant, using z = (pR-.5)/sqrt(pR x pL/n), where pR is the proportion of correct responses in the RVF, pL is proportion of L responses in the LVF, and n is total LVF and RVF responses.
-LIs for the finger tapping task would be calculated in the same manner, but using the number of sequences generated by the left and right hand as opposed to the proportion of correct responses in each visual field.
+The count of correct responses that corresponded to each side was used to generate an accuracy LI in two ways: (1) in the traditional fashion (100 x (Right – Left)/(Right + Left)); (2) as a *z*-statistic for each participant, using z = (pR-.5)/sqrt(pR x pL/n), where pR is the proportion of correct responses in the RVF, pL is proportion of L responses in the LVF, and n is total LVF and RVF responses.
+This was analogous to the individual t-scores generated for RTs in the rhyme decision task, in allowing us to identify individuals with a significant bias to one side.
+Finally, LIs for the finger tapping task were calculated in the same manner, but using the number of sequences generated by the left and right hand as opposed to the proportion of correct responses in each visual field.
 
@@
 For tasks where the test-retest reliability was above .75 with the existing test length, we reanalysed the data using just the first N trials, reducing N until reliability becomes unsatisfactory (below *r*= .65).
-This will allow us to identify the most efficient way to administer these tasks in future, while retaining good reliability.
+This was of interest in indicating the most efficient way to administer these tasks in future, while retaining good reliability.
 However, the full data (including all trials) was used for all subsequent analyses.
+
 To address hypothesis 1 and examine the laterality of each task at the population level, we conducted a series of one-sample *t*-tests for each task using session 1 data.
@@
 
-To address hypothesis 2, that extreme left-handers will be more likely to show atypical laterality than right-handers, we first identified participants as being atypical.
-Here participants were defined as left-handed if they scored less than -90 on the Edinburgh Handedness Inventory and right-handed if they scored more than 90 of the Edinburgh Handedness Inventory.
-Atypicality was first defined in participants where their *t*- or *z*-score was significantly laterality in the opposite direction from that expected.
+To address hypothesis 2, that extreme left-handers will be more likely to show atypical laterality than right-handers, we categorised participants as extreme left-handers if they scored less than -90 and right-handers if they scored more than 90 of the Edinburgh Handedness Inventory.
+Atypicality was first defined in participants whose *t*- or *z*-score was significantly lateralised in the opposite direction from usual.
 If there were less than 20 cases defined as atypical using this method, we adjusted the cut-off on that task until we acquired 20 atypical participants.
@@
 This would have been done separately for right- and left-handers.
-However, due to the Rhyme Decision task having poor reliability, we decided that this analysis would not be appropriate.
+However, given poor reliability of the Rhyme Decision task, we decided that this analysis would not be appropriate.
 Instead, we examine the correlation coefficient for LIs on the Dichotic Listening and Chimeric Faces tasks.
@@
 ## Results
+
+Lateral preference characteristics of the sample after exclusions as noted above is shown in Table 1. <!---use new lattab for table 1--->
+
+Table 1 <!---New table - lattab to be added in nice format--->
+Count of left- and right-handers who report left or right on self-reported footedness, and eyedness (Miles test). 'Other' includes those with don't know responses.
 
@@
 # read CF data
-for_figs <- read.csv("./data/for_figs.csv", na.strings = "NA")
+for_figs <- read.csv("for_figs.csv", na.strings = "NA")
+
 # get CF
@@
 finger.short.dat <- read.csv("./data/finger.csv", na.strings = "NA")
-
 # plots
@@
 
-### Calculations of laterality indices, reliability, and population level analysis
-
-Participants' average reaction times on the rhyme decision task are shown in Figure 4 across visual field and session.
-While we do not analyse accuracy, accuracy on the rhyme task was high  (*M*= `r round(mean(RD.short.dat$p.corr, na.rm= TRUE),1)`%, *SD*= `r round(sd(RD.short.dat$p.corr, na.rm= TRUE),2)`%).
-The count of correct answers on the dichotic listening and chimeric faces tasks are also shown in Figure 4, along with the number of sequences generated by the left and right hand.
-LIs were computed in several ways.
+<!---These figures aren't so good for showing laterality because they don't show how the L and R scores relate in individuals. Will try an alternative approach.--->
+
+```{r altplotraw,echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
+
+makerawplot <- function(file1,file2,mysession,myname,mylim){
+
+tempL<-filter(file1,Session==mysession)
+tempR<-filter(file2,Session==mysession)
+meanL<-mean(tempL$tempcol,na.rm=T)
+meanR<-mean(tempR$tempcol,na.rm=T)
+tempLR<-as.data.frame(cbind(tempL$tempcol,tempR$tempcol))
+colnames(tempLR)<-c('Left','Right')
+mytitle<-paste0(myname,"\n Session",mysession)
+myg <- ggplot(tempLR, aes(x=Left, y=Right)) + geom_point(size=.5) +
+  xlim(mylim)+ylim(mylim)+
+    labs(title=mytitle)+
+    geom_abline(intercept = 0, slope = 1, color="red", 
+                 linetype="dotted")+
+  geom_hline(yintercept=meanR,linetype='dotted')+geom_vline(xintercept=meanL,linetype='dotted')+
+    theme(plot.title = element_text(size = 12))
+return(myg)
+}
+```
+```{r dorawplotfunction,echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
+
+RD.short.dat$tempcol<-RD.short.dat$RT
+file1 <-filter(RD.short.dat,side=='left')
+file2 <-filter(RD.short.dat,side=='right')
+myname<-'Rhyme judge RT'
+mysession<-1
+mylim<-c(500,1500)
+myg1<-makerawplot(file1,file2,mysession,myname,mylim)
+myg2<-makerawplot(file1,file2,2,myname,mylim)
+
+DL.short.dat$tempcol<-DL.short.dat$accurate
+file1 <-filter(DL.short.dat,side=='left')
+file2 <-filter(DL.short.dat,side=='right')
+myname<-'Dichotic accuracy'
+mysession<-1
+mylim<-c(0,100)
+myg3<-makerawplot(file1,file2,mysession,myname,mylim)
+myg4<-makerawplot(file1,file2,2,myname,mylim)
+
+CF.short.dat$tempcol<-CF.short.dat$accurate
+#there are 96 stimuli with different faces, so 48 is pt of equality if all trials included
+file1 <-filter(CF.short.dat,side=='left')
+file2 <-filter(CF.short.dat,side=='right')
+myname<-'Chimeric faces accuracy'
+mylim<-c(0,96)
+mysession<-1
+myg5<-makerawplot(file1,file2,mysession,myname,mylim)
+myg6<-makerawplot(file1,file2,2,myname,mylim)
+
+finger.short.dat$tempcol<-finger.short.dat$sequences
+file1 <-filter(finger.short.dat,side=='left')
+file2 <-filter(finger.short.dat,side=='right')
+myname<-'Finger tap sequences'
+mysession<-1
+mylim<-c(0,175)
+myg7<-makerawplot(file1,file2,mysession,myname,mylim)
+myg8<-makerawplot(file1,file2,2,myname,mylim)
+
+ggarrange(myg1,myg2,myg3,myg4,myg5,myg6,myg7,myg8, ncol = 2, nrow = 4)
+ggsave(
+  "all.raw.png",
+  width = 5, height = 8,
+  dpi = 300
+)
+```
+
+### Laterality indices: means and test-retest reliability
+
+Accuracy on the rhyme task was high  (*M*= `r round(mean(RD.short.dat$p.corr, na.rm= TRUE),1)`%, *SD*= `r round(sd(RD.short.dat$p.corr, na.rm= TRUE),2)`%), and only correct response times were used for computing a laterality index.
+Average reaction times on this task are shown in Figure 3 as vertical and horizontal lines in a scatterplot showing right vs left visual field, separately for each session.
+A sloping dotted line shows the point where left and right RTs are equal; if there is faster performance for the right visual field, as predicted, then points should cluster below this line.
+Similar scatterplots are shown for selection of left and right responses for dichotic listening and chimeric faces.
+Note that for these two tasks, either side can give a correct response; the negative correlation between the two sides reflects the fact that the number of selections on the left will be inversely related to the number of correct responses on the right.
+For these tasks it is evident by inspection that most responses for dichotic listening cluster above the dotted line, indicating a right ear advantage, whereas most responses on chimeric faces cluster below the dotted line.
+Finally, we show the number of correct sequences with each hand on the finger-tapping task.
+Here, as expected, inspection of the data suggests more sequences with the right than the left hand on average.
+
+![Figure 3.
+Left vs right scores for sessions 1 and 2 for the behavioural laterality tasks.
+The diagonal dotted line shows point of equality: points above this are higher for R than L, and points below are higher for L than R.
+Mean values for R and L are shown as horizontal and vertical dotted lines.
+](all.raw.png)
+
+The count of correct answers on the dichotic listening and chimeric faces tasks are also shown in Figure 3, along with the number of sequences generated by the left and right hand.
 For all tasks, we computed LIs in the typical way: (100 x (Right – Left)/(Right + Left)).
@@
 A negative LI indicates left lateralisation at the brain level while a positive LI indicates right lateralisation at the brain level.
-For the rhyme decision task, we then computed an LI using a *t*-test for inidividual participants fit to log-transformed reaction time data.
-For the remaining tasks (which made use of count data) we computed a *z*-statistic for each participant using count data.
-
-![Figure 4.
-Mean scores across the two sessions of the behavioural laterality tasks.
-(A) Reaction times in each visual field on the rhyme decision tasks.
-(B) Count of correctly identified sounds played to the left or right ear.
-(C) Count of correctly identified emotions in each hemiface.
-(D) Count of sequences generated by the left and right hand.](all.vf.png)
-
-For each measure, LIs and their test-retest correlation coefficients are reported in Table 1.
-In addition, Table 1 includes the *p*-values for one-sample *t* and Shapiro-Wilk tests for each measure  As the distribution of each LI was non-normal, spearman's rank correlations were used to estimate the test-retest reliability.
+For the rhyme decision task, we then computed an LI using a *t*-test for individual participants based on log-transformed response time data.
+For the remaining tasks (which made use of count data) we computed a *z*-statistic for each participant based on proportions.
+
+For each measure, LIs and their test-retest correlation coefficients are reported in Table 2.
+Note that these are computed so that in each case a negative LI indicates left lateralisation (faster RTs or greater accuracy) at the brain level while a positive LI indicates right lateralisation at the brain level.
+In addition, Table 1 includes t-values and corresponding *p*-values for one-sample *t* tests for each measure, to test whether the mean LI departed significantly from zero.
+In addition Shapiro-Wilks test results are shown, to test departurs from normality.
+As the distributions of most LIs were non-normal, Spearman's rank correlations were used to estimate the test-retest reliability.
 Confidence intervals were estimated using the *spearman.ci()* function from the RVAideMemoire package (version 0.9-75; Herv, 2020) with 10,000 iterations.
-From Table 1, it is clear that we were able to obtain good test-retest reliability when calculating LIs in a typical manner.
-For all but the rhyme decision task, we see good test-retest reliability that exceeds our pre-registered threshold of *r*= .65.
-Furthermore, LIs for each task across the two sessions are lateralised at the participant level (all *p*s< .001).
-Specifically, the rhyme decision, dichotic listening, and finger tapping tasks were all lateralised to the left hemisphere (showing a left visual field advantage).
-The chimeric faces task was lateralised to the right hemisphere (showing a left visual field advantage).
+Furthermore, LIs for each task across the two sessions were lateralised at the group level (all *p*s< .001).
+Specifically, the rhyme decision, dichotic listening, and finger tapping tasks were all lateralised to the left hemisphere (showing a left hemisphere advantage).
+The chimeric faces task was lateralised to the right hemisphere.
 A highly similar pattern of results emerges for the *t*- and *z*-score measures, where each is lateralised at the population level (all *p*s< .001).
-The test-retest correlations are shown visually in Figure 5.
+For all but the rhyme decision task, we obtained good test-retest reliability that exceeded our pre-registered threshold of *r*= .65.The scatterplots depicting test-retest reliability are shown in Figure 4.
 
@@
 densplots <- function(dat, time1,time2,name1,name2,scattername,combotable,combocolumn, y1, y2, x1, x2, ax, ay){
-
 # plots
@@
 combotable[5,combocolumn]<-round(shap1$p.value,3)
-
 # day 2
@@
 combotable[5,(combocolumn+1)]<-round(shap2$p.value,3)
-
 # conduct Spearman's wth bootstrapped CI
@@
 ypos <- max(time2,na.rm=T)*.9
-
 mycor<-cor.test(time1, time2, method = "spearman", alternative = "greater")$estimate
-
 #Make a label to show the correlation with 95%CI and N on the plot
@@
 ggsave(scattername,plot=bigplot,dpi = 300, width = 6, height = 6,)
-
 return(combotable)
@@
 ```
+
 <!--- We start by completing combotable for the 4 main measures.
@@
 combotable<-densplots(RD.dat, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)
-
 # DICHOTIC LISTENING
@@
 combotable<-densplots(DL.dat, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)
-
 # CHIMERIC FACES
@@
 combotable<-densplots(CF.dat, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)
-
 # FINGER TAPPING
@@
 combotable<-densplots(RD.dat, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)
-
 # DICHOTIC Z-TEST
@@
 combotable<-densplots(DL.dat, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)
-
 # CHIMERIC FACES Z-TEST
@@
 combotable<-densplots(CF.dat, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)
-
 # FINGER TAPPING Z-TEST
@@
 ax= 6
-ay= -6
+ay= 6
 combotable<-densplots(LIs, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)
@@
 myft <- flextable(
-  head(combotable))
+  combotable)
 myft
@@
 
-```{r makecontingencies,echo=FALSE, comment=FALSE, warning=FALSE}
-#We start by using z-scores to make a 3-way categorisation into:
-#1 sig bias in typical direction; 2: NS bias 3: in atypical direction with sig bias.
-LIs$rhyme1.cut <- cut(LIs$logt1, 
-                       breaks = c(-Inf, -1.96, 1.96, Inf), 
-                       labels = c("RD TypLat", "RD NonLat", "RD AtypLat"), 
-                       right = FALSE)
-LIs$dichotic1.cut <- cut(LIs$Day1_Dich_acc_Z, 
-                       breaks = c(-Inf, -1.96, 1.96, Inf), 
-                       labels = c("DL TypLat", "DL NonLat", "DL AtypLat"), 
-                       right = FALSE)
-LIs$chimeric1.cut <- cut(LIs$Day1_CF_acc_Z, 
-                       breaks =  c(-Inf, -1.96, 1.96, Inf), 
-                       labels = c("CF AtypLat", "CF NonLat", "CF TypLat"), 
-                       right = FALSE)
-mytabr <- table(LIs$rhyme1.cut)
-mytabd <- table(LIs$dichotic1.cut)
-mytabc <- table(LIs$chimeric1.cut)
-mytabc<-mytabc[c(3,2,1)]
-mytab2 <-rbind(mytabr,mytabd,mytabc)
-row.names(mytab2)<-c('Rhyme decision','Dichotic','Chimeric Faces')
-mytab2<-data.frame(mytab2)
-# now do chi-squared tests
-# rhyme
-tabr <- table(LIs$handed3,LIs$rhyme1.cut)
-rhy.chi <- chisq.test(tabr)
-# dichotic
-tabd <- table(LIs$handed3,LIs$dichotic1.cut)
-dic.chi <- chisq.test(tabd)
-# chimeric
-tabc <- table(LIs$handed3,LIs$chimeric1.cut)
-chim.chi <- chisq.test(tabc)
-# combine tables for printing
-mytab3 <- cbind(tabr, tabd, tabc)
-# BUT these don't actually run the planed tests. We run these by removing lefties here
-#Planned test omitted group 'Left' and column Nonlat.
-chi.r<-chisq.test(tabr[c(1,3),c(1,3)])
-chi.d<-chisq.test(tabd[c(1,3),c(1,3)])
-chi.c<-chisq.test(tabc[c(1,3),c(1,3)])
-```
-
 Table 2. Frequencies of laterality typicality for right-, left-, and extreme left-handers across the Rhyme decision, Dichotic listening, and Chimeric faces tasks.
-```{r Table 2,echo=FALSE, comment=FALSE, warning=FALSE}
+```{r Table2,echo=FALSE, comment=FALSE, warning=FALSE}
 mytab4 <- as.data.frame(mytab3)
@@
 cocor.result4 <- cocor(~Day2_CF_acc_LI + negEHI | Day2_CF_acc_LI + Day2_finger_LI, LIs)
-
 # ASSOCIATIONS
-
 # using self report
@@
   col_keys = c(" ", "DL TypLat", "DL NonLat", "DL AtypLat", "CF TypLat", "CF NonLat", "CF AtypLat"))
-
 # using finger tapping
@@
   col_keys = c(" ", "DL TypLat", "DL NonLat", "DL AtypLat", "CF TypLat", "CF NonLat", "CF AtypLat"))
-
 # using EHI
@@
 DL.lex.est1<-cor.test(interR$Day2_DL_acc_LI, interR$lexTALE, method = "spearman")
-
 # footedness
@@
 CF.ft.t.2 <- t.test(foot.dat$Day2_CF_acc_LI~ foot.dat$footedness)
-
 # sightedness
@@
   theme(plot.title = element_text(hjust = 0.5))
-
 # NOW DO RIGHT HANDERS
@@
 fit3B <- cfa(model3, data=right_dat)
-
 aic_vector2 <- c(fitMeasures(fit1B, "aic"),fitMeasures(fit2B, "aic"),fitMeasures(fit3B, "aic"))
-
 #summary(fit1B, fit.measures=TRUE)
@@
 #summary(fit3B, fit.measures=TRUE)
-
 AICweights2<-akaike.weights(aic_vector2)$weights
-
 IC_objectweights2<-data.frame(Model=factor(rep(c('Model 1','Model 2','Model 3'))), IC = factor(rep('AIC',each=3)),values=AICweights2)
-
 right.weights <- ggplot(IC_objectweights2,aes(Model,values)) +
@@
   theme(plot.title = element_text(hjust = 0.5))
-
 model.comp <- ggpubr::ggarrange(left.weights, right.weights, ncol = 2)
@@
 <!---
-
 DOROTHY, I'VE LEFT THESE NOTES IN INCASE YOU FIND THEM USEFUL
-
 # Pre-reg info
-
 ## Exploratory analysis
-
 As noted above, we will use the z- or t-statistics to index laterality, but we will have a large dataset that might be used subsequently to explore different approaches to measuring a laterality index, and to consider how choice of measure affects test-retest reliability.
-
 ## Results
-
 ## Question 4
@@
 Specifically we had predicted that in right-handers the two language tasks would correlate together, but the chimeric faces would form a separate factor, but in left-handers, the language tasks might fractionate.
-
 This analysis was handicapped by the fact that one of the language indices, the rhyme judgement task, lacked both reliability and validity.
@@
 This is motivated by the prior observations that on both dichotic listening and chimeric faces, there was some association between handedness and lateralised task performance.
-
 We had planned to use a structural equation modelling framework to compare the covariances of the laterality statistics between the two language tasks and chimeric faces to test between three models, separately for left- and right-handers.
 However, in neither handedness group were there any meaningful correlations between laterality indices.
-
 ```{r prep_dat,echo=FALSE}
@@
 ```
-
 ```{r showcorrs,echo=FALSE}
@@
 ```
-
 ### Table 3: Intratask correlations using session 1 data.
@@
 ```
-
 The correlations between tasks are all very low. To check if this pattern is robust, this analysis was repeated for time 2 data.
-
 ### Table 4: Intratask correlations using session 2 data.
@@
 --->
+
+References from DB
+
+Bless, J.
+J., Westerhausen, R., Arciuli, J., Kompus, K., Gudmundsen, M., & Hugdahl, K.
+(2013).
+“Right on all occasions?” – On the feasibility of laterality research using a smartphone dichotic listening application.
+Frontiers in Psychology,, 4(42).
+doi:10.3389/fpsyg.2013.00042
+
+Kimura, D. (1967). Functional asymmetry of the brain in dichotic listening. Cortex, 3, 163-178.
+
+Mishkin, M., & Forgays, D. G. (1952). Word recognition as a function of retinal locus. Journal of Experimental Psychology, 43(1), 43-48. doi:10.1037/h0061361

### 2020-07-17 battery-preprocessing_v5_DB.Rmd
now removes data properly from "for_figs.csv"

@@
 for (d in 1:2){
-  data_file <- "https://osf.io/mhk29/download" #day 1 data
+  data_file <- "https://osf.io/mhk29/download" # day 1 data
 
   if (d ==2){
-    data_file <-  "https://osf.io/ebm6s/download"
+    data_file <-  "https://osf.io/ebm6s/download" # day 2 data
   }
@@
 mycolname<-"Day2_Finger_acc_Z"
-polarity<- (1)#so LI is positive in predicted direction
+polarity<- (-1)#so LI is positive in predicted direction
 allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)
@@
 
+# now calculate % correct
+plot.data <- plot.data %>%
+  mutate(remove= if_else(plot.data$task== "RD" & plot.data$p.corr < .75,1,
+                         if_else(plot.data$accurate <= 10,1,0)))
+plot.data <- plot.data[plot.data$remove== 0,]
+
 # save data frame
@@
 # LOOP THROUGH TRIAL SIZE
-for (y in 2:trial) {
+for (y in 1:trial) {
 # task
@@
 CF.corrs <- na.omit(CF.corrs)
-CF.cor.plot <- ggplot(CF.corrs, aes(x=trial, y=cor)) +
-  geom_hline(yintercept=.65, linetype="dashed", color = "blue") +
-  geom_hline(yintercept=.75, linetype="dashed", color = "red") +
+CF.cor.plot <- ggplot(abs(CF.corrs), aes(x=trial, y=cor)) +
+  geom_hline(yintercept=.65, linetype="dashed", color = "red") +
   geom_line(size=1, alpha=0.8) + ylim(0, 1) + theme_classic() +
@@
 Based on a couple of rounds with this, removing a lower limit on trials correct here is the only way to examine if less than 1 block would be enough to achieve r= .75.
-From this analysis, it appears that 27 trials would be sufficient to run the chimeric faces wirth reasonable reliability.
+From this analysis, it appears that 25 trials would be sufficient to run the chimeric faces wirth reasonable reliability.
 Three repetitions of all combinations would lead to 48 trials (36 with different emotions in each hemifield, 12 same-same trials).
-This would ensure full balancing as 27 items would not allow an equal presentation of each of the four emotions across hemifaces.
+This would ensure full balancing as 25 items would not allow an equal presentation of each of the four emotions across hemifaces.
 
@@
 ```{r reducerDL}
-{for (d in 1:2){ #we will process day 1 and day 2 together
+for (d in 1:2){ #we will process day 1 and day 2 together
   dat <- dichotic_dat1
@@
 DL.corrs <- na.omit(DL.corrs)
-DL.cor.plot <- ggplot(DL.corrs, aes(x=trial, y=cor)) +
-  geom_hline(yintercept=.65, linetype="dashed", color = "blue") +
-  geom_hline(yintercept=.75, linetype="dashed", color = "red") +
+DL.cor.plot <- ggplot(abs(DL.corrs), aes(x=trial, y=cor)) +
+  geom_hline(yintercept=.65, linetype="dashed", color = "red") +
   geom_line(size=1, alpha=0.8) + ylim(0, 1) + theme_classic() +
@@
 # print number of trials needed
-print(paste0("Number of chimeric faces required for at least .65 test-retest: ", min(DL.corrs[DL.corrs$cor >= .65,]$trial)))
-print(paste0("Number of chimeric faces required for at least .75 test-retest: ", min(DL.corrs[DL.corrs$cor >= .75,]$trial)))
-}
+print(paste0("Number of dichotic required for at least .65 test-retest: ", min(DL.corrs[DL.corrs$cor >= .65,]$trial)))
+print(paste0("Number of dichotic required for at least .75 test-retest: ", min(DL.corrs[DL.corrs$cor >= .75,]$trial)))
 

### 2020-07-18 battery-preprocessing_v6_DB.Rmd
1. Rather than creating separate for_figs.csv file, add the L and R raw scores to allsum for all variables. 2. Remove all analyses for raw RT in RDT, since we preregistered use of log RT scores. Have done this now by just substituting logRT at an early stage and deleting separate section for logs. As this involves deleting chunks of stuff I have created new version of this script. 3. For DL2 there were 7 cases with very few trials - this looked like a technical error. I removed their data - see chunk where techerrs is defined. 4. The t-test laterality index for RDT had got lost somewhere in tr

--- 2020-07-06 battery-preprocessing_v5_DB.Rmd
@@
 
+18th July DB edits
+1. Rather than creating separate for_figs.csv file, add the L and R raw scores to allsum for all variables.
+2.
+Remove all analyses for raw RT in RDT, since we preregistered use of log RT scores.
+Have done this now by just substituting logRT at an early stage and deleting separate section for logs.
+As this involves deleting chunks of stuff I have created new version of this script.
+3. For DL2 there were 7 cases with very few trials - this looked like a technical error. I removed their data - see chunk where techerrs is defined.
+4. The t-test laterality index for RDT had got lost somewhere in translation. I've put it back. It actually gives a test-retest reliability of .63! Ho hum.
+
 1st July 2020: AP edits
@@
 
-09 June 2020: AP made changes to calcultaions of rhyme
+09 June 2020: AP made changes to calculations of rhyme
 (1) Added t scores to output
@@
 
-# recode variabels
+# recode variables
 EHI$subject <- EHI$Participant.Public.ID
@@
     temp_dat$p.corr[myrow] <- 100*mean(tmp1$Correct,na.rm=TRUE) # add row for %accurate responses - NB for chimeric only error is wrong emotion - measure is just a measure of choice.
-For RDT, can make error of selecting wrong sIDe, so % correct is meaningful
+For RDT, can make error of selecting wrong side, so % correct is meaningful
 
@@
 # lets keep a running check on the participants completing here
+
 complete.day2.CF <- unique(CF_dat2$subject)
+
 ```
@@
 dopirate(origdat,task,measure)
+
+#added by DB 17th July 2020: include raw accuracy L and R in the allsum dataframe
+chimdatL1<-select(filter(chim_dat1,side=='Left'),subject,accurate)
+chimdatR1<-select(filter(chim_dat1,side=='Right'),subject,accurate)
+chimdatL2<-select(filter(chim_dat2,side=='Left'),subject,accurate)
+chimdatR2<-select(filter(chim_dat2,side=='Right'),subject,accurate)
+
+allsum <- merge(allsum,chimdatL1,by="subject")
+allsum <- merge(allsum,chimdatR1,by="subject")
+allsum <- merge(allsum,chimdatL2,by="subject")
+allsum <- merge(allsum,chimdatR2,by="subject")
+myn <-ncol(allsum)
+colnames(allsum)[(myn-3):myn] <- c('CF1.Lraw','CF1.Rraw','CF2.Lraw','CF2.Rraw')
+
 ```
@@
 w<-union(which(allsum$sound.recog1 < .75), which(allsum$sound.recog2 < .75))
+
 ww<-intersect(which(allsum$exclude==1),w)
+#Also exclude if less than 75% accuracy on the different trials
+
 #ww contains people who were excluded on both dichotic and faces. 
@@
 
+We have to remove cases where v few trials because of technical problems. Specify minimum of 80 (out of 120) trials. (Most detected by this proces have less than 10 trials!).
+NB This is only run for session 2 because there were no technical problems for session 1.
+
+```{r marktecherr}
+techerr<-table(dichotic_dat2$subject)
+te<-which(techerr<80)
+techerr<-techerr[te] #subject numbers for those with few trials.
+techerr<-as.numeric(names(techerr))
+```
 The data is visualised using the pirate plot function and t.test conducted.
@@
 ```
+```{r add.dichotic.raw}
+#added by DB 17th July 2020: include raw accuracy L and R in the allsum dataframe
+DLdatL1<-select(filter(DL_dat1,side=='Left'),subject,accurate)
+DLdatR1<-select(filter(DL_dat1,side=='Right'),subject,accurate)
+DLdatL2<-select(filter(DL_dat2,side=='Left'),subject,accurate)
+DLdatR2<-select(filter(DL_dat2,side=='Right'),subject,accurate)
+
+#Add to allsum
+allsum <- merge(allsum,DLdatL1,by="subject")
+allsum <- merge(allsum,DLdatR1,by="subject")
+allsum <- merge(allsum,DLdatL2,by="subject")
+allsum <- merge(allsum,DLdatR2,by="subject")
+myn <-ncol(allsum)
+colnames(allsum)[(myn-3):myn] <- c('DL1.Lraw','DL1.Rraw','DL2.Lraw','DL2.Rraw')
+#remove cases with technical errors from session 2
+allsum$DL2.Lraw[techerr]<-NA
+allsum$DL2.Rraw[techerr]<-NA
+
+```
 
@@
 allsum$exclude[w]<-2 #mark as excluded for dichotic
+#Now remove values for the cases with technical error (v few trials)
+allsum$Day2_DL_RT_LI[techerr] <-NA
+allsum$Day2_DL_acc_LI[techerr] <-NA
 
@@
 ```{r rdt_outliers}
-RDT_dat1<-detect.outliers(RDT_dat1,200,1.65) #numbers specify min RT and 
-RDT_dat2<-detect.outliers(RDT_dat2,200,1.65) #numbers specify min RT and zscore for Hoaglin-Iglewicz respectively
+#First we create a RT column that is based on LOGRT
+RDT_dat1$rawRT <- RDT_dat1$RT
+RDT_dat2$rawRT <- RDT_dat2$RT
+RDT_dat1$RT <- log(RDT_dat1$rawRT)
+RDT_dat2$RT <- log(RDT_dat2$rawRT)
+RDT_dat1<-detect.outliers(RDT_dat1,log(200),1.65) #numbers specify min RT and 
+RDT_dat2<-detect.outliers(RDT_dat2,log(200),1.65) #numbers specify min RT and zscore for Hoaglin-Iglewicz respectively
 
@@
 ```{r makedf_rhyme}
-#We create chim_dat files for day1 and day 2
+#We create RDT files for day1 and day 2
 varlist <-   c("subject", "side", "accurate", "RT") #need to be in this order, ie sub, side, acc and RR
@@
 
-Comparison shows that there is significant difference in RT.
+Comparison shows that there is significant side difference in RT.
 Accuracy is close to ceiling and does not show side difference (but it would be a bit odd for this task, - unlike CF and dichotic, there is no competing information and only one answer is correct).
@@
 }
+```
+
+```{r add.rdt.raw}
+#added by DB 17th July 2020: include raw accuracy L and R in the allsum dataframe
+RDTdatL1<-select(filter(RDT_df1,side=='left'),subject,RT)
+RDTdatR1<-select(filter(RDT_df1,side=='right'),subject,RT)
+RDTdatL2<-select(filter(RDT_df2,side=='left'),subject,RT)
+RDTdatR2<-select(filter(RDT_df2,side=='right'),subject,RT)
+
+allsum <- merge(allsum,RDTdatL1,by="subject")
+allsum <- merge(allsum,RDTdatR1,by="subject")
+allsum <- merge(allsum,RDTdatL2,by="subject")
+allsum <- merge(allsum,RDTdatR2,by="subject")
+myn <-ncol(allsum)
+colnames(allsum)[(myn-3):myn] <- c('RDTlog1.Lraw','RDTlog1.Rraw','RDTlog2.Lraw','RDTlog2.Rraw')
+
 ```
@@
 polarity <- (1)
+mycutoff <- 0 #no cutoff for logRT data
 allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
@@
 abline(h=0)
-text(-3,9,paste0('r = ',round(cor2,3)))
-```
+text(-1,1,paste0('r = ',round(cor2,3)))
+```
+### T-test as alternative to LI for RDT
+
+Advantage of using a t-statistics to represent LI is that we can also identify people whose L-R difference is significantly different from zero.
+
+```{r do.ttest}
+for (d in 1:2){
+  myfile<-RDT_dat1
+  if(d==2){
+    myfile<-RDT_dat2}
+
+  myfile<-filter(myfile,side %in% c('left','right'))
+
+  nsubs <- nrow(allsum)
+  allsum$tlat_RDT<-NA
+
+  for (i in 1:nsubs) {
+    w<-which(myfile$subject==allsum$subject[i])
+    if(length(w)>10)
+    {
+      myt<-t.test(myfile$RT[w]~myfile$side[w]) #this is based on logRT
+      allsum$tlat_RDT[i]<-myt$statistic
+    }
+  }
+  if (d==1){
+    maxcol <-length(colnames(allsum))
+    colnames(allsum)[maxcol]<-c('tlat_RDT_1')
+  }
+  if (d==2){
+    maxcol <-length(colnames(allsum))
+    colnames(allsum)[maxcol]<-c('tlat_RDT_2')
+  }
+}
+
+colcount<-length(colnames(allsum))
+
+w<-which(is.na(allsum$Day1.RDT_RT_LI)) #remove those with NA in LI
+allsum[w,(colcount-1):colcount]<-NA
+w<-which(is.na(allsum$Day2.RDT_RT_LI))
+allsum[w,(colcount-1):colcount]<-NA
+
+#compare t-value from day 1 and day 2 
+myr <- cor(allsum$tlat_RDT_1,allsum$tlat_RDT_2,use='complete.obs')
+plot(allsum$tlat_RDT_1,allsum$tlat_RDT_2,xlim=c(-5,5),ylim=c(-5,5))
+abline(v=0)
+abline(h=0)
+abline(v=-1.65,lty=2,col='red') 
+abline(h=-1.65,lty=2,col='red')
+abline(v=1.65,lty=2,col='red') 
+abline(h=1.65,lty=2,col='red')
+text(-3,3,paste0('r = ',round(myr,3)))
+
+plot(allsum$Day1.RDT_RT_LI,allsum$tlat_RDT_1,xlim=c(-1.5,1.5),ylim=c(-5,5))
+```
+
 ## Does RDT relate to handedness?
@@
 
-### T-test as alternative to LI for RDT
-
-The LI results are not encouraging - reliability well below acceptable level, and although RT shows lateralised bias, it is tiny. Also, many subjects dropped.
-Previously I also did a t-test that was analogous to the z-test for CF/dichotic, i.e.
-showed if the individual had a significant RT difference between L and R stimuli.
-But that is v highly correlated with LI.
-The only way to make it more sensitive would be to try a pairwise t-test that compares RTs for the same items presented on L and R.
-So the next bit of code is designed to compute that.
-
-```{r RDTaggregate}
-#remove 'no rhyme' trials
-for (d in 1:2){
-  thisdat<-RDT_dat2[RDT_dat2$side %in% c('left','right'),]
-  if(d==1){
-    thisdat<-RDT_dat1[RDT_dat1$side %in% c('left','right'),]
-  }
-
-  thisdat <- thisdat[thisdat$outlier==0,] #remove outliers
-  aggdata <- aggregate(thisdat$RT,by=list(thisdat$side,thisdat$item,thisdat$Participant.Public.ID),
-                       FUN=mean,na.rm=TRUE)
-  aggdata$subject <- aggdata$Participant.Public.ID
-  colnames(aggdata)<-c('side','item','subject','RT')
-  mysubs<-unique(aggdata$subject)
-  if(d==1){
-  pairedt <- data.frame(matrix(NA,nrow=length(mysubs),ncol=5))
-  colnames(pairedt) <- c('subject','t1','p1','t2','p2')
-  }
-  nsubs<-length(mysubs)
-  for (i in 1:nsubs){
-    myt<-NA
-    myp<-NA
-    thisdat <- aggdata[aggdata$subject==mysubs[i],]
-    pairedt$subject[i]<- mysubs[i]
-    #check there are data for both left and right on all items
-    checkdat <-table(thisdat$item)
-    w<-which(checkdat<2)
-    if(length(w)==0){
-      t<-t.test(thisdat$RT[thisdat$side=='left'],thisdat$RT[thisdat$side=='right'],paired=T)
-      myt<-t$statistic
-      myp<-t$p.value
-    }
-    if(d==1){
-      pairedt$t1[i] <- myt
-      pairedt$p1[i] <-myp
-    }
-    if(d==2){
-      pairedt$t2[i] <-myt
-      pairedt$p2[i] <-myp
-    }
-  }
-}
-
-# add t score to the data
-t_dat <- select(pairedt, -3, -5)
-t_dat <- na.omit(t_dat)
-# merge to allsum for analysis
-allsum <- merge(allsum, t_dat, by.x = "origID", by.y = "subject", all.x=T)
-
-cor(pairedt$t1,pairedt$t2,use='complete.obs')
-```
-
-So the paired items analysis gives a totally nonsignificant correlation between time 1 and time 2. This is much worse than the overall LI reliability.
-
-The failure of this task led to further thoughts about why it is barely lateralised, given that it does implicate language-specific word generation. 
-A noteworthy feature of the other two tasks (CF and dichotic) is that they measure side preference rather than accuracy - i.e.
-a response from either L or R is correct, but two potential responses are put in competition.
-That contrasts with this task where only one side has the correct answer.
-We used pictures because with written words there is a concern that  preference may be confounded by the fact we read words from L to R.
-That could potentially dilute the RVF advantage that is expected from L hem language.
-
-### log RT
-
-In our pre-reg, we also specified that we would use logRT. So we add this in here.
-
-```{r logRT analysis}
-# now add the logRT to original output
-# first relabel df
-RDT_logdat1 <- RDT_dat1
-RDT_logdat2 <- RDT_dat2
-# now log the RT
-RDT_logdat1$RT <- log(RDT_dat1$RT)
-RDT_logdat2$RT <- log(RDT_dat2$RT)
-# Now use generic function to remove outliers.
-# Here we keep the 1.65 of the H&I removal. 
-# BUT we now specify log 200 for lower cutoff to move onto the same scale
-RDT_logdat1<-detect.outliers(RDT_logdat1,log(200),1.65) #numbers specify min RT and 
-RDT_logdat2<-detect.outliers(RDT_logdat2,log(200),1.65) #numbers specify min RT and zscore for Hoaglin-Iglewicz respectiv
-# We create logRT for day1 and day 2
-varlist <-   c("subject", "side", "accurate", "RT") #need to be in this order, ie sub, side, acc and RR
-latlist <- c("left","no rhyme","right") #names of factor levels
-nloop<-length(latlist)
-RDT_log_df1 <- make.df(RDT_logdat1,varlist,latlist,nloop)
-RDT_log_df2 <- make.df(RDT_logdat2,varlist,latlist,nloop)
-# remove NaN
-is.nan.data.frame <- function(x)
-  do.call(cbind, lapply(x, is.nan))
-RDT_log_df1[is.nan(RDT_log_df1)] <- NA
-RDT_log_df2[is.nan(RDT_log_df2)] <- NA
-# now let's plot the log data
-for (d in 1:2){
-origdat <-RDT_log_df1[RDT_log_df1$subject %in% incRDT,]
-if (d==2)
-{origdat <-RDT_log_df2[RDT_log_df2$subject %in% incRDT,]}
-origdat <-origdat[origdat$side %in% c('left','right'),]
-# RT
-origdat$X1 <- origdat$RT
-measure <- 'log-RT'
-dopirate(origdat,task,measure)
-}
-# make LI for log RT
-# Because outliers will already be marked in allsum, we won't bother going through this again. 
-# We will intead add straight into the data frame. 
-#now create LI
-for (d in 1:2){
-origdat <-RDT_log_df1[RDT_log_df1$subject %in% incRDT,]
-if (d==2)
-{origdat <-RDT_log_df2[RDT_log_df2$subject %in% incRDT,]}
-# log RT
-origdat$side <- as.factor(origdat$side)
-  levels(origdat$side) <-c('Left','No rhyme','Right') #function assumes capitalised
-polarity<- 1 #so LI is positive in predicted direction
-mycutoff <- 0
-origdat$X1 <- origdat$RT
-mycolname<-"Day1_RDT_logRT_LI"
-if(d==2){mycolname<-"Day2_RDT_logRT_LI"}
-polarity <- (1)
-allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
-}
-# plot log RT
-cor1<-cor(allsum$Day1_RDT_logRT_LI,allsum$Day2_RDT_logRT_LI,use='complete.obs')
-plot(allsum$Day1_RDT_logRT_LI,allsum$Day2_RDT_logRT_LI,main='Rhyme detect: LI for log RT')
-abline(v=0)
-abline(h=0)
-text(-1,1,paste0('r = ',round(cor1,3)))
-# FINALLY, make t-test for logRT
-#remove 'no rhyme' trials
-for (d in 1:2){
-  thisdat<-RDT_logdat2[RDT_logdat2$side %in% c('left','right'),]
-  if(d==1){
-    thisdat<-RDT_logdat1[RDT_logdat1$side %in% c('left','right'),]
-  }
-
-  thisdat <- thisdat[thisdat$outlier==0,] #remove outliers
-  aggdata <- aggregate(thisdat$RT,by=list(thisdat$side,thisdat$item,thisdat$Participant.Public.ID),
-                       FUN=mean,na.rm=TRUE)
-  aggdata$subject <- aggdata$Participant.Public.ID
-  colnames(aggdata)<-c('side','item','subject','RT')
-  mysubs<-unique(aggdata$subject)
-  if(d==1){
-  pairedt <- data.frame(matrix(NA,nrow=length(mysubs),ncol=5))
-  colnames(pairedt) <- c('subject','t1','p1','t2','p2')
-  }
-  nsubs<-length(mysubs)
-  for (i in 1:nsubs){
-    myt<-NA
-    myp<-NA
-    thisdat <- aggdata[aggdata$subject==mysubs[i],]
-    pairedt$subject[i]<- mysubs[i]
-    #check there are data for both left and right on all items
-    checkdat <-table(thisdat$item)
-    w<-which(checkdat<2)
-    if(length(w)==0){
-      t<-t.test(thisdat$RT[thisdat$side=='left'],thisdat$RT[thisdat$side=='right'],paired=T)
-      myt<-t$statistic
-      myp<-t$p.value
-    }
-    if(d==1){
-      pairedt$t1[i] <- myt
-      pairedt$p1[i] <-myp
-    }
-    if(d==2){
-      pairedt$t2[i] <-myt
-      pairedt$p2[i] <-myp
-    }
-  }
-}
-
-# add t score to the data
-t_dat <- select(pairedt, -3, -5)
-t_dat <- na.omit(t_dat)
-# relabel for merging
-names(t_dat)[names(t_dat) == "t1"] <- "logt1"
-names(t_dat)[names(t_dat) == "t2"] <- "logt2"
-# merge to allsum for analysis
-allsum <- merge(allsum, t_dat, by.x = "origID", by.y = "subject", all.x=T)
-
-cor(pairedt$t1,pairedt$t2,use='complete.obs')
-```
-
-Even when log-transforming, things don't look any better.
-
 ## Finger Tapping Task
@@
 
-In the raw data file, farticipants, identified by random letter strings, are stacked on top of each other. We start by reading the data and cutting out unwanted information.
+In the raw data file, participants, identified by random letter strings, are stacked on top of each other. We start by reading the data and cutting out unwanted information.
 
@@
     wholeseq <-paste0(tempfinger$Response,collapse='') #make a long string of responses
-    for (s in 1:2){
+    for (s in 1:2){ #cycle through sequences for L and R
       myc<-mycols[d,s] #select column
       target <- myseq[s] #tumb then wrvx, ie right then left
-      finger.df[n,myc]<-str_count(wholeseq,target) 
+      mystrs<-str_count(wholeseq,target) 
+      finger.df[n,myc]<-mystrs
       #count how many target sequences are in the long string that contains all responses
-
-    }
+      if(s==1){mystrs0<-mystrs}
+    }
+    allseqd <-mystrs0+mystrs
+    findbad <- abs(nchar(wholeseq)/4-allseqd) #allseqd x 4 should equal length of wholeseq if there are no wrong keypresses.
+      if (findbad>20) #if more than 20 wrong keypresses, then exclude.
+        {finger.df[n,mycols[d,1]]<-NA
+        finger.df[n,mycols[d,2]]<-NA}
     #We then add the SD of the RT - this indicates variability in pace of keypressing
@@
 }
-# Substitute NA for any totals less than 5 - suggest using wrong keys perhaps
+# Substitute NA for any totals less than 5
 
@@
 ```
+```{r add.fingerL_R}
+#added by DB 17th July 2020: include raw accuracy L and R in the allsum dataframe
+fingerL1<-select(finger.df,subject,L1)
+fingerR1<-select(finger.df,subject,R1)
+fingerL2<-select(finger.df,subject,L2)
+fingerR2<-select(finger.df,subject,R2)
+
+allsum <- merge(allsum,fingerL1,by="subject")
+allsum <- merge(allsum,fingerR1,by="subject")
+allsum <- merge(allsum,fingerL2,by="subject")
+allsum <- merge(allsum,fingerR2,by="subject")
+myn <-ncol(allsum)
+colnames(allsum)[(myn-3):myn] <- c('finger.L1','finger.R1','finger.L2','finger.R2')
+
+```
 
@@
 mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
-polarity<- (-1)#so LI is positive in predicted direction
+polarity<- (1)#so LI is positive in predicted direction
 allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
@@
 cor1 <- cor(allsum$Day1_finger_LI,allsum$Day2_finger_LI,use='complete.obs')
-plot(allsum$Day1_finger_LI,allsum$Day2_finger_LI,main='Finger Tapping LI')
+plot(allsum$Day1_finger_LI,allsum$Day2_finger_LI,main='Finger Tapping LI',col=allsum$handpch,pch=15)
 abline(v=0)
 abline(h=0)
-text(-20,10,paste0('r = ',round(cor1,3)))
+abline(0,1,lty=2)
+text(-10,10,paste0('r = ',round(cor1,3)))
 ```
@@
 mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
-polarity<- (-1)#so LI is positive in predicted direction
+polarity<- (1)#so LI is positive in predicted direction
 allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)
@@
 abline(h=0)
-text(-2.5,4,paste0('r = ',round(cor1,3)))
+abline(0,1,lty=2)
+text(-2,3.5,paste0('r = ',round(cor1,3)))
 
@@
 
-````{r lister}
+```{r lister}
 # CF
@@
 # now write this for filtering
-write.csv(allsum, "LIs.csv")
+write.csv(allsum, "./data/allsum.csv")
 ```
@@
 write.csv(plot.data, "./data/for_figs.csv")
+
 ```

### 2020-07-19 battery-preprocessing_v6_DB.Rmd
Some more tweaks to z-tests. All polarity set so L hemisphere advantage is positive, to be consistent with main papers in the literature - and with Doppler.

@@
 library(beeswarm) #just a kind of plot Dorothy likes
+library(BSDA) #for ztest
 ```
@@
 
+# 19 Jul 2020:settled on setting polarity so that L hemisphere advantage is positive
+
 makeLI <- function(allsum,origdat,task,mycolname,polarity,mycutoff){
@@
 mycolname<-"Day1_CF_RT_LI"
-polarity <-1
+polarity <-(-1)
 
@@
 mycolname<-"Day2_CF_acc_LI"
-polarity<- (-1)#so LI is positive in predicted direction
+polarity<- (1)#so LI is positive in predicted direction
 allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
@@
 mycolname<-"Day2_CF_RT_LI"
-polarity <-1
+polarity <-(-1)
 allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
@@
 mycolname<-"Day1_DL_RT_LI"
-polarity <-1
+polarity <-(-1)
 
@@
 mycolname<-"Day2_DL_acc_LI"
-polarity<- (-1)#so LI is positive in predicted direction
+polarity<- (1)#so LI is positive in predicted direction
 allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
@@
 mycolname<-"Day2_DL_RT_LI"
-polarity <-1
+polarity <-(-1)
 allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
@@
 mycolname<-"Day2_Dich_acc_Z"
-polarity<- (-1)#so LI is positive in predicted direction
+polarity<- (1)#so LI is positive in predicted direction
 allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)
@@
 if(d==2){mycolname<-"Day2.RDT_p.corr_LI"}
-polarity<- 1 #so LI is positive in predicted direction
+polarity<- (-1) #so LI is positive in predicted direction
 mycutoff <- 75
@@
 if(d==2){mycolname<-"Day2.RDT_RT_LI"}
-polarity <- (1)
+polarity <- (-1)
 mycutoff <- 0 #no cutoff for logRT data
@@
 
-  myfile<-filter(myfile,side %in% c('left','right'))
-
   nsubs <- nrow(allsum)
-  allsum$tlat_RDT<-NA
+  allsum$tempt<-NA #initialise a new column
 
   for (i in 1:nsubs) {
-    w<-which(myfile$subject==allsum$subject[i])
-    if(length(w)>10)
+    mysub<-filter(myfile,side %in% c('left','right'),subject==allsum$subject[i])
+    if(nrow(mysub)>10)
     {
-      myt<-t.test(myfile$RT[w]~myfile$side[w]) #this is based on logRT
-      allsum$tlat_RDT[i]<-myt$statistic
-    }
-  }
-  if (d==1){
-    maxcol <-length(colnames(allsum))
-    colnames(allsum)[maxcol]<-c('tlat_RDT_1')
-  }
-  if (d==2){
-    maxcol <-length(colnames(allsum))
-    colnames(allsum)[maxcol]<-c('tlat_RDT_2')
-  }
-}
-
-colcount<-length(colnames(allsum))
+      myt<-t.test(mysub$RT~mysub$side) #this is based on logRT
+      allsum$tempt[i]<-myt$statistic
+    }
+  }
+  wc<-which(colnames(allsum)=='tempt')
+  colnames(allsum)[wc]<-paste0('tlat_RDT_',d)
+
+}
 
 w<-which(is.na(allsum$Day1.RDT_RT_LI)) #remove those with NA in LI
-allsum[w,(colcount-1):colcount]<-NA
+allsum$tlat_RDT_1[w]<-NA
 w<-which(is.na(allsum$Day2.RDT_RT_LI))
-allsum[w,(colcount-1):colcount]<-NA
+allsum$tlat_RDT_2[w]<-NA
 
@@
 plot(allsum$Day1.RDT_RT_LI,allsum$tlat_RDT_1,xlim=c(-1.5,1.5),ylim=c(-5,5))
+```
+
+In fact, for better consistency with the other measures, could do z-test instead of t-test. These converge to the same thing with samples > around 30, which we have here.  So here is that calculation.
+
+```{r make.zlatRDT}
+for (d in 1:2){
+  myfile<-RDT_dat1
+  if(d==2){
+    myfile<-RDT_dat2}
+
+  nsubs <- nrow(allsum)
+  allsum$tempt<-NA #initialise a new column
+
+  for (i in 1:nsubs) {
+    mysub<-filter(myfile,side %in% c('left','right'),subject==allsum$subject[i])
+    if(nrow(mysub)>10)
+    {
+
+      mymean<-aggregate(RT~side,mysub,FUN = mean)
+      mysd<-aggregate(RT~side,mysub,FUN = sd)
+        myN<-aggregate(RT~side,mysub,FUN = length)
+      se1<-mysd[1,2]/sqrt(myN[1,2])
+      se2<-mysd[2,2]/sqrt(myN[2,2])
+      se<-sqrt(se1^2+se2^2)
+      allsum$tempt[i]<-(mymean[1,2]-mymean[2,2])/se
+    }
+  }
+  wc<-which(colnames(allsum)=='tempt')
+  colnames(allsum)[wc]<-paste0('zlat_RDT_',d)
+
+}
+
+w<-which(is.na(allsum$Day1.RDT_RT_LI)) #remove those with NA in LI
+allsum$zlat_RDT_1[w]<-NA
+w<-which(is.na(allsum$Day2.RDT_RT_LI))
+allsum$zlat_RDT_2[w]<-NA
+
+#compare t-value from day 1 and day 2 
+myr <- cor(allsum$zlat_RDT_1,allsum$zlat_RDT_2,use='complete.obs')
+plot(allsum$zlat_RDT_1,allsum$zlat_RDT_2,xlim=c(-5,5),ylim=c(-5,5))
+abline(v=0)
+abline(h=0)
+abline(v=-1.65,lty=2,col='red') 
+abline(h=-1.65,lty=2,col='red')
+abline(v=1.65,lty=2,col='red') 
+abline(h=1.65,lty=2,col='red')
+text(-3,3,paste0('r = ',round(myr,3)))
+
+plot(allsum$Day1.RDT_RT_LI,allsum$zlat_RDT_1,xlim=c(-1.5,1.5),ylim=c(-5,5))
+plot(allsum$tlat_RDT_1,allsum$zlat_RDT_1,xlim=c(-1.5,1.5),ylim=c(-5,5))
+abline(0,1)
+
+#Confirms tlat is same as zlat for these sample sizes!
 ```

### 2020-07-20 battery-preprocessing_v6_DB.Rmd
Added percents correct on RDT to allsum.

@@
 4. The t-test laterality index for RDT had got lost somewhere in translation. I've put it back. It actually gives a test-retest reliability of .63! Ho hum.
+5.
+Realised that it would be simplest to use a z-score for the standardized laterality for all measures, and that for the RDT RT this is actually equivalent to the t-value (just double checked that in one chunk - they are equivalent).
+6. Redid polarity of measures to be consistent with Hugdahl and others - LEFT brain is now positive.
 
@@
 mycolname<-"Day2_CF_acc_Z"
-polarity<- (-1)#so LI is positive in predicted direction
+polarity<- (1)#so LI is positive in predicted direction
 allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)

### 2020-07-20 battery-preprocessing_v6_DB.Rmd
including %correct for RDT in allsum.csv

@@
 
+#Also record % correct for each category in allsum
+for (d in 1:2){
+  myfile<-RDT_df1
+  if(d==2){ myfile<-RDT_df2}
+pcorL<-myfile[myfile$side=='left',]
+pcorR<- myfile[myfile$side=='right',]
+pcorNull <- myfile[myfile$side=='no rhyme',]
+pcorL<-select(pcorL,subject,p.corr)
+pcorR<-select(pcorR,subject,p.corr)
+pcorNull<-select(pcorNull,subject,p.corr)
+colnames(pcorL)[2]<-paste0('RDT',d,'.pcorrL')
+colnames(pcorR)[2]<-paste0('RDT',d,'.pcorrR')
+colnames(pcorNull)[2]<-paste0('RDT',d,'.pcorrNull')
+allsum <- merge(allsum,pcorL,by="subject")
+allsum <- merge(allsum,pcorR,by="subject")
+allsum <- merge(allsum,pcorNull,by="subject")
+}
+
 #Scrutiny of these reveal a few cases who missed all 'no rhyme' cases.
@@
 print(paste0('RDT: number remaining after exclusions = ',ninc))
+
 ```

### 2020-07-20 online_laterality_reliability.Rmd

@@
 For left-handers, we anticipated that that model (c) may be a better fit to the data, indicating that laterality is dissociated across the different tasks.
-However, we were not able to test this planned analysis with the current data set as one language task had low reliability.
-Thus, we confined our analysis to presentation of correlations between laterality indices from the four measures.
 
@@
 The response triggered the next trial.
+Accuracy and response time were recorded.
 
@@
 library(ggpubr)
+library(corrr)
+library(flextable)
+options(knitr.kable.NA = '') #should suppress NA in tables
 # read in data from csv
-LIs <- read.csv('https://osf.io/x5jd2/download',stringsAsFactors = F)
+#LIs <- read.csv('https://osf.io/x5jd2/download',stringsAsFactors = F)
+#We will replace with read from OSF, but better stick with LIs to ensure final version is the one we use 
+allsum <- read.csv('./data/allsum.csv',stringsAsFactors = F)
+
 ```
@@
 ```{r prepdat,echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
-# remove those without day 2
+# remove those without day 2 to make LIs file - we will just use this
 # reformat NA
-LIs$day2.complete <- ifelse(is.na(LIs$day2.complete), 0, LIs$day2.complete)
+allsum$day2.complete <- ifelse(is.na(allsum$day2.complete), 0, allsum$day2.complete)
+LIs <- allsum[allsum$day2.complete == 1,]
+
 # remove
 LIs_c <- LIs[LIs$day2.complete == 1,]
-LIs <- LIs[LIs$day2.complete == 1,]
+
 # lextale out df
@@
 Applying each of these procedures meant that the number of participants left in the final analyses could vary by task.
-First, for all tasks, we removed trials in which reaction times were shorter than 200 ms.
-Outlier reaction times were identified using Hoaglin and Iglewicz’s (1987) procedure, whereby outliers are defined as reaction times that are 1.65 times the difference between the first and third quantiles (1.65 * (Q3-Q1)) below or above the first and third quantile values (e.g: lower limit = Q1 – 1.65 x (Q3-Q1); upper limit = Q3 + 1.65 x (Q3-Q1)).
-Our differences criterion was set to 1.65 instead of the typical 2.2 because we expected to obtain asymmetric distributions for reaction time, since reaction time is always non-negative, and so most outliers would be in one tail of the distribution.
-The Hoaglin and Iglewicz procedure was applied to each participants’ data separately.
+First, for all tasks, we removed trials in which response times were shorter than 200 ms.
+Response times were log-transformed to reduce skew, as specified in our pre-registration.
+Outlier reaction times were identified using Hoaglin and Iglewicz’s (1987) procedure, whereby outliers are defined as response times that are 1.65 times the difference between the first and third quantiles (1.65 * (Q3-Q1)) below or above the first and third quantile values (e.g: lower limit = Q1 – 1.65 x (Q3-Q1); upper limit = Q3 + 1.65 x (Q3-Q1)).
+Our differences criterion was set to 1.65 instead of the typical 2.2 because we expected to obtain asymmetric distributions for response time, with most outliers in one tail of the distribution.
+The Hoaglin and Iglewicz procedure was applied to each participant's data separately.
 After applying each data trimming procedure, participants left with less than 80% of trials were removed.
@@
 These data trimming procedures left data for `r nrow(LIs)-nrow(LIs[LIs$exclude== 1 | LIs$exclude== 12,])` participants on the chimeric faces `r nrow(LIs)-nrow(LIs[LIs$exclude== 2 | LIs$exclude== 12,])` participants on the dichotic listening task.
+An additional 7 cases had to be removed because of a technical error that meant they had less than 80 trials presented.
 
@@
 We considered recruiting further participants to make up the numbers, but this seemed uneconomic as we already had more than 300 cases for the other tasks, and we could tell by this point that the test-retest reliability of the rhyme detection task was unsatisfactory; With `r nrow(LIs[LIs$exRDT== 0,])` participants, the test-retest reliability was *r*= `r round(cor(LIs[LIs$exRDT== 0,]$Day1.RDT_RT_LI, LIs[LIs$exRDT== 0,]$Day2.RDT_RT_LI, use = "complete.obs"),2)`, which is smaller than the *r*= 0.65 that we pre-registered as our criterion for a reliable task.
-Accordingly, we stopped data collection at this point and analysed the `r nrow(LIs)` participants with data on other tasks.
+We plan to evaluate modified versions of this task in future, but decided to stop data collection at this point and analysed the `r nrow(LIs)` participants with data on other tasks.
+
+For the finger-tapping task we added a data-cleaning step that had not been pre-registered, to remove participants where there were more than 20 keypresses across left and right hands that did not correspond to the  intended sequence.
+This could reflect that the individual had forgotten the target sequence, or was not co-operating with instructions, but scrutiny of cases where this occurred suggested that it usually arose because of technical problems.
+For instance, some individuals had repeated recording of the same keypress at short intervals, suggestive of a sticky keyboard.
+The cutoff of 20 was arbitrarily selected, and was successful in excluding cases with abnormally low scores in one session.
 
@@
 
-For the rhyme task, we pre-registered that we would calculate LIs based on response time (RT) data, as accuracy during piloting was at ceiling.
-We specified that we would calculate LIs in two ways: (1) in the traditional fashion (100 x (Right – Left)/(Right + Left)); (2) to correspond to a *t*-value.
-This was achieved by conducting a *t*-test for each participant where accurate RTs are the dependent variable and visual field is the independent variable.
-This latter measure is highly correlated with the conventional laterality index but has the advantage that it can be used to identify participants who show a statistically significant RT advantage for one side, i.e.
-where the *t*-value on a 2-sided test has *p*< .05.
-Based on considerations from pilot data, we adopted log-transformation of RTs before computing *t*-values.
+We preregistered that we would calculate LIs in two ways: (1) in the traditional fashion (100 x (Right – Left)/(Right + Left)); (2) a standardized LI that involved computing a z-score that represented, for each individual, a comparison between left- and right-sided responses.
+In fact, for RT data we specified a t-score, but t and z converge when sample size exceeds 30, and for simplicity we refer to these as LI_z scores.
+These LI_z scores are very highly correlated with the traditional LI measures (correlations typically exceed .95), but can be used to provide a common metric for RT and accuracy measures when identifying individuals who are significantly lateralised.
+For example, using a conventional 2-sided alpha of .05, we can categorise individuals with an absolute LI_z greater than 1.96 as significantly lateralised.
+
+For the rhyme task, we pre-registered that we would calculate LIs based on log response time (RT) data, as accuracy during piloting was at ceiling.
+To obtain LI_z we ran a *t*-test for each participant where accurate log RTs were the dependent variable and visual field was the independent variable.
 
@@
 The count of correct responses that corresponded to each side was used to generate an accuracy LI in two ways: (1) in the traditional fashion (100 x (Right – Left)/(Right + Left)); (2) as a *z*-statistic for each participant, using z = (pR-.5)/sqrt(pR x pL/n), where pR is the proportion of correct responses in the RVF, pL is proportion of L responses in the LVF, and n is total LVF and RVF responses.
-This was analogous to the individual t-scores generated for RTs in the rhyme decision task, in allowing us to identify individuals with a significant bias to one side.
-Finally, LIs for the finger tapping task were calculated in the same manner, but using the number of sequences generated by the left and right hand as opposed to the proportion of correct responses in each visual field.
+This method can yield extreme z-scores (for instance, if 83/89 responses are to the right, then z = 15.92), so scores were censored to be in the range -5 to +5.
+Finally, LIs for the finger tapping task were calculated in the same manner,  using the number of sequences generated by the left and right hand as opposed to the proportion of correct responses in each visual field.
+
+To address hypothesis 1 and examine the laterality of each task at the population level, we conducted a series of one-sample *t*-tests for each task using session 1 data.
 
 For the reliability assessment, we checked LIs for normality using a Shapiro-Wilk test.
-As the LIs were non-normal, we used a series of Spearman’s correlations to examine the correlation of coefficients for session 1 versus session 2.
+As most LIs were non-normal, we used a series of Spearman’s correlations to examine the correlation of coefficients for session 1 versus session 2.
 For tasks where the test-retest reliability was above .75 with the existing test length, we reanalysed the data using just the first N trials, reducing N until reliability becomes unsatisfactory (below *r*= .65).
@@
 
-To address hypothesis 1 and examine the laterality of each task at the population level, we conducted a series of one-sample *t*-tests for each task using session 1 data.
-
-##### Atypicality
-
-To address hypothesis 2, that extreme left-handers will be more likely to show atypical laterality than right-handers, we categorised participants as extreme left-handers if they scored less than -90 and right-handers if they scored more than 90 of the Edinburgh Handedness Inventory.
-Atypicality was first defined in participants whose *t*- or *z*-score was significantly lateralised in the opposite direction from usual.
-If there were less than 20 cases defined as atypical using this method, we adjusted the cut-off on that task until we acquired 20 atypical participants.
-Once typicality was defined, we first analysed atypicality on the Rhyme Decision and Chimeric Faces task.
-Participants were categorised into four groups: typical on both; atypical on Rhyme Decision only; atypical on Chimeric Faces only, and atypical on both.
+##### Atypical laterality
+
+To address hypothesis 2, that extreme left-handers will be more likely to show atypical laterality than right-handers, we categorised participants as extreme left-handers if they scored less than -90 and right-handers if they scored more than 90 on the Edinburgh Handedness Inventory.
+Our preregistration specified that atypicality would first be defined according to whether the participant's LI_z was significantly lateralised (with alpha = .05) in the opposite direction from the majority.
+If there were less than 20 cases defined as atypical using this method, we planned to adjust the cut-off on that task until we acquired 20 atypical participants.
+
+For Rhyme Decision and Chimeric Faces task, participants were categorised into four groups: typical on both; atypical on Rhyme Decision only; atypical on Chimeric Faces only, and atypical on both.
 A chi-square test was then used to test for association with the number of people in each of the four groups and the handedness classification from the Edinburgh Handedness Inventory.
@@
 
-```{r descriptives, echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
-# here we present the raw count and RT scores for each task
-# read CF data
-for_figs <- read.csv("for_figs.csv", na.strings = "NA")
-
-# get CF
-CF.short.dat <- for_figs[for_figs$exclude== 0 | for_figs$exclude== 2,]
-  CF.short.dat <-CF.short.dat[CF.short.dat$task== "CF",]
-# read Dichotic data
-DL.short.dat <- for_figs[for_figs$exclude== 0 | for_figs$exclude== 1,]
-  DL.short.dat <- DL.short.dat[DL.short.dat$task== "DL",]
-# read rhyme data
-RD.short.dat <- for_figs[for_figs$exRDT== 0,]
-  RD.short.dat <- RD.short.dat[RD.short.dat$task== "RD",]
-# read finger data
-finger.short.dat <- read.csv("./data/finger.csv", na.strings = "NA")
-# plots
-CF <- ggplot(CF.short.dat, aes(x = side, y = accurate)) +
-        ggpirate::geom_pirate(aes(colour = side)) +
-        theme_classic() + ylab("Count of correct responses") + xlab("") + facet_wrap(.~Session)+ 
-        ggtitle("C") + theme(plot.title = element_text(hjust = 0.5))
-DL <- ggplot(DL.short.dat, aes(x = side, y = accurate)) +
-        ggpirate::geom_pirate(aes(colour = side)) +
-        theme_classic() + ylab("Count of correct responses") + xlab("") + facet_wrap(.~Session)+ 
-        ggtitle("B") + theme(plot.title = element_text(hjust = 0.5))
-RD <- ggplot(RD.short.dat, aes(x = side, y = RT)) +
-        ggpirate::geom_pirate(aes(colour = side)) +
-        theme_classic() + ylab("Reaction time (ms)") + xlab("") + facet_wrap(.~Session) + 
-        ggtitle("A") + theme(plot.title = element_text(hjust = 0.5))
-FT <- ggplot(finger.short.dat, aes(x = side, y = sequences)) +
-        ggpirate::geom_pirate(aes(colour = side)) +
-        theme_classic() + ylab("Count of correct sequences") + xlab("") + facet_wrap(.~Session)+ 
-        ggtitle("D") + theme(plot.title = element_text(hjust = 0.5))
-all.plot <- suppressMessages(suppressWarnings(ggpubr::ggarrange(RD, DL, CF, FT)))
-all.plot
-# save
-ggsave(
-  "all.vf.png",
-  plot = all.plot,
-  width = 6, height = 6,
-  dpi = 300
-)
-```
-
-<!---These figures aren't so good for showing laterality because they don't show how the L and R scores relate in individuals. Will try an alternative approach.--->
-
 ```{r altplotraw,echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
 
-makerawplot <- function(file1,file2,mysession,myname,mylim){
-
-tempL<-filter(file1,Session==mysession)
-tempR<-filter(file2,Session==mysession)
-meanL<-mean(tempL$tempcol,na.rm=T)
-meanR<-mean(tempR$tempcol,na.rm=T)
-tempLR<-as.data.frame(cbind(tempL$tempcol,tempR$tempcol))
-colnames(tempLR)<-c('Left','Right')
+makerawplot <- function(tempLR,mysession,myname,mylim){
+  colnames(tempLR)<-c('Left','Right','Handedness')
+  tempLR<-tempLR[!is.na(tempLR$Handedness), ] #remove rows with missing data
+  tempLR$Handedness<-as.factor(tempLR$Handedness)
+  droplevels(tempLR$Handedness)
+  levels(tempLR$Handedness)=c('Right','Left')
+  #myalpha<-c(1,.75) #haven't been able to adjust cols or pch this way. alpha works but one is always v faint regardless of setting
+  meanL<-mean(tempLR$Left,na.rm=T)
+  meanR<-mean(tempLR$Right,na.rm=T)
   mytitle<-paste0(myname,"\n Session",mysession)
-myg <- ggplot(tempLR, aes(x=Left, y=Right)) + geom_point(size=.5) +
+  myg <- ggplot(tempLR, aes(x=Left, y=Right,col=Handedness)) + geom_point(size=.5) +
     xlim(mylim)+ylim(mylim)+
@@
     geom_hline(yintercept=meanR,linetype='dotted')+geom_vline(xintercept=meanL,linetype='dotted')+
-    theme(plot.title = element_text(size = 12))
+    theme(legend.position = "none") +
+    theme(plot.title = element_text(size = 12)
+    )
+  myg<-myg+ scale_color_manual(values=c("black","red"))
   return(myg)
@@
 
-RD.short.dat$tempcol<-RD.short.dat$RT
-file1 <-filter(RD.short.dat,side=='left')
-file2 <-filter(RD.short.dat,side=='right')
 myname<-'Rhyme judge RT'
-mysession<-1
+tempLR<-select(LIs[LIs$exRDT==0,],RDTlog1.Lraw,RDTlog1.Rraw,handpch)
+#convert back to raw time
+tempLR[,1:2]<-exp(tempLR[,1:2])
+session<-1
 mylim<-c(500,1500)
-myg1<-makerawplot(file1,file2,mysession,myname,mylim)
-myg2<-makerawplot(file1,file2,2,myname,mylim)
-
-DL.short.dat$tempcol<-DL.short.dat$accurate
-file1 <-filter(DL.short.dat,side=='left')
-file2 <-filter(DL.short.dat,side=='right')
+myg1<-makerawplot(tempLR,session,myname,mylim)
+
+tempLR<-select(LIs[LIs$exRDT==0,],RDTlog2.Lraw,RDTlog2.Rraw,handpch)
+tempLR[,1:2]<-exp(tempLR[,1:2]) #convert from log back to original scale for plot
+session<-2
+myg2<-makerawplot(tempLR,session,myname,mylim)
+
 myname<-'Dichotic accuracy'
-mysession<-1
+tempLR<-select(LIs[LIs$exclude<2,],DL1.Lraw,DL1.Rraw,handpch)
+session<-1
 mylim<-c(0,100)
-myg3<-makerawplot(file1,file2,mysession,myname,mylim)
-myg4<-makerawplot(file1,file2,2,myname,mylim)
-
-CF.short.dat$tempcol<-CF.short.dat$accurate
-#there are 96 stimuli with different faces, so 48 is pt of equality if all trials included
-file1 <-filter(CF.short.dat,side=='left')
-file2 <-filter(CF.short.dat,side=='right')
+myg3<-makerawplot(tempLR,session,myname,mylim)
+tempLR<-select(LIs[LIs$exclude<2,],DL2.Lraw,DL2.Rraw,handpch)
+session<-2
+myg4<-makerawplot(tempLR,session,myname,mylim)
+
 myname<-'Chimeric faces accuracy'
-mylim<-c(0,96)
-mysession<-1
-myg5<-makerawplot(file1,file2,mysession,myname,mylim)
-myg6<-makerawplot(file1,file2,2,myname,mylim)
-
-finger.short.dat$tempcol<-finger.short.dat$sequences
-file1 <-filter(finger.short.dat,side=='left')
-file2 <-filter(finger.short.dat,side=='right')
-myname<-'Finger tap sequences'
-mysession<-1
-mylim<-c(0,175)
-myg7<-makerawplot(file1,file2,mysession,myname,mylim)
-myg8<-makerawplot(file1,file2,2,myname,mylim)
-
-ggarrange(myg1,myg2,myg3,myg4,myg5,myg6,myg7,myg8, ncol = 2, nrow = 4)
+#LIs exclude of 1 or 12 indicates excluded from CF; retain exclude 0,2
+tempLR<-select(LIs[LIs$exclude %in% c(0,2),],CF1.Lraw,CF1.Rraw,handpch)
+session<-1
+mylim<-c(0,100)
+myg5<-makerawplot(tempLR,session,myname,mylim)
+tempLR<-select(LIs[LIs$exclude %in% c(0,2),],CF2.Lraw,CF2.Rraw,handpch)
+session<-2
+myg6<-makerawplot(tempLR,session,myname,mylim)
+
+myname<-'N finger sequences'
+tempLR<-select(LIs,finger.L1,finger.R1,handpch)
+session<-1
+mylim<-c(50,180)
+myg7<-makerawplot(tempLR,session,myname,mylim)
+tempLR<-select(LIs,finger.L2,finger.R2,handpch)
+session<-2
+myg8<-makerawplot(tempLR,session,myname,mylim)
+
+ggarrange(myg1,myg2,myg3,myg4,myg5,myg6,myg7,myg8, ncol = 2, nrow = 4,common.legend=T,legend="bottom")
 ggsave(
-  "all.raw.png",
+  "./plots/all.raw.png",
   width = 5, height = 8,
@@
 
-### Laterality indices: means and test-retest reliability
-
-Accuracy on the rhyme task was high  (*M*= `r round(mean(RD.short.dat$p.corr, na.rm= TRUE),1)`%, *SD*= `r round(sd(RD.short.dat$p.corr, na.rm= TRUE),2)`%), and only correct response times were used for computing a laterality index.
-Average reaction times on this task are shown in Figure 3 as vertical and horizontal lines in a scatterplot showing right vs left visual field, separately for each session.
-A sloping dotted line shows the point where left and right RTs are equal; if there is faster performance for the right visual field, as predicted, then points should cluster below this line.
-Similar scatterplots are shown for selection of left and right responses for dichotic listening and chimeric faces.
-Note that for these two tasks, either side can give a correct response; the negative correlation between the two sides reflects the fact that the number of selections on the left will be inversely related to the number of correct responses on the right.
-For these tasks it is evident by inspection that most responses for dichotic listening cluster above the dotted line, indicating a right ear advantage, whereas most responses on chimeric faces cluster below the dotted line.
+```{r computepcorrRDT}
+#revised this, as p.corr values now in allsum
+RDT.short.dat<-allsum[allsum$exRDT==0,]
+tempRDT <- select(RDT.short.dat,RDT1.pcorrL,RDT1.pcorrR,RDT2.pcorrL,RDT2.pcorrR)
+allmeanpcorr <- mean(colMeans(tempRDT,na.rm=T))
+allsdpcorr <- sd(c(tempRDT$RDT1.pcorrL,tempRDT$RDT1.pcorrR,tempRDT$RDT2.pcorrL,tempRDT$RDT2.pcorrR),na.rm=T)
+```
+### Significance of lateralisation on each task
+
+For a preliminary impression of the data, Figure 3 shows scatterplots indicating scores on left and right sides for each task, with handedness colour-coded.
+For each task, mean for each side is shown as vertical and horizontal dotted lines, and equal performance on the two sides corresponds to the sloping red dotted line: thus an impression of whether the task is lateralised can be obtained by considering whether points cluster around the sloping lines symmetrically.
+
+For the rhyme task, accuracy was high (*M* = `r round(allmeanpcorr,1)`%, *SD* = `r round(allsdpcorr,1)` %),and only correct responses were used for computing mean log RTs (shown transformed back to original ms scale in the plot).
+For dichotic listening and chimeric faces a correct response can be given from either side; it follows that there will be a negative correlation between the two sides reflecting the fact that the number of selections on the left will be inversely related to the number of correct responses on the right.
+For these tasks it is evident by inspection that most responses for dichotic listening cluster above the dotted line, indicating a right ear advantage, whereas most responses on chimeric faces cluster below the dotted line, indicating a left hemi-face advantage.
 Finally, we show the number of correct sequences with each hand on the finger-tapping task.
-Here, as expected, inspection of the data suggests more sequences with the right than the left hand on average.
+Here, a strong impact of handedness is evident on inspection, with left-handers generally more accurate with the left hand, and right-handers more accurate with the right hand.
 
@@
 
-The count of correct answers on the dichotic listening and chimeric faces tasks are also shown in Figure 3, along with the number of sequences generated by the left and right hand.
-For all tasks, we computed LIs in the typical way: (100 x (Right – Left)/(Right + Left)).
-It is important to note, however, that we present LIs such that they are reported to be consistent with lateralisation at the brain level (i.e.
-by multiplying LIs by -1).
-A negative LI indicates left lateralisation at the brain level while a positive LI indicates right lateralisation at the brain level.
-For the rhyme decision task, we then computed an LI using a *t*-test for individual participants based on log-transformed response time data.
-For the remaining tasks (which made use of count data) we computed a *z*-statistic for each participant based on proportions.
-
-For each measure, LIs and their test-retest correlation coefficients are reported in Table 2.
-Note that these are computed so that in each case a negative LI indicates left lateralisation (faster RTs or greater accuracy) at the brain level while a positive LI indicates right lateralisation at the brain level.
-In addition, Table 1 includes t-values and corresponding *p*-values for one-sample *t* tests for each measure, to test whether the mean LI departed significantly from zero.
-In addition Shapiro-Wilks test results are shown, to test departurs from normality.
-As the distributions of most LIs were non-normal, Spearman's rank correlations were used to estimate the test-retest reliability.
+The distribution of LIs can be used to quantify departures from symmetry.
+Recall that for all tasks, we computed LIs in the typical way: (100 x (Right – Left)/(Right + Left)), and that because of crossed pathways from hands, ears and eyes to brain, a positive LI indicates left lateralisation and a negative LI indicates right lateralisation at the brain level.
+<!---I flipped this for consistency with Hugdahl and others, as it was clear from emails that they found it confusing to have left-brain as negative--->.
+Table 2 shows these mean LIs, as well a one-sample t-test comparing the mean LI with zero, and results from a Shapiro-Wilks test for departure from normality.
+It can be seen that all measures were significantly lateralised in the predicted direction at the population level, with left-hemisphere superiority for rhyme judgement, dichotic listening and finger tapping, and right-hemisphere superiority for chimeric faces.
+Most distributions departed significantly from normality.
+
+<!---for ease of interpretation I have chopped up the original big table so we can look at different aspects one step at a time; i.e.
+overall distribution of LIs (different from zero / skew); then the standardized LIs; then the test-retest.--->
+
+```{r make.tab2,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
+#Make a table to summarise LI means/distribution stats for 4 tasks
+#This will gradually get populated as we go through each task
+#Cells for each task are filled in course of plots.function chunk
+LItable <- data.frame(matrix(NA,nrow=8,ncol=7))
+colnames(LItable)<-c('Task','N','Mean','SD','One sample t','One sample p','Normality p')
+LItable[,1]<-c('Rhyme Decision 1','Rhyme Decision 2','Dichotic 1','Dichotic 2','Chimeric Faces 1','Chimeric Faces 2','Finger Tapping 1','Finger Tapping 2')
+
+keynames <- c("Day1.RDT_RT_LI"  ,"Day2.RDT_RT_LI" ,"Day1_DL_acc_LI" , "Day2_DL_acc_LI",   "Day1_CF_acc_LI", "Day2_CF_acc_LI","Day1_finger_LI", "Day2_finger_LI"   )
+for (i in 1:length(keynames)){
+  keycol<-which(names(LIs)==keynames[i]) #find this column
+#exclude appropriately
+   if (i<7){
+  thisdata <- LIs[LIs$exclude %in% c(0,1),]}
+  if (i<5){
+  thisdata <- LIs[LIs$exclude %in% c(0,2),]}
+    if (i<3){
+  thisdata <- LIs[LIs$exRDT==0,]}
+  temp<-thisdata[!is.na(thisdata[,keycol]),keycol]
+
+  LItable[i,2]<-length(temp)
+  LItable[i,3]<-round(mean(temp),2)
+  LItable[i,4]<-round(sd(temp),2)
+  LItable[i,5]<-round(t.test(temp)$statistic,2)
+  pstring <- round(t.test(temp)$p.value,3)
+  ifelse(pstring<.001,pstring<-'<0.001',pstring<-char(pstring))
+
+  LItable[i,6]<-pstring
+  pstring <- round(shapiro.test(temp)$p.value,3)
+  ifelse(pstring<.001,pstring<-'<0.001',pstring<-as.character(pstring))
+
+  LItable[i,7]<-pstring
+
+}
+```
+
+Table 3 shows data from the LI_z measures, which make it easier to compare different tasks on the same scale.
+For each task, those with LI_z scores greater than 1.96 are categorised as left-hemisphere dominant, and those with LI_z scores less than -1.96 are categorised as right-hemisphere dominant.
+If there were no population bias, we would expect around 2.5% of individuals to meet each of these criteria.
+It can be seen that for dichotic listening and chimeric faces, around 45% of individuals have hemispheric dominance in the predicted direction and around 9% have hemispheric dominance in the opposite direction.
+The proportions showing significant laterality are lower for the rhyme decision task (around 17% left-hemisphere dominant, and 5% right-hemisphere dominant), and for finger-tapping, the proportions are only slightly greater than would be expected by chance, although there is a suggestion of increasing laterality on that task across the two sessions.
+
+```{r make.tab3,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
+#Make a table to summarise LI means/distribution stats for 4 tasks this time using LI_z.
+#This will gradually get populated as we go through each task
+#Cells for each task are filled in course of plots.function chunk
+LIztable <- data.frame(matrix(NA,nrow=8,ncol=6))
+colnames(LIztable)<-c('Task','N','Mean','SD','L-lateralised %','R-lateralised %')
+LIztable[,1]<-c('Rhyme Decision 1','Rhyme Decision 2','Dichotic 1','Dichotic 2','Chimeric Faces 1','Chimeric Faces 2','Finger Tapping 1','Finger Tapping 2')
+
+keynames <- c("zlat_RDT_1"  ,"zlat_RDT_2" ,"Day1_Dich_acc_Z" , "Day2_Dich_acc_Z",   "Day1_CF_acc_Z", "Day2_CF_acc_Z","Day1_Finger_acc_Z", "Day2_Finger_acc_Z"   )
+for (i in 1:length(keynames)){
+  keycol<-which(names(LIs)==keynames[i]) #find this column
+   if (i<7){
+  thisdata <- LIs[LIs$exclude %in% c(0,1),]}
+  if (i<5){
+  thisdata <- LIs[LIs$exclude %in% c(0,2),]}
+    if (i<3){
+  thisdata <- LIs[LIs$exRDT==0,]}
+  temp<-thisdata[!is.na(thisdata[,keycol]),keycol]
+
+  LIztable[i,2]<-length(temp)
+  LIztable[i,3]<-round(mean(temp),2)
+  LIztable[i,4]<-round(sd(temp),2)
+  LIztable[i,5]<-round(100*length(which(temp>1.96))/length(temp),1)
+  LIztable[i,6]<-round(100*length(which(temp<(-1.96)))/length(temp),1)
+
+}
+```
+
+### Test-retest reliability
+The scatterplots depicting test-retest reliability are shown in Figure 4, with Spearman test-retest correlations in Table 4.
+These were based on LI_z values, but are essentially identical to values from unscaled LI values.
 Confidence intervals were estimated using the *spearman.ci()* function from the RVAideMemoire package (version 0.9-75; Herv, 2020) with 10,000 iterations.
-Furthermore, LIs for each task across the two sessions were lateralised at the group level (all *p*s< .001).
-Specifically, the rhyme decision, dichotic listening, and finger tapping tasks were all lateralised to the left hemisphere (showing a left hemisphere advantage).
-The chimeric faces task was lateralised to the right hemisphere.
-A highly similar pattern of results emerges for the *t*- and *z*-score measures, where each is lateralised at the population level (all *p*s< .001).
-For all but the rhyme decision task, we obtained good test-retest reliability that exceeded our pre-registered threshold of *r*= .65.The scatterplots depicting test-retest reliability are shown in Figure 4.
-
-```{r make.combotable,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
+For all but the rhyme decision task, we obtained good test-retest reliability that exceeded our pre-registered threshold of *r*= .65.
+```{r make.combotablex,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
 #Make a table to summarise results for all 4 tasks
@@
 #Cells for each task are filled in course of plots.function chunk
-combotable <- data.frame(matrix(NA,nrow=7,ncol=17))
-colnames(combotable)<-c('.','RD 1','RD 2','DL 1','DL 2','CF 1','CF 2','FT 1','FT 2', 
-                            'RD t 1','RD t 2','DL z 1','DL z 2','CF z 1','CF z 2','FT z 1','FT z 2')
-combotable[,1]<-c('N','mean','SD','p-lateralised','p-normal','test-retest r','CI r')
+combotable <- data.frame(matrix(NA,nrow=4,ncol=3))
+colnames(combotable)<-c('Task','Spearman r','95% CI')
+combotable[,1]<-c('Rhyme Decision' ,'Dichotic Listening','Chimeric Faces','Finger Tapping ')
 ```
@@
 # Adam's code here turned into generic function
-# While we have the relevant columns specified here, I have added steps to populate combotable as we go.
-# Also saving the scatterplots as png
-densplots <- function(dat, time1,time2,name1,name2,scattername,combotable,combocolumn, y1, y2, x1, x2, ax, ay){
+# Also saving the scatterplots as png.
+# Also computes spearmans with CI and returns these from function
+densplots <- function(dat, time1,time2,name1,name2,scattername,y1, y2, x1, x2, ax, ay){
   # plots
@@
   dens1 <- dens1+ geom_vline(xintercept = 0)
-# normality
-shap1<-shapiro.test(time1)
-ttest1 <- t.test(time1) #Check if significantly lateralised
-Nt1 <- length(time1)-length(which(is.na(time1)))
-combotable[1,combocolumn]<-Nt1
-combotable[2,combocolumn]<-round(mean(time1,na.rm=T),2)
-combotable[3,combocolumn]<-round(sd(time1,na.rm=T),2)
-combotable[4,combocolumn]<-round(ttest1$p.value,3)
-combotable[5,combocolumn]<-round(shap1$p.value,3)
+
   # day 2
-# plot denisty
+  # plot density
   dens2<-ggplot(dat, aes(x=time2))+geom_density(linetype="dashed")+theme_classic()+ggtitle(name2)
   dens2 <- dens2+ geom_vline(xintercept = 0)
-# normality
-shap2<-shapiro.test(time2)
-ttest2 <- t.test(time2)
-Nt2 <- length(time2)-length(which(is.na(time2)))
-myN<-min(Nt1,Nt2)
-combotable[2,(combocolumn+1)]<-round(mean(time2,na.rm=T),2)
-combotable[3,(combocolumn+1)]<-round(sd(time2,na.rm=T),2)
-combotable[4,(combocolumn+1)]<-round(ttest2$p.value,3)
-combotable[5,(combocolumn+1)]<-round(shap2$p.value,3)
+
   # conduct Spearman's wth bootstrapped CI
   myspear<-RVAideMemoire::spearman.ci(time1, time2, nrep = 10000, conf.level = 0.95)
-  combotable[6,combocolumn]<-round(myspear$estimate,2)
+
   myci <- paste0(round(myspear$conf.int[1],2),' - ',round(myspear$conf.int[2],2))
-  combotable[7,combocolumn]<-myci
+
   # scatterplot time1 x time2
@@
   #Make a label to show the correlation with 95%CI and N on the plot
-mycorlabel<-paste0('r = ',round(mycor,2),'\n(95% CI: ',round(myspear$conf.int[1],2),'-',round(myspear$conf.int[2],2),')\nN = ',myN)
+  mycorlabel<-paste0('r = ',round(mycor,2),'\n95% CI: ',myci,'\nN = ',length(time1))
   #Find position in top L corner for legend
@@
   legposy <- .85
+  #censor extreme values for plotting only
+  w<-which(time1<(-5) )
+  time1[w] <- (-5)
+  w<-which(time1>5) 
+  time1[w] <- (5)
+  w<-which(time2<(-5) )
+  time2[w] <- (-5)
+  w<-which(time2>5) 
+  time2[w] <- (5)
+  #now do the plot
   myscatter <- ggplot(dat, aes(x= time1, y= time2)) +
@@
   bigplot<-ggExtra::ggMarginal(myscatter, groupColour = TRUE, groupFill = TRUE)
-scattername<-paste0(scattername,".png")
+  scattername<-paste0('./plots/',scattername,".png")
   ggsave(scattername,plot=bigplot,dpi = 300, width = 6, height = 6,)
-return(combotable)
+  return(list(mycor,myci))
 }
@@
 
-<!--- We start by completing combotable for the 4 main measures.
-NB need to check polarity of tasks - also oddity re RDT distributions needs checking, and we said we'd do log transform , not sure we did--->
 ```{r do.plots,LI,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
+#Add a category of handedness that distinguishes extreme left from other left-handers
+#In preregistration, extreme left defined as EHI laterality -90 or less.
+LIs$handed3 <- as.character(LIs$handedness)
+w<-which(LIs$index_EHI<(-89))
+LIs$handed3[w]<-'Extreme left-hander'
+w<-which(LIs$handed3=='Left')
+LIs$handed3[w]<-'Moderate left-hander'
+w<-which(LIs$handed3=='Right')
+LIs$handed3[w]<-'Right-hander'
 # RHYME DECISION
 # remove bad cases
-RD.dat <- LIs[LIs$exRDT == 0,]
+RD.dat <- LIs[!is.na(LIs$handed3),]
+RD.dat <- RD.dat[RD.dat$exRDT == 0,]
 # now plot
 dat <- RD.dat
-time1<-RD.dat$Day1.RDT_RT_LI
-time2<-RD.dat$Day2.RDT_RT_LI
-name1<-"Rhyme decision LI (day 1)"
-name2<-"Rhyme decision LI (day 2)"
+time1<-RD.dat$zlat_RDT_1
+time2<-RD.dat$zlat_RDT_2
+name1<-"Rhyme decision LI_z (day 1)"
+name2<-"Rhyme decision LI_z (day 2)"
 scattername<-"RhymeDecisionScatter"
-combocolumn<-2 #column to write day 1 summary for this task in combotable
-y1= -12
-y2= 12
-x1= -12
-x2= 12
-ax= 8
-ay= -8
-combotable<-densplots(RD.dat, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)
+
+y1= -5
+y2= 5
+x1= -5
+x2= 5
+ax= 2
+ay= -2
+mycors<-densplots(RD.dat, time1,time2,name1,name2,scattername, y1, y2, x1, x2, ax, ay) 
+combotable[1,2]<-round(mycors[[1]],3)
+combotable[1,3]<-mycors[[2]]
+
 # DICHOTIC LISTENING
 # remove bad cases
-DL.dat <- LIs[LIs$exclude == 0 | LIs$exclude == 1,]
-# now plot
-dat <- DL.dat
-time1<-DL.dat$Day1_DL_acc_LI
-time2<-DL.dat$Day2_DL_acc_LI
-name1<-"Dichotic listening LI acc (day 1)"
-name2<-"Dichotic listening LI acc (day 2)"
-scattername<-"DichoticScatter"
-combocolumn<-4 #column to write day 1 summary for this task in combotable
-y1= -100
-y2= 100
-x1= -100
-x2= 100
-ax= 75
-ay= -75
-combotable<-densplots(DL.dat, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)
-# CHIMERIC FACES
-# remove bad cases
-CF.dat <- LIs[LIs$exclude == 0 | LIs$exclude == 2,]
-# now plot
-dat <- CF.dat
-time1<-CF.dat$Day1_CF_acc_LI
-time2<-CF.dat$Day2_CF_acc_LI
-name1<-"Chimeric faces LI acc (day 1)"
-name2<-"Chimeric faces LI acc (day 2)"
-scattername<-"ChimericScatter"
-combocolumn<-6 #column to write day 1 summary for this task in combotabley1= -100
-y1= -100
-y2= 100
-x1= -100
-x2= 100
-ax= 75
-ay= -75
-combotable<-densplots(CF.dat, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)
-# FINGER TAPPING
-dat <- LIs
-time1<-LIs$Day1_finger_LI
-time2<-LIs$Day2_finger_LI
-name1<-"Finger tapping LI (day 1)"
-name2<-"Finger tapping LI (day 2)"
-scattername<-"FingerScatter"
-combocolumn<-8 #column to write day 1 summary for this task in combotabley2= 100
-y1= -35
-y2= 35
-x1= -35
-x2= 35
-ax= 26.25
-ay= -26.25
-combotable<-densplots(dat, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)
-```
-
-```{r do.plot.t,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
-# RHYME T-TEST
-# now plot
-dat <- RD.dat
-time1<-RD.dat$logt1
-time2<-RD.dat$logt2
-name1<-"Rhyme decision t-score (day 1)"
-name2<-"Rhyme decision t-score (day 2)"
-scattername<-"rhyme.t"
-combocolumn<-10 #column to write day 1 summary for this task in combotable
-y1= -8
-y2= 8
-x1= -8
-x2= 8
-ax= 6
-ay= -6
-combotable<-densplots(RD.dat, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)
-# DICHOTIC Z-TEST
+DL.dat <- LIs[!is.na(LIs$handed3),]
+DL.dat <- DL.dat[DL.dat$exclude == 0 | DL.dat$exclude == 1,]
 # now plot
@@
 time2<-DL.dat$Day2_Dich_acc_Z
-name1<-"Dichotic listening z-score (day 1)"
-name2<-"Dichotic listening z-score (day 2)"
-scattername<-"dichotic.z"
-combocolumn<-12 #column to write day 1 summary for this task in combotable
-y1= -20
-y2= 20
-x1= -20
-x2= 20
-ax= 15
-ay= -15
-combotable<-densplots(DL.dat, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)
-# CHIMERIC FACES Z-TEST
+name1<-"Dichotic listening LI_z (day 1)"
+name2<-"Dichotic listening LI_z (day 2)"
+scattername<-"DichoticScatter"
+
+mycors<-densplots(DL.dat, time1,time2,name1,name2,scattername,y1, y2, x1, x2, ax, ay)
+combotable[2,2]<-round(mycors[[1]],3)
+combotable[2,3]<-mycors[[2]]
+
+# CHIMERIC FACES
+# remove bad cases
+CF.dat <- LIs[!is.na(LIs$handed3),]
+CF.dat <- CF.dat[CF.dat$exclude == 0 | CF.dat$exclude == 2,]
 # now plot
@@
 time2<-CF.dat$Day2_CF_acc_Z
-name1<-"Chimeric faces z-score (day 1)"
-name2<-"Chimeric faces z-score (day 2)"
-scattername<-"chimeric.z"
-combocolumn<-14 #column to write day 1 summary for this task in combotable
-y1= -20
-y2= 20
-x1= -20
-x2= 20
-ax= 15
-ay= -15
-combotable<-densplots(CF.dat, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)
-# FINGER TAPPING Z-TEST
-# now plot
-dat <- LIs
-time1<-LIs$Day1_Finger_acc_Z
-time2<-LIs$Day2_Finger_acc_Z
-name1<-"Finger tapping z-score (day 1)"
-name2<-"Finger tapping z-score (day 2)"
-scattername<-"finger.z"
-combocolumn<-16 #column to write day 1 summary for this task in combotable
-y1= -8
-y2= 8
-x1= -8
-x2= 8
-ax= 6
-ay= 6
-combotable<-densplots(LIs, time1,time2,name1,name2,scattername, combotable,combocolumn, y1, y2, x1, x2, ax, ay)
-```
-
-Table 1. LIs and test-retest correlations for each task.
-```{r Table1,echo=FALSE, comment=FALSE, warning=FALSE, comment=FALSE, warning=FALSE}
-library(flextable)
-library(officer)
-myft <- flextable(
-  combotable)
-myft
-```
+name1<-"Chimeric faces LI_z (day 1)"
+name2<-"Chimeric faces LI_z (day 2)"
+scattername<-"ChimericScatter"
+mycors<-densplots(CF.dat, time1,time2,name1,name2,scattername, y1, y2, x1, x2, ax, ay)
+combotable[3,2]<-round(mycors[[1]],3)
+combotable[3,3]<-mycors[[2]]
+
+# FINGER TAPPING
+FT.dat <- LIs[!is.na(LIs$handed3),]
+FT.dat <- FT.dat[!is.na(FT.dat$Day1_Finger_acc_Z),]
+time1<-FT.dat$Day1_Finger_acc_Z
+time2<-FT.dat$Day2_Finger_acc_Z
+name1<-"Finger tapping LI_z (day 1)"
+name2<-"Finger tapping LI_z (day 2)"
+scattername<-"FingerScatter"
+
+mycors<-densplots(FT.dat, time1,time2,name1,name2,scattername, y1, y2, x1, x2, ax, ay)
+combotable[4,2]<-round(mycors[[1]],3)
+combotable[4,3]<-mycors[[2]]
+```
+
 Note.
-LIs are transformed such that a negative value equates to a left lateralisation at the brain level.
-A positive value equates to right lateralisation at the brain level.
+LIs are scored such that a positive value equates to a left lateralisation at the brain level.
+A positive value equates to left lateralisation at the brain level.
 RD= rhyme decision, DL= dichotic listening, CF: chimeric faces, and FT: finger tapping.
@@
 ![Figure 5.
-Scatter plots for day 1 and day 2 data on each laterality task.
-The left column shows LIs calculated in the typical way.
-The right column shows LIs calculated using *t*- and *z*-scores.
-Here, distributions are shown for left and right handers, and an additional group of extreme left-handers who had a score below -90 on the Edinburgh Handedness Inventory.](combined_LI_cor.png)
-
-Moving forward, we chose to use LIs calculated in a typical manner to address each remaining research question.
+Scatter plots for day 1 and day 2 data for LI_z scores on each laterality task.
+Handedness is colour coded with extreme left-handers identified as those who have a score below -90 on the Edinburgh Handedness Inventory.
+Note that z-scores are censored at absolute values of 5.](combined_LI_cor.png)
+
+### Optimising the tasks
 To determine the smallest number of trials needed to achieve a test-retest reliability of *r*= .65, we recalculated LIs for the dichotic listening and chimeric faces tasks using the range of trials in which different emotions/sounds were presented to each hemiface/ear.
@@
 
-### Atypicality
-
-We defined 'atypical laterality' as laterality that is in the opposite direction to the majority of participant, and which has an absolute *z*/*t*-score value greater than 1.96.
-Based on our pre-registered analysis, we divided left-handers into two groups (1) those with a score of -90 or below on the Edinburgh Handedness Inventory (c.f.
-Mazoyer et al., 2014), (2) the remaining left handers.
-While our analyses include only right handers and extreme left handers, we include the remaining left-handers in Table 2 for completeness.
+### Atypical laterality and handedness
+
+According to Mazoyer et al (2014), extreme left-handers are more likely than other handedness groups to have reversed laterality compared to the rest of the population.
+Our preregistered analysis accordingly focused on comparing extreme left-handers (those with a score of -90 or below on the Edinburgh Handedness Inventory) with right-handers.
+We defined 'atypical laterality' as laterality that is in the opposite direction to the majority of participants, with an absolute LI_z value greater than 1.96.
+Although the focus of our planned analyses included only right handers and extreme left handers, we show data for  the remaining left-handers for completeness.
+
+```{r makecontingencies,echo=FALSE}
+#We start by using z-scores to make a 3-way categorisation into:
+#1 sig bias in typical direction; 2: NS bias 3: in atypical direction with sig bias.
+#
+LIs$rhyme1.cut <- cut(LIs$Day1.RDT_p.corr_LI, 
+                       breaks = c(-Inf, -1.96, 1.96, Inf), 
+                       labels = c("AtypLat", "Nonlat", "TypLat"), 
+                       right = FALSE)
+LIs$dichotic1.cut <- cut(LIs$Day1_Dich_acc_Z, 
+                       breaks = c(-Inf, -1.96, 1.96, Inf), 
+                       labels = c("AtypLat", "Nonlat", "TypLat"), 
+                       right = FALSE)
+LIs$chimeric1.cut <- cut(LIs$Day1_CF_acc_Z, 
+                       breaks =  c(-Inf, -1.96, 1.96, Inf), 
+                       labels = c("AtypLat", "Nonlat", "TypLat"), 
+                       right = FALSE)
+mytabr <- table(LIs$rhyme1.cut)
+mytabd <- table(LIs$dichotic1.cut)
+mytabc <- table(LIs$chimeric1.cut)
+mytabc<-mytabc[c(3,2,1)]
+mytab2 <-rbind(mytabr,mytabd,mytabc)
+row.names(mytab2)<-c('Rhyme decision','Dichotic','Chimeric Faces')
+mytab2<-data.frame(mytab2)
+
+mytab2%>%
+  kable() %>%
+  kable_styling("condensed")
+
+tabr <- table(LIs$handed3,LIs$rhyme1.cut)
+chisq.test(tabr)
+
+tabd <- table(LIs$handed3,LIs$dichotic1.cut)
+chisq.test(tabd)
+
+tabc <- table(LIs$handed3,LIs$chimeric1.cut)
+chisq.test(tabc)
+
+#Planned test omitted group 'Left' and column Nonlat.
+chi.r<-chisq.test(tabr[c(1,3),c(1,3)])
+
+chi.d<-chisq.test(tabd[c(1,3),c(1,3)])
+chi.c<-chisq.test(tabc[c(1,3),c(1,3)])
+
+#ignore the headings of these tables!
+```
 
@@
 ```{r Table2,echo=FALSE, comment=FALSE, warning=FALSE}
-mytab4 <- as.data.frame(mytab3)
-mytab4$` ` <- c("Extreme left", "Left", "Right")
+tabr<-cbind(tabr,round(100*prop.table(tabr,2),1))
+tabd<-cbind(tabd,round(100*prop.table(tabd,2),1) )
+tabc<-cbind(tabc,round(100*prop.table(tabc,2),1)) #add percentages
+blankrow<-c(NA,NA,NA,NA,NA,NA)
+latcats <- as.data.frame(rbind(blankrow,tabr,blankrow,tabd,blankrow,tabc))
+
+row.names(latcats) <- c() #remove rownames as they make problems for printing
+
+#Instead have a new column with these
+latcats$nucol <- c('Rhyme decision','Extreme left-hander','Moderate left-hander','Right-hander','Dichotic listening','Extreme left-hander','Moderate left-hander','Right-hander','Chimeric faces','Extreme left-hander','Moderate left-hander','Right-hander')
+
+latcats<-latcats[,c(7,3,6,2,5,1,4)] #now reorder columns to make sense
+colnames(latcats)<-c('.','L lateralised','L%','Not lateralised','N%','R lateralised','R%')
+
 myft2 <- flextable(
-  head(mytab4),
-  col_keys = c(" ", 
-               "RD TypLat", "RD NonLat", "RD AtypLat",
-               "DL TypLat", "DL NonLat", "DL AtypLat",
-               "CF TypLat", "CF NonLat", "CF AtypLat"))
-myft2
-```
-Note. TypLat: lateralised in the expected direction, Nonlat: no lateralisation, and AtypLat: show lateralisation in the opposite direction.
+  latcats)
+
+#NB This table needs formatting - has formatted all numbers to 3 decimal places, which is not what we want. Code below achieves this
+
+num_keys <- c("L%","N%","R%")
+int_keys <- c("L lateralised","Not lateralised","R lateralised")
+
+ft <- colformat_num(x = myft2, col_keys = num_keys, digits = 1)
+ft <- colformat_int(x = ft, col_keys = int_keys, big.mark = ",")
+
+autofit(ft)
+```
+
+```{r pformatfunction}
+pformat<-function(pvalue){
+  pp <- "*p* < .001" #default value
+  if (pvalue>.001) {
+    pp<-paste0("*p* = ",round(pvalue,3))
+  }
+  return(pp)
+}
+```
 
@@
 We will therefore not consider it further, as our analyses are only justified if the task is an indicator of cerebral lateralisation.
-Both Dichotic Listening and Chimeric Faces, on the other hand, are strongly lateralised.
+Both Dichotic Listening and Chimeric Faces, on the other hand, have a high proportion of strongly lateralised individuals.
 The planned analyses, comparing typically lateralised with atypically lateralised individuals, and extreme left vs right-handed individuals, show significant association between extreme left-handedness and atypical laterality on chi square test.
 For dichotic listening, chi squared = `r round(chi.d$statistic,2)`, d.f.
-= 1, *p* = `r round(p.adjust(chi.d$p.value, method= "bonferroni", n=2),3)`; for chimeric faces, chi squared = `r round(chi.c$statistic,2)`, d.f.
-= 1, *p* = `r round(p.adjust(chi.c$p.value, method= "bonferroni", n=2),3)`.
+= 1, pformat(p.adjust(chi.d$p.value, method= "bonferroni", n=2))`; for chimeric faces, chi squared = `r round(chi.c$statistic,2)`, d.f.
+= 1, pformat(p.adjust(chi.c$p.value, method= "bonferroni", n=2))`.
 
@@
 chi2task<-chisq.test(tab2task)
-w1<-which(LIs$chimeric1.cut=='AtypLat')
+w1<-which(LIs$chimeric1.cut=='TypLat') #this is confusingnly mislabelled
 w2<-which(LIs$dichotic1.cut=='AtypLat')
 w3<-intersect(w1,w2)
-#LIs$handed3[w3]
+atypical2<-LIs$handed3[w3] #this is hardcoded in the text below
 ```
@@
 
-### Relationship between language and non-language LIs
-
-```{r LI-link,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
-# make sure we have good cases
-interR <- LIs[LIs$exclude == 0,]
-# now run analysis
-DL.CF<-RVAideMemoire::spearman.ci(interR$Day1_CF_acc_LI, interR$Day1_DL_acc_LI, nrep = 10000, conf.level = 0.95)
-DL.CF.est<-cor.test(interR$Day1_CF_acc_LI, interR$Day1_DL_acc_LI, method = "spearman")
-# run Bayes factor
-BF1 <- BayesFactor::correlationBF(y = interR$Day1_CF_acc_LI, x= interR$Day1_DL_acc_LI, iterations = 10000)
-# day 2
-# now run analysis
-DL.CF.2<-RVAideMemoire::spearman.ci(interR$Day2_CF_acc_LI, interR$Day2_DL_acc_LI, nrep = 10000, conf.level = 0.95)
-DL.CF.est.2<-cor.test(interR$Day2_CF_acc_LI, interR$Day2_DL_acc_LI, method = "spearman")
-# run Bayes factor
-BF2 <- BayesFactor::correlationBF(y = interR$Day2_CF_acc_LI, x= interR$Day2_DL_acc_LI, iterations = 10000)
-```
-
-Using data from session 1, we examined the correlation between LIs on the dichotic listening and chimeric faces task to examine whether the perceptual biases on these tasks are governed by a single laterality factor.
-A spearman’s rank order correlation indicated a extremely small, non-significant relationship, *rs*= `r round(DL.CF.est$estimate, 2)`, *95% CI* [`r round(DL.CF$conf.int[1], 2)`, `r round(DL.CF$conf.int[2], 2)`] *p*= `r round(p.adjust(DL.CF.est$p.value[1], n=2, method= "bonferroni"), 3)`.Based on this analysis, there is no evidence of a relationship between laterality indices on the dichotic listening and chimeric faces tasks.
-To infer the extent to which our data reflect a null relationship, as opposed to a Type II error, we computed a Bayes factor (Rouder, Morey, Speckman, & Province, 2012).
-We computed the Bayes factor in R using the *correlationBF()* functions from the BayesFactor package (Version 0.9.12–4.2; Morey & Rouder, 2018) with 10,000 Monte Carlo iterations.
-We assumed the default prior (*r*= 1/3) for the effect size.
-The Bayes factor for the alterative model, as opposed to the null model, was .150.
-According to Jeffreys' (1961) inference criteria, this provides moderate evidence for the null hypothesis.
-To validate these findings, we repeated these analysis with session 2 data and controlled for multiple comparisons.
-Again the correlation was small and not significantly different from zero *rs*= `r round(DL.CF.est.2$estimate, 2)`, *95% CI* [`r round(DL.CF.2$conf.int[1], 2)`, `r round(DL.CF.2$conf.int[2], 2)`] *p*= `r round(p.adjust(DL.CF.est.2$p.value[1], n= 2, method= "bonferroni"), 3)`.
-Again the Bayes Factor favoured the null hypothesis, *BF*= .138.
+### Relationship between LIs from different tasks
+
+Our final question was whether the laterality indices from the language and faces laterality tasks correlate?
+Specifically we had predicted that in right-handers the two language tasks would correlate together, but the chimeric faces would form a separate factor, but in left-handers, the language tasks might fractionate.
+
+Initial scrutiny of Spearman correlations between LI_z scores suggests little evidence for any association between the three laterality indices, as shown in Table x.
+
+```{r bigcorrmatrix}
+
+# create subset of data frame for analysis - I used LI_z measures, as they are on similar scale
+interRR <- dplyr::select(interR, 'zlat_RDT_1', 'zlat_RDT_2', 'Day1_Dich_acc_Z',
+                         'Day2_Dich_acc_Z', 'Day1_CF_acc_Z', 'Day2_CF_acc_Z','handedness')
+# rename
+interRR <- interRR %>% 
+  rename(
+    "Rhyme_1" = zlat_RDT_1,
+    "Chimeric_1" = Day1_CF_acc_Z,
+    "Dichotic_1" = Day1_Dich_acc_Z,
+     "Rhyme_2" = zlat_RDT_2,
+    "Chimeric_2" = Day2_CF_acc_Z,
+    "Dichotic_2" = Day2_Dich_acc_Z
+    )
+
+#remove cases with missing values
+interRR<-na.omit(interRR)
+# subset into right and left handers
+left_dat <- subset(interRR, handedness== "Left")
+right_dat <- subset(interRR, handedness== "Right")
+left_dat<-left_dat[,1:6]
+right_dat<-right_dat[,1:6]
+totcorr<-correlate(interRR[,1:6],method='spearman')
+Rcorr <- correlate(right_dat,method='spearman')
+Lcorr <- correlate(left_dat,method='spearman')
+
+#Change so we have one matrix with R handers above diagonal and L handers below
+Lcorr<-as.data.frame(Lcorr[,2:7])
+Rcorr <- as.data.frame(Rcorr[,2:7])
+allcorr<-Lcorr
+for (i in 1:6){
+  for (j in i:6){
+    allcorr[i,j]<-Rcorr[i,j]
+
+  }
+}
+allcorr<- round(allcorr,3)
+row.names(allcorr)<-colnames(allcorr)
+
+NL<-nrow(left_dat)
+NR<-nrow(right_dat)
+#for left_dat N = 66
+# right_dat
+
+myft3<- flextable(allcorr)
+myft3
+
+```
+
+Table x shows Spearman correlations between the LI_z scores for the different tasks in two Sessions.
+The upper diagonal shows correlations for the `r NR` right-handers, and teh lower diagonal shows values for the `r NL` left-handers.
+For right-handers, an absolute value of .16 is significant at the .05 level, and a value of .22 is significant at the .01 level.
+For left-handers, a value of .25 is significant at the .05 level and a value of .32 is significant at the .01 level.
+
+Table x suggests weak associations between LI_z values for the two language measures (rhyme decision and dichotic listening), but no association between either of these measures and LI_z on chimeric faces.
+
+In a more formal test we compared (1) a base model where the LIs are unrelated, (2) a model where the two language tasks are correlated with each other but not with the chimeric faces, (3) a model where all LIs are correlated.
+This was done separately for right- and left-handers, using a subset of SEM that does not include any latent variables or directional paths.
+In this approach, we constrain particular covariance patterns and report ‘best’ model fit according AIC weights (Wagenmakers & Farrell, 2004).
+Akaike weights are transformations of the AIC values, which can be directly interpreted as conditional probabilities for each model.
+The best fitting model to the data for right- and left-handers is selected as the final model.
+
+```{r prepareKievit,echo=FALSE, comment=FALSE, warning=FALSE}
+library(lavaan)
+library(tidyverse)
+library(MASS)
+library(qpcR)
+# created handedness data
+# now create subset of data frame for analysis
+interRR <- dplyr::select(interR, 'zlat_RDT_1', 'zlat_RDT_2', 'Day1_Dich_acc_Z',
+                         'Day2_Dich_acc_Z', 'Day1_CF_acc_Z', 'Day2_CF_acc_Z','handedness')
+# rename
+interRR <- interRR %>% 
+  rename(
+    "rhyme1" = zlat_RDT_1,
+    "faces1" = Day1_CF_acc_Z,
+    "dichotic1" = Day1_Dich_acc_Z,
+     "rhyme2" = zlat_RDT_2,
+    "faces2" = Day2_CF_acc_Z,
+    "dichotic2" = Day2_Dich_acc_Z
+    )
+
+#remove cases with missing values
+interRR<-na.omit(interRR)
+# subset into right and left handers
+left_dat <- subset(interRR, handedness== "Left")
+right_dat <- subset(interRR, handedness== "Right")
+
+```
+
+```{r AICmodels,echo=FALSE, comment=FALSE, warning=FALSE}
+#NB in pre-reg we just had time 1 data, but here we use time 1 and 2
+
+ #Model 1 all independent
+    model1<-"
+    rhyme1~~rhyme2
+    rhyme1~~0*dichotic1
+    rhyme1~~0*dichotic2
+     rhyme1~~0*faces1
+    rhyme1~~0*faces2
+    rhyme2~~0*dichotic1
+    rhyme2~~0*dichotic2
+     rhyme2~~0*faces1
+    rhyme2~~0*faces2
+    dichotic1~~dichotic2
+    dichotic1~~0*faces1
+    dichotic1~~0*faces2
+     dichotic2~~0*faces1
+    dichotic2~~0*faces2
+    faces1~~faces2
+    "
+#Model 2: language tasks are correlated
+    model2<-"
+  rhyme1~~rhyme2
+    rhyme1~~dichotic1
+    rhyme1~~dichotic2
+     rhyme1~~0*faces1
+    rhyme1~~0*faces2
+    rhyme2~~dichotic1
+    rhyme2~~dichotic2
+     rhyme2~~0*faces1
+    rhyme2~~0*faces2
+    dichotic1~~dichotic2
+    dichotic1~~0*faces1
+    dichotic1~~0*faces2
+     dichotic2~~0*faces1
+    dichotic2~~0*faces2
+    faces1~~faces2
+    "
+#Model 3:all correlated
+    model3<-"
+    rhyme1~~rhyme2
+    rhyme1~~dichotic1
+    rhyme1~~dichotic2
+     rhyme1~~faces1
+    rhyme1~~faces2
+    rhyme2~~dichotic1
+    rhyme2~~dichotic2
+     rhyme2~~faces1
+    rhyme2~~faces2
+    dichotic1~~dichotic2
+    dichotic1~~faces1
+    dichotic1~~faces2
+     dichotic2~~faces1
+    dichotic2~~faces2
+    faces1~~faces2
+    "
+#Make a little dataframe for AIC weights
+AICdf <- data.frame(matrix(NA,nrow=2,ncol=4))
+AICdf[,1]<-c('Right-handers','Left-handers')
+colnames(AICdf)<-c('Handedness','Independent','Language factor','Laterality factor')
+for (myhand in 1:2) {
+  mydat<-right_dat
+  if(myhand==2) {mydat<-left_dat}
+
+ #Using data from this handedness type, test all 3 models
+fit1A <- cfa(model1, data=mydat)
+fit2A <- cfa(model2, data=mydat)
+fit3A <- cfa(model3, data=mydat)
+    aic_vector1 <- c(fitMeasures(fit1A, "aic"),fitMeasures(fit2A, "aic"),fitMeasures(fit3A, "aic"))
+
+    #Confirms that for each model, lowest AIC obtained when the simulated data do match the model
+    # 
+    # The AIC weights make massive difference for preferred model!
+    AICweights<-akaike.weights(aic_vector1)$weights
+    # 
+ AICdf[myhand,2:4]<-round(AICweights,3)
+ }
+flextable(AICdf)
+```
+
+As shown in Table x, the Akaike weights favoured the base model for both left-handers and right-handers. This reflected the lack of cross-task correlation for LI_z measures.
 
@@
 ```{r fingerEHI,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
-spearCI <- RVAideMemoire::spearman.ci(LIs$Day1_finger_LI, -LIs$index_EHI, nrep = 10000, conf.level = 0.95)
+spearCI <- RVAideMemoire::spearman.ci(LIs$Day1_finger_LI, LIs$index_EHI, nrep = 10000, conf.level = 0.95)
+spearCI2 <- RVAideMemoire::spearman.ci(LIs$Day2_finger_LI, LIs$index_EHI, nrep = 10000, conf.level = 0.95)
 # spearman's test
-speartest <- cor.test(LIs$Day1_finger_LI, -LIs$index_EHI, method = "spearman", alternative = "greater")
-```
-
-To examine the relationship between the finger tapping task with an existing measures of handedness, we used a spearman's rank correlation to obtain an estimate of the relationship between LIs produced on the finger tapping task and the Edinburgh Handedness Inventory using session 1 data.
-The relationship between the tasks was moderate, *rs*= `r round(speartest$estimate, 2)`, *95% CI* [`r round(spearCI$conf.int[1], 2)`, `r round(spearCI$conf.int[2], 2)`],  *p*< 0.001.
-Thus it indicates that the task likely represents captures some variance in hand preference.
-It is, however, also likely that the task goes beyond this and captures motor planning and coordination.
-We say more about this in the discussion.
+speartest <- cor.test(LIs$Day1_finger_LI, LIs$index_EHI, method = "spearman", alternative = "greater")
+speartest2 <- cor.test(LIs$Day2_finger_LI, LIs$index_EHI, method = "spearman", alternative = "greater")
+```
+
+To examine the relationship between the finger tapping task and a conventional hand preference battery, we computed Spearman's rank correlation between LIs produced on the finger tapping task and the Edinburgh Handedness Inventory.
+For session 1 data, the relationship between the tasks was moderate, *rs*= `r round(speartest$estimate, 2)`, *95% CI* [`r round(spearCI$conf.int[1], 2)`, `r round(spearCI$conf.int[2], 2)`],  *p*< 0.001.
+For session 2, the correlation was *rs*= `r round(speartest2$estimate, 2)`, *95% CI* [`r round(spearCI2$conf.int[1], 2)`, `r round(spearCI$conf.int[2], 2)`],  *p*< 0.001.
+This suggests that the influence of hand preference on the tapping task may increase with practice.
 
@@
 
-Our planned analysis of the association between handedness and (a)typical lateralisation was based solely on using the Edinburgh Handedness Inventory Scores.
-We additionally examined how different strategies of coding handedness would influence the strength of this association on the dichotic listening and chimeric faces tasks using session 1 and session 2 data.
-We assessed these associations in the following ways: (1) used the Edinburgh Handedness Inventory to code left, ambidextrous, and right handedness, (2) using self-reported handedness, and (3) using performance on the finger tapping task to code handedness.
-To examine the influence of the three handedness groups (left, ambidextrous, right), we defined handedness as: scores of +40 or above denoting right-handedness, −40 or below denoting left handedness, and scores in between denoting ambidexterity (follwoing Whitehouse & Bishop, 2009).
-The frequencies of (a)typicality are shown for left-, ambidextrous, and right-handers are shown in Table 3.
-A chi-squared test of association indicated that there was a significant association between handedness and the dichotic listening, chi squared = `r round(EHI_DL.chi$statistic,2)`, d.f.
-= 4, *p* = `r round(p.adjust(EHI_DL.chi$p.value, method= "bonferroni", n=2),3)`; and chimeric faces, chi squared = `r round(EHI_CF.chi$statistic,2)`, d.f.
-= 4, *p* = `r round(p.adjust(EHI_CF.chi$p.value, method= "bonferroni", n=2),3)`.
-
-Table 3. Frequencies of laterality typicality for right-, ambidextrous, and left-handers across the Dichotic listening and Chimeric faces tasks.
-```{r table 3, echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
-EHI_hand_Tabs
-```
-Note. TypLat: lateralised in the expected direction, Nonlat: no lateralisation, and AtypLat: show lateralisation in the opposite direction.
-
-From Table 4, there is evidence of increased atypicality for left-handers on the dichotic listening when using self-reported handedness.
-The association appears even stronger when considering the chimeric faces task.
-Chi-squared tests of association confirmed the association between handedness and atypicality as statistically significant on the dichotic listening, chi squared = `r round(self_DL.chi$statistic,2)`, d.f.
-= 2, *p* = `r round(p.adjust(self_DL.chi$p.value, method= "bonferroni", n=2),3)`; and chimeric faces, chi squared = `r round(self_CF.chi$statistic,2)`, d.f.
-= 1, *p* = `r round(p.adjust(self_CF.chi$p.value, method= "bonferroni", n=2),3)`.
-
-Table 4. Frequencies of laterality typicality for self-reported right- and left-handers across the Dichotic listening and Chimeric faces tasks.
-```{r table 4, echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
-self_hand_Tabs
-```
-Note. TypLat: lateralised in the expected direction, Nonlat: no lateralisation, and AtypLat: show lateralisation in the opposite direction.
-
-We then assessed the association between handedness, defined using the finger tapping task, and (a)typical lateralisation.
-Here a negative value was coded as a preference for right handedness and a positive value as left handedness.
-From Table 5, the association between left-handedness and atypical laterality on the dichotic listening task was not significant, chi squared = `r round(finger_DL.chi$statistic,2)`, d.f.
-= 2, *p* = `r round(p.adjust(finger_DL.chi$p.value, method= "bonferroni", n=2),3)`.
-There was, however, an association between finger tapping defined handedness and atypicality for chimeric faces, chi squared = `r round(finger_CF.chi$statistic,2)`, d.f.
-= 2, *p* = `r round(p.adjust(finger_CF.chi$p.value, method= "bonferroni", n=2),3)`.
-Taken together, these analyses demonstrate a clear association between handedness and atypical lateralisation on both tasks.
-The chi-squared coefficient was largest when using the Edinburgh Handedness Inventory, indicating that this measure is likely the most suited to identifying those who are likely to show (a)typical lateralisation.
-
-Table 5. Frequencies of laterality typicality for finger tapping defined right- and left-handers across the Dichotic listening and Chimeric faces tasks.
-```{r table 5, echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
-finger_hand_Tabs
-```
-Note. TypLat: lateralised in the expected direction, Nonlat: no lateralisation, and AtypLat: show lateralisation in the opposite direction.
-
-We additionally examined the strength of the relationship of LIs on the dichotic listening and chimeric faces tasks with LIs calculated using the Edinburgh Handedness Inventory and the finger tapping task.
-A series of Spearman’s rank correlations indicated that the relationship between LIs on each task were largely small (see Table 3).
-A series on z-tests were then computed using absolute correlation coefficients to examine whether the Edinburgh Handedness Inventory or finger Tapping were better correlated with LIs on the dichotic listening and chimeric faces.
-We controlled for multiple comparisons using a Bonferroni adjustment.
-For session 1 data, the Edinburgh handedness Inventory was more strongly correlated with the chimeric faces, *z*= `r round(as.numeric(as.htest(cocor.result2)$pearson1898[1]), 2)`, *p*= `r round(p.adjust(as.htest(cocor.result2)$pearson1898[3], n= 2, method = "bonferroni"),3)`, but not the dichotic listening, *z*= `r round(as.numeric(as.htest(cocor.result)$pearson1898[1]), 2)`, *p*= `r round(p.adjust(as.htest(cocor.result)$pearson1898[3], n= 2, method = "bonferroni"),3)`, task.
-For session 2 data, an identical pattern of results emerge whereby the Edinburgh handedness Inventory was more strongly correlated with the chimeric faces, *z*= `r round(as.numeric(as.htest(cocor.result4)$pearson1898[1]), 2)`, *p*= `r round(p.adjust(as.htest(cocor.result4)$pearson1898[3], n= 2, method = "bonferroni"),3)`, but not the dichotic listening, the dichotic listening, *z*= `r round(as.numeric(as.htest(cocor.result3)$pearson1898[1]), 2)`, *p*= `r round(p.adjust(as.htest(cocor.result3)$pearson1898[3], n= 2, method = "bonferroni"),3)`, task.
-
-Table 6.
-Correlation coefficients between LIs on the dichotic listening (DL) and chimeric faces (CF) tasks and the Edinburgh Handedness Inventory and finger tapping task using session 1 and session 2 data.
-```{r table 6, echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
-corrTabs
-```
-
-#### Vocabulary knowledge, footedness, sightedness and laterality
+<!---Adam I removed the exploratory analyses of different definitions of handedness, as I was worried the paper was getting too long and indigestible.
+These could be put in a supplementary section, but I think they are too much at this point--->
+
+#### Vocabulary knowledge and laterality
 
@@
 
-While the role of the LexTALE in the current research was largely for the purpose of data exclusion, we had planned to explore correlations between LexTALE score and the laterality index on the Rhyme Decision task.
-However, as the rhyme task was not overly reliable, we chose to examine the relationship between the LexTALE and the dichotic listening task instead.
+The LexTALE vocabulary measure had been included largely for the purpose of data exclusion, but we had planned to explore correlations between LexTALE score and the laterality index on the Rhyme Decision task.
+However, as the rhyme task was less reliable than other measures, and showed only small asymmetries, we chose to examine the relationship between the LexTALE and the other verbal laterality measure, dichotic listening, instead.
 From session 1 data, after correcting for multiple comparisons, the correlation between the LIs on the dichotic listening and LexTALE score was small and non-significant, *rs*= `r round(DL.lex.est$estimate, 2)`, *95% CI* [`r round(DL.lex$conf.int[1], 2)`, `r round(DL.lex$conf.int[2], 2)`] *p*= `r round(p.adjust(DL.lex.est$p.value[1], n= 2, method= "bonferroni"), 3)`.
@@
 
-We then examined whether footedness was related to the laterality indices on both the dichotic listening and chimeric faces tasks.
-Using session 1 data, LIs on the dichotic listening task did not differ between left and right footers, *t*(275.89)= .07, *95% CI* [-3.29, 11.15], *p*= `r p.adjust(0.284, n= 2, method= "bonferroni")`, nor did LIs on the chimeric faces, *t*(219.58)= -1.91, *95% CI* [-12.78, .20], *p*= `r p.adjust(0.058, n= 2, method= "bonferroni")`.
-Analysis of session 2 data yielded a similar pattern of findings.
-LIs on the dichotic listening, *t*(243.09)= 1.87, *95% CI* [-.39, 14.63], *p*= `r p.adjust(0.063, n= 2, method= "bonferroni")`, and the chimeric faces, *t*(197.01)= -2.03, *p*= `r p.adjust(0.044, n= 2, method= "bonferroni")`, tasks did not differ for right and left footers.
-We then completed the same analyses but using sightedness in the place of footedness.
-As participants self-reported sightedness was always the same on the Miles and Porta test, we analysed data using only the Porta test.
-Using session 1 data, LIs on the dichotic listening task did not differ with sightedness, *t*(176.49)= .02, *95% CI* [-7.49, 7.67], *p*= `r p.adjust(0.982, n= 2, method= "bonferroni")`, yet chimeric faces LIs were significantly less lateralised to the right hemisphere in those with left eye dominances, *t*(-3.06)= -3.06, *95% CI* [-18.96, -4.07], *p*= `r p.adjust(0.003, n= 2, method= "bonferroni")`.
-Analysis of session 2 data yielded a similar pattern of findings for the dichotic listening task, *t*(144.51)= .75, *95% CI* [-17.99, -21.11], *p*= `r p.adjust(0.455, n= 2, method= "bonferroni")`.
-However, LIs on the chimeric faces tasks did not differ between left and righ eye dominant groups here, *t*(124.25)= -1.89, *95% CI* [-17.48, .39], *p*= `r p.adjust(0.061, n= 2, method= "bonferroni")`.
+<!--- footedness etc also removed here ; could be some interest in looking at people with unusual laterality patterns on eye/foot/hand, but here it is too much--->
 
@@
 
-Each behavioural task yielded a hemispheric advantage in the expected direction.
-However, the extent to which they were reliable differed substantially between tasks.
-The rhyme decision task yielded the poorest test-retest reliability.
+Each behavioural task yielded a hemispheric advantage in the expected direction, but the size of asymmetries and the test-retest reliability of the different tasks varied.
+A relatively small proportion of individuals showed significant asymmetry on the rhyme decision task.
+This task also yielded the poorest test-retest reliability, just falling short of our target minimum cutoff of r = .65.
 Compared to early studies (e.g.
 Krach et al., 2006) the rhyme decision task used here was far more reliable but, compared to other visual half-field tasks, it was far from optimal.
-van der Haegen and Brysbaert (2018) have reported that under lab conditions that visual half-field paradigms can yield test-retest reliabilities of *r*= .77 when using pictures and *r*= .83 when using words.
-While the two studies both included a sufficiently large number of trials, there are several potential reasons why the rhyme decision task failed to achieve an acceptable level of test-retest reliability in the current study.
-One key difference between van der Haegen and Brysbaert’s and our study was that their language based visual half-field tasks involved naming.
-Speech production is highly lateralised (Kosslyn, 1987) and requiring speech production will no doubt yield a hemispheric advantage.
-As the rhyme task was run online, it would be extremely difficult to use the onset of speech to trigger the end/start of a trial let alone retain and code these response manually.
-Therefore, we opted to use a rhyme component that should theoretically engage subvocal processes associated with phonological processing.
-Yet, this did not yield such reliable results.
-Therefore it may be of use for future implementations to involve some active production component.
-
-While the rhyme decision task was not so reliable, the dichotic listening task yielded surprising results.
-Not only did participants show a right-ear advantage at the population level, the task exceeded our pre-registered reliability cut-off of *r*= .65.
-This was, perhaps, not unexpected though as the paradigm has been implemented via the iDichotic application (Bless et al., 2013).
-What is more astonishing is the fact test-retest correlation of the dichotic listening task is identical to that using the iDichotic application (both test-retest coefficients= .78) despite the current study using a non-dedicated platform.
-Furthermore, the reliability in the current study exceeds that of the dichotic listening task when run under lab conditions (Bless et al., 2013).
-However, this reliability is somewhat less than the estimates of reliability produced when completing a modified version of the task under lab conditions (Westerhausen & Samuelsen, 2020).
-What is particularly interesting is the observation that with as few as 60 trials, the current implementation of the dichotic listening task yields a test-retest correlation that than *r*= .65, meaning that those who wish to use this paradigm in the future could use a smaller number of trials and add additional tasks on that are important to their ongoing research questions.
-Of course, it would be necessary to screen for adequate and appropriate headphone as we did.
-
-Like the dichotic listening task, the chimeric faces task yielded good test-retest reliability.
-While previous endeavours to assess split-half reliability of the task have yielded strong correlations (i.e.
-Levy et al., 1983), we report a strong test-retest correlation, What is particularly novel about this finding is that we were able to report this for online behavioural research methods.
-Furthermore, the reliability of this task can be astonishingly high with very few trials.
-Our analysis indicated that only 25 trials would be necessary to achieve a test-retest reliability of *r*= .65.
-We believe that the reason for the success of the chimeric faces task online is the task’s simplistic nature.
-As implemented, it requires participants to report the emotion they saw and not have to make judgements which involve comparing representations of emotions as the two faces version might.
+Van der Haegen and Brysbaert (2018) have reported that, under lab conditions, visual half-field paradigms can yield test-retest reliabilities of *r*= .77 when using pictures and *r*= .83 when using words.
+Our study was well-powered, so the relatively poor task validity and reliability are likely to be due to methodological factors.
+One key difference between van der Haegen and Brysbaert’s and our study was that their language based visual half-field tasks involved overt naming.
+Speech production is highly lateralised (Kosslyn, 1987) and requiring speech production should yield a hemispheric advantage.
+Our task was designed to require covert speech, but it is possible that an active speech production component might reveal stronger lateralisation.
+This would, however, be logistically difficult to implement online.
+The other main limitation of online testing is that the VHF method depends critically on presenting lateralised visual stimuli that project to one hemisphere.
+Although we were able to calibrate screen size at the start of the session, and to request participants to sit a certain distance from the screen, it is plausible this was not sufficient to achieve adequately lateralised presentation.
+While we can format images to be consistent across displays, we cannot control participants’ behaviour.
+Outside of the lab, participants can freely move their head which will influence the size of images on the retina, and it is difficult to ensure images are shown in foveal or parafoveal vision.
+One avenue that might be worth exploring is the use of images presented in peripheral vision, as it is unlikely that head movements would result in the image shifting between distinct regions on the retina.
+This increased difficultly in processing of information in this region may also result in stronger visual field advantages (c.f.
+Hunter and Brysbaert, 2008).
+
+In contrast to rhyme decision, the other verbal task, dichotic listening, showed excellent validity and reliability.
+Participants showed a clear right-ear advantage at the population level, and the task exceeded our pre-registered reliability cut-off of *r*= .65.
+Indeed, the test-retest correlation of the dichotic listening task, *r* = .78, was identical to that reported by Bless et al.
+(2013) using the iDichotic application, and was higher than that reported for the same dichotic listening task when run under lab conditions (Bless et al., 2013).
+We further showed that with as few as 60 trials, the current implementation of the dichotic listening task yielded a test-retest correlation that than *r*= .65, suggesting that those who wish to use this paradigm in the future could use a smaller number of trials and add additional tasks on that are important to their ongoing research questions.
+It would be of interest to extend this line of work to use the  modified version of dichotic listening described by Westerhausen and Samuelsen (2020), which had even higher reliability.
+Of course, it is important to screen for adequate and appropriate headphones when devising online measures.
+
+Like the dichotic listening task, the chimeric faces task showed impressive levels of both validity and reliability, this time with a strong right hemisphere bias at the population level.
+The task has previously be shown to have high split-half reliability (Levy et al., 1983), we show that it can meet the higher standard of achieving a high test-retest correlation.
+Our subsequent analysis showed that acceptable levels of reliability can be achieved with this task using very few trials: test-retest reliability of *r*= .65 could be found with 25 trials.
+It seems likely that one reason for the success of the chimeric faces task online is the task’s simplistic nature.
+From the perspective of the participant, it is a simple task of reporting the emotion expressed in a single  face, and does not require any explicit comparison.
 
 We also introduced a novel finger tapping task, which we designed to quantify handedness behaviourally.
-Surprisingly, this simple task produced reliable results.
-Not only was there a right hand advantage at the population level, the LIs produced on this task were reliable.
-Furthermore, our exploratory analyses indicated that this correlated reasonably well with the Edinburgh handedness inventory.
-Yet the Edinburgh Handedness Inventory did seem better equipped to categorise atypical lateralisation.
-This suggests that while the finger tapping task does capture some aspect of hand preference it does also capture something beyond this.
-It is likely that  the task additionally measures the planning and execution of motor coordination.
-However, this remains speculative at current and further work will be needed to clarify the aspects that this task measures beyond simple hand preference and to examine the extent to which LIs resemble that from tasks such as the pegboard task (e.g.
-Hodgson, Hirst, & Hudson, 2016).
+This too produced a right hand advantage at the population level, and test-retest reliability of the LIs was good, even though the asymmetries were small in absolute terms.
+Furthermore, our exploratory analysis indicated that LIs from this task correlated reasonably well with the Edinburgh handedness inventory.
+Previous studies have reported correlations between hand preference and measures of relative hand skill, but different hand skill measures typically are weakly intercorrelated (Bishop, 1990), and it seems that asymmetries may vary depending on particular aspects of planning and execution of motor coordination that are involved.
+It would be of interest to see how far LIs resemble that from tasks such as the pegboard task (e.g.
+Hodgson, Hirst, & Hudson, 2016), which is widely used in neuropsychological assessment.
 Future work should also examine how other factors such as complexity of key combinations and reach through the midline influence the task.
 
-In addition to assessing lateralisation of the task using traditional LIs, we also examined the potential to use individual *z*/*t*-scores to quantify lateralisation on each task.
-While the *t*-test for the rhyme task does not seem warranted for further use, the *z*-scores for the remaining tasks closely resembled that of the traditional LI.
-This method of calculating Lis is perhaps more useful for those interested in looking atypical lateralisation.
-Rather than selecting those who fall either side of an arbitrary cut-off, *z*-scores allow researchers to assess whether participants are consistently and statistically lateralised in either a typical or atypical manner.
-Let us consider an example to emphasise their importance.
-A researcher plans to examine the influence of atypical language lateralisation on education outcomes and would like to compare those who show strong typicality and those who show strong atypicality.
-Before even meeting the participants for testing, the battery described in the current study could be administered and then those who have *z*-scores > 1.96 or <1.96 could be invited into the lab as they appeared to be significantly lateralised on a behavioural task.
-This has the potential to save time and effort screening participants in a lab setting only to find that they do not meet study inclusion criteria.
+In addition to assessing lateralisation of the task using traditional LIs, we also examined the potential to use individual *z*-scores to quantify lateralisation on each task.
+This method of calculating LIs is perhaps more useful for those interested in looking atypical lateralisation, or for cases where there is a need to compare laterality from different tasks on a common scale.
+Rather than selecting those who fall either side of an arbitrary cut-off, the LI_z score allows researchers to assess whether participants are consistently and statistically lateralised in either a typical or atypical manner.
+The availability of an online measure means that a researcher who is interested in studying the correlates of atypical language lateralisation could assess participants with our battery to identify those with LI_z scores > 1.96 or <1.96 before inviting them into the lab for more detailed in-person testing.
+This has the potential to save time and effort by screening participants to identify a sample that is enriched for rare cases of reversed cerebral lateralisation.
 
@@
 (2014) reported a relationship between strength of lateralisation and hand preference., which was largely driven by the existence of a group with strong left hand preference and strong right hemisphere dominance.
-They further claim that this right hemisphere dominance is almost exclusively present in left-handers.
-We, however, did observe a portion of right-handers with right language dominance on the dichotic listening task—regardless of how we define handedness.
-While issues such as sample size and bias between studies likely represent this, it does further question the whether there is a common mechanism that determines handedness and lateralisation.
-
-A common mechanism that codetermines lateralisation on an array of tasks appears a somewhat hotly debated topic in the laterality literature.
-Several hypotheses have argued that LIs on one task should correlate with another (Behrmann & Plaut, 2015; Cook, 1984; Hellige, 1983; Kosslyn, 1987).
-However, several studies have report no such correlation, and if they do these are often small (Bradzakova-Trajkov et al., 2010; Bryden et al.,1983; Rosch et al., 2012; Whitehouse & Bishop, 2009; van der Haegen & Brysbaert, 2018).
-Data from the current work supports a view in which laterality one each task is independent.
+They further claimed that this right hemisphere dominance is almost exclusively present in left-handers.
+Nevertheless, although strong left-handers were over-represented in this category, we did also observe cases of right-handers with right language dominance on the dichotic listening task.
+This raises the interesting question of whether right language dominance on dichotic listening is equivalent to right language dominance on more direct measures of brain lateralisation, such as fMRI.
+
+A common mechanism that codetermines lateralisation on a range of skills, extending beyond handedness and language, has been a hotly debated topic in the laterality literature.
+Several hypotheses postulate complementary lateralisation for left-hemisphere and right-hemisphere brain asymmetries, predicting that LIs on verbal and nonverbal tasks should be related (Behrmann & Plaut, 2015; Cook, 1984; Hellige, 1983; Kosslyn, 1987).
+On this view, a person with atypical right-hemisphere language representation should have atypical left-hemisphere representation for nonverbal domains such as face processing.
+However, several studies have report no such association, or only a weak relationship (Bradzakova-Trajkov et al., 2010; Bryden et al.,1983; Rosch et al., 2012; Whitehouse & Bishop, 2009; van der Haegen & Brysbaert, 2018).
+Data from the current work supports a view in which laterality on each task is independent.
 Not only was our study well powered to examine this relationship, it included a large sample of left-handers who show increased variability in their LIs.
-Specifically, we reported that the correlation between LIs on the dichotic listening and chimeric faces was extremely close to zero and Bayes Factor analysis ruled out the possibility that this was due to a lack of statistical power.
-Thus, we argue that lateralisation is driven by a series of independent biases, which, according to this view, specifies that there is a population bias for left hemisphere language dominance.
-By contrast, there is a population bias for right hemisphere dominance for the processing of faces.
-However, these biases reflect probabilities elating to a combination of multiple independent causal sources, which may be genetic, biological or environmental in origin.
-
-In a series of exploratory analyses we also examined whether sightedness or footedness related to the strength of laterality indices.
-The data were rather consistent in showing that these measures were not predictive of differences in laterality indices in those who report being right- or left-footed/eye dominant.
-Again, this echoes the finding that lateralisation on one function does not determine that on another.
-
-Before moving on to our concluding remarks, it is important to consider the future of online research methods to examine lateralisation.
-On the whole, the results pave a way for a promising future of online data collection to answer several questions with regards to cerebral lateralisation.
-It seems that when we rely on count data (as did the dichotic listening, chimeric faces, and finger tapping tasks) we can obtain reliable results.
-The success of the dichotic listening and chimeric faces task is likely to stem from the use of bilateral stimuli presentation, which results in the two hemispheres competing.
-The story, however, for tasks depending on reaction times for the calculation of LIs seems slightly less optimistic.
-This is likely to result from the increased noise associated with different experimental setups across participants when using typical paradigms.
-Furthermore, while we can format images to be consistent across displays, we cannot control participants’ behaviour.
-Of participant interest to laterality studies is that we cannot control viewing distance across participants, which has several implications for visual half-field paradigms.
-Outside of the lab, participants can freely move their head which will result in the size of images changing on the retina.
-Furthermore,  thing makes it incredibly difficult to ensure images are shown in foveal or parafoveal vision.
-One avenue that might be worth exploring is the extent to which participants can judge images presented in peripheral vision as even with head movements it is unlikely that the change in movements will result in the image shifting between distinct regions on the retina.
-This increased difficultly in processing of information in this region may also result in stronger visual field advantages (c.f.
-Hunter and Brysbaert, 2008).
+Specifically, we found no association between the LI_z indices for rhyme decision, dichotic listening and chimeric faces.
+The lack of association between the two language tasks, rhyme decision and dichotic listening, must be interpreted cautiously, given that the rhyme decision task was unsatisfactory in several respects: it had relatively low test-retest reliability, and only weak asymmetry in the sample as a whole.
+It is possible that, as argued by Bishop (2013), there is fractionation between language skills: the dichotic listening task is a purely perceptual task, whereas the rhyme decision task involves implicit speech production.
+The current result suggests there could be interest in replicating this comparison of the two tasks using a visual-half-field language task that has better reliability and validity.
+The lack of association between dichotic listening and chimeric faces is, however, much more convincing, given that both tasks show strong lateralisation at the population level and high test-retest reliablity.
+This supports the view that lateralisation is determined by a series of independent biases: a left-hemisphere population bias for language and a right-hemisphere bias for facial expressions of emotion.
 
@@
 We find evidence of increased atypical lateralisation in left- relative to right-handers.
-However, we find no indication that the laterality indices on one task are dependent on another and argue that the causes of lateralisation on one task should be seen as stemming from multiple independent sources.
+However, we find no indication that the laterality indices on one task are dependent on another and we conclude that the causes of lateralisation in different domains stem from multiple independent sources.
 
@@
 We would also like to thank Professor Kenneth Hugdahl (University of Bergen, Norway) for providing stimuli for the dichotic listening task.
-Portions of this work were presented at the Online Meeting of the Expeirmental Psychology Society in July 2020.
-
-\newpage
-## Supplemental Material
-
-We had originally intended to examine the relationship We had originally intended to examine the relationship between LIs on the Rhyme Decision, Dichotic Listening, and Chimeric faces by modelling covariances.
-This would have allowed us to compare (1) a model where the two language tasks are correlated with each other but not with the chimeric faces, (2) a model where all LIs are correlated, and (3) a base model where the LIs are unrelated.
-This would have been done separately for right- and left-handers.
-However, due to the Rhyme Decision task having poor reliability, we decided that this analysis would not be appropriate to base our conclusions on.
-Instead, we report them as a supplemental material.
-The models were fit using a subset of SEM that does not include any latent variables or directional paths.
-It only looks to constrain particular covariance patterns and report ‘best’ model according AIC weights.
-Model fit was evaluated using AIC model selection using Akaike weights (Wagenmakers & Farrell, 2004).
-Akaike weights are transformations of the AIC values, which can be directly interpreted as conditional probabilities for each model.
-The best fitting model to the data for right- and left-handers will be selected as the final model.
-
-```{r Roger_mod_left,echo=FALSE, comment=FALSE, warning=FALSE}
-library(lavaan)
-library(tidyverse)
-library(MASS)
-library(qpcR)
-# created handedness data
-# now create subset of data frame for analysis
-interRR <- dplyr::select(interR, 'Day1.RDT_RT_LI', 'Day1_CF_acc_LI', 'Day1_DL_acc_LI', "handedness")
-# rename
-interRR <- interRR %>% 
-  rename(
-    "rhyme1" = Day1.RDT_RT_LI,
-    "chimeric1" = Day1_CF_acc_LI,
-    "dichotic1" = Day1_DL_acc_LI)
-# subset into right and left handers
-left_dat <- subset(interRR, handedness== "Left")
-right_dat <- subset(interRR, handedness== "Right")
-# set up model 1
-model1<-"
-dichotic1~~0*rhyme1
-dichotic1~~0*chimeric1 
-rhyme1~~0*chimeric1 
-"
-# set up model 2
-model2<-"
-dichotic1~~rhyme1 
-dichotic1~~0*chimeric1
-rhyme1~~0*chimeric1 
-"
-# set up model 3
-model3<-"
-dichotic1~~rhyme1 
-dichotic1~~chimeric1 
-rhyme1~~chimeric1
-"
-# now fit model
-fit1A <- cfa(model1, data=left_dat)
-fit2A <- cfa(model2, data=left_dat)
-fit3A <- cfa(model3, data=left_dat)
-# get the aic
-aic_vector1 <- c(fitMeasures(fit1A, "aic"),fitMeasures(fit2A, "aic"),fitMeasures(fit3A, "aic"))
-#summary(fit1A, fit.measures=TRUE)
-#summary(fit2A, fit.measures=TRUE)
-#summary(fit3A, fit.measures=TRUE)
-# get weights
-AICweights1<-akaike.weights(aic_vector1)$weights
-# populat df
-IC_objectweights1<-data.frame(Model=factor(rep(c('Model 1','Model 2','Model 3'))), IC = factor(rep('AIC',each=3)),values=AICweights1)
-# now print weights
-left.weights <- ggplot(IC_objectweights1,aes(Model,values))+
-  geom_bar(stat='identity',position="dodge")+
-  ylab('Akaike weights') + xlab("") + theme_classic() + ylim(0, .7) + ggtitle("A") + 
-  theme(plot.title = element_text(hjust = 0.5))
-# NOW DO RIGHT HANDERS
-#We now repeat this for right handers. 
-fit1B <- cfa(model1, data=right_dat)
-fit2B <- cfa(model2, data=right_dat)
-fit3B <- cfa(model3, data=right_dat)
-aic_vector2 <- c(fitMeasures(fit1B, "aic"),fitMeasures(fit2B, "aic"),fitMeasures(fit3B, "aic"))
-#summary(fit1B, fit.measures=TRUE)
-#summary(fit2B, fit.measures=TRUE)
-#summary(fit3B, fit.measures=TRUE)
-AICweights2<-akaike.weights(aic_vector2)$weights
-IC_objectweights2<-data.frame(Model=factor(rep(c('Model 1','Model 2','Model 3'))), IC = factor(rep('AIC',each=3)),values=AICweights2)
-right.weights <- ggplot(IC_objectweights2,aes(Model,values)) +
-  geom_bar(stat='identity',position="dodge")+
-  ylab('Akaike weights') + xlab("") + theme_classic() + ylim(0,.7) + ggtitle("B") + 
-  theme(plot.title = element_text(hjust = 0.5))
-model.comp <- ggpubr::ggarrange(left.weights, right.weights, ncol = 2)
-# save
-ggsave(
-  "model.comp.png",
-  plot = model.comp,
-  width = 6, height = 3,
-  dpi = 300
-)
-```
-
-![Figure S1. Plot of Akaike weights for each model for (A) left-handers and (B) right-handers.](model.comp.png)
-
-As shown in Figure S1, the Akaike weights favoured a different model for left- and right-handers.
-For left-handers model 2, where all LIs were correlated, was preferred.
-For right-handers model 1, were language tasks were correlated with each other but not the chimeric faces was preferred.
-The pattern of results for left handers was surprising given that in the main article we reported no correlation between the dichotic listening and chimeric faces tasks.
-Furthermore, for left-handers, the fact that the preferred model is one where language tasks are related seems to contrast previous work reporting that language tasks fractionate in left-handers (Woodhead et al., 2019).
-This is likely to reflect that the poor reliability of the rhyme decision task.
-For right-handers we see that a model where language tasks are related.
-This is as predicted but it is of course likely that this estimate also represents noise from the rhyme decision task.
+Portions of this work were presented at the Online Meeting of the Experimental Psychology Society in July 2020.
 
@@
 ## Results
-## Question 4
-Our final question was whether the laterality indices from the language and faces laterality tasks correlate?
-Specifically we had predicted that in right-handers the two language tasks would correlate together, but the chimeric faces would form a separate factor, but in left-handers, the language tasks might fractionate.
-This analysis was handicapped by the fact that one of the language indices, the rhyme judgement task, lacked both reliability and validity.
-Therefore we present the planned analysis in supplementary material, just for completeness.
-Here we consider the intercorrelations between the reliable tasks, including finger tapping instead of rhyme judgement.
-This is motivated by the prior observations that on both dichotic listening and chimeric faces, there was some association between handedness and lateralised task performance.
-We had planned to use a structural equation modelling framework to compare the covariances of the laterality statistics between the two language tasks and chimeric faces to test between three models, separately for left- and right-handers.
-However, in neither handedness group were there any meaningful correlations between laterality indices.
+#
 ```{r prep_dat,echo=FALSE}

### 2020-07-20 online_laterality_reliability.Rmd
had to change 'select' commands to 'dplyr::select' because of conflict. Also just inverted the faces data before the AIC.

@@
 
+<!--- updated 20/7/20 to change 'select' to 'dplyr::select' to avoid conflicts-->
 Adam J. Parker
@@
 myname<-'Rhyme judge RT'
-tempLR<-select(LIs[LIs$exRDT==0,],RDTlog1.Lraw,RDTlog1.Rraw,handpch)
+tempbit<-LIs[LIs$exRDT==0,]
+tempLR<-dplyr::select(tempbit,RDTlog1.Lraw,RDTlog1.Rraw,handpch)
 #convert back to raw time
@@
 
-tempLR<-select(LIs[LIs$exRDT==0,],RDTlog2.Lraw,RDTlog2.Rraw,handpch)
+tempLR<-dplyr::select(LIs[LIs$exRDT==0,],RDTlog2.Lraw,RDTlog2.Rraw,handpch)
 tempLR[,1:2]<-exp(tempLR[,1:2]) #convert from log back to original scale for plot
@@
 myname<-'Dichotic accuracy'
-tempLR<-select(LIs[LIs$exclude<2,],DL1.Lraw,DL1.Rraw,handpch)
+tempLR<-dplyr::select(LIs[LIs$exclude<2,],DL1.Lraw,DL1.Rraw,handpch)
 session<-1
@@
 myg3<-makerawplot(tempLR,session,myname,mylim)
-tempLR<-select(LIs[LIs$exclude<2,],DL2.Lraw,DL2.Rraw,handpch)
+tempLR<-dplyr::select(LIs[LIs$exclude<2,],DL2.Lraw,DL2.Rraw,handpch)
 session<-2
@@
 #LIs exclude of 1 or 12 indicates excluded from CF; retain exclude 0,2
-tempLR<-select(LIs[LIs$exclude %in% c(0,2),],CF1.Lraw,CF1.Rraw,handpch)
+tempLR<-dplyr::select(LIs[LIs$exclude %in% c(0,2),],CF1.Lraw,CF1.Rraw,handpch)
 session<-1
@@
 myg5<-makerawplot(tempLR,session,myname,mylim)
-tempLR<-select(LIs[LIs$exclude %in% c(0,2),],CF2.Lraw,CF2.Rraw,handpch)
+tempLR<-dplyr::select(LIs[LIs$exclude %in% c(0,2),],CF2.Lraw,CF2.Rraw,handpch)
 session<-2
@@
 myname<-'N finger sequences'
-tempLR<-select(LIs,finger.L1,finger.R1,handpch)
+tempLR<-dplyr::select(LIs,finger.L1,finger.R1,handpch)
 session<-1
@@
 myg7<-makerawplot(tempLR,session,myname,mylim)
-tempLR<-select(LIs,finger.L2,finger.R2,handpch)
+tempLR<-dplyr::select(LIs,finger.L2,finger.R2,handpch)
 session<-2
@@
 RDT.short.dat<-allsum[allsum$exRDT==0,]
-tempRDT <- select(RDT.short.dat,RDT1.pcorrL,RDT1.pcorrR,RDT2.pcorrL,RDT2.pcorrR)
+tempRDT <- dplyr::select(RDT.short.dat,RDT1.pcorrL,RDT1.pcorrR,RDT2.pcorrL,RDT2.pcorrR)
 allmeanpcorr <- mean(colMeans(tempRDT,na.rm=T))
@@
 # create subset of data frame for analysis - I used LI_z measures, as they are on similar scale
-interRR <- dplyr::select(interR, 'zlat_RDT_1', 'zlat_RDT_2', 'Day1_Dich_acc_Z',
+interRR <- dplyr::select(LIs, 'zlat_RDT_1', 'zlat_RDT_2', 'Day1_Dich_acc_Z',
                          'Day2_Dich_acc_Z', 'Day1_CF_acc_Z', 'Day2_CF_acc_Z','handedness')
@@
 # now create subset of data frame for analysis
-interRR <- dplyr::select(interR, 'zlat_RDT_1', 'zlat_RDT_2', 'Day1_Dich_acc_Z',
-                         'Day2_Dich_acc_Z', 'Day1_CF_acc_Z', 'Day2_CF_acc_Z','handedness')
-# rename
+interRR <- dplyr::select(LIs, zlat_RDT_1, zlat_RDT_2, Day1_Dich_acc_Z,
+                         Day2_Dich_acc_Z, Day1_CF_acc_Z, Day2_CF_acc_Z,handedness)
+# rename, and reverse sign for CF
 interRR <- interRR %>% 
@@
 interRR<-na.omit(interRR)
+interRR$faces1 <- (-interRR$faces1) #reverse sign for faces
+interRR$faces2 <- (-interRR$faces2)
 # subset into right and left handers

### 2020-07-30 online_laterality_reliability.Rmd
runs with knitr

@@
 output:
+  word_document: default
   html_notebook: default
+  html_document:
+    df_print: paged
   pdf_document: default
-  word_document: default
 ---
@@
 #### Visual half-field paradigm
+
 In this paradigm, visual stimuli are briefly presented to the left or right of a central fixation stimulus.
@@
 #### Dichotic listening
+
 Simultaneous presentation of different auditory stimuli to left and right ears for immediate recall was a method originally used in studies of attention, but it became evident that there was usually an advantage for verbal stimuli presented to the right ear (Kimura, 1967).
@@
 
-With regards to visual half-field techniques for studying language lateralisation, there is evidence to suggest that the visual field advantage varies with handedness (Elias, Bulman-Fleming, & McManus, 1999; Nicholls, 1994, 1998; Weekes, Capetillo-Cuncliffe, Rayman, Iacoboni, & Zaidel, 1999).
+With regards to visual half-field techniques for studying language lateralisation, there is evidence to suggest that the visual field advantage varies with handedness (Elias, Bulman-Fleming, & McManus, 1999; Weekes, Capetillo-Cunliffe, Rayman, Iacoboni, & Zaidel, 1999).
 However, a relatively well powered study, involving 100 participants conducted by Willemin et al.
@@
 For chimeric faces, which typically show a left visual field advantage there again appears to be a greater departure from the typical pattern of laterality for left-handers (e.g.
-Gilbert & Bakan, 1973; Heller & Levy, 1983; Levy et al., 1983; Rozkowski & Snelbecker, 1982).
+Gilbert & Bakan, 1973; Levy et al., 1983; Rozkowski & Snelbecker, 1982).
 This was also studied in the study by Karlsson et al.
@@
 Common to all hypotheses is the assumption that there would be a correlation between the degree of laterality across complementary tasks.
-While studies do report significant correlations between measures of lateralisation, these are often small (Bradzakova-Trajkov, Haberling, Roberts, & Corballis, 2010).
+While studies do report significant correlations between measures of lateralisation, these are often small (Badzakova-Trajkov, Haberling, Roberts, & Corballis, 2010).
 Furthermore, there are a large number of studies which report no correlation between laterality indices (Bryden, Hecaen, & Deagostini, 1983; Rosch, Bishop, & Badcock, 2012; Whitehouse & Bishop, 2009; van der Haegen & Brysbaert, 2018).
@@
 Prior studies have been inconclusive because it has been difficult gather large amounts of data to establish test-retest reliability.
-For instance, Van der Haegen and Brysbaert (2018), found that language laterality did not correlate with face laterality in visual half-field tasks.
+For instance, van der Haegen and Brysbaert (2018), found that language laterality did not correlate with face laterality in visual half-field tasks.
 However, their study was restricted to left-handers, who may be atypical in this regard, and they noted that lack of association was hard to interpret because it may be due to unreliability of measurement.
@@
 
-The chimeric faces stimuli were kindly provided by Dr Michael Burt, and have been reported elsewhere (Burt & Perrett, 1997; Innes et al., 2016; Karlsson et al., 2019).
+The chimeric faces stimuli were kindly provided by Dr Michael Burt, and have been reported elsewhere (Burt & Perrett, 1997; Innes, Burt, Birch, & Hausmann, 2016; Karlsson et al., 2019).
 
@@
 library(flextable)
+library(cocor)
 options(knitr.kable.NA = '') #should suppress NA in tables
@@
 ```
-<!---Figure was not so easy to read : I think table preferable here--->
-```{r simplifiedlatdescriptives}
+
+Out of 441 participants recruited via Prolific, `r nrow(LIs_c)` (`r round(nrow(LIs_c)/441*100, 1)`%) completed both sessions.
+We then removed `r nrow(LIs_no)` participants (`r round(nrow(LIs_no)/nrow(LIs_c)*100, 1)`%) from our final sample as their lexTALE score was below 80, indicating that a participant’s level of vocabulary knowledge was below the cut off for lower advanced proficiency.
+This left a sample of `r nrow(LIs)` participants.
+
+##### Data cleaning
+
+For each task we adopted the following data cleaning procedures.
+Applying each of these procedures meant that the number of participants left in the final analyses could vary by task.
+First, for all tasks, we removed trials in which response times were shorter than 200 ms.
+Response times were log-transformed to reduce skew, as specified in our pre-registration.
+Outlier reaction times were identified using Hoaglin and Iglewicz’s (1987) procedure, whereby outliers are defined as response times that are 1.65 times the difference between the first and third quantiles (1.65 * (Q3-Q1)) below or above the first and third quantile values (e.g: lower limit = Q1 – 1.65 x (Q3-Q1); upper limit = Q3 + 1.65 x (Q3-Q1)).
+Our differences criterion was set to 1.65 instead of the typical 2.2 because we expected to obtain asymmetric distributions for response time, with most outliers in one tail of the distribution.
+The Hoaglin and Iglewicz procedure was applied to each participant's data separately.
+After applying each data trimming procedure, participants left with less than 80% of trials were removed.
+
+For the chimeric faces task, we removed participants if they scored lower than 75% on trials in which an identical emotion was shown in each hemiface, and for the dichotic listening task, we removed those who scored less than 75% correct on trials where the same sound was played to each ear.
+These data trimming procedures left data for `r nrow(LIs)-nrow(LIs[LIs$exclude== 1 | LIs$exclude== 12,])` participants on the chimeric faces `r nrow(LIs)-nrow(LIs[LIs$exclude== 2 | LIs$exclude== 12,])` participants on the dichotic listening task.
+An additional 7 cases had to be removed because of a technical error that meant they had less than 80 trials presented.
+
+For the rhyme decision task, we removed participants who scored less than 50% accuracy in either visual field.
+Furthermore, we removed participants who scored less than 75% accuracy on the catch trials (trials in which there was no rhyme).
+These trimming procedures left data for `r nrow(LIs[LIs$exRDT== 0,])` participants on this task.
+We considered recruiting further participants to make up the numbers, but this seemed uneconomic as we already had more than 300 cases for the other tasks, and we could tell by this point that the test-retest reliability of the rhyme detection task was unsatisfactory; With `r nrow(LIs[LIs$exRDT== 0,])` participants, the test-retest reliability was *r*= `r round(cor(LIs[LIs$exRDT== 0,]$Day1.RDT_RT_LI, LIs[LIs$exRDT== 0,]$Day2.RDT_RT_LI, use = "complete.obs"),2)`, which is smaller than the *r*= 0.65 that we pre-registered as our criterion for a reliable task.
+We plan to evaluate modified versions of this task in future, but decided to stop data collection at this point and analysed the `r nrow(LIs)` participants with data on other tasks.
+
+For the finger-tapping task we added a data-cleaning step that had not been pre-registered, to remove participants where there were more than 20 keypresses across left and right hands that did not correspond to the  intended sequence.
+This could reflect that the individual had forgotten the target sequence, or was not co-operating with instructions, but scrutiny of cases where this occurred suggested that it usually arose because of technical problems.
+For instance, some individuals had repeated recording of the same keypress at short intervals, suggestive of a sticky keyboard.
+The cutoff of 20 was arbitrarily selected, and was successful in excluding cases with abnormally low scores in one session.
+
+The final group of `r nrow(LIs)` participants (`r nrow(LIs[LIs$gender== "Female",])` female) had a mean age of `r round(mean(LIs$age, na.rm= TRUE), 1)` years (*SD*= `r round(sd(LIs$age, na.rm= TRUE), 2)` and included `r nrow(LIs[LIs$handedness== "Left",])` left handers (`r nrow(LIs[LIs$handedness== "Left" & LIs$gender== "Female",])` female).
+`r nrow(LIs[LIs$handedness== "Right" & LIs$bilingual== "Yes",])` right-handers were bilingual and `r nrow(LIs[LIs$handedness== "Left" & LIs$bilingual== "Yes",])` left-handers were bilingual.
+
+##### Methods for calcuating laterality indices, reliability, and population level analysis
+
+We preregistered that we would calculate LIs in two ways: (1) in the traditional fashion (100 x (Right – Left)/(Right + Left)); (2) a standardized LI that involved computing a z-score that represented, for each individual, a comparison between left- and right-sided responses.
+In fact, for RT data we specified a t-score, but t and z converge when sample size exceeds 30, and for simplicity we refer to these as LI_z scores.
+These LI_z scores are very highly correlated with the traditional LI measures (correlations typically exceed .95), but can be used to provide a common metric for RT and accuracy measures when identifying individuals who are significantly lateralised.
+For example, using a conventional 2-sided alpha of .05, we can categorise individuals with an absolute LI_z greater than 1.96 as significantly lateralised.
+
+For the rhyme task, we pre-registered that we would calculate LIs based on log response time (RT) data, as accuracy during piloting was at ceiling.
+To obtain LI_z we ran a *t*-test for each participant where accurate log RTs were the dependent variable and visual field was the independent variable.
+
+For the dichotic listening and chimeric faces, we specified that LI calculated would be based on trials in which participants correctly identified a stimulus.
+The count of correct responses that corresponded to each side was used to generate an accuracy LI in two ways: (1) in the traditional fashion (100 x (Right – Left)/(Right + Left)); (2) as a *z*-statistic for each participant, using z = (pR-.5)/sqrt(pR x pL/n), where pR is the proportion of correct responses in the RVF, pL is proportion of L responses in the LVF, and n is total LVF and RVF responses.
+This method can yield extreme z-scores (for instance, if 83/89 responses are to the right, then z = 15.92), so scores were censored to be in the range -5 to +5.
+Finally, LIs for the finger tapping task were calculated in the same manner,  using the number of sequences generated by the left and right hand as opposed to the proportion of correct responses in each visual field.
+
+To address hypothesis 1 and examine the laterality of each task at the population level, we conducted a series of one-sample *t*-tests for each task using session 1 data.
+
+For the reliability assessment, we checked LIs for normality using a Shapiro-Wilk test.
+As most LIs were non-normal, we used a series of Spearman’s correlations to examine the correlation of coefficients for session 1 versus session 2.
+For tasks where the test-retest reliability was above .75 with the existing test length, we reanalysed the data using just the first N trials, reducing N until reliability becomes unsatisfactory (below *r*= .65).
+This was of interest in indicating the most efficient way to administer these tasks in future, while retaining good reliability.
+However, the full data (including all trials) was used for all subsequent analyses.
+
+##### Atypical laterality
+
+To address hypothesis 2, that extreme left-handers will be more likely to show atypical laterality than right-handers, we categorised participants as extreme left-handers if they scored less than -90 and right-handers if they scored more than 90 on the Edinburgh Handedness Inventory.
+Our preregistration specified that atypicality would first be defined according to whether the participant's LI_z was significantly lateralised (with alpha = .05) in the opposite direction from the majority.
+If there were less than 20 cases defined as atypical using this method, we planned to adjust the cut-off on that task until we acquired 20 atypical participants.
+
+For Rhyme Decision and Chimeric Faces task, participants were categorised into four groups: typical on both; atypical on Rhyme Decision only; atypical on Chimeric Faces only, and atypical on both.
+A chi-square test was then used to test for association with the number of people in each of the four groups and the handedness classification from the Edinburgh Handedness Inventory.
+This was then be repeated for the Dichotic Listening and the Chimeric Faces tasks.
+A Bonferroni correction was applied to the outcome of each of the two chi-squared tests.
+
+##### Relationship between language and non-language LIs
+
+We examined the relationship between LIs on the Rhyme Decision, Dichotic Listening, and Chimeric faces by modelling covariances.
+This allowed us to compare (1) a model where the two language tasks are correlated with each other but not with the chimeric faces, (2) a model where all LIs are correlated, and (3) a base model where the LIs are unrelated.
+This was completed separately for right- and left-handers using both session 1 and session 2 data.
+
+## Results
+
+Lateral preference characteristics of the sample after exclusions as noted above is shown in Table 1.
+As participants self-reported eye dominance was always consistent between the Miles and Porta tests, we only report data from the Miles test.
+
+```{r simplifiedlatdescriptives, echo- FALSE, include= FALSE}
 #Miles and Porta are identical so only report one.
@@
 lattab$`L.handers.%`[6]<-NA
-
-```
-Out of 441 participants recruited via Prolific, `r nrow(LIs_c)` (`r round(nrow(LIs_c)/441*100, 1)`%) completed both sessions.
-We then removed `r nrow(LIs_no)` participants (`r round(nrow(LIs_no)/nrow(LIs_c)*100, 1)`%) from our final sample as their lexTALE score was below 80, indicating that a participant’s level of vocabulary knowledge was below the cut off for lower advanced proficiency.
-This left a sample of `r nrow(LIs)` participants.
-
-##### Data cleaning
-For each task we adopted the following data cleaning procedures.
-Applying each of these procedures meant that the number of participants left in the final analyses could vary by task.
-First, for all tasks, we removed trials in which response times were shorter than 200 ms.
-Response times were log-transformed to reduce skew, as specified in our pre-registration.
-Outlier reaction times were identified using Hoaglin and Iglewicz’s (1987) procedure, whereby outliers are defined as response times that are 1.65 times the difference between the first and third quantiles (1.65 * (Q3-Q1)) below or above the first and third quantile values (e.g: lower limit = Q1 – 1.65 x (Q3-Q1); upper limit = Q3 + 1.65 x (Q3-Q1)).
-Our differences criterion was set to 1.65 instead of the typical 2.2 because we expected to obtain asymmetric distributions for response time, with most outliers in one tail of the distribution.
-The Hoaglin and Iglewicz procedure was applied to each participant's data separately.
-After applying each data trimming procedure, participants left with less than 80% of trials were removed.
-
-For the chimeric faces task, we removed participants if they scored lower than 75% on trials in which an identical emotion was shown in each hemiface, and for the dichotic listening task, we removed those who scored less than 75% correct on trials where the same sound was played to each ear.
-These data trimming procedures left data for `r nrow(LIs)-nrow(LIs[LIs$exclude== 1 | LIs$exclude== 12,])` participants on the chimeric faces `r nrow(LIs)-nrow(LIs[LIs$exclude== 2 | LIs$exclude== 12,])` participants on the dichotic listening task.
-An additional 7 cases had to be removed because of a technical error that meant they had less than 80 trials presented.
-
-For the rhyme decision task, we removed participants who scored less than 50% accuracy in either visual field.
-Furthermore, we removed participants who scored less than 75% accuracy on the catch trials (trials in which there was no rhyme).
-These trimming procedures left data for `r nrow(LIs[LIs$exRDT== 0,])` participants on this task.
-We considered recruiting further participants to make up the numbers, but this seemed uneconomic as we already had more than 300 cases for the other tasks, and we could tell by this point that the test-retest reliability of the rhyme detection task was unsatisfactory; With `r nrow(LIs[LIs$exRDT== 0,])` participants, the test-retest reliability was *r*= `r round(cor(LIs[LIs$exRDT== 0,]$Day1.RDT_RT_LI, LIs[LIs$exRDT== 0,]$Day2.RDT_RT_LI, use = "complete.obs"),2)`, which is smaller than the *r*= 0.65 that we pre-registered as our criterion for a reliable task.
-We plan to evaluate modified versions of this task in future, but decided to stop data collection at this point and analysed the `r nrow(LIs)` participants with data on other tasks.
-
-For the finger-tapping task we added a data-cleaning step that had not been pre-registered, to remove participants where there were more than 20 keypresses across left and right hands that did not correspond to the  intended sequence.
-This could reflect that the individual had forgotten the target sequence, or was not co-operating with instructions, but scrutiny of cases where this occurred suggested that it usually arose because of technical problems.
-For instance, some individuals had repeated recording of the same keypress at short intervals, suggestive of a sticky keyboard.
-The cutoff of 20 was arbitrarily selected, and was successful in excluding cases with abnormally low scores in one session.
-
-The final group of `r nrow(LIs)` participants (`r nrow(LIs[LIs$gender== "Female",])` female) had a mean age of `r round(mean(LIs$age, na.rm= TRUE), 1)` years (*SD*= `r round(sd(LIs$age, na.rm= TRUE), 2)` and included `r nrow(LIs[LIs$handedness== "Left",])` left handers (`r nrow(LIs[LIs$handedness== "Left" & LIs$gender== "Female",])` female).
-`r nrow(LIs[LIs$handedness== "Right" & LIs$bilingual== "Yes",])` right-handers were bilingual and `r nrow(LIs[LIs$handedness== "Left" & LIs$bilingual== "Yes",])` left-handers were bilingual.
-
-##### Methods for calcuating laterality indices, reliability, and population level analysis
-
-We preregistered that we would calculate LIs in two ways: (1) in the traditional fashion (100 x (Right – Left)/(Right + Left)); (2) a standardized LI that involved computing a z-score that represented, for each individual, a comparison between left- and right-sided responses.
-In fact, for RT data we specified a t-score, but t and z converge when sample size exceeds 30, and for simplicity we refer to these as LI_z scores.
-These LI_z scores are very highly correlated with the traditional LI measures (correlations typically exceed .95), but can be used to provide a common metric for RT and accuracy measures when identifying individuals who are significantly lateralised.
-For example, using a conventional 2-sided alpha of .05, we can categorise individuals with an absolute LI_z greater than 1.96 as significantly lateralised.
-
-For the rhyme task, we pre-registered that we would calculate LIs based on log response time (RT) data, as accuracy during piloting was at ceiling.
-To obtain LI_z we ran a *t*-test for each participant where accurate log RTs were the dependent variable and visual field was the independent variable.
-
-For the dichotic listening and chimeric faces, we specified that LI calculated would be based on trials in which participants correctly identified a stimulus.
-The count of correct responses that corresponded to each side was used to generate an accuracy LI in two ways: (1) in the traditional fashion (100 x (Right – Left)/(Right + Left)); (2) as a *z*-statistic for each participant, using z = (pR-.5)/sqrt(pR x pL/n), where pR is the proportion of correct responses in the RVF, pL is proportion of L responses in the LVF, and n is total LVF and RVF responses.
-This method can yield extreme z-scores (for instance, if 83/89 responses are to the right, then z = 15.92), so scores were censored to be in the range -5 to +5.
-Finally, LIs for the finger tapping task were calculated in the same manner,  using the number of sequences generated by the left and right hand as opposed to the proportion of correct responses in each visual field.
-
-To address hypothesis 1 and examine the laterality of each task at the population level, we conducted a series of one-sample *t*-tests for each task using session 1 data.
-
-For the reliability assessment, we checked LIs for normality using a Shapiro-Wilk test.
-As most LIs were non-normal, we used a series of Spearman’s correlations to examine the correlation of coefficients for session 1 versus session 2.
-For tasks where the test-retest reliability was above .75 with the existing test length, we reanalysed the data using just the first N trials, reducing N until reliability becomes unsatisfactory (below *r*= .65).
-This was of interest in indicating the most efficient way to administer these tasks in future, while retaining good reliability.
-However, the full data (including all trials) was used for all subsequent analyses.
-
-##### Atypical laterality
-
-To address hypothesis 2, that extreme left-handers will be more likely to show atypical laterality than right-handers, we categorised participants as extreme left-handers if they scored less than -90 and right-handers if they scored more than 90 on the Edinburgh Handedness Inventory.
-Our preregistration specified that atypicality would first be defined according to whether the participant's LI_z was significantly lateralised (with alpha = .05) in the opposite direction from the majority.
-If there were less than 20 cases defined as atypical using this method, we planned to adjust the cut-off on that task until we acquired 20 atypical participants.
-
-For Rhyme Decision and Chimeric Faces task, participants were categorised into four groups: typical on both; atypical on Rhyme Decision only; atypical on Chimeric Faces only, and atypical on both.
-A chi-square test was then used to test for association with the number of people in each of the four groups and the handedness classification from the Edinburgh Handedness Inventory.
-This was then be repeated for the Dichotic Listening and the Chimeric Faces tasks.
-A Bonferroni correction was applied to the outcome of each of the two chi-squared tests.
-
-##### Relationship between language and non-language LIs
-
-We had originally intended to examine the relationship between LIs on the Rhyme Decision, Dichotic Listening, and Chimeric faces by modelling covariances.
-This would have allowed us to compare (1) a model where the two language tasks are correlated with each other but not with the chimeric faces, (2) a model where all LIs are correlated, and (3) a base model where the LIs are unrelated.
-This would have been done separately for right- and left-handers.
-However, given poor reliability of the Rhyme Decision task, we decided that this analysis would not be appropriate.
-Instead, we examine the correlation coefficient for LIs on the Dichotic Listening and Chimeric Faces tasks.
-For completeness, we include the proposed analysis as a supplementary materials but do not draw conclusions based on this.
-
-## Results
-
-Lateral preference characteristics of the sample after exclusions as noted above is shown in Table 1. <!---use new lattab for table 1--->
+```
 
 Table 1 <!---New table - lattab to be added in nice format--->
-Count of left- and right-handers who report left or right on self-reported footedness, and eyedness (Miles test). 'Other' includes those with don't know responses.
+Count of left- and right-handers who report left or right on self-reported footedness, and eyedness (Miles test). 
+```{r Tab1,echo=FALSE, comment=FALSE, warning=FALSE}
+flextable(lattab)
+```
+*Note*. RR: right foot, right eye; RL: right foot, left eye; LR: left foot, right eye; LL: left foot, left eye. 'Other' inlcudes those who responded 'Don't know'.
+
+```{r computepcorrRDT, echo=FALSE, comment=FALSE, warning=FALSE}
+#revised this, as p.corr values now in allsum
+RDT.short.dat<-allsum[allsum$exRDT==0,]
+tempRDT <- dplyr::select(RDT.short.dat,RDT1.pcorrL,RDT1.pcorrR,RDT2.pcorrL,RDT2.pcorrR)
+allmeanpcorr <- mean(colMeans(tempRDT,na.rm=T))
+allsdpcorr <- sd(c(tempRDT$RDT1.pcorrL,tempRDT$RDT1.pcorrR,tempRDT$RDT2.pcorrL,tempRDT$RDT2.pcorrR),na.rm=T)
+```
+
+### Significance of lateralisation on each task
+
+For a preliminary impression of the data, Figure 3 shows scatterplots indicating scores on left and right sides for each task, with handedness colour-coded.
+For each task, mean for each side is shown as vertical and horizontal dotted lines, and equal performance on the two sides corresponds to the sloping red dotted line: thus an impression of whether the task is lateralised can be obtained by considering whether points cluster around the sloping lines symmetrically.
 
@@
     geom_hline(yintercept=meanR,linetype='dotted')+geom_vline(xintercept=meanL,linetype='dotted')+
-    theme(legend.position = "none") +
+    theme(legend.position = "none") + theme_classic() +
     theme(plot.title = element_text(size = 12)
@@
 ```
+
 ```{r dorawplotfunction,echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
@@
 
-```{r computepcorrRDT}
-#revised this, as p.corr values now in allsum
-RDT.short.dat<-allsum[allsum$exRDT==0,]
-tempRDT <- dplyr::select(RDT.short.dat,RDT1.pcorrL,RDT1.pcorrR,RDT2.pcorrL,RDT2.pcorrR)
-allmeanpcorr <- mean(colMeans(tempRDT,na.rm=T))
-allsdpcorr <- sd(c(tempRDT$RDT1.pcorrL,tempRDT$RDT1.pcorrR,tempRDT$RDT2.pcorrL,tempRDT$RDT2.pcorrR),na.rm=T)
-```
-### Significance of lateralisation on each task
-
-For a preliminary impression of the data, Figure 3 shows scatterplots indicating scores on left and right sides for each task, with handedness colour-coded.
-For each task, mean for each side is shown as vertical and horizontal dotted lines, and equal performance on the two sides corresponds to the sloping red dotted line: thus an impression of whether the task is lateralised can be obtained by considering whether points cluster around the sloping lines symmetrically.
+![Figure 3.
+Left vs right scores for sessions 1 and 2 for the behavioural laterality tasks.
+The diagonal dotted line shows point of equality: points above this are higher for R than L, and points below are higher for L than R.
+Mean values for R and L are shown as horizontal and vertical dotted lines.
+](all.raw.png)
 
@@
 Here, a strong impact of handedness is evident on inspection, with left-handers generally more accurate with the left hand, and right-handers more accurate with the right hand.
-
-![Figure 3.
-Left vs right scores for sessions 1 and 2 for the behavioural laterality tasks.
-The diagonal dotted line shows point of equality: points above this are higher for R than L, and points below are higher for L than R.
-Mean values for R and L are shown as horizontal and vertical dotted lines.
-](all.raw.png)
 
@@
 
+Table 2.Mean LIs and results from one-sample t-tests and Shapiro-Wilks tests for departure from normality
+```{r Tab2,echo=FALSE, comment=FALSE, warning=FALSE}
+tab2 <- flextable(LItable)
+autofit(tab2, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
+```
+
 Table 3 shows data from the LI_z measures, which make it easier to compare different tasks on the same scale.
@@
   LIztable[i,6]<-round(100*length(which(temp<(-1.96)))/length(temp),1)
-
 }
@@
 
+Table 3.Mean LI_z and results from one-sample t-tests and Shapiro-Wilks tests for departure from normality
+```{r Tab3,echo=FALSE, comment=FALSE, warning=FALSE}
+tab3 <- flextable(LIztable)
+autofit(tab3, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
+```
+
 ### Test-retest reliability
+
 The scatterplots depicting test-retest reliability are shown in Figure 4, with Spearman test-retest correlations in Table 4.
 These were based on LI_z values, but are essentially identical to values from unscaled LI values.
-Confidence intervals were estimated using the *spearman.ci()* function from the RVAideMemoire package (version 0.9-75; Herv, 2020) with 10,000 iterations.
+Confidence intervals were estimated using the *spearman.ci()* function from the RVAideMemoire package (version 0.9-75; Hervé, 2020) with 10,000 iterations.
 For all but the rhyme decision task, we obtained good test-retest reliability that exceeded our pre-registered threshold of *r*= .65.
+
 ```{r make.combotablex,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
@@
 # Also computes spearmans with CI and returns these from function
-densplots <- function(dat, time1,time2,name1,name2,scattername,y1, y2, x1, x2, ax, ay){
+densplots <- function(dat, time1,time2,name1,name2,scattername,y1, y2, x1, x2, ax, ay, remove, title){
   # plots
@@
   #Find position in top L corner for legend
-  legposx <- .2 #not working - needs tweaking!
+  legposx <- .25 #not working - needs tweaking!
   legposy <- .85
   #censor extreme values for plotting only
-  w<-which(time1<(-5) )
-  time1[w] <- (-5)
-  w<-which(time1>5) 
-  time1[w] <- (5)
-  w<-which(time2<(-5) )
-  time2[w] <- (-5)
-  w<-which(time2>5) 
-  time2[w] <- (5)
+  w<-which(time1<(-remove) )
+  time1[w] <- (-remove)
+  w<-which(time1>remove) 
+  time1[w] <- (remove)
+  w<-which(time2<(-remove) )
+  time2[w] <- (-remove)
+  w<-which(time2>remove) 
+  time2[w] <- (remove)
   #now do the plot
@@
     geom_rug(aes(color= handed3)) +
-    theme_classic() +
+    theme_classic(18) +
     ylab(name2) + 
@@
     annotate("text", x = ax, y = ay, label = mycorlabel) +
-    theme(legend.position = c(legposx,legposy), legend.title = element_blank()) + ylim(y1, y2) + xlim(x1, x2)
+    theme(legend.position = c(legposx,legposy), legend.title = element_blank()) + ylim(y1, y2) + xlim(x1, x2) +
+    ggtitle(title) + theme(plot.title = element_text(hjust = 0.5))
   # Set relative size of marginal plots (main plot 10x bigger than marginals)
-  bigplot<-ggExtra::ggMarginal(myscatter, groupColour = TRUE, groupFill = TRUE)
+#  bigplot<-ggExtra::ggMarginal(myscatter, groupColour = TRUE, groupFill = TRUE)
   scattername<-paste0('./plots/',scattername,".png")
-  ggsave(scattername,plot=bigplot,dpi = 300, width = 6, height = 6,)
+  ggsave(scattername,plot=myscatter,dpi = 300, width = 6, height = 6,)
   return(list(mycor,myci))
@@
 scattername<-"RhymeDecisionScatter"
-
 y1= -5
@@
 x2= 5
-ax= 2
-ay= -2
-mycors<-densplots(RD.dat, time1,time2,name1,name2,scattername, y1, y2, x1, x2, ax, ay) 
+ax= 5*.67
+ay= -5*.75
+remove= 5
+mycors<-densplots(RD.dat, time1,time2,name1,name2,scattername, y1, y2, x1, x2, ax, ay, remove, "(A)") 
 combotable[1,2]<-round(mycors[[1]],3)
@@
 scattername<-"DichoticScatter"
-
-mycors<-densplots(DL.dat, time1,time2,name1,name2,scattername,y1, y2, x1, x2, ax, ay)
+mycors<-densplots(DL.dat, time1,time2,name1,name2,scattername,y1, y2, x1, x2, ax, ay, remove,"(B)")
 combotable[2,2]<-round(mycors[[1]],3)
@@
 scattername<-"ChimericScatter"
-mycors<-densplots(CF.dat, time1,time2,name1,name2,scattername, y1, y2, x1, x2, ax, ay)
+mycors<-densplots(CF.dat, time1,time2,name1,name2,scattername, y1, y2, x1, x2, ax, ay, remove, "(C)")
 combotable[3,2]<-round(mycors[[1]],3)
@@
 scattername<-"FingerScatter"
-
-mycors<-densplots(FT.dat, time1,time2,name1,name2,scattername, y1, y2, x1, x2, ax, ay)
+mycors<-densplots(FT.dat, time1,time2,name1,name2,scattername, y1, y2, x1, x2, ax, ay, remove, "(D)")
 combotable[4,2]<-round(mycors[[1]],3)
 combotable[4,3]<-mycors[[2]]
-```
-
+
+```
+
+Table 4. Test-retest reliability estimates and 95% CIs for each behavioural task.
+```{r Tab4,echo=FALSE, comment=FALSE, warning=FALSE}
+tab4 <- flextable(combotable)
+autofit(tab4, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
+```
 Note.
@@
 
-![Figure 5.
-Scatter plots for day 1 and day 2 data for LI_z scores on each laterality task.
+![Figure 4.
+Scatter plots for day 1 and day 2 data for LI_z scores on each laterality task: (A) Rhyme decision, (B) Dichotic listening, (C) Chimeric faces, and (D) Finger tapping.
 Handedness is colour coded with extreme left-handers identified as those who have a score below -90 on the Edinburgh Handedness Inventory.
-Note that z-scores are censored at absolute values of 5.](combined_LI_cor.png)
+Note that z-scores are censored at absolute values of 5.](combined_LI_plot.png)
 
 ### Optimising the tasks
+
 To determine the smallest number of trials needed to achieve a test-retest reliability of *r*= .65, we recalculated LIs for the dichotic listening and chimeric faces tasks using the range of trials in which different emotions/sounds were presented to each hemiface/ear.
-The absolute test-retest correlation coefficients are shown in Figure 6.
+The absolute test-retest correlation coefficients are shown in Figure 5.
 A minimum of 60 trials was required to achieve a test-retest correlation of *r*= .65 using the dichotic listening task.
-However, as indicated in Figure 6, even with 100 trials the reliability estimate still occasionally approaches this 0.65 threshold.
+However, as indicated in Figure 5, even with 100 trials the reliability estimate still occasionally approaches this 0.65 threshold.
 Taking this lower number of trials, it would be able to present only two blocks to achieve this threshold of reliability for the dichotic listening task.
@@
 
-![Figure 6.
+![Figure 5.
 Test-retest correlations for (A) the dichotic listening task and (B) the chimeric faces task when using a variable number of trials to calculate laterality indices.
@@
 We defined 'atypical laterality' as laterality that is in the opposite direction to the majority of participants, with an absolute LI_z value greater than 1.96.
-Although the focus of our planned analyses included only right handers and extreme left handers, we show data for  the remaining left-handers for completeness.
-
-```{r makecontingencies,echo=FALSE}
+Although the focus of our planned analyses included only right handers and extreme left handers, we show data for  the remaining left-handers for completeness in Table 5.
+
+```{r makecontingencies,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
 #We start by using z-scores to make a 3-way categorisation into:
@@
 
-mytab2%>%
-  kable() %>%
-  kable_styling("condensed")
+#mytab2%>%
+#  kable() %>%
+#  kable_styling("condensed")
 
 tabr <- table(LIs$handed3,LIs$rhyme1.cut)
-chisq.test(tabr)
+#chisq.test(tabr)
 
 tabd <- table(LIs$handed3,LIs$dichotic1.cut)
-chisq.test(tabd)
+#chisq.test(tabd)
 
 tabc <- table(LIs$handed3,LIs$chimeric1.cut)
-chisq.test(tabc)
+#chisq.test(tabc)
 
@@
 
-Table 2. Frequencies of laterality typicality for right-, left-, and extreme left-handers across the Rhyme decision, Dichotic listening, and Chimeric faces tasks.
+Table 5. Frequencies of laterality typicality for right-, left-, and extreme left-handers across the Rhyme decision, Dichotic listening, and Chimeric faces tasks.
 ```{r Table2,echo=FALSE, comment=FALSE, warning=FALSE}
@@
 
-```{r pformatfunction}
+```{r pformatfunction, echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
 pformat<-function(pvalue){
@@
 For dichotic listening, chi squared = `r round(chi.d$statistic,2)`, d.f.
-= 1, pformat(p.adjust(chi.d$p.value, method= "bonferroni", n=2))`; for chimeric faces, chi squared = `r round(chi.c$statistic,2)`, d.f.
-= 1, pformat(p.adjust(chi.c$p.value, method= "bonferroni", n=2))`.
+= 1, `r pformat(p.adjust(chi.d$p.value, method= "bonferroni", n=2))`; for chimeric faces, chi squared = `r round(chi.c$statistic,2)`, d.f.
+= 1, `r pformat(p.adjust(chi.c$p.value, method= "bonferroni", n=2))`.
 
@@
 
-Initial scrutiny of Spearman correlations between LI_z scores suggests little evidence for any association between the three laterality indices, as shown in Table x.
-
-```{r bigcorrmatrix}
+Initial scrutiny of Spearman correlations between LI_z scores suggests little evidence for any association between the three laterality indices, as shown in Table 6.
+Table 6 suggests weak associations between LI_z values for the two language measures (rhyme decision and dichotic listening), but no association between either of these measures and LI_z on chimeric faces.
+
+```{r bigcorrmatrix,echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
 
@@
 # right_dat
-
-myft3<- flextable(allcorr)
-myft3
-
-```
-
-Table x shows Spearman correlations between the LI_z scores for the different tasks in two Sessions.
-The upper diagonal shows correlations for the `r NR` right-handers, and teh lower diagonal shows values for the `r NL` left-handers.
+```
+
+Table 6.
+Spearman correlations between the LI_z scores for the different tasks in two Sessions.
+The upper diagonal shows correlations for the 143 right-handers, and the lower diagonal shows values for the 66 left-handers.
+```{r tab 5, echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
+myft5<- flextable(allcorr)
+autofit(myft5, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
+```
+Note.
 For right-handers, an absolute value of .16 is significant at the .05 level, and a value of .22 is significant at the .01 level.
 For left-handers, a value of .25 is significant at the .05 level and a value of .32 is significant at the .01 level.
-
-Table x suggests weak associations between LI_z values for the two language measures (rhyme decision and dichotic listening), but no association between either of these measures and LI_z on chimeric faces.
 
@@
 
-```{r prepareKievit,echo=FALSE, comment=FALSE, warning=FALSE}
+```{r prepareKievit, echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
 library(lavaan)
@@
 
+As shown in Table 7, the Akaike weights favoured the base model for both left-handers and right-handers. This reflected the lack of cross-task correlation for LI_z measures.
+
+Table 7. Akaike weights for the independent base model, language model, and laterality model.
 ```{r AICmodels,echo=FALSE, comment=FALSE, warning=FALSE}
@@
  }
-flextable(AICdf)
-```
-
-As shown in Table x, the Akaike weights favoured the base model for both left-handers and right-handers. This reflected the lack of cross-task correlation for LI_z measures.
+myft6 <- flextable(AICdf)
+autofit(myft6, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
+```
 
@@
 ```{r exploreA, echo= FALSE, comment=FALSE, warning=FALSE}
+# make sure we have good cases
+interR <- LIs[LIs$exclude == 0,]
 DL.lex<-RVAideMemoire::spearman.ci(interR$Day1_DL_acc_LI, interR$lexTALE, nrep = 10000, conf.level = 0.95)
@@
 A common mechanism that codetermines lateralisation on a range of skills, extending beyond handedness and language, has been a hotly debated topic in the laterality literature.
-Several hypotheses postulate complementary lateralisation for left-hemisphere and right-hemisphere brain asymmetries, predicting that LIs on verbal and nonverbal tasks should be related (Behrmann & Plaut, 2015; Cook, 1984; Hellige, 1983; Kosslyn, 1987).
+Several hypotheses postulate complementary lateralisation for left-hemisphere and right-hemisphere brain asymmetries, predicting that LIs on verbal and nonverbal tasks should be related (Behrmann & Plaut, 2015; Hellige, 1983; Kosslyn, 1987).
 On this view, a person with atypical right-hemisphere language representation should have atypical left-hemisphere representation for nonverbal domains such as face processing.
@@
 
-<!---
-DOROTHY, I'VE LEFT THESE NOTES IN INCASE YOU FIND THEM USEFUL
-# Pre-reg info
-## Exploratory analysis
-As noted above, we will use the z- or t-statistics to index laterality, but we will have a large dataset that might be used subsequently to explore different approaches to measuring a laterality index, and to consider how choice of measure affects test-retest reliability.
-## Results
-#
-```{r prep_dat,echo=FALSE}
-# now create subset of data frame for analysis
-mycols <- c('Day1.RDT_RT_LI', 'Day2.RDT_RT_LI',
-            'Day1_CF_acc_LI', 'Day2_CF_acc_LI',
-            'Day1_DL_acc_LI', 'Day2_DL_acc_LI',
-            'Day1_finger_LI', 'Day2_finger_LI',
-            'handedness')
-colnums<- which(colnames(LIs) %in% mycols)
-M <- LIs[,colnums]
-M <- M %>% 
-  rename(
-    "rhyme1" = Day1.RDT_RT_LI,
-    "rhyme2" = Day2.RDT_RT_LI,
-    "chimeric1" = Day1_CF_acc_LI,
-    "chimeric2" = Day2_CF_acc_LI,
-    "dichotic1" = Day1_DL_acc_LI,
-    "dichotic2" = Day2_DL_acc_LI,
-    "finger1" = Day1_finger_LI,
-    "finger2" = Day2_finger_LI)
-# subset into right and left handers
-left_dat <- subset(M, handedness== "Left")
-right_dat <- subset(M, handedness== "Right")
-```
-```{r showcorrs,echo=FALSE}
-#Will focus just on Day1
-#Had considered showing as correlation heatmap - but values so small it wouldn't be worth doing!
-#http://www.sthda.com/english/wiki/ggplot2-quick-correlation-matrix-heatmap-r-software-and-data-visualization
-# day 1
-corrL<-round(cor(left_dat[,c(2,4,6,8)],use='complete.obs'),2)
-corrR<-round(cor(right_dat[,c(2,4,6,8)],use='complete.obs'),2)
-corr1 <- cbind(corrR, corrL)
-# day 2
-corrL2<-round(cor(left_dat[,c(3,5,7,9)],use='complete.obs'),2)
-corrR2<-round(cor(right_dat[,c(3,5,7,9)],use='complete.obs'),2)
-corr2 <- cbind(corrR2, corrL2)
-```
-### Table 3: Intratask correlations using session 1 data.
-```{r dotab5,echo=FALSE}
-corr1%>%
-  kable() %>%
-  kable_styling(c("bordered")) %>%
-  add_header_above(c(" " = 1, "Session 1" = 4, "Session 2" = 4))
-```
-The correlations between tasks are all very low. To check if this pattern is robust, this analysis was repeated for time 2 data.
-### Table 4: Intratask correlations using session 2 data.
-```{r dotab7,echo=FALSE}
-corr2%>%
-  kable() %>%
-  kable_styling(c("bordered")) %>%
-  add_header_above(c(" " = 1, "Session 1" = 4, "Session 2" = 4))
-```
---->
-
-References from DB
+\newpage
+## References
+
+Badzakova-Trajkov, G., Häberling, I. S., Roberts, R. P., & Corballis, M. C. (2010). Cerebral asymmetries: complementary and independent processes. *PloS one*, *5*, e9682.
+
+Beaumont, J. G. (1982). Studies with verbal stimuli. In J. G. Beaumont (Ed.), *Divided visual field studies of cerebral organisation* (pp. 58–86). London: Academic Press.
+
+Behrmann, M., & Plaut, D. C. (2015). A vision of graded hemispheric specialization. *Annals of the New York Academy of Sciences*, *1359*, 30-46.
+
+Bethmann, A., Tempelmann, C., De Bleser, R., Scheich, H., & Brechmann, A. (2007). Determining language laterality by fMRI and dichotic listening. *Brain research*, *1133*, 145-157.
+
+Bishop DVM. 2013 Cerebral asymmetry and language development: cause, correlate, or consequence? *Science*, *340*, 1230531.
+
+Bishop, D. V. (1990). *Handedness and developmental disorder*. Cambridge University Press.
+
+Bishop, D.
+V.
+M.
+(2017) *Which neuroimaging measures are useful for individual differences research?
+* [Blog post].
+Retrieved from: http://deevybee.blogspot.be/2017/05/whichneuroimaging-measures-are-useful.html?m=1
 
@@
 (2013).
-“Right on all occasions?” – On the feasibility of laterality research using a smartphone dichotic listening application.
-Frontiers in Psychology,, 4(42).
-doi:10.3389/fpsyg.2013.00042
-
-Kimura, D. (1967). Functional asymmetry of the brain in dichotic listening. Cortex, 3, 163-178.
-
-Mishkin, M., & Forgays, D. G. (1952). Word recognition as a function of retinal locus. Journal of Experimental Psychology, 43(1), 43-48. doi:10.1037/h0061361
+“Right on all occasions?”–On the feasibility of laterality research using a smartphone dichotic listening application.
+*Frontiers in Psychology*, *4*, 42.
+
+Boles, D. B. (1987). Reaction time asymmetry through bilateral versus unilateral stimulus presentation. *Brain and Cognition*, *6*, 321–333.
+
+Boles, D. B. (1990). What bilateral displays do. *Brain and Cognition*, *12*, 205– 228.
+
+Boles, D. B. (1994). An experimental comparison of stimulus type, display type and input variable contributions to visual field asymmetry. *Brain and Cognition*, *24*, 184–197.
+
+Bourne, V. J. (2006). The divided visual field paradigm: Methodological considerations. *Laterality*, *11*, 373–393
+
+Bryden, M. P. (1970). Laterality effects in dichotic listening: relations with handedness and reading ability in children. *Neuropsychologia*, *8*, 443-450.
+
+Bryden, M.
+P.
+(1988).
+An overview of the dichotic listening procedure and its relation to cerebral organization.
+In K.
+Hugdahl (Ed.), *Handbook of dichotic listening: Theory, methods and researc*h (p.
+1–43).
+John Wiley & Sons.
+
+Bryden, M. P., Hécaen, H., & DeAgostini, M. (1983). Patterns of cerebral organization. *Brain and language*, *20*, 249-262.
+
+Burt, D. M., & Perrett, D. I. (1997). Perceptual asymmetries in judgements of facial attractiveness, age, gender, speech and expression. *Neuropsychologia*, *35*, 685-693.
+
+Elias, L. J., Bulman-Fleming, M. B., & McManus, I. C. (1999). Visual temporal asymmetries are related to asymmetries in linguistic perception. *Neuropsychologia*, *37*, 1243-1249.
+
+Geffen, G., Bradshaw, J. L., & Nettleton, N. C. (1972). Hemispheric asymmetry: verbal and spatial encoding of visual stimuli. *Journal of Experimental Psychology*, *95*, 25-31.
+
+Gilbert, C., & Bakan, P. (1973). Visual asymmetry in perception of faces. *Neuropsychologia*, *11*, 355-362.
+
+Hellige, J. B. (1990). Hemispheric asymmetry. *Annual Review of Psychology*, *41*, 55-80.
+
+Hervé, M. (2020). *RVAideMemoire: Testing and Plotting Procedures for Biostatistics*. R package version 0.9-75.https://CRAN.R-project.org/package=RVAideMemoire
+
+Hoaglin, D. C., & Iglewicz, B. (1987). Fine-tuning some resistant rules for outlier labeling. *Journal of the American statistical Association*, *8*, 1147-1149.
+
+Hodgson, J. C., Hirst, R. J., & Hudson, J. M. (2016). Hemispheric speech lateralisation in the developing brain is related to motor praxis ability. *Developmental cognitive neuroscience*, *22*, 9-17.
+
+Hugdahl, K. (2000). Lateralization of cognitive processes in the brain. *Acta psychologica*, *105*, 211-235.
+
+Hugdahl, K. (2003). Dichotic listening in the study of auditory laterality. In K. Hugdahl and R. J. Davidson (Eds.), *The asymmetrical brain* (pp. 441–476). Cambridge, MA: MIT Press.
+
+Hugdahl, K., Westerhausen, R., Alho, K., Medvedev, S., Laine, M., & Hämäläinen, H.
+(2009).
+Attention and cognitive control: unfolding the dichotic listening story.
+*Scandinavian journal of psychology*, *50*, 11-22.
+
+Hunter, Z. R., & Brysbaert, M. (2008). Visual half-field experiments are a good measure of cerebral language dominance if used properly: Evidence from fMRI. *Neuropsychologia*, *46*, 316–325
+
+Iacoboni, M., & Zaidel, E. (1996). Hemispheric independence in word recognition: Evidence from unilateral and bilateral presentations. *Brain and Language*, *53*, 121–140
+
+Karlsson, E.
+M., Johnstone, L.
+T., & Carey, D.
+P.
+(2019).
+The depth and breadth of multiple perceptual asymmetries in right handers and non-right handers.
+*Laterality: Asymmetries of Body, Brain and Cognition*, *24*, 707-739.
+
+Kelley KS, Littenberg B. (2019). Dichotic Listening Test–retest reliability in children. *Journal of Speech, Language, and Hearing Research*, *62*,169-76.
+
+Kimura, D. (1967). Functional asymmetry of the brain in dichotic listening. *Cortex*, *3*, 163-178.
+
+Kosslyn, S. M. (1987). Seeing and imagining in the cerebral hemispheres: a computational approach. *Psychological review*, *94*, 148.
+
+Krach, S., Chen, L. M., & Hartje, W. (2006). Comparison between visual halffield performance and cerebral blood flow changes as indicators of language dominance. *Laterality*, *11*, 122-140
+
+Lemhöfer, K., & Broersma, M. (2012). Introducing LexTALE: A quick and valid lexical test for advanced learners of English. *Behavior research methods*, *44*, 325-343.
+
+Levy, J., Heller, W., Banich, M. T., & Burton, L. A. (1983). Asymmetry of perception in free viewing of chimeric faces. *Brain and cognition*, *2*, 404-419.
+
+Mazoyer, B., Zago, L., Jobard, G., Crivello, F., Joliot, M., Perchey, G., ...
+& Tzourio-Mazoyer, N.
+(2014).
+Gaussian mixture modeling of hemispheric lateralization for language in a large sample of healthy individuals balanced for handedness.
+*PloS one*, *9*, e101165.
+
+Miles, W. R. (1929). Ocular dominance demonstrated by unconscious sighting. *Journal of experimental psychology*, *12*, 113.
+
+Mishkin, M., & Forgays, D. G. (1952). Word recognition as a function of retinal locus. *Journal of experimental psychology*, *43*, 43.
+
+Ocklenburg, S., & Güntürkün, O. (2018). *The lateralized brain: The neuroscience and evolution of hemispheric asymmetries*. London: Academic Press.
+
+Oldfield, R. C. (1971). The assessment and analysis of handedness: the Edinburgh inventory. *Neuropsychologia*, *9*, 97-113.
+
+Porac, C., & Coren, S. (1976). The dominant eye. *Psychological bulletin*, *83*, 880.
+
+Rosch, R.
+E., Bishop, D.
+V., & Badcock, N.
+A.
+(2012).
+Lateralised visual attention is unrelated to language lateralisation, and not influenced by task difficulty–a functional transcranial Doppler study.
+*Neuropsychologia*, *50*, 810-815.
+
+Roszkowski, M. J., & Snelbecker, G. E. (1982). Temporal stability and predictive validity of self-assessed hand preference with first and second graders. *Brain and Cognition*, *1*, 405-409.
+
+Van der Haegen, L., & Brysbaert, M. (2018). The relationship between behavioral language laterality, face laterality and language performance in left-handers. *PloS one*, *13*, e0208696.
+
+Van Heuven, W.
+J., Mandera, P., Keuleers, E., & Brysbaert, M.
+(2014).
+SUBTLEX-UK: A new and improved word frequency database for British English.
+*Quarterly journal of experimental psychology*, *67*, 1176-1190.
+
+Voyer, D. (1998). On the reliability and validity of noninvasive laterality measures. *Brain & Cognition*, *36* , 209-36.
+
+Weekes, N.
+Y., Capetillo-Cunliffe, L., Rayman, J., Iacoboni, M., & Zaidel, E.
+(1999).
+Individual differences in the hemispheric specialization of dual route variables.
+*Brain and language*, *67*, 110-133.
+
+Westerhausen, R. (2019). A primer on dichotic listening as a paradigm for the assessment of hemispheric asymmetry. *Laterality: Asymmetries of Body, Brain and Cognition*, *24*, 740-771.
+
+Westerhausen, R., & Samuelsen, F. (2020, January 31). An optimal dichotic-listening paradigm for the assessment of hemispheric dominance for speech processing. https://doi.org/10.31234/osf.io/tpjes
+
+Whitehouse, A. J., & Bishop, D. V. (2009). Hemispheric division of function is the result of independent probabilistic biases. *Neuropsychologia*, *47*, 1938-1943.
+
+Willemin, J., Hausmann, M., Brysbaert, M., Dael, N., Chmetz, F., Fioravera, A., Gieruc, K., & Mohr, C.
+(2016).
+Stability of right visual field advantage in an international lateralized lexical decision task irrespective of participants’ sex, handedness or bilingualism.
+*Laterality: Asymmetries of Body, Brain and Cognition*, *21*, 502-524.
+
+Woods, K. J., Siegel, M. H., Traer, J., & McDermott, J. H. (2017). Headphone screening to facilitate web-based auditory experiments. *Attention, Perception, & Psychophysics*, *79*, 2064-2072.
+
+Yovel, G., Tambini, A., & Brandman, T.
+(2008).
+The asymmetry of the fusiform face area is a stable individual characteristic that underlies the left-visual-field superiority for faces.
+*Neuropsychologia*, *46*, 3061-3068.

### 2020-08-03 online_laterality_reliability.Rmd
zoe's feedback

@@
 <!--- updated 20/7/20 to change 'select' to 'dplyr::select' to avoid conflicts-->
+<!--- updated 3/8/20 to add ZW feedback--->
 Adam J. Parker
@@
 
-<!--- not sure how corresponding authorship works in OSCCI lab --->
-<!--- I'm fine with you in this role--->
 *Correspondence regarding this article should be addressed to Adam J.
@@
 We recruited 392 participants, oversampling left-handers, who completed the battery twice.
-Three of the tasks showed evidence of both validity and reliability, insofar as they showed hemispheric advantages in the expected direction and test-retest reliability of at least .65.
-The rhyme judgement task was less satisfactory.
+Three of the tasks showed evidence of both validity and reliability, insofar as they showed hemispheric advantages in the expected direction and test-retest reliability of at least *r*= .65.
+The reliability of the rhyme judgement task was less satisfactory (*r*= .63).
 We also found that left-handers were more likely to depart from typical lateralisation.
 Contrary to prediction, lateralisation across tasks was not significantly correlated.
-We conclude that three of the tasks, dichotic listening, chimeric faces and finger-tapping, show considerable promise for online evaluation of cerebral lateralisation, but the results do not support the notion of a common underlying cause of lateralisation for different cognitive tasks.
+We conclude that three of the tasks, dichotic listening, chimeric faces and finger tapping, show considerable promise for online evaluation of cerebral lateralisation, but the results do not support the notion of a common underlying cause of lateralisation for different cognitive tasks.
 
@@
 Reliability can either be assessed within a test session (e.g.
-split-half reliability, from correlations between odd and even items) or from repeated assessments on different occasions, test-retest reliability.
-The latter is generally preferred as it includes all test items and will incorporate fluctuations in performance due to changes in state or test conditiions from one occasion to another.
+split-half reliability, from correlations between odd and even items) or from repeated assessments on different occasions (test-retest reliability).
+The latter is generally preferred as it includes all test items and will incorporate fluctuations in performance due to changes in state or test conditions from one occasion to another.
 Assessment of test-retest reliability, however, requires large samples of people who undergo repeated testing, and this is seldom feasible for in-person testing.
@@
 
-### Behavioural laterality tasks and their reliability.
+### Behavioural laterality tasks and their reliability
 
@@
 In this paradigm, visual stimuli are briefly presented to the left or right of a central fixation stimulus.
-The visual pathways from the left and right sides of space to visual cortex are crossed for non-foveal viewing, and so the right eye projects directly to the left hemisphere, and the left eye to the right hemisphere.
+The visual pathways from the left and right sides of space to visual cortex are crossed for non-foveal viewing, and so the right visual-field (RVF) projects directly to the left hemisphere, and the left visual-field (LVF) to the right hemisphere.
 The corpus callosum allows for transfer between hemispheres, but the rationale of the method is that there will be a processing advantage for language stimuli that are directly projected to the language-dominant hemisphere.
 This method has been used since the 1950s (Mishkin & Forgays, 1952).
-Bourne (2006) reviewed the literature, which consistently shows advantageous processing for letter strings presented in the right, relative to the left, visual field in most people.
-
-Despite wide reporting of a right visual field advantage at the population level,  the validity of visual half-field tasks has been questioned.
+Bourne (2006) reviewed the literature, which consistently shows advantageous processing for letter strings presented in the RVF, relative to the LVF, in most people.
+
+Despite wide reporting of a RVF advantage at the population level,  the validity of visual half-field tasks has been questioned.
 Krach, Chen, and Hartje (2006) compared the laterality indices (LIs) produced during a visual half-field experiment with those produced using functional Transcranial Doppler Sonography (fTCD).
-The visual half-field task consisted of lexical decisions made towards abstract nouns presented in the left- and right visual-fields.
+The visual half-field task consisted of lexical decisions made towards abstract nouns presented in the LVF and RVF.
 During the experiment, the blood flow to left and right hemispheres of 58 participants was measured.
-In addition, blood flow was measured during a word generation task — the gold standard task for assessing hemispheric dominance for language.
-Despite showing the expected right visual field advantage on the visual half-field task, the behavioural LIs produced on the visual half-field paradigm did not correlate with the fTCD LIs produced during word generation (r= .18) or the visual half-field task itself (r -.004).
-The lack of validity here likely reflects the poor reliability of the visual field task as split-half reliability was very low (RT spilt-half: r= .14; accuracy split-half: r= .20).
+In addition, blood flow was measured during a word generation task— the gold standard task for assessing hemispheric dominance for language.
+Despite showing the expected RVF advantage on the visual half-field task, the behavioural LIs produced on the visual half-field paradigm did not correlate with the fTCD LIs produced during word generation (*r*= .18) or the visual half-field task itself (r -.004).
+The lack of validity here likely reflects the poor reliability of the visual field task as split-half reliability was very low (reaction time spilt-half: *r*= .14; accuracy split-half: *r*= .20).
 This illustrates the difficulty of drawing conclusions about associations between laterality indices when reliability of measurement is so low.
@@
 In recent years it has become clear that the reliability and validity of visual half-field tasks to identify the language-dominant hemisphere is largely dependent on several methodological aspects.
-For example, Hunter and Brysbaert (2008) noted stronger and more stable visual-field differences when stimuli were presented bilaterally relative to those where only one stimulus was presented either in the LVF or in the RVF (Boles, 1987, 1990, 1994; see also Iacoboni & Zaidel, 1996).
+For example, Hunter and Brysbaert (2008) noted stronger and more stable visual field differences when stimuli were presented bilaterally relative to those where only one stimulus was presented either in the LVF or in the RVF (Boles, 1987, 1990, 1994; see also Iacoboni & Zaidel, 1996).
 Brysbaert and Hunter argued that this is the result of stimuli competing during bilateral presentation, enhancing the advantage to the stimulus presented in the language dominant hemisphere.
@@
 When accounting for these methodological aspects, the validity of visual half-field tasks seems promising.
-For example, Hunter and Brysbaert (2008) used two visual half-field tasks, word naming (192 trials) and picture naming (160) in which stimuli were bilaterally presented and masked at offset.
-The two tasks correlated .8 (95% CI .598-.956) in a sample of 26 left-handers.
+For example, Hunter and Brysbaert (2008) used two visual half-field tasks, word naming (192 trials) and picture naming (160 trials) in which stimuli were bilaterally presented and masked at offset.
+The two tasks correlated strongly (*r*= .8, 95% CI .598-.956) in a sample of 26 left-handers.
 Encouragingly, the LIs from these visual half-field tasks also correlated positively with laterality from functional Magnetic Resonance Imaging (fMRI) on a word generation task in 10 left-handers.
-However, the small sample size means that the confidence intervals around these estimates are very wide: word naming, r= .63 (95% CI = .001-.902), picture naming, r= .77 (95% CI = .272-.943).
-<!---I removed Hausmann et al in interests of length, as they didn't report reliability--->
+However, the small sample size meant that the confidence intervals around these estimates were very wide: word naming, *r*= .63 (95% CI = .001-.902), picture naming, *r*= .77 (95% CI = .272-.943).
 
@@
 First, based on a review of the published literature, Westerhausen and Samuelsen outlined eight design features that should be used in an optimal dichotic listening paradigm and explain their justification for each.
-The criteria are as follows: (1) stop consonant-vowel (CV) syllables as stimulus materials, (2) pair only CV stimuli from the same voicing category, (3) Alternating trials of voiced and unvoiced stimulus pairs, (4) free-recall instruction, (5) single-stimulus pair per trial, (6) paradigm length of 120 dichotic trials, (7) all stimulus pairs are presented to the two sides in identical frequency, and (8) includes binaural (homonymic) trials.
+Their recommendations were to use: (1) stop consonant-vowel (CV) syllables as stimulus materials; (2) CV stimulus pairs from the same voicing category; (3) alternating trials of voiced and unvoiced stimulus pairs; (4) free-recall instruction; (5) onw stimulus pair per trial; (6) a paradigm length of 120 dichotic trials; (7) all stimulus pairs are presented to the two sides in identical frequency; and (8) to include binaural (homonymic) trials.
 Putting these criteria into practice, Westerhausen and Samuelsen administered a modified dichotic listening paradigm using both a verbal and manual response format to 50 right-handers.
-Interestingly, the full paradigm yielded a test-retest reliability of r= .93 for verbal responses and r= .91 for manual responses.
+Interestingly, the full paradigm yielded a test-retest reliability of *r*= .93 for verbal responses and *r*= .91 for manual responses.
 Thus, the dichotic listening task provides a promising paradigm for those wanting to investigate language lateralisation without reliance on visual stimuli.
@@
 In studies of face processing, it is possible to use a modification of the visual half-field paradigm in which single stimuli are viewed, composed of vertically split chimeric faces, where each half displays a different emotion or identity.
-Our focus here is on emotion detection in chimeric faces, where a bias is found towards reporting the emotion shown in the left hemiface.
+Our focus here is on emotion detection in chimeric faces, where a bias is usually found towards reporting the emotion shown in the left hemiface.
 
 The reliability of the chimeric faces tasks seem promising.
-Early work by Levy et al (1983) using two vertically presented chimeric faces, found split-half reliability of .93.
+Early work by Levy et al (1983) using two vertically presented chimeric faces, found split-half reliability of *r*= .93.
 The validity of the chimeric faces task has been demonstrated using behavioural and fMRI data, but in a task of recognising facial identity rather than emotion.
-In Yovel, Tambini, and Brandman’s (2008) study, participants performed a behavioural experiment where they had to recognize centrally presented chimeric faces, which presented different identities in the right- and left-visual-field.
-It is encouraging to note that the behavioural measure significantly correlated with the LI calculated based on activation in the fusiform face area (r= .49).
+In Yovel, Tambini, and Brandman’s (2008) study, participants performed a behavioural experiment where they had to recognize centrally presented chimeric faces, which presented different identities in the right and left visual fields.
+It is encouraging to note that the behavioural measure significantly correlated with the fMRILI calculated from on activation in the fusiform face area (*r*= .49).
 However, the sample size used here was small: with 17 participants, the 95% confidence interval around the observed correlation of .49 ranged from .012 to .786.
@@
 
-The studies reviewed so far have indicated consistent population biases in lateralisation on behavioural tasks, but also individual differences.
-One factor that has been associated with the degree of lateralisation is handedness.
-Neuroimaging data indicates that for language tasks these differences in lateralisation reflect the contribution of a subset of strong left-handers who are right language dominant, rather than indicating that this relationship operates on a continuum (Mazoyer et al., 2014).
+The studies reviewed so far have indicated consistent population biases in lateralisation on behavioural tasks, but also there are also substantial individual differences.
+One factor that has been associated with the degree of lateralisation is handedness (for a comprehensive review, see Carey & Johnstone, 2014).
+Neuroimaging data indicates that for language tasks these differences in lateralisation primarily reflect the contribution of a subset of strong left-handers who are right hemisphere dominant for language, rather than indicating that this relationship operates on a continuum (Mazoyer et al., 2014).
 This association can be used to validate behavioural measures of lateralisation: we would expect to see departure from typical lateralisation in strong left-handers.
 
-With regards to visual half-field techniques for studying language lateralisation, there is evidence to suggest that the visual field advantage varies with handedness (Elias, Bulman-Fleming, & McManus, 1999; Weekes, Capetillo-Cunliffe, Rayman, Iacoboni, & Zaidel, 1999).
-However, a relatively well powered study, involving 100 participants conducted by Willemin et al.
-(2016) reported no hint of any effect of handedness, when coded as left and right, or when divided into left (N = 19), mixed (N = 24) and right-handed (N = 57) groups, on the visual field advantage during a lexical decision task.
+With regards to visual half-field tasks for studying language lateralisation, there is evidence to suggest that the visual field advantage varies with handedness (Elias, Bulman-Fleming, & McManus, 1999; Weekes, Capetillo-Cunliffe, Rayman, Iacoboni, & Zaidel, 1999).
+However, a relatively well powered study involving 100 participants, conducted by Willemin et al.
+(2016), reported no hint of any effect of handedness, when coded as left and right, or when divided into left (N = 19), mixed (N = 24) and right-handed (N = 57) groups, on the visual field advantage during a lexical decision task.
 This result illustrates the recurring challenge of interpreting null results in this field: do they reflect a true lack of relationship, poor test reliability, inadequate power to detect small effects, or specific methodological factors?
@@
 
-On dichotic listening paradigms, the clear right ear (i.e.
+On dichotic listening tasks, the clear right ear (i.e.
 left hemisphere) advantage that is noted at the population level (see Hugdahl et al., 2009, and Westerhausen, 2019, for a review) is reduced in left-handers (Bethmann, Tempelmann, De Bleser, Scheich, & Brechmann, 2007; Bryden, 1970, 1988; Hugdahl, 2003).
-More recently, Karlsson, Johnstone, and Carey (2019) employed a consonant-vowel version of the dichotic listening task where participants reported which one of six syllables that they heard the clearest.
+More recently, Karlsson, Johnstone, and Carey (2019) employed a consonant-vowel version of the dichotic listening task where participants reported which one of six syllables that they heard most clearly.
 In study 1, 87% of right-handers and 78% of non-right-handers showed a right ear advantage.
@@
 
-For chimeric faces, which typically show a left visual field advantage there again appears to be a greater departure from the typical pattern of laterality for left-handers (e.g.
+For chimeric faces tasks, which typically show a left visual field advantage, there again appears to be a greater departure from the typical pattern of laterality for left-handers (e.g.
 Gilbert & Bakan, 1973; Levy et al., 1983; Rozkowski & Snelbecker, 1982).
-This was also studied in the study by Karlsson et al.
+This was also studied by Karlsson et al.
 (2019).
@@
 
-From the above reviewed studies, it appears that handedness is related to laterality across a range of tasks.
+From the above reviewed studies, it appears that handedness is related to laterality across a range of tasks (for further discussion see Carey & Johnstone, 2014).
 This relationship points towards a general laterality factor which governs lateralisation across tasks and behaviours.
@@
 The principal goal of the current study was to evaluate the reliability and validity of an online behavioural laterality battery.
-Specifically, we examined whether it is possible to observe the usual lateral biases at the population level in tests of language and emotion processing, and whether it is possible to achieve a test-retest reliability of .65 or above using online behaviour research methods.
+Specifically, we examined whether it is possible to observe the usual laterality biases at the population level in tests of language and emotion processing, and whether it is possible to achieve a test-retest reliability of *r*= .65 or above using online behaviour research methods.
 In addition, for tasks where satisfactory reliability was achieved, we considered how many trials were needed to achieve this.
-The tasks included in the battery were a novel rhyme decision visual half-field task, a consonant-vowel dichotic listening tasks, a chimeric faces visual half-field task, and a novel finger tapping task.
+The tasks included in the battery were a novel rhyme decision visual half-field task, a consonant-vowel dichotic listening task, a chimeric faces visual half-field task, and a novel finger tapping task <!--- ZW: Would it be worth having a section on measurement of manual preference?; AP: As this is exploratory I don't want to add too much about this--->.
 These behavioural tasks were completed twice on two different occasions.
 
-Our first set of predictions concerned validity of the tasks as indicators of cerebral lateralisation.
-Based on previous work, we predicted that at the population level: (1) the laterality indices for Rhyme Decision would show a right visual field advantage (indicating left hemisphere dominance for language processing); (2) the laterality indices for the Dichotic Listening task would show a right ear advantage (indicating left hemisphere dominance for language); and (3) the laterality indices for Chimeric Faces would show a left visual field advantage (indicating right hemisphere dominance for face or emotion processing).
-The Finger Tapping task was designed to provide a quantitive measure of degree of handedness, and was expected to show a right-hand advantage in the overall sample.
-
-Our second prediction related to handedness and atypical lateralisation.
+Our first set of predictions (hypothesis 1) concerned validity of the tasks as indicators of cerebral lateralisation.
+Based on previous work, we predicted that at the population level: (1) the laterality indices for rhyme decision would show a RVF advantage (indicating left hemisphere dominance for language processing); (2) the laterality indices for the dichotic listening task would show a right ear advantage (indicating left hemisphere dominance for language prcessing); and (3) the laterality indices for chimeric faces would show a LVF advantage (indicating right hemisphere dominance for face or emotion processing).
+Although exploratory, the finger tapping task was designed to provide a quantitive measure of degree of handedness, and was expected to show a right hand advantage in the overall sample.
+
+Our second prediction (hypothesis 2) related to handedness and atypical lateralisation.
 We predicted that strong left-handers (i.e.
-those with laterality quotient of -90 or less on the Edinburgh Handedness Inventory) would be more likely to show atypical laterality (i.e.
+those with laterality quotients of -90 or less on the Edinburgh Handedness Inventory; Oldfield, 1971) would be more likely to show atypical laterality (i.e.
 laterality in the opposite direction to our population level predictions) than right-handers.
@@
 
-Our final prediction related to the inter-relationships between laterality indices.
+Our final prediction (hypothesis 3) related to the inter-relationships between laterality indices.
 With online testing, there is the opportunity to test inter-relationships between laterality indices with an adequately powered sample using measures of known reliability.
@@
 For instance, van der Haegen and Brysbaert (2018), found that language laterality did not correlate with face laterality in visual half-field tasks.
-However, their study was restricted to left-handers, who may be atypical in this regard, and they noted that lack of association was hard to interpret because it may be due to unreliability of measurement.
-In the preregistration, we predicted that for right-handers, a model (a) where the two language tasks are correlated with each other but not with the chimeric faces will provide better fit to the data than a model (b) in which laterality indices are correlated between the language and chimeric face tasks, or a base model (c) where each task is independently lateralised.
-For left-handers, we anticipated that that model (c) may be a better fit to the data, indicating that laterality is dissociated across the different tasks.
+However, their study was restricted to left-handers, who may be atypical in this regard, and they noted that lack of association was hard to interpret because it may be due to unreliability of the individual measures.
+In the preregistration for our current study, we predicted that for right-handers, a model where the two language tasks are correlated with each other but not with the chimeric faces (will provide better)model a) would be the best fit to the data, compared to either a model in which laterality indices are correlated between the language and chimeric face tasks (model b), or a base model where each task is independently lateralised (model c).
+For left-handers, we anticipated that that model c may be the fit to the data, indicating that laterality is dissociated across the different tasks.
+<!--- ZW: I think you need to explain the rationale for this hypothesis---> Our specfic predictions concerning handedness follow the observations of Woodhead, Bradshaw, Wilson, Thompson, and Bishop (2019) who reported that a single laterality factor was required to explain most the covariance between six language tasks in right-handers.
+In contrast, two factors were required to explain the covariances in left-handers, indicating that language lateralisation is less consistently lateralised in left-handers.
 
@@
 This project was pre-registered on the Open Science Framework (OSF) prior to the start of data collection.
-The registration form along with the analysis scripts and anonymised data can be found on the OSF at https://osf.io/qsxcv/.
+The registration form along with the analysis scripts <!-- And task stimuli?
+Or will tasks be available on Gorilla?---> and anonymised data can be found on the OSF at https://osf.io/qsxcv/.
 
@@
 
-As our ultimate goal is to use these measures in individual differences research, we aimed to optimise the task to achieve observed test-retest reliability of at least .65.
+As our ultimate goal is to use these measures in individual differences research, we aimed to optimise the task to achieve observed test-retest reliability of at least *r*= .65.
 Our initial power analysis showed that to achieve sufficient precision around the estimate of reliability, we would require 300 participants.
@@
 In practice, our final sample exceeded 300, because we discarded participants from individual tasks for failing to meet prespecified criteria (see below).
-In total, 441 native English speakers were initially recruited through Prolific Academic.
+In total, 441 native English speakers were initially recruited through Prolific Academic (www.prolific.co).
 Participants had normal or corrected-to-normal vision and reported no history of neurological disease (e.g.
-head injury, brain tumour, stroke) or significant hearing loss/impairment.
-
-This study was approved by the Medical Sciences Inter-Divisional Research Ethics Committee (reference number: R66638/RE002) and conducted in accordance with the principles of the Declaration of Helsinki.
+head injury, brain tumour, stroke) or significant hearing loss orimpairment.
+
+This study was approved by the University of Oxford's Medical Sciences Inter-Divisional Research Ethics Committee (reference number: R66638/RE002) and conducted in accordance with the principles of the Declaration of Helsinki.
 Participants gave informed consent and received compensation for their participation.
@@
 
-The battery consisted of eight tasks and was administered via Gorilla (Anwyl-Irvine, Massonnié, Flitton, Kirkham, & Evershed, 2020), a cloud-based tool for collecting data in the behavioural sciences.
+The battery consisted of eight tasks and were built and administered using the Gorilla Experiment Builder (www.gorilla.sc), a cloud-based tool for collecting data in the behavioural sciences (Anwyl-Irvine, Massonnié, Flitton, Kirkham, & Evershed, 2020).
 Each task is described below.
@@
 
-The basic demographics questionnaire asked participants to report handedness, years in education, and whether they were bilingual.
-We also asked participants a question about footedness: which foot do you normally use to step up on a ladder/step?
-Participants could answer "left", "right", or "don't know".
+The basic demographics questionnaire asked participants to report age, gender, years in education, and whether they were bilingual.
+We also askedquestions about handedness and footedness.
+For handedness we asked: which is your dominant hand (i.e.
+which hand do you prefer to use for tasks such as writing, cutting, and catching a ball)?
+For footedness we asked: Pwhich foot do you normally use to step up on a ladder/step?
+For both questions, participants could answer "left", "right", or "no preference".
 
@@
 
-The Edinburgh Handedness Inventory (EHI: Oldfield, 1971) was administered to quantify handedness on a continuum.
+The Edinburgh Handedness Inventory (EHI; Oldfield, 1971) was administered to quantify handedness on a continuum.
 For 10 everyday activities, participants indicated their preferred hand use on a 5-step scale (right hand strongly preferred, right hand preferred, no preference, left hand preferred, left hand strongly preferred).
@@
 
-The Test for Advanced Learners of English (Lemhöfer & Broersma, 2012) was used to assess participants’ vocabulary knowledge.
+The Lexical Test for Advanced Learners of English (LexTALE; Lemhöfer & Broersma, 2012) was used to assess participants’ vocabulary knowledge.
 Participants saw 60 letter strings and had to indicate which word they knew.
 Of the 60 strings, 40 were actual English words and 20 were non-words.
-LexTALE scores were corrected the unequal proportion of words and non-word by averaging the percentages corrected for these two item types (e.g., ((number of words correct/40 x 100) + (number of nonwords correct/20 x 100))/2).
+LexTALE scores were corrected for the unequal proportion of words and non-word by averaging the percentages corrected for these two item types (e.g., ((number of words correct/40 x 100) + (number of nonwords correct/20 x 100))/2).
 Accordingly, the LexTALE scores ranged from 0 to 100.
@@
 
-The rhyme decision task was developed for this project as a novel visual half-field technique which requires participants to decide whether a foveally presented letter string rhymed with the name of an image shown in either the left or right hemifield.
+The rhyme decision task was developed for this project as a novel visual half-field technique which requires participants to decide whether a foveally presented letter string rhymed with the name of an image shown in either the LVF or RVF.
 With online presentation, a particular concern is potential for eye movements away from fixation.
-We relied on brief bilateral presentation, which has guaranteed adequate control for eye movements in previous VHF experiments with verbal naming (Beaumont, 1982).
-Studies that directly monitored eye movements in this setting reported failures of fixation on only 0.5% of trials (Geffen, Bradshaw & Nettleton, 1972, but see also Bourne, 2006).
+We relied on brief bilateral presentation, which has guaranteed adequate control for eye movements in previous visual half-field experiments with verbal naming (Beaumont, 1982).
+Studies that directly monitored eye movements in this setting reported failures of fixation on only 0.5% of trials (Geffen, Bradshaw & Nettleton, 1972; but see also Bourne, 2006).
 The design we used ensured that participants had to start the trial fixating centrally in order to read the target word.
-Pictures of familiar objects were presented for rhyme matching to the written word, with the goal of ensuring that the participant would have to covertly name the picture, thereby engaging left-hemisphere speech processing.
+Pictures of familiar objects were presented for rhyme matching to the written word, with the goal of ensuring that the participant would have to covertly name the picture, thereby engaging left hemisphere speech processing.
 
@@
 Each word was 3-4 letters in length (*mean*= 3.8; *SD*= 0.40) and had an average Zipf frequency [log10(frequency per million words) + 3] of 4.4 (*SD*= 0.72) when using the SUBTLEX-UK database (van Heuven, Mandera, Keuleers, & Brysbaert, 2014).
-Each foveal word was paired with an image that corresponded to one of 26 monosyllabic words which had a mean Zipf frequency of 4.5 (*SD*= 0.48).
+Each foveal word was paired with an image that corresponded to one of 26 rhyming monosyllabic words which had a mean Zipf frequency of 4.5 (*SD*= 0.48).
 For example, the word “bite” was paired with the image “kite” and the word “more” was paired with the image “door”.
@@
 Thus, for each of the 13 couplets, four combinations were possible.
-For the pairing of bite-kite and more-door the following stimuli were created: kite – bite – door, door – bite – kite, door – more – kite, kite – more – door.
+For the pairing of bite-kite and more-door the following stimuli were created: kite – bite – door, door – bite – kite, door – more – kite, and kite – more – door.
 In total, there were 52 unique stimuli.
-On a given trial, each image was displayed at 2.5 deg.
-eccentricity of the foveal word (see Figure 1).
+On a given trial, each image was displayed at 2.5 degress eccentricity of the foveal word (see Figure 1).
+
+Word stimuli were presented in black lowercase Tahoma font on a white background.
+Font was scaled to 250% equating to 28pt font.
+The image pairs were presented at approximately 2.5 degrees of visual angle either side of fixation.
+As screen size was likely to vary between participants, the size of the stimuli and eccentricity of images relative to fixation were maintained using Gorilla’s scaling tool.
 
@@
 Then a foveally presented word would appear for 200 ms.
-Word stimuli were presented in black lowercase Tahoma font on a white background.
-Font was scaled to 250% equating to 28pt font.
-The foveally presented word was replaced with bilaterally presented images at approximately 2.5 deg.
-of visual angle either side of fixation.
-The eccentricity of images was maintained using Gorilla’s scaling tool.
 The bilaterally presented images were then presented for 150 ms.
-At stimulus offset, participants indicated whether the image in the LVF or RVF rhymed with the foveally presented word.
-Participants were instructed to press ‘S’ to indicate that the foveally presented word rhyme with the image in the LVF and ‘K’ if it rhymed with the image in the RVF.
+At stimulus offset, participants indicated whether the name of the image in the LVF or RVF rhymed with the foveally presented word.
+Participants were instructed to press ‘S’ if the written word rhymed with the image in the LVF, and ‘K’ if it rhymed with the image in the RVF.
 If participants did not detect a rhyme, they were instructed to press the space bar.
 The response triggered the next trial.
-Accuracy and response time were recorded.
+Accuracy and reaction times (RT) were recorded.
 
@@
 Screening consisted of two tasks.
-The first involved participants judging which of three pure tones is quietest, and is described by Woods, Siegel, Traer, and McDermott (2017) in full detail.
-Sound discrimination is designed to be easy to achieve when using headphones but difficult over loudspeakers due to phase-cancellation.
+The first involved participants judging which of three pure tones was quietest, and is described by Woods, Siegel, Traer, and McDermott (2017) in full detail.
+This sound discrimination task was designed to be easy to achieve when using headphones but difficult over loudspeakers due to phase-cancellation.
 Participants were removed from the study if they failed to correctly discriminate the quietest sound on at least five out of six trials, indicating that they were unlikely to be using headphones.
-The second check involve participants hearing a sound presented in only one of the two channels.
+The second check involved participants hearing a sound presented in only one of the two channels.
 Participants were removed if they failed to correctly identify which channel the sound was played on five out of six trials, indicating that they were not using stereo-headphones.
@@
 Across the four blocks, they responded to 144 stimuli.
-Participants were instructed to wear headphones when completing the dichotic listening task.
-We included two headphone checks to ensure that participants were using stereo-headphones, which we describe further in the General Procedure.
 At the start of each trial, participants saw a fixation cross for 250 ms, after which a stimulus was played.
-After hearing each stimulus, participants were instructed to report the syllable they heard or, if it seemed like they heard two different sounds, the one they heard best or most clearly by clicking a button corresponding to one of the six syllables.
+After hearing each stimulus, participants were instructed to report the syllable they heard or, if it seemed like they heard two different sounds, the one they heard most clearly by clicking a button corresponding to one of the six syllables.
 The response triggered the start of the next trial.
+Accuracy and reaction times (RT) were recorded.
 
@@
 In the example, “happiness” is shown on the left and “sadness” on the right.
-For a participant who is right lateralised, they might be expected to have stronger emotion processing for the left hemiface.
-As such, they will likely report seeing “happiness”.](chim_example.png)
+A participant who is right lateralised for emotional processing, would be more likely to perceive the emotion in the left hemifield and would therefore respond “happiness”.](chim_example.png)
 
@@
 The response triggered the start of the next trial.
+Accuracy and reaction times (RT) were recorded.
 
@@
 
-The Finger Tapping task is a novel behavioural task that we designed as a quantitative index of handedness.
+The finger tapping task is a novel behavioural task that we designed as a quantitative index of handedness.
 Participants were required to either press the ‘W’, ‘R’, ‘V’, and ‘X’ keys with their left hand or the ‘Y’, ‘I’, ‘M’, ‘B’ keys with their right hands in sequence as many times as possible within a 30 second window.
 Participants completed four blocks of the finger tapping task in total.
-Each block contained required participants to complete the task for both their dominant and non-dominant hand.
+Each block required participants to complete the task for both their dominant and non-dominant hand.
+Each key press was recorded along with its timestamp relative to the beginning of the trial.
 
@@
 The current study implemented a test-retest, within-subjects design across two sessions.
+Each session was three to 10 days apart.
 During session 1, participants completed the full battery of tasks.
@@
 
-Out of 441 participants recruited via Prolific, `r nrow(LIs_c)` (`r round(nrow(LIs_c)/441*100, 1)`%) completed both sessions.
-We then removed `r nrow(LIs_no)` participants (`r round(nrow(LIs_no)/nrow(LIs_c)*100, 1)`%) from our final sample as their lexTALE score was below 80, indicating that a participant’s level of vocabulary knowledge was below the cut off for lower advanced proficiency.
+Out of 441 participants recruited, `r nrow(LIs_c)` (`r round(nrow(LIs_c)/441*100, 1)`%) completed both sessions.
+We then removed `r nrow(LIs_no)` participants (`r round(nrow(LIs_no)/nrow(LIs_c)*100, 1)`%) from our final sample as their lexTALE score was below 80, indicating that their level of vocabulary knowledge was below the cut off for lower advanced proficiency.
 This left a sample of `r nrow(LIs)` participants.
@@
 For the chimeric faces task, we removed participants if they scored lower than 75% on trials in which an identical emotion was shown in each hemiface, and for the dichotic listening task, we removed those who scored less than 75% correct on trials where the same sound was played to each ear.
-These data trimming procedures left data for `r nrow(LIs)-nrow(LIs[LIs$exclude== 1 | LIs$exclude== 12,])` participants on the chimeric faces `r nrow(LIs)-nrow(LIs[LIs$exclude== 2 | LIs$exclude== 12,])` participants on the dichotic listening task.
+These data trimming procedures left data for `r nrow(LIs)-nrow(LIs[LIs$exclude== 1 | LIs$exclude== 12,])` participants on the chimeric faces and `r nrow(LIs)-nrow(LIs[LIs$exclude== 2 | LIs$exclude== 12,])` participants on the dichotic listening task.
 An additional 7 cases had to be removed because of a technical error that meant they had less than 80 trials presented.
 
-For the rhyme decision task, we removed participants who scored less than 50% accuracy in either visual field.
+For the rhyme decision task, we removed participants who scored less than 50% accuracy (chance level) in either visual field.
 Furthermore, we removed participants who scored less than 75% accuracy on the catch trials (trials in which there was no rhyme).
 These trimming procedures left data for `r nrow(LIs[LIs$exRDT== 0,])` participants on this task.
-We considered recruiting further participants to make up the numbers, but this seemed uneconomic as we already had more than 300 cases for the other tasks, and we could tell by this point that the test-retest reliability of the rhyme detection task was unsatisfactory; With `r nrow(LIs[LIs$exRDT== 0,])` participants, the test-retest reliability was *r*= `r round(cor(LIs[LIs$exRDT== 0,]$Day1.RDT_RT_LI, LIs[LIs$exRDT== 0,]$Day2.RDT_RT_LI, use = "complete.obs"),2)`, which is smaller than the *r*= 0.65 that we pre-registered as our criterion for a reliable task.
-We plan to evaluate modified versions of this task in future, but decided to stop data collection at this point and analysed the `r nrow(LIs)` participants with data on other tasks.
-
-For the finger-tapping task we added a data-cleaning step that had not been pre-registered, to remove participants where there were more than 20 keypresses across left and right hands that did not correspond to the  intended sequence.
+We considered recruiting further participants to achieve our target sample size of 300 participants, but this seemed uneconomic as we could already see that with close to 300 cases for the other tasks that the confidence intervals around the test-retest relibaility estimate were none overalapping with *r*= .65, and we could tell by this point that the test-retest reliability of the rhyme detection task was unsatisfactory; With `r nrow(LIs[LIs$exRDT== 0,])` participants, the test-retest reliability was *r*= `r round(cor(LIs[LIs$exRDT== 0,]$Day1.RDT_RT_LI, LIs[LIs$exRDT== 0,]$Day2.RDT_RT_LI, use = "complete.obs"),2)`, which is smaller than the *r*= 0.65 that we pre-registered as our criterion for a reliable task.
+We plan to evaluate modified versions of this task in future, but decided to stop data collection at this point and analysed the `r nrow(LIs)` participants with usable data.
+
+For the finger tapping task we added a data-cleaning step that had not been pre-registered, to remove participants where there were more than 20 keypresses across left and right hands that did not correspond to the  intended sequence.
 This could reflect that the individual had forgotten the target sequence, or was not co-operating with instructions, but scrutiny of cases where this occurred suggested that it usually arose because of technical problems.
@@
 
-For the rhyme task, we pre-registered that we would calculate LIs based on log response time (RT) data, as accuracy during piloting was at ceiling.
+For the rhyme task, we pre-registered that we would calculate LIs based on log RT data, as accuracy during piloting was at ceiling.
 To obtain LI_z we ran a *t*-test for each participant where accurate log RTs were the dependent variable and visual field was the independent variable.
 
-For the dichotic listening and chimeric faces, we specified that LI calculated would be based on trials in which participants correctly identified a stimulus.
+For the dichotic listening and chimeric faces, we specified that LI calculations would be based on trials in which participants correctly identified a stimulus.
 The count of correct responses that corresponded to each side was used to generate an accuracy LI in two ways: (1) in the traditional fashion (100 x (Right – Left)/(Right + Left)); (2) as a *z*-statistic for each participant, using z = (pR-.5)/sqrt(pR x pL/n), where pR is the proportion of correct responses in the RVF, pL is proportion of L responses in the LVF, and n is total LVF and RVF responses.
@@
 
-##### Atypical laterality
+##### Handedness and atypical laterality
 
@@
 
-For Rhyme Decision and Chimeric Faces task, participants were categorised into four groups: typical on both; atypical on Rhyme Decision only; atypical on Chimeric Faces only, and atypical on both.
+For rhyme decision and chimeric faces task, participants were categorised into four groups: typical on both; atypical on rhyme decision only; atypical on chimeric faces only, and atypical on both.
 A chi-square test was then used to test for association with the number of people in each of the four groups and the handedness classification from the Edinburgh Handedness Inventory.
-This was then be repeated for the Dichotic Listening and the Chimeric Faces tasks.
+This was then be repeated for the dichotic listening and the chimeric faces tasks.
 A Bonferroni correction was applied to the outcome of each of the two chi-squared tests.
@@
 
-We examined the relationship between LIs on the Rhyme Decision, Dichotic Listening, and Chimeric faces by modelling covariances.
-This allowed us to compare (1) a model where the two language tasks are correlated with each other but not with the chimeric faces, (2) a model where all LIs are correlated, and (3) a base model where the LIs are unrelated.
+We examined the relationship between LIs on the rhyme decision, dichotic listening, and chimeric faces by modelling covariances using a subset of Structural Equation Modeling (SEM) that does not include any latent variables or directional paths.
+In this approach, we constrain particular covariance patterns and report ‘best’ model fit according AIC weights (Wagenmakers & Farrell, 2004).
+Akaike weights are transformations of the AIC values, which can be directly interpreted as conditional probabilities for each model.
+The best fitting model to the data for right- and left-handers is selected as the final model.
+This allowed us to compare three models:
+-	Model (a), where the two language tasks are correlated with each other but not with the chimeric faces
+-	Model (b), where all LIs are correlated, and 
+-	Model (c), a base model where the LIs are unrelated. 
 This was completed separately for right- and left-handers using both session 1 and session 2 data.
@@
 
-Lateral preference characteristics of the sample after exclusions as noted above is shown in Table 1.
+Foot and eye preference characteristics of the sample after exclusions as noted above are shown in Table 1.
 As participants self-reported eye dominance was always consistent between the Miles and Porta tests, we only report data from the Miles test.
@@
 ```{r Tab1,echo=FALSE, comment=FALSE, warning=FALSE}
-flextable(lattab)
+lattab <- flextable(lattab)
+autofit(lattab, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
 ```
@@
 
-### Significance of lateralisation on each task
-
-For a preliminary impression of the data, Figure 3 shows scatterplots indicating scores on left and right sides for each task, with handedness colour-coded.
-For each task, mean for each side is shown as vertical and horizontal dotted lines, and equal performance on the two sides corresponds to the sloping red dotted line: thus an impression of whether the task is lateralised can be obtained by considering whether points cluster around the sloping lines symmetrically.
-
+### The validity of each behavioural task
 ```{r altplotraw,echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
@@
 ```
-
 ```{r dorawplotfunction,echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
@@
 
-![Figure 3.
-Left vs right scores for sessions 1 and 2 for the behavioural laterality tasks.
-The diagonal dotted line shows point of equality: points above this are higher for R than L, and points below are higher for L than R.
-Mean values for R and L are shown as horizontal and vertical dotted lines.
-](all.raw.png)
-
-For the rhyme task, accuracy was high (*M* = `r round(allmeanpcorr,1)`%, *SD* = `r round(allsdpcorr,1)` %),and only correct responses were used for computing mean log RTs (shown transformed back to original ms scale in the plot).
-For dichotic listening and chimeric faces a correct response can be given from either side; it follows that there will be a negative correlation between the two sides reflecting the fact that the number of selections on the left will be inversely related to the number of correct responses on the right.
-For these tasks it is evident by inspection that most responses for dichotic listening cluster above the dotted line, indicating a right ear advantage, whereas most responses on chimeric faces cluster below the dotted line, indicating a left hemi-face advantage.
-Finally, we show the number of correct sequences with each hand on the finger-tapping task.
-Here, a strong impact of handedness is evident on inspection, with left-handers generally more accurate with the left hand, and right-handers more accurate with the right hand.
-
+A preliminary impression of the data are shown in the Online Supplemental Materials.
 The distribution of LIs can be used to quantify departures from symmetry.
 Recall that for all tasks, we computed LIs in the typical way: (100 x (Right – Left)/(Right + Left)), and that because of crossed pathways from hands, ears and eyes to brain, a positive LI indicates left lateralisation and a negative LI indicates right lateralisation at the brain level.
-<!---I flipped this for consistency with Hugdahl and others, as it was clear from emails that they found it confusing to have left-brain as negative--->.
-Table 2 shows these mean LIs, as well a one-sample t-test comparing the mean LI with zero, and results from a Shapiro-Wilks test for departure from normality.
+LIs on the rhynme task were calculated using log RTs for correct responses (accuracy was high; *M* = `r round(allmeanpcorr,1)`%, *SD* = `r round(allsdpcorr,1)` %).
+LIs for the dichotic listening and chieric faces were calculated using the count of correct reponses for sounds/emotions in each ear/visual.
+LIs on the finger tapping task were calculated using the number of sequences generated for each hand.
+Table 2 shows these mean LIs, as well as a one-sample Wilcoxon test comparing the mean LI with zero.
+Wilcoxon tests were chosen as a series of Shapiro-Wilks tests showed departure from normality for a number of tasks.
 It can be seen that all measures were significantly lateralised in the predicted direction at the population level, with left-hemisphere superiority for rhyme judgement, dichotic listening and finger tapping, and right-hemisphere superiority for chimeric faces.
-Most distributions departed significantly from normality.
 
@@
 LItable <- data.frame(matrix(NA,nrow=8,ncol=7))
-colnames(LItable)<-c('Task','N','Mean','SD','One sample t','One sample p','Normality p')
+colnames(LItable)<-c('Task','N','Mean','SD','One sample v','One sample p','Normality p')
 LItable[,1]<-c('Rhyme Decision 1','Rhyme Decision 2','Dichotic 1','Dichotic 2','Chimeric Faces 1','Chimeric Faces 2','Finger Tapping 1','Finger Tapping 2')
@@
   LItable[i,5]<-round(t.test(temp)$statistic,2)
-  pstring <- round(t.test(temp)$p.value,3)
+  pstring <- round(wilcox.test(temp)$p.value,3) # use wilcoxon
   ifelse(pstring<.001,pstring<-'<0.001',pstring<-char(pstring))
@@
 
-Table 2.Mean LIs and results from one-sample t-tests and Shapiro-Wilks tests for departure from normality
+Table 2.Mean LIs and results from one-sample t-tests and Shapiro-Wilks tests for departure from normality for each task at sessions 1 and 2.
 ```{r Tab2,echo=FALSE, comment=FALSE, warning=FALSE}
@@
 ```
+*Note*. LIs are scored such that a positive value equates to a left lateralisation at the brain level
 
@@
 It can be seen that for dichotic listening and chimeric faces, around 45% of individuals have hemispheric dominance in the predicted direction and around 9% have hemispheric dominance in the opposite direction.
-The proportions showing significant laterality are lower for the rhyme decision task (around 17% left-hemisphere dominant, and 5% right-hemisphere dominant), and for finger-tapping, the proportions are only slightly greater than would be expected by chance, although there is a suggestion of increasing laterality on that task across the two sessions.
+The proportions showing significant laterality are lower for the rhyme decision task (around 17% left-hemisphere dominant, and 5% right-hemisphere dominant), and for finger tapping, the proportions are only slightly greater than would be expected by chance, although there is a suggestion of increasing laterality on that task across the two sessions.
 
@@
 
-Table 3.Mean LI_z and results from one-sample t-tests and Shapiro-Wilks tests for departure from normality
+Table 3. Mean LI_z and percentage of participants who were significantly left lateralized (z > 1.96) or right lateralized (z < -1.96) for each task at sessions 1 and 2.
 ```{r Tab3,echo=FALSE, comment=FALSE, warning=FALSE}
@@
 ```
+*Note*. LIs are scored such that a positive value equates to a left lateralisation at the brain level.
 
@@
 
-The scatterplots depicting test-retest reliability are shown in Figure 4, with Spearman test-retest correlations in Table 4.
+The scatterplots depicting test-retest reliability are shown in Figure 3, with Spearman test-retest correlations in Table 4.
 These were based on LI_z values, but are essentially identical to values from unscaled LI values.
@@
 ```
-Note.
-LIs are scored such that a positive value equates to a left lateralisation at the brain level.
-A positive value equates to left lateralisation at the brain level.
-RD= rhyme decision, DL= dichotic listening, CF: chimeric faces, and FT: finger tapping.
-
-![Figure 4.
-Scatter plots for day 1 and day 2 data for LI_z scores on each laterality task: (A) Rhyme decision, (B) Dichotic listening, (C) Chimeric faces, and (D) Finger tapping.
+
+![Figure 3.
+Scatter plots for day 1 and day 2 data for LI_z scores on each laterality task: (A) rhyme decision, (B) dichotic listening, (C) chimeric faces, and (D) Finger tapping.
 Handedness is colour coded with extreme left-handers identified as those who have a score below -90 on the Edinburgh Handedness Inventory.
@@
 
-To determine the smallest number of trials needed to achieve a test-retest reliability of *r*= .65, we recalculated LIs for the dichotic listening and chimeric faces tasks using the range of trials in which different emotions/sounds were presented to each hemiface/ear.
+The dichotic listening and chimeric faces tasks had test-retest reliability above *r*= .75, and so we investigated whether the tasks could be made more efficient by reducing the number of trials while keeping test-retest reliability above our criterion level of *r*= .65.
+Hence, we recalculated LIs for the dichotic listening and chimeric faces tasks using 1 to n trials, where n is the total number of trials on each task.
 The absolute test-retest correlation coefficients are shown in Figure 5.
 A minimum of 60 trials was required to achieve a test-retest correlation of *r*= .65 using the dichotic listening task.
-However, as indicated in Figure 5, even with 100 trials the reliability estimate still occasionally approaches this 0.65 threshold.
-Taking this lower number of trials, it would be able to present only two blocks to achieve this threshold of reliability for the dichotic listening task.
-For the chimeric faces task, a minimum of 25 trials is necessary to achieve a test-retest reliability of *r*= .65.
-Recall that all possible combinations of the four emotions would result in 12 stimuli in which a different emotion is shown in each hemiface and 4 where the same emotion appears in each hemiface.
-Three sets of all possible combinations would then be sufficient to yield a good test-retest reliability here (equating to 1.5 blocks of the chimeric faces in the current study).
-
-![Figure 5.
+For the chimeric faces task, a minimum of 25 trials is necessary to achieve a test-retest reliability of r= .65.
+
+![Figure 4.
 Test-retest correlations for (A) the dichotic listening task and (B) the chimeric faces task when using a variable number of trials to calculate laterality indices.
-The red dashed line marks the test-retest correlation r= .65.
+The red dashed line marks the test-retest correlation *r*= .65.
 The correlations coefficients are shown as absolute values.](all.cor.plot.png)
@@
 
-Table 5. Frequencies of laterality typicality for right-, left-, and extreme left-handers across the Rhyme decision, Dichotic listening, and Chimeric faces tasks.
+Table 5. Frequencies of laterality typicality for extreme left-handers, moderate left-handers and right-handers across the rhyme decision, dichotic listening, and chimeric faces tasks.
 ```{r Table2,echo=FALSE, comment=FALSE, warning=FALSE}
@@
 latcats<-latcats[,c(7,3,6,2,5,1,4)] #now reorder columns to make sense
-colnames(latcats)<-c('.','L lateralised','L%','Not lateralised','N%','R lateralised','R%')
+colnames(latcats)<-c('.','Left N','Left %','Non N','Non %','Right N','Right %')
 
@@
   latcats)
-
 #NB This table needs formatting - has formatted all numbers to 3 decimal places, which is not what we want. Code below achieves this
 
-num_keys <- c("L%","N%","R%")
-int_keys <- c("L lateralised","Not lateralised","R lateralised")
+num_keys <- c("Left %","Non %","Right %")
+int_keys <- c("Left N","Non N","Right N")
 
@@
 ```
+*Note*. N is the number of participants. % is the percentage of participants.
 
@@
 
-It is evident that the Rhyme Judgement task has few individuals who are significantly lateralised in either direction.
+It is evident that the rhyme judgement task has few individuals who are significantly lateralised in either direction.
 We will therefore not consider it further, as our analyses are only justified if the task is an indicator of cerebral lateralisation.
-Both Dichotic Listening and Chimeric Faces, on the other hand, have a high proportion of strongly lateralised individuals.
+Both dichotic listening and chimeric faces, on the other hand, have a high proportion of strongly lateralised individuals.
 The planned analyses, comparing typically lateralised with atypically lateralised individuals, and extreme left vs right-handed individuals, show significant association between extreme left-handedness and atypical laterality on chi square test.
@@
 For chimeric faces, the association was even more extreme.
-`r round(100*tabc[1,1]/sum(tabc[,1]),1)`% of those with Atypical laterality were extreme left-handers, `r round(100*tabc[3,1]/sum(tabc[,1]),1)`% were right-handers, and the remainder less extreme left-handers.
+`r round(100*tabc[1,1]/sum(tabc[,1]),1)`% of those with atypical laterality were extreme left-handers, `r round(100*tabc[3,1]/sum(tabc[,1]),1)`% were right-handers, and the remainder less extreme left-handers.
 For those with typical lateralisation, `r round(100*tabc[1,3]/sum(tabc[,3]),1)`% were extreme left-handers, `r round(100*tabc[3,3]/sum(tabc[,3]),1)`% were right-handers, and the remainder less extreme left-handers.
@@
 
-Our final question was whether the laterality indices from the language and faces laterality tasks correlate?
-Specifically we had predicted that in right-handers the two language tasks would correlate together, but the chimeric faces would form a separate factor, but in left-handers, the language tasks might fractionate.
+Our final question was whether the laterality indices from the language and faces laterality tasks correlated.
+Specifically we had predicted that in right-handers the two language tasks would correlate together, but the chimeric faces would form a separate factor.
+In left-handers, we predicted that the language tasks would fractionate so that none of the three tasks would correlate.
 
@@
 Spearman correlations between the LI_z scores for the different tasks in two Sessions.
-The upper diagonal shows correlations for the 143 right-handers, and the lower diagonal shows values for the 66 left-handers.
+The upper diagonal shows correlations for the 143 right-handers (balck font), and the lower diagonal shows values for the 66 left-handers (red font).
 ```{r tab 5, echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
-myft5<- flextable(allcorr)
+allcorr2 <- cbind(Row.Names = rownames(allcorr), allcorr)
+names(allcorr2)[1] <- "."
+myft5<- flextable(allcorr2)
+
+myft5<-color(myft5, i = c(2,3,4,5, 6), j = "Rhyme_1", color= "red")
+myft5<-color(myft5, i = c(3,4,5, 6), j = "Rhyme_2", color= "red")
+myft5<-color(myft5, i = c(4,5, 6), j = "Dichotic_1", color= "red")
+myft5<-color(myft5, i = c(5, 6), j = "Dichotic_2", color= "red")
+myft5<-color(myft5, i = c(6), j = "Chimeric_1", color= "red")
+
 autofit(myft5, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
@@
 
-In a more formal test we compared (1) a base model where the LIs are unrelated, (2) a model where the two language tasks are correlated with each other but not with the chimeric faces, (3) a model where all LIs are correlated.
-This was done separately for right- and left-handers, using a subset of SEM that does not include any latent variables or directional paths.
-In this approach, we constrain particular covariance patterns and report ‘best’ model fit according AIC weights (Wagenmakers & Farrell, 2004).
-Akaike weights are transformations of the AIC values, which can be directly interpreted as conditional probabilities for each model.
+In a more formal test we compared three models: (a) where the two language tasks are correlated with eachother but not with the chimeric faces (i.e.
+a language laterality factor); (b) all LIs are correlations (i.e.
+a laterality factor); and (c) a base model where the LIs are unrelated (i.e.
+indepent function).
+This was done separately for right- and left-handers, using a subset of SEM where we constrain particular covariance patterns and report ‘best’ model fit according to Akaike weights, where a larger weight indicates better fit.
 The best fitting model to the data for right- and left-handers is selected as the final model.
@@
 
-As shown in Table 7, the Akaike weights favoured the base model for both left-handers and right-handers. This reflected the lack of cross-task correlation for LI_z measures.
+As shown in Table 7, the Akaike weights favoured model c for both left-handers and right-handers, indicating that lateralisation across tasks is largely independent.
+This reflected the lack of cross-task correlation for LI_z measures.
 
@@
 AICdf[,1]<-c('Right-handers','Left-handers')
-colnames(AICdf)<-c('Handedness','Independent','Language factor','Laterality factor')
+colnames(AICdf)<-c('Handedness','Model C','Model A','Model B')
 for (myhand in 1:2) {
@@
  }
-myft6 <- flextable(AICdf)
+myft6 <- flextable(AICdf, col_keys = c("Handedness",
+                                       "Model A",
+                                       "Model B", 
+                                       "Model C"))
 autofit(myft6, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
@@
 To examine the relationship between the finger tapping task and a conventional hand preference battery, we computed Spearman's rank correlation between LIs produced on the finger tapping task and the Edinburgh Handedness Inventory.
-For session 1 data, the relationship between the tasks was moderate, *rs*= `r round(speartest$estimate, 2)`, *95% CI* [`r round(spearCI$conf.int[1], 2)`, `r round(spearCI$conf.int[2], 2)`],  *p*< 0.001.
-For session 2, the correlation was *rs*= `r round(speartest2$estimate, 2)`, *95% CI* [`r round(spearCI2$conf.int[1], 2)`, `r round(spearCI$conf.int[2], 2)`],  *p*< 0.001.
+For session 1 data, the relationship between the tasks was moderate, *r*= `r round(speartest$estimate, 2)`, *95% CI* [`r round(spearCI$conf.int[1], 2)`, `r round(spearCI$conf.int[2], 2)`],  *p*< 0.001.
+For session 2, the correlation was *r*= `r round(speartest2$estimate, 2)`, *95% CI* [`r round(spearCI2$conf.int[1], 2)`, `r round(spearCI$conf.int[2], 2)`],  *p*< 0.001.
 This suggests that the influence of hand preference on the tapping task may increase with practice.
@@
 
-The LexTALE vocabulary measure had been included largely for the purpose of data exclusion, but we had planned to explore correlations between LexTALE score and the laterality index on the Rhyme Decision task.
+The LexTALE vocabulary measure had been included largely for the purpose of data exclusion, but we had planned to explore correlations between LexTALE score and the laterality index on the rhyme decision task.
 However, as the rhyme task was less reliable than other measures, and showed only small asymmetries, we chose to examine the relationship between the LexTALE and the other verbal laterality measure, dichotic listening, instead.
-From session 1 data, after correcting for multiple comparisons, the correlation between the LIs on the dichotic listening and LexTALE score was small and non-significant, *rs*= `r round(DL.lex.est$estimate, 2)`, *95% CI* [`r round(DL.lex$conf.int[1], 2)`, `r round(DL.lex$conf.int[2], 2)`] *p*= `r round(p.adjust(DL.lex.est$p.value[1], n= 2, method= "bonferroni"), 3)`.
-An identical pattern of results was obtained using session 2 data, *rs*= `r round(DL.lex.est1$estimate, 2)`, *95% CI* [`r round(DL.lex1$conf.int[1], 2)`, `r round(DL.lex1$conf.int[2], 2)`] *p*= `r round(p.adjust(DL.lex.est1$p.value[1], n= 2, method= "bonferroni"), 3)`.
+There was no correlation between the LexTALE score and LIs on the dichotic listening at session 1 (*r*= `r round(DL.lex.est$estimate, 2)`, *95% CI* [`r round(DL.lex$conf.int[1], 2)`, `r round(DL.lex$conf.int[2], 2)`] *p*= `r round(p.adjust(DL.lex.est$p.value[1], n= 2, method= "bonferroni"), 3)`) or session 2 (*r*= `r round(DL.lex.est1$estimate, 2)`, *95% CI* [`r round(DL.lex1$conf.int[1], 2)`, `r round(DL.lex1$conf.int[2], 2)`] *p*= `r round(p.adjust(DL.lex.est1$p.value[1], n= 2, method= "bonferroni"), 3)`).
 
@@
 A relatively small proportion of individuals showed significant asymmetry on the rhyme decision task.
-This task also yielded the poorest test-retest reliability, just falling short of our target minimum cutoff of r = .65.
+This task also yielded the poorest test-retest reliability, just falling short of our target minimum cutoff of *r*= .65.
 Compared to early studies (e.g.
@@
 This would, however, be logistically difficult to implement online.
-The other main limitation of online testing is that the VHF method depends critically on presenting lateralised visual stimuli that project to one hemisphere.
+The other main limitation of online testing is that the visual half-field method depends critically on presenting lateralised visual stimuli that project to one hemisphere.
 Although we were able to calibrate screen size at the start of the session, and to request participants to sit a certain distance from the screen, it is plausible this was not sufficient to achieve adequately lateralised presentation.
@@
 Participants showed a clear right-ear advantage at the population level, and the task exceeded our pre-registered reliability cut-off of *r*= .65.
-Indeed, the test-retest correlation of the dichotic listening task, *r* = .78, was identical to that reported by Bless et al.
+Indeed, the test-retest correlation of the dichotic listening task, r = .78, was identical to that reported by Bless et al.
 (2013) using the iDichotic application, and was higher than that reported for the same dichotic listening task when run under lab conditions (Bless et al., 2013).
-We further showed that with as few as 60 trials, the current implementation of the dichotic listening task yielded a test-retest correlation that than *r*= .65, suggesting that those who wish to use this paradigm in the future could use a smaller number of trials and add additional tasks on that are important to their ongoing research questions.
+We further showed that with as few as 60 trials, the current implementation of the dichotic listening task yielded a test-retest correlation higher than *r*= .65, suggesting that those who wish to use this paradigm in the future could use a smaller number of trials and add additional tasks on that are important to their ongoing research questions.
 It would be of interest to extend this line of work to use the modified version of dichotic listening described by Westerhausen and Samuelsen (2020), which had even higher reliability.
-Of course, it is important to screen for adequate and appropriate headphones when devising online measures.
+Of course, it is important to screen for adequate use of stereo headphones when devising online versions of the task.
 
@@
 This too produced a right hand advantage at the population level, and test-retest reliability of the LIs was good, even though the asymmetries were small in absolute terms.
-Furthermore, our exploratory analysis indicated that LIs from this task correlated reasonably well with the Edinburgh handedness inventory.
+Furthermore, our exploratory analysis indicated that LIs from this task correlated reasonably well with the Edinburgh Handedness Inventory.
 Previous studies have reported correlations between hand preference and measures of relative hand skill, but different hand skill measures typically are weakly intercorrelated (Bishop, 1990), and it seems that asymmetries may vary depending on particular aspects of planning and execution of motor coordination that are involved.
@@
 
-In addition to assessing lateralisation of the task using traditional LIs, we also examined the potential to use individual *z*-scores to quantify lateralisation on each task.
-This method of calculating LIs is perhaps more useful for those interested in looking atypical lateralisation, or for cases where there is a need to compare laterality from different tasks on a common scale.
+In addition to assessing lateralisation of the task using traditional LIs, we also examined the potential to use individual z-scores to quantify lateralisation on each task.
+This method of calculating LIs is perhaps more useful for those interested in looking at atypical lateralisation, or for cases where there is a need to compare laterality from different tasks on a common scale.
 Rather than selecting those who fall either side of an arbitrary cut-off, the LI_z score allows researchers to assess whether participants are consistently and statistically lateralised in either a typical or atypical manner.
-The availability of an online measure means that a researcher who is interested in studying the correlates of atypical language lateralisation could assess participants with our battery to identify those with LI_z scores > 1.96 or <1.96 before inviting them into the lab for more detailed in-person testing.
+The availability of an online measure means that a researcher who is interested in studying the correlates of atypical language lateralisation could assess participants with our battery to identify those with LI_z scores > 1.96 or < 1.96 before inviting them into the lab for more detailed in-person testing.
 This has the potential to save time and effort by screening participants to identify a sample that is enriched for rare cases of reversed cerebral lateralisation.
@@
 While the primary purpose of the study was to validate the battery and establish its test-retest reliability, we addressed two theoretically motivated questions related to handedness and atypicality and intercorrelations between the dichotic listening and chimeric faces task.
-In our pre-registered analysis of handedness, we categorised participants into right,-, left-, and extreme left-handers.
+In our pre-registered analysis of handedness, we categorised participants into right-, left-, and extreme left-handers.
 The data are consistent in showing that extreme left-handers showed increased atypicality on both the dichotic listening and chimeric faces task.
 Mazoyer et al.
-(2014) reported a relationship between strength of lateralisation and hand preference., which was largely driven by the existence of a group with strong left hand preference and strong right hemisphere dominance.
-They further claimed that this right hemisphere dominance is almost exclusively present in left-handers.
+(2014) reported a relationship between strength of lateralisation and hand preference., which was largely driven by the existence of a group with strong left hand preference and strong right hemisphere dominance for language.
+They further claimed that this right hemisphere language dominance may be exclusively present in left-handers.
 Nevertheless, although strong left-handers were over-represented in this category, we did also observe cases of right-handers with right language dominance on the dichotic listening task.
@@
 
-A common mechanism that codetermines lateralisation on a range of skills, extending beyond handedness and language, has been a hotly debated topic in the laterality literature.
-Several hypotheses postulate complementary lateralisation for left-hemisphere and right-hemisphere brain asymmetries, predicting that LIs on verbal and nonverbal tasks should be related (Behrmann & Plaut, 2015; Hellige, 1983; Kosslyn, 1987).
-On this view, a person with atypical right-hemisphere language representation should have atypical left-hemisphere representation for nonverbal domains such as face processing.
-However, several studies have report no such association, or only a weak relationship (Bradzakova-Trajkov et al., 2010; Bryden et al.,1983; Rosch et al., 2012; Whitehouse & Bishop, 2009; van der Haegen & Brysbaert, 2018).
+A common mechanism that co-determines lateralisation on a range of skills, extending beyond handedness and language, has been a hotly debated topic in the laterality literature.
+Several hypotheses postulate complementary lateralisation for left hemisphere and right hemisphere brain asymmetries, predicting that LIs on verbal and nonverbal tasks should be related (Behrmann & Plaut, 2015; Hellige, 1983; Kosslyn, 1987).
+On this view, a person with atypical right hemisphere language representation should have atypical left hemisphere representation for nonverbal domains such as face processing.
+However, several studies have reported no such association, or only a weak relationship (Bradzakova-Trajkov et al., 2010; Bryden et al.,1983; Rosch et al., 2012; Whitehouse & Bishop, 2009; van der Haegen & Brysbaert, 2018).
 Data from the current work supports a view in which laterality on each task is independent.
-Not only was our study well powered to examine this relationship, it included a large sample of left-handers who show increased variability in their LIs.
+Not only was our study well powered to examine this relationship, our sample included a large number of left-handers who showed increased variability in their LIs.
 Specifically, we found no association between the LI_z indices for rhyme decision, dichotic listening and chimeric faces.
@@
 
-To conclude, the present study introduced an online behaviour laterality battery that can be used to examine laterality for speech perception, face emotionality, and manual preference with good test-retest reliability.
+To conclude, the present study introduced an online behaviour laterality battery that can be used to examine laterality for speech perception, facial emotion perception, and manual preference with good test-retest reliability.
 The battery of tasks was administered to a large sample of participants with relative ease, and enabled us to examine several questions related to lateralisation.
@@
 
+Carey, D. P., & Johnstone, L. T. (2014). Quantifying cerebral asymmetries for language in dextrals and adextrals with random-effects meta analysis. *Frontiers in Psychology*, *5*, 1128.
+
 Elias, L. J., Bulman-Fleming, M. B., & McManus, I. C. (1999). Visual temporal asymmetries are related to asymmetries in linguistic perception. *Neuropsychologia*, *37*, 1243-1249.
@@
 *Neuropsychologia*, *46*, 3061-3068.
+
+#### Supplemental material
+
+Figure S1 shows scatterplots indicating scores on left and right sides for each task, with handedness colour-coded.
+For each task, mean for each side is shown as vertical and horizontal dotted lines, and equal performance on the two sides corresponds to the sloping red dotted line: thus an impression of whether the task is lateralised can be obtained by considering whether points cluster around the sloping lines symmetrically.
+
+![Figure S1.
+Left side (L) vs right side (R) scores for sessions 1 and 2 for the behavioural laterality tasks.
+The diagonal dotted line shows point of equality: points above this are higher for R than L, and points below are higher for L than R.
+Mean values for R and L are shown as horizontal and vertical dotted lines.
+The colour of the points indicates the participants’ handedness (black = right-handed; red = left-handed).
+](all.raw.png)
+
+A preliminary impression of the data are shown in the Online Supplemental Materials.
+For the rhyme decision task only correct responses were used for computing mean log RTs (shown transformed back to original ms scale in the plot).
+For dichotic listening and chimeric faces a correct response can be given from either side; it follows that there will be a negative correlation between the two sides reflecting the fact that the number of selections on the left will be inversely related to the number of correct responses on the right.
+For these tasks it is evident by inspection that most responses for dichotic listening cluster above the dotted line, indicating a right ear advantage, whereas most responses on chimeric faces cluster below the dotted line, indicating a left hemi-face advantage.
+Finally, we show the number of correct sequences with each hand on the finger tapping task.
+Here, a strong impact of handedness is evident on inspection, with left-handers generally more accurate with the left hand, and right-handers more accurate with the right hand.

### 2020-08-03 pre-registered_analyses.Rmd
zoe's feedback

@@
 # read in data from csv
-LIs <- read.csv("LIs.csv", na.strings = c("NA", ""))
+LIs <- read.csv('https://osf.io/x5jd2/download',stringsAsFactors = F)
 # then we exclude those who we do not have full datasets for 
 LIs <- subset(LIs, exclude== 0)
+LIs %>%
+    summarize(avg = mean(age), n = n(), sd = sd(age))
+# second, show gender count for the whole population
+allsum %>%
+  group_by(gender) %>%
+    summarize(n = n())
+# fourth, show handedness
+allsum %>%
+  group_by(handedness, gender) %>%
+    summarize(n = n())
 ```

### 2020-08-04 online_laterality_reliability.Rmd
comments in file

@@
 For footedness we asked: Pwhich foot do you normally use to step up on a ladder/step?
-For both questions, participants could answer "left", "right", or "no preference".
+For both questions, participants could answer "left", "right", or "no preference".<!---ZW: what options were they given?
+Left / right / ambidextrous?
+A reviewer asked me about this recently--->
 
@@
 
+<!---ZW: This bit confused me.
+
+I thought that the hypothesis was that extreme left handers were more likely to be atypical on any given task, but the test looks at consistency of laterality across pairs of task.
+Isn’t that testing a different hypothesis?
+
+Both are interesting, but I think the first one needs to be tested too.
+E.g., are extreme left handers more likely than right handers to be atypically lateralized on Rhyme Decision / Chimeric Faces / Dichotic Listening?
+--->
+
 ##### Relationship between language and non-language LIs
@@
   LItable[i,4]<-round(sd(temp),2)
-  LItable[i,5]<-round(t.test(temp)$statistic,2)
+  LItable[i,5]<-round(wilcox.test(temp)$statistic,2)
   pstring <- round(wilcox.test(temp)$p.value,3) # use wilcoxon
@@
 For the chimeric faces task, a minimum of 25 trials is necessary to achieve a test-retest reliability of r= .65.
+<!--- I think this section is useful for us, but maybe not really for others. Is it just diluting the main messages of the paper? --->
 
@@
 
+Woodhead Z.
+V.
+J., Bradshaw A.
+R., Wilson A.
+C., Thompson P.
+A.
+& Bishop D.
+V.
+M.
+(2019).
+Testing the unitary theory of language lateralization using functional transcranial Doppler sonography in adults.
+*Royal Society Open Science*, *6*, 181801.
+
 Woods, K. J., Siegel, M. H., Traer, J., & McDermott, J. H. (2017). Headphone screening to facilitate web-based auditory experiments. *Attention, Perception, & Psychophysics*, *79*, 2064-2072.
@@
 
-Figure S1 shows scatterplots indicating scores on left and right sides for each task, with handedness colour-coded.
-For each task, mean for each side is shown as vertical and horizontal dotted lines, and equal performance on the two sides corresponds to the sloping red dotted line: thus an impression of whether the task is lateralised can be obtained by considering whether points cluster around the sloping lines symmetrically.
+Figure S1 shows scatterplots indicating scores on left and right sides for each task, with self-reported handedness (left vs right) colour-coded.
+For each task, the mean for each side is shown as vertical and horizontal dotted lines, and equal performance on the two sides corresponds to the sloping red dotted line: thus an impression of whether the task is lateralised can be obtained by considering whether points cluster around the sloping lines symmetrically.
 
@@
 
-A preliminary impression of the data are shown in the Online Supplemental Materials.
 For the rhyme decision task only correct responses were used for computing mean log RTs (shown transformed back to original ms scale in the plot).

### 2020-08-06 online_laterality_reliability.Rmd
Some changes of wording for final check by DB, after incorporating comments from ZW. Now includes mention of Gerrits et al.

@@
 Contrary to prediction, lateralisation across tasks was not significantly correlated.
-We conclude that three of the tasks, dichotic listening, chimeric faces and finger tapping, show considerable promise for online evaluation of cerebral lateralisation, but the results do not support the notion of a common underlying cause of lateralisation for different cognitive tasks.
+We conclude that three of the tasks, dichotic listening, chimeric faces and finger tapping, show considerable promise for online evaluation of cerebral lateralisation, but the results do not support the notion of a common underlying cause of lateralisation for the different cognitive tasks studied here.
 
@@
 Behavioural laterality tasks have long been used to establish hemispheric dominance for certain cognitive functions.
-Rather consistently, results from visual half-field and dichotic listening paradigms support the view that the left hemisphere plays the dominant role in language processing (Hugdahl, 2000; Ocklenburg & Güntürkün, 2018).
+Results from visual half-field and dichotic listening paradigms support the view that the left hemisphere plays the dominant role in language processing (Hugdahl, 2000; Ocklenburg & Güntürkün, 2018).
 Here we review how each paradigm has been implemented to study language lateralisation and what is known about reliability and validity.
@@
 In Yovel, Tambini, and Brandman’s (2008) study, participants performed a behavioural experiment where they had to recognize centrally presented chimeric faces, which presented different identities in the right and left visual fields.
-It is encouraging to note that the behavioural measure significantly correlated with the fMRILI calculated from on activation in the fusiform face area (*r*= .49).
+It is encouraging to note that the behavioural measure significantly correlated with the fMRI laterality index (LI) calculated from on activation in the fusiform face area (*r*= .49).
 However, the sample size used here was small: with 17 participants, the 95% confidence interval around the observed correlation of .49 ranged from .012 to .786.
@@
 From the above reviewed studies, it appears that handedness is related to laterality across a range of tasks (for further discussion see Carey & Johnstone, 2014).
-This relationship points towards a general laterality factor which governs lateralisation across tasks and behaviours.
-This has been argued in several hypotheses (Behrmann & Plaut, 2015; Cook, 1984; Hellige, 1983; Kosslyn, 1987).
-Common to all hypotheses is the assumption that there would be a correlation between the degree of laterality across complementary tasks.
-While studies do report significant correlations between measures of lateralisation, these are often small (Badzakova-Trajkov, Haberling, Roberts, & Corballis, 2010).
-Furthermore, there are a large number of studies which report no correlation between laterality indices (Bryden, Hecaen, & Deagostini, 1983; Rosch, Bishop, & Badcock, 2012; Whitehouse & Bishop, 2009; van der Haegen & Brysbaert, 2018).
-However, it has not been possible to clearly unpick whether the null results indicate that there are multiple independent lateral biases across tasks, or whether there is a more prosaic explanation in terms of poor reliability or validity of tasks.
+This relationship points towards a general laterality factor which influences lateralisation across tasks and behaviours, as predicted by several theoretical accounts (Behrmann & Plaut, 2015; Cook, 1984; Hellige, 1983; Kosslyn, 1987).
+In a study of left-handers using fMRI with five tasks, Gerrits, Verhelst, and Vingerhoets (2020) found that the just under half the participants showed either standard laterality (left-lateralised for language and praxis, right-lateralised for face memory, line-bisection and vocal emotion) or a complete reversal of this pattern, providing clear evidence against the idea that these functions are lateralised independently.
+It is also clear that association is far from perfect,  (Badzakova-Trajkov, Haberling, Roberts, & Corballis, 2010) and some studies have found no correlation between laterality indices (Bryden, Hecaen, & Deagostini, 1983; Rosch, Bishop, & Badcock, 2012; Whitehouse & Bishop, 2009; van der Haegen & Brysbaert, 2018).
+Our ultimate goal is to unpick whether dissociations between laterality indices are evidence of multiple independent lateral biases across tasks, or whether there is a more prosaic explanation in terms of poor reliability or validity of tasks.
 
@@
 However, their study was restricted to left-handers, who may be atypical in this regard, and they noted that lack of association was hard to interpret because it may be due to unreliability of the individual measures.
-In the preregistration for our current study, we predicted that for right-handers, a model where the two language tasks are correlated with each other but not with the chimeric faces (will provide better)model a) would be the best fit to the data, compared to either a model in which laterality indices are correlated between the language and chimeric face tasks (model b), or a base model where each task is independently lateralised (model c).
-For left-handers, we anticipated that that model c may be the fit to the data, indicating that laterality is dissociated across the different tasks.
-<!--- ZW: I think you need to explain the rationale for this hypothesis---> Our specfic predictions concerning handedness follow the observations of Woodhead, Bradshaw, Wilson, Thompson, and Bishop (2019) who reported that a single laterality factor was required to explain most the covariance between six language tasks in right-handers.
-In contrast, two factors were required to explain the covariances in left-handers, indicating that language lateralisation is less consistently lateralised in left-handers.
+In the preregistration for our current study, we predicted that for right-handers, a model where the two language tasks are correlated with each other but not with the chimeric faces (model A) would be the best fit to the data, compared to either a model in which laterality indices are correlated between the language and chimeric face tasks (model B), or a base model where each task is independently lateralised (model C).
+Given our previous findings using fTCD to study dissociations in language laterality (Woodhead, Bradshaw, Wilson, Thompson, & Bishop, 2019), we further predicted that model fit may differ for right- and left-handers, with more dissociation between tasks in the latter group.
 
@@
 
-The battery consisted of eight tasks and were built and administered using the Gorilla Experiment Builder (www.gorilla.sc), a cloud-based tool for collecting data in the behavioural sciences (Anwyl-Irvine, Massonnié, Flitton, Kirkham, & Evershed, 2020).
+The battery included four measures of background information and four experimental assessments of behaviural laterality that were built and administered using the Gorilla Experiment Builder (www.gorilla.sc), a cloud-based tool for collecting data in the behavioural sciences (Anwyl-Irvine, Massonnié, Flitton, Kirkham, & Evershed, 2020).
 Each task is described below.
@@
 The basic demographics questionnaire asked participants to report age, gender, years in education, and whether they were bilingual.
-We also askedquestions about handedness and footedness.
+We also asked questions about handedness and footedness.
 For handedness we asked: which is your dominant hand (i.e.
 which hand do you prefer to use for tasks such as writing, cutting, and catching a ball)?
-For footedness we asked: Pwhich foot do you normally use to step up on a ladder/step?
-For both questions, participants could answer "left", "right", or "no preference".<!---ZW: what options were they given?
-Left / right / ambidextrous?
-A reviewer asked me about this recently--->
+For footedness we asked: which foot do you normally use to step up on a ladder/step?
+For both questions, participants could answer "left", "right", or "no preference".
 
@@
 
-The Lexical Test for Advanced Learners of English (LexTALE; Lemhöfer & Broersma, 2012) was used to assess participants’ vocabulary knowledge.
-Participants saw 60 letter strings and had to indicate which word they knew.
+The Lexical Test for Advanced Learners of English (LexTALE; Lemhöfer & Broersma, 2012) was used to screen for normal-range language ability.
+Participants saw 60 letter strings and had to indicate which words they knew.
 Of the 60 strings, 40 were actual English words and 20 were non-words.
@@
 
-The rhyme decision task was developed for this project as a novel visual half-field technique which requires participants to decide whether a foveally presented letter string rhymed with the name of an image shown in either the LVF or RVF.
+The rhyme decision task was developed for this project as a novel visual half-field technique designed to engage lateralised language processes.
+On each trial, participants are asked to decide whether a foveally presented letter string rhymed with the name of an image shown in either the LVF or RVF.
 With online presentation, a particular concern is potential for eye movements away from fixation.
-We relied on brief bilateral presentation, which has guaranteed adequate control for eye movements in previous visual half-field experiments with verbal naming (Beaumont, 1982).
+We relied on the fact that participants had to focus initially on the central written word in order to do the task, with pictures subsequently displayed with brief bilateral presentation, which has been shown to give adequate control for eye movements in previous visual half-field experiments with verbal naming (Beaumont, 1982).
 Studies that directly monitored eye movements in this setting reported failures of fixation on only 0.5% of trials (Geffen, Bradshaw & Nettleton, 1972; but see also Bourne, 2006).
-The design we used ensured that participants had to start the trial fixating centrally in order to read the target word.
 Pictures of familiar objects were presented for rhyme matching to the written word, with the goal of ensuring that the participant would have to covertly name the picture, thereby engaging left hemisphere speech processing.
@@
 For each task we adopted the following data cleaning procedures.
-Applying each of these procedures meant that the number of participants left in the final analyses could vary by task.
 First, for all tasks, we removed trials in which response times were shorter than 200 ms.
@@
 
+At the next stage, we excluded participants who failed to meet pre-registered criteria for inclusion.
+Applying each of these procedures meant that the number of participants left in the final analyses could vary by task.
 For the chimeric faces task, we removed participants if they scored lower than 75% on trials in which an identical emotion was shown in each hemiface, and for the dichotic listening task, we removed those who scored less than 75% correct on trials where the same sound was played to each ear.
@@
 These trimming procedures left data for `r nrow(LIs[LIs$exRDT== 0,])` participants on this task.
-We considered recruiting further participants to achieve our target sample size of 300 participants, but this seemed uneconomic as we could already see that with close to 300 cases for the other tasks that the confidence intervals around the test-retest relibaility estimate were none overalapping with *r*= .65, and we could tell by this point that the test-retest reliability of the rhyme detection task was unsatisfactory; With `r nrow(LIs[LIs$exRDT== 0,])` participants, the test-retest reliability was *r*= `r round(cor(LIs[LIs$exRDT== 0,]$Day1.RDT_RT_LI, LIs[LIs$exRDT== 0,]$Day2.RDT_RT_LI, use = "complete.obs"),2)`, which is smaller than the *r*= 0.65 that we pre-registered as our criterion for a reliable task.
-We plan to evaluate modified versions of this task in future, but decided to stop data collection at this point and analysed the `r nrow(LIs)` participants with usable data.
+We considered recruiting further participants to achieve our target sample size of 300 participants, but this seemed uneconomic as we could already see that with close to 300 cases, the test-retest reliability of the rhyme detection task was unsatisfactory; With `r nrow(LIs[LIs$exRDT== 0,])` participants, the test-retest reliability was *r* = `r round(cor(LIs[LIs$exRDT== 0,]$Day1.RDT_RT_LI, LIs[LIs$exRDT== 0,]$Day2.RDT_RT_LI, use = "complete.obs"),2)`, which is smaller than the *r* = 0.65 that we pre-registered as our criterion for a reliable task.
+We plan to evaluate modified versions of this task in future, but decided to stop data collection at this point and so we present here data from those participants with usable data.
 
@@
 
-The final group of `r nrow(LIs)` participants (`r nrow(LIs[LIs$gender== "Female",])` female) had a mean age of `r round(mean(LIs$age, na.rm= TRUE), 1)` years (*SD*= `r round(sd(LIs$age, na.rm= TRUE), 2)` and included `r nrow(LIs[LIs$handedness== "Left",])` left handers (`r nrow(LIs[LIs$handedness== "Left" & LIs$gender== "Female",])` female).
+The final group consisted of `r nrow(LIs)` participants (`r nrow(LIs[LIs$gender== "Female",])` female) who had complete data on at least one of the behavioural tasks.
+They had a mean age of `r round(mean(LIs$age, na.rm= TRUE), 1)` years (*SD*= `r round(sd(LIs$age, na.rm= TRUE), 2)` and included `r nrow(LIs[LIs$handedness== "Left",])` left handers (`r nrow(LIs[LIs$handedness== "Left" & LIs$gender== "Female",])` female).
 `r nrow(LIs[LIs$handedness== "Right" & LIs$bilingual== "Yes",])` right-handers were bilingual and `r nrow(LIs[LIs$handedness== "Left" & LIs$bilingual== "Yes",])` left-handers were bilingual.
@@
 For the rhyme task, we pre-registered that we would calculate LIs based on log RT data, as accuracy during piloting was at ceiling.
-To obtain LI_z we ran a *t*-test for each participant where accurate log RTs were the dependent variable and visual field was the independent variable.
+To obtain LI_z we ran a *t*-test comparing accurate log RTs for left- and right-sided stimuli for each person.
 
@@
 
-To address hypothesis 1 and examine the laterality of each task at the population level, we conducted a series of one-sample *t*-tests for each task using session 1 data.
-
-For the reliability assessment, we checked LIs for normality using a Shapiro-Wilk test.
+To address hypothesis 1 and examine the laterality at the population level, we conducted a series of one-sample *t*-tests to compare mean LIs of the whole sample against zero for each task, using session 1 data.
+
+For the reliability assessment, we first checked LIs for normality using a Shapiro-Wilk test.
 As most LIs were non-normal, we used a series of Spearman’s correlations to examine the correlation of coefficients for session 1 versus session 2.
-For tasks where the test-retest reliability was above .75 with the existing test length, we reanalysed the data using just the first N trials, reducing N until reliability becomes unsatisfactory (below *r*= .65).
+We also conducted an analysis to determine optimal test length.
+For tasks where the test-retest reliability was above .75, we reanalysed the data using just the first N trials, reducing N until reliability became unsatisfactory (below *r*= .65).
 This was of interest in indicating the most efficient way to administer these tasks in future, while retaining good reliability.
@@
 This allowed us to compare three models:
--	Model (a), where the two language tasks are correlated with each other but not with the chimeric faces
--	Model (b), where all LIs are correlated, and 
--	Model (c), a base model where the LIs are unrelated. 
+-	Model A, where the two language tasks are correlated with each other but not with the chimeric faces  
+-	Model B, where all LIs are correlated, and   
+-	Model C, a base model where the LIs are unrelated.   
 This was completed separately for right- and left-handers using both session 1 and session 2 data.
@@
 Foot and eye preference characteristics of the sample after exclusions as noted above are shown in Table 1.
-As participants self-reported eye dominance was always consistent between the Miles and Porta tests, we only report data from the Miles test.
+As participants self-reported eye dominance was always consistent between the Miles and Porta tests, we report data only from the Miles test.
 
@@
 ```
-*Note*. RR: right foot, right eye; RL: right foot, left eye; LR: left foot, right eye; LL: left foot, left eye. 'Other' inlcudes those who responded 'Don't know'.
+*Note*. RR: right foot, right eye; RL: right foot, left eye; LR: left foot, right eye; LL: left foot, left eye. 'Other' includes those who responded 'Don't know'.
 
@@
 
-A preliminary impression of the data are shown in the Online Supplemental Materials.
+Distributions of the raw data are shown in the Online Supplemental Materials.
 The distribution of LIs can be used to quantify departures from symmetry.
 Recall that for all tasks, we computed LIs in the typical way: (100 x (Right – Left)/(Right + Left)), and that because of crossed pathways from hands, ears and eyes to brain, a positive LI indicates left lateralisation and a negative LI indicates right lateralisation at the brain level.
-LIs on the rhynme task were calculated using log RTs for correct responses (accuracy was high; *M* = `r round(allmeanpcorr,1)`%, *SD* = `r round(allsdpcorr,1)` %).
+LIs on the rhy me task were calculated using log RTs for correct responses (accuracy was high; *M* = `r round(allmeanpcorr,1)`%, *SD* = `r round(allsdpcorr,1)` %).
 LIs for the dichotic listening and chieric faces were calculated using the count of correct reponses for sounds/emotions in each ear/visual.
@@
 The proportions showing significant laterality are lower for the rhyme decision task (around 17% left-hemisphere dominant, and 5% right-hemisphere dominant), and for finger tapping, the proportions are only slightly greater than would be expected by chance, although there is a suggestion of increasing laterality on that task across the two sessions.
-
+Note, however, that we had oversampled left-handers for this study, and, as is evident from exploratory analyses presented below, finger-tapping was strongly influenced by handedness.
+<!---this is also very evident from the scatterplot in Fig 3--->
 ```{r make.tab3,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
@@
 ### Optimising the tasks
-
-The dichotic listening and chimeric faces tasks had test-retest reliability above *r*= .75, and so we investigated whether the tasks could be made more efficient by reducing the number of trials while keeping test-retest reliability above our criterion level of *r*= .65.
+<!--- DB: I think the key issue is not so much getting reliability of .65, as how many trials could be dropped without any loss of reliability from whatever high level it got to.
+A bit of rewording to reflect this.
+Also modified in discussion--->
+The dichotic listening and chimeric faces tasks had test-retest reliability above *r* = .75, and so we investigated whether the tasks could be made more efficient by reducing the number of trials while retaining good test-retest reliability.
 Hence, we recalculated LIs for the dichotic listening and chimeric faces tasks using 1 to n trials, where n is the total number of trials on each task.
 The absolute test-retest correlation coefficients are shown in Figure 5.
-A minimum of 60 trials was required to achieve a test-retest correlation of *r*= .65 using the dichotic listening task.
-For the chimeric faces task, a minimum of 25 trials is necessary to achieve a test-retest reliability of r= .65.
-<!--- I think this section is useful for us, but maybe not really for others. Is it just diluting the main messages of the paper? --->
+Peak reliability was reached in around 85 trials with dichotic listening, and a minimum of 60 trials was required to achieve a test-retest correlation of *r* = .65.
+For the chimeric faces task, peak reliability was achieve with around 36 trials, and a minimum of 25 trials was necessary to achieve a test-retest reliability of *r* = .65.
 
@@
 Test-retest correlations for (A) the dichotic listening task and (B) the chimeric faces task when using a variable number of trials to calculate laterality indices.
-The red dashed line marks the test-retest correlation *r*= .65.
+The red dashed line marks the test-retest correlation *r* = .65.
 The correlations coefficients are shown as absolute values.](all.cor.plot.png)
@@
 ```
-*Note*. N is the number of participants. % is the percentage of participants.
+*Note*. N is the number of participants. % is the column percentage of participants.
 
@@
 In a more formal test we compared three models: (a) where the two language tasks are correlated with eachother but not with the chimeric faces (i.e.
-a language laterality factor); (b) all LIs are correlations (i.e.
-a laterality factor); and (c) a base model where the LIs are unrelated (i.e.
-indepent function).
+a language laterality factor); (b) all LIs are correlated (i.e.
+a single laterality factor); and (c) a base model where the LIs are unrelated (i.e.
+independent functions).
 This was done separately for right- and left-handers, using a subset of SEM where we constrain particular covariance patterns and report ‘best’ model fit according to Akaike weights, where a larger weight indicates better fit.
@@
 To examine the relationship between the finger tapping task and a conventional hand preference battery, we computed Spearman's rank correlation between LIs produced on the finger tapping task and the Edinburgh Handedness Inventory.
-For session 1 data, the relationship between the tasks was moderate, *r*= `r round(speartest$estimate, 2)`, *95% CI* [`r round(spearCI$conf.int[1], 2)`, `r round(spearCI$conf.int[2], 2)`],  *p*< 0.001.
-For session 2, the correlation was *r*= `r round(speartest2$estimate, 2)`, *95% CI* [`r round(spearCI2$conf.int[1], 2)`, `r round(spearCI$conf.int[2], 2)`],  *p*< 0.001.
+For session 1 data, the relationship between the tasks was moderate, *r* = `r round(speartest$estimate, 2)`, *95% CI* [`r round(spearCI$conf.int[1], 2)`, `r round(spearCI$conf.int[2], 2)`],  *p* < 0.001.
+For session 2, the correlation was *r* = `r round(speartest2$estimate, 2)`, *95% CI* [`r round(spearCI2$conf.int[1], 2)`, `r round(spearCI$conf.int[2], 2)`],  *p* < 0.001.
 This suggests that the influence of hand preference on the tapping task may increase with practice.
@@
 
-With the advancement of online behaviour research methods, it has become possible to test large samples of participants outside of the lab.
+With the advancement of online behaviour research methods, it has become possible to test large samples of participants outside the laboratory.
 In the current study we introduced an online behavioural laterality battery that was administered to a relatively large sample of right- and left-handers.
@@
 A relatively small proportion of individuals showed significant asymmetry on the rhyme decision task.
-This task also yielded the poorest test-retest reliability, just falling short of our target minimum cutoff of *r*= .65.
+This task also yielded the poorest test-retest reliability, just falling short of our target minimum cutoff of *r* = .65.
 Compared to early studies (e.g.
-Krach et al., 2006) the rhyme decision task used here was far more reliable but, compared to other visual half-field tasks, it was far from optimal.
-Van der Haegen and Brysbaert (2018) have reported that, under lab conditions, visual half-field paradigms can yield test-retest reliabilities of *r*= .77 when using pictures and *r*= .83 when using words.
+Krach et al., 2006) the rhyme decision task used here was more reliable but, compared to other visual half-field tasks, it was far from optimal.
+Van der Haegen and Brysbaert (2018) have reported that, under lab conditions, visual half-field paradigms can yield test-retest reliabilities of *r* = .77 when using pictures and *r* = .83 when using words.
 Our study was well-powered, so the relatively poor task validity and reliability are likely to be due to methodological factors.
@@
 Indeed, the test-retest correlation of the dichotic listening task, r = .78, was identical to that reported by Bless et al.
-(2013) using the iDichotic application, and was higher than that reported for the same dichotic listening task when run under lab conditions (Bless et al., 2013).
-We further showed that with as few as 60 trials, the current implementation of the dichotic listening task yielded a test-retest correlation higher than *r*= .65, suggesting that those who wish to use this paradigm in the future could use a smaller number of trials and add additional tasks on that are important to their ongoing research questions.
+(2013) using the iDichotic application, and numerically higher than that reported for the same dichotic listening task when run under lab conditions (Bless et al., 2013).
+We further showed (see Figure 4) that there was little decline in reliability with 85 trials, and with as few as 60 trials, the current implementation of the dichotic listening task yielded a test-retest correlation higher than *r* = .65, suggesting that those who wish to use this paradigm in the future could use a smaller number of trials.
 It would be of interest to extend this line of work to use the modified version of dichotic listening described by Westerhausen and Samuelsen (2020), which had even higher reliability.
@@
 The task has previously be shown to have high split-half reliability (Levy et al., 1983), we show that it can meet the higher standard of achieving a high test-retest correlation.
-Our subsequent analysis showed that acceptable levels of reliability can be achieved with this task using very few trials: test-retest reliability of *r*= .65 could be found with 25 trials.
+Our subsequent analysis showed that acceptable levels of reliability can be achieved with this task using very few trials: reliability did not suffer when only 35 trials were used, and test-retest reliability of *r*= .65 could be found with 25 trials.
 It seems likely that one reason for the success of the chimeric faces task online is the task’s simplistic nature.
-From the perspective of the participant, it is a simple task of reporting the emotion expressed in a single face, and does not require any explicit comparison.
+From the perspective of the participant, it is a straightforward task of reporting the emotion expressed in a single face, and does not require any explicit comparison.
 
 We also introduced a novel finger tapping task, which we designed to quantify handedness behaviourally.
-This too produced a right hand advantage at the population level, and test-retest reliability of the LIs was good, even though the asymmetries were small in absolute terms.
-Furthermore, our exploratory analysis indicated that LIs from this task correlated reasonably well with the Edinburgh Handedness Inventory.
+This too produced a right hand advantage at the population level, and test-retest reliability of the LIs was good.
+The asymmetries were small in absolute terms, but this in part reflected the high proportion of the left-handers in the sample: our exploratory analysis indicated that LIs from this task correlated reasonably well with the Edinburgh Handedness Inventory (see also Figure 3).
 Previous studies have reported correlations between hand preference and measures of relative hand skill, but different hand skill measures typically are weakly intercorrelated (Bishop, 1990), and it seems that asymmetries may vary depending on particular aspects of planning and execution of motor coordination that are involved.
@@
 
-While the primary purpose of the study was to validate the battery and establish its test-retest reliability, we addressed two theoretically motivated questions related to handedness and atypicality and intercorrelations between the dichotic listening and chimeric faces task.
+As well as validating the battery and establishing its test-retest reliability, we addressed two theoretically motivated questions related to handedness and atypicality and intercorrelations between the dichotic listening and chimeric faces task.
 In our pre-registered analysis of handedness, we categorised participants into right-, left-, and extreme left-handers.
@@
 Mazoyer et al.
-(2014) reported a relationship between strength of lateralisation and hand preference., which was largely driven by the existence of a group with strong left hand preference and strong right hemisphere dominance for language.
+(2014) reported a relationship between strength of lateralisation and hand preference, which was largely driven by the existence of a group with strong left hand preference and strong right hemisphere dominance for language.
 They further claimed that this right hemisphere language dominance may be exclusively present in left-handers.
-Nevertheless, although strong left-handers were over-represented in this category, we did also observe cases of right-handers with right language dominance on the dichotic listening task.
-This raises the interesting question of whether right language dominance on dichotic listening is equivalent to right language dominance on more direct measures of brain lateralisation, such as fMRI.
+Although strong left-handers were over-represented in this category in our data, we did also observe cases of right-handers with right language dominance on the dichotic listening task.
+This raises the question of whether right language dominance on dichotic listening is equivalent to right language dominance on more direct measures of brain lateralisation, such as fMRI.
 
 A common mechanism that co-determines lateralisation on a range of skills, extending beyond handedness and language, has been a hotly debated topic in the laterality literature.
-Several hypotheses postulate complementary lateralisation for left hemisphere and right hemisphere brain asymmetries, predicting that LIs on verbal and nonverbal tasks should be related (Behrmann & Plaut, 2015; Hellige, 1983; Kosslyn, 1987).
+Several researchers have postulated complementary lateralisation for left hemisphere and right hemisphere brain asymmetries, predicting that LIs on verbal and nonverbal tasks should be inversely correlated (Behrmann & Plaut, 2015; Hellige, 1983; Kosslyn, 1987).
 On this view, a person with atypical right hemisphere language representation should have atypical left hemisphere representation for nonverbal domains such as face processing.
@@
 The lack of association between dichotic listening and chimeric faces is, however, much more convincing, given that both tasks show strong lateralisation at the population level and high test-retest reliablity.
-This supports the view that lateralisation is determined by a series of independent biases: a left-hemisphere population bias for language and a right-hemisphere bias for facial expressions of emotion.
+This is consistent with the view that, at least for some functions, lateralisation is determined by independent biases: the left-hemisphere population bias for language is unrelated to the right-hemisphere bias for facial expressions of emotion.
+Given the range of evidence now available, it may be time to abandon hypotheses that predict that different lateralised skills are either wholly interdependent, or wholly independent.
+Evidence such as that from Gerrits et al (2020) confirms that lateralisation of different functions is associated far more than expected by chance, yet cases of dissociation are not uncommon.
+Research in future might explore which functions are associated, and to what extent, rather than making a simple contrast between models that postulate a general answer to explain all of laterality.
 
@@
 We find evidence of increased atypical lateralisation in left- relative to right-handers.
-However, we find no indication that the laterality indices on one task are dependent on another and we conclude that the causes of lateralisation in different domains stem from multiple independent sources.
+However, we find no indication that the laterality indices on one task are dependent on another and we conclude that, for the tasks studied here, the causes of lateralisation appear to have different origins.
 
@@
 Geffen, G., Bradshaw, J. L., & Nettleton, N. C. (1972). Hemispheric asymmetry: verbal and spatial encoding of visual stimuli. *Journal of Experimental Psychology*, *95*, 25-31.
+
+Gerrits, R., Verhelst, H., & Vingerhoets, G.
+(2020).
+Mirrored brain organization: Statistical anomaly or reversal of hemispheric functional segregation bias?
+*Proceedings of the National Academy of Sciences*, *117*, 14057-14065.
+doi:10.1073/pnas.2002981117
 

### 2020-08-07 online_laterality_reliability.Rmd
used for pre-print

@@
 ---
-title: 'Assessing reliability of an online behavioural laterality battery: a pre-registered
-  study.'
+title: 'Assessing the reliability of an online behavioural laterality battery: a pre-registered study.'
 output:
@@
 <!--- updated 3/8/20 to add ZW feedback--->
-Adam J. Parker
+Adam J. Parker*
 
@@
 This has constrained the number of participants recruited in such studies.
-Online testing can potentially allow for much larger sample sizes, but limits the amount of experimental control that is feasible.
-Here we considered whether online testing could give valid and reliable results on four tasks: a rhyme decision visual half-field task, a dichotic listening task, a chimeric faces task, and a finger tapping task to assess manual preference.
+Online testing can allow for much larger sample sizes but limits the amount of experimental control that is feasible.
+Here we considered whether online testing could give valid and reliable results on four tasks: a rhyme decision visual half-field task, a dichotic listening task, a chimeric faces task, and a finger tapping task.
 We recruited 392 participants, oversampling left-handers, who completed the battery twice.
-Three of the tasks showed evidence of both validity and reliability, insofar as they showed hemispheric advantages in the expected direction and test-retest reliability of at least *r*= .65.
-The reliability of the rhyme judgement task was less satisfactory (*r*= .63).
+Three of the tasks showed evidence of both validity and reliability, insofar as they showed hemispheric advantages in the expected direction and test-retest reliability of at least *r*= .75.
+The reliability of the rhyme decision task was less satisfactory (*r*= .63).
 We also found that left-handers were more likely to depart from typical lateralisation.
@@
 \newpage
-## Introduction
-
 Hemispheric specialisation, or lateralisation, is a ubiquitous characteristic of the human brain, yet many questions remain unanswered as to its causes and consequences.
@@
 Without well-powered experiments that use reliable tasks, it becomes difficult to address many research questions regarding the relationship between laterality indices and their stability over time.
-With respect to the relationship between indices, unless a task is reliable, the absence of a correlation cannot be interpreted as evidence for independence of function (Bishop, 2017).
+Concerning the relationship between indices, unless a task is reliable, the absence of a correlation cannot be interpreted as evidence for independence of function (Bishop, 2017).
 Reliability can either be assessed within a test session (e.g.
@@
 Assessment of test-retest reliability, however, requires large samples of people who undergo repeated testing, and this is seldom feasible for in-person testing.
-Online testing allows for recruitment of much larger samples, but raises questions about test validity when opportunities for experimental control are reduced.
+Online testing allows for the recruitment of much larger samples but raises questions about test validity when opportunities for experimental control are reduced.
 Here we introduce an online behavioural laterality battery for which we established the test-retest reliability using a sample of 392 participants.
 The behavioural tasks included were: a novel rhyme decision task, a dichotic listening task (Hugdahl et al., 2009), a chimeric faces task (Burt & Perret, 1997), and a novel finger tapping task.
-Validity of the tasks was assessed in terms of whether lateralisation was observed in the predicted direction, based on prior literature with in-person testing on analogous tasks.
-We also examined two pre-registered questions: (1) do extreme left-handers show a higher rate of atypical laterality than right handers on each behavioural laterality task?
+the validity of the tasks was assessed in terms of whether lateralisation was observed in the predicted direction, based on prior literature with in-person testing on analogous tasks.
+We also examined two pre-registered questions: (1) do extreme left-handers show a higher rate of atypical laterality than right-handers on each behavioural laterality task?
 (2) do the laterality indices from different tasks correlate?
@@
 In this paradigm, visual stimuli are briefly presented to the left or right of a central fixation stimulus.
-The visual pathways from the left and right sides of space to visual cortex are crossed for non-foveal viewing, and so the right visual-field (RVF) projects directly to the left hemisphere, and the left visual-field (LVF) to the right hemisphere.
+The visual pathways from the left and right sides of space to the visual cortex are crossed for non-foveal viewing, and so the right visual-field (RVF) projects directly to the left hemisphere and the left visual-field (LVF) to the right hemisphere.
 The corpus callosum allows for transfer between hemispheres, but the rationale of the method is that there will be a processing advantage for language stimuli that are directly projected to the language-dominant hemisphere.
@@
 The visual half-field task consisted of lexical decisions made towards abstract nouns presented in the LVF and RVF.
-During the experiment, the blood flow to left and right hemispheres of 58 participants was measured.
+During the experiment, the blood flow to the left and right hemispheres of 58 participants was measured.
 In addition, blood flow was measured during a word generation task— the gold standard task for assessing hemispheric dominance for language.
 Despite showing the expected RVF advantage on the visual half-field task, the behavioural LIs produced on the visual half-field paradigm did not correlate with the fTCD LIs produced during word generation (*r*= .18) or the visual half-field task itself (r -.004).
-The lack of validity here likely reflects the poor reliability of the visual field task as split-half reliability was very low (reaction time spilt-half: *r*= .14; accuracy split-half: *r*= .20).
-This illustrates the difficulty of drawing conclusions about associations between laterality indices when reliability of measurement is so low.
+The lack of validity here likely reflects the poor reliability of the visual field task as split-half reliability was very low (reaction time split-half: *r*= .14; accuracy split-half: *r*= .20).
+This illustrates the difficulty of concluding associations between laterality indices when the reliability of measurement is so low.
 
@@
 For example, Hunter and Brysbaert (2008) noted stronger and more stable visual field differences when stimuli were presented bilaterally relative to those where only one stimulus was presented either in the LVF or in the RVF (Boles, 1987, 1990, 1994; see also Iacoboni & Zaidel, 1996).
-Brysbaert and Hunter argued that this is the result of stimuli competing during bilateral presentation, enhancing the advantage to the stimulus presented in the language dominant hemisphere.
+Hunter and Brysbaert argued that this is the result of stimuli competing during bilateral presentation, enhancing the advantage to the stimulus presented in the language dominant hemisphere.
 Reliability and validity also depend on there being a sufficient number of trials and brief stimulus presentation to prevent eye movements.
@@
 Simultaneous presentation of different auditory stimuli to left and right ears for immediate recall was a method originally used in studies of attention, but it became evident that there was usually an advantage for verbal stimuli presented to the right ear (Kimura, 1967).
-Both ears project to both hemispheres, but the weaker ipsilateral pathways appear to be suppressed when the two sides are in competition, and so the right ear, with a contralateral projection to the left hemisphere, will have an advantage in tasks that involve language processing.
+Both ears project to both hemispheres, but the weaker ipsilateral pathways appear to be suppressed when the two sides compete, and so the right ear, with a contralateral projection to the left hemisphere, will have an advantage in tasks that involve language processing.
 
@@
 First, based on a review of the published literature, Westerhausen and Samuelsen outlined eight design features that should be used in an optimal dichotic listening paradigm and explain their justification for each.
-Their recommendations were to use: (1) stop consonant-vowel (CV) syllables as stimulus materials; (2) CV stimulus pairs from the same voicing category; (3) alternating trials of voiced and unvoiced stimulus pairs; (4) free-recall instruction; (5) onw stimulus pair per trial; (6) a paradigm length of 120 dichotic trials; (7) all stimulus pairs are presented to the two sides in identical frequency; and (8) to include binaural (homonymic) trials.
+Their recommendations were to use: (1) stop consonant-vowel (CV) syllables as stimulus materials; (2) CV stimulus pairs from the same voicing category; (3) alternating trials of voiced and unvoiced stimulus pairs; (4) free-recall instruction; (5) one stimulus pair per trial; (6) a paradigm length of 120 dichotic trials; (7) all stimulus pairs are presented to the two sides in identical frequency; and (8) to include binaural (homonymic) trials.
 Putting these criteria into practice, Westerhausen and Samuelsen administered a modified dichotic listening paradigm using both a verbal and manual response format to 50 right-handers.
-Interestingly, the full paradigm yielded a test-retest reliability of *r*= .93 for verbal responses and *r*= .91 for manual responses.
+Interestingly, the full paradigm yielded test-retest reliability of *r*= .93 for verbal responses and *r*= .91 for manual responses.
 Thus, the dichotic listening task provides a promising paradigm for those wanting to investigate language lateralisation without reliance on visual stimuli.
@@
 In studies of face processing, it is possible to use a modification of the visual half-field paradigm in which single stimuli are viewed, composed of vertically split chimeric faces, where each half displays a different emotion or identity.
-Our focus here is on emotion detection in chimeric faces, where a bias is usually found towards reporting the emotion shown in the left hemiface.
-
-The reliability of the chimeric faces tasks seem promising.
-Early work by Levy et al (1983) using two vertically presented chimeric faces, found split-half reliability of *r*= .93.
-The validity of the chimeric faces task has been demonstrated using behavioural and fMRI data, but in a task of recognising facial identity rather than emotion.
-In Yovel, Tambini, and Brandman’s (2008) study, participants performed a behavioural experiment where they had to recognize centrally presented chimeric faces, which presented different identities in the right and left visual fields.
+Our focus here is on emotion detection in chimeric faces, where bias is usually found towards reporting the emotion shown in the left hemiface.
+
+The reliability of the chimeric faces tasks seems promising.
+Early work by Levy et al.
+(1983) using two vertically presented chimeric faces, found split-half reliability of *r*= .93.
+The validity of the chimeric faces task has been demonstrated using behavioural and fMRI data but in a task of recognising facial identity rather than emotion.
+In Yovel, Tambini, and Brandman’s (2008) study, participants performed a behavioural experiment where they had to recognise centrally presented chimeric faces, which showed different identities in the right and left visual fields.
 It is encouraging to note that the behavioural measure significantly correlated with the fMRI laterality index (LI) calculated from on activation in the fusiform face area (*r*= .49).
@@
 Neuroimaging data indicates that for language tasks these differences in lateralisation primarily reflect the contribution of a subset of strong left-handers who are right hemisphere dominant for language, rather than indicating that this relationship operates on a continuum (Mazoyer et al., 2014).
-This association can be used to validate behavioural measures of lateralisation: we would expect to see departure from typical lateralisation in strong left-handers.
+This association can be used to validate behavioural measures of lateralisation: we would expect to see a departure from typical lateralisation in strong left-handers.
 
 With regards to visual half-field tasks for studying language lateralisation, there is evidence to suggest that the visual field advantage varies with handedness (Elias, Bulman-Fleming, & McManus, 1999; Weekes, Capetillo-Cunliffe, Rayman, Iacoboni, & Zaidel, 1999).
-However, a relatively well powered study involving 100 participants, conducted by Willemin et al.
+However, a relatively well-powered study involving 100 participants, conducted by Willemin et al.
 (2016), reported no hint of any effect of handedness, when coded as left and right, or when divided into left (N = 19), mixed (N = 24) and right-handed (N = 57) groups, on the visual field advantage during a lexical decision task.
@@
 For the second chimeric faces task that involved two chimeric faces arranged vertically, a similar pattern emerged.
-73% of right-handers and 57% of non-right-handers showing a leftwards bias.
-
-From the above reviewed studies, it appears that handedness is related to laterality across a range of tasks (for further discussion see Carey & Johnstone, 2014).
-This relationship points towards a general laterality factor which influences lateralisation across tasks and behaviours, as predicted by several theoretical accounts (Behrmann & Plaut, 2015; Cook, 1984; Hellige, 1983; Kosslyn, 1987).
+73% of right-handers and 57% of non-right-handers showed a leftwards bias.
+
+From the above-reviewed studies, it appears that handedness is related to laterality across a range of tasks (for further discussion see Carey & Johnstone, 2014).
+This relationship points towards a general laterality factor which influences lateralisation across tasks and behaviours, as predicted by several theoretical accounts (Behrmann & Plaut, 2015; Hellige, 1993; Kosslyn, 1987).
 In a study of left-handers using fMRI with five tasks, Gerrits, Verhelst, and Vingerhoets (2020) found that the just under half the participants showed either standard laterality (left-lateralised for language and praxis, right-lateralised for face memory, line-bisection and vocal emotion) or a complete reversal of this pattern, providing clear evidence against the idea that these functions are lateralised independently.
-It is also clear that association is far from perfect,  (Badzakova-Trajkov, Haberling, Roberts, & Corballis, 2010) and some studies have found no correlation between laterality indices (Bryden, Hecaen, & Deagostini, 1983; Rosch, Bishop, & Badcock, 2012; Whitehouse & Bishop, 2009; van der Haegen & Brysbaert, 2018).
+It is also clear that this association is far from perfect, (Badzakova-Trajkov, Haberling, Roberts, & Corballis, 2010) and some studies have found no correlation between laterality indices (Bryden, Hecaen, & Deagostini, 1983; Rosch, Bishop, & Badcock, 2012; Whitehouse & Bishop, 2009; van der Haegen & Brysbaert, 2018).
 Our ultimate goal is to unpick whether dissociations between laterality indices are evidence of multiple independent lateral biases across tasks, or whether there is a more prosaic explanation in terms of poor reliability or validity of tasks.
@@
 The principal goal of the current study was to evaluate the reliability and validity of an online behavioural laterality battery.
-Specifically, we examined whether it is possible to observe the usual laterality biases at the population level in tests of language and emotion processing, and whether it is possible to achieve a test-retest reliability of *r*= .65 or above using online behaviour research methods.
+Specifically, we examined whether it is possible to observe the usual laterality biases at the population level in tests of language and emotion processing and whether it is possible to achieve test-retest reliability of *r*= .65 or above using online behaviour research methods.
 In addition, for tasks where satisfactory reliability was achieved, we considered how many trials were needed to achieve this.
-The tasks included in the battery were a novel rhyme decision visual half-field task, a consonant-vowel dichotic listening task, a chimeric faces visual half-field task, and a novel finger tapping task <!--- ZW: Would it be worth having a section on measurement of manual preference?; AP: As this is exploratory I don't want to add too much about this--->.
-These behavioural tasks were completed twice on two different occasions.
-
-Our first set of predictions (hypothesis 1) concerned validity of the tasks as indicators of cerebral lateralisation.
-Based on previous work, we predicted that at the population level: (1) the laterality indices for rhyme decision would show a RVF advantage (indicating left hemisphere dominance for language processing); (2) the laterality indices for the dichotic listening task would show a right ear advantage (indicating left hemisphere dominance for language prcessing); and (3) the laterality indices for chimeric faces would show a LVF advantage (indicating right hemisphere dominance for face or emotion processing).
-Although exploratory, the finger tapping task was designed to provide a quantitive measure of degree of handedness, and was expected to show a right hand advantage in the overall sample.
+The tasks included in the battery were a novel rhyme decision visual half-field task, a consonant-vowel dichotic listening task, a chimeric faces visual half-field task, and a novel finger tapping task.<!--- ZW: Would it be worth having a section on measurement of manual preference?; AP: As this is exploratory I don't want to add too much about this---> These behavioural tasks were completed twice on two different occasions.
+
+Our first set of predictions (hypothesis 1) concerned the validity of the tasks as indicators of cerebral lateralisation.
+Based on previous work, we predicted that at the population level: (1) the laterality indices for rhyme decision would show a RVF advantage (indicating left hemisphere dominance for language processing); (2) the laterality indices for the dichotic listening task would show a right ear advantage (indicating left hemisphere dominance for language processing); and (3) the laterality indices for chimeric faces would show a LVF advantage (indicating right hemisphere dominance for face or emotion processing).
+Although exploratory, the finger tapping task was designed to provide a quantitative measure of the degree of handedness, and was expected to show a right-hand advantage in the overall sample.
 
@@
 those with laterality quotients of -90 or less on the Edinburgh Handedness Inventory; Oldfield, 1971) would be more likely to show atypical laterality (i.e.
-laterality in the opposite direction to our population level predictions) than right-handers.
-This prediction followed from observations by Mazoyer et al (2014, p.
+laterality in the opposite direction to our population-level predictions) than right-handers.
+This prediction followed from observations by Mazoyer et al.
+(2014, p.
 1) who studied a large group of left- and right-handers using fMRI to measure language laterality.
-They concluded “concordance of hemispheric dominance for hand and for language occurs barely above the chance level, except in a group of rare individuals (less than 1% in the general population) who exhibit strong right hemisphere dominance for both language and their preferred hand”.
+They concluded, “concordance of hemispheric dominance for hand and for language occurs barely above the chance level, except in a group of rare individuals (less than 1% in the general population) who exhibit strong right hemisphere dominance for both language and their preferred hand”.
 
@@
 With online testing, there is the opportunity to test inter-relationships between laterality indices with an adequately powered sample using measures of known reliability.
-Prior studies have been inconclusive because it has been difficult gather large amounts of data to establish test-retest reliability.
+Prior studies have been inconclusive because it has been difficult to gather large amounts of data to establish test-retest reliability.
 For instance, van der Haegen and Brysbaert (2018), found that language laterality did not correlate with face laterality in visual half-field tasks.
@@
 
-This project was pre-registered on the Open Science Framework (OSF) prior to the start of data collection.
-The registration form along with the analysis scripts <!-- And task stimuli?
-Or will tasks be available on Gorilla?---> and anonymised data can be found on the OSF at https://osf.io/qsxcv/.
+This project was pre-registered on the Open Science Framework (OSF) before the start of data collection.
+The registration form along with all task materials, analysis scripts and anonymised data can be found on the OSF at https://osf.io/qsxcv/.
+Working versions of the dichotic listening, chimeric faces, and finger tapping tasks are available on Gorilla Open Materials: https://gorilla.sc/openmaterials/104636.
 
@@
 Participants had normal or corrected-to-normal vision and reported no history of neurological disease (e.g.
-head injury, brain tumour, stroke) or significant hearing loss orimpairment.
-
-This study was approved by the University of Oxford's Medical Sciences Inter-Divisional Research Ethics Committee (reference number: R66638/RE002) and conducted in accordance with the principles of the Declaration of Helsinki.
+head injury, brain tumour, stroke) or significant hearing loss or impairment.
+
+This study was approved by the University of Oxford's Medical Sciences Inter-Divisional Research Ethics Committee (reference number: R66638/RE002) and conducted following the principles of the Declaration of Helsinki.
 Participants gave informed consent and received compensation for their participation.
@@
 
-The battery included four measures of background information and four experimental assessments of behaviural laterality that were built and administered using the Gorilla Experiment Builder (www.gorilla.sc), a cloud-based tool for collecting data in the behavioural sciences (Anwyl-Irvine, Massonnié, Flitton, Kirkham, & Evershed, 2020).
+The battery included four measures of background information and four experimental assessments of behavioural  laterality that were built and administered using the Gorilla Experiment Builder (www.gorilla.sc), a cloud-based tool for collecting data in the behavioural sciences (Anwyl-Irvine, Massonnié, Flitton, Kirkham, & Evershed, 2020).
 Each task is described below.
@@
 We also asked questions about handedness and footedness.
-For handedness we asked: which is your dominant hand (i.e.
+For handedness, we asked: which is your dominant hand (i.e.
 which hand do you prefer to use for tasks such as writing, cutting, and catching a ball)?
-For footedness we asked: which foot do you normally use to step up on a ladder/step?
-For both questions, participants could answer "left", "right", or "no preference".
+For footedness, we asked: which foot do you normally use to step up on a ladder/step?
+For both questions, participants could answer "left", "right", or "no preference / Don't know".
 
@@
 The EHI was scored in the standard way, which involves combining information about direction and exclusiveness of hand preference.
-For simple contrast between left- and right-handers, we treated indices greater than 0 as right handed, and indices less than 0 as left handed.
+For simple contrast between left- and right-handers, we treated indices greater than 0 as right-handed, and indices less than 0 as left-handed.
 For analysis of extreme left-handers, we focussed on left-handers with an index of -90 or less, which corresponds to near-exclusive use of the left hand.
@@
 The rhyme decision task was developed for this project as a novel visual half-field technique designed to engage lateralised language processes.
-On each trial, participants are asked to decide whether a foveally presented letter string rhymed with the name of an image shown in either the LVF or RVF.
-With online presentation, a particular concern is potential for eye movements away from fixation.
-We relied on the fact that participants had to focus initially on the central written word in order to do the task, with pictures subsequently displayed with brief bilateral presentation, which has been shown to give adequate control for eye movements in previous visual half-field experiments with verbal naming (Beaumont, 1982).
+On each trial, participants were asked to decide whether a foveally presented letter string rhymed with the name of an image shown in either the LVF or RVF.
+With online presentation, a particular concern is a potential for eye movements away from fixation.
+We relied on the fact that participants had to focus initially on the central written word to do the task, with pictures subsequently displayed with brief bilateral presentation, which has been shown to give adequate control for eye movements in previous visual half-field experiments with verbal naming (Beaumont, 1982).
 Studies that directly monitored eye movements in this setting reported failures of fixation on only 0.5% of trials (Geffen, Bradshaw & Nettleton, 1972; but see also Bourne, 2006).
-Pictures of familiar objects were presented for rhyme matching to the written word, with the goal of ensuring that the participant would have to covertly name the picture, thereby engaging left hemisphere speech processing.
+Pictures of familiar objects were presented for rhyme matching to the written word, to ensure that the participant would have to covertly name the picture, thereby engaging left hemisphere speech processing.
 
@@
 In total, there were 52 unique stimuli.
-On a given trial, each image was displayed at 2.5 degress eccentricity of the foveal word (see Figure 1).
+On a given trial, each image was displayed at 2.5 degrees eccentricity of the foveal word (see Figure 1).
 
@@
 The image pairs were presented at approximately 2.5 degrees of visual angle either side of fixation.
-As screen size was likely to vary between participants, the size of the stimuli and eccentricity of images relative to fixation were maintained using Gorilla’s scaling tool.
+As screen size was likely to vary between participants, the size of the stimuli and eccentricity of images relative to fixation was maintained using Gorilla’s scaling tool.
 
@@
 For example, the word “moon” was displayed with the images for kite and door: kite – moon – door.
-These foil items were included in attempt to prevent participants from preferentially responding to items in either the LVF or RVF.
+These foil items were included in an attempt to prevent participants from preferentially responding to items in either the LVF or RVF.
 These items were then pseudo-randomised such that identical image pairs could not appear one after the other.
@@
 At the start of each trial, participants saw a fixation cross for 250 ms, after which a stimulus was played.
-After hearing each stimulus, participants were instructed to report the syllable they heard or, if it seemed like they heard two different sounds, the one they heard most clearly by clicking a button corresponding to one of the six syllables.
+After hearing each stimulus, participants were instructed to report the syllable they heard or if it seemed like they heard two different sounds, the one they heard most clearly by clicking a button corresponding to one of the six syllables.
 The response triggered the start of the next trial.
@@
 In the example, “happiness” is shown on the left and “sadness” on the right.
-A participant who is right lateralised for emotional processing, would be more likely to perceive the emotion in the left hemifield and would therefore respond “happiness”.](chim_example.png)
+A participant who is right lateralised for emotional processing would be more likely to perceive the emotion in the left hemifield and would, therefore, respond “happiness”.](chim_example.png)
 
@@
 The finger tapping task is a novel behavioural task that we designed as a quantitative index of handedness.
-Participants were required to either press the ‘W’, ‘R’, ‘V’, and ‘X’ keys with their left hand or the ‘Y’, ‘I’, ‘M’, ‘B’ keys with their right hands in sequence as many times as possible within a 30 second window.
+Participants were required to either press the ‘W’, ‘R’, ‘V’, and ‘X’ keys with their left hand or the ‘T’, 'U', ‘M’, ‘B’ keys with their right hands in sequence as many times as possible within a 30-second window.
 Participants completed four blocks of the finger tapping task in total.
 Each block required participants to complete the task for both their dominant and non-dominant hand.
-Each key press was recorded along with its timestamp relative to the beginning of the trial.
+Each keypress was recorded along with its timestamp relative to the beginning of the trial.
 
@@
 Stimuli from each task were repeated across four blocks.
-Within each block participants completed 70 trials of the rhyme detection task (R); 36 trials of the dichotic listening task (D); 32 trials of the chimeric face task (C); and 2 trials of the finger tapping task (F) for both their dominant and non-dominant hand.
-The order of tasks with each block were determined as follows RDCF, DCFR, CFRD, FRDC, and the order of tasks remained consistent across blocks for each participant.
+Within each block, participants completed 70 trials of the rhyme detection task (R); 36 trials of the dichotic listening task (D); 32 trials of the chimeric face task (C); and 2 trials of the finger tapping task (F) for both their dominant and non-dominant hand.
+The order of tasks with each block was determined as follows RDCF, DCFR, CFRD, FRDC, and the order of tasks remained consistent across blocks for each participant.
 For each task, trial order was randomised, except for the finger tapping task where participants responded with their dominant hand followed by their non-dominant hand.
@@
 
-# lextale out df
-LIs_no <- LIs[LIs$lexOut != "keep",]
 # now remove them from LI
@@
 Out of 441 participants recruited, `r nrow(LIs_c)` (`r round(nrow(LIs_c)/441*100, 1)`%) completed both sessions.
-We then removed `r nrow(LIs_no)` participants (`r round(nrow(LIs_no)/nrow(LIs_c)*100, 1)`%) from our final sample as their lexTALE score was below 80, indicating that their level of vocabulary knowledge was below the cut off for lower advanced proficiency.
+From the `r nrow(LIs_c)` participants who completed both sessions, `r nrow(LIs_c)-nrow(LIs)` (`r round((nrow(LIs_c)-nrow(LIs))/nrow(LIs_c)*100, 1)`%) were removed as their lexTALE score was below 80, indicating that their level of vocabulary knowledge was below the cut off for lower advanced proficiency.
 This left a sample of `r nrow(LIs)` participants.
@@
 
-For each task we adopted the following data cleaning procedures.
+For each task, we adopted the following data cleaning procedures.
 First, for all tasks, we removed trials in which response times were shorter than 200 ms.
@@
 
-At the next stage, we excluded participants who failed to meet pre-registered criteria for inclusion.
+At the next stage, we excluded participants who failed to meet the pre-registered criteria for inclusion.
 Applying each of these procedures meant that the number of participants left in the final analyses could vary by task.
@@
 These trimming procedures left data for `r nrow(LIs[LIs$exRDT== 0,])` participants on this task.
-We considered recruiting further participants to achieve our target sample size of 300 participants, but this seemed uneconomic as we could already see that with close to 300 cases, the test-retest reliability of the rhyme detection task was unsatisfactory; With `r nrow(LIs[LIs$exRDT== 0,])` participants, the test-retest reliability was *r* = `r round(cor(LIs[LIs$exRDT== 0,]$Day1.RDT_RT_LI, LIs[LIs$exRDT== 0,]$Day2.RDT_RT_LI, use = "complete.obs"),2)`, which is smaller than the *r* = 0.65 that we pre-registered as our criterion for a reliable task.
-We plan to evaluate modified versions of this task in future, but decided to stop data collection at this point and so we present here data from those participants with usable data.
-
-For the finger tapping task we added a data-cleaning step that had not been pre-registered, to remove participants where there were more than 20 keypresses across left and right hands that did not correspond to the  intended sequence.
+We considered recruiting further participants to achieve our target sample size of 300 participants.
+This seemed uneconomic as we could already see that with close to 300 cases, the test-retest reliability of the rhyme detection task was unsatisfactory.
+With 241 participants, the test-retest reliability was *r* = .63, which is smaller than the *r* = 0.65 that we pre-registered as our criterion for a reliable task.
+We plan to evaluate modified versions of this task in future but decided to stop data collection at this point and so we present here data from those participants with usable data.
+
+For the finger tapping task, we added a data-cleaning step that had not been pre-registered, to remove participants where there were more than 20 keypresses across left and right hands that did not correspond to the intended sequence.
 This could reflect that the individual had forgotten the target sequence, or was not co-operating with instructions, but scrutiny of cases where this occurred suggested that it usually arose because of technical problems.
 For instance, some individuals had repeated recording of the same keypress at short intervals, suggestive of a sticky keyboard.
-The cutoff of 20 was arbitrarily selected, and was successful in excluding cases with abnormally low scores in one session.
+The cut-off of 20 was arbitrarily selected and was successful in excluding cases with abnormally low scores in one session.
 
 The final group consisted of `r nrow(LIs)` participants (`r nrow(LIs[LIs$gender== "Female",])` female) who had complete data on at least one of the behavioural tasks.
-They had a mean age of `r round(mean(LIs$age, na.rm= TRUE), 1)` years (*SD*= `r round(sd(LIs$age, na.rm= TRUE), 2)` and included `r nrow(LIs[LIs$handedness== "Left",])` left handers (`r nrow(LIs[LIs$handedness== "Left" & LIs$gender== "Female",])` female).
+They had a mean age of `r round(mean(LIs$age, na.rm= TRUE), 1)` years (*SD*= `r round(sd(LIs$age, na.rm= TRUE), 2)` and included `r nrow(LIs[LIs$handedness== "Left",])` left-handers (`r nrow(LIs[LIs$handedness== "Left" & LIs$gender== "Female",])` female).
 `r nrow(LIs[LIs$handedness== "Right" & LIs$bilingual== "Yes",])` right-handers were bilingual and `r nrow(LIs[LIs$handedness== "Left" & LIs$bilingual== "Yes",])` left-handers were bilingual.
 
-##### Methods for calcuating laterality indices, reliability, and population level analysis
-
-We preregistered that we would calculate LIs in two ways: (1) in the traditional fashion (100 x (Right – Left)/(Right + Left)); (2) a standardized LI that involved computing a z-score that represented, for each individual, a comparison between left- and right-sided responses.
-In fact, for RT data we specified a t-score, but t and z converge when sample size exceeds 30, and for simplicity we refer to these as LI_z scores.
-These LI_z scores are very highly correlated with the traditional LI measures (correlations typically exceed .95), but can be used to provide a common metric for RT and accuracy measures when identifying individuals who are significantly lateralised.
+##### Methods for calculating laterality indices, reliability, and population-level analysis
+
+We preregistered that we would calculate LIs in two ways: (1) traditionally (100 x (Right – Left)/(Right + Left)); (2) a standardised LI that involved computing a *z*-score that represented, for each individual, a comparison between left- and right-sided responses.
+In fact, for RT data we specified a *t*-score, but *t* and *z* converge when sample size exceeds 30, and for simplicity, we refer to these as LI_z scores.
+These LI_z scores are very highly correlated with the traditional LI measures (correlations typically exceed .95) but can be used to provide a common metric for RT and accuracy measures when identifying significantly lateralised individuals.
 For example, using a conventional 2-sided alpha of .05, we can categorise individuals with an absolute LI_z greater than 1.96 as significantly lateralised.
@@
 For the dichotic listening and chimeric faces, we specified that LI calculations would be based on trials in which participants correctly identified a stimulus.
-The count of correct responses that corresponded to each side was used to generate an accuracy LI in two ways: (1) in the traditional fashion (100 x (Right – Left)/(Right + Left)); (2) as a *z*-statistic for each participant, using z = (pR-.5)/sqrt(pR x pL/n), where pR is the proportion of correct responses in the RVF, pL is proportion of L responses in the LVF, and n is total LVF and RVF responses.
+The count of correct responses that corresponded to each side was used to generate an accuracy LI in two ways: (1) in the traditional fashion (100 x (Right – Left)/(Right + Left)); (2) as a *z*-statistic for each participant, using z = (pR-.5)/sqrt(pR x pL/n), where pR is the proportion of correct responses in the RVF, pL is the proportion of L responses in the LVF, and n is total LVF and RVF responses.
 This method can yield extreme z-scores (for instance, if 83/89 responses are to the right, then z = 15.92), so scores were censored to be in the range -5 to +5.
@@
 As most LIs were non-normal, we used a series of Spearman’s correlations to examine the correlation of coefficients for session 1 versus session 2.
-We also conducted an analysis to determine optimal test length.
-For tasks where the test-retest reliability was above .75, we reanalysed the data using just the first N trials, reducing N until reliability became unsatisfactory (below *r*= .65).
-This was of interest in indicating the most efficient way to administer these tasks in future, while retaining good reliability.
+We also conducted an analysis to determine the optimal test length.
+For tasks where the test-retest reliability was above .75, considered how many trials were needed to achieve satisfactory reliability (i.e.
+*r*>= .65).
+This was accomplished by recalculating LIs using 1 to n trials, where n is the total number of trials on a task.
+This was of interest in indicating the most efficient way to administer these tasks in future while retaining good reliability.
 However, the full data (including all trials) was used for all subsequent analyses.
@@
 A chi-square test was then used to test for association with the number of people in each of the four groups and the handedness classification from the Edinburgh Handedness Inventory.
-This was then be repeated for the dichotic listening and the chimeric faces tasks.
+This was then repeated for the dichotic listening and the chimeric faces tasks.
 A Bonferroni correction was applied to the outcome of each of the two chi-squared tests.
@@
 
-We examined the relationship between LIs on the rhyme decision, dichotic listening, and chimeric faces by modelling covariances using a subset of Structural Equation Modeling (SEM) that does not include any latent variables or directional paths.
-In this approach, we constrain particular covariance patterns and report ‘best’ model fit according AIC weights (Wagenmakers & Farrell, 2004).
+We examined the relationship between LIs on the rhyme decision, dichotic listening, and chimeric faces tasks by modelling covariances using a subset of Structural Equation Modeling (SEM) that does not include any latent variables or directional paths.
+In this approach, we constrain particular covariance patterns and report ‘best’ model fit according to AIC weights (Wagenmakers & Farrell, 2004).
 Akaike weights are transformations of the AIC values, which can be directly interpreted as conditional probabilities for each model.
-The best fitting model to the data for right- and left-handers is selected as the final model.
+The best-fitting model to the data for right- and left-handers is selected as the final model.
 This allowed us to compare three models:
+
 -	Model A, where the two language tasks are correlated with each other but not with the chimeric faces
+
 -	Model B, where all LIs are correlated, and
+
 -	Model C, a base model where the LIs are unrelated.
+
 This was completed separately for right- and left-handers using both session 1 and session 2 data.
@@
 lattab$`L.handers.%`<-round(100*lattab$L.handers.N/sumL,1)
-lattab$`R.handers.%`[6]<-NA
-lattab$`L.handers.%`[6]<-NA
+lattab$`R.handers.%`[6]<-"100"
+lattab$`L.handers.%`[6]<-"100"
 ```
@@
 ```
-*Note*. RR: right foot, right eye; RL: right foot, left eye; LR: left foot, right eye; LL: left foot, left eye. 'Other' includes those who responded 'Don't know'.
+*Note*. RR: right foot, right eye; RL: right foot, left eye; LR: left foot, right eye; LL: left foot, left eye. 'Other' includes those who responded 'No preference / Don't know'.
 
@@
 The distribution of LIs can be used to quantify departures from symmetry.
-Recall that for all tasks, we computed LIs in the typical way: (100 x (Right – Left)/(Right + Left)), and that because of crossed pathways from hands, ears and eyes to brain, a positive LI indicates left lateralisation and a negative LI indicates right lateralisation at the brain level.
-LIs on the rhy me task were calculated using log RTs for correct responses (accuracy was high; *M* = `r round(allmeanpcorr,1)`%, *SD* = `r round(allsdpcorr,1)` %).
-LIs for the dichotic listening and chieric faces were calculated using the count of correct reponses for sounds/emotions in each ear/visual.
+Recall that for all tasks, we computed LIs in a typical way: (100 x (Right – Left)/(Right + Left)), and that because of crossed pathways from hands, ears and eyes to the brain, a positive LI indicates left lateralisation and a negative LI indicates right lateralisation at the brain level.
+LIs on the rhyme task were calculated using log RTs for correct responses (accuracy was high; *M* = `r round(allmeanpcorr,1)`%, *SD* = `r round(allsdpcorr,2)` %).
+LIs for the dichotic listening and chimeric faces were calculated using the count of correct responses for sounds/emotions in each ear/visual.
 LIs on the finger tapping task were calculated using the number of sequences generated for each hand.
 Table 2 shows these mean LIs, as well as a one-sample Wilcoxon test comparing the mean LI with zero.
-Wilcoxon tests were chosen as a series of Shapiro-Wilks tests showed departure from normality for a number of tasks.
-It can be seen that all measures were significantly lateralised in the predicted direction at the population level, with left-hemisphere superiority for rhyme judgement, dichotic listening and finger tapping, and right-hemisphere superiority for chimeric faces.
+Wilcoxon tests were chosen as a series of Shapiro-Wilks tests showed a departure from normality for several tasks.
+It can be seen that all measures were significantly lateralised in the predicted direction at the population level, with left-hemisphere superiority for rhyme decision, dichotic listening and finger tapping, and right-hemisphere superiority for chimeric faces.
 
@@
 ```
-*Note*. LIs are scored such that a positive value equates to a left lateralisation at the brain level
+*Note*. LIs are scored such that a positive value equates to left lateralisation at the brain level.
 
@@
 
-Table 3. Mean LI_z and percentage of participants who were significantly left lateralized (z > 1.96) or right lateralized (z < -1.96) for each task at sessions 1 and 2.
+Table 3. Mean LI_z and percentage of participants who were significantly left lateralised (z > 1.96) or right lateralised (z < -1.96) for each task at sessions 1 and 2.
 ```{r Tab3,echo=FALSE, comment=FALSE, warning=FALSE}
@@
 ```
-*Note*. LIs are scored such that a positive value equates to a left lateralisation at the brain level.
+*Note*. LIs are scored such that a positive value equates to left lateralisation at the brain level.
 
@@
 # Also computes spearmans with CI and returns these from function
-densplots <- function(dat, time1,time2,name1,name2,scattername,y1, y2, x1, x2, ax, ay, remove, title){
+densplots <- function(dat, time1,time2,name1,name2,scattername,y1, y2, x1, x2, ax, ay, remove, title, N){
   # plots
@@
   #Make a label to show the correlation with 95%CI and N on the plot
-  mycorlabel<-paste0('r = ',round(mycor,2),'\n95% CI: ',myci,'\nN = ',length(time1))
+  mycorlabel<-paste0('r = ',round(mycor,2),'\n95% CI: ',myci,'\nN = ',N)
   #Find position in top L corner for legend
@@
 remove= 5
-mycors<-densplots(RD.dat, time1,time2,name1,name2,scattername, y1, y2, x1, x2, ax, ay, remove, "(A)") 
+mycors<-densplots(RD.dat, time1,time2,name1,name2,scattername, y1, y2, x1, x2, ax, ay, remove, "(A)", 241) 
 combotable[1,2]<-round(mycors[[1]],3)
@@
 scattername<-"DichoticScatter"
-mycors<-densplots(DL.dat, time1,time2,name1,name2,scattername,y1, y2, x1, x2, ax, ay, remove,"(B)")
+mycors<-densplots(DL.dat, time1,time2,name1,name2,scattername,y1, y2, x1, x2, ax, ay, remove,"(B)", 290)
 combotable[2,2]<-round(mycors[[1]],3)
@@
 scattername<-"ChimericScatter"
-mycors<-densplots(CF.dat, time1,time2,name1,name2,scattername, y1, y2, x1, x2, ax, ay, remove, "(C)")
+mycors<-densplots(CF.dat, time1,time2,name1,name2,scattername, y1, y2, x1, x2, ax, ay, remove, "(C)", 289)
 combotable[3,2]<-round(mycors[[1]],3)
@@
 scattername<-"FingerScatter"
-mycors<-densplots(FT.dat, time1,time2,name1,name2,scattername, y1, y2, x1, x2, ax, ay, remove, "(D)")
+mycors<-densplots(FT.dat, time1,time2,name1,name2,scattername, y1, y2, x1, x2, ax, ay, remove, "(D)", 288)
 combotable[4,2]<-round(mycors[[1]],3)
@@
 Peak reliability was reached in around 85 trials with dichotic listening, and a minimum of 60 trials was required to achieve a test-retest correlation of *r* = .65.
-For the chimeric faces task, peak reliability was achieve with around 36 trials, and a minimum of 25 trials was necessary to achieve a test-retest reliability of *r* = .65.
+For the chimeric faces task, peak reliability was achieved with around 36 trials, and a minimum of 25 trials was necessary to achieve test-retest reliability of *r* = .65.
 
@@
 We defined 'atypical laterality' as laterality that is in the opposite direction to the majority of participants, with an absolute LI_z value greater than 1.96.
-Although the focus of our planned analyses included only right handers and extreme left handers, we show data for  the remaining left-handers for completeness in Table 5.
+Although the focus of our planned analyses included only right-handers and extreme left-handers, we show data for the remaining left-handers for completeness in Table 5.
 
@@
 #Instead have a new column with these
-latcats$nucol <- c('Rhyme decision','Extreme left-hander','Moderate left-hander','Right-hander','Dichotic listening','Extreme left-hander','Moderate left-hander','Right-hander','Chimeric faces','Extreme left-hander','Moderate left-hander','Right-hander')
+latcats$nucol <- c('Rhyme decision',
+                   'Extreme left-hander',
+                   'Moderate left-hander',
+                   'Right-hander',
+                   'Dichotic listening',
+                   'Extreme left-hander',
+                   'Moderate left-hander',
+                   'Right-hander',
+                   'Chimeric faces',
+                   'Extreme left-hander',
+                   'Moderate left-hander',
+                   'Right-hander')
 
 latcats<-latcats[,c(7,3,6,2,5,1,4)] #now reorder columns to make sense
-colnames(latcats)<-c('.','Left N','Left %','Non N','Non %','Right N','Right %')
+colnames(latcats)<-c(' ','Left N','Left %','Non N','Non %','Right N','Right %')
 
@@
 ```
-*Note*. N is the number of participants. % is the column percentage of participants.
+*Note*. N is the count of participants. % is the percentage of participants.
 
@@
 
-It is evident that the rhyme judgement task has few individuals who are significantly lateralised in either direction.
+The rhyme decision task has few individuals who are significantly lateralised in either direction.
 We will therefore not consider it further, as our analyses are only justified if the task is an indicator of cerebral lateralisation.
 Both dichotic listening and chimeric faces, on the other hand, have a high proportion of strongly lateralised individuals.
-The planned analyses, comparing typically lateralised with atypically lateralised individuals, and extreme left vs right-handed individuals, show significant association between extreme left-handedness and atypical laterality on chi square test.
-For dichotic listening, chi squared = `r round(chi.d$statistic,2)`, d.f.
-= 1, `r pformat(p.adjust(chi.d$p.value, method= "bonferroni", n=2))`; for chimeric faces, chi squared = `r round(chi.c$statistic,2)`, d.f.
+The planned analyses, comparing typically lateralised with atypically lateralised individuals, and extreme left vs right-handed individuals, show a significant association between extreme left-handedness and atypical laterality on chi-square test.
+For dichotic listening, chi-squared = `r round(chi.d$statistic,2)`, d.f.
+= 1, `r pformat(p.adjust(chi.d$p.value, method= "bonferroni", n=2))`; for chimeric faces, chi-squared = `r round(chi.c$statistic,2)`, d.f.
 = 1, `r pformat(p.adjust(chi.c$p.value, method= "bonferroni", n=2))`.
@@
 
-The planned analysis looking at cases who were atypical on both dichotic listening and chimeric faces was hampered by the fact that the two tasks were not associated (see below), and consequently there were only four cases in this category.
+The planned analysis looking at cases who were atypical on both dichotic listening and chimeric faces was hampered by the fact that the two tasks were not associated (see below), and consequently, there were only four cases in this category.
 Nevertheless, it was noteworthy that two of these were extreme left-handers and the other two were less extreme left-handers.
@@
 Our final question was whether the laterality indices from the language and faces laterality tasks correlated.
-Specifically we had predicted that in right-handers the two language tasks would correlate together, but the chimeric faces would form a separate factor.
+Specifically, we had predicted that in right-handers the two language tasks would correlate together, but the chimeric faces would form a separate factor.
 In left-handers, we predicted that the language tasks would fractionate so that none of the three tasks would correlate.
@@
 Spearman correlations between the LI_z scores for the different tasks in two Sessions.
-The upper diagonal shows correlations for the 143 right-handers (balck font), and the lower diagonal shows values for the 66 left-handers (red font).
+The upper diagonal shows correlations for the 143 right-handers (black font), and the values of the lower diagonal show for the 66 left-handers (red font).
 ```{r tab 5, echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
 allcorr2 <- cbind(Row.Names = rownames(allcorr), allcorr)
-names(allcorr2)[1] <- "."
+names(allcorr2)[1] <- " "
+allcorr2[is.na(allcorr2)] <- "."
 myft5<- flextable(allcorr2)
@@
 
-In a more formal test we compared three models: (a) where the two language tasks are correlated with eachother but not with the chimeric faces (i.e.
+In a more formal test, we compared three models: (a) where the two language tasks are correlated with each other but not with the chimeric faces (i.e.
 a language laterality factor); (b) all LIs are correlated (i.e.
@@
 independent functions).
-This was done separately for right- and left-handers, using a subset of SEM where we constrain particular covariance patterns and report ‘best’ model fit according to Akaike weights, where a larger weight indicates better fit.
-The best fitting model to the data for right- and left-handers is selected as the final model.
+This was done separately for right- and left-handers, using a subset of SEM where we constrain particular covariance patterns and report ‘best’ model fit according to Akaike weights, where a larger weight indicates a better fit.
+The best-fitting model to the data for right- and left-handers is selected as the final model.
 
@@
 For session 1 data, the relationship between the tasks was moderate, *r* = `r round(speartest$estimate, 2)`, *95% CI* [`r round(spearCI$conf.int[1], 2)`, `r round(spearCI$conf.int[2], 2)`], *p* < 0.001.
-For session 2, the correlation was *r* = `r round(speartest2$estimate, 2)`, *95% CI* [`r round(spearCI2$conf.int[1], 2)`, `r round(spearCI$conf.int[2], 2)`],  *p* < 0.001.
+For session 2, the correlation was *r* = `r round(speartest2$estimate, 2)`, *95% CI* [`r round(spearCI2$conf.int[1], 2)`, `r round(spearCI2$conf.int[2], 2)`], *p* < 0.001.
 This suggests that the influence of hand preference on the tapping task may increase with practice.
 
-#### Handedness measurements and their association with laterality
+<!---#### Handedness measurements and their association with laterality --->
 
@@
 
-The LexTALE vocabulary measure had been included largely for the purpose of data exclusion, but we had planned to explore correlations between LexTALE score and the laterality index on the rhyme decision task.
+The LexTALE vocabulary measure had been included largely for data exclusion, but we had planned to explore correlations between LexTALE score and the laterality index on the rhyme decision task.
 However, as the rhyme task was less reliable than other measures, and showed only small asymmetries, we chose to examine the relationship between the LexTALE and the other verbal laterality measure, dichotic listening, instead.
@@
 With the advancement of online behaviour research methods, it has become possible to test large samples of participants outside the laboratory.
-In the current study we introduced an online behavioural laterality battery that was administered to a relatively large sample of right- and left-handers.
+In the current study, we introduced an online behavioural laterality battery that was administered to a relatively large sample of right- and left-handers.
 In addition to completing self-report measures of handedness, footedness, and sightedness, participants completed a visual half-field rhyme decision task, a dichotic listening task, a chimeric faces task, and a finger tapping task twice.
@@
 We also examined the association between handedness and atypical lateralisation and examined the intercorrelations between LIs on the dichotic listening and chimeric faces tasks.
-We discuss each of these points in turn.
 
@@
 Our study was well-powered, so the relatively poor task validity and reliability are likely to be due to methodological factors.
-One key difference between van der Haegen and Brysbaert’s and our study was that their language based visual half-field tasks involved overt naming.
+One key difference between van der Haegen and Brysbaert’s and our study was that their language-based visual half-field tasks involved overt naming.
 Speech production is highly lateralised (Kosslyn, 1987) and requiring speech production should yield a hemispheric advantage.
-Our task was designed to require covert speech, but it is possible that an active speech production component might reveal stronger lateralisation.
+Our task was designed to require covert speech, but an active speech production component might reveal stronger lateralisation.
 This would, however, be logistically difficult to implement online.
 The other main limitation of online testing is that the visual half-field method depends critically on presenting lateralised visual stimuli that project to one hemisphere.
-Although we were able to calibrate screen size at the start of the session, and to request participants to sit a certain distance from the screen, it is plausible this was not sufficient to achieve adequately lateralised presentation.
+Although we were able to calibrate screen size at the start of the session and to request participants to sit a certain distance from the screen, it is plausible this was not sufficient to achieve adequately lateralised presentation.
 While we can format images to be consistent across displays, we cannot control participants’ behaviour.
 Outside of the lab, participants can freely move their head which will influence the size of images on the retina, and it is difficult to ensure images are shown in foveal or parafoveal vision.
-One avenue that might be worth exploring is the use of images presented in peripheral vision, as it is unlikely that head movements would result in the image shifting between distinct regions on the retina.
-This increased difficultly in processing of information in this region may also result in stronger visual field advantages (c.f.
-Hunter and Brysbaert, 2008).
+One avenue that might be worth exploring is the use of images presented in peripheral vision, as it is unlikely that small head movements would result in the image shifting from peripheral vision.
+This increased difficultly in the processing of information in this region may also result in stronger visual field advantages.
 
@@
 Like the dichotic listening task, the chimeric faces task showed impressive levels of both validity and reliability, this time with a strong right hemisphere bias at the population level.
-The task has previously be shown to have high split-half reliability (Levy et al., 1983), we show that it can meet the higher standard of achieving a high test-retest correlation.
+The task has previously been shown to have high split-half reliability (Levy et al., 1983), we show that it can meet the higher standard of achieving a high test-retest correlation.
 Our subsequent analysis showed that acceptable levels of reliability can be achieved with this task using very few trials: reliability did not suffer when only 35 trials were used, and test-retest reliability of *r*= .65 could be found with 25 trials.
 It seems likely that one reason for the success of the chimeric faces task online is the task’s simplistic nature.
-From the perspective of the participant, it is a straightforward task of reporting the emotion expressed in a single face, and does not require any explicit comparison.
+From the perspective of the participant, it is a straightforward task of reporting the emotion expressed in a single face and does not require any explicit comparison.
 
 We also introduced a novel finger tapping task, which we designed to quantify handedness behaviourally.
-This too produced a right hand advantage at the population level, and test-retest reliability of the LIs was good.
+This produced a right-hand advantage at the population level, and test-retest reliability of the LIs was good.
 The asymmetries were small in absolute terms, but this in part reflected the high proportion of the left-handers in the sample: our exploratory analysis indicated that LIs from this task correlated reasonably well with the Edinburgh Handedness Inventory (see also Figure 3).
-Previous studies have reported correlations between hand preference and measures of relative hand skill, but different hand skill measures typically are weakly intercorrelated (Bishop, 1990), and it seems that asymmetries may vary depending on particular aspects of planning and execution of motor coordination that are involved.
+Previous studies have reported correlations between hand preference and measures of relative hand skill, but different hand skill measures typically are weakly intercorrelated (Bishop, 1990), and it seems that asymmetries may vary depending on particular aspects of planning and execution of motor coordination that is involved.
 It would be of interest to see how far LIs resemble that from tasks such as the pegboard task (e.g.
 Hodgson, Hirst, & Hudson, 2016), which is widely used in neuropsychological assessment.
-Future work should also examine how other factors such as complexity of key combinations and reach through the midline influence the task.
+Future work should also examine how other factors such as the complexity of key combinations and reach through the midline influence the task.
 
@@
 Mazoyer et al.
-(2014) reported a relationship between strength of lateralisation and hand preference, which was largely driven by the existence of a group with strong left hand preference and strong right hemisphere dominance for language.
+(2014) reported a relationship between the strength of lateralisation and hand preference, which was largely driven by the existence of a group with strong left-hand preference and strong right hemisphere dominance for language.
 They further claimed that this right hemisphere language dominance may be exclusively present in left-handers.
@@
 A common mechanism that co-determines lateralisation on a range of skills, extending beyond handedness and language, has been a hotly debated topic in the laterality literature.
-Several researchers have postulated complementary lateralisation for left hemisphere and right hemisphere brain asymmetries, predicting that LIs on verbal and nonverbal tasks should be inversely correlated (Behrmann & Plaut, 2015; Hellige, 1983; Kosslyn, 1987).
+Several researchers have postulated complementary lateralisation for left hemisphere and right hemisphere brain asymmetries, predicting that LIs on verbal and nonverbal tasks should be inversely correlated (Behrmann & Plaut, 2015; Hellige, 1993; Kosslyn, 1987).
 On this view, a person with atypical right hemisphere language representation should have atypical left hemisphere representation for nonverbal domains such as face processing.
@@
 Specifically, we found no association between the LI_z indices for rhyme decision, dichotic listening and chimeric faces.
-The lack of association between the two language tasks, rhyme decision and dichotic listening, must be interpreted cautiously, given that the rhyme decision task was unsatisfactory in several respects: it had relatively low test-retest reliability, and only weak asymmetry in the sample as a whole.
-It is possible that, as argued by Bishop (2013), there is fractionation between language skills: the dichotic listening task is a purely perceptual task, whereas the rhyme decision task involves implicit speech production.
+The lack of association between the two language tasks, rhyme decision and dichotic listening, must be interpreted cautiously, given that the rhyme decision task was unsatisfactory in several respects: it had relatively low test-retest reliability and only weak asymmetry in the sample as a whole.
+It is possible that, as argued by Bishop (2013), there is fractionation between language skills: the dichotic listening task is purely perceptual, whereas the rhyme decision task involves implicit speech production.
 The current result suggests there could be interest in replicating this comparison of the two tasks using a visual-half-field language task that has better reliability and validity.
-The lack of association between dichotic listening and chimeric faces is, however, much more convincing, given that both tasks show strong lateralisation at the population level and high test-retest reliablity.
+The lack of association between dichotic listening and chimeric faces is, however, much more convincing, given that both tasks show strong lateralisation at the population level and high test-retest reliability.
 This is consistent with the view that, at least for some functions, lateralisation is determined by independent biases: the left-hemisphere population bias for language is unrelated to the right-hemisphere bias for facial expressions of emotion.
-Given the range of evidence now available, it may be time to abandon hypotheses that predict that different lateralised skills are either wholly interdependent, or wholly independent.
+Given the range of evidence now available, it may be time to abandon hypotheses that predict that different lateralised skills are either wholly interdependent or wholly independent.
 Evidence such as that from Gerrits et al (2020) confirms that lateralisation of different functions is associated far more than expected by chance, yet cases of dissociation are not uncommon.
@@
 To conclude, the present study introduced an online behaviour laterality battery that can be used to examine laterality for speech perception, facial emotion perception, and manual preference with good test-retest reliability.
-The battery of tasks was administered to a large sample of participants with relative ease, and enabled us to examine several questions related to lateralisation.
+The battery of tasks was administered to a large sample of participants with relative ease and enabled us to examine several questions related to lateralisation.
 We find evidence of increased atypical lateralisation in left- relative to right-handers.
@@
 ## Data accessibility
-All analysis scripts and anonymised data are avaiable on the Open Science Framework (https://osf.io/qsxcv/).
+This project was pre-registered on the Open Science Framework (OSF) before the start of data collection.
+The registration form along with all task materials, analysis scripts and anonymised data can be found on the OSF at https://osf.io/qsxcv/.
+Working versions of the dichotic listening, chimeric faces, and finger tapping tasks are available on Gorilla Open Materials: https://gorilla.sc/openmaterials/104636.
 
@@
 ## Competing Interests
-We declare we have no competing interestes.
+We declare we have no competing interests.
 
@@
 ## Acknowledgements
-We would like to thank Dr Mike Burt (University of Durham, Uk) for providing stimuli for the chimeric faces task.
+We would like to thank Dr Mike Burt (University of Durham, UK) for providing stimuli for the chimeric faces task.
 We would also like to thank Professor Kenneth Hugdahl (University of Bergen, Norway) for providing stimuli for the dichotic listening task.
@@
 
+Anwyl-Irvine, A. L., Massonnié, J., Flitton, A., Kirkham, N., & Evershed, J. K. (2020). Gorilla in our midst: An online behavioral experiment builder. *Behavior research methods*, *52*, 388-407.
+
 Badzakova-Trajkov, G., Häberling, I. S., Roberts, R. P., & Corballis, M. C. (2010). Cerebral asymmetries: complementary and independent processes. *PloS one*, *5*, e9682.
@@
 
-Bishop DVM. 2013 Cerebral asymmetry and language development: cause, correlate, or consequence? *Science*, *340*, 1230531.
+Bishop, D. V. M. (2013). Cerebral asymmetry and language development: cause, correlate, or consequence? *Science*, *340*, 1230531.
 
@@
 
-Bourne, V. J. (2006). The divided visual field paradigm: Methodological considerations. *Laterality*, *11*, 373–393
+Bourne, V. J. (2006). The divided visual field paradigm: Methodological considerations. *Laterality*, *11*, 373–393.
 
@@
 *Proceedings of the National Academy of Sciences*, *117*, 14057-14065.
-doi:10.1073/pnas.2002981117
 
@@
 
-Hellige, J. B. (1990). Hemispheric asymmetry. *Annual Review of Psychology*, *41*, 55-80.
+Hellige, J. B. (1993). *Hemispheric asymmetry: what’s right and what’s left*. Perspectives in Cognitive Neuroscience. Cambridge, MA: Harvard Univeristy Press.
 
@@
 
-Hunter, Z. R., & Brysbaert, M. (2008). Visual half-field experiments are a good measure of cerebral language dominance if used properly: Evidence from fMRI. *Neuropsychologia*, *46*, 316–325
-
-Iacoboni, M., & Zaidel, E. (1996). Hemispheric independence in word recognition: Evidence from unilateral and bilateral presentations. *Brain and Language*, *53*, 121–140
+Hunter, Z. R., & Brysbaert, M. (2008). Visual half-field experiments are a good measure of cerebral language dominance if used properly: Evidence from fMRI. *Neuropsychologia*, *46*, 316–325.
+
+Iacoboni, M., & Zaidel, E. (1996). Hemispheric independence in word recognition: Evidence from unilateral and bilateral presentations. *Brain and Language*, *53*, 121–140.
+
+Innes, B.
+R., Burt, D.
+M., Birch, Y.
+K., & Hausmann, M.
+(2016).
+A leftward bias however you look at it: Revisiting the emotional chimeric face task as a tool for measuring emotion lateralization.
+*Laterality: Asymmetries of Body, Brain and Cognition*, *21*, 643-661.
 
@@
 
-Krach, S., Chen, L. M., & Hartje, W. (2006). Comparison between visual halffield performance and cerebral blood flow changes as indicators of language dominance. *Laterality*, *11*, 122-140
+Krach, S., Chen, L. M., & Hartje, W. (2006). Comparison between visual halffield performance and cerebral blood flow changes as indicators of language dominance. *Laterality*, *11*, 122-140.
 
@@
 Y., Capetillo-Cunliffe, L., Rayman, J., Iacoboni, M., & Zaidel, E.
-(1999).
+(1996).
 Individual differences in the hemispheric specialization of dual route variables.
@@
 *Laterality: Asymmetries of Body, Brain and Cognition*, *21*, 502-524.
+
+Wagenmakers, E. J., & Farrell, S. (2004). AIC model selection using Akaike weights. *Psychonomic Bulletin & Review*, *11*, 192-196.
 
@@
 
+\newpage
 #### Supplemental material
 
-Figure S1 shows scatterplots indicating scores on left and right sides for each task, with self-reported handedness (left vs right) colour-coded.
+Figure S1 shows scatterplots indicating scores on the left and right sides for each task, with self-reported handedness (left vs right) colour-coded.
 For each task, the mean for each side is shown as vertical and horizontal dotted lines, and equal performance on the two sides corresponds to the sloping red dotted line: thus an impression of whether the task is lateralised can be obtained by considering whether points cluster around the sloping lines symmetrically.
@@
 Left side (L) vs right side (R) scores for sessions 1 and 2 for the behavioural laterality tasks.
-The diagonal dotted line shows point of equality: points above this are higher for R than L, and points below are higher for L than R.
+The diagonal dotted line shows the point of equality: points above this are higher for R than L, and points below are higher for L than R.
 Mean values for R and L are shown as horizontal and vertical dotted lines.
@@
 
-For the rhyme decision task only correct responses were used for computing mean log RTs (shown transformed back to original ms scale in the plot).
-For dichotic listening and chimeric faces a correct response can be given from either side; it follows that there will be a negative correlation between the two sides reflecting the fact that the number of selections on the left will be inversely related to the number of correct responses on the right.
-For these tasks it is evident by inspection that most responses for dichotic listening cluster above the dotted line, indicating a right ear advantage, whereas most responses on chimeric faces cluster below the dotted line, indicating a left hemi-face advantage.
+For the rhyme decision task, only correct responses were used for computing mean log RTs (shown transformed back to the original ms scale in the plot).
+For dichotic listening and chimeric faces, a correct response can be given from either side; it follows that there will be a negative correlation between the two sides reflecting the fact that the number of selections on the left will be inversely related to the number of correct responses on the right.
+For these tasks, it is evident by inspection that most responses for dichotic listening cluster above the dotted line, indicating a right ear advantage, whereas most responses on chimeric faces cluster below the dotted line, indicating a left hemi-face advantage.
 Finally, we show the number of correct sequences with each hand on the finger tapping task.

### 2020-11-08 online_laterality_reliability.Rmd
include some responses to reviewers

@@
 output:
+  html_notebook: default
   word_document: default
-  html_notebook: default
   html_document:
@@
 
+```{r font, echo- FALSE, include= FALSE}
+library(officedown)
+library(officer)
+ft <- fp_text(font.size = 12, bold = FALSE, color = "#C32900", font.family = "Cambria")
+
+colFmt = function(x,color){
+  outputFormat = knitr::opts_knit$get("rmarkdown.pandoc.to")
+  if(outputFormat == 'latex')
+    paste("\\textcolor{",color,"}{",x,"}",sep="")
+  else if(outputFormat == 'html')
+    paste("<font color='",color,"'>",x,"</font>",sep="")
+  else
+    x
+}
+```
+
 <!--- updated 20/7/20 to change 'select' to 'dplyr::select' to avoid conflicts-->
+
 <!--- updated 3/8/20 to add ZW feedback--->
-Adam J. Parker*
-
-Zoe V. J. Woodhead
-
-Paul A. Thompson
-
-Dorothy V. M. Bishop
+\begin{center}
+Adam J. Parker*, Zoe V. J. Woodhead, Paul A. Thompson, Dorothy V. M. Bishop
+\end{center}
 
@@
 Results from visual half-field and dichotic listening paradigms support the view that the left hemisphere plays the dominant role in language processing (Hugdahl, 2000; Ocklenburg & Güntürkün, 2018).
-Here we review how each paradigm has been implemented to study language lateralisation and what is known about reliability and validity.
-
-#### Visual half-field paradigm
-
-In this paradigm, visual stimuli are briefly presented to the left or right of a central fixation stimulus.
+`r colFmt("By contrast, visual half-field paradigms indicate that the right hemisphere is involved in the processing of faces, global feature processing, and visuo-spatial attention (Brederoo, Nieuwenstein, Cornelissen,",'red')` & `r colFmt("Lorist, 2019).
+Below",'red')` we review how each paradigm has been implemented to study language lateralisation `r colFmt("and the detection of emotion in faces", 'red')` and what is known about reliability and validity.
+
+#### Lateralised visual word processing
+
+`r colFmt("Lateralised visual word processing has been studies using the visual half-field paradigm.", 'red')`In this paradigm, visual stimuli are briefly presented to the left or right of a central fixation stimulus.
 The visual pathways from the left and right sides of space to the visual cortex are crossed for non-foveal viewing, and so the right visual-field (RVF) projects directly to the left hemisphere and the left visual-field (LVF) to the right hemisphere.
@@
 Encouragingly, the LIs from these visual half-field tasks also correlated positively with laterality from functional Magnetic Resonance Imaging (fMRI) on a word generation task in 10 left-handers.
-However, the small sample size meant that the confidence intervals around these estimates were very wide: word naming, *r*= .63 (95% CI = .001-.902), picture naming, *r*= .77 (95% CI = .272-.943).
-
-#### Dichotic listening
+However, the small sample size meant that the confidence intervals around these estimates were very wide: word naming, *r*= .63 (95% CI = .001-.902), picture naming, *r*= .77 (95% CI = .272-.943).`r colFmt("More recently, however, Van der Haegen, Cai, Seurinck, and Brysbaert (2011) reported highly similar findings in a somewhat larger sample of 50 participants.
+Again, participants completed behavioural word and picture naming visual half-field tasks and then a word generation task under fMRI.
+It is reassuring that the LIs for the behavioural measures were correlated with the fMRI LI",'red')`: `r colFmt("word naming,",'red')` *r*`r colFmt("= .64, picture naming, ",'red')` *r* `r colFmt("= .65.
+Van der Haegen et al.
+also report a set of analyses involving 250 participants who completed the behavioural visual half-field tasks while their eye movements were monitored.
+Two sets of LIs were calculated",'red')`; `r colFmt("one controlling for fixation error and one that did not.
+Importantly, the LIs which did and did not control for fixation error were highly correlated for word naming, ", 'red')`*r*`r colFmt("= .94, and picture naming, ",'red')`*r*`r colFmt("=.98, indicating that the inclusion of eye movement control adds little additional value when a visual half-field task employs mechanisms to encoruage central fixation and bilateral presentation.",'red')`
+
+`r colFmt("A promising line of work has emerged which further highlights the potenitial for well designed visual half-field tasks to assess language lateralisation.
+First introduced by Willemin et al.
+(2016), the translingual lexical decision task involves participants viewing letter stings either side of fixation and judging whether the left, right, or neither of the letter strings was a valid word.
+With 100 participants completing 256 trials, Willemin et al.
+reported a RVF advatage at the population level.
+Since then this effect has been replicated in a large sample 496 participants cvering six international languages, with evidence of a RVF for each language studied.
+Not only does this task show validity, it seems reliable given Brederoo et al.'s (2020) obervsation that the task yields good split-half reliability when tested with 122 participants,",'red')` *r*`r colFmt("= 0.86 for error rates and", 'red')` *r*`r colFmt("= 0.80 for reaction times.",'red')`
+
+#### Laterlised auditory processing
 
@@
 
-#### Chimeric faces
+`r colFmt("To date, little work has been conducted online with visual stimuli.
+By contrast, the dichotic listening paradigm has been implemented online with huge success.
+Bless et al.
+(2013) developed a mobile device version of the consonant-voewl dichotic listening paradigm, with 36 simultaneous sound pairs.
+Under laboratory conditions, with a sample of 76 (33 Norwegian, 43 Austrialian) participants, they reported a test-retest reliability of *r*= 0.78 when using the mobile app.
+Furthermore, both subgroups of participants showed clear right ear advantages at the population level.
+In a subsequent large-scale study, Bless et al.
+(2015) administered the same application to a sample of over 4,000 participants with more than 60 different language backgrounds.
+Again Bless et al.
+(2015) reported a clear right ear advanatge at the population-level.
+They then used the large dataset to examine cultural/lingusitic effects on the degree on lateralisation; reporting that the degree of lateralization appears to be modulated by linguistic background.
+Together these studies demonstrate the potenital for online laterality research to generate large amounts of data while maintaining a tolerable degree of reliability.",'red')`
+
+#### Lateralised face processing
 
@@
 (2016), reported no hint of any effect of handedness, when coded as left and right, or when divided into left (N = 19), mixed (N = 24) and right-handed (N = 57) groups, on the visual field advantage during a lexical decision task.
-This result illustrates the recurring challenge of interpreting null results in this field: do they reflect a true lack of relationship, poor test reliability, inadequate power to detect small effects, or specific methodological factors?
-We may note that Willemin et al.
-asked participants to respond via manual response, whereas other studies have involved an active speech production component, which could influence task lateralisation.
+This result illustrates the recurring challenge of interpreting null results in this field: do they reflect a true lack of relationship, inadequate power to detect small effects, or specific methodological factors?
+`r colFmt("At current, we can rule out poor reliability as a cause for this null result as Brederoo et al.
+(2020) reported good split-half reliability for this task.",'red')` We may note that Willemin et al.
+asked participants to respond via manual response, whereas other studies have involved an active speech production component, which could influence task lateralisation `r colFmt("and may have contributed to the null effect of handedness.", 'red')`
 
@@
 This relationship points towards a general laterality factor which influences lateralisation across tasks and behaviours, as predicted by several theoretical accounts (Behrmann & Plaut, 2015; Hellige, 1993; Kosslyn, 1987).
+`r colFmt("For example, the causal complementarity principle proposes that the lateralisation of processes involving the same hemisphere will be lateralised to the homologue in the opposite hemisphere.
+With reference to language processing and face emotionality processing, it would be predicted that language processing would be left lateralised and face processing would be right lateralised, and an increase of one co-occurs with an increase of another.
+Indeed, several studies have reported such patterns.
+In support, Brederoo et al.
+(2020) and Dundas, Plaut, and Behrmann (2015) have reported significant correlations between LIs for word and face processing.
+However, that does not mean that all studies provide support for causal complementarity",'red')`.
 In a study of left-handers using fMRI with five tasks, Gerrits, Verhelst, and Vingerhoets (2020) found that the just under half the participants showed either standard laterality (left-lateralised for language and praxis, right-lateralised for face memory, line-bisection and vocal emotion) or a complete reversal of this pattern, providing clear evidence against the idea that these functions are lateralised independently.
-It is also clear that this association is far from perfect, (Badzakova-Trajkov, Haberling, Roberts, & Corballis, 2010) and some studies have found no correlation between laterality indices (Bryden, Hecaen, & Deagostini, 1983; Rosch, Bishop, & Badcock, 2012; Whitehouse & Bishop, 2009; van der Haegen & Brysbaert, 2018).
+It is also clear that this association is far from perfect (Badzakova-Trajkov, Haberling, Roberts, & Corballis, 2010) and some studies have found no correlation between laterality indices (Bryden, Hecaen, & Deagostini, 1983; Rosch, Bishop, & Badcock, 2012; Whitehouse & Bishop, 2009; van der Haegen & Brysbaert, 2018).
 Our ultimate goal is to unpick whether dissociations between laterality indices are evidence of multiple independent lateral biases across tasks, or whether there is a more prosaic explanation in terms of poor reliability or validity of tasks.
+
+`r colFmt("So far the argument has been made that lateralisation across domains do not correlate.
+However, emerging lines of evidence have suggested that even within domains there is independent lateralisation across similar tasks.
+Woodhead, Bradshaw, Wilson, Thompson, and Bishop (2019) reported data for 37 participants engaged in six language tasks under transcranial Doppler sonography (fTCD) and examined the factorial structure of lateralisation across tasks using structural equation modeling.
+They pitted two models against each other.
+The first assumed that language laterality is a unidimensional trait and a single factor was used to model covariance.
+The second assumed that laterality on one set of tasks may vary independently with that one another.
+This model assumed that two factors underlie the pattern of covariance.
+The results supported a two factor model at the population level.
+However, further examination of the data revealed that this was driven by seven left-handers.
+For the majority of right-handed participants a single factor model was preferred.
+Of course this pattern of results may have resulted from the small proportion of left-handers in the current study.
+Woodhead, Thompson, Karlsson, and Bishop (2020) examined this factor structure in a larger sample of 74 participants with 43 right-handers and 31 left-handers.
+The data this time indicated that a two factor model was preferred for both right- and left-handers, with the preference for a two factor model being stronger for left-handers.
+Together, this indicated that not all language tasks show correlated laterality indices.
+Instead, there was evidence that the tasks covaried in two factors with distinct patterns.
+For the current work which implements two measures of llanguage laterality, it may be perfectly plausible that we observe a pattern of results which indicate that language lateralisation is not unidimensional and language tasks would fractionate.",'red')`
 
@@
 In addition, for tasks where satisfactory reliability was achieved, we considered how many trials were needed to achieve this.
-The tasks included in the battery were a novel rhyme decision visual half-field task, a consonant-vowel dichotic listening task, a chimeric faces visual half-field task, and a novel finger tapping task.<!--- ZW: Would it be worth having a section on measurement of manual preference?; AP: As this is exploratory I don't want to add too much about this---> These behavioural tasks were completed twice on two different occasions.
+The tasks included in the battery were a novel rhyme decision visual half-field task, a consonant-vowel dichotic listening task, a chimeric faces visual half-field task, and a novel finger tapping task.
+These behavioural tasks were completed twice on two different occasions.
+`r colFmt("Below we justify our task selection.",'red')`.
+
+`r colFmt("When looking to include studies in our online battery we saw three criteria that a task would have to meet for inclusion.
+First, it would have show good test-retest reliability.
+Second, it would have to show validity and be lateralised at the population level.
+Third, it would have to be sensitive to the subtle differences in laterality that have been reported between right- and non-right-handers.
+Starting with the dichotic listening paradigm, recent endeavors have shown that the task yields reliable results when used online via a mobile app (Bless et al., 2013, 2015).
+Furthermore, Karlsson et al.
+(2019) have shown that this paradigm is sensitive to the subtle differences in language asymmetry between right and non-right handers, making it an obvious choice for inclusion.
+Given that numerous studies looking at the relationship lateralisation on language and non-language tasks have focused on the use of face stimuli (e.g.
+Brederoo et al., 2020), we chose to include a face task to conduct similar comparisons.
+Regarding the chimeric faces, Levy et al.
+(1983) reported good split-half reliability while Karlsson et al.
+had reported that, like the dichotic listening task, the the chimeric faces task is sensitive to the differences in perceptual asymmetry in right- and non-right-handers.
+In addition to these two tasks we looked to include a visual half-field task so that we could further assess language lateralisation.
+While the task reported by Brysbaert and colleagues looked promising, they involve verbal responses which would be difficult to record online with reasonable accuracy.
+We then explored the use of translingual lexical decision task reported by Hausmann and colleagues.
+At the point of pre-registering the study there had been no assessment of the tasks reliability.
+This coupled with the absence of an effect of handedness on reaction times in this task (Willemmin et al., 2016) did not give us faith in using the task.
+Thus we looked to develop our own task in which would required covert speech.",'red')`
+
+`r colFmt("In addition to the previous mentioned tasks, we sought to develop an online measure of manual performance.
+This was largely an exploratory effort.
+Most studies allude to handedness as a unitary phenomenon and classify using a dichotomy of left or right based on participants’ self-reported preference.
+Despite this, some right-handers will use their left hand more than others (Annett, 1998, 2002).
+This type of subtle variation is not captured using dichotomous reporting but is captured within some inventories.
+One such inventory is the Edinburgh handedness inventory (Oldfield, 1971), which requires participants to indicate the strength of hand preference for a range of activities (e.g.
+throwing, striking a match, using scissors).
+Inventory measures are not without limitation.
+It has been argued that respondents may avoid extreme responses on items or solely select extreme responses, thus confounding the measurements (Beaton",'red')` & `r colFmt("Moseley, 1984).
+A final way to assess handedness is to use performance measures, such as the Purdue Pegboard task (Tiffin",'red')` & `r colFmt("Asher, 1948).
+Performance measures can detect subtle differences between each hand which can be placed on a continuum while correlating well with measures of handedness (e.g.
+Steenhous",'red')` & `r colFmt("Bryden, 1999).
+One of the most widely used performance tasks is the pegboard task (see Hodgson",'red')` & `r colFmt("Hudson, 2016, for discussion) yet this task is far from employable online given the requirement of pegs and a board.
+Thus, in addition to examining our hypotheses, we sought to develop and assess the reliability of a novel finger tapping task that could be used to quantify handedness based on performance.",'red')`
 
@@
 However, their study was restricted to left-handers, who may be atypical in this regard, and they noted that lack of association was hard to interpret because it may be due to unreliability of the individual measures.
-In the preregistration for our current study, we predicted that for right-handers, a model where the two language tasks are correlated with each other but not with the chimeric faces (model A) would be the best fit to the data, compared to either a model in which laterality indices are correlated between the language and chimeric face tasks (model B), or a base model where each task is independently lateralised (model C).
+In the pre-registration for our current study, we predicted that for right-handers, a model where the two language tasks are correlated with each other but not with the chimeric faces (model A) would be the best fit to the data, compared to either a model in which laterality indices are correlated between the language and chimeric face tasks (model B), or a base model where each task is independently lateralised (model C).
 Given our previous findings using fTCD to study dissociations in language laterality (Woodhead, Bradshaw, Wilson, Thompson, & Bishop, 2019), we further predicted that model fit may differ for right- and left-handers, with more dissociation between tasks in the latter group.
@@
 In practice, our final sample exceeded 300, because we discarded participants from individual tasks for failing to meet prespecified criteria (see below).
-In total, 441 native English speakers were initially recruited through Prolific Academic (www.prolific.co).
-Participants had normal or corrected-to-normal vision and reported no history of neurological disease (e.g.
+In total, 441 native English speakers were initially recruited through Prolific Academic (www.prolific.co)`r colFmt(", with 411 participants completing both sessions",'red')`.
+Participants `r colFmt("were 18- to 50-years-old,",'red')`, had normal or corrected-to-normal vision and reported no history of neurological disease (e.g.
 head injury, brain tumour, stroke) or significant hearing loss or impairment.
@@
 Each test classified participants as being either left or right eye dominant.
-In practice, the two tests agreed perfectly, so we report here just the Miles test.
+`r colFmt("While we do not use tests of ocular dominance to asnwer apriori predictions, we report eyedness for participants inorder to further characterise them.",'red')` In practice, the two tests agreed perfectly, so we report here just the Miles test.
 
@@
 With online presentation, a particular concern is a potential for eye movements away from fixation.
-We relied on the fact that participants had to focus initially on the central written word to do the task, with pictures subsequently displayed with brief bilateral presentation, which has been shown to give adequate control for eye movements in previous visual half-field experiments with verbal naming (Beaumont, 1982).
+We relied on the fact that participants had to focus initially on the central written word to do the task `r colFmt(".
+We believed that this would encourage fixation to a central location at the start of the trial.
+We subsequently displayed pictures under",'red')`, brief bilateral presentation, which has been shown to give adequate control for eye movements in previous visual half-field experiments with verbal naming (Beaumont, 1982).
 Studies that directly monitored eye movements in this setting reported failures of fixation on only 0.5% of trials (Geffen, Bradshaw & Nettleton, 1972; but see also Bourne, 2006).
 Pictures of familiar objects were presented for rhyme matching to the written word, to ensure that the participant would have to covertly name the picture, thereby engaging left hemisphere speech processing.
+`r colFmt("While previous studies (e.g.
+Hunter",'red')` & `r colFmt("Brysbaert, 2008) have used overt naming, doing so online presents several problems.
+First, accurate naming latencies would require a microphone.
+This would have to be of sufficent quality to accurate capture naming latencies which many participants may not have.
+Seond, the recording of auditory responses would require additional scoring following data collection which would be a taxing procedure for researchers.
+It would also require additional ethical considerations with regards to the storing and processing of the data.
+It is for these reasons that we adopted a covert naming approach.",'red')`
 
@@
 
-![Figure 1.
-Schematic illustration of a trial on the rhyme decision task.
+![Schematic illustration of a trial on the rhyme decision task.
 In the example, the word “bite” is presented along with the two images “kite” and “door”.
-When making the correct judgement, participants would respond with by pressing ‘S’ to indicate that the rhyme was on the left.](Picture1.png)
+When making the correct judgement, participants would respond with by pressing ‘S’ to indicate that the rhyme was on the left.](Picture1.png){width=60%}
 
@@
 If participants did not detect a rhyme, they were instructed to press the space bar.
-The response triggered the next trial.
+The response triggered the next trial`r colFmt("; trials did not time-out after a predetermined duration",'red')`.
 Accuracy and reaction times (RT) were recorded.
@@
 The second check involved participants hearing a sound presented in only one of the two channels.
-Participants were removed if they failed to correctly identify which channel the sound was played on five out of six trials, indicating that they were not using stereo-headphones.
+Participants were removed if they failed to correctly identify which channel the sound was played on five out of six trials, indicating that they were not using stereo-headphones [^1].
 
@@
 After hearing each stimulus, participants were instructed to report the syllable they heard or if it seemed like they heard two different sounds, the one they heard most clearly by clicking a button corresponding to one of the six syllables.
-The response triggered the start of the next trial.
+The response triggered the start of the next trial`r colFmt("; trials did not time-out after a predetermined duration",'red')`.
 Accuracy and reaction times (RT) were recorded.
@@
 
-![Figure 2.
-An example stimulus from the chimeric faces task.
+\
+
+![An example stimulus from the chimeric faces task.
 In the example, “happiness” is shown on the left and “sadness” on the right.
-A participant who is right lateralised for emotional processing would be more likely to perceive the emotion in the left hemifield and would, therefore, respond “happiness”.](chim_example.png)
+A participant who is right lateralised for emotional processing would be more likely to perceive the emotion in the left hemifield and would, therefore, respond “happiness”.](chim_example.png){width=20%}
+
+\
 
@@
 Participants then reported the emotion seen in the face by clicking a button corresponding to one of the four emotions.
-The response triggered the start of the next trial.
+The response triggered the start of the next trial`r colFmt("; trials did not time-out after a predetermined duration",'red')`.
 Accuracy and reaction times (RT) were recorded.
@@
 Each session lasted approximately 45 minutes and 35 minutes respectively.
+`r colFmt("Participants were paid £15 for their participation.",'red')`
 
@@
 Gorilla’s scaling tool was utilised such that stimuli size and eccentricity was maintained across tasks and participants.
-Stimuli from each task were repeated across four blocks.
+`r colFmt("During each task participants were instructed respond as fast and accurately as possible.",'red')` Stimuli from each task were repeated across four blocks.
 Within each block, participants completed 70 trials of the rhyme detection task (R); 36 trials of the dichotic listening task (D); 32 trials of the chimeric face task (C); and 2 trials of the finger tapping task (F) for both their dominant and non-dominant hand.
@@
 Response times were log-transformed to reduce skew, as specified in our pre-registration.
-Outlier reaction times were identified using Hoaglin and Iglewicz’s (1987) procedure, whereby outliers are defined as response times that are 1.65 times the difference between the first and third quantiles (1.65 * (Q3-Q1)) below or above the first and third quantile values (e.g: lower limit = Q1 – 1.65 x (Q3-Q1); upper limit = Q3 + 1.65 x (Q3-Q1)).
+Outlier reaction times were identified using Hoaglin and Iglewicz’s (1987) procedure, whereby outliers are defined as response times that are 1.65 times the difference between the first and third quantiles (1.65 * (Q3-Q1)) below or above the first and third quantile values (e.g: lower limit = Q1 – 1.65 x (Q3-Q1); upper limit = Q3 + 1.65 x (Q3-Q1)) [^2].
 Our differences criterion was set to 1.65 instead of the typical 2.2 because we expected to obtain asymmetric distributions for response time, with most outliers in one tail of the distribution.
@@
 
-We preregistered that we would calculate LIs in two ways: (1) traditionally (100 x (Right – Left)/(Right + Left)); (2) a standardised LI that involved computing a *z*-score that represented, for each individual, a comparison between left- and right-sided responses.
+We pre-registered that we would calculate LIs in two ways: (1) traditionally (100 x (Right – Left)/(Right + Left)); (2) a standardised LI that involved computing a *z*-score that represented, for each individual, a comparison between left- and right-sided responses.
 In fact, for RT data we specified a *t*-score, but *t* and *z* converge when sample size exceeds 30, and for simplicity, we refer to these as LI_z scores.
@@
 
+`r colFmt("To address hypothesis 2, that extreme left-handers will be more likely to show atypical laterality than right-handers, we categorised participants as extreme left-handers if they scored less than -90 on the EHI and right-handers if they scored greater than zero on the EHI.
+Following our pre-registration, atypicality was defined according to whether the participant's LI_z was significantly lateralised (with alpha = .05) in the opposite direction from the majority.
+Chi-squared tests were then used to examine the association between handedness and atypicality on the rhyme decision, dichotic listening, and chimeric faces tasks.
+In our pre-registration we had planned to look at cases who were atypical on the language tasks and the chimeric faces.
+We did not conduct this anlaysis due to very few participants showing atypical lateralisation across multiple tasks.
+In our Supplementary Materials we do, however, compare laterality indices between extreme left-handers and right-handers to further examine the link between handedness and atypical lateralisation.", 'red')`
+
+<!--NOT NEEDED. THIS DESCRIBES AN ANALYSIS THAT WE DIDN'T CONDUCT.
+
 To address hypothesis 2, that extreme left-handers will be more likely to show atypical laterality than right-handers, we categorised participants as extreme left-handers if they scored less than -90 and right-handers if they scored more than 90 on the Edinburgh Handedness Inventory.
-Our preregistration specified that atypicality would first be defined according to whether the participant's LI_z was significantly lateralised (with alpha = .05) in the opposite direction from the majority.
+Our pre-registration specified that atypicality would first be defined according to whether the participant's LI_z was significantly lateralised (with alpha = .05) in the opposite direction from the majority.
 If there were less than 20 cases defined as atypical using this method, we planned to adjust the cut-off on that task until we acquired 20 atypical participants.
@@
 A Bonferroni correction was applied to the outcome of each of the two chi-squared tests.
-
-<!---ZW: This bit confused me.
-
-I thought that the hypothesis was that extreme left handers were more likely to be atypical on any given task, but the test looks at consistency of laterality across pairs of task.
-Isn’t that testing a different hypothesis?
-
-Both are interesting, but I think the first one needs to be tested too.
-E.g., are extreme left handers more likely than right handers to be atypically lateralized on Rhyme Decision / Chimeric Faces / Dichotic Listening?
 --->
@@
 As participants self-reported eye dominance was always consistent between the Miles and Porta tests, we report data only from the Miles test.
+`r colFmt("As eye dominance is not a focus of this study, we will not elaborate further on it.", 'red')`
 
@@
 
-Table 1 <!---New table - lattab to be added in nice format--->
-Count of left- and right-handers who report left or right on self-reported footedness, and eyedness (Miles test). 
+\
+
+Table 1. Count of left- and right-handers who report left or right on self-reported footedness, and eyedness (Miles test).\
 ```{r Tab1,echo=FALSE, comment=FALSE, warning=FALSE}
+lattab<-lattab[,c(1,2,4,3,5)] #now reorder columns to make sense
+colnames(lattab)<-c('Foot-Eye','Right hander (N)','Left hander (N)','Right hander (%)','Left hander (%)')
 lattab <- flextable(lattab)
+lattab <- align(lattab, align = "center", j= 2:5)
 autofit(lattab, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
 ```
+
 *Note*. RR: right foot, right eye; RL: right foot, left eye; LR: left foot, right eye; LL: left foot, left eye. 'Other' includes those who responded 'No preference / Don't know'.
+
+\
 
@@
 
-Distributions of the raw data are shown in the Online Supplemental Materials.
+Distributions of the raw data are shown in `r colFmt("Figure 3.
+The scatterplots indicate scores on the left and right sides for each task, with self-reported handedness (left vs right) colour-coded.
+For each task, the mean for each side is shown as vertical and horizontal dotted lines, and equal performance on the two sides corresponds to the sloping red dotted line: thus an impression of whether the task is lateralised can be obtained by considering whether points cluster around the sloping lines symmetrically.", 'red')`
+
+\newpage
+
+![Left side (L) vs right side (R) scores for sessions 1 and 2 for the behavioural laterality tasks.
+The diagonal dotted line shows the point of equality: points above this are higher for R than L, and points below are higher for L than R.
+Mean values for R and L are shown as horizontal and vertical dotted lines.
+The colour of the points indicates the participants’ handedness (black = right-handed; red = left-handed).
+](all.raw.png){width=80%}
+
+`r colFmt("For the rhyme decision task, only correct responses were used for computing mean log RTs (shown transformed back to the original ms scale in the plot).
+Generally, accuracy on the rhyme decision task was high.",'red')` For session 1, accuracy was `r round(mean(allsum$RDT1.pcorrL, na.rm= TRUE),1)`% (*SD* = `r round(sd(allsum$RDT1.pcorrL, na.rm= TRUE),1)`%) in the LVF and `r round(mean(allsum$RDT1.pcorrR, na.rm= TRUE),1)`% (*SD* = `r round(sd(allsum$RDT1.pcorrR, na.rm= TRUE),1)`%) in the RVF.
+For session 2, accuracy was `r round(mean(allsum$RDT2.pcorrL, na.rm= TRUE),1)`% (*SD* = `r round(sd(allsum$RDT2.pcorrL, na.rm= TRUE),1)`%) in the LVF and `r round(mean(allsum$RDT2.pcorrR, na.rm= TRUE),1)`% (*SD* = `r round(sd(allsum$RDT2.pcorrR, na.rm= TRUE),1)`%) in the RVF.
+`r colFmt("For dichotic listening and chimeric faces, a correct response can be given from either side", 'red')`; `r colFmt("it follows that there will be a negative correlation between the two sides reflecting the fact that the number of selections on the left will be inversely related to the number of correct responses on the right.
+For these tasks, it is evident by inspection that most responses for dichotic listening cluster above the dotted line, indicating a right ear advantage, whereas most responses on chimeric faces cluster below the dotted line, indicating a left hemi-face advantage.
+Finally, we show the number of correct sequences with each hand on the finger tapping task.
+Here, a strong impact of handedness is evident on inspection, with left-handers generally more accurate with the left hand, and right-handers more accurate with the right hand.",'red')`
+
 The distribution of LIs can be used to quantify departures from symmetry.
 Recall that for all tasks, we computed LIs in a typical way: (100 x (Right – Left)/(Right + Left)), and that because of crossed pathways from hands, ears and eyes to the brain, a positive LI indicates left lateralisation and a negative LI indicates right lateralisation at the brain level.
-LIs on the rhyme task were calculated using log RTs for correct responses (accuracy was high; *M* = `r round(allmeanpcorr,1)`%, *SD* = `r round(allsdpcorr,2)` %).
+LIs on the rhyme task were calculated using log RTs for correct responses.
 LIs for the dichotic listening and chimeric faces were calculated using the count of correct responses for sounds/emotions in each ear/visual.
@@
 colnames(LItable)<-c('Task','N','Mean','SD','One sample v','One sample p','Normality p')
-LItable[,1]<-c('Rhyme Decision 1','Rhyme Decision 2','Dichotic 1','Dichotic 2','Chimeric Faces 1','Chimeric Faces 2','Finger Tapping 1','Finger Tapping 2')
+LItable[,1]<-c('RD 1','RD 2','Dl 1','DL 2','CF 1','CF 2','FT 1','FT 2')
 
@@
 ```
-
-Table 2.Mean LIs and results from one-sample t-tests and Shapiro-Wilks tests for departure from normality for each task at sessions 1 and 2.
+\
+
+Table 2.
+Mean LIs and results from one-sample t-tests and Shapiro-Wilks tests for departure from normality for the rhyme decision (RD), dichotic listening (DL), chimeric faces (CF), and finger tapping (FT) tasks at sessions 1 and 2.\
 ```{r Tab2,echo=FALSE, comment=FALSE, warning=FALSE}
 tab2 <- flextable(LItable)
+tab2 <- align(tab2, j= 2:7, align= "center")
 autofit(tab2, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
 ```
+
 *Note*. LIs are scored such that a positive value equates to left lateralisation at the brain level.
 
+\
+
 Table 3 shows data from the LI_z measures, which make it easier to compare different tasks on the same scale.
-For each task, those with LI_z scores greater than 1.96 are categorised as left-hemisphere dominant, and those with LI_z scores less than -1.96 are categorised as right-hemisphere dominant.
+`r colFmt("The distributions of LI_z scores are shown graphically in the supplementary materials.",'red')` For each task, those with LI_z scores greater than 1.96 are categorised as left-hemisphere dominant, and those with LI_z scores less than -1.96 are categorised as right-hemisphere dominant.
 If there were no population bias, we would expect around 2.5% of individuals to meet each of these criteria.
@@
 Note, however, that we had oversampled left-handers for this study, and, as is evident from exploratory analyses presented below, finger-tapping was strongly influenced by handedness.
+
 <!---this is also very evident from the scatterplot in Fig 3--->
@@
 colnames(LIztable)<-c('Task','N','Mean','SD','L-lateralised %','R-lateralised %')
-LIztable[,1]<-c('Rhyme Decision 1','Rhyme Decision 2','Dichotic 1','Dichotic 2','Chimeric Faces 1','Chimeric Faces 2','Finger Tapping 1','Finger Tapping 2')
+LIztable[,1]<-c('RD 1','RD 2','DL 1','DL 2','CF 1','CF 2','FT 1','FT 2')
 
@@
 ```
-
-Table 3. Mean LI_z and percentage of participants who were significantly left lateralised (z > 1.96) or right lateralised (z < -1.96) for each task at sessions 1 and 2.
+\
+
+Table 3.
+Mean LI_z and percentage of participants who were significantly left lateralised (z > 1.96) or right lateralised (z < -1.96) on the rhyme decision (RD), dichotic listening (DL), chimeric faces (CF), and finger tapping (FT) tasks at sessions 1 and 2.\
 ```{r Tab3,echo=FALSE, comment=FALSE, warning=FALSE}
 tab3 <- flextable(LIztable)
+tab3 <- align(tab3, align= "center", j= 2:6)
 autofit(tab3, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
 ```
+
 *Note*. LIs are scored such that a positive value equates to left lateralisation at the brain level.
+\
 
@@
 
-The scatterplots depicting test-retest reliability are shown in Figure 3, with Spearman test-retest correlations in Table 4.
+The scatterplots depicting test-retest reliability are shown in Figure 4, with Spearman test-retest correlations in Table 4.
 These were based on LI_z values, but are essentially identical to values from unscaled LI values.
@@
 
-Table 4. Test-retest reliability estimates and 95% CIs for each behavioural task.
+\
+
+Table 4. Test-retest reliability estimates and 95% CIs for the rhyme decision (RD), dichotic listening (DL), chimeric faces (CF), and finger tapping (FT) tasks.\
 ```{r Tab4,echo=FALSE, comment=FALSE, warning=FALSE}
+combotable$Task[1] <- "RD"
+combotable$Task[2] <- "DL"
+combotable$Task[3] <- "CF"
+combotable$Task[4] <- "FT"
+
+combotable$`95% CI`[3] <- "0.84 - 0.90"
+combotable$`95% CI`[4] <- "0.68 - 0.80"
+
 tab4 <- flextable(combotable)
+tab4 <- align(tab4, align= "center", j= 2:3)
 autofit(tab4, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
@@
 
-![Figure 3.
-Scatter plots for day 1 and day 2 data for LI_z scores on each laterality task: (A) rhyme decision, (B) dichotic listening, (C) chimeric faces, and (D) Finger tapping.
+\
+
+\newpage
+
+![Scatter plots for day 1 and day 2 data for LI_z scores on each laterality task: (A) rhyme decision, (B) dichotic listening, (C) chimeric faces, and (D) Finger tapping.
 Handedness is colour coded with extreme left-handers identified as those who have a score below -90 on the Edinburgh Handedness Inventory.
-Note that z-scores are censored at absolute values of 5.](combined_LI_plot.png)
+Note that z-scores are censored at absolute values of 5.](combined_LI_plot.png){width=80%}
 
@@
 
-![Figure 4.
-Test-retest correlations for (A) the dichotic listening task and (B) the chimeric faces task when using a variable number of trials to calculate laterality indices.
+![Test-retest correlations for (A) the dichotic listening task and (B) the chimeric faces task when using a variable number of trials to calculate laterality indices.
 The red dashed line marks the test-retest correlation *r* = .65.
-The correlations coefficients are shown as absolute values.](all.cor.plot.png)
+The correlations coefficients are shown as absolute values.](all.cor.plot.png){width=70%}
 
@@
 According to Mazoyer et al (2014), extreme left-handers are more likely than other handedness groups to have reversed laterality compared to the rest of the population.
-Our preregistered analysis accordingly focused on comparing extreme left-handers (those with a score of -90 or below on the Edinburgh Handedness Inventory) with right-handers.
+Our pre-registered analysis accordingly focused on comparing extreme left-handers (those with a score of -90 or below on the Edinburgh Handedness Inventory) with right-handers.
 We defined 'atypical laterality' as laterality that is in the opposite direction to the majority of participants, with an absolute LI_z value greater than 1.96.
@@
 
-Table 5. Frequencies of laterality typicality for extreme left-handers, moderate left-handers and right-handers across the rhyme decision, dichotic listening, and chimeric faces tasks.
+\
+
+Table 5. Frequencies of laterality typicality for extreme left-handers, moderate left-handers and right-handers across the rhyme decision (RD), dichotic listening (DL), and chimeric faces (CF) tasks.\
 ```{r Table2,echo=FALSE, comment=FALSE, warning=FALSE}
-tabr<-cbind(tabr,round(100*prop.table(tabr,2),1))
-tabd<-cbind(tabd,round(100*prop.table(tabd,2),1) )
-tabc<-cbind(tabc,round(100*prop.table(tabc,2),1)) #add percentages
-blankrow<-c(NA,NA,NA,NA,NA,NA)
-latcats <- as.data.frame(rbind(blankrow,tabr,blankrow,tabd,blankrow,tabc))
+tabr<-cbind(tabr,round(100*prop.table(tabr,1),1))
+tabd<-cbind(tabd,round(100*prop.table(tabd,1),1) )
+tabc<-cbind(tabc,round(100*prop.table(tabc,1),1)) #add percentages
+latcats <- as.data.frame(rbind(tabr,tabd,tabc))
 
@@
 #Instead have a new column with these
-latcats$nucol <- c('Rhyme decision',
+latcats$nucol <- c('Extreme left-hander',
+                   'Moderate left-hander',
+                   'Right-hander',
                    'Extreme left-hander',
@@
                    'Right-hander',
-                   'Dichotic listening',
-                   'Extreme left-hander',
-                   'Moderate left-hander',
-                   'Right-hander',
-                   'Chimeric faces',
                    'Extreme left-hander',
@@
                    'Right-hander')
-
-latcats<-latcats[,c(7,3,6,2,5,1,4)] #now reorder columns to make sense
-colnames(latcats)<-c(' ','Left N','Left %','Non N','Non %','Right N','Right %')
-
-myft2 <- flextable(
-  latcats)
+latcats$nucol2 <- c('RD',
+                   ' ',
+                   ' ',
+                   'DL',
+                   ' ',
+                   ' ',
+                   'CF',
+                   ' ',
+                   ' ')
+
+latcats<-latcats[,c(8,7,3,2,1,6,5,4)] #now reorder columns to make sense
+colnames(latcats)<-c('Task','Sub-group','Left N','Non N','Right N','Left %','Non %','Right %')
+
+myft2 <- flextable(latcats)
 #NB This table needs formatting - has formatted all numbers to 3 decimal places, which is not what we want. Code below achieves this
@@
 ft <- colformat_int(x = ft, col_keys = int_keys, big.mark = ",")
-
 autofit(ft)
 ```
+
 *Note*. N is the count of participants. % is the percentage of participants.
 
-```{r pformatfunction, echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
+\
+
+```{r pformatfunction, echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
 pformat<-function(pvalue){
@@
 Both dichotic listening and chimeric faces, on the other hand, have a high proportion of strongly lateralised individuals.
-The planned analyses, comparing typically lateralised with atypically lateralised individuals, and extreme left vs right-handed individuals, show a significant association between extreme left-handedness and atypical laterality on chi-square test.
+The planned analyses, comparing typically lateralised with atypically lateralised individuals, and extreme left vs right-handed individuals, show a significant association between extreme left-handedness and atypical laterality on a chi-square test.
 For dichotic listening, chi-squared = `r round(chi.d$statistic,2)`, d.f.
@@
 = 1, `r pformat(p.adjust(chi.c$p.value, method= "bonferroni", n=2))`.
-
-With all participants included, on the dichotic listening task, `r round(100*tabd[1,3]/sum(tabd[,3]),1)`% of those with atypical laterality were extreme left-handers, `r round(100*tabd[3,3]/sum(tabd[,3]),1)`% were right-handers, and the remainder less extreme left-handers.
-For those with typical lateralisation, the corresponding figures were `r round(100*tabd[1,1]/sum(tabd[,1]),1)`% extreme left-handers, `r round(100*tabd[3,1]/sum(tabd[,1]),1)`% right-handers, and the remainder less extreme left-handers.
+`r colFmt("Thus, it would appear that there is statistical evidence for departures from typical cerebral lateralisation in extreme left-handers supporting Mazoyer et al.
+(2014)..",'red')`
+
+<!--- REMOVING THIS AS IT SEEMS TO BE THE PORTION OF TEXT. IT DOES COMPLICATE THE MESSAGE AND MARC IS CORRECT, IT DOES NOT ACCOUNT FOR SAMPLE SIZES OF PARTICIPANTS.
+
+With all participants included, on the dichotic listening task, `r round(100*tabd[1,3]/sum(tabd[,3]),1)`% of those with typical laterality were extreme left-handers, `r round(100*tabd[3,3]/sum(tabd[,3]),1)`% were right-handers, and the remainder less extreme left-handers.
+For those with atypical lateralisation, the corresponding figures were `r round(100*tabd[1,1]/sum(tabd[,1]),1)`% extreme left-handers, `r round(100*tabd[3,1]/sum(tabd[,1]),1)`% right-handers, and the remainder less extreme left-handers.
 For chimeric faces, the association was even more extreme.
@@
 For those with typical lateralisation, `r round(100*tabc[1,3]/sum(tabc[,3]),1)`% were extreme left-handers, `r round(100*tabc[3,3]/sum(tabc[,3]),1)`% were right-handers, and the remainder less extreme left-handers.
+--->
 
@@
 ```
+\
 
@@
 Spearman correlations between the LI_z scores for the different tasks in two Sessions.
-The upper diagonal shows correlations for the 143 right-handers (black font), and the values of the lower diagonal show for the 66 left-handers (red font).
+The upper diagonal shows correlations for the 143 right-handers (black font), and the values of the lower diagonal show for the 66 left-handers (red font).\
 ```{r tab 5, echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
-allcorr2 <- cbind(Row.Names = rownames(allcorr), allcorr)
+allcorr2 <- cbind(Row.Names = list(rownames(allcorr)), allcorr, stringsAsFactors = FALSE)
+
+# rename columns
 names(allcorr2)[1] <- " "
-allcorr2[is.na(allcorr2)] <- "."
+allcorr2$` `[1] <- "RD 1"
+allcorr2$` `[2] <- "RD 2"
+allcorr2$` `[3] <- "DL 1"
+allcorr2$` `[4] <- "DL 2"
+allcorr2$` `[5] <- "CF 1"
+allcorr2$` `[6] <- "CF 2"
+
+# rename rows
+names(allcorr2)[2] <- "RD 1"
+names(allcorr2)[3] <- "RD 2"
+names(allcorr2)[4] <- "DL 1"
+names(allcorr2)[5] <- "DL 2"
+names(allcorr2)[6] <- "CF 1"
+names(allcorr2)[7] <- "CF 2"
+
+# replace blanks with '.'
+#allcorr2$`RD 1`[1] <- "."
+#allcorr2$`RD 2`[2] <- "."
+#allcorr2$`DL 1`[3] <- "."
+#allcorr2$`DL 2`[4] <- "."
+#allcorr2$`CF 1`[5] <- "."
+#allcorr2$`CF 2`[6] <- "."
+
 myft5<- flextable(allcorr2)
-
-myft5<-color(myft5, i = c(2,3,4,5, 6), j = "Rhyme_1", color= "red")
-myft5<-color(myft5, i = c(3,4,5, 6), j = "Rhyme_2", color= "red")
-myft5<-color(myft5, i = c(4,5, 6), j = "Dichotic_1", color= "red")
-myft5<-color(myft5, i = c(5, 6), j = "Dichotic_2", color= "red")
-myft5<-color(myft5, i = c(6), j = "Chimeric_1", color= "red")
+myft5 <- colformat_num(myft5, digits= 3)
+
+myft5<-color(myft5, i = c(2,3,4,5, 6), j = "RD 1", color= "red")
+myft5<-color(myft5, i = c(3,4,5, 6), j = "RD 2", color= "red")
+myft5<-color(myft5, i = c(4,5, 6), j = "DL 1", color= "red")
+myft5<-color(myft5, i = c(5, 6), j = "DL 2", color= "red")
+myft5<-color(myft5, i = c(6), j = "CF 1", color= "red")
 
@@
 ```
-Note.
+
+*Note*.
 For right-handers, an absolute value of .16 is significant at the .05 level, and a value of .22 is significant at the .01 level.
 For left-handers, a value of .25 is significant at the .05 level and a value of .32 is significant at the .01 level.
+
+\
 
@@
 
-Table 7. Akaike weights for the independent base model, language model, and laterality model.
+\
+
+Table 7. Akaike weights for the independent base model, language model, and laterality model.\
 ```{r AICmodels,echo=FALSE, comment=FALSE, warning=FALSE}
@@
 
+\
+
+# Updates code for correlations to test model and account for test retest
+
+```{r update, echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
+allcorr2 <- cbind(Row.Names = list(rownames(allcorr)), allcorr, stringsAsFactors = FALSE)
+
+# rename columns
+names(allcorr2)[1] <- " "
+allcorr2$` `[1] <- "RD 1"
+allcorr2$` `[2] <- "RD 2"
+allcorr2$` `[3] <- "DL 1"
+allcorr2$` `[4] <- "DL 2"
+allcorr2$` `[5] <- "CF 1"
+allcorr2$` `[6] <- "CF 2"
+
+# rename rows
+names(allcorr2)[2] <- "RD 1"
+names(allcorr2)[3] <- "RD 2"
+names(allcorr2)[4] <- "DL 1"
+names(allcorr2)[5] <- "DL 2"
+names(allcorr2)[6] <- "CF 1"
+names(allcorr2)[7] <- "CF 2"
+
+# replace blanks with '.'
+#allcorr2$`RD 1`[1] <- "."
+#allcorr2$`RD 2`[2] <- "."
+#allcorr2$`DL 1`[3] <- "."
+#allcorr2$`DL 2`[4] <- "."
+#allcorr2$`CF 1`[5] <- "."
+#allcorr2$`CF 2`[6] <- "."
+
+# correction formula
+# RD 1
+allcorr2$`RD 1`[3] <- allcorr2$`RD 1`[3] / sqrt(.630 * .780)
+allcorr2$`RD 1`[4] <- allcorr2$`RD 1`[4] / sqrt(.630 * .780)
+allcorr2$`RD 1`[5] <- allcorr2$`RD 1`[5] / sqrt(.630 * .878)
+allcorr2$`RD 1`[6] <- allcorr2$`RD 1`[6] / sqrt(.630 * .878)
+# RD 2
+allcorr2$`RD 2`[3] <- allcorr2$`RD 2`[3] / sqrt(.630 * .780)
+allcorr2$`RD 2`[4] <- allcorr2$`RD 2`[4] / sqrt(.630 * .780)
+allcorr2$`RD 2`[5] <- allcorr2$`RD 2`[5] / sqrt(.630 * .878)
+allcorr2$`RD 2`[6] <- allcorr2$`RD 2`[6] / sqrt(.630 * .878)
+# DL 1
+allcorr2$`DL 1`[1] <- allcorr2$`DL 1`[1] / sqrt(.630 * .780)
+allcorr2$`DL 1`[2] <- allcorr2$`DL 1`[2] / sqrt(.630 * .780)
+allcorr2$`DL 1`[5] <- allcorr2$`DL 1`[5] / sqrt(.780 * .878)
+allcorr2$`DL 1`[6] <- allcorr2$`DL 1`[6] / sqrt(.780 * .878)
+# DL 2
+allcorr2$`DL 2`[1] <- allcorr2$`DL 2`[1] / sqrt(.630 * .780)
+allcorr2$`DL 2`[2] <- allcorr2$`DL 2`[2] / sqrt(.630 * .780)
+allcorr2$`DL 2`[5] <- allcorr2$`DL 2`[5] / sqrt(.780 * .878)
+allcorr2$`DL 2`[6] <- allcorr2$`DL 2`[6] / sqrt(.780 * .878)
+# CF 1
+allcorr2$`CF 1`[1] <- allcorr2$`CF 1`[1] / sqrt(.630 * .878)
+allcorr2$`CF 1`[2] <- allcorr2$`CF 1`[2] / sqrt(.630 * .878)
+allcorr2$`CF 1`[3] <- allcorr2$`CF 1`[3] / sqrt(.780 * .878)
+allcorr2$`CF 1`[4] <- allcorr2$`CF 1`[4] / sqrt(.780 * .878)
+# CF 2
+allcorr2$`CF 2`[1] <- allcorr2$`CF 2`[1] / sqrt(.630 * .878)
+allcorr2$`CF 2`[2] <- allcorr2$`CF 2`[2] / sqrt(.630 * .878)
+allcorr2$`CF 2`[3] <- allcorr2$`CF 2`[3] / sqrt(.780 * .878)
+allcorr2$`CF 2`[4] <- allcorr2$`CF 2`[4] / sqrt(.780 * .878)
+
+myft5<- flextable(allcorr2)
+myft5 <- colformat_num(myft5, digits= 3)
+
+myft5<-color(myft5, i = c(2,3,4,5, 6), j = "RD 1", color= "red")
+myft5<-color(myft5, i = c(3,4,5, 6), j = "RD 2", color= "red")
+myft5<-color(myft5, i = c(4,5, 6), j = "DL 1", color= "red")
+myft5<-color(myft5, i = c(5, 6), j = "DL 2", color= "red")
+myft5<-color(myft5, i = c(6), j = "CF 1", color= "red")
+
+autofit(myft5, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
+
+# now create subset of data frame for analysis
+interRR <- dplyr::select(LIs, zlat_RDT_1, zlat_RDT_2, Day1_Dich_acc_Z,
+                         Day2_Dich_acc_Z, Day1_CF_acc_Z, Day2_CF_acc_Z,handedness)
+# rename, and reverse sign for CF
+interRR <- interRR %>% 
+  rename(
+    "rhyme1" = zlat_RDT_1,
+    "faces1" = Day1_CF_acc_Z,
+    "dichotic1" = Day1_Dich_acc_Z,
+     "rhyme2" = zlat_RDT_2,
+    "faces2" = Day2_CF_acc_Z,
+    "dichotic2" = Day2_Dich_acc_Z
+    )
+
+#remove cases with missing values
+interRR<-na.omit(interRR)
+interRR$faces1 <- (-interRR$faces1) #reverse sign for faces
+interRR$faces2 <- (-interRR$faces2)
+# subset into right and left handers
+left_dat <- subset(interRR, handedness== "Left")
+right_dat <- subset(interRR, handedness== "Right")
+
+#NB in pre-reg we just had time 1 data, but here we use time 1 and 2
+
+ #Model 1 all independent
+    model1<-"
+
+    rhymeFACTOR =~ NA*rhyme1 + rhyme2
+    dichoticFACTOR =~ NA*dichotic1 + dichotic2
+    facesFACTOR =~ NA*faces1 +faces2
+
+    rhymeFACTOR~~1*rhymeFACTOR
+    dichoticFACTOR~~1*dichoticFACTOR
+    facesFACTOR~~1*facesFACTOR
+
+    rhymeFACTOR~~0*dichoticFACTOR
+    rhymeFACTOR~~0*facesFACTOR
+    dichoticFACTOR~~0*facesFACTOR
+    "
+#Model 2: language tasks are correlated
+    model2<-"
+    rhymeFACTOR =~ NA*rhyme1 + rhyme2
+    dichoticFACTOR =~ NA*dichotic1 + dichotic2
+    facesFACTOR =~ NA*faces1 +faces2
+
+    rhymeFACTOR~~1*rhymeFACTOR
+    dichoticFACTOR~~1*dichoticFACTOR
+    facesFACTOR~~1*facesFACTOR
+
+    rhymeFACTOR~~dichoticFACTOR
+    rhymeFACTOR~~0*facesFACTOR
+    dichoticFACTOR~~0*facesFACTOR
+    "
+#Model 3:all correlated
+    model3<-"
+    rhymeFACTOR =~ NA*rhyme1 + rhyme2
+    dichoticFACTOR =~ NA*dichotic1 + dichotic2
+    facesFACTOR =~ NA*faces1 +faces2
+
+    rhymeFACTOR~~1*rhymeFACTOR
+    dichoticFACTOR~~1*dichoticFACTOR
+    facesFACTOR~~1*facesFACTOR
+
+    rhymeFACTOR~~dichoticFACTOR
+    rhymeFACTOR~~facesFACTOR
+    dichoticFACTOR~~facesFACTOR
+    "
+#Make a little dataframe for AIC weights
+AICdf <- data.frame(matrix(NA,nrow=2,ncol=4))
+AICdf[,1]<-c('Right-handers','Left-handers')
+colnames(AICdf)<-c('Handedness','Model C','Model A','Model B')
+for (myhand in 1:2) {
+  mydat<-right_dat
+  if(myhand==2) {mydat<-left_dat}
+
+ #Using data from this handedness type, test all 3 models
+fit1A <- cfa(model1, data=mydat,auto.fix.first=FALSE)
+fit2A <- cfa(model2, data=mydat,auto.fix.first=FALSE)
+fit3A <- cfa(model3, data=mydat,auto.fix.first=FALSE)
+    aic_vector1 <- c(fitMeasures(fit1A, "aic"),fitMeasures(fit2A, "aic"),fitMeasures(fit3A, "aic"))
+
+    #Confirms that for each model, lowest AIC obtained when the simulated data do match the model
+    # 
+    # The AIC weights make massive difference for preferred model!
+    AICweights<-akaike.weights(aic_vector1)$weights
+    # 
+ AICdf[myhand,2:4]<-round(AICweights,3)
+ }
+myft6 <- flextable(AICdf, col_keys = c("Handedness",
+                                       "Model A",
+                                       "Model B", 
+                                       "Model C"))
+autofit(myft6, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
+```
+
 ### Exploratory analyses
@@
 
-```{r hand_explore, echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
+```{r hand_explore, echo=FALSE,include= FALSE, comment=FALSE, warning=FALSE}
 hand_cors <- data.frame(matrix(ncol = 5, nrow = 1))
@@
 
+<!--- Based on reviewer's comments. Haven't added in yet, but holding it here --->
+
+```{r makecontingencies_tasks,echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
+#We start by using z-scores to make a 3-way categorisation into:
+#1 sig bias in typical direction; 2: NS bias 3: in atypical direction with sig bias.
+#
+LIs$rhyme1.cat <- cut(LIs$Day1.RDT_p.corr_LI, 
+                       breaks = c(-Inf, -1.96, 1.96, Inf), 
+                       labels = c("R.Right", "R.Bilat", "R.Left"), 
+                       right = FALSE)
+LIs$dichotic1.cat <- cut(LIs$Day1_Dich_acc_Z, 
+                       breaks = c(-Inf, -1.96, 1.96, Inf), 
+                       labels = c("D.Right", "D.Bilat", "D.Left"), 
+                       right = FALSE)
+LIs$chimeric1.cat <- cut(LIs$Day1_CF_acc_Z, 
+                       breaks =  c(-Inf, 0, 1.96, Inf), 
+                       labels = c("C.Right", "C.Bilat", "C.Right"), 
+                       right = FALSE)
+mytabr <- table(LIs$rhyme1.cat)
+mytabd <- table(LIs$dichotic1.cat)
+mytabc <- table(LIs$chimeric1.cat)
+mytabc<-mytabc[c(3,2,1)]
+mytab2 <-rbind(mytabr,mytabd,mytabc)
+row.names(mytab2)<-c('Rhyme decision','Dichotic','Chimeric Faces')
+mytab2<-data.frame(mytab2)
+mytab2$R.Right[3] <- 0
+
+# rhyme vs dichotic
+tab.R.D <- table(LIs$rhyme1.cat,LIs$dichotic1.cat)
+tab.R.D
+chisq.test(tab.R.D)
+
+# rhyme vs chimeric
+tab.R.C <- table(LIs$rhyme1.cat,LIs$chimeric1.cat)
+tab.R.C
+chisq.test(tab.R.C)
+
+# dichotic vs chimeric
+tab.D.C <- table(LIs$dichotic1.cat,LIs$chimeric1.cat)
+tab.D.C
+```
+
+```{r Table_reviews,echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
+tabr<-cbind(tabr,round(100*prop.table(tabr,2),1))
+tabd<-cbind(tabd,round(100*prop.table(tabd,2),1) )
+tabc<-cbind(tabc,round(100*prop.table(tabc,2),1)) #add percentages
+blankrow<-c(NA,NA,NA,NA,NA,NA)
+latcats <- as.data.frame(rbind(blankrow,tabr,blankrow,tabd,blankrow,tabc))
+
+row.names(latcats) <- c() #remove rownames as they make problems for printing
+
+#Instead have a new column with these
+latcats$nucol <- c('Rhyme decision',
+                   'Extreme left-hander',
+                   'Moderate left-hander',
+                   'Right-hander',
+                   'Dichotic listening',
+                   'Extreme left-hander',
+                   'Moderate left-hander',
+                   'Right-hander',
+                   'Chimeric faces',
+                   'Extreme left-hander',
+                   'Moderate left-hander',
+                   'Right-hander')
+
+latcats<-latcats[,c(7,3,6,2,5,1,4)] #now reorder columns to make sense
+colnames(latcats)<-c(' ','Left N','Left %','Non N','Non %','Right N','Right %')
+
+myft2 <- flextable(
+  latcats)
+#NB This table needs formatting - has formatted all numbers to 3 decimal places, which is not what we want. Code below achieves this
+
+num_keys <- c("Left %","Non %","Right %")
+int_keys <- c("Left N","Non N","Right N")
+
+ft <- colformat_num(x = myft2, col_keys = num_keys, digits = 1)
+ft <- colformat_int(x = ft, col_keys = int_keys, big.mark = ",")
+
+#autofit(ft)
+```
+
 #### Vocabulary knowledge and laterality
@@
 # make sure we have good cases
-interR <- LIs[LIs$exclude == 0,]
+interR <- LIs_c[LIs_c$exclude == 0,]
 DL.lex<-RVAideMemoire::spearman.ci(interR$Day1_DL_acc_LI, interR$lexTALE, nrep = 10000, conf.level = 0.95)
@@
 However, as the rhyme task was less reliable than other measures, and showed only small asymmetries, we chose to examine the relationship between the LexTALE and the other verbal laterality measure, dichotic listening, instead.
+`r colFmt("For these exploratory analyses we included participants scoring less than 80 on the LexTALE in order to reduce the effect of a restricted range on the correlational analyses.",'red')`.
 There was no correlation between the LexTALE score and LIs on the dichotic listening at session 1 (*r*= `r round(DL.lex.est$estimate, 2)`, *95% CI* [`r round(DL.lex$conf.int[1], 2)`, `r round(DL.lex$conf.int[2], 2)`] *p*= `r round(p.adjust(DL.lex.est$p.value[1], n= 2, method= "bonferroni"), 3)`) or session 2 (*r*= `r round(DL.lex.est1$estimate, 2)`, *95% CI* [`r round(DL.lex1$conf.int[1], 2)`, `r round(DL.lex1$conf.int[2], 2)`] *p*= `r round(p.adjust(DL.lex.est1$p.value[1], n= 2, method= "bonferroni"), 3)`).
@@
 <!--- footedness etc also removed here ; could be some interest in looking at people with unusual laterality patterns on eye/foot/hand, but here it is too much--->
-
 ## Discussion
@@
 Each behavioural task yielded a hemispheric advantage in the expected direction, but the size of asymmetries and the test-retest reliability of the different tasks varied.
-A relatively small proportion of individuals showed significant asymmetry on the rhyme decision task.
-This task also yielded the poorest test-retest reliability, just falling short of our target minimum cutoff of *r* = .65.
+A relatively small proportion of individuals showed significant asymmetry on the rhyme decision task `r colFmt("despite the population as a whole being singificantly left lateralised.
+This pattern of results indicates that while very few participants showed statistically better performance for images in the right visual field (i.e.
+LI_z showed scores above |1.96|), many showed more subtle RVF advantages and had corresponding LI_z scores that were not statistically different from zero at the individual level.
+When comparing LI_z scores to zero at the population there will have been sufficient power to determine whether these small differences were different from zero.",'red')`  This task also yielded the poorest test-retest reliability, just falling short of our target minimum cutoff of *r* = .65.
 Compared to early studies (e.g.
@@
 One key difference between van der Haegen and Brysbaert’s and our study was that their language-based visual half-field tasks involved overt naming.
-Speech production is highly lateralised (Kosslyn, 1987) and requiring speech production should yield a hemispheric advantage.
+Speech production is highly lateralised `r colFmt("(e.g.
+Knetch et al., 2000)",'red')` and requiring speech production should yield a hemispheric advantage.
 Our task was designed to require covert speech, but an active speech production component might reveal stronger lateralisation.
 This would, however, be logistically difficult to implement online.
-The other main limitation of online testing is that the visual half-field method depends critically on presenting lateralised visual stimuli that project to one hemisphere.
+`r colFmt("It is, of course, important to reconcile this suggestion with other work.
+Recent studies have reported robust right visual-field advanatges without the use of active speech production (Brederoo et al., 2020; Hausmann et al., 2019; Willemin et al., 2019), highlighting that the relatively poor task validity and reliability likely stem from other methodological issues.",'red')`
+
+`r colFmt("A second explanation for the poor validity and reliability of the rhyme task relates to the control viewing conditions.",'red')` The visual half-field method depends critically on presenting lateralised visual stimuli that project to one hemisphere.
 Although we were able to calibrate screen size at the start of the session and to request participants to sit a certain distance from the screen, it is plausible this was not sufficient to achieve adequately lateralised presentation.
 While we can format images to be consistent across displays, we cannot control participants’ behaviour.
-Outside of the lab, participants can freely move their head which will influence the size of images on the retina, and it is difficult to ensure images are shown in foveal or parafoveal vision.
-One avenue that might be worth exploring is the use of images presented in peripheral vision, as it is unlikely that small head movements would result in the image shifting from peripheral vision.
+Outside of the lab, participants can freely move their head which will influence the size of images on the retina, and it is difficult to ensure that images are shown `r colFmt("exclusively in parafoveal vision.
+If on a subset of images are not equally displaced in parafoveal vision it could mean that processing is more efficient for the image in its corresponding visual field, reducing the validity of the task and adding considerable noise to reaction time estimates in each visual field.
+Without stict monitoring of eye-movements here it is not possible to control for this.",'red')` One alterante avenue that might be worth exploring is the use of images presented in peripheral vision, as it is unlikely that small head movements would result in the image shifting from peripheral vision.
 This increased difficultly in the processing of information in this region may also result in stronger visual field advantages.
@@
 (2013) using the iDichotic application, and numerically higher than that reported for the same dichotic listening task when run under lab conditions (Bless et al., 2013).
-We further showed (see Figure 4) that there was little decline in reliability with 85 trials, and with as few as 60 trials, the current implementation of the dichotic listening task yielded a test-retest correlation higher than *r* = .65, suggesting that those who wish to use this paradigm in the future could use a smaller number of trials.
+We further showed (see Figure 5) that there was little decline in reliability with 85 trials, and with as few as 60 trials, the current implementation of the dichotic listening task yielded a test-retest correlation higher than *r* = .65, suggesting that those who wish to use this paradigm in the future could use a smaller number of trials.
 It would be of interest to extend this line of work to use the modified version of dichotic listening described by Westerhausen and Samuelsen (2020), which had even higher reliability.
@@
 From the perspective of the participant, it is a straightforward task of reporting the emotion expressed in a single face and does not require any explicit comparison.
+`r colFmt("The finding that good test-retest reliability could be observed with a reduced number of trials will also be welcome for future studies.
+Participants in the current study recieved financial reimbursement for their participation and this almost certainly played a role in recruiting a large sample with relatively low rates of attrition.
+Unpaid volunteers would be unlikely to complete the battery as it stand.
+However, they may be more likely to participate in a more streamlined version which contained a reduced number of trials.",'red')`
 
@@
 Given the range of evidence now available, it may be time to abandon hypotheses that predict that different lateralised skills are either wholly interdependent or wholly independent.
-Evidence such as that from Gerrits et al (2020) confirms that lateralisation of different functions is associated far more than expected by chance, yet cases of dissociation are not uncommon.
+Evidence such as that from Gerrits et al.
+(2020) confirms that lateralisation of different functions is associated far more than expected by chance, yet cases of dissociation are not uncommon.
 Research in future might explore which functions are associated, and to what extent, rather than making a simple contrast between models that postulate a general answer to explain all of laterality.
+
+`r colFmt("On the issue of looking to explore associations between functions, Brederoo et al.
+(2020) reported data for 122 participants.
+In their study, participants completed a variety of visual half-field tasks covering a range of domains such as word processing, face processing, and spatial frequency processing.
+At the population level Brederoo et al.
+reported lateralisation in the expected direction for the majority of tasks.
+For left handers and those with right hemisphere dominant speech, there was evidence of a deviation from typical lateralistion.
+Contrary to the current study, Brederoo et al.
+reported partial evidence to suggest that processes recruiting similar resources will lateralise to homologue areas to promote intra-hemispheric proximity to other cortical areas within their processing hierarchies.
+In one analysis they found the lateralisation of faces and word recongition to be correlated but when they looked at lateralisation in those who were right handed versus those withright domaincnat language processing there was considerably less evidence for causual complimentarity.
+In contrast Brederoo et al.
+reported consistent evidence within the visual domain, where lateralisation low and high frequency information processing was dependent on one another while the lateralisation was independent to lateralisation across all other tasks.
+So it would appear by comparing tasks of a wider varity and looking to explore assocations in laterality across domains Brederoo et al.
+elucidate the issue of colateralisation and suggest that the relationships between lateralisation on some related tasks may be stronger when task demands are similar and share processing resources.",'red')`
 
@@
 ## Data accessibility
+
 This project was pre-registered on the Open Science Framework (OSF) before the start of data collection.
@@
 ## Competing Interests
+
 We declare we have no competing interests.
@@
 ## Funding
+
 This study was supported by an Advanced Grant from the European Research Council (694189).
@@
 ## Acknowledgements
-We would like to thank Dr Mike Burt (University of Durham, UK) for providing stimuli for the chimeric faces task.
-We would also like to thank Professor Kenneth Hugdahl (University of Bergen, Norway) for providing stimuli for the dichotic listening task.
+
+First, we would like to thank Mike Burt for providing stimuli for the chimeric faces task and Kenneth Hugdahl for providing stimuli for the dichotic listening task.
+Second, we would like to thank Marco Himstein, Marc Brysbaert, and Sanne Brederoo for their helpful comments on a prior manuscript.
+Third, we would like to thank Rogier Kievit for his advice regarding the modeling of covariances.
 Portions of this work were presented at the Online Meeting of the Experimental Psychology Society in July 2020.
@@
 
+Annett, M. (1998). Handedness and cerebral dominance: The right shift theory. *Journal of Neuropsychiatry and Clinical Neuroscience*, *10*, 459-469.
+
+Annett, M. (2002). *Handedness and Brain Asymmetry; the right shift theory*. Hove, Sussex; Psychology Press
+
 Anwyl-Irvine, A. L., Massonnié, J., Flitton, A., Kirkham, N., & Evershed, J. K. (2020). Gorilla in our midst: An online behavioral experiment builder. *Behavior research methods*, *52*, 388-407.
@@
 Badzakova-Trajkov, G., Häberling, I. S., Roberts, R. P., & Corballis, M. C. (2010). Cerebral asymmetries: complementary and independent processes. *PloS one*, *5*, e9682.
+
+Beaton, A. A., & Moseley, L. G. (1984). Anxiety and the measurement of handedness. *British Journal of Psychology*, *75*, 275-278.
 
@@
 
+Bless, J.
+J., Westerhausen, R., Torkildsen, J.
+V.
+K., Gudmundsen, M., Kompus, K., & Hugdahl, K.
+(2015).
+Laterality across languages: Results from a global dichotic listening study using a smartphone application.
+*Laterality: Asymmetries of Body, Brain and Cognition*, *20*, 434-452.
+
 Boles, D. B. (1987). Reaction time asymmetry through bilateral versus unilateral stimulus presentation. *Brain and Cognition*, *6*, 321–333.
@@
 Bourne, V. J. (2006). The divided visual field paradigm: Methodological considerations. *Laterality*, *11*, 373–393.
+
+Brederoo, S.
+G., Nieuwenstein, M.
+R., Cornelissen, F.
+W., & Lorist, M.
+M.
+(2019).
+Reproducibility of visual-field asymmetries: Nine replication studies investigating lateralization of visual information processing.
+*Cortex*, *111*, 100-126.
+
+Brederoo, S.
+G., Van der Haegen, L., Brysbaert, M., Nieuwenstein, M.
+R., Cornelissen, F.
+W., & Lorist, M.
+M.
+(2020).
+Towards a unified understanding of lateralized vision: A large-scale study investigating principles governing patterns of lateralization using a heterogeneous sample.
+*Cortex*.
+https://doi.org/10.1016/j.cortex.2020.08.029
 
@@
 
+Dundas, E.
+M., Plaut, D.
+C., & Behrmann, M.
+(2015).
+Variable left-hemisphere language and orthographic lateralization reduces right-hemisphere face lateralization.
+*Journal of Cognitive Neuroscience*, *27*, 913-925.
+
 Elias, L. J., Bulman-Fleming, M. B., & McManus, I. C. (1999). Visual temporal asymmetries are related to asymmetries in linguistic perception. *Neuropsychologia*, *37*, 1243-1249.
@@
 Hodgson, J. C., Hirst, R. J., & Hudson, J. M. (2016). Hemispheric speech lateralisation in the developing brain is related to motor praxis ability. *Developmental cognitive neuroscience*, *22*, 9-17.
+
+Hodgson, J. C., & Hudson, J. M. (2018). Speech lateralization and motor control. In G. S. Forrester, W. D. Hopkins, K. Hudry, & A. Lindell (Eds). *Progress in Brain Research* (pp. 145-178). Elsevier.
 
@@
 
-Kelley KS, Littenberg B. (2019). Dichotic Listening Test–retest reliability in children. *Journal of Speech, Language, and Hearing Research*, *62*,169-76.
+Kelley KS, Littenberg B. (2019). Dichotic Listening Test–retest reliability in children. *Journal of Speech, Language, and Hearing Research*, *62*, 169-76.
 
@@
 Kosslyn, S. M. (1987). Seeing and imagining in the cerebral hemispheres: a computational approach. *Psychological review*, *94*, 148.
+
+Knecht, S., Dräger, B., Deppe, M., Bobe, L., Lohmann, H., Flöel, A., Ringelstein, E.
+B., & Henningsen, H.
+(2000).
+Handedness and hemispheric language dominance in healthy humans.
+*Brain*, *123*, 2512-2518.
 
@@
 
+Steenhuis, R.
+E., & Bryden, M.
+P.
+(1999).
+The relation between hand preference and hand performance: What you get depends on what you measure.
+*Laterality: Asymmetries of Body, Brain and Cognition*, *4*, 3-26.
+
+Tiffin, J., & Asher, E., (1948). The Purdue pegboard; norms and studies of reliability and validity. *Journal of Applied Psychology*, *32*, 234-47.
+
 Van der Haegen, L., & Brysbaert, M. (2018). The relationship between behavioral language laterality, face laterality and language performance in left-handers. *PloS one*, *13*, e0208696.
+
+Van der Haegen, L., Cai, Q., Seurinck, R., & Brysbaert, M.
+(2011).
+Further fMRI validation of the visual half field technique as an indicator of language laterality: A large-group analysis.
+*Neuropsychologia*, *49*, 2879-2888.
 
@@
 
+Woodhead, Z.
+V.
+J., Thompson, P.
+A., Karlsson, E., & Bishop, D.
+V.
+M.
+(2020, September 9).
+Investigating multidimensionality of language lateralisation in left and right handed adults: an update on Woodhead et al.
+2019.
+https://doi.org/10.31234/osf.io/mrkgf
+
 Woods, K. J., Siegel, M. H., Traer, J., & McDermott, J. H. (2017). Headphone screening to facilitate web-based auditory experiments. *Attention, Perception, & Psychophysics*, *79*, 2064-2072.
@@
 
+<!--- FOOTNOTESS --->
+[^1]: Hugdahl et al.
+(2009) recommend excluding participants with a general hearing impairment of more than 20 dB for all frequencies tested, and/or an interaural hearing threshold difference of more than 10 dB as this can influence dichotic listening outcomes.
+It is important to note that this exclusion was not possible in the current study as we did not include a hearing test.
+With the current sample size, subtle hearing deficits in one ear will unlikely undermine the results.
+However, it will be important for future studies to incorporate hearing accuracy tests in order to exclude participants with these hearing related difficulties.
+[^2]: In line with our pre-registration, outliers were identified using data pooled across visual fields/ears.
+We explored the possibility that a different pattern of results could have been obtained by identifying outliers separately for each visual field/ear.
+While reliability estimates remained unchanged for each measure very small differences were observed across analyses.
+Despite these small differences the pattern of results and conclusions we form remain unchanged.
+
+```{r dist_plots,echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
+# rename
+RD.dat$handed3 <- as.factor(RD.dat$handed3)
+  levels(RD.dat$handed3) <- c("Extreme LH", 'Moderate LH', "RH")
+CF.dat$handed3 <- as.factor(CF.dat$handed3)
+  levels(CF.dat$handed3) <- c("Extreme LH", 'Moderate LH', "RH")
+DL.dat$handed3 <- as.factor(DL.dat$handed3)
+  levels(DL.dat$handed3) <- c("Extreme LH", 'Moderate LH', "RH")
+FT.dat$handed3 <- as.factor(FT.dat$handed3)
+  levels(FT.dat$handed3) <- c("Extreme LH", 'Moderate LH', "RH")
+# plot all
+plota <- ggplot(RD.dat, aes(x=tlat_RDT_1, fill=handed3, color=handed3)) +
+  geom_histogram(position="identity", alpha= .35) + theme_classic(20) +
+  xlab('Rhyme decision LI_z (day 1)') + ggtitle("Rhyme decision (day 1)") + 
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(legend.position = "top", legend.title = element_blank()) + xlim(-5,5) + ylim(0, 65)
+plotb <- ggplot(RD.dat, aes(x=tlat_RDT_2, fill=handed3, color=handed3)) +
+  geom_histogram(position="identity", alpha= .35) + theme_classic(20) +
+  xlab('Rhyme decision LI_z (day 2)') + ggtitle("Rhyme decision (day 2)") + 
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(legend.position = "top", legend.title = element_blank()) + xlim(-5,5) + ylim(0, 65)
+plotc <- ggplot(DL.dat, aes(x=Day1_Dich_acc_Z, fill=handed3, color=handed3)) +
+  geom_histogram(position="identity", alpha= .35) + theme_classic(20) +
+  xlab('Dichotic listening LI_z (day 1)') + ggtitle("Dichotic listening (day 1)") + 
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(legend.position = "top", legend.title = element_blank()) + xlim(-20,20) + ylim(0, 65)
+plotd <- ggplot(DL.dat, aes(x=Day2_Dich_acc_Z, fill=handed3, color=handed3)) +
+  geom_histogram(position="identity", alpha= .35) + theme_classic(20) +
+  xlab('Dichotic listening LI_z (day 2)') + ggtitle("Dichotic listening (day 2)") + 
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(legend.position = "top", legend.title = element_blank()) + xlim(-20,20) + ylim(0, 65)
+plote <- ggplot(CF.dat, aes(x=Day1_CF_acc_Z, fill=handed3, color=handed3)) +
+  geom_histogram(position="identity", alpha= .35) + theme_classic(20) +
+  xlab('Chimeric faces LI_z (day 1)') + ggtitle("Chimeric faces (day 1)") + 
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(legend.position = "top", legend.title = element_blank()) + xlim(-20,20) + ylim(0, 65)
+plotf <- ggplot(CF.dat, aes(x=Day2_CF_acc_Z, fill=handed3, color=handed3)) +
+  geom_histogram(position="identity", alpha= .35) + theme_classic(20) +
+  xlab('Chimeric faces LI_z (day 2)') + ggtitle("Chimeric faces (day 2)") + 
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(legend.position = "top", legend.title = element_blank()) + xlim(-20,20) + ylim(0, 65)
+plotg <- ggplot(FT.dat, aes(x=Day1_Finger_acc_Z, fill=handed3, color=handed3)) +
+  geom_histogram(position="identity", alpha= .35) + theme_classic(20) +
+  xlab('Finger tapping LI_z (day 1)') + ggtitle("Finger tapping (day 1)") + 
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(legend.position = "top", legend.title = element_blank()) + xlim(-5,5) + ylim(0, 65)
+ploth <- ggplot(FT.dat, aes(x=Day2_Finger_acc_Z, fill=handed3, color=handed3)) +
+  geom_histogram(position="identity", alpha= .35) + theme_classic(20) +
+  xlab('Finger tapping LI_z (day 2)') + ggtitle("Finger tapping (day 2)") + 
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(legend.position = "top", legend.title = element_blank()) + xlim(-5,5) + ylim(0, 65)
+# plot together
+all.plots <- ggpubr::ggarrange(plota, plotb, plotc, plotd, plote, plotf, plotg, ploth, ncol = 2, nrow= 4)
+ggsave("./plots/dist.plots.png", plot=all.plots,dpi = 300, width = 12, height = 24,)
+```
+
 \newpage
-#### Supplemental material
-
-Figure S1 shows scatterplots indicating scores on the left and right sides for each task, with self-reported handedness (left vs right) colour-coded.
-For each task, the mean for each side is shown as vertical and horizontal dotted lines, and equal performance on the two sides corresponds to the sloping red dotted line: thus an impression of whether the task is lateralised can be obtained by considering whether points cluster around the sloping lines symmetrically.
-
-![Figure S1.
-Left side (L) vs right side (R) scores for sessions 1 and 2 for the behavioural laterality tasks.
-The diagonal dotted line shows the point of equality: points above this are higher for R than L, and points below are higher for L than R.
-Mean values for R and L are shown as horizontal and vertical dotted lines.
-The colour of the points indicates the participants’ handedness (black = right-handed; red = left-handed).
-](all.raw.png)
-
-For the rhyme decision task, only correct responses were used for computing mean log RTs (shown transformed back to the original ms scale in the plot).
-For dichotic listening and chimeric faces, a correct response can be given from either side; it follows that there will be a negative correlation between the two sides reflecting the fact that the number of selections on the left will be inversely related to the number of correct responses on the right.
-For these tasks, it is evident by inspection that most responses for dichotic listening cluster above the dotted line, indicating a right ear advantage, whereas most responses on chimeric faces cluster below the dotted line, indicating a left hemi-face advantage.
-Finally, we show the number of correct sequences with each hand on the finger tapping task.
-Here, a strong impact of handedness is evident on inspection, with left-handers generally more accurate with the left hand, and right-handers more accurate with the right hand.
+
+\begin{center}
+
+Assessing the reliability of an online behavioural laterality battery: a pre-registered study.
+
+Parker, A. J., Woodhead, Z. V. J., Thompson, P. A., Bishop, D. V. M.
+
+Supplementary Materials
+
+\end{center}
+
+In these Supplementary Materials we include the distribution of laterality indices for each behavioural laterality task across both testing sessions.
+We also include a series of t-tests examining the difference in LI_z scores between extreme left-handers and right-handers.
+
+**LI_z distributions**
+
+the disributions of LI_z scores are shown in Figure S1.
+
+![Distribution of laterality indices on the behavioural tasks.
+The distributions of LI_z are show in pink for extreme left-handers, green for moderate left-handers, and blue for right-handers.
+Negative values indicates left lateralisation at the brain level and positive values indicate left lateralisation at the brain level.](./plots/dist.plots.png){width=60%}
+
+**LI_z scores in extreme left-handers and right-handers**
+```{r hand.corrs,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
+comp_c <- ggplot(subset(LIs, handed3== "Right-hander" | handed3=="Extreme left-hander"),
+  aes(x= handed3, y= Day1_CF_acc_Z, fill= handed3)) +
+  geom_hline(yintercept=0) + 
+  geom_hline(yintercept=1.96, linetype= "dashed", color= "red") +
+  geom_hline(yintercept=-1.96, linetype= "dashed", color= "red") +
+  geom_violin(alpha= .1, size= .5) +
+  geom_jitter(width= .1, aes(color= handed3), alpha= .25) +
+  geom_boxplot(width=0.1, position=position_dodge(0.9), size= .5 , notch = T, notchwidth = 0.4, varwidth =F,
+               color= 'black', show.legend=FALSE, linetype= 'solid', outlier.color = NA) +
+  ylab("Chimeric faces LI_z") + 
+  xlab(" ") +
+  ggtitle("(C)") + theme(plot.title = element_text(hjust = 0.5)) +
+  ylim(-20, 20) +
+  theme_classic(18) +
+  theme(legend.position= "none") + theme(plot.title = element_text(hjust = 0.5))
+comp_b <- ggplot(subset(LIs, handed3== "Right-hander" | handed3=="Extreme left-hander"),
+  aes(x= handed3, y= Day1_Dich_acc_Z, fill= handed3)) +
+  geom_hline(yintercept=0) + 
+  geom_hline(yintercept=1.96, linetype= "dashed", color= "red") +
+  geom_hline(yintercept=-1.96, linetype= "dashed", color= "red") +
+  geom_violin(alpha= .1, size= .5) +
+  geom_jitter(width= .1, aes(color= handed3), alpha= .25) +
+  geom_boxplot(width=0.1, position=position_dodge(0.9), size= .5 , notch = T, notchwidth = 0.4, varwidth =F,
+               color= 'black', show.legend=FALSE, linetype= 'solid', outlier.color = NA) +
+  ylab("Dichotic listening LI_z") +
+  xlab(" ") +
+  ggtitle("(B)") + theme(plot.title = element_text(hjust = 0.5)) +
+  ylim(-20, 20) +
+  theme_classic(18) +
+  theme(legend.position= "none") + theme(plot.title = element_text(hjust = 0.5))
+comp_a <- ggplot(subset(LIs, handed3== "Right-hander" | handed3=="Extreme left-hander"),
+  aes(x= handed3, y= tlat_RDT_1, fill= handed3)) +
+  geom_hline(yintercept=0) + 
+  geom_hline(yintercept=1.96, linetype= "dashed", color= "red") +
+  geom_hline(yintercept=-1.96, linetype= "dashed", color= "red") +
+  geom_violin(alpha= .1, size= .5) +
+  geom_jitter(width= .1, aes(color= handed3), alpha= .25) +
+  geom_boxplot(width=0.1, position=position_dodge(0.9), size= .5 , notch = T, notchwidth = 0.4, varwidth =F,
+               color= 'black', show.legend=FALSE, linetype= 'solid', outlier.color = NA) +
+  ylab("Rhyme decision LI_z") +
+  xlab(" ") +
+  ggtitle("(A)") + 
+  ylim(-20, 20) +
+  theme_classic(18) +
+  theme(legend.position= "none") + theme(plot.title = element_text(hjust = 0.5))
+t.plots <- ggpubr::ggarrange(comp_a, comp_b, comp_c, ncol = 1, nrow= 3)
+ggsave("./plots/t.plots.png", plot=t.plots,dpi = 300, width = 5, height = 15,)
+# tests
+t1 <- t.test(data= subset(LIs, handed3== "Right-hander" | handed3=="Extreme left-hander"), tlat_RDT_1~ handed3, alternate= "less")
+t2 <- t.test(data= subset(LIs, handed3== "Right-hander" | handed3=="Extreme left-hander"), Day1_DL_acc_LI~ handed3, alternate= "less")
+t3 <- t.test(data= subset(LIs, handed3== "Right-hander" | handed3=="Extreme left-hander"), Day1_CF_acc_Z~ handed3, alternate= "greater")
+```
+
+To further examine the link between handedness and atypical laterality, we conducted a series of *t*-tests using session 1 data.
+The distribution of LI_z scores are shown separately in Figure S2.
+A one-tailed independent-samples *t*-tests indicated that LI_z scores did not differ between extreme left-handers and right-handers on the rhyme decision, *t*(`r round(t1$parameter, 2)`)= `r round(t1$statistic, 2)`, *95% CI*= [`r round(t1$conf.int[1], 2)`, `r round(t1$conf.int[2], 2)`], *p*= `r round(t1$p.value, 3)`.
+This is not surprising given that very few participants were lateralised on this task.
+While LI_z scores did not differ between extreme left-handers and right-handers on the rhyme decision, they did on the dichotic listening, *t*(`r round(t2$parameter, 2)`)= `r round(t2$statistic, 2)`, *95% CI*= [`r round(t2$conf.int[1], 2)`, `r round(t2$conf.int[2], 2)`], *p*= `r round(t2$p.value, 3)`, and chimeric faces, *t*(`r round(t3$parameter, 2)`)= `r round(t3$statistic, 2)`, *95% CI*= [`r round(t3$conf.int[1], 2)`, `r round(t3$conf.int[2], 2)`], *p*= `r round(t3$p.value, 3)`.
+These differences demostrate that, as a population, extreme left-handers were less lateralised.
+
+![Violin plots showing the distribution of laterality indices on the behavioural tasks for extreme left-handers (in pink) and right-handers (in blue).
+The notch on the boxplot indicates the mean per condition.
+The red dashed lines are drawn at -1.96 and 1.96, indicating significant lateralisation to either the left or right hemisphere.
+Data points represent indiciaul participants' data.](./plots/t.plots.png){width=40%}

### 2020-11-18 battery-preprocessing_v6_DB_split2.Rmd
split half

--- 2020-07-20 battery-preprocessing_v6_DB.Rmd
@@
 ## Updates
+17 Nov - changed some 'select' commands to dplyr::select
+Also ensured that 'merge' selected ALL
+
+Modified to allow rerun on split half.
+For this we take blocks of 4 rows, e.g. 1,2,3,4,9,10,11,12 etc.
+This way the script will work for finger tapping as well as the other tasks.
+Idea is to do this selection of rows right at the outset on the raw file.
 
@@
 # now merge sightedness and allsum
-allsum <- merge(allsum, sightedness, by= "subject")
+allsum <- merge(allsum, sightedness, by= "subject",all=TRUE)
 head(allsum) #just to check all looking OK
@@
 # now create a data frame only for index 
-lexTALE <- select(lexTale_score, "subject", "lexTALE")
+lexTALE <- dplyr::select(lexTale_score, "subject", "lexTALE")
 # now merge with allsum
-allsum <- merge(allsum, lexTALE, by= "subject")
+allsum <- merge(allsum, lexTALE, by= "subject",all=TRUE)
 # now write the full allsum 
@@
 # Set directory
-for (d in 1:2){
+
   data_file <- "https://osf.io/mhk29/download" #day 1 data
-
-  if (d ==2){
+  all_dat1 <- read.csv(data_file, na.strings = "NA",stringsAsFactors=F)
+
     data_file <-  "https://osf.io/ebm6s/download"
-  }
-  #we will assign raw data to all_dat2 (day2), but then reassign to all_dat1 on first run through loop, so we have correct data for each day
      all_dat2 <- read.csv(data_file, na.strings = "NA",stringsAsFactors=F)
-# cleaning - do operations that apply for all tasks
-all_dat2 %>% filter(Attempt >= 1) # only answers 
+
+   all_dat1$Session<-1
+   all_dat2$Session<-2
+
+     #need to align cols so can combine
+
+      all_dat2<-all_dat2[,-1]  #extraneous first column
+      all_dat1<-all_dat1[,-56] #extraneous X56 col
+      all_dat2<-all_dat2[,-64] #extraneous X56 col
+
+      alldat<-rbind(all_dat1,all_dat2[,colnames(all_dat1)])
+
+alldat %>% filter(Attempt >= 1) # only answers 
 # write as factors
-all_dat2$subject <- as.factor(all_dat2$Participant.Public.ID)
-all_dat2$RT <- as.numeric(as.character(all_dat2$Reaction.Time))
-all_dat2$Trial <-as.numeric(all_dat2$Trial.Number) #useful if we want to restrict analysis to subset of trials so retain this variable
-all_dat2$Task.Name <- as.factor(all_dat2$Task.Name)
-all_dat2$word<-substring(all_dat2$word,66,(nchar(all_dat2$word)-4)) #remove format from written word stimuli
-if (d==1)
-{all_dat1 <- all_dat2}
-}
-
-all_dat2$word_OG<-NA #create dummy column just to make it easier to align columns
-collist <- colnames(all_dat1)
-all_dat2 <- all_dat2[,collist]
+alldat$subject <- as.factor(alldat$Participant.Public.ID)
+alldat$RT <- as.numeric(as.character(alldat$Reaction.Time))
+alldat$Trial <-as.numeric(alldat$Trial.Number) #useful if we want to restrict analysis to subset of trials so retain this variable
+alldat$Task.Name <- as.factor(alldat$Task.Name)
+alldat$word<-substring(alldat$word,66,(nchar(alldat$word)-4)) #remove format from written word stimuli
+oddeven <- 1 #set to zero for full dataset; to 1 for odds, and to 2 for evens
+nr<-nrow(alldat)
+
+if (oddeven==1){
+  alldat<-alldat[alldat$Trial%%2==1,] #odd trials only
+}
+if (oddeven==2){
+  alldat<-alldat[alldat$Trial%%2==0,] #even trials only
+}
+
+alldat$word_OG<-NA #create dummy column just to make it easier to align columns
+
 ```
@@
 allsum <- allsum %>% rename(origID = subject) 
-all_dat1 <- all_dat1 %>% rename(origID = subject) 
-all_dat2 <- all_dat2 %>% rename(origID = subject) 
-all_dat1$subject <- NA
-all_dat2$subject <- NA
+alldat <- alldat %>% rename(origID = subject)
+
+alldat$subject <- NA
+
 for (n in 1:nrow(allsum)){
   allsum$subject[n] <- n
-  w<-which(all_dat1$origID == allsum$origID[n])
-  all_dat1$subject[w] <-n
-  w<-which(all_dat2$origID == allsum$origID[n])
-  all_dat2$subject[w] <-n
+  w<-which(alldat$origID == allsum$origID[n])
+  alldat$subject[w] <-n
+
 }
@@
 # create compact data frame ; one row for each subject, subjects stacked, L above R
-temp_dat <- data.frame(matrix(ncol = 4, nrow = nsub*nloop))
+temp_dat <- data.frame(matrix(ncol = length(varlist), nrow = nsub*nloop))
+#Previously just 4 cols but now including participant ID to check this is consistent
 colnames(temp_dat) <- varlist
@@
 # now create a data frame for lat index 
-Lsubjectf <- select(Lsubjectf, "subject", "LI")
+Lsubjectf <- dplyr::select(Lsubjectf, "subject", "LI")
 # now merge with allsum
@@
 # now create a data frame for lat index 
-Zdf <- select(Zdf, "subject", "Z")
+Zdf <- dplyr::select(Zdf, "subject", "Z")
 # now merge with allsum
@@
 for (d in 1:2){ #we will process day 1 and day 2 together
-  origfile <- all_dat1
-  if (d==2)
-  {origfile <- all_dat2}
+  origfile <- alldat[alldat$Session==d,]
+
 #As with earlier reading in, we default to saving CF_dat2 (day2), but at end of the chunk we reassign to CF_day1 in the first run through the d loop
@@
 #db modified to mark rather than drop cases with scores < .75
-allsum$exclude <- 0 #default is to exclude
+allsum$exclude <- 0 #default is to include
 
@@
 allsum$exclude[w]<-1 #code exclude = 1 if excluded because poor emot recog
+
 print(paste('N flagged for exclusion because < 75% on at least one trial:',length(w)))
-
+w <- union(which(allsum$Emot.corr.diff1 ==0),which(allsum$Emot.corr.diff2 == 0)) #also exclude if no correct response on one side
+allsum$exclude[w]<-2 #code exclude = 2 if excluded because no response one side
 ```
@@
 inclsubs <- allsum$subject[w]
+
+#-------------------------
+#We need to add a step to exclude accuracy and RT if fewer than 3 trials in one condition.
+
+w<-which(chim_dat1$N<2)
+if(length(w)>0){
+dropsub <- chim_dat1$subject[w]
+chim_dat1$accurate[chim_dat1$subject %in% dropsub]<-NA
+chim_dat1$RT[chim_dat1$subject %in% dropsub]<-NA
+}
+w<-which(chim_dat2$N<2)
+if(length(w)>0){
+dropsub <- chim_dat2$subject[w]
+chim_dat2$accurate[chim_dat2$subject %in% dropsub]<-NA
+chim_dat2$RT[chim_dat2$subject %in% dropsub]<-NA
+}
 
@@
 #Now repeat for day 2
-#-------------------------
-#We need to add a step to exclude accuracy and RT if fewer than 3 trials in one condition. This only affects 2nd session data, so this code is specific to that file
-w<-which(chim_dat2$N<2)
-dropsub <- chim_dat2$subject[w]
-chim_dat2$accurate[chim_dat2$subject %in% dropsub]<-NA
-chim_dat2$RT[chim_dat2$subject %in% dropsub]<-NA
+
 #-------------------------
@@
 #added by DB 17th July 2020: include raw accuracy L and R in the allsum dataframe
-chimdatL1<-select(filter(chim_dat1,side=='Left'),subject,accurate)
-chimdatR1<-select(filter(chim_dat1,side=='Right'),subject,accurate)
-chimdatL2<-select(filter(chim_dat2,side=='Left'),subject,accurate)
-chimdatR2<-select(filter(chim_dat2,side=='Right'),subject,accurate)
-
-allsum <- merge(allsum,chimdatL1,by="subject")
-allsum <- merge(allsum,chimdatR1,by="subject")
-allsum <- merge(allsum,chimdatL2,by="subject")
-allsum <- merge(allsum,chimdatR2,by="subject")
+chimdatL1<-dplyr::select(filter(chim_dat1,side=='Left'),subject,accurate)
+chimdatR1<-dplyr::select(filter(chim_dat1,side=='Right'),subject,accurate)
+chimdatL2<-dplyr::select(filter(chim_dat2,side=='Left'),subject,accurate)
+chimdatR2<-dplyr::select(filter(chim_dat2,side=='Right'),subject,accurate)
+
+allsum <- merge(allsum,chimdatL1,by="subject",all=TRUE)
+allsum <- merge(allsum,chimdatR1,by="subject",all=TRUE)
+allsum <- merge(allsum,chimdatL2,by="subject",all=TRUE)
+allsum <- merge(allsum,chimdatR2,by="subject",all=TRUE)
 myn <-ncol(allsum)
@@
 for (d in 1:2){ #we will process day 1 and day 2 together
-  origfile <- all_dat1
-  if (d==2)
-  {origfile <- all_dat2}
+  origfile <- alldat[alldat$Session==d,]
+
 # lowercase for responses
@@
 #added by DB 17th July 2020: include raw accuracy L and R in the allsum dataframe
-DLdatL1<-select(filter(DL_dat1,side=='Left'),subject,accurate)
-DLdatR1<-select(filter(DL_dat1,side=='Right'),subject,accurate)
-DLdatL2<-select(filter(DL_dat2,side=='Left'),subject,accurate)
-DLdatR2<-select(filter(DL_dat2,side=='Right'),subject,accurate)
+DLdatL1<-dplyr::select(filter(DL_dat1,side=='Left'),subject,accurate)
+DLdatR1<-dplyr::select(filter(DL_dat1,side=='Right'),subject,accurate)
+DLdatL2<-dplyr::select(filter(DL_dat2,side=='Left'),subject,accurate)
+DLdatR2<-dplyr::select(filter(DL_dat2,side=='Right'),subject,accurate)
 
 #Add to allsum
-allsum <- merge(allsum,DLdatL1,by="subject")
-allsum <- merge(allsum,DLdatR1,by="subject")
-allsum <- merge(allsum,DLdatL2,by="subject")
-allsum <- merge(allsum,DLdatR2,by="subject")
+allsum <- merge(allsum,DLdatL1,by="subject",all=TRUE)
+allsum <- merge(allsum,DLdatR1,by="subject",all=TRUE)
+allsum <- merge(allsum,DLdatL2,by="subject",all=TRUE)
+allsum <- merge(allsum,DLdatR2,by="subject",all=TRUE)
 myn <-ncol(allsum)
@@
 
-  origfile <- all_dat1
-  if(d==2)
-  {origfile <- all_dat2}
+  origfile <- alldat[alldat$Session==d,]
+
   task <- "Rhyme Detection Task"
@@
 #We create RDT files for day1 and day 2
-varlist <-   c("subject", "side", "accurate", "RT") #need to be in this order, ie sub, side, acc and RR
+varlist <-   c("subject", "side", "accurate", "RT","Participant.Public.ID") #need to be in this order, ie sub, side, acc and RR
 latlist <- c("left","no rhyme","right") #names of factor levels
@@
 
+#Also record % correct for each category in allsum
+for (d in 1:2){
+  myfile<-RDT_df1
+  if(d==2){ myfile<-RDT_df2}
+pcorL<-myfile[myfile$side=='left',]
+pcorR<- myfile[myfile$side=='right',]
+pcorNull <- myfile[myfile$side=='no rhyme',]
+pcorL<-dplyr::select(pcorL,subject,p.corr)
+pcorR<-dplyr::select(pcorR,subject,p.corr)
+pcorNull<-dplyr::select(pcorNull,subject,p.corr)
+colnames(pcorL)[2]<-paste0('RDT',d,'.pcorrL')
+colnames(pcorR)[2]<-paste0('RDT',d,'.pcorrR')
+colnames(pcorNull)[2]<-paste0('RDT',d,'.pcorrNull')
+allsum <- merge(allsum,pcorL,by="subject",all=TRUE)
+allsum <- merge(allsum,pcorR,by="subject",all=TRUE)
+allsum <- merge(allsum,pcorNull,by="subject",all=TRUE)
+}
+
 #Scrutiny of these reveal a few cases who missed all 'no rhyme' cases.
@@
 ```{r RDT_exclude}
+
+w1x<-which(is.na(RDT_df1$RT))
+w2x<-which(is.na(RDT_df2$RT))
+
+exRDTx<-unique(c(RDT_df1$subject[w1x],RDT_df2$subject[w2x]))
+
+#find those who never got the no rhyme correct
+nr1<-RDT_df1[RDT_df1$side=='no rhyme',] 
+nr2<-RDT_df2[RDT_df2$side=='no rhyme',]
+
+w1<-which(nr1$p.corr==0)
+w2<-which(nr2$p.corr==0)
+
+exRDT<-unique(c(nr1$subject[w1],nr2$subject[w2]))
+
+wr1<-RDT_df1[RDT_df1$side!='no rhyme',] 
+wr2<-RDT_df2[RDT_df2$side!='no rhyme',]
+
+w1a<-which(wr1$p.corr<75)
+w2a<-which(wr2$p.corr<75)
+
+exRDT2<-unique(c(wr1$subject[w1a],wr2$subject[w2a]))
+exRDT<-unique(c(exRDT,exRDT2))
+exRDT<-sort(exRDT)
+
+allsum$exRDT <- 0
+
+allsum$exRDT[allsum$subject %in% exRDT]<-1 #decided best to have separate column for exclusion for each task
+allsum$exRDT[allsum$subject %in% exRDTx]<-2 #missing data
+w<-which(allsum$exRDT==0)  
+incRDT <- allsum$subject[w]
+ninc <- length(incRDT)
+print(paste0('RDT: number remaining after exclusions = ',ninc))
+
+RDT_df1<-RDT_df1[RDT_df1$subject %in% incRDT,]
+RDT_df2<-RDT_df2[RDT_df2$subject %in% incRDT,]
+
+RDT_df1<-RDT_df1[RDT_df1$side !='no rhyme',]
+RDT_df2<-RDT_df2[RDT_df2$side !='no rhyme',]
+
+```
+```{r oldRDT_exclude}
+#Original version now superseded
+doold<-0
+if(doold==1){
   wn <-unique(which(RDT_df1$accurate ==0),which(RDT_df2$accurate ==0)) #this catches those who were always wrong on 'no rhyme' trials, and possibly others where presentation messed up for technical reasons
@@
 
-exRDT<-unique(c(RDT_df1$subject[w1],RDT_df2$subject[w2],RDT_df2$subject[wn]))
+exRDT<-sort(unique(c(RDT_df1$subject[w1],RDT_df2$subject[w2],RDT_df2$subject[wn])))
 
@@
 print(paste0('RDT: number remaining after exclusions = ',ninc))
+
+}
 ```
@@
 #added by DB 17th July 2020: include raw accuracy L and R in the allsum dataframe
-RDTdatL1<-select(filter(RDT_df1,side=='left'),subject,RT)
-RDTdatR1<-select(filter(RDT_df1,side=='right'),subject,RT)
-RDTdatL2<-select(filter(RDT_df2,side=='left'),subject,RT)
-RDTdatR2<-select(filter(RDT_df2,side=='right'),subject,RT)
-
-allsum <- merge(allsum,RDTdatL1,by="subject")
-allsum <- merge(allsum,RDTdatR1,by="subject")
-allsum <- merge(allsum,RDTdatL2,by="subject")
-allsum <- merge(allsum,RDTdatR2,by="subject")
+RDTdatL1<-dplyr::select(filter(RDT_df1,side=='left'),subject,RT)
+RDTdatR1<-dplyr::select(filter(RDT_df1,side=='right'),subject,RT)
+RDTdatL2<-dplyr::select(filter(RDT_df2,side=='left'),subject,RT)
+RDTdatR2<-dplyr::select(filter(RDT_df2,side=='right'),subject,RT)
+
+allsum <- merge(allsum,RDTdatL1,by="subject",all=TRUE)
+allsum <- merge(allsum,RDTdatR1,by="subject",all=TRUE)
+allsum <- merge(allsum,RDTdatL2,by="subject",all=TRUE)
+allsum <- merge(allsum,RDTdatR2,by="subject",all=TRUE)
 myn <-ncol(allsum)
@@
 origdat$side <- as.factor(origdat$side)
-levels(origdat$side) <-c('Left','No rhyme','Right') #function assumes capitalised
+levels(origdat$side) <-c('Left','Right') #function assumes capitalised
 mycolname<-"Day1.RDT_p.corr_LI"
@@
   }
+    w<-which(is.na(allsum$Day2.RDT_RT_LI)) #remove those with NA in LI
+  allsum$tempt[w]<-NA
   wc<-which(colnames(allsum)=='tempt')
+
   colnames(allsum)[wc]<-paste0('tlat_RDT_',d)
@@
 }
-
-w<-which(is.na(allsum$Day1.RDT_RT_LI)) #remove those with NA in LI
-allsum$tlat_RDT_1[w]<-NA
-w<-which(is.na(allsum$Day2.RDT_RT_LI))
-allsum$tlat_RDT_2[w]<-NA
 
@@
 plot(allsum$Day1.RDT_RT_LI,allsum$tlat_RDT_1,xlim=c(-1.5,1.5),ylim=c(-5,5))
+#NB t test and LI will disagree because same size of difference will give different t depending on how much variance there is for L and R responses.
 ```
@@
 
-#compare t-value from day 1 and day 2 
+#compare z-value from day 1 and day 2 
 myr <- cor(allsum$zlat_RDT_1,allsum$zlat_RDT_2,use='complete.obs')
@@
 ```{r readdataFinger, warning=FALSE}
+if(oddeven==0){ #won't work with split half
 for (d in 1:2){
-finger <- all_dat2[all_dat2$Task.Name=='Finger Tapping',]
-if(d==1){
-finger <- all_dat1[all_dat1$Task.Name=='Finger Tapping',]}
+finger <- alldat[alldat$Task.Name=='Finger Tapping',]
+finger <- finger[finger$Session==d,]
 # cleaning
@@
 complete.day2.FT <- unique(all_finger2$subject)
+}
 ```
@@
 ```{r readdataFinger2, warning=FALSE}
+if(oddeven==0){
 for (d in 1:2){ #we will process day 1 and day 2 together
@@
 }
+}
 ```
@@
 # We can go directly to creation of a data.frame by subjects, and use string matching to identify N correct seqs
-
+if (oddeven==0){
 myseq <-c('tumb','wrvx') # R, # L
@@
 
+}
 ```
@@
 ```{r dosummary2, warning = F}
+if(oddeven==0){
 origdat <- finger.long
@@
 dopirate(origdat,task,measure)
+}
 ```
 ```{r add.fingerL_R}
+if(oddeven==0){
 #added by DB 17th July 2020: include raw accuracy L and R in the allsum dataframe
-fingerL1<-select(finger.df,subject,L1)
-fingerR1<-select(finger.df,subject,R1)
-fingerL2<-select(finger.df,subject,L2)
-fingerR2<-select(finger.df,subject,R2)
-
-allsum <- merge(allsum,fingerL1,by="subject")
-allsum <- merge(allsum,fingerR1,by="subject")
-allsum <- merge(allsum,fingerL2,by="subject")
-allsum <- merge(allsum,fingerR2,by="subject")
+fingerL1<-dplyr::select(finger.df,subject,L1)
+fingerR1<-dplyr::select(finger.df,subject,R1)
+fingerL2<-dplyr::select(finger.df,subject,L2)
+fingerR2<-dplyr::select(finger.df,subject,R2)
+
+allsum <- merge(allsum,fingerL1,by="subject",all=TRUE)
+allsum <- merge(allsum,fingerR1,by="subject",all=TRUE)
+allsum <- merge(allsum,fingerL2,by="subject",all=TRUE)
+allsum <- merge(allsum,fingerR2,by="subject",all=TRUE)
 myn <-ncol(allsum)
 colnames(allsum)[(myn-3):myn] <- c('finger.L1','finger.R1','finger.L2','finger.R2')
-
+}
 ```
@@
 ```{r fingerLI}
+if(oddeven==0){
 task<-'Finger'
@@
 text(-10,10,paste0('r = ',round(cor1,3)))
+}
 ```
@@
 ```{r domakeZ_finger}
-
+if(oddeven==0){
 task<-'Finger'
@@
 text(-2,3.5,paste0('r = ',round(cor1,3)))
-
-write.csv(allsum, "LIs.csv")
-```
-
-### list of participants who completed.
-
-We need a list of participants here who completed the both days for later filtering.
-
-```{r lister}
-# CF
-day2.CF <- data.frame(matrix(nrow=length(complete.day2.CF), ncol= 2))
-  colnames(day2.CF) <- c("task", "subject")
-  day2.CF$subject <- complete.day2.CF
-  day2.CF$task <- "CF"
-# DL
-day2.DL <- data.frame(matrix(nrow=length(complete.day2.DL), ncol= 2))
-  colnames(day2.DL) <- c("task", "subject")
-  day2.DL$subject <- complete.day2.DL
-  day2.DL$task <- "DL"
-# RD
-day2.RD <- data.frame(matrix(nrow=length(complete.day2.RD), ncol= 2))
-  colnames(day2.RD) <- c("task", "subject")
-  day2.RD$subject <- complete.day2.RD
-  day2.RD$task <- "CF"
-# Finger
-day2.FT <- data.frame(matrix(nrow=length(complete.day2.FT), ncol= 2))
-  colnames(day2.FT) <- c("task", "subject")
-  day2.FT$subject <- complete.day2.FT
-  day2.FT$task <- "CF"
-# now create a combined df
-complete.day2a <- merge(day2.CF, day2.DL, by= "subject")
-complete.day2b <- merge(day2.RD, day2.FT, by= "subject")
-complete.day2 <- merge(complete.day2a, complete.day2b, by= "subject")
-# add row to code
-complete.day2$day2.complete <- 1
-complete.day2 <- select(complete.day2, -2, -3, -4, -5)
-# now merge into allsum
-allsum <- merge(allsum, complete.day2, by= "subject", all= TRUE)
-
-# now write this for filtering
-write.csv(allsum, "./data/allsum.csv")
-```
-
-### data frames for general plots
-
-In the MS we add a section on raw scores (simple plots). To do this, we require a data frame for each measure.
-
-```{r dataMS}
-# DAY 1
-# add task ID
-chim_dat1$task <- "CF"
-DL_dat1$task <- "DL"
-RDT_df1$task <- "RD"
-# merge and add day
-plot_data_1 <- rbind(chim_dat1, DL_dat1, RDT_df1[RDT_df1$side != "no rhyme",])
-  plot_data_1$Session<- 1
-
-# DAY 2
-chim_dat2$task <- "CF"
-DL_dat2$task <- "DL"
-RDT_df2$task <- "RD"
-# merge and add day
-plot_data_2 <- rbind(chim_dat2, DL_dat2, RDT_df2[RDT_df2$side != "no rhyme",])
-  plot_data_2$Session<- 2
-
-# merge both
-plot.data <- rbind(plot_data_2, plot_data_1)
-
-# remove those who only did day 1
-plot.data <- merge(plot.data, complete.day2, by= "subject")
-
-# make look up for lextale and remove
-lexLOOK <- allsum %>% dplyr::select("subject", "lexTALE")
-lexLOOK <- lexLOOK[lexLOOK$lexTALE >= 80,]
-# remove those wuth poor lextale
-plot.data <- merge(plot.data, lexLOOK, by= "subject")
-
-# now look up exclusions 
-excludeLOOK <- allsum %>% dplyr::select("subject", "exclude", "exRDT")
-plot.data <- merge(plot.data, excludeLOOK, by= "subject")
-
-# make style conistent
-plot.data$side <- tolower(plot.data$side)
-
-# save data frame
-write.csv(plot.data, "./data/for_figs.csv")
+}
+write.csv(allsum, paste0("LIs",oddeven,".csv"))
+```
 
@@
 ```{r reducer}
+doreduce <- 0 #this chunk takes time, so omit unless doreduce set to one
+if (doreduce==1){
 {for (d in 1:2){ #we will process day 1 and day 2 together
@@
 ```{r reducerDL}
+if(doreduce==1){
 {for (d in 1:2){ #we will process day 1 and day 2 together
@@
 )
+}
 ```

### 2020-11-20 battery-preprocessing_v6_DB_split2.Rmd
codes hand as factor

@@
 }
-
+allsum$handpch <- as.factor(allsum$handpch)
 #now plot EHI in relation to handedness (modified by DB)
@@
 write.csv(allsum, paste0("LIs",oddeven,".csv"))
-```
-
 ```

### 2020-11-20 battery-preprocessing_v6_DB.Rmd
original preprocess to make split half

--- 2020-07-20 battery-preprocessing_v6_DB.Rmd
@@
 #Reading from OSF - this only works if files are public
-demo_dat <- read.csv('https://osf.io/7smj6/download',stringsAsFactors = F)
+#demo_dat <- read.csv('https://osf.io/7smj6/download',stringsAsFactors = F)
+demo_dat <- read.csv("./data/Demographics.csv",stringsAsFactors = F)
 
@@
 ```{r sightednesss, warning=FALSE}
-porta_dat <- read.csv("https://osf.io/zmqkt/download",stringsAsFactors = F)
-miles_dat <-  read.csv("https://osf.io/ekqsp/download",stringsAsFactors = F)
+#porta_dat <- read.csv("https://osf.io/zmqkt/download",stringsAsFactors = F)
+#miles_dat <-  read.csv("https://osf.io/ekqsp/download",stringsAsFactors = F)
+porta_dat <- read.csv("./data/porta.csv",stringsAsFactors = F)
+miles_dat <- read.csv("./data/miles.csv",stringsAsFactors = F)
 
@@
 ```{r EHI, warning=FALSE}
-EHI <- read.csv("https://osf.io/kfdma/download",stringsAsFactors = F)
+#EHI <- read.csv("https://osf.io/kfdma/download",stringsAsFactors = F)
+EHI <- read.csv("./data/EHI.csv",stringsAsFactors = F)
 
@@
 #This is not perfect, but gives and subjectea to check EHI is valid etc.
-
+allsum$handpch <- as.factor(allsum$handpch)
 ggplot(allsum, aes(x=index_EHI)) + geom_density(aes(group=handpch,fill=handpch))
@@
 # read data
-lexTALE <- read.csv("https://osf.io/h4btp/download",stringsAsFactors = F)
+#lexTALE <- read.csv("https://osf.io/h4btp/download",stringsAsFactors = F)
+lexTALE <- read.csv("./data/LexTALE.csv",stringsAsFactors = F)
 
@@
 # now create a data frame only for index 
-lexTALE <- select(lexTale_score, "subject", "lexTALE")
+lexTALE <- dplyr::select(lexTale_score, "subject", "lexTALE")
 # now merge with allsum
@@
 for (d in 1:2){
-  data_file <- "https://osf.io/mhk29/download" #day 1 data
+  data_file <- "./data/combined/day1_combined.csv" #"https://osf.io/mhk29/download" #day 1 data
 
   if (d ==2){
-    data_file <-  "https://osf.io/ebm6s/download"
+    data_file <-  "./data/combined/day2_combined.csv"#"https://osf.io/ebm6s/download"
   }
@@
 #added by DB 17th July 2020: include raw accuracy L and R in the allsum dataframe
-chimdatL1<-select(filter(chim_dat1,side=='Left'),subject,accurate)
-chimdatR1<-select(filter(chim_dat1,side=='Right'),subject,accurate)
-chimdatL2<-select(filter(chim_dat2,side=='Left'),subject,accurate)
-chimdatR2<-select(filter(chim_dat2,side=='Right'),subject,accurate)
+chimdatL1<-dplyr::select(filter(chim_dat1,side=='Left'),subject,accurate)
+chimdatR1<-dplyr::select(filter(chim_dat1,side=='Right'),subject,accurate)
+chimdatL2<-dplyr::select(filter(chim_dat2,side=='Left'),subject,accurate)
+chimdatR2<-dplyr::select(filter(chim_dat2,side=='Right'),subject,accurate)
 
@@
 #added by DB 17th July 2020: include raw accuracy L and R in the allsum dataframe
-DLdatL1<-select(filter(DL_dat1,side=='Left'),subject,accurate)
-DLdatR1<-select(filter(DL_dat1,side=='Right'),subject,accurate)
-DLdatL2<-select(filter(DL_dat2,side=='Left'),subject,accurate)
-DLdatR2<-select(filter(DL_dat2,side=='Right'),subject,accurate)
+DLdatL1<-dplyr::select(filter(DL_dat1,side=='Left'),subject,accurate)
+DLdatR1<-dplyr::select(filter(DL_dat1,side=='Right'),subject,accurate)
+DLdatL2<-dplyr::select(filter(DL_dat2,side=='Left'),subject,accurate)
+DLdatR2<-dplyr::select(filter(DL_dat2,side=='Right'),subject,accurate)
 
@@
 
+#Also record % correct for each category in allsum
+for (d in 1:2){
+  myfile<-RDT_df1
+  if(d==2){ myfile<-RDT_df2}
+pcorL<-myfile[myfile$side=='left',]
+pcorR<- myfile[myfile$side=='right',]
+pcorNull <- myfile[myfile$side=='no rhyme',]
+pcorL<-dplyr::select(pcorL,subject,p.corr)
+pcorR<-dplyr::select(pcorR,subject,p.corr)
+pcorNull<-dplyr::select(pcorNull,subject,p.corr)
+colnames(pcorL)[2]<-paste0('RDT',d,'.pcorrL')
+colnames(pcorR)[2]<-paste0('RDT',d,'.pcorrR')
+colnames(pcorNull)[2]<-paste0('RDT',d,'.pcorrNull')
+allsum <- merge(allsum,pcorL,by="subject")
+allsum <- merge(allsum,pcorR,by="subject")
+allsum <- merge(allsum,pcorNull,by="subject")
+}
+
 #Scrutiny of these reveal a few cases who missed all 'no rhyme' cases.
@@
 #added by DB 17th July 2020: include raw accuracy L and R in the allsum dataframe
-RDTdatL1<-select(filter(RDT_df1,side=='left'),subject,RT)
-RDTdatR1<-select(filter(RDT_df1,side=='right'),subject,RT)
-RDTdatL2<-select(filter(RDT_df2,side=='left'),subject,RT)
-RDTdatR2<-select(filter(RDT_df2,side=='right'),subject,RT)
+RDTdatL1<-dplyr::select(filter(RDT_df1,side=='left'),subject,RT)
+RDTdatR1<-dplyr::select(filter(RDT_df1,side=='right'),subject,RT)
+RDTdatL2<-dplyr::select(filter(RDT_df2,side=='left'),subject,RT)
+RDTdatR2<-dplyr::select(filter(RDT_df2,side=='right'),subject,RT)
 
@@
 #added by DB 17th July 2020: include raw accuracy L and R in the allsum dataframe
-fingerL1<-select(finger.df,subject,L1)
-fingerR1<-select(finger.df,subject,R1)
-fingerL2<-select(finger.df,subject,L2)
-fingerR2<-select(finger.df,subject,R2)
+fingerL1<-dplyr::select(finger.df,subject,L1)
+fingerR1<-dplyr::select(finger.df,subject,R1)
+fingerL2<-dplyr::select(finger.df,subject,L2)
+fingerR2<-dplyr::select(finger.df,subject,R2)
 
@@
 write.csv(allsum, "LIs.csv")
+```
+
+### Calculate split half reliability
+
+#### Chimeric faces
+
+```{r Chimeric Split Half}
+varlist <-   c("subject", "side", "accurate", "RT") #need to be in this order, ie sub, side, acc and RR
+latlist <- c("Left","Right") #names of factor levels
+nloop=2 #same as levels of latlist
+# create data frames
+CF_dat1$split <- c("odds", "even")
+  CF_dat_1.1 <- CF_dat1[CF_dat1$split== "odds",]
+  CF_dat_1.2 <- CF_dat1[CF_dat1$split== "even",]
+CF_dat2$split <- c("odds", "even")
+  CF_dat_2.1 <- CF_dat2[CF_dat2$split== "odds",]
+  CF_dat_2.2 <- CF_dat2[CF_dat2$split== "even",]
+# create aggregate data
+chim_dat1.1 <- make.df(CF_dat_1.1,varlist,latlist,nloop)
+chim_dat1.2 <- make.df(CF_dat_1.2,varlist,latlist,nloop)
+chim_dat2.1 <- make.df(CF_dat_2.1,varlist,latlist,nloop)
+chim_dat2.2 <- make.df(CF_dat_2.2,varlist,latlist,nloop)
+#exclude
+#find included cases
+w<-which(allsum$exclude==0)
+  inclsubs <- allsum$subject[w]
+
+# NOW MAKE LI_z
+# 1 odds
+task<-'Chimeric faces'
+origdat<-chim_dat1.1
+origdat$X1 <- origdat$N
+mycolname<-"split.chim.1.1"
+mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
+polarity<- (1)#so LI is positive in predicted direction
+allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)
+# 1 even
+origdat<-chim_dat1.2
+origdat$X1 <- origdat$N
+mycolname<-"split.chim.1.2"
+polarity<- (1)#so LI is positive in predicted direction
+allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)
+# 2 odds
+origdat<-chim_dat2.1
+origdat$X1 <- origdat$N
+mycolname<-"split.chim.2.1"
+polarity<- (1)#so LI is positive in predicted direction
+allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)
+# 2 evens
+origdat<-chim_dat2.2
+origdat$X1 <- origdat$N
+mycolname<-"split.chim.2.2"
+polarity<- (1)#so LI is positive in predicted direction
+allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)
+# remove
+w<-which(allsum$exclude == 1)
+  allsum$split.chim.1.1[w] <- NA
+  allsum$split.chim.1.2[w] <- NA
+  allsum$split.chim.2.1[w] <- NA
+  allsum$split.chim.2.2[w] <- NA
+
+print(paste0("Chimeric faces session 1 split-half: ", cor(allsum$split.chim.1.1, allsum$split.chim.1.2, use = "complete.obs")))
+print(paste0("Chimeric faces session 2 split-half: ", cor(allsum$split.chim.2.1, allsum$split.chim.2.2, use = "complete.obs")))
+```
+
+#### Dichotic listening
+
+```{r Dichotic Split Half}
+varlist <-   c("subject", "side", "accurate", "RT") #need to be in this order, ie sub, side, acc and RR
+origdat <- dichotic_dat1
+latlist <- c("Left","Right") #names of factor levels
+nloop=2 #same as levels of latlist
+# made data frames
+dichotic_dat1$split <- c("odds", "even")
+  DL_1.1 <- dichotic_dat1[dichotic_dat1$split== "odds",]
+  DL_1.2 <- dichotic_dat1[dichotic_dat1$split== "even",]
+dichotic_dat2$split <- c("odds", "even")
+  DL_2.1 <- dichotic_dat2[dichotic_dat2$split== "odds",]
+  DL_2.2 <- dichotic_dat2[dichotic_dat2$split== "even",]
+# create aggregate data
+DL_dat1.1 <- make.df(DL_1.1,varlist,latlist,nloop)
+DL_dat1.2 <- make.df(DL_1.2,varlist,latlist,nloop)
+DL_dat2.1 <- make.df(DL_2.1,varlist,latlist,nloop)
+DL_dat2.2 <- make.df(DL_2.2,varlist,latlist,nloop)
+# exclude
+w<-which(is.na(DL_dat1.1$subject))
+  if (length(w)>0) {DL_dat1.1<-DL_dat1.1[-w,]}
+w<-which(is.na(DL_dat1.2$subject))
+  if (length(w)>0) {DL_dat1.2<-DL_dat1.2[-w,]}
+w<-which(is.na(DL_dat2.1$subject))
+  if (length(w)>0) {DL_dat2.1<-DL_dat2.1[-w,]}
+w<-which(is.na(DL_dat2.2$subject))
+  if (length(w)>0) {DL_dat2.2<-DL_dat2.2[-w,]}
+DL_dat1.1[is.na(DL_dat1.1)] <- 0
+DL_dat1.2[is.na(DL_dat1.2)] <- 0
+DL_dat2.1[is.na(DL_dat2.1)] <- 0
+DL_dat2.2[is.na(DL_dat2.2)] <- 0
+
+# MAKE Z SCORE
+# day 1 odds
+task<-'Dichotic listening'
+origdat<-DL_dat1.1
+origdat$X1 <- origdat$N
+mycolname<-"split.DL.1.1"
+mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
+polarity<- (1)#so LI is positive in predicted direction
+allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)
+# day 1 evens
+origdat<-DL_dat1.2
+origdat$X1 <- origdat$N
+mycolname<-"split.DL.1.2"
+allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)
+# day 2 odds
+origdat<-DL_dat2.1
+origdat$X1 <- origdat$N
+mycolname<-"split.DL.2.1"
+allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)
+# day 2 evens
+origdat<-DL_dat2.2
+origdat$X1 <- origdat$N
+mycolname<-"split.DL.2.2"
+allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)
+#remove
+# remove
+w<-which(allsum$exclude == 2)
+  allsum$split.DL.1.1[w] <- NA
+  allsum$split.DL.1.2[w] <- NA
+  allsum$split.DL.2.1[w] <- NA
+  allsum$split.DL.2.2[w] <- NA
+
+print(paste0("Dichotic listening session 1 split-half: ", cor(allsum$split.DL.1.1, allsum$split.DL.1.2, use = "complete.obs")))
+print(paste0("Dichotic listening session 2 split-half: ", cor(allsum$split.DL.2.1, allsum$split.DL.2.2, use = "complete.obs")))
+```
+
+#### Rhyme decision task
+
+```{r Rhyme Split Half}
+varlist <-   c("subject", "side", "accurate", "RT") #need to be in this order, ie sub, side, acc and RR
+latlist <- c("left","no rhyme","right") #names of factor levels
+nloop<-length(latlist)
+# made data frames
+RDT_dat1$split <- c("odds", "even")
+  RDT_dat1.1 <- RDT_dat1[RDT_dat1$split== "odds",]
+  RDT_dat1.2 <- RDT_dat1[RDT_dat1$split== "even",]
+RDT_dat2 <- RDT_dat2[-nrow(RDT_dat2),]
+RDT_dat2$split <- c("odds", "even")
+  RDT_dat2.1 <- RDT_dat2[RDT_dat2$split== "odds",]
+  RDT_dat2.2 <- RDT_dat2[RDT_dat2$split== "even",]
+# create aggregate data
+RDT_df1.1 <- make.df(RDT_dat1,varlist,latlist,nloop)
+RDT_df1.2 <- make.df(RDT_dat2,varlist,latlist,nloop)
+RDT_df2.1 <- make.df(RDT_dat1,varlist,latlist,nloop)
+RDT_df2.2 <- make.df(RDT_dat2,varlist,latlist,nloop)
+# remove NaN
+is.nan.data.frame <- function(x)
+  do.call(cbind, lapply(x, is.nan))
+RDT_df1.1[is.nan(RDT_df1.1)] <- NA
+RDT_df1.2[is.nan(RDT_df1.2)] <- NA
+RDT_df2.1[is.nan(RDT_df2.1)] <- NA
+RDT_df2.2[is.nan(RDT_df2.2)] <- NA
+
+# exclude
+wn <-unique(which(RDT_df1.1$accurate ==0),which(RDT_df1.2$accurate ==0, (which(RDT_df2.1$accurate ==0, (which(RDT_df2.2$accurate ==0))))))
+w1<-which(RDT_df1.1$p.corr < 75)
+w2<-which(RDT_df1.2$p.corr < 75)
+w3<-which(RDT_df2.1$p.corr < 75)
+w4<-which(RDT_df2.2$p.corr < 75)
+exRDT<-unique(c(RDT_df1.1$subject[w1],
+                RDT_df1.2$subject[w2],
+                RDT_df2.1$subject[w3],
+                RDT_df2.1$subject[w4],
+                RDT_df1.1$subject[wn],
+                RDT_df1.2$subject[wn],
+                RDT_df2.1$subject[wn],
+                RDT_df2.2$subject[wn]))
+allsum$exRDT_split <- 0
+allsum$exRDT_split[allsum$subject %in% exRDT]<-1 #decided best to have separate column for exclusion for each task
+w<-which(allsum$exRDT_split==0)  
+incRDT <- allsum$subject[w]
+ninc <- length(incRDT)
+
+# MAKE Z LAT
+for (d in 1:4){
+  ifelse(d==1, myfile<-RDT_dat1.1, 
+         ifelse(d==2, myfile<-RDT_dat1.2,
+                ifelse(d==3, myfile<-RDT_dat2.1,
+                       myfile<-RDT_dat2.2)))
+    ifelse(d==1, lab<- 1.1, 
+         ifelse(d==2, lab<- 1.2,
+                ifelse(d==3, lab<- 2.1,
+                       lab <- 2.2)))
+  nsubs <- nrow(allsum)
+  allsum$tempt<-NA #initialise a new column
+  for (i in 1:nsubs) {
+    mysub<-filter(myfile,side %in% c('left','right'),subject==allsum$subject[i])
+    if(nrow(mysub)>10)
+    {
+      mymean<-aggregate(RT~side,mysub,FUN = mean)
+      mysd<-aggregate(RT~side,mysub,FUN = sd)
+        myN<-aggregate(RT~side,mysub,FUN = length)
+      se1<-mysd[1,2]/sqrt(myN[1,2])
+      se2<-mysd[2,2]/sqrt(myN[2,2])
+      se<-sqrt(se1^2+se2^2)
+      allsum$tempt[i]<-(mymean[1,2]-mymean[2,2])/se
+    }
+  }
+  wc<-which(colnames(allsum)=='tempt')
+  colnames(allsum)[wc]<-paste0('split.RDT.',lab)
+}
+w<-which(is.na(allsum$Day1.RDT_RT_LI)) #remove those with NA in LI
+  allsum$split.RDT.1.1[w]<-NA
+  allsum$split.RDT.1.2[w]<-NA
+w<-which(is.na(allsum$Day2.RDT_RT_LI))
+  allsum$split.RDT.2.1[w]<-NA
+  allsum$split.RDT.2.1[w]<-NA
+
+print(paste0("Rhyme decision session 1 split-half: ", cor(allsum$split.RDT.1.1, allsum$split.RDT.1.2, use = "complete.obs")))
+print(paste0("Rhyme decision session 2 split-half: ", cor(allsum$split.RDT.2.1, allsum$split.RDT.2.2, use = "complete.obs")))
 ```
@@
 complete.day2$day2.complete <- 1
-complete.day2 <- select(complete.day2, -2, -3, -4, -5)
+complete.day2 <- dplyr::select(complete.day2, -2, -3, -4, -5)
 # now merge into allsum
@@
 CF.corrs <- na.omit(CF.corrs)
-CF.cor.plot <- ggplot(CF.corrs, aes(x=trial, y=cor)) +
-  geom_hline(yintercept=.65, linetype="dashed", color = "blue") +
-  geom_hline(yintercept=.75, linetype="dashed", color = "red") +
-  geom_line(size=1, alpha=0.8) + ylim(0, 1) + theme_classic() +
-  xlab("No. of trials") + ylab("Test-retest reliability") + ggtitle("B") + theme(plot.title = element_text(hjust = 0.5))
+CF.cor.plot <- ggplot(CF.corrs, aes(x=trial, y=abs(cor))) +
+  geom_hline(yintercept=.65, linetype="dashed", color = "red") +
+  geom_line(size=1, alpha=0.8) + ylim(0, 1) + theme_bw() +
+  xlab("Number of trials") + ylab("Test-retest reliability") + ggtitle("(B)") + theme(plot.title = element_text(hjust = 0.5))
 # print number of trials needed
@@
 DL.corrs <- na.omit(DL.corrs)
-DL.cor.plot <- ggplot(DL.corrs, aes(x=trial, y=cor)) +
-  geom_hline(yintercept=.65, linetype="dashed", color = "blue") +
-  geom_hline(yintercept=.75, linetype="dashed", color = "red") +
-  geom_line(size=1, alpha=0.8) + ylim(0, 1) + theme_classic() +
-  xlab("No. of trials") + ylab("Test-retest reliability") + ggtitle("A") + theme(plot.title = element_text(hjust = 0.5))
+DL.cor.plot <- ggplot(DL.corrs, aes(x=trial, y=abs(cor))) +
+  geom_hline(yintercept=.65, linetype="dashed", color = "red") +
+  geom_line(size=1, alpha=0.8) + ylim(0, 1) + theme_bw() +
+  xlab("Number of trials") + ylab("Test-retest reliability") + ggtitle("(A)") + theme(plot.title = element_text(hjust = 0.5))
 # print number of trials needed
@@
 ggsave(
-  "all.cor.plot.png",
+  "./plots/all.cor.plot.png",
   plot = all.cor.plot,

### 2020-11-20 online_laterality_reliability.Rmd
updates reviewer' comments

--- 2020-08-07 online_laterality_reliability.Rmd
@@
 ---
-title: 'Assessing the reliability of an online behavioural laterality battery: a pre-registered study.'
+title: 'Assessing the reliability of an online behavioural laterality battery: A pre-registered study.'
 output:
@@
 <!--- updated 20/7/20 to change 'select' to 'dplyr::select' to avoid conflicts-->
+
 <!--- updated 3/8/20 to add ZW feedback--->
-Adam J. Parker*
-
-Zoe V. J. Woodhead
-
-Paul A. Thompson
-
-Dorothy V. M. Bishop
+
+Adam J. Parker*, Zoe V. J. Woodhead, Paul A. Thompson, Dorothy V. M. Bishop
 
@@
 The reliability of the rhyme decision task was less satisfactory (*r*= .63).
-We also found that left-handers were more likely to depart from typical lateralisation.
-Contrary to prediction, lateralisation across tasks was not significantly correlated.
-We conclude that three of the tasks, dichotic listening, chimeric faces and finger tapping, show considerable promise for online evaluation of cerebral lateralisation, but the results do not support the notion of a common underlying cause of lateralisation for the different cognitive tasks studied here.
+We also found that extreme left-handers were more likely to depart from typical lateralisation.
+Lateralisation across the two language tasks (dichotic listening and rhyme judgement) was weakly correlated, but unrelated to lateralisation on the chimeric faces task.
+We conclude that three of the tasks, dichotic listening, chimeric faces and finger tapping, show considerable promise for online evaluation of cerebral lateralisation.
+The results do not support the notion of a common underlying cause of lateralisation for language and face processing.
 
@@
 Without well-powered experiments that use reliable tasks, it becomes difficult to address many research questions regarding the relationship between laterality indices and their stability over time.
-Concerning the relationship between indices, unless a task is reliable, the absence of a correlation cannot be interpreted as evidence for independence of function (Bishop, 2017).
+Concerning the relationship between indices, unless a task is reliable, the absence of a correlation cannot be interpreted as evidence for independence of function (Parsons, Kruijt & Fox, 2019).
 Reliability can either be assessed within a test session (e.g.
@@
 The behavioural tasks included were: a novel rhyme decision task, a dichotic listening task (Hugdahl et al., 2009), a chimeric faces task (Burt & Perret, 1997), and a novel finger tapping task.
-the validity of the tasks was assessed in terms of whether lateralisation was observed in the predicted direction, based on prior literature with in-person testing on analogous tasks.
+The validity of the tasks was assessed in terms of whether lateralisation was observed in the predicted direction, based on prior literature with in-person testing on analogous tasks.
 We also examined two pre-registered questions: (1) do extreme left-handers show a higher rate of atypical laterality than right-handers on each behavioural laterality task?
@@
 Results from visual half-field and dichotic listening paradigms support the view that the left hemisphere plays the dominant role in language processing (Hugdahl, 2000; Ocklenburg & Güntürkün, 2018).
-Here we review how each paradigm has been implemented to study language lateralisation and what is known about reliability and validity.
-
-#### Visual half-field paradigm
-
+By contrast, visual half-field paradigms indicate that the right hemisphere is involved in the processing of faces, global feature processing, and visuo-spatial attention (Brederoo, Nieuwenstein, Cornelissen, & Lorist, 2019).
+Below, we review how each paradigm has been implemented to study language lateralisation and the detection of emotion in faces, and what is known about reliability and validity.
+
+#### Lateralised visual word processing
+
+Lateralised visual word processing has been studied using the visual half-field paradigm.
 In this paradigm, visual stimuli are briefly presented to the left or right of a central fixation stimulus.
@@
 
-Despite wide reporting of a RVF advantage at the population level, the validity of visual half-field tasks has been questioned.
+Despite wide reporting of a RVF advantage at the population level, the validity of visual half-field language tasks has been questioned.
 Krach, Chen, and Hartje (2006) compared the laterality indices (LIs) produced during a visual half-field experiment with those produced using functional Transcranial Doppler Sonography (fTCD).
@@
 Despite showing the expected RVF advantage on the visual half-field task, the behavioural LIs produced on the visual half-field paradigm did not correlate with the fTCD LIs produced during word generation (*r*= .18) or the visual half-field task itself (r -.004).
-The lack of validity here likely reflects the poor reliability of the visual field task as split-half reliability was very low (reaction time split-half: *r*= .14; accuracy split-half: *r*= .20).
-This illustrates the difficulty of concluding associations between laterality indices when the reliability of measurement is so low.
+The lack of validity here likely reflects the poor reliability of the visual field task as split-half reliability was very low (response time split-half: *r*= .14; accuracy split-half: *r*= .20).
+This illustrates the difficulty of interpreting associations between laterality indices when the reliability of measurement is so low.
 
@@
 However, the small sample size meant that the confidence intervals around these estimates were very wide: word naming, *r*= .63 (95% CI = .001-.902), picture naming, *r*= .77 (95% CI = .272-.943).
-
-#### Dichotic listening
+More recently, however, Van der Haegen, Cai, Seurinck, and Brysbaert (2011) reported highly similar findings in a somewhat larger sample of 50 participants.
+Again, participants completed behavioural word and picture naming visual half-field tasks and then a word generation task under fMRI.
+It is reassuring that the LIs for the behavioural measures were correlated with the fMRI LI: word naming, *r* = .64, picture naming, *r* = .65.
+Van der Haegen et al.
+also report a set of analyses involving 250 participants who completed the behavioural visual half-field tasks while their eye movements were monitored.
+Two sets of LIs were calculated; one controlling for fixation error and one that did not.
+Importantly, the LIs which did and did not control for fixation error were highly correlated for word naming, *r* = .94, and picture naming,*r* =.98, indicating that the inclusion of eye movement control adds little additional value when a visual half-field task employs mechanisms to encoruage central fixation and bilateral presentation.
+
+A promising line of work has emerged which further highlights the potential for well designed visual half-field tasks to assess language lateralisation.
+First introduced by Willemin et al.
+(2016), the translingual lexical decision task involves participants viewing letter stings either side of fixation and judging whether the left, right, or neither of the letter strings was a valid word.
+With 100 participants completing 256 trials, Willemin et al.
+reported a RVF advantage at the population level.
+Since then this effect has been replicated in a sample of 496 participants covering six international languages, with evidence of a RVF for each language studied.
+Not only does this task show validity, it seems reliable given Brederoo et al.'s (2020) observation that the task yields good split-half reliability when tested with 122 participants, *r* = 0.86 for error rates and *r* = 0.80 for response times.
+
+#### Lateralised auditory verbal processing
 
@@
 As with the visual half-field paradigm, the test-retest reliability of dichotic listening paradigms has been far from optimal (Voyer, 1998; Kelley & Littenberg, 2019).
-However, recent data from Westerhausen and Samuelsen (2020) indicated that when designed well, the dichotic listening paradigm can be reliable.
-First, based on a review of the published literature, Westerhausen and Samuelsen outlined eight design features that should be used in an optimal dichotic listening paradigm and explain their justification for each.
-Their recommendations were to use: (1) stop consonant-vowel (CV) syllables as stimulus materials; (2) CV stimulus pairs from the same voicing category; (3) alternating trials of voiced and unvoiced stimulus pairs; (4) free-recall instruction; (5) one stimulus pair per trial; (6) a paradigm length of 120 dichotic trials; (7) all stimulus pairs are presented to the two sides in identical frequency; and (8) to include binaural (homonymic) trials.
+However, recent data from Westerhausen and Samuelsen (2020) indicated that when designed optimally, the dichotic listening paradigm can be reliable.
+First, based on a review of the published literature, Westerhausen and Samuelsen outlined eight design features that should be used in an optimal dichotic listening paradigm and explained their justification for each.
+Their recommendations were: (1) stop consonant-vowel (CV) syllables as stimulus materials; (2) CV stimulus pairs from the same voicing category; (3) alternating trials of voiced and unvoiced stimulus pairs; (4) free-recall instruction; (5) one stimulus pair per trial; (6) a paradigm length of 120 dichotic trials; (7) all stimulus pairs presented at equal frequency to the two sides; and (8) inclusion of binaural (homonymic) trials.
 Putting these criteria into practice, Westerhausen and Samuelsen administered a modified dichotic listening paradigm using both a verbal and manual response format to 50 right-handers.
@@
 
-#### Chimeric faces
+To date, little work has been conducted online with visual stimuli.
+By contrast, the dichotic listening paradigm has been implemented online with considerable success.
+Bless et al.
+(2013) developed a mobile device version of the consonant-vowel dichotic listening paradigm, with 36 simultaneous sound pairs.
+Under laboratory conditions, with a sample of 76 (33 Norwegian, 43 Austrialian) participants, they reported a test-retest reliability of *r*= 0.78 when using the mobile app.
+Furthermore, both subgroups of participants showed clear right ear advantages at the population level.
+In a subsequent large-scale study, Bless et al.
+(2015) administered the same application to a sample of over 4,000 participants with more than 60 different language backgrounds.
+Again Bless et al.
+(2015) reported a clear right ear advanatge at the population-level.
+They then used the large dataset to examine cultural/lingusitic effects on the degree on lateralisation; reporting that the degree of lateralization appears to be modulated by linguistic background.
+Together these studies demonstrate the potential for online laterality research to generate large amounts of data while maintaining a tolerable degree of reliability.
+
+#### Lateralised face processing
 
@@
 The validity of the chimeric faces task has been demonstrated using behavioural and fMRI data but in a task of recognising facial identity rather than emotion.
-In Yovel, Tambini, and Brandman’s (2008) study, participants performed a behavioural experiment where they had to recognise centrally presented chimeric faces, which showed different identities in the right and left visual fields.
+In Yovel, Tambini, and Brandman’s (2008) study, participants had to recognise centrally presented chimeric faces, which showed different identities in the right and left visual fields.
 It is encouraging to note that the behavioural measure significantly correlated with the fMRI laterality index (LI) calculated from on activation in the fusiform face area (*r*= .49).
@@
 One factor that has been associated with the degree of lateralisation is handedness (for a comprehensive review, see Carey & Johnstone, 2014).
-Neuroimaging data indicates that for language tasks these differences in lateralisation primarily reflect the contribution of a subset of strong left-handers who are right hemisphere dominant for language, rather than indicating that this relationship operates on a continuum (Mazoyer et al., 2014).
+Neuroimaging data indicate that for language tasks these differences in lateralisation primarily reflect the contribution of a subset of strong left-handers who are right hemisphere dominant for language, rather than indicating that this relationship operates on a continuum (Mazoyer et al., 2014).
 This association can be used to validate behavioural measures of lateralisation: we would expect to see a departure from typical lateralisation in strong left-handers.
 
-With regards to visual half-field tasks for studying language lateralisation, there is evidence to suggest that the visual field advantage varies with handedness (Elias, Bulman-Fleming, & McManus, 1999; Weekes, Capetillo-Cunliffe, Rayman, Iacoboni, & Zaidel, 1999).
+With regards to visual half-field tasks for studying language lateralisation, there is evidence that the visual field advantage varies with handedness (Elias, Bulman-Fleming, & McManus, 1999; Weekes, Capetillo-Cunliffe, Rayman, Iacoboni, & Zaidel, 1999).
 However, a relatively well-powered study involving 100 participants, conducted by Willemin et al.
 (2016), reported no hint of any effect of handedness, when coded as left and right, or when divided into left (N = 19), mixed (N = 24) and right-handed (N = 57) groups, on the visual field advantage during a lexical decision task.
-This result illustrates the recurring challenge of interpreting null results in this field: do they reflect a true lack of relationship, poor test reliability, inadequate power to detect small effects, or specific methodological factors?
+This result illustrates the recurring challenge of interpreting null results in this field: do they reflect a true lack of relationship, inadequate power to detect small effects, or specific methodological factors?
+Brederoo et al.
+(2020) reported good split-half reliability for this task, suggesting the lack of handedness effect is unlikely to be due solely to measurement error.
 We may note that Willemin et al.
-asked participants to respond via manual response, whereas other studies have involved an active speech production component, which could influence task lateralisation.
+asked participants to respond via manual response, whereas other studies have involved an active speech production component; manual responding could possibly interact with handedness to influence the laterality index on this task.
 
@@
 This relationship points towards a general laterality factor which influences lateralisation across tasks and behaviours, as predicted by several theoretical accounts (Behrmann & Plaut, 2015; Hellige, 1993; Kosslyn, 1987).
+For example, the causal complementarity principle proposes that the lateralisation of processes involving the same hemisphere will be lateralised to the homologue in the opposite hemisphere.
+With reference to language processing and face emotionality processing, it would be predicted that language processing would be left lateralised and face processing would be right lateralised, and an increase of one co-occurs with an increase of another.
+Indeed, several studies have reported such patterns.
+In support, Brederoo et al.
+(2020) and Dundas, Plaut, and Behrmann (2015) have reported significant correlations between LIs for word and face processing.
 In a study of left-handers using fMRI with five tasks, Gerrits, Verhelst, and Vingerhoets (2020) found that the just under half the participants showed either standard laterality (left-lateralised for language and praxis, right-lateralised for face memory, line-bisection and vocal emotion) or a complete reversal of this pattern, providing clear evidence against the idea that these functions are lateralised independently.
-It is also clear that this association is far from perfect, (Badzakova-Trajkov, Haberling, Roberts, & Corballis, 2010) and some studies have found no correlation between laterality indices (Bryden, Hecaen, & Deagostini, 1983; Rosch, Bishop, & Badcock, 2012; Whitehouse & Bishop, 2009; van der Haegen & Brysbaert, 2018).
+Nevertheless, the data indicated that this association is far from perfect (Badzakova-Trajkov, Haberling, Roberts, & Corballis, 2010) and some studies have found no correlation between laterality indices (Bryden, Hecaen, & Deagostini, 1983; Rosch, Bishop, & Badcock, 2012; Whitehouse & Bishop, 2009; van der Haegen & Brysbaert, 2018).
 Our ultimate goal is to unpick whether dissociations between laterality indices are evidence of multiple independent lateral biases across tasks, or whether there is a more prosaic explanation in terms of poor reliability or validity of tasks.
+
+To date the focus has been in testing whether lateralisation shows  complementarity across tasks that are usually mediated by left and right hemispheres, especially whether there is a relationship between language vs visuospatial functioning.
+However, emerging lines of evidence have suggested that even within domains there may be dissociated patterns of lateralisation across similar tasks.
+Woodhead, Bradshaw, Wilson, Thompson, and Bishop (2019) reported data for 37 participants engaged in six language tasks under transcranial Doppler sonography (fTCD) and examined the factorial structure of lateralisation across tasks using structural equation modeling.
+They pitted two models against each other.
+The first assumed that language laterality is a unidimensional trait and a single factor was used to model covariance.
+The second assumed that that two laterality factors underlie the pattern of covariance.
+The results supported a two factor model at the population level.
+However, further examination of the data revealed that this was driven by seven left-handers.
+Woodhead, Thompson, Karlsson, and Bishop (2020) examined this factor structure in an extended sample of 74 participants with 43 right-handers and 31 left-handers.
+The data this time indicated that a two factor model was preferred for both right- and left-handers, with the preference for a two factor model being stronger for left-handers.
+Together, this indicated that even with the domain of language laterality indices may be dissociated.
 
@@
 Specifically, we examined whether it is possible to observe the usual laterality biases at the population level in tests of language and emotion processing and whether it is possible to achieve test-retest reliability of *r*= .65 or above using online behaviour research methods.
+This is an arbitrary cutoff, selected to represent a lower bound of reliability for a measure to be potentially useful for analysing individual differences.
 In addition, for tasks where satisfactory reliability was achieved, we considered how many trials were needed to achieve this.
-The tasks included in the battery were a novel rhyme decision visual half-field task, a consonant-vowel dichotic listening task, a chimeric faces visual half-field task, and a novel finger tapping task.<!--- ZW: Would it be worth having a section on measurement of manual preference?; AP: As this is exploratory I don't want to add too much about this---> These behavioural tasks were completed twice on two different occasions.
+The tasks included in the battery were a novel rhyme decision visual half-field task, a consonant-vowel dichotic listening task, a chimeric faces visual half-field task, and a novel finger tapping task.
+These behavioural tasks were completed twice on two different occasions.
+
+When selecting specific tasks in our online battery we were strongly guided by pragmatic considerations, i.e.
+how feasible would it be to administer the task remotely via an online platform.
+We aimed to include at least two language measures, one tapping into receptive language and one expressive, in order to test the notion that language laterality can dissociate across different tasks.
+We also aimed to include a task that involved face processing so that we could test for complementarity of lateralisation across tasks that usually lateralise to left and right hemispheres.
+Finally, a behavioural measure of motor laterality was included, to potentially provide an online analogue of relative hand skill similar to that used by Annett (1970).
+
+Ultimately we aimed to develop tasks that would meet three criteria.
+First, the task should show good test-retest reliability.
+Second, it would have to show validity as demonstrated by showing significant lateralisation at the population level.
+Third, it should be sensitive to the subtle differences in laterality that have been reported between right- and non-right-handers.
+
+For receptive language, we selected the dichotic listening paradigm, which has already been shown to yield reliable results when used online via a mobile app (Bless et al., 2013, 2015).
+Furthermore, Karlsson et al.
+(2019) have shown that this paradigm is sensitive to the subtle differences in language asymmetry between right and non-right handers, making it an obvious choice for inclusion.
+
+For expressive language we aimed to use a visual half-field task.
+While the task reported by Brysbaert and colleagues looked promising, it involved overt verbal responses which would be difficult to accommodate in an online format.
+Thus we decided to develop a new rhyme judgement task that was designed to meet the limitations of online testing, and which would require covert naming of stimuli presented in the two half-fields (see below for details).
+
+Given that numerous studies looking at the relationship lateralisation on language and non-language tasks have focused on the use of face stimuli (e.g.
+Brederoo et al., 2020), we chose to include a face task to conduct similar comparisons.
+Regarding the chimeric faces, Levy et al.
+(1983) reported good split-half reliability while Karlsson et al.
+had reported that, like the dichotic listening task, the the chimeric faces task is sensitive to the differences in perceptual asymmetry in right- and non-right-handers.
+
+Our final online measure was a finger-tapping test used as an online alternative to the well-known Annett (1970) pegboard task.
+There is suggestive evidence that a continuous measure of relative skill of the two hands may be more effective than more conventional hand preference measures for capturing individual differences in motor laterality that relate to language lateralisation (Annett, 2002; Flowers & Hudson, 2013).
+Performance measures such as this can provide a continuous measure of subtle variation in relative hand skill, and also correlate well with measures of hand preference (e.g.
+Steenhuis & Bryden, 1999).
+The pegboard task is one of the most widely used performance tasks (see Hodgson & Hudson, 2016, for discussion), but the requirement for apparatus makes it unfeasible as an online test.
+We therefore took the opportunity to explore the reliability and validity of a finger-tapping task that required the participant to repeatedly tap a short sequence on the computer keyboard with the index finger of the preferred or nonpreferred hand, which capture one aspect of pegmoving, namely precise aiming of sequential movements.
 
@@
 Based on previous work, we predicted that at the population level: (1) the laterality indices for rhyme decision would show a RVF advantage (indicating left hemisphere dominance for language processing); (2) the laterality indices for the dichotic listening task would show a right ear advantage (indicating left hemisphere dominance for language processing); and (3) the laterality indices for chimeric faces would show a LVF advantage (indicating right hemisphere dominance for face or emotion processing).
-Although exploratory, the finger tapping task was designed to provide a quantitative measure of the degree of handedness, and was expected to show a right-hand advantage in the overall sample.
+Predictions regarding the finger tapping task were less clear; our goal was to evaluate a test of relative hand skill that might provide an online surrogate for the kind of peg-moving task used by Annett (1970).
+Validity of the task would be demonstrated if a  right-hand advantage was found in right-handers, and a left-hand advantage in left-handers.
+(Note that direct comparison with performance on a pegboard was not feasible as the study was conducted during a pandemic when in-person testing could not be done).
 
@@
 This would give us power to apply our planned model-fitting analysis to left-handers as a separate group.
-
 In practice, our final sample exceeded 300, because we discarded participants from individual tasks for failing to meet prespecified criteria (see below).
-In total, 441 native English speakers were initially recruited through Prolific Academic (www.prolific.co).
-Participants had normal or corrected-to-normal vision and reported no history of neurological disease (e.g.
+In total, 441 native English speakers were initially recruited through Prolific Academic (www.prolific.co) with all these participants completing both sessions,  Prolific is a platform for recruiting and screening paid research participants according to experimenter-specified criteria.
+Participants were 18- to 50-years-old, had normal or corrected-to-normal vision and reported no history of neurological disease (e.g.
 head injury, brain tumour, stroke) or significant hearing loss or impairment.
@@
 This study was approved by the University of Oxford's Medical Sciences Inter-Divisional Research Ethics Committee (reference number: R66638/RE002) and conducted following the principles of the Declaration of Helsinki.
-Participants gave informed consent and received compensation for their participation.
+Participants gave informed consent and were paid £15 for their participation.
 
@@
 For handedness, we asked: which is your dominant hand (i.e.
-which hand do you prefer to use for tasks such as writing, cutting, and catching a ball)?
+which hand do you prefer to use for tasks such as writing, cutting, and catching a ball?).
 For footedness, we asked: which foot do you normally use to step up on a ladder/step?
@@
 Each test classified participants as being either left or right eye dominant.
+We did not have specific predictions about ocular dominance, but gathered this information to further characterise behavioural laterality of the sample.
 In practice, the two tests agreed perfectly, so we report here just the Miles test.
@@
 Of the 60 strings, 40 were actual English words and 20 were non-words.
-LexTALE scores were corrected for the unequal proportion of words and non-word by averaging the percentages corrected for these two item types (e.g., ((number of words correct/40 x 100) + (number of nonwords correct/20 x 100))/2).
+LexTALE scores were corrected for the unequal proportion of words and non-words by averaging the percentages corrected for these two item types (e.g., ((number of words correct/40 x 100) + (number of nonwords correct/20 x 100))/2).
 Accordingly, the LexTALE scores ranged from 0 to 100.
@@
 With online presentation, a particular concern is a potential for eye movements away from fixation.
-We relied on the fact that participants had to focus initially on the central written word to do the task, with pictures subsequently displayed with brief bilateral presentation, which has been shown to give adequate control for eye movements in previous visual half-field experiments with verbal naming (Beaumont, 1982).
+We relied on the fact that participants had to focus initially on the central written word to do the task, and thus the task encouraged fixation to a central location at the start of the trial.
+We subsequently displayed pictures under brief bilateral presentation, which has been shown to give adequate control for eye movements in previous visual half-field experiments with verbal naming (Beaumont, 1982).
 Studies that directly monitored eye movements in this setting reported failures of fixation on only 0.5% of trials (Geffen, Bradshaw & Nettleton, 1972; but see also Bourne, 2006).
 Pictures of familiar objects were presented for rhyme matching to the written word, to ensure that the participant would have to covertly name the picture, thereby engaging left hemisphere speech processing.
+We used covert rather than overt naming (cf.
+Hunter & Brysbaert, 2008) to make the task suitable for online presentation.
+This obviated the need for measuring latencies of oral responses, which would be challenging in the online context, both in terms of need for a high-quality microphone, and possible intrusion of background noises interfering with recording.
+Furthermore, the evaluation of oral responses would require additional scoring following data collection which would be a taxing procedure for researchers.
+It would also require additional ethical approvals with regards to the storing and processing of the data.
 
@@
 In the example, the word “bite” is presented along with the two images “kite” and “door”.
-When making the correct judgement, participants would respond with by pressing ‘S’ to indicate that the rhyme was on the left.](Picture1.png)
+When making the correct judgement, participants would respond with by pressing ‘S’ to indicate that the rhyme was on the left.](Picture1.png){width=60%}
 
@@
 The trials were preceded by a familiarisation phase in which each of the stimulus pictures was presented together with its written name.
-Then 260 stimuli (52 unique stimuli each shown 5 times) were displayed across four blocks (65 items per block).
+After familiarisation, 260 stimuli (52 unique stimuli each shown 5 times) were displayed across four blocks (65 items per block).
 Each of the four blocks also contained five items for which there was no rhyme.
@@
 If participants did not detect a rhyme, they were instructed to press the space bar.
-The response triggered the next trial.
-Accuracy and reaction times (RT) were recorded.
+The response triggered the next trial; note that the program waited until a response occurred: trials did not time-out after a predetermined duration.
+Accuracy and response times (RT) were recorded.
 
@@
 After hearing each stimulus, participants were instructed to report the syllable they heard or if it seemed like they heard two different sounds, the one they heard most clearly by clicking a button corresponding to one of the six syllables.
-The response triggered the start of the next trial.
-Accuracy and reaction times (RT) were recorded.
+The response triggered the start of the next trial; trials did not time-out after a predetermined duration.
+Accuracy and response times (RT) were recorded.
 
@@
 In the example, “happiness” is shown on the left and “sadness” on the right.
-A participant who is right lateralised for emotional processing would be more likely to perceive the emotion in the left hemifield and would, therefore, respond “happiness”.](chim_example.png)
+A participant who is right lateralised for emotional processing would be more likely to perceive the emotion in the left hemifield and would, therefore, respond “happiness”.](chim_example.png){width=20%}
 
@@
 Participants then reported the emotion seen in the face by clicking a button corresponding to one of the four emotions.
-The response triggered the start of the next trial.
-Accuracy and reaction times (RT) were recorded.
+The response triggered the start of the next trial; trials did not time-out after a predetermined duration.
+Accuracy and response times (RT) were recorded.
 
@@
 
-The finger tapping task is a novel behavioural task that we designed as a quantitative index of handedness.
+The finger tapping task is a novel behavioural task that we designed as a quantitative index of handedness analogous to a peg-moving task.
 Participants were required to either press the ‘W’, ‘R’, ‘V’, and ‘X’ keys with their left hand or the ‘T’, 'U', ‘M’, ‘B’ keys with their right hands in sequence as many times as possible within a 30-second window.
@@
 During session 1, participants completed the full battery of tasks.
-Self -report measures were completed first, followed by the behavioural laterality tasks.
+Self-report measures were completed first, followed by the behavioural laterality tasks.
 During session 2, participants completed only the behavioural laterality tasks.
 Hence, each participant provided data for the rhyme decision, dichotic listening, chimeric faces, and finger tapping tasks twice.
-Each session lasted approximately 45 minutes and 35 minutes respectively.
-
-When completing the behavioural tasks participants were instructed to sit approximately 50 cm from the screen.
-Gorilla’s scaling tool was utilised such that stimuli size and eccentricity was maintained across tasks and participants.
+The two sessions lasted approximately 45 minutes and 35 minutes respectively.
+
+When completing the behavioural tasks, participants were instructed to sit approximately 50 cm from the screen.
+Gorilla’s scaling tool, which requires the participant to hold a credit card on the screen while adjusting size to match an image on the screen, was utilised such that stimuli size and eccentricity was maintained across tasks and participants.
+During each task participants were instructed to respond as fast and accurately as possible.
 Stimuli from each task were repeated across four blocks.
@@
     geom_text(aes(label=Count), vjust=-0.3, position = position_dodge(0.9), size=3.5)+
-  theme_classic() + theme(legend.position = "top") + ylab("Count") + xlab("") + ylim(0,220)
+  theme_bw() + theme(legend.position = "top") + ylab("Count") + xlab("") + ylim(0,220)
 ggsave(
@@
 Response times were log-transformed to reduce skew, as specified in our pre-registration.
-Outlier reaction times were identified using Hoaglin and Iglewicz’s (1987) procedure, whereby outliers are defined as response times that are 1.65 times the difference between the first and third quantiles (1.65 * (Q3-Q1)) below or above the first and third quantile values (e.g: lower limit = Q1 – 1.65 x (Q3-Q1); upper limit = Q3 + 1.65 x (Q3-Q1)).
+Outlier response times were identified using Hoaglin and Iglewicz’s (1987) procedure, whereby outliers are defined as response times that are 1.65 times the difference between the first and third quantiles (1.65 * (Q3-Q1)) below or above the first and third quantile values (e.g: lower limit = Q1 – 1.65 x (Q3-Q1); upper limit = Q3 + 1.65 x (Q3-Q1)).
+This criterion was applied to all responses, before subdivision into left- and right-sided responses.
 Our differences criterion was set to 1.65 instead of the typical 2.2 because we expected to obtain asymmetric distributions for response time, with most outliers in one tail of the distribution.
@@
 
-To address hypothesis 2, that extreme left-handers will be more likely to show atypical laterality than right-handers, we categorised participants as extreme left-handers if they scored less than -90 and right-handers if they scored more than 90 on the Edinburgh Handedness Inventory.
-Our preregistration specified that atypicality would first be defined according to whether the participant's LI_z was significantly lateralised (with alpha = .05) in the opposite direction from the majority.
-If there were less than 20 cases defined as atypical using this method, we planned to adjust the cut-off on that task until we acquired 20 atypical participants.
-
-For rhyme decision and chimeric faces task, participants were categorised into four groups: typical on both; atypical on rhyme decision only; atypical on chimeric faces only, and atypical on both.
-A chi-square test was then used to test for association with the number of people in each of the four groups and the handedness classification from the Edinburgh Handedness Inventory.
-This was then repeated for the dichotic listening and the chimeric faces tasks.
-A Bonferroni correction was applied to the outcome of each of the two chi-squared tests.
-
-<!---ZW: This bit confused me.
-
-I thought that the hypothesis was that extreme left handers were more likely to be atypical on any given task, but the test looks at consistency of laterality across pairs of task.
-Isn’t that testing a different hypothesis?
-
-Both are interesting, but I think the first one needs to be tested too.
-E.g., are extreme left handers more likely than right handers to be atypically lateralized on Rhyme Decision / Chimeric Faces / Dichotic Listening?
---->
+To address hypothesis 2, that extreme left-handers will be more likely to show atypical laterality than right-handers, we categorised participants as extreme left-handers if they scored less than -90 on the EHI and right-handers if they scored greater than zero on the EHI.
+Following our pre-registration, atypicality was defined according to whether the participant's LI_z was significantly lateralised (with alpha = .05) in the opposite direction from the majority.
+Chi-square tests were then used to examine the association between handedness and atypicality on the rhyme decision, dichotic listening, and chimeric faces tasks.
+In our pre-registration we had planned to look at cases who were atypical on the language tasks and the chimeric faces.
+We did not conduct this analysis because few participants showed atypical lateralisation across multiple tasks.
+In our Supplementary Materials we do, however, compare laterality indices between extreme left-handers and right-handers to further examine the link between handedness and atypical lateralisation.
 
@@
 
-We examined the relationship between LIs on the rhyme decision, dichotic listening, and chimeric faces tasks by modelling covariances using a subset of Structural Equation Modeling (SEM) that does not include any latent variables or directional paths.
+We examined the relationship between LIs on the rhyme decision, dichotic listening, and chimeric faces tasks by modelling covariances using a simplified version of Structural Equation Modeling (SEM) that used the sum of LIs from the two sessions of each task to represent each of three factors, minimising the impact of measurement error.
 In this approach, we constrain particular covariance patterns and report ‘best’ model fit according to AIC weights (Wagenmakers & Farrell, 2004).
-Akaike weights are transformations of the AIC values, which can be directly interpreted as conditional probabilities for each model.
-The best-fitting model to the data for right- and left-handers is selected as the final model.
-This allowed us to compare three models:
+Akaike weights are transformations of the AIC values, which can be directly interpreted as conditional probabilities for each model, with the best-fitting model being selected.
+We  compared three models:
 
@@
 
-This was completed separately for right- and left-handers using both session 1 and session 2 data.
+This was completed separately for right- and left-handers, as well as for the whole combined group.
 
@@
 
-Foot and eye preference characteristics of the sample after exclusions as noted above are shown in Table 1.
-As participants self-reported eye dominance was always consistent between the Miles and Porta tests, we report data only from the Miles test.
+Descriptive data on foot and eye preference characteristics of the sample after exclusions as noted above are shown in Table 1.
+As these aspects of laterality are not a focus of this study, we will not elaborate further on these data.
 
@@
 
-Table 1 <!---New table - lattab to be added in nice format--->
-Count of left- and right-handers who report left or right on self-reported footedness, and eyedness (Miles test). 
+Table 1. Count of left- and right-handers who report left or right on self-reported footedness, and eyedness (Miles test).
 ```{r Tab1,echo=FALSE, comment=FALSE, warning=FALSE}
+lattab<-lattab[,c(1,2,4,3,5)] #now reorder columns to make sense
+colnames(lattab)<-c('Foot-Eye','Right hander (N)','Left hander (N)','Right hander (%)','Left hander (%)')
 lattab <- flextable(lattab)
+lattab <- align(lattab, align = "center", j= 2:5)
 autofit(lattab, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
@@
     geom_hline(yintercept=meanR,linetype='dotted')+geom_vline(xintercept=meanL,linetype='dotted')+
-    theme(legend.position = "none") + theme_classic() +
+    theme(legend.position = "none") + theme_bw() +
     theme(plot.title = element_text(size = 12)
     )
-  myg<-myg+ scale_color_manual(values=c("black","red"))
+  #myg<-myg+ scale_color_manual(values=c("black","red"))
   return(myg)
@@
 
-Distributions of the raw data are shown in the Online Supplemental Materials.
+Distributions of the raw data are shown in Figure 3.
+The scatterplots indicate scores on the left and right sides for each task, with self-reported handedness (left vs right) colour-coded.
+For each task, the mean for each side is shown as vertical and horizontal dotted lines, and equal performance on the two sides corresponds to the sloping red dotted line: thus an impression of whether the task is lateralised can be obtained by considering whether points cluster around the sloping lines symmetrically.
+
+\newpage
+
+![Figure 3.
+Left side (L) vs right side (R) scores for sessions 1 and 2 for the behavioural laterality tasks.
+The diagonal dotted line shows the point of equality: points above this are higher for R than L, and points below are higher for L than R.
+Mean values for R and L are shown as horizontal and vertical dotted lines.
+The colour of the points indicates the participants’ handedness (pink = right-handed; blue = left-handed).
+](./plots/all.raw.png){width=80%}
+
+For the rhyme decision task, only correct responses were used for computing mean log RTs (shown transformed back to the original ms scale in the plot).
+Generally, accuracy on the rhyme decision task was high.
+For session 1, accuracy was `r round(mean(allsum$RDT1.pcorrL, na.rm= TRUE),1)`% (*SD* = `r round(sd(allsum$RDT1.pcorrL, na.rm= TRUE),1)`%) in the LVF and `r round(mean(allsum$RDT1.pcorrR, na.rm= TRUE),1)`% (*SD* = `r round(sd(allsum$RDT1.pcorrR, na.rm= TRUE),1)`%) in the RVF.
+For session 2, accuracy was `r round(mean(allsum$RDT2.pcorrL, na.rm= TRUE),1)`% (*SD* = `r round(sd(allsum$RDT2.pcorrL, na.rm= TRUE),1)`%) in the LVF and `r round(mean(allsum$RDT2.pcorrR, na.rm= TRUE),1)`% (*SD* = `r round(sd(allsum$RDT2.pcorrR, na.rm= TRUE),1)`%) in the RVF.
+For dichotic listening and chimeric faces, a correct response can be given from either side; it follows that there will be a negative correlation between the two sides reflecting the fact that the number of selections on the left will be inversely related to the number of correct responses on the right.
+For these tasks, it is evident by inspection that most responses for dichotic listening cluster above the dotted line, indicating a right ear advantage, whereas most responses on chimeric faces cluster below the dotted line, indicating a left hemi-face advantage.
+Finally, we show the number of correct sequences with each hand on the finger tapping task.
+Here, a strong impact of handedness is evident on inspection, with left-handers generally more accurate with the left hand, and right-handers more accurate with the right hand.
+
 The distribution of LIs can be used to quantify departures from symmetry.
 Recall that for all tasks, we computed LIs in a typical way: (100 x (Right – Left)/(Right + Left)), and that because of crossed pathways from hands, ears and eyes to the brain, a positive LI indicates left lateralisation and a negative LI indicates right lateralisation at the brain level.
-LIs on the rhyme task were calculated using log RTs for correct responses (accuracy was high; *M* = `r round(allmeanpcorr,1)`%, *SD* = `r round(allsdpcorr,2)` %).
+LIs on the rhyme task were calculated using log RTs for correct responses.
 LIs for the dichotic listening and chimeric faces were calculated using the count of correct responses for sounds/emotions in each ear/visual.
@@
 It can be seen that all measures were significantly lateralised in the predicted direction at the population level, with left-hemisphere superiority for rhyme decision, dichotic listening and finger tapping, and right-hemisphere superiority for chimeric faces.
-
-<!---for ease of interpretation I have chopped up the original big table so we can look at different aspects one step at a time; i.e.
-overall distribution of LIs (different from zero / skew); then the standardized LIs; then the test-retest.--->
 
@@
 
-Table 2.Mean LIs and results from one-sample t-tests and Shapiro-Wilks tests for departure from normality for each task at sessions 1 and 2.
+Table 2.
+Mean LIs and results from one-sample t-tests and Shapiro-Wilks tests for departure from normality for the rhyme decision (RD), dichotic listening (DL), chimeric faces (CF), and finger tapping (FT) tasks at sessions 1 and 2.
 ```{r Tab2,echo=FALSE, comment=FALSE, warning=FALSE}
 tab2 <- flextable(LItable)
+tab2 <- align(tab2, j= 2:7, align= "center")
 autofit(tab2, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
@@
 Table 3 shows data from the LI_z measures, which make it easier to compare different tasks on the same scale.
+The distributions of LI_z scores are shown graphically in the supplementary materials.
 For each task, those with LI_z scores greater than 1.96 are categorised as left-hemisphere dominant, and those with LI_z scores less than -1.96 are categorised as right-hemisphere dominant.
@@
 
-Table 3. Mean LI_z and percentage of participants who were significantly left lateralised (z > 1.96) or right lateralised (z < -1.96) for each task at sessions 1 and 2.
+Table 3.
+Mean LI_z and percentage of participants who were significantly left lateralised (z > 1.96) or right lateralised (z < -1.96) on the rhyme decision (RD), dichotic listening (DL), chimeric faces (CF), and finger tapping (FT) tasks at sessions 1 and 2.
 ```{r Tab3,echo=FALSE, comment=FALSE, warning=FALSE}
 tab3 <- flextable(LIztable)
+tab3 <- align(tab3, align= "center", j= 2:6)
 autofit(tab3, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
@@
   # plots
-  dens1<-ggplot(dat, aes(x=time1))+geom_density(linetype="dashed")+theme_classic()+ggtitle(name1)
+  dens1<-ggplot(dat, aes(x=time1))+geom_density(linetype="dashed")+theme_bw()+ggtitle(name1)
   dens1 <- dens1+ geom_vline(xintercept = 0)
@@
   # plot density
-  dens2<-ggplot(dat, aes(x=time2))+geom_density(linetype="dashed")+theme_classic()+ggtitle(name2)
+  dens2<-ggplot(dat, aes(x=time2))+geom_density(linetype="dashed")+theme_bw()+ggtitle(name2)
   dens2 <- dens2+ geom_vline(xintercept = 0)
@@
   #now do the plot
-  myscatter <- ggplot(dat, aes(x= time1, y= time2)) +
+  myscatter <<- ggplot(dat, aes(x= time1, y= time2)) +
     geom_hline(yintercept=0, color = "black") +
@@
     geom_rug(aes(color= handed3)) +
-    theme_classic(18) +
+    theme_bw(18) +
     ylab(name2) + 
@@
 time2<-RD.dat$zlat_RDT_2
-name1<-"Rhyme decision LI_z (day 1)"
-name2<-"Rhyme decision LI_z (day 2)"
+name1<-"Rhyme decision LI_z (session 1)"
+name2<-"Rhyme decision LI_z (session 2)"
 scattername<-"RhymeDecisionScatter"
@@
 remove= 5
-mycors<-densplots(RD.dat, time1,time2,name1,name2,scattername, y1, y2, x1, x2, ax, ay, remove, "(A)", 241) 
+title= "(A)"
+N=241
+mycors<-densplots(RD.dat, time1,time2,name1,name2,scattername, y1, y2, x1, x2, ax, ay, remove, title, N) 
 combotable[1,2]<-round(mycors[[1]],3)
 combotable[1,3]<-mycors[[2]]
+RDT.plot <- myscatter
 
@@
 time2<-DL.dat$Day2_Dich_acc_Z
-name1<-"Dichotic listening LI_z (day 1)"
-name2<-"Dichotic listening LI_z (day 2)"
+name1<-"Dichotic listening LI_z (session 1)"
+name2<-"Dichotic listening LI_z (session 2)"
 scattername<-"DichoticScatter"
-mycors<-densplots(DL.dat, time1,time2,name1,name2,scattername,y1, y2, x1, x2, ax, ay, remove,"(B)", 290)
+title="(B)"
+N=290
+mycors<-densplots(DL.dat, time1,time2,name1,name2,scattername,y1, y2, x1, x2, ax, ay, remove,title,N)
 combotable[2,2]<-round(mycors[[1]],3)
 combotable[2,3]<-mycors[[2]]
+DL.plot <- myscatter
 
@@
 time2<-CF.dat$Day2_CF_acc_Z
-name1<-"Chimeric faces LI_z (day 1)"
-name2<-"Chimeric faces LI_z (day 2)"
+name1<-"Chimeric faces LI_z (session 1)"
+name2<-"Chimeric faces LI_z (session 2)"
 scattername<-"ChimericScatter"
-mycors<-densplots(CF.dat, time1,time2,name1,name2,scattername, y1, y2, x1, x2, ax, ay, remove, "(C)", 289)
+title="(C)"
+N=289
+mycors<-densplots(CF.dat, time1,time2,name1,name2,scattername, y1, y2, x1, x2, ax, ay, remove,title,N)
 combotable[3,2]<-round(mycors[[1]],3)
 combotable[3,3]<-mycors[[2]]
+CF.plot <- myscatter
 
@@
 time2<-FT.dat$Day2_Finger_acc_Z
-name1<-"Finger tapping LI_z (day 1)"
-name2<-"Finger tapping LI_z (day 2)"
+name1<-"Finger tapping LI_z (session 1)"
+name2<-"Finger tapping LI_z (session 2)"
 scattername<-"FingerScatter"
-mycors<-densplots(FT.dat, time1,time2,name1,name2,scattername, y1, y2, x1, x2, ax, ay, remove, "(D)", 288)
+title="(D)"
+N=288
+mycors<-densplots(FT.dat, time1,time2,name1,name2,scattername, y1, y2, x1, x2, ax, ay, remove,title,N)
 combotable[4,2]<-round(mycors[[1]],3)
 combotable[4,3]<-mycors[[2]]
-
-```
-
-Table 4. Test-retest reliability estimates and 95% CIs for each behavioural task.
+FT.plot <- myscatter
+
+combine.plots <- ggarrange(RDT.plot, DL.plot, CF.plot, FT.plot)
+ggsave("./plots/all_LIs.png", combine.plots, width = 12, height = 12)
+```
+
+Table 4. Test-retest reliability estimates and 95% CIs for the rhyme decision (RD), dichotic listening (DL), chimeric faces (CF), and finger tapping (FT) tasks.
 ```{r Tab4,echo=FALSE, comment=FALSE, warning=FALSE}
+combotable$Task[1] <- "RD"
+combotable$Task[2] <- "DL"
+combotable$Task[3] <- "CF"
+combotable$Task[4] <- "FT"
+
+combotable$`95% CI`[3] <- "0.84 - 0.90"
+combotable$`95% CI`[4] <- "0.68 - 0.80"
+
 tab4 <- flextable(combotable)
+tab4 <- align(tab4, align= "center", j= 2:3)
 autofit(tab4, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
@@
 
-![Figure 3.
-Scatter plots for day 1 and day 2 data for LI_z scores on each laterality task: (A) rhyme decision, (B) dichotic listening, (C) chimeric faces, and (D) Finger tapping.
+\newpage
+
+![Figure 4.
+Scatter plots for session 1 and session 2 data for LI_z scores on each laterality task: (A) rhyme decision, (B) dichotic listening, (C) chimeric faces, and (D) Finger tapping.
 Handedness is colour coded with extreme left-handers identified as those who have a score below -90 on the Edinburgh Handedness Inventory.
-Note that z-scores are censored at absolute values of 5.](combined_LI_plot.png)
+Note that z-scores are censored at absolute values of 5.](./plots/all_LIs.png){width=80%}
 
 ### Optimising the tasks
-<!--- DB: I think the key issue is not so much getting reliability of .65, as how many trials could be dropped without any loss of reliability from whatever high level it got to.
-A bit of rewording to reflect this.
-Also modified in discussion--->
+
 The dichotic listening and chimeric faces tasks had test-retest reliability above *r* = .75, and so we investigated whether the tasks could be made more efficient by reducing the number of trials while retaining good test-retest reliability.
@@
 
-![Figure 4.
+![Figure 5.
 Test-retest correlations for (A) the dichotic listening task and (B) the chimeric faces task when using a variable number of trials to calculate laterality indices.
 The red dashed line marks the test-retest correlation *r* = .65.
-The correlations coefficients are shown as absolute values.](all.cor.plot.png)
+The correlations coefficients are shown as absolute values.](./plots/all.cor.plot.png){width=70%}
 
@@
 #
-LIs$rhyme1.cut <- cut(LIs$Day1.RDT_p.corr_LI, 
+LIs$rhyme1.cut <- cut(LIs$zlat_RDT_1, 
                        breaks = c(-Inf, -1.96, 1.96, Inf), 
@@
 
-Table 5. Frequencies of laterality typicality for extreme left-handers, moderate left-handers and right-handers across the rhyme decision, dichotic listening, and chimeric faces tasks.
+Table 5. Frequencies of laterality typicality for extreme left-handers, moderate left-handers and right-handers across the rhyme decision (RD), dichotic listening (DL), and chimeric faces (CF) tasks.
 ```{r Table2,echo=FALSE, comment=FALSE, warning=FALSE}
+tabr<-cbind(tabr,round(100*prop.table(tabr,1),1))
+tabd<-cbind(tabd,round(100*prop.table(tabd,1),1) )
+tabc<-cbind(tabc,round(100*prop.table(tabc,1),1)) #add percentages
+latcats <- as.data.frame(rbind(tabr,tabd,tabc))
+
+row.names(latcats) <- c() #remove rownames as they make problems for printing
+
+#Instead have a new column with these
+latcats$nucol <- c('Extreme left-hander',
+                   'Moderate left-hander',
+                   'Right-hander',
+                   'Extreme left-hander',
+                   'Moderate left-hander',
+                   'Right-hander',
+                   'Extreme left-hander',
+                   'Moderate left-hander',
+                   'Right-hander')
+latcats$nucol2 <- c('RD',
+                   ' ',
+                   ' ',
+                   'DL',
+                   ' ',
+                   ' ',
+                   'CF',
+                   ' ',
+                   ' ')
+
+latcats<-latcats[,c(8,7,3,2,1,6,5,4)] #now reorder columns to make sense
+colnames(latcats)<-c('Task','Sub-group','Left N','Non N','Right N','Left %','Non %','Right %')
+
+myft2 <- flextable(latcats)
+#NB This table needs formatting - has formatted all numbers to 3 decimal places, which is not what we want. Code below achieves this
+
+num_keys <- c("Left %","Non %","Right %")
+int_keys <- c("Left N","Non N","Right N")
+
+ft <- colformat_num(x = myft2, col_keys = num_keys, digits = 1)
+ft <- colformat_int(x = ft, col_keys = int_keys, big.mark = ",")
+autofit(ft)
+```
+*Note*. N is the count of participants. % is the percentage of participants.
+
+```{r pformatfunction, echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
+pformat<-function(pvalue){
+  pp <- "*p* < .001" #default value
+  if (pvalue>.001) {
+    pp<-paste0("*p* = ",round(pvalue,3))
+  }
+  return(pp)
+}
+```
+
+The rhyme decision task has few individuals who are significantly lateralised in either direction.
+We will therefore not consider it further, as our analyses are only justified if the task is an indicator of cerebral lateralisation.
+Both dichotic listening and chimeric faces, on the other hand, have a high proportion of strongly lateralised individuals.
+The planned analyses, comparing typically lateralised with atypically lateralised individuals, and extreme left vs right-handed individuals, show a significant association between extreme left-handedness and atypical laterality on a chi-square test.
+For dichotic listening, chi-square = `r round(chi.d$statistic,2)`, d.f.
+= 1, `r pformat(p.adjust(chi.d$p.value, method= "bonferroni", n=2))`; for chimeric faces, chi-square = `r round(chi.c$statistic,2)`, d.f.
+= 1, `r pformat(p.adjust(chi.c$p.value, method= "bonferroni", n=2))`.
+These data provide statistical support for the prediction from Mazoyer et al (2014) that extreme left-handers have an increased rate of atypical cerebral lateralisation.
+
+```{r contingency2tasks,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
+tab2task <- table(LIs$dichotic1.cut,LIs$chimeric1.cut)
+chi2task<-chisq.test(tab2task)
+w1<-which(LIs$chimeric1.cut=='TypLat') #this is confusingnly mislabelled
+w2<-which(LIs$dichotic1.cut=='AtypLat')
+w3<-intersect(w1,w2)
+atypical2<-LIs$handed3[w3] #this is hardcoded in the text below
+```
+
+The planned analysis looking at cases who were atypical on both dichotic listening and chimeric faces was hampered by the fact that the two tasks were not associated (see below), and consequently, there were only four cases in this category.
+Nevertheless, it was noteworthy that two of these were extreme left-handers and the other two were less extreme left-handers.
+
+### Relationship between LIs from different tasks
+
+Our final question was whether the laterality indices from the language and faces laterality tasks correlated.
+Specifically, we had predicted that in right-handers the two language tasks would correlate together, but the chimeric faces would form a separate factor.
+In left-handers, we predicted that the language tasks would fractionate so that none of the three tasks would correlate.
+
+Initial scrutiny of Spearman correlations between LI_z scores suggests weak associations between LI_z values for the two language measures (rhyme decision and dichotic listening), but no association between either of these measures and LI_z on chimeric faces.
+Note, however, that correlations will be influenced by test reliability; disattenuated correlations (divided by the square root of the product of the reliabilities of the two measures) will be higher (Muchinsky, 1996).
+
+```{r bigcorrmatrix,echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
+
+# create subset of data frame for analysis - I used LI_z measures, as they are on similar scale
+interRR <- dplyr::select(LIs, 'zlat_RDT_1', 'zlat_RDT_2', 'Day1_Dich_acc_Z',
+                         'Day2_Dich_acc_Z', 'Day1_CF_acc_Z', 'Day2_CF_acc_Z','handedness')
+# rename
+interRR <- interRR %>% 
+  rename(
+    "Rhyme_1" = zlat_RDT_1,
+    "Chimeric_1" = Day1_CF_acc_Z,
+    "Dichotic_1" = Day1_Dich_acc_Z,
+     "Rhyme_2" = zlat_RDT_2,
+    "Chimeric_2" = Day2_CF_acc_Z,
+    "Dichotic_2" = Day2_Dich_acc_Z
+    )
+
+#remove cases with missing values
+interRR<-na.omit(interRR)
+# subset into right and left handers
+left_dat <- subset(interRR, handedness== "Left")
+right_dat <- subset(interRR, handedness== "Right")
+left_dat<-left_dat[,1:6]
+right_dat<-right_dat[,1:6]
+totcorr<-correlate(interRR[,1:6],method='spearman')
+Rcorr <- correlate(right_dat,method='spearman')
+Lcorr <- correlate(left_dat,method='spearman')
+
+#Change so we have one matrix with R handers above diagonal and L handers below
+Lcorr<-as.data.frame(Lcorr[,2:7])
+Rcorr <- as.data.frame(Rcorr[,2:7])
+allcorr<-Lcorr
+for (i in 1:6){
+  for (j in i:6){
+    allcorr[i,j]<-Rcorr[i,j]
+
+  }
+}
+allcorr<- round(allcorr,3)
+row.names(allcorr)<-colnames(allcorr)
+
+NL<-nrow(left_dat)
+NR<-nrow(right_dat)
+#for left_dat N = 66
+# right_dat
+```
+
+Table 6.
+Spearman correlations between the LI_z scores for the different tasks in two Sessions.
+The upper diagonal shows correlations for the `r NR` right-handers (black font), and the values of the lower diagonal show for the `r NL` left-handers (red font).
+```{r tab 5, echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
+allcorr2 <- cbind(Row.Names = list(rownames(allcorr)), allcorr, stringsAsFactors = FALSE)
+
+# rename columns
+names(allcorr2)[1] <- " "
+allcorr2$` `[1] <- "RD 1"
+allcorr2$` `[2] <- "RD 2"
+allcorr2$` `[3] <- "DL 1"
+allcorr2$` `[4] <- "DL 2"
+allcorr2$` `[5] <- "CF 1"
+allcorr2$` `[6] <- "CF 2"
+
+# rename rows
+names(allcorr2)[2] <- "RD 1"
+names(allcorr2)[3] <- "RD 2"
+names(allcorr2)[4] <- "DL 1"
+names(allcorr2)[5] <- "DL 2"
+names(allcorr2)[6] <- "CF 1"
+names(allcorr2)[7] <- "CF 2"
+
+# replace blanks with '.'
+#allcorr2$`RD 1`[1] <- "."
+#allcorr2$`RD 2`[2] <- "."
+#allcorr2$`DL 1`[3] <- "."
+#allcorr2$`DL 2`[4] <- "."
+#allcorr2$`CF 1`[5] <- "."
+#allcorr2$`CF 2`[6] <- "."
+
+myft5<- flextable(allcorr2)
+myft5 <- colformat_num(myft5, digits= 3)
+
+myft5<-color(myft5, i = c(2,3,4,5, 6), j = "RD 1", color= "red")
+myft5<-color(myft5, i = c(3,4,5, 6), j = "RD 2", color= "red")
+myft5<-color(myft5, i = c(4,5, 6), j = "DL 1", color= "red")
+myft5<-color(myft5, i = c(5, 6), j = "DL 2", color= "red")
+myft5<-color(myft5, i = c(6), j = "CF 1", color= "red")
+
+autofit(myft5, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
+```
+*Note*.
+For right-handers, an absolute value of .16 is significant at the .05 level, and a value of .22 is significant at the .01 level.
+For left-handers, a value of .25 is significant at the .05 level and a value of .32 is significant at the .01 level.
+
+In a more formal test, we compared three models: (a) where the two language tasks are correlated with each other but not with the chimeric faces (i.e.
+a language laterality factor); (b) all LIs are correlated (i.e.
+a single laterality factor); and (c) a base model where the LIs are unrelated (i.e.
+independent functions).
+We departed from our pregistered analysis in two respects on the basis of reviewer recommendations.
+First we show results for the full sample as well as for separate handedness groups; second we adopt an alternative approach to the Akaike weights analysis that takes measurement error into account by creating factor scores based on the summed scores for the two test sessions.
+
+```{r prepareKievit, echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
+library(lavaan)
+library(tidyverse)
+library(MASS)
+library(qpcR)
+# created handedness data
+# now create subset of data frame for analysis
+interRR <- dplyr::select(LIs, split.RDT.1.1, split.RDT.1.2, split.RDT.2.1, split.RDT.2.2,
+                              split.DL.1.1, split.DL.1.2, split.DL.2.1, split.DL.2.2,
+                              split.chim.1.1, split.chim.1.2, split.chim.2.1, split.chim.2.2, handedness)
+# rename, and reverse sign for CF
+interRR <- interRR %>% 
+  rename(
+    "rhyme11" = split.RDT.1.1,
+    "faces11" = split.chim.1.1,
+    "dichotic11" = split.DL.1.1,
+     "rhyme12" = split.RDT.1.2,
+    "faces12" = split.chim.1.2,
+    "dichotic12" = split.DL.1.2,
+      "rhyme21" = split.RDT.2.1,
+    "faces21" = split.chim.2.1,
+    "dichotic21" = split.DL.2.1,
+     "rhyme22" = split.RDT.2.2,
+    "faces22" = split.chim.2.2,
+    "dichotic22" = split.DL.2.2
+    )
+
+#remove cases with missing values
+interRR<-na.omit(interRR)
+interRR$faces11 <- (-interRR$faces11) 
+interRR$faces12 <- (-interRR$faces12)
+interRR$faces21 <- (-interRR$faces21) 
+interRR$faces22 <- (-interRR$faces22)
+# subset into right and left handers
+left_dat <- subset(interRR, handedness== "Left")
+right_dat <- subset(interRR, handedness== "Right")
+
+```
+
+As shown in Table 7, the Akaike weights favoured model B for both left-handers and right-handers, as well as for both groups combined, confirming that there is association for laterality indices for the two language tasks, but not between language latearlity and laterality for face processing of emotion.
+
+Table 7. Akaike weights for the language model, independent base model, and laterality model.
+```{r disattenuateR, ,echo=FALSE, comment=FALSE, warning=FALSE}
+# First check range of scores on all variables for left and right-handers to address point re restriction of range
+rangedf <- data.frame(matrix(NA,nrow=3,ncol=4))
+colnames(rangedf)<-c('Llower','Lupper','Rlower','Rupper')
+rownames(rangedf)<-c('Rhyme','Dichotic','Chimeric')
+thisrow<-0
+for (myc in c(1,3,5)){
+  thisrow<-thisrow+1
+  rangedf[thisrow,1:2]<-range(left_dat[myc:(myc+1)])
+  rangedf[thisrow,3:4]<-range(right_dat[myc:(myc+1)])
+}
+
+#Now for combined handedness group, compute disattenuated correlations
+#We already have computed totcorr as symmetric matrix
+totcorr<-as.data.frame(totcorr)
+r.rhyme<-totcorr[2,2]
+r.dich<-totcorr[4,4]
+r.chim<-totcorr[6,6]
+
+nutot<-totcorr
+for (c in 2:3){ #col 1 is name of task
+  for (r in 3:4){
+    nutot[r,c]<-totcorr[r,c]/sqrt(r.rhyme*r.dich)
+  }
+    for (r in 5:6){
+    nutot[r,c]<-totcorr[r,c]/sqrt(r.rhyme*r.chim)
+  }
+}
+for (c in 4:5){
+  for (r in 5:6){
+    nutot[r,c]<-totcorr[r,c]/sqrt(r.dich*r.chim)
+  }
+}
+
+#nutot shows disattenuated correlations below diagonal
+```
+
+```{r newAkaike_factor,echo=FALSE, comment=FALSE, warning=FALSE}
+#New version of Akaike test combines 2 test scores for each into a factor to deal with measurement error
+
+#Model 1:all correlated
+    model1F<-"
+    rhymeFACTOR =~ NA*rhyme11 + rhyme12 + rhyme21 + rhyme 22
+    dichoticFACTOR =~ NA*dichotic11 + dichotic12 + dichotic21 + dichotic22
+    facesFACTOR =~ NA*faces11 +faces12 + faces21 +faces22
+
+    rhymeFACTOR~~1*rhymeFACTOR
+    dichoticFACTOR~~1*dichoticFACTOR
+    facesFACTOR~~1*facesFACTOR
+
+    rhymeFACTOR~~dichoticFACTOR
+    rhymeFACTOR~~facesFACTOR
+    dichoticFACTOR~~facesFACTOR
+    "
+
+#Model 2: language tasks are correlated
+    model2F<-"
+    rhymeFACTOR =~ NA*rhyme11 + rhyme12 + rhyme21 + rhyme 22
+    dichoticFACTOR =~ NA*dichotic11 + dichotic12 + dichotic21 + dichotic22
+    facesFACTOR =~ NA*faces11 +faces12 + faces21 +faces22
+
+    rhymeFACTOR~~1*rhymeFACTOR
+    dichoticFACTOR~~1*dichoticFACTOR
+    facesFACTOR~~1*facesFACTOR
+
+    rhymeFACTOR~~dichoticFACTOR
+    rhymeFACTOR~~0*facesFACTOR
+    dichoticFACTOR~~0*facesFACTOR
+    "
+ #Model 3 all independent
+    model3F<-"
+
+    rhymeFACTOR =~ NA*rhyme11 + rhyme12 + rhyme21 + rhyme 22
+    dichoticFACTOR =~ NA*dichotic11 + dichotic12 + dichotic21 + dichotic22
+    facesFACTOR =~ NA*faces11 +faces12 + faces21 +faces22
+
+    rhymeFACTOR~~1*rhymeFACTOR
+    dichoticFACTOR~~1*dichoticFACTOR
+    facesFACTOR~~1*facesFACTOR
+
+    rhymeFACTOR~~0*dichoticFACTOR
+    rhymeFACTOR~~0*facesFACTOR
+    dichoticFACTOR~~0*facesFACTOR
+    "
+#Make a little dataframe for AIC weights
+AICdf <- data.frame(matrix(NA,nrow=3,ncol=4))
+AICdf[,1]<-c('Whole sample','Right-handers','Left-handers')
+colnames(AICdf)<-c('Handedness','Model 1F','Model 2F','Model 3F')
+for (myhand in 1:3) {
+  mydat<-interRR
+   if(myhand==2) {mydat<-right_dat}
+  if(myhand==3) {mydat<-left_dat}
+
+ #Using data from this handedness type, test all 3 models
+fit1A <- cfa(model1F, data=mydat,auto.fix.first=FALSE)
+fit2A <- cfa(model2F, data=mydat,auto.fix.first=FALSE)
+fit3A <- cfa(model3F, data=mydat,auto.fix.first=FALSE)
+    aic_vector1 <- c(fitMeasures(fit1A, "aic"),fitMeasures(fit2A, "aic"),fitMeasures(fit3A, "aic"))
+
+    #Confirms that for each model, lowest AIC obtained when the simulated data do match the model
+    # 
+    # The AIC weights make massive difference for preferred model!
+    AICweights<-akaike.weights(aic_vector1)$weights
+    # 
+ AICdf[myhand,2:4]<-round(AICweights,3)
+ }
+myft6 <- flextable(AICdf, col_keys = c("Handedness",
+                                       "Model 1F",
+                                       "Model 2F", 
+                                       "Model 3F"))
+autofit(myft6, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
+```
+
+### Exploratory analyses
+
+#### Finger tapping and Edinburgh Handedness Inventory
+
+```{r fingerEHI,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
+spearCI <- RVAideMemoire::spearman.ci(LIs$Day1_finger_LI, LIs$index_EHI, nrep = 10000, conf.level = 0.95)
+spearCI2 <- RVAideMemoire::spearman.ci(LIs$Day2_finger_LI, LIs$index_EHI, nrep = 10000, conf.level = 0.95)
+# spearman's test
+speartest <- cor.test(LIs$Day1_finger_LI, LIs$index_EHI, method = "spearman", alternative = "greater")
+speartest2 <- cor.test(LIs$Day2_finger_LI, LIs$index_EHI, method = "spearman", alternative = "greater")
+```
+
+To examine the relationship between the finger tapping task and a conventional hand preference battery, we computed Spearman's rank correlation between LIs produced on the finger tapping task and the Edinburgh Handedness Inventory.
+For session 1 data, the relationship between the tasks was moderate, *r* = `r round(speartest$estimate, 2)`, *95% CI* [`r round(spearCI$conf.int[1], 2)`, `r round(spearCI$conf.int[2], 2)`], *p* < 0.001.
+For session 2, the correlation was *r* = `r round(speartest2$estimate, 2)`, *95% CI* [`r round(spearCI2$conf.int[1], 2)`, `r round(spearCI2$conf.int[2], 2)`], *p* < 0.001.
+This suggests that the influence of hand preference on the tapping task may increase with practice.
+
+<!--- Based on reviewer's comments. Haven't added in yet, but holding it here --->
+
+```{r makecontingencies_tasks,echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
+#We start by using z-scores to make a 3-way categorisation into:
+#1 sig bias in typical direction; 2: NS bias 3: in atypical direction with sig bias.
+#
+LIs$rhyme1.cat <- cut(LIs$zlat_RDT_1, 
+                       breaks = c(-Inf, -1.96, 1.96, Inf), 
+                       labels = c("R.Right", "R.Bilat", "R.Left"), 
+                       right = FALSE)
+LIs$dichotic1.cat <- cut(LIs$Day1_Dich_acc_Z, 
+                       breaks = c(-Inf, -1.96, 1.96, Inf), 
+                       labels = c("D.Right", "D.Bilat", "D.Left"), 
+                       right = FALSE)
+LIs$chimeric1.cat <- cut(LIs$Day1_CF_acc_Z, 
+                       breaks =  c(-Inf, 0, 1.96, Inf), 
+                       labels = c("C.Right", "C.Bilat", "C.Right"), 
+                       right = FALSE)
+mytabr <- table(LIs$rhyme1.cat)
+mytabd <- table(LIs$dichotic1.cat)
+mytabc <- table(LIs$chimeric1.cat)
+mytabc<-mytabc[c(3,2,1)]
+mytab2 <-rbind(mytabr,mytabd,mytabc)
+row.names(mytab2)<-c('Rhyme decision','Dichotic','Chimeric Faces')
+mytab2<-data.frame(mytab2)
+mytab2$R.Right[3] <- 0
+
+# rhyme vs dichotic
+tab.R.D <- table(LIs$rhyme1.cat,LIs$dichotic1.cat)
+tab.R.D
+chisq.test(tab.R.D)
+
+# rhyme vs chimeric
+tab.R.C <- table(LIs$rhyme1.cat,LIs$chimeric1.cat)
+tab.R.C
+chisq.test(tab.R.C)
+
+# dichotic vs chimeric
+tab.D.C <- table(LIs$dichotic1.cat,LIs$chimeric1.cat)
+tab.D.C
+```
+
+```{r Table_reviews,echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
 tabr<-cbind(tabr,round(100*prop.table(tabr,2),1))
@@
 
-autofit(ft)
-```
-*Note*. N is the count of participants. % is the percentage of participants.
-
-```{r pformatfunction, echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
-pformat<-function(pvalue){
-  pp <- "*p* < .001" #default value
-  if (pvalue>.001) {
-    pp<-paste0("*p* = ",round(pvalue,3))
-  }
-  return(pp)
-}
-```
-
-The rhyme decision task has few individuals who are significantly lateralised in either direction.
-We will therefore not consider it further, as our analyses are only justified if the task is an indicator of cerebral lateralisation.
-Both dichotic listening and chimeric faces, on the other hand, have a high proportion of strongly lateralised individuals.
-The planned analyses, comparing typically lateralised with atypically lateralised individuals, and extreme left vs right-handed individuals, show a significant association between extreme left-handedness and atypical laterality on chi-square test.
-For dichotic listening, chi-squared = `r round(chi.d$statistic,2)`, d.f.
-= 1, `r pformat(p.adjust(chi.d$p.value, method= "bonferroni", n=2))`; for chimeric faces, chi-squared = `r round(chi.c$statistic,2)`, d.f.
-= 1, `r pformat(p.adjust(chi.c$p.value, method= "bonferroni", n=2))`.
-
-With all participants included, on the dichotic listening task, `r round(100*tabd[1,3]/sum(tabd[,3]),1)`% of those with atypical laterality were extreme left-handers, `r round(100*tabd[3,3]/sum(tabd[,3]),1)`% were right-handers, and the remainder less extreme left-handers.
-For those with typical lateralisation, the corresponding figures were `r round(100*tabd[1,1]/sum(tabd[,1]),1)`% extreme left-handers, `r round(100*tabd[3,1]/sum(tabd[,1]),1)`% right-handers, and the remainder less extreme left-handers.
-For chimeric faces, the association was even more extreme.
-`r round(100*tabc[1,1]/sum(tabc[,1]),1)`% of those with atypical laterality were extreme left-handers, `r round(100*tabc[3,1]/sum(tabc[,1]),1)`% were right-handers, and the remainder less extreme left-handers.
-For those with typical lateralisation, `r round(100*tabc[1,3]/sum(tabc[,3]),1)`% were extreme left-handers, `r round(100*tabc[3,3]/sum(tabc[,3]),1)`% were right-handers, and the remainder less extreme left-handers.
-
-```{r contingency2tasks,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
-tab2task <- table(LIs$dichotic1.cut,LIs$chimeric1.cut)
-chi2task<-chisq.test(tab2task)
-w1<-which(LIs$chimeric1.cut=='TypLat') #this is confusingnly mislabelled
-w2<-which(LIs$dichotic1.cut=='AtypLat')
-w3<-intersect(w1,w2)
-atypical2<-LIs$handed3[w3] #this is hardcoded in the text below
-```
-
-The planned analysis looking at cases who were atypical on both dichotic listening and chimeric faces was hampered by the fact that the two tasks were not associated (see below), and consequently, there were only four cases in this category.
-Nevertheless, it was noteworthy that two of these were extreme left-handers and the other two were less extreme left-handers.
-
-### Relationship between LIs from different tasks
-
-Our final question was whether the laterality indices from the language and faces laterality tasks correlated.
-Specifically, we had predicted that in right-handers the two language tasks would correlate together, but the chimeric faces would form a separate factor.
-In left-handers, we predicted that the language tasks would fractionate so that none of the three tasks would correlate.
-
-Initial scrutiny of Spearman correlations between LI_z scores suggests little evidence for any association between the three laterality indices, as shown in Table 6.
-Table 6 suggests weak associations between LI_z values for the two language measures (rhyme decision and dichotic listening), but no association between either of these measures and LI_z on chimeric faces.
-
-```{r bigcorrmatrix,echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
-
-# create subset of data frame for analysis - I used LI_z measures, as they are on similar scale
-interRR <- dplyr::select(LIs, 'zlat_RDT_1', 'zlat_RDT_2', 'Day1_Dich_acc_Z',
-                         'Day2_Dich_acc_Z', 'Day1_CF_acc_Z', 'Day2_CF_acc_Z','handedness')
-# rename
-interRR <- interRR %>% 
-  rename(
-    "Rhyme_1" = zlat_RDT_1,
-    "Chimeric_1" = Day1_CF_acc_Z,
-    "Dichotic_1" = Day1_Dich_acc_Z,
-     "Rhyme_2" = zlat_RDT_2,
-    "Chimeric_2" = Day2_CF_acc_Z,
-    "Dichotic_2" = Day2_Dich_acc_Z
-    )
-
-#remove cases with missing values
-interRR<-na.omit(interRR)
-# subset into right and left handers
-left_dat <- subset(interRR, handedness== "Left")
-right_dat <- subset(interRR, handedness== "Right")
-left_dat<-left_dat[,1:6]
-right_dat<-right_dat[,1:6]
-totcorr<-correlate(interRR[,1:6],method='spearman')
-Rcorr <- correlate(right_dat,method='spearman')
-Lcorr <- correlate(left_dat,method='spearman')
-
-#Change so we have one matrix with R handers above diagonal and L handers below
-Lcorr<-as.data.frame(Lcorr[,2:7])
-Rcorr <- as.data.frame(Rcorr[,2:7])
-allcorr<-Lcorr
-for (i in 1:6){
-  for (j in i:6){
-    allcorr[i,j]<-Rcorr[i,j]
-
-  }
-}
-allcorr<- round(allcorr,3)
-row.names(allcorr)<-colnames(allcorr)
-
-NL<-nrow(left_dat)
-NR<-nrow(right_dat)
-#for left_dat N = 66
-# right_dat
-```
-
-Table 6.
-Spearman correlations between the LI_z scores for the different tasks in two Sessions.
-The upper diagonal shows correlations for the 143 right-handers (black font), and the values of the lower diagonal show for the 66 left-handers (red font).
-```{r tab 5, echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
-allcorr2 <- cbind(Row.Names = rownames(allcorr), allcorr)
-names(allcorr2)[1] <- " "
-allcorr2[is.na(allcorr2)] <- "."
-myft5<- flextable(allcorr2)
-
-myft5<-color(myft5, i = c(2,3,4,5, 6), j = "Rhyme_1", color= "red")
-myft5<-color(myft5, i = c(3,4,5, 6), j = "Rhyme_2", color= "red")
-myft5<-color(myft5, i = c(4,5, 6), j = "Dichotic_1", color= "red")
-myft5<-color(myft5, i = c(5, 6), j = "Dichotic_2", color= "red")
-myft5<-color(myft5, i = c(6), j = "Chimeric_1", color= "red")
-
-autofit(myft5, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
-```
-Note.
-For right-handers, an absolute value of .16 is significant at the .05 level, and a value of .22 is significant at the .01 level.
-For left-handers, a value of .25 is significant at the .05 level and a value of .32 is significant at the .01 level.
-
-In a more formal test, we compared three models: (a) where the two language tasks are correlated with each other but not with the chimeric faces (i.e.
-a language laterality factor); (b) all LIs are correlated (i.e.
-a single laterality factor); and (c) a base model where the LIs are unrelated (i.e.
-independent functions).
-This was done separately for right- and left-handers, using a subset of SEM where we constrain particular covariance patterns and report ‘best’ model fit according to Akaike weights, where a larger weight indicates a better fit.
-The best-fitting model to the data for right- and left-handers is selected as the final model.
-
-```{r prepareKievit, echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
-library(lavaan)
-library(tidyverse)
-library(MASS)
-library(qpcR)
-# created handedness data
-# now create subset of data frame for analysis
-interRR <- dplyr::select(LIs, zlat_RDT_1, zlat_RDT_2, Day1_Dich_acc_Z,
-                         Day2_Dich_acc_Z, Day1_CF_acc_Z, Day2_CF_acc_Z,handedness)
-# rename, and reverse sign for CF
-interRR <- interRR %>% 
-  rename(
-    "rhyme1" = zlat_RDT_1,
-    "faces1" = Day1_CF_acc_Z,
-    "dichotic1" = Day1_Dich_acc_Z,
-     "rhyme2" = zlat_RDT_2,
-    "faces2" = Day2_CF_acc_Z,
-    "dichotic2" = Day2_Dich_acc_Z
-    )
-
-#remove cases with missing values
-interRR<-na.omit(interRR)
-interRR$faces1 <- (-interRR$faces1) #reverse sign for faces
-interRR$faces2 <- (-interRR$faces2)
-# subset into right and left handers
-left_dat <- subset(interRR, handedness== "Left")
-right_dat <- subset(interRR, handedness== "Right")
-
-```
-
-As shown in Table 7, the Akaike weights favoured model c for both left-handers and right-handers, indicating that lateralisation across tasks is largely independent.
-This reflected the lack of cross-task correlation for LI_z measures.
-
-Table 7. Akaike weights for the independent base model, language model, and laterality model.
-```{r AICmodels,echo=FALSE, comment=FALSE, warning=FALSE}
-#NB in pre-reg we just had time 1 data, but here we use time 1 and 2
-
- #Model 1 all independent
-    model1<-"
-    rhyme1~~rhyme2
-    rhyme1~~0*dichotic1
-    rhyme1~~0*dichotic2
-     rhyme1~~0*faces1
-    rhyme1~~0*faces2
-    rhyme2~~0*dichotic1
-    rhyme2~~0*dichotic2
-     rhyme2~~0*faces1
-    rhyme2~~0*faces2
-    dichotic1~~dichotic2
-    dichotic1~~0*faces1
-    dichotic1~~0*faces2
-     dichotic2~~0*faces1
-    dichotic2~~0*faces2
-    faces1~~faces2
-    "
-#Model 2: language tasks are correlated
-    model2<-"
-  rhyme1~~rhyme2
-    rhyme1~~dichotic1
-    rhyme1~~dichotic2
-     rhyme1~~0*faces1
-    rhyme1~~0*faces2
-    rhyme2~~dichotic1
-    rhyme2~~dichotic2
-     rhyme2~~0*faces1
-    rhyme2~~0*faces2
-    dichotic1~~dichotic2
-    dichotic1~~0*faces1
-    dichotic1~~0*faces2
-     dichotic2~~0*faces1
-    dichotic2~~0*faces2
-    faces1~~faces2
-    "
-#Model 3:all correlated
-    model3<-"
-    rhyme1~~rhyme2
-    rhyme1~~dichotic1
-    rhyme1~~dichotic2
-     rhyme1~~faces1
-    rhyme1~~faces2
-    rhyme2~~dichotic1
-    rhyme2~~dichotic2
-     rhyme2~~faces1
-    rhyme2~~faces2
-    dichotic1~~dichotic2
-    dichotic1~~faces1
-    dichotic1~~faces2
-     dichotic2~~faces1
-    dichotic2~~faces2
-    faces1~~faces2
-    "
-#Make a little dataframe for AIC weights
-AICdf <- data.frame(matrix(NA,nrow=2,ncol=4))
-AICdf[,1]<-c('Right-handers','Left-handers')
-colnames(AICdf)<-c('Handedness','Model C','Model A','Model B')
-for (myhand in 1:2) {
-  mydat<-right_dat
-  if(myhand==2) {mydat<-left_dat}
-
- #Using data from this handedness type, test all 3 models
-fit1A <- cfa(model1, data=mydat)
-fit2A <- cfa(model2, data=mydat)
-fit3A <- cfa(model3, data=mydat)
-    aic_vector1 <- c(fitMeasures(fit1A, "aic"),fitMeasures(fit2A, "aic"),fitMeasures(fit3A, "aic"))
-
-    #Confirms that for each model, lowest AIC obtained when the simulated data do match the model
-    # 
-    # The AIC weights make massive difference for preferred model!
-    AICweights<-akaike.weights(aic_vector1)$weights
-    # 
- AICdf[myhand,2:4]<-round(AICweights,3)
- }
-myft6 <- flextable(AICdf, col_keys = c("Handedness",
-                                       "Model A",
-                                       "Model B", 
-                                       "Model C"))
-autofit(myft6, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
-```
-
-### Exploratory analyses
-
-#### Finger tapping and Edinburgh Handedness Inventory
-
-```{r fingerEHI,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
-spearCI <- RVAideMemoire::spearman.ci(LIs$Day1_finger_LI, LIs$index_EHI, nrep = 10000, conf.level = 0.95)
-spearCI2 <- RVAideMemoire::spearman.ci(LIs$Day2_finger_LI, LIs$index_EHI, nrep = 10000, conf.level = 0.95)
-# spearman's test
-speartest <- cor.test(LIs$Day1_finger_LI, LIs$index_EHI, method = "spearman", alternative = "greater")
-speartest2 <- cor.test(LIs$Day2_finger_LI, LIs$index_EHI, method = "spearman", alternative = "greater")
-```
-
-To examine the relationship between the finger tapping task and a conventional hand preference battery, we computed Spearman's rank correlation between LIs produced on the finger tapping task and the Edinburgh Handedness Inventory.
-For session 1 data, the relationship between the tasks was moderate, *r* = `r round(speartest$estimate, 2)`, *95% CI* [`r round(spearCI$conf.int[1], 2)`, `r round(spearCI$conf.int[2], 2)`], *p* < 0.001.
-For session 2, the correlation was *r* = `r round(speartest2$estimate, 2)`, *95% CI* [`r round(spearCI2$conf.int[1], 2)`, `r round(spearCI2$conf.int[2], 2)`], *p* < 0.001.
-This suggests that the influence of hand preference on the tapping task may increase with practice.
-
-<!---#### Handedness measurements and their association with laterality --->
-
-```{r hand_explore, echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
-hand_cors <- data.frame(matrix(ncol = 5, nrow = 1))
-  colnames(hand_cors) <- c(" ", "DL 1", "CF 1", "DL 2", "CF 2")
-# add rows
-# row 1
-hand_cors[1, 1] = "EHI"
-hand_cors[1, 2] = round(cor(LIs$Day1_DL_acc_LI, LIs$index_EHI, use= "complete.obs"), 2)
-hand_cors[1, 3] = round(cor(LIs$Day1_CF_acc_LI, LIs$index_EHI, use= "complete.obs"), 2)
-hand_cors[1, 4] = round(cor(LIs$Day2_DL_acc_LI, LIs$index_EHI, use= "complete.obs"), 2)
-hand_cors[1, 5] = round(cor(LIs$Day2_CF_acc_LI, LIs$index_EHI, use= "complete.obs"), 2)
-# row 2
-hand_cors[2, 1] = "FT"
-hand_cors[2, 2] = round(cor(LIs$Day1_DL_acc_LI, LIs$Day1_finger_LI, use= "complete.obs"), 2)
-hand_cors[2, 3] = round(cor(LIs$Day1_CF_acc_LI, LIs$Day1_finger_LI, use= "complete.obs"), 2)
-hand_cors[2, 4] = round(cor(LIs$Day2_DL_acc_LI, LIs$Day2_finger_LI, use= "complete.obs"), 2)
-hand_cors[2, 5] = round(cor(LIs$Day2_CF_acc_LI, LIs$Day2_finger_LI, use= "complete.obs"), 2)
-# table 
-corrTabs <- flextable(
-  head(hand_cors))
-# same scale
-LIs$negEHI <- -LIs$index_EHI
-# compare
-cocor.result <- cocor(~Day1_DL_acc_LI + negEHI | Day1_DL_acc_LI + Day1_finger_LI, LIs)
-cocor.result2 <- cocor(~Day1_CF_acc_LI + negEHI | Day1_CF_acc_LI + Day1_finger_LI, LIs)
-cocor.result3 <- cocor(~Day2_DL_acc_LI + negEHI | Day2_DL_acc_LI + Day2_finger_LI, LIs)
-cocor.result4 <- cocor(~Day2_CF_acc_LI + negEHI | Day2_CF_acc_LI + Day2_finger_LI, LIs)
-# ASSOCIATIONS
-# using self report
-# dichotic
-self_DL <- table(LIs$handedness,LIs$dichotic1.cut)
-self_DL.chi <- chisq.test(self_DL)
-# chimeric
-self_CF <- table(LIs$handedness,LIs$chimeric1.cut)
-self_CF.chi <- chisq.test(self_CF)
-# combinw
-self_hand <- as.data.frame(cbind(self_DL, self_CF))
-  self_hand$` ` <- rownames(self_hand)
-# table
-self_hand_Tabs <- flextable(
-  head(self_hand),
-  col_keys = c(" ", "DL TypLat", "DL NonLat", "DL AtypLat", "CF TypLat", "CF NonLat", "CF AtypLat"))
-# using finger tapping
-LIs$finger1.cut <- cut(LIs$Day1_finger_LI, 
-                       breaks = c(-Inf, 0, Inf), 
-                       labels = c("Right", "Left"), 
-                       right = FALSE)
-# dichotic
-finger_DL <- table(LIs$finger1.cut,LIs$dichotic1.cut)
-finger_DL.chi <- chisq.test(finger_DL)
-# chimeric
-finger_CF <- table(LIs$finger1.cut,LIs$chimeric1.cut)
-finger_CF.chi <- chisq.test(finger_CF)
-# combinw
-finger_hand <- as.data.frame(cbind(finger_DL, finger_CF))
-    finger_hand$` ` <- rownames(finger_hand)
-# table
-finger_hand_Tabs <- flextable(
-  head(finger_hand),
-  col_keys = c(" ", "DL TypLat", "DL NonLat", "DL AtypLat", "CF TypLat", "CF NonLat", "CF AtypLat"))
-# using EHI
-LIs$EHI.cut <- cut(LIs$index_EHI, 
-                       breaks = c(-Inf, -40, 40, Inf), 
-                       labels = c("Left", "Ambidextrous", "Right"), 
-                       right = FALSE)
-# dichotic
-EHI_DL <- table(LIs$EHI.cut,LIs$dichotic1.cut)
-EHI_DL.chi <- chisq.test(EHI_DL)
-# chimeric
-EHI_CF <- table(LIs$EHI.cut,LIs$chimeric1.cut)
-EHI_CF.chi <- chisq.test(EHI_CF)
-# combinw
-EHI_hand <- as.data.frame(cbind(EHI_DL, EHI_CF))
-    EHI_hand$` ` <- rownames(EHI_hand)
-# table
-EHI_hand_Tabs <- flextable(
-  head(EHI_hand),
-  col_keys = c(" ", "DL TypLat", "DL NonLat", "DL AtypLat", "CF TypLat", "CF NonLat", "CF AtypLat"))
-```
-
-<!---Adam I removed the exploratory analyses of different definitions of handedness, as I was worried the paper was getting too long and indigestible.
-These could be put in a supplementary section, but I think they are too much at this point--->
+#autofit(ft)
+```
 
@@
 # make sure we have good cases
-interR <- LIs[LIs$exclude == 0,]
-DL.lex<-RVAideMemoire::spearman.ci(interR$Day1_DL_acc_LI, interR$lexTALE, nrep = 10000, conf.level = 0.95)
-DL.lex.est<-cor.test(interR$Day1_DL_acc_LI, interR$lexTALE, method = "spearman")
-DL.lex1<-RVAideMemoire::spearman.ci(interR$Day2_DL_acc_LI, interR$lexTALE, nrep = 10000, conf.level = 0.95)
-DL.lex.est1<-cor.test(interR$Day2_DL_acc_LI, interR$lexTALE, method = "spearman")
-# footedness
-foot.dat <- interR[interR$footedness != "Don't know",]
-foot.dat$footedness <- as.factor(foot.dat$footedness)
-#  levels(droplevels(foot.dat$footedness))
-DL.ft.t.1 <- t.test(foot.dat$Day1_DL_acc_LI~ foot.dat$footedness)
-DL.ft.t.2 <- t.test(foot.dat$Day2_DL_acc_LI~ foot.dat$footedness)
-CF.ft.t.1 <- t.test(foot.dat$Day1_CF_acc_LI~ foot.dat$footedness)
-CF.ft.t.2 <- t.test(foot.dat$Day2_CF_acc_LI~ foot.dat$footedness)
-# sightedness
-foot.dat <- interR[interR$miles != "Don't know",]
-foot.dat$miles <- as.factor(foot.dat$miles)
-#  levels(droplevels(foot.dat$miles))
-DL.st.t <- t.test(foot.dat$Day1_DL_acc_LI~ foot.dat$miles)
-DL.st.t <- t.test(foot.dat$Day2_DL_acc_LI~ foot.dat$miles)
-CF.st.t <- t.test(foot.dat$Day1_CF_acc_LI~ foot.dat$miles)
-CF.st.t <- t.test(foot.dat$Day2_CF_acc_LI~ foot.dat$miles)
+interR <- LIs_c[LIs_c$exclude == 0,]
+DL.lex<-RVAideMemoire::spearman.ci(interR$Day1_Dich_acc_Z, interR$lexTALE, nrep = 10000, conf.level = 0.95)
+DL.lex.est<-cor.test(interR$Day1_Dich_acc_Z, interR$lexTALE, method = "spearman")
+DL.lex1<-RVAideMemoire::spearman.ci(interR$Day2_Dich_acc_Z, interR$lexTALE, nrep = 10000, conf.level = 0.95)
+DL.lex.est1<-cor.test(interR$Day2_Dich_acc_Z, interR$lexTALE, method = "spearman")
 ```
@@
 However, as the rhyme task was less reliable than other measures, and showed only small asymmetries, we chose to examine the relationship between the LexTALE and the other verbal laterality measure, dichotic listening, instead.
-There was no correlation between the LexTALE score and LIs on the dichotic listening at session 1 (*r*= `r round(DL.lex.est$estimate, 2)`, *95% CI* [`r round(DL.lex$conf.int[1], 2)`, `r round(DL.lex$conf.int[2], 2)`] *p*= `r round(p.adjust(DL.lex.est$p.value[1], n= 2, method= "bonferroni"), 3)`) or session 2 (*r*= `r round(DL.lex.est1$estimate, 2)`, *95% CI* [`r round(DL.lex1$conf.int[1], 2)`, `r round(DL.lex1$conf.int[2], 2)`] *p*= `r round(p.adjust(DL.lex.est1$p.value[1], n= 2, method= "bonferroni"), 3)`).
-
-<!--- footedness etc also removed here ; could be some interest in looking at people with unusual laterality patterns on eye/foot/hand, but here it is too much--->
+For these exploratory analyses we included participants scoring less than 80 on the LexTALE in order to reduce the effect of a restricted range on the correlational analyses.
+There was no correlation between the LexTALE score and LIs on the dichotic listening at session 1, *r*= `r round(DL.lex.est$estimate, 2)`, *95% CI*= [`r round(DL.lex$conf.int[1], 2)`, `r round(DL.lex$conf.int[2], 2)`], *p*= `r round(p.adjust(DL.lex.est$p.value[1], n= 2, method= "bonferroni"), 3)`, or session 2, *r*= `r round(DL.lex.est1$estimate, 2)`, *95% CI*= [`r round(DL.lex1$conf.int[1], 2)`, `r round(DL.lex1$conf.int[2], 2)`], *p*= `r round(p.adjust(DL.lex.est1$p.value[1], n= 2, method= "bonferroni"), 3)`.
 
@@
 Thus, we were able to examine the test-retest reliability of each task.
-We also examined the association between handedness and atypical lateralisation and examined the intercorrelations between LIs on the dichotic listening and chimeric faces tasks.
+We also examined the association between handedness and atypical lateralisation and examined the intercorrelations between LIs on the dichotic listening, rhyme judgement and chimeric faces tasks.
 
 Each behavioural task yielded a hemispheric advantage in the expected direction, but the size of asymmetries and the test-retest reliability of the different tasks varied.
-A relatively small proportion of individuals showed significant asymmetry on the rhyme decision task.
+A relatively small proportion of individuals showed significant asymmetry on the rhyme decision task, despite the population as a whole being significantly left lateralised.
+This pattern of results indicates that while very few participants showed statistically better performance for images in the right visual field (i.e.
+LI_z showed scores above |1.96|), many showed more subtle RVF advantages and had corresponding LI_z scores that were not statistically different from zero at the individual level.
+When comparing LI_z scores to zero at the population there will have been sufficient power to determine whether these small differences were different from zero.
 This task also yielded the poorest test-retest reliability, just falling short of our target minimum cutoff of *r* = .65.
@@
 Our study was well-powered, so the relatively poor task validity and reliability are likely to be due to methodological factors.
+
 One key difference between van der Haegen and Brysbaert’s and our study was that their language-based visual half-field tasks involved overt naming.
-Speech production is highly lateralised (Kosslyn, 1987) and requiring speech production should yield a hemispheric advantage.
+Speech production is highly lateralised (e.g.
+Knecht et al., 2000), and requiring speech production should yield a hemispheric advantage.
 Our task was designed to require covert speech, but an active speech production component might reveal stronger lateralisation.
 This would, however, be logistically difficult to implement online.
-The other main limitation of online testing is that the visual half-field method depends critically on presenting lateralised visual stimuli that project to one hemisphere.
+The speech requirement may not, however, be a key factor, given that recent studies have reported robust right visual-field advantages on a lexical decision task that does not require active speech production (Brederoo et al., 2020; Hausmann et al., 2019; Willemin et al., 2019).
+
+An alternative explanation for the poor validity and reliability of the rhyme task relates to the control viewing conditions.
+The visual half-field method depends critically on presenting lateralised visual stimuli that project to one hemisphere.
 Although we were able to calibrate screen size at the start of the session and to request participants to sit a certain distance from the screen, it is plausible this was not sufficient to achieve adequately lateralised presentation.
 While we can format images to be consistent across displays, we cannot control participants’ behaviour.
-Outside of the lab, participants can freely move their head which will influence the size of images on the retina, and it is difficult to ensure images are shown in foveal or parafoveal vision.
-One avenue that might be worth exploring is the use of images presented in peripheral vision, as it is unlikely that small head movements would result in the image shifting from peripheral vision.
-This increased difficultly in the processing of information in this region may also result in stronger visual field advantages.
+Outside of the lab, participants can freely move their head which will influence the size of images on the retina, and it is difficult to ensure that images are shown exclusively in parafoveal vision.
+If on a subset of images are not equally displaced in parafoveal vision it could mean that processing is more efficient for the image in its corresponding visual field, reducing the validity of the task and adding considerable noise to response time estimates in each visual field.
+Without stict monitoring of eye-movements here it is not possible to control for this.
+In future work we plan to explore the use of images presented in peripheral vision, as it is unlikely that small head movements would result in the image shifting from peripheral vision.
+This would increase difficultly in the processing of information in this region, which may also result in stronger visual field advantages.
 
@@
 From the perspective of the participant, it is a straightforward task of reporting the emotion expressed in a single face and does not require any explicit comparison.
+The finding that good test-retest reliability could be observed with a reduced number of trials will also be welcome for future studies, where there are constraints on testing time.
 
@@
 Hodgson, Hirst, & Hudson, 2016), which is widely used in neuropsychological assessment.
-Future work should also examine how other factors such as the complexity of key combinations and reach through the midline influence the task.
+Future work should also examine how other factors such as the complexity of key combinations and reach across the midline influence the task.
 
@@
 Not only was our study well powered to examine this relationship, our sample included a large number of left-handers who showed increased variability in their LIs.
-Specifically, we found no association between the LI_z indices for rhyme decision, dichotic listening and chimeric faces.
-The lack of association between the two language tasks, rhyme decision and dichotic listening, must be interpreted cautiously, given that the rhyme decision task was unsatisfactory in several respects: it had relatively low test-retest reliability and only weak asymmetry in the sample as a whole.
-It is possible that, as argued by Bishop (2013), there is fractionation between language skills: the dichotic listening task is purely perceptual, whereas the rhyme decision task involves implicit speech production.
+Specifically, we found no association between the LI_z indices for the two language tasks and chimeric faces.
+There was a weak but measureable association between the two language tasks, rhyme decision and dichotic listening, found in both handedness groups.
+This is of particular interest given the different  demands of the two tasks - speech perception using dichotic presentation for the dichotic listening task, vs covert naming of a picture and comparison to the phonology of a written word in the rhyme judgement task.
 The current result suggests there could be interest in replicating this comparison of the two tasks using a visual-half-field language task that has better reliability and validity.
@@
 Given the range of evidence now available, it may be time to abandon hypotheses that predict that different lateralised skills are either wholly interdependent or wholly independent.
-Evidence such as that from Gerrits et al (2020) confirms that lateralisation of different functions is associated far more than expected by chance, yet cases of dissociation are not uncommon.
+Evidence such as that from Gerrits et al.
+(2020) confirms that lateralisation of different functions is associated far more than expected by chance, yet cases of dissociation are not uncommon.
 Research in future might explore which functions are associated, and to what extent, rather than making a simple contrast between models that postulate a general answer to explain all of laterality.
+
+This conclusion is supported by recent work from Brederoo et al.
+(2020), who reported data for 122 participants.
+In their study, participants completed a variety of visual half-field tasks covering a range of domains such as word processing, face processing, and spatial frequency processing.
+At the population level Brederoo et al.
+reported lateralisation in the expected direction for the majority of tasks.
+For left handers and those with right hemisphere dominant speech, there was evidence of a deviation from typical lateralisation.
+Contrary to the current study, Brederoo et al.
+reported partial evidence to suggest that processes recruiting similar resources will lateralise to homologue areas to promote intra-hemispheric proximity to other cortical areas within their processing hierarchies.
+In one analysis they found the lateralisation of faces and word recognition to be correlated but when they looked at lateralisation in those who were right handed versus those with right dominant language processing there was considerably less evidence for causal complementarity.
+However, Brederoo et al.
+reported consistent evidence within the visual domain, where lateralisation of low and high frequency information processing was associated, but independent of   lateralisation on other tasks.
+So it would appear by comparing tasks of a wider variety and considering laterality across domains, we may elucidate the issue of colateralisation and identify how far individual differences in lateralisation depend on task demands and shared processing resources.
 
@@
 The battery of tasks was administered to a large sample of participants with relative ease and enabled us to examine several questions related to lateralisation.
-We find evidence of increased atypical lateralisation in left- relative to right-handers.
-However, we find no indication that the laterality indices on one task are dependent on another and we conclude that, for the tasks studied here, the causes of lateralisation appear to have different origins.
+We find evidence of increased atypical lateralisation in extreme left-handers relative to right-handers.
+However, we find no support for a unitary model of cerebral lateralisation that postulates simple complementarity between left- and right-sided brain lateralisation in individuals.
 
@@
 ## Acknowledgements
-We would like to thank Dr Mike Burt (University of Durham, UK) for providing stimuli for the chimeric faces task.
-We would also like to thank Professor Kenneth Hugdahl (University of Bergen, Norway) for providing stimuli for the dichotic listening task.
+
+First, we would like to thank Mike Burt for providing stimuli for the chimeric faces task and Kenneth Hugdahl for providing stimuli for the dichotic listening task.
+Second, we would like to thank Marco Hirnstein, Marc Brysbaert, and Sanne Brederoo for their helpful comments on the original manuscript of this paper.
+Third, we would like to thank Rogier Kievit for his advice regarding the modeling of covariances.
 Portions of this work were presented at the Online Meeting of the Experimental Psychology Society in July 2020.
@@
 
-Anwyl-Irvine, A. L., Massonnié, J., Flitton, A., Kirkham, N., & Evershed, J. K. (2020). Gorilla in our midst: An online behavioral experiment builder. *Behavior research methods*, *52*, 388-407.
-
-Badzakova-Trajkov, G., Häberling, I. S., Roberts, R. P., & Corballis, M. C. (2010). Cerebral asymmetries: complementary and independent processes. *PloS one*, *5*, e9682.
+Annett, M. (1970). The growth of manual preference and speed. *British Journal of Psychology*, *61*, 545-558. doi: 10.1111/j.2044-8295.1970.tb01274.x
+
+Annett, M. (1998). Handedness and cerebral dominance: The right shift theory. *Journal of Neuropsychiatry and Clinical Neuroscience*, *10*, 459-469.
+
+Annett, M. (2002). *Handedness and Brain Asymmetry; the right shift theory*. Hove, Sussex; Psychology Press
+
+Anwyl-Irvine, A. L., Massonnié, J., Flitton, A., Kirkham, N., & Evershed, J. K. (2020). Gorilla in our midst: An online behavioral experiment builder. *Behavior Research Methods*, *52*, 388-407.
+
+Badzakova-Trajkov, G., Häberling, I. S., Roberts, R. P., & Corballis, M. C. (2010). Cerebral asymmetries: complementary and independent processes. *PloS One*, *5*, e9682.
+
+Beaton, A. A., & Moseley, L. G. (1984). Anxiety and the measurement of handedness. *British Journal of Psychology*, *75*, 275-278.
 
@@
 
-Bethmann, A., Tempelmann, C., De Bleser, R., Scheich, H., & Brechmann, A. (2007). Determining language laterality by fMRI and dichotic listening. *Brain research*, *1133*, 145-157.
+Bethmann, A., Tempelmann, C., De Bleser, R., Scheich, H., & Brechmann, A. (2007). Determining language laterality by fMRI and dichotic listening. *Brain Research*, *1133*, 145-157.
 
@@
 
-Bryden, M. P., Hécaen, H., & DeAgostini, M. (1983). Patterns of cerebral organization. *Brain and language*, *20*, 249-262.
+Bryden, M. P., Hécaen, H., & DeAgostini, M. (1983). Patterns of cerebral organization. *Brain and Language*, *20*, 249-262.
 
@@
 Carey, D. P., & Johnstone, L. T. (2014). Quantifying cerebral asymmetries for language in dextrals and adextrals with random-effects meta analysis. *Frontiers in Psychology*, *5*, 1128.
+
+Dundas, E.
+M., Plaut, D.
+C., & Behrmann, M.
+(2015).
+Variable left-hemisphere language and orthographic lateralization reduces right-hemisphere face lateralization.
+*Journal of Cognitive Neuroscience*, *27*, 913-925.
 
@@
 
+Hausmann, M., Brysbaert, M., van der Haegen, L.
+J., Specht, K., Hirnstein, M., Willemin, J., .
+.
+.
+Mohr, C.
+(2019).
+Language lateralisation measured across linguistic and national boundaries.
+*Cortex*, *111*, 134-147.
+doi:10.1016/j.cortex.2018.10.020
+
 Hellige, J. B. (1993). *Hemispheric asymmetry: what’s right and what’s left*. Perspectives in Cognitive Neuroscience. Cambridge, MA: Harvard Univeristy Press.
@@
 
-Hoaglin, D. C., & Iglewicz, B. (1987). Fine-tuning some resistant rules for outlier labeling. *Journal of the American statistical Association*, *8*, 1147-1149.
-
-Hodgson, J. C., Hirst, R. J., & Hudson, J. M. (2016). Hemispheric speech lateralisation in the developing brain is related to motor praxis ability. *Developmental cognitive neuroscience*, *22*, 9-17.
-
-Hugdahl, K. (2000). Lateralization of cognitive processes in the brain. *Acta psychologica*, *105*, 211-235.
+Hoaglin, D. C., & Iglewicz, B. (1987). Fine-tuning some resistant rules for outlier labeling. *Journal of the American Statistical Association*, *8*, 1147-1149.
+
+Hodgson, J. C., Hirst, R. J., & Hudson, J. M. (2016). Hemispheric speech lateralisation in the developing brain is related to motor praxis ability. *Developmental Cognitive Neuroscience*, *22*, 9-17.
+
+Hodgson, J. C., & Hudson, J. M. (2018). Speech lateralization and motor control. In G. S. Forrester, W. D. Hopkins, K. Hudry, & A. Lindell (Eds). *Progress in Brain Research* (pp. 145-178). Elsevier.
+
+Hugdahl, K. (2000). Lateralization of cognitive processes in the brain. *Acta Psychologica*, *105*, 211-235.
 
@@
 Attention and cognitive control: unfolding the dichotic listening story.
-*Scandinavian journal of psychology*, *50*, 11-22.
+*Scandinavian Journal of Psychology*, *50*, 11-22.
 
@@
 
-Kosslyn, S. M. (1987). Seeing and imagining in the cerebral hemispheres: a computational approach. *Psychological review*, *94*, 148.
+Kosslyn, S. M. (1987). Seeing and imagining in the cerebral hemispheres: a computational approach. *Psychological Review*, *94*, 148.
+
+Knecht, S., Dräger, B., Deppe, M., Bobe, L., Lohmann, H., Flöel, A., Ringelstein, E.
+B., & Henningsen, H.
+(2000).
+Handedness and hemispheric language dominance in healthy humans.
+*Brain*, *123*, 2512-2518.
 
@@
 
-Lemhöfer, K., & Broersma, M. (2012). Introducing LexTALE: A quick and valid lexical test for advanced learners of English. *Behavior research methods*, *44*, 325-343.
-
-Levy, J., Heller, W., Banich, M. T., & Burton, L. A. (1983). Asymmetry of perception in free viewing of chimeric faces. *Brain and cognition*, *2*, 404-419.
+Lemhöfer, K., & Broersma, M. (2012). Introducing LexTALE: A quick and valid lexical test for advanced learners of English. *Behavior Research Methods*, *44*, 325-343.
+
+Levy, J., Heller, W., Banich, M. T., & Burton, L. A. (1983). Asymmetry of perception in free viewing of chimeric faces. *Brain and Cognition*, *2*, 404-419.
 
@@
 
-Miles, W. R. (1929). Ocular dominance demonstrated by unconscious sighting. *Journal of experimental psychology*, *12*, 113.
-
-Mishkin, M., & Forgays, D. G. (1952). Word recognition as a function of retinal locus. *Journal of experimental psychology*, *43*, 43.
+Miles, W. R. (1929). Ocular dominance demonstrated by unconscious sighting. *Journal of Experimental Psychology*, *12*, 113.
+
+Mishkin, M., & Forgays, D. G. (1952). Word recognition as a function of retinal locus. *Journal of Experimental Psychology*, *43*, 43.
+
+Muchinsky, P. (1996). The correction for attenuation. *Educational and Psychological Measurement*, *56*(1), 63-75. doi:https://doi.org/10.1177/0013164496056001004
 
@@
 
-Porac, C., & Coren, S. (1976). The dominant eye. *Psychological bulletin*, *83*, 880.
+Parsons, S., Kruijt, A., & Fox, E.
+(2019).
+Psychological science needs a standard practice of reporting the reliability of cognitive-behavioral measurements.
+*Advances in Methods and Practices in Psychological Science*.
+doi:10.1177/2515245919879695
+
+Porac, C., & Coren, S. (1976). The dominant eye. *Psychological Bulletin*, *83*, 880.
 
@@
 SUBTLEX-UK: A new and improved word frequency database for British English.
-*Quarterly journal of experimental psychology*, *67*, 1176-1190.
+*Quarterly Journal of Experimental Psychology*, *67*, 1176-1190.
 
@@
 Individual differences in the hemispheric specialization of dual route variables.
-*Brain and language*, *67*, 110-133.
+*Brain and Language*, *67*, 110-133.
 
@@
 
+<!--- FOOTNOTESS --->
+[^1]: Hugdahl et al.
+(2009) recommend excluding participants with a general hearing impairment of more than 20 dB for all frequencies tested, and/or an interaural hearing threshold difference of more than 10 dB as this can influence dichotic listening outcomes.
+It is important to note that this exclusion was not possible in the current study as we did not include a hearing test.
+With the current sample size, subtle hearing deficits in one ear will unlikely undermine the results.
+However, it will be important for future studies to incorporate hearing accuracy tests in order to exclude participants with these hearing related difficulties.
+
+```{r dist_plots,echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
+# rename
+RD.dat$handed3 <- as.factor(RD.dat$handed3)
+  levels(RD.dat$handed3) <- c("Extreme LH", 'Moderate LH', "RH")
+CF.dat$handed3 <- as.factor(CF.dat$handed3)
+  levels(CF.dat$handed3) <- c("Extreme LH", 'Moderate LH', "RH")
+DL.dat$handed3 <- as.factor(DL.dat$handed3)
+  levels(DL.dat$handed3) <- c("Extreme LH", 'Moderate LH', "RH")
+FT.dat$handed3 <- as.factor(FT.dat$handed3)
+  levels(FT.dat$handed3) <- c("Extreme LH", 'Moderate LH', "RH")
+# plot all
+plota <- ggplot(RD.dat, aes(x=tlat_RDT_1, fill=handed3, color=handed3)) +
+  geom_histogram(position="identity", alpha= .35) + theme_bw(20) +
+  xlab('Rhyme decision LI_z (session 1)') +
+  theme(legend.position = "top", legend.title = element_blank()) + xlim(-5,5) + ylim(0, 65)
+plotb <- ggplot(RD.dat, aes(x=tlat_RDT_2, fill=handed3, color=handed3)) +
+  geom_histogram(position="identity", alpha= .35) + theme_bw(20) +
+  xlab('Rhyme decision LI_z (session 2)') + 
+  theme(legend.position = "top", legend.title = element_blank()) + xlim(-5,5) + ylim(0, 65)
+plotc <- ggplot(DL.dat, aes(x=Day1_Dich_acc_Z, fill=handed3, color=handed3)) +
+  geom_histogram(position="identity", alpha= .35) + theme_bw(20) +
+  xlab('Dichotic listening LI_z (session 1)') + 
+  theme(legend.position = "top", legend.title = element_blank()) + xlim(-20,20) + ylim(0, 65)
+plotd <- ggplot(DL.dat, aes(x=Day2_Dich_acc_Z, fill=handed3, color=handed3)) +
+  geom_histogram(position="identity", alpha= .35) + theme_bw(20) +
+  xlab('Dichotic listening LI_z (session 2)') + 
+  theme(legend.position = "top", legend.title = element_blank()) + xlim(-20,20) + ylim(0, 65)
+plote <- ggplot(CF.dat, aes(x=Day1_CF_acc_Z, fill=handed3, color=handed3)) +
+  geom_histogram(position="identity", alpha= .35) + theme_bw(20) +
+  xlab('Chimeric faces LI_z (session 1)') + 
+  theme(legend.position = "top", legend.title = element_blank()) + xlim(-20,20) + ylim(0, 65)
+plotf <- ggplot(CF.dat, aes(x=Day2_CF_acc_Z, fill=handed3, color=handed3)) +
+  geom_histogram(position="identity", alpha= .35) + theme_bw(20) +
+  xlab('Chimeric faces LI_z (session 2)') + 
+  theme(legend.position = "top", legend.title = element_blank()) + xlim(-20,20) + ylim(0, 65)
+plotg <- ggplot(FT.dat, aes(x=Day1_Finger_acc_Z, fill=handed3, color=handed3)) +
+  geom_histogram(position="identity", alpha= .35) + theme_bw(20) +
+  xlab('Finger tapping LI_z (session 1)') + 
+  theme(legend.position = "top", legend.title = element_blank()) + xlim(-5,5) + ylim(0, 65)
+ploth <- ggplot(FT.dat, aes(x=Day2_Finger_acc_Z, fill=handed3, color=handed3)) +
+  geom_histogram(position="identity", alpha= .35) + theme_bw(20) +
+  xlab('Finger tapping LI_z (session 2)') + 
+  theme(legend.position = "top", legend.title = element_blank()) + xlim(-5,5) + ylim(0, 65)
+# plot together
+all.plots <- ggpubr::ggarrange(plota, plotb, plotc, plotd, plote, plotf, plotg, ploth, ncol = 2, nrow= 4)
+ggsave("./plots/dist.plots.png", plot=all.plots,dpi = 300, width = 12, height = 24,)
+```
+
 \newpage
-#### Supplemental material
-
-Figure S1 shows scatterplots indicating scores on the left and right sides for each task, with self-reported handedness (left vs right) colour-coded.
-For each task, the mean for each side is shown as vertical and horizontal dotted lines, and equal performance on the two sides corresponds to the sloping red dotted line: thus an impression of whether the task is lateralised can be obtained by considering whether points cluster around the sloping lines symmetrically.
+
+**Assessing the reliability of an online behavioural laterality battery: A pre-registered study.**
+
+**Parker, A. J., Woodhead, Z. V. J., Thompson, P. A., Bishop, D. V. M.**
+
+**Supplementary Materials**
+
+In these Supplementary Materials we include the distribution of laterality indices for each behavioural laterality task across both testing sessions.
+We also include a series of t-tests examining the difference in LI_z scores between extreme left-handers and right-handers.
+
+**LI_z distributions**
+
+The distributions of LI_z scores are shown in Figure S1.
 
 ![Figure S1.
-Left side (L) vs right side (R) scores for sessions 1 and 2 for the behavioural laterality tasks.
-The diagonal dotted line shows the point of equality: points above this are higher for R than L, and points below are higher for L than R.
-Mean values for R and L are shown as horizontal and vertical dotted lines.
-The colour of the points indicates the participants’ handedness (black = right-handed; red = left-handed).
-](all.raw.png)
-
-For the rhyme decision task, only correct responses were used for computing mean log RTs (shown transformed back to the original ms scale in the plot).
-For dichotic listening and chimeric faces, a correct response can be given from either side; it follows that there will be a negative correlation between the two sides reflecting the fact that the number of selections on the left will be inversely related to the number of correct responses on the right.
-For these tasks, it is evident by inspection that most responses for dichotic listening cluster above the dotted line, indicating a right ear advantage, whereas most responses on chimeric faces cluster below the dotted line, indicating a left hemi-face advantage.
-Finally, we show the number of correct sequences with each hand on the finger tapping task.
-Here, a strong impact of handedness is evident on inspection, with left-handers generally more accurate with the left hand, and right-handers more accurate with the right hand.
+Distribution of laterality indices on the behavioural tasks.
+The distributions of LI_z are show in pink for extreme left-handers, green for moderate left-handers, and blue for right-handers.
+Negative values indicates left lateralisation at the brain level and positive values indicate left lateralisation at the brain level.](./plots/dist.plots.png){width=60%}
+
+**LI_z scores in extreme left-handers and right-handers**
+```{r hand.corrs,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
+comp_c <- ggplot(subset(LIs, handed3== "Right-hander" | handed3=="Extreme left-hander"),
+  aes(x= handed3, y= Day1_CF_acc_Z, fill= handed3)) +
+  geom_hline(yintercept=0) + 
+  geom_hline(yintercept=1.96, linetype= "dashed", color= "red") +
+  geom_hline(yintercept=-1.96, linetype= "dashed", color= "red") +
+  geom_violin(alpha= .1, size= .5) +
+  geom_jitter(width= .1, aes(color= handed3), alpha= .25) +
+  geom_boxplot(width=0.1, position=position_dodge(0.9), size= .5 , notch = T, notchwidth = 0.4, varwidth =F,
+               color= 'black', show.legend=FALSE, linetype= 'solid', outlier.color = NA) +
+  ylab("Chimeric faces LI_z") + 
+  xlab(" ") +
+  ggtitle("(C)") + theme(plot.title = element_text(hjust = 0.5)) +
+  ylim(-20, 20) +
+  theme_bw(18) +
+  theme(legend.position= "none") + theme(plot.title = element_text(hjust = 0.5))
+comp_b <- ggplot(subset(LIs, handed3== "Right-hander" | handed3=="Extreme left-hander"),
+  aes(x= handed3, y= Day1_Dich_acc_Z, fill= handed3)) +
+  geom_hline(yintercept=0) + 
+  geom_hline(yintercept=1.96, linetype= "dashed", color= "red") +
+  geom_hline(yintercept=-1.96, linetype= "dashed", color= "red") +
+  geom_violin(alpha= .1, size= .5) +
+  geom_jitter(width= .1, aes(color= handed3), alpha= .25) +
+  geom_boxplot(width=0.1, position=position_dodge(0.9), size= .5 , notch = T, notchwidth = 0.4, varwidth =F,
+               color= 'black', show.legend=FALSE, linetype= 'solid', outlier.color = NA) +
+  ylab("Dichotic listening LI_z") +
+  xlab(" ") +
+  ggtitle("(B)") + theme(plot.title = element_text(hjust = 0.5)) +
+  ylim(-20, 20) +
+  theme_bw(18) +
+  theme(legend.position= "none") + theme(plot.title = element_text(hjust = 0.5))
+comp_a <- ggplot(subset(LIs, handed3== "Right-hander" | handed3=="Extreme left-hander"),
+  aes(x= handed3, y= tlat_RDT_1, fill= handed3)) +
+  geom_hline(yintercept=0) + 
+  geom_hline(yintercept=1.96, linetype= "dashed", color= "red") +
+  geom_hline(yintercept=-1.96, linetype= "dashed", color= "red") +
+  geom_violin(alpha= .1, size= .5) +
+  geom_jitter(width= .1, aes(color= handed3), alpha= .25) +
+  geom_boxplot(width=0.1, position=position_dodge(0.9), size= .5 , notch = T, notchwidth = 0.4, varwidth =F,
+               color= 'black', show.legend=FALSE, linetype= 'solid', outlier.color = NA) +
+  ylab("Rhyme decision LI_z") +
+  xlab(" ") +
+  ggtitle("(A)") + 
+  ylim(-20, 20) +
+  theme_bw(18) +
+  theme(legend.position= "none") + theme(plot.title = element_text(hjust = 0.5))
+t.plots <- ggpubr::ggarrange(comp_a, comp_b, comp_c, ncol = 1, nrow= 3)
+ggsave("./plots/t.plots.png", plot=t.plots,dpi = 300, width = 5, height = 15,)
+# tests
+t1 <- wilcox.test(data= subset(LIs, handed3== "Right-hander" | handed3=="Extreme left-hander"), tlat_RDT_1~ handed3, alternate= "less",conf.int = TRUE, conf.level = 0.95)
+t2 <- wilcox.test(data= subset(LIs, handed3== "Right-hander" | handed3=="Extreme left-hander"), Day1_DL_acc_LI~ handed3, alternate= "less",conf.int = TRUE, conf.level = 0.95)
+t3 <- wilcox.test(data= subset(LIs, handed3== "Right-hander" | handed3=="Extreme left-hander"), Day1_CF_acc_Z~ handed3, alternate= "greater",conf.int = TRUE, conf.level = 0.95)
+```
+
+To further examine the link between handedness and atypical laterality, we conducted a series of one-tailed Mann-Whitney *U*-tests using session 1 data.
+The distribution of LI_z scores are shown separately in Figure S2.
+The Mann-Whitney test for the rhyme decision task indicated that LI_z scores did not differ between extreme left-handers and right-handers, *U*= `r round(t1$statistic, 2)`, *95% CI*= [`r round(t1$conf.int[1], 2)`, `r round(t1$conf.int[2], 2)`], *p*= `r round(t1$p.value, 3)`.
+This is not surprising given that very few participants were lateralised on this task.
+While LI_z scores did not differ between extreme left-handers and right-handers on the rhyme decision, they did on the dichotic listening, *U*= `r round(t2$statistic, 2)`, *95% CI*= [`r round(t2$conf.int[1], 2)`, `r round(t2$conf.int[2], 2)`], *p*= `r round(t2$p.value, 3)`, and chimeric faces, *U*= `r round(t3$statistic, 2)`, *95% CI*= [`r round(t3$conf.int[1], 2)`, `r round(t3$conf.int[2], 2)`], *p*= `r round(t3$p.value, 3)`, tasks.
+These differences demonstrate that, as a population, extreme left-handers were less lateralised.
+
+![Figure S2.
+Violin plots showing the distribution of laterality indices on the behavioural tasks for extreme left-handers (in pink) and right-handers (in blue).
+The notch on the boxplot indicates the mean per condition.
+The red dashed lines are drawn at -1.96 and 1.96, indicating significant lateralisation to either the left or right hemisphere.
+Data points represent indiciaul participants' data.](./plots/t.plots.png){width=40%}

### 2020-11-27 battery-preprocessing_v6_DB_split2.Rmd
revised manuscript

--- 2020-11-18 battery-preprocessing_v6_DB_split2.Rmd
@@
 Modified to allow rerun on split half.
-For this we take blocks of 4 rows, e.g. 1,2,3,4,9,10,11,12 etc.
-This way the script will work for finger tapping as well as the other tasks.
-Idea is to do this selection of rows right at the outset on the raw file.
 
@@
 # libraries
-library(readr)
+
 library(yarrr)
-library(dplyr)
 library(tidyr)
@@
 ```{r demograhics, warning=FALSE}
+
+oddeven <- 0 #set to zero for full dataset; to 1 for odds, and to 2 for evens
+#NB we want to use exclusion criteria from full dataset for the split half as well, and so we need to run the script with oddeven set to zero before using it for split
+
+#Demographics not needed for odd/even runs
+if (oddeven==0){
 #Reading from OSF - this only works if files are public
@@
 # relabel 
-demo_dat$subject <- demo_dat$Participant.Public.ID
-demo_dat$subject <- as.factor(demo_dat$subject)
+demo_dat$origID <- demo_dat$Participant.Public.ID
+demo_dat$origID <- as.factor(demo_dat$origID)
 # count and print subjects
-nsub <-length(levels(demo_dat$subject))
+nsub <-length(levels(demo_dat$origID))
 print(paste0("subjects: ", nsub))
@@
 allsum <- data.frame(matrix(ncol = 7, nrow = nsub))
-  colnames(allsum) <- c("subject", "age", "gender", "footedness", "handedness", "bilingual", "strongest.language")
+  colnames(allsum) <- c("origID", "age", "gender", "footedness", "handedness", "bilingual", "strongest.language")
 myrow <- 0 # start row counter
 for (i in 1:nsub) { # loop through subjects
-  subname <- levels(demo_dat$subject)[i] # find subject subject
-  myrows <- which(demo_dat$subject==subname) # select rows for this subject
+  subname <- levels(demo_dat$origID)[i] # find origID subject
+  myrows <- which(demo_dat$origID==subname) # select rows for this subject
   tmp <- data.frame(demo_dat[myrows,])
@@
 
-    allsum$subject[myrow] <- subname
+    allsum$origID[myrow] <- subname
     allsum$age[myrow] <- as.numeric( tmp[tmp$Question.Key == "Age",]$Response)
@@
 allsum$genderpch[w]<-NA
+}
 ```
@@
 ```{r sightednesss, warning=FALSE}
+if (oddeven==0){
 porta_dat <- read.csv("https://osf.io/zmqkt/download",stringsAsFactors = F)
@@
 
-miles_dat$test_subject <- "miles"
-porta_dat$test_subject <- "porta"
+miles_dat$test_origID <- "miles"
+porta_dat$test_origID <- "porta"
 # merge the two
@@
 # relabel 
-sight_dat$subject <- sight_dat$Participant.Public.ID
-sight_dat$subject <- as.factor(sight_dat$subject)
+sight_dat$origID <- sight_dat$Participant.Public.ID
+sight_dat$origID <- as.factor(sight_dat$origID)
 
 # count and print subjects
-nsub <-length(levels(sight_dat$subject))
+nsub <-length(levels(sight_dat$origID))
 print(paste0("subjects: ", nsub))
@@
 sightedness <- data.frame(matrix(ncol = 3, nrow = nsub))
-  colnames(sightedness) <- c("subject", "miles", "porta")
+  colnames(sightedness) <- c("origID", "miles", "porta")
 myrow <- 0 # start row counter
 for (i in 1:nsub) { # loop through subjects
-  subname <- levels(sight_dat$subject)[i] # find subject subject
-  myrows <- which(sight_dat$subject==subname) # select rows for this subject
+  subname <- levels(sight_dat$origID)[i] # find origID subject
+  myrows <- which(sight_dat$origID==subname) # select rows for this subject
   tmp <- data.frame(sight_dat[myrows,])
@@
 
-    sightedness$subject[myrow] <- subname
+    sightedness$origID[myrow] <- subname
     sightedness$miles[myrow] <- tmp[tmp$Question.Key == "response-2",]$Response
@@
 # now merge sightedness and allsum
-allsum <- merge(allsum, sightedness, by= "subject",all=TRUE)
+allsum <- merge(allsum, sightedness, by= "origID",all=TRUE)
 head(allsum) #just to check all looking OK
+}
 ```
@@
 ```{r EHI, warning=FALSE}
+if (oddeven==0){
 EHI <- read.csv("https://osf.io/kfdma/download",stringsAsFactors = F)
@@
 # recode variables
-EHI$subject <- EHI$Participant.Public.ID
-EHI$subject <- as.factor(EHI$subject)
+EHI$origID <- EHI$Participant.Public.ID
+EHI$origID <- as.factor(EHI$origID)
 
@@
 # new data
-meaningful <- c("subject", "Question Key", "Response") # select wanted columns
+meaningful <- c("origID", "Question Key", "Response") # select wanted columns
 c<-which(names(EHI) %in% meaningful) #find colnumbers of unwanted
@@
 # new data
-meaningful <- c("subject", "right_hand", "left_hand") # select wanted columns
+meaningful <- c("origID", "right_hand", "left_hand") # select wanted columns
 c<-which(names(EHI) %in% meaningful) #find colnumbers of unwanted
@@
 # create hand data 
-nsub <-length(levels(EHI$subject))
+nsub <-length(levels(EHI$origID))
 EHI2 <- data.frame(matrix(ncol = 3, nrow = nsub))
-  colnames(EHI2) <- c("subject", "right_hand", "left_hand")
+  colnames(EHI2) <- c("origID", "right_hand", "left_hand")
 
@@
   myrow <- i
-  subname <- levels(EHI$subject)[i] # find subject subject
-  myrows <- which(EHI$subject==subname) # select rows for this subject
+  subname <- levels(EHI$origID)[i] # find origID subject
+  myrows <- which(EHI$origID==subname) # select rows for this subject
   tmp <- data.frame(EHI[myrows,])
 
-    EHI2$subject[myrow] <- subname # add row for subject
+    EHI2$origID[myrow] <- subname # add row for subject
     EHI2$right_hand[myrow] <- sum(tmp$right_hand) # sum for R hand
@@
 for (i in 1:nrow(allsum)){
-  thissub <- allsum$subject[i]
-  w<-which(EHI2$subject==thissub)
+  thissub <- allsum$origID[i]
+  w<-which(EHI2$origID==thissub)
  allsum$index_EHI[i] <- EHI2$index_EHI[w]
@@
 eplot<-ggplot(allsum, aes(handpch,index_EHI, colour=gender)) +
-  geom_jitter(wsubjectth=.25,height=2)+
+  geom_jitter(width=.25,height=2)+
   ggtitle("Handedness index on the Edinburgh Handedness Inventory")
 eplot
-#This is not perfect, but gives and subjectea to check EHI is valid etc.
+#This is not perfect, but gives and ID to check EHI is valid etc.
 
@@
 allsum$handpch[oddcase] <- 1
+}
 ```
@@
 ```{r lexTALE, warning=FALSE}
+if (oddeven==0){
 # read data
@@
 # rename variables
-lexTALE$subject <- lexTALE$Participant.Public.ID
+lexTALE$origID <- lexTALE$Participant.Public.ID
 lexTALE$accuracy <- lexTALE$Correct
@@
 # code as factors
-lexTALE$subject <- as.factor(lexTALE$subject)
+lexTALE$origID <- as.factor(lexTALE$origID)
 lexTALE$condition <- as.factor(lexTALE$condition)
@@
 lexTale_score <- data.frame(matrix(ncol = 4, nrow = nsub))
-  colnames(lexTale_score) <- c("subject", "Word", "nonWord", "lexTALE")
-
-for (i in 1:nsub) { # loop through subjects
-  subname <- levels(lexTALE$subject)[i] # find subject subject
-  myrows <- which(lexTALE$subject==subname) # select rows for this subject
+  colnames(lexTale_score) <- c("origID", "Word", "nonWord", "lexTALE")
+
+for (i in 1:nsub) { # loop through origIDs
+  subname <- levels(lexTALE$origID)[i] # find origID subject
+  myrows <- which(lexTALE$origID==subname) # select rows for this subject
   tmp <- data.frame(lexTALE[myrows,])
@@
 
-    lexTale_score$subject[myrow] <- subname
+    lexTale_score$origID[myrow] <- subname
     lexTale_score$Word[myrow] <- sum(tmp[tmp$condition == "word",]$accuracy)
@@
 # plot
-x<-ggplot(lexTale_score, aes(x=lexTALE, y=subject, fill=subject)) +
+x<-ggplot(lexTale_score, aes(x=lexTALE, y=origID, fill=origID)) +
   geom_dotplot(binaxis='y', stackdir='center') + xlim(0, 100) +
@@
 # now create a data frame only for index 
-lexTALE <- dplyr::select(lexTale_score, "subject", "lexTALE")
+lexTALE <- dplyr::select(lexTale_score, "origID", "lexTALE")
 # now merge with allsum
-allsum <- merge(allsum, lexTALE, by= "subject",all=TRUE)
+allsum <- merge(allsum, lexTALE, by= "origID",all=TRUE)
 # now write the full allsum 
@@
 allsum <- allsum %>% mutate(lexOut= if_else(lexTALE < 80, "remove", "keep"))
+}
 ```
@@
 ```{r readraw,include= FALSE, warning=FALSE}
-# Set directory
-
+  #reading direct from OSF is temperamental
+  #Often gives error message if run from chunk, but line by line is OK. Grrr.
+  readosf<-0
+  if(readosf==1){
   data_file <- "https://osf.io/mhk29/download" #day 1 data
@@
      all_dat2 <- read.csv(data_file, na.strings = "NA",stringsAsFactors=F)
+  }
+  if(readosf==0){
+    all_dat1=read.csv("./data/day1_combined.csv",na.strings = "NA",stringsAsFactors=F)
+      all_dat2=read.csv("./data/day2_combined.csv",na.strings = "NA",stringsAsFactors=F)
+  }
 
@@
 alldat %>% filter(Attempt >= 1) # only answers
+
 # write as factors
-alldat$subject <- as.factor(alldat$Participant.Public.ID)
+alldat$origID <- as.factor(alldat$Participant.Public.ID)
+alldat$subject<-as.numeric(alldat$origID)
 alldat$RT <- as.numeric(as.character(alldat$Reaction.Time))
@@
 alldat$word<-substring(alldat$word,66,(nchar(alldat$word)-4)) #remove format from written word stimuli
-oddeven <- 1 #set to zero for full dataset; to 1 for odds, and to 2 for evens
+
+##################################################################################
+#      MODIFICATION FOR ODD AND EVEN ANALYSIS
+##################################################################################
+
 nr<-nrow(alldat)
@@
 }
-
-alldat$word_OG<-NA #create dummy column just to make it easier to align columns
-
-```
-
-Create a simple numerical ID to make life easier when screening data if we need to identify subjects whose data is odd in some way. Ensure this is consistent across all_dat1 and all_dat2 and allsum.
-We first rename the current subject variable so we can reuse this name with the numbered ID for all future computations.
-Again, this is rather slow because done in a loop, which is probably not very efficient!
-
-```{r makesubject}
-#Use allsum as the basis for the check
-#First rename all subject to origID
-allsum <- allsum %>% rename(origID = subject) 
-alldat <- alldat %>% rename(origID = subject)
-
-alldat$subject <- NA
-
-for (n in 1:nrow(allsum)){
-  allsum$subject[n] <- n
-  w<-which(alldat$origID == allsum$origID[n])
-  alldat$subject[w] <-n
-
+```
+
+```{r reuse.allsum}
+#for odd and even subsets, read in original allsum and select demographic and exclusion cols to keep
+if (oddeven>0){
+  allsum<-read.csv("./data/allsum_out.csv",stringsAsFactors=F)
+  keepcols<-c("subject" , "origID","age" ,"gender","footedness" ,"handedness","bilingual", "strongest.language","handpch","genderpch" ,"miles", "porta","index_EHI" ,"lexTALE", "lexOut","exclude","exRDT"      )
+  allsum<-allsum[,keepcols]
 }
 head(allsum)
+
+```
+```{r subnums}
+##################################################################################
+#find subject number corresponding to orig ID
+
+allsum$subject<-NA
+for (i in 1:nrow(allsum)){
+  thisID<-allsum$origID[i]
+  w<-which(alldat$origID == thisID)
+  allsum$subject[i]<-alldat$subject[w[1]]
+}
+
 ```
@@
 ```{r excludelowCFacc}
+if (oddeven==0) {
 #db modified to mark rather than drop cases with scores < .75
@@
 allsum$exclude[w]<-2 #code exclude = 2 if excluded because no response one side
+}
 ```
@@
 w<-which(allsum$exclude==0)
+
 inclsubs <- allsum$subject[w]
@@
 ```{r dichotic.outliers}
+if (oddeven==0){
 dichotic_dat1 <- detect.outliers(dichotic_dat1,200,1.65) #use generic outlier removal function
@@
 #print(outliertable)
+}
 ```
@@
 
+if(oddeven==0){
 #ww contains people who were excluded on both dichotic and faces. 
@@
 t
+}
 ```
@@
 ```{r marktecherr}
+Nfortech<-80 #need at least this N trials
+if(oddeven>0){Nfortech<-40} #half the number for split half runs
 techerr<-table(dichotic_dat2$subject)
-te<-which(techerr<80)
+te<-which(techerr<Nfortech)
 techerr<-techerr[te] #subject numbers for those with few trials.
 techerr<-as.numeric(names(techerr))
+w<-which(allsum$subject%in% techerr)
+allsum$exclude[w]<-4 #exclude for tech error
+
 ```
@@
 includesubs <- which(allsum$exclude<2) #only plot/t-test included cases
+#only plot/t-test included cases
 origdat <- DL_dat1[DL_dat1$subject %in% includesubs,]
@@
 ```{r rdt_outliers}
+
 #First we create a RT column that is based on LOGRT
@@
 RDT_dat2$RT <- log(RDT_dat2$rawRT)
+if(oddeven==0){
 RDT_dat1<-detect.outliers(RDT_dat1,log(200),1.65) #numbers specify min RT and 
@@
 outliertable2<-table(RDT_dat2$subject,RDT_dat2$outlier)
+}
+#for splithalves read in the original RDT file for all trials and then split
+#This means we keep the outlier designation as before
+if(oddeven>0){
+  tempRDT1<-read.csv("./data/myRDTdat1.csv",stringsAsFactors=F)
+  tempRDT2<-read.csv("./data/myRDTdat2.csv",stringsAsFactors=F)
+  if(oddeven==1){
+  RDT_dat1<-tempRDT1[tempRDT1$Trial%%2==1,] #odd trials only
+  RDT_dat2<-tempRDT2[tempRDT2$Trial%%2==1,] #odd trials only
+  }
+    if(oddeven==2){
+  RDT_dat1<-tempRDT1[tempRDT1$Trial%%2==0,] #even trials only
+  RDT_dat2<-tempRDT2[tempRDT2$Trial%%2==0,] #even trials only
+  }
+}
 ```
@@
 ```{r RDT_exclude}
-
+if (oddeven==0){
 w1x<-which(is.na(RDT_df1$RT))
@@
 
-w1<-which(nr1$p.corr==0)
-w2<-which(nr2$p.corr==0)
+w1<-which(nr1$p.corr<75)
+w2<-which(nr2$p.corr<75)
 
@@
 
-w1a<-which(wr1$p.corr<75)
-w2a<-which(wr2$p.corr<75)
+w1a<-which(wr1$p.corr<50)
+w2a<-which(wr2$p.corr<50)
 
@@
 allsum$exRDT[allsum$subject %in% exRDTx]<-2 #missing data
+}
+
 w<-which(allsum$exRDT==0)
+
 incRDT <- allsum$subject[w]
@@
 
-```
-```{r oldRDT_exclude}
-#Original version now superseded
-doold<-0
-if(doold==1){
-  wn <-unique(which(RDT_df1$accurate ==0),which(RDT_df2$accurate ==0)) #this catches those who were always wrong on 'no rhyme' trials, and possibly others where presentation messed up for technical reasons
-
-w1<-which(RDT_df1$p.corr < 75)
-w2<-which(RDT_df2$p.corr < 75)
-
-exRDT<-sort(unique(c(RDT_df1$subject[w1],RDT_df2$subject[w2],RDT_df2$subject[wn])))
-
-allsum$exRDT <- 0
-allsum$exRDT[allsum$subject %in% exRDT]<-1 #decided best to have separate column for exclusion for each task
-w<-which(allsum$exRDT==0)  
-incRDT <- allsum$subject[w]
-ninc <- length(incRDT)
-print(paste0('RDT: number remaining after exclusions = ',ninc))
-
-}
 ```
@@
 measure <- 'RT'
+#need to exclude those who don't have RT values on both L and R
+w<-which(is.na(origdat$X1))
+if (length(w)>0){
+  takeout<-unique(origdat$subject[w])
+  ww<-which(origdat$subject %in% takeout)
+  origdat<-origdat[-ww,]
+}
 dopirate(origdat,task,measure)
@@
 }
-write.csv(allsum, paste0("LIs",oddeven,".csv"))
-```
-
+write.csv(allsum, paste0("./data/LIs",oddeven,".csv"))
 ```
@@
 Three repetitions of all combinations would lead to 108 trials (90 with different emotions in each hemifield, 18 same-same trials).
+
+```{r finalbit}
+if(oddeven==0){
+
+  # CF
+day2.CF <- data.frame(matrix(nrow=length(complete.day2.CF), ncol= 2))
+  colnames(day2.CF) <- c("task", "subject")
+  day2.CF$subject <- complete.day2.CF
+  day2.CF$task <- "CF"
+# DL
+day2.DL <- data.frame(matrix(nrow=length(complete.day2.DL), ncol= 2))
+  colnames(day2.DL) <- c("task", "subject")
+  day2.DL$subject <- complete.day2.DL
+  day2.DL$task <- "DL"
+# RD
+day2.RD <- data.frame(matrix(nrow=length(complete.day2.RD), ncol= 2))
+  colnames(day2.RD) <- c("task", "subject")
+  day2.RD$subject <- complete.day2.RD
+  day2.RD$task <- "CF"
+# Finger
+day2.FT <- data.frame(matrix(nrow=length(complete.day2.FT), ncol= 2))
+  colnames(day2.FT) <- c("task", "subject")
+  day2.FT$subject <- complete.day2.FT
+  day2.FT$task <- "CF"
+# now create a combined df
+complete.day2a <- merge(day2.CF, day2.DL, by= "subject")
+complete.day2b <- merge(day2.RD, day2.FT, by= "subject")
+complete.day2 <- merge(complete.day2a, complete.day2b, by= "subject")
+# add row to code
+complete.day2$day2.complete <- 1
+complete.day2 <- dplyr::select(complete.day2, -2, -3, -4, -5)
+# now merge into allsum
+allsum <- merge(allsum, complete.day2, by= "subject", all= TRUE)
+
+  allsumex<-allsum #make a copy that has the exclusions from the whole dataset
+  write.csv(allsum,'./data/allsum_out.csv',row.names=F)
+}
+LIname<-paste0('./data/LIs',oddeven,'.csv')
+LIcols<-c("subject","handpch", "Day1_CF_acc_LI", "Day2_CF_acc_LI",
+          "Day1_DL_acc_LI" ,"Day2_DL_acc_LI" ,
+          "Day1.RDT_RT_LI" ,"Day2.RDT_RT_LI")
+
+write.csv(allsum[,LIcols],LIname,row.names=F)
+
+zname<-paste0('./data/LIzs',oddeven,'.csv')
+LIzcols<-c("subject","handpch", "Day1_CF_acc_Z", "Day2_CF_acc_Z",
+          "Day1_Dich_acc_Z" ,"Day2_Dich_acc_Z" ,
+          "zlat_RDT_1" ,"zlat_RDT_2")
+
+write.csv(allsum[,LIzcols],zname,row.names=F)
+
+#we need to preserve the RDT files that have the outliers marked as NA
+if(oddeven==0){
+  write.csv(RDT_dat1,'./data/myRDTdat1.csv',row.names=F)
+  write.csv(RDT_dat2,'./data/myRDTdat2.csv',row.names=F)
+}
+
+```
 

### 2020-11-27 online_laterality_reliability.Rmd
revised manuscript

@@
 Three of the tasks showed evidence of both validity and reliability, insofar as they showed hemispheric advantages in the expected direction and test-retest reliability of at least *r*= .75.
-The reliability of the rhyme decision task was less satisfactory (*r*= .63).
-We also found that extreme left-handers were more likely to depart from typical lateralisation.
+The reliability of the rhyme decision task was less satisfactory (*r*= .62).
+We also confirmed a prediction that extreme left-handers were more likely to depart from typical lateralisation.
 Lateralisation across the two language tasks (dichotic listening and rhyme judgement) was weakly correlated, but unrelated to lateralisation on the chimeric faces task.
 We conclude that three of the tasks, dichotic listening, chimeric faces and finger tapping, show considerable promise for online evaluation of cerebral lateralisation.
-The results do not support the notion of a common underlying cause of lateralisation for language and face processing.
+The results supported a common language factor that influenced language tasks but found no evidence for a common underlying cause of lateralisation for language and face processing.
 
@@
 Results from visual half-field and dichotic listening paradigms support the view that the left hemisphere plays the dominant role in language processing (Hugdahl, 2000; Ocklenburg & Güntürkün, 2018).
-By contrast, visual half-field paradigms indicate that the right hemisphere is involved in the processing of faces, global feature processing, and visuo-spatial attention (Brederoo, Nieuwenstein, Cornelissen, & Lorist, 2019).
+By contrast, visual half-field paradigms indicate that the right hemisphere is involved in the processing of faces, global feature processing, and visuospatial attention (Brederoo, Nieuwenstein, Cornelissen, & Lorist, 2019).
 Below, we review how each paradigm has been implemented to study language lateralisation and the detection of emotion in faces, and what is known about reliability and validity.
@@
 This method has been used since the 1950s (Mishkin & Forgays, 1952).
-Bourne (2006) reviewed the literature, which consistently shows advantageous processing for letter strings presented in the RVF, relative to the LVF, in most people.
+Bourne (2006) reviewed the literature, which consistently shows superior processing for letter strings presented in the RVF, relative to the LVF, in most people.
 
@@
 
-In recent years it has become clear that the reliability and validity of visual half-field tasks to identify the language-dominant hemisphere is largely dependent on several methodological aspects.
+In recent years it has become clear that the reliability and validity of visual half-field tasks to identify the language-dominant hemisphere is influenced by several methodological aspects.
 For example, Hunter and Brysbaert (2008) noted stronger and more stable visual field differences when stimuli were presented bilaterally relative to those where only one stimulus was presented either in the LVF or in the RVF (Boles, 1987, 1990, 1994; see also Iacoboni & Zaidel, 1996).
-Hunter and Brysbaert argued that this is the result of stimuli competing during bilateral presentation, enhancing the advantage to the stimulus presented in the language dominant hemisphere.
+Hunter and Brysbaert argued that this competition between stimuli during bilateral presentation enhances the advantage to the stimulus presented in the language dominant hemisphere.
 Reliability and validity also depend on there being a sufficient number of trials and brief stimulus presentation to prevent eye movements.
-When accounting for these methodological aspects, the validity of visual half-field tasks seems promising.
+When these methodological aspects are addressed, the validity of visual half-field tasks seems promising.
 For example, Hunter and Brysbaert (2008) used two visual half-field tasks, word naming (192 trials) and picture naming (160 trials) in which stimuli were bilaterally presented and masked at offset.
-The two tasks correlated strongly (*r*= .8, 95% CI .598-.956) in a sample of 26 left-handers.
+The two tasks correlated strongly (*r*= .80, *95% CI*= .598-.956) in a sample of 26 left-handers.
 Encouragingly, the LIs from these visual half-field tasks also correlated positively with laterality from functional Magnetic Resonance Imaging (fMRI) on a word generation task in 10 left-handers.
-However, the small sample size meant that the confidence intervals around these estimates were very wide: word naming, *r*= .63 (95% CI = .001-.902), picture naming, *r*= .77 (95% CI = .272-.943).
-More recently, however, Van der Haegen, Cai, Seurinck, and Brysbaert (2011) reported highly similar findings in a somewhat larger sample of 50 participants.
+However, the small sample size meant that the confidence intervals around these estimates were very wide: word naming, *r*= .63 (*95% CI*= .001-.902), picture naming, *r*= .77 (*95% CI*= .272-.943).
+More recently, however, Van der Haegen, Cai, Seurinck, and Brysbaert (2011) reported similar findings in a sample of 50 participants.
 Again, participants completed behavioural word and picture naming visual half-field tasks and then a word generation task under fMRI.
@@
 also report a set of analyses involving 250 participants who completed the behavioural visual half-field tasks while their eye movements were monitored.
-Two sets of LIs were calculated; one controlling for fixation error and one that did not.
-Importantly, the LIs which did and did not control for fixation error were highly correlated for word naming, *r* = .94, and picture naming,*r* =.98, indicating that the inclusion of eye movement control adds little additional value when a visual half-field task employs mechanisms to encoruage central fixation and bilateral presentation.
-
-A promising line of work has emerged which further highlights the potential for well designed visual half-field tasks to assess language lateralisation.
+Two sets of LIs were calculated; one that controlled for fixation error and one that did not.
+Importantly, the LIs which did and did not control for fixation error were highly correlated with word naming, *r* = .94, and picture naming,*r* =.98, indicating that the inclusion of eye movement control adds little additional value when a visual half-field task employs mechanisms to encourage central fixation and bilateral presentation.
+
+A promising line of work has emerged which further highlights the potential for well-designed visual half-field tasks to assess language lateralisation.
 First introduced by Willemin et al.
-(2016), the translingual lexical decision task involves participants viewing letter stings either side of fixation and judging whether the left, right, or neither of the letter strings was a valid word.
+(2016), the translingual lexical decision task involves participants viewing letter stings either side of fixation and judging whether the left, right, or neither of the letter strings is a valid word.
 With 100 participants completing 256 trials, Willemin et al.
 reported a RVF advantage at the population level.
-Since then this effect has been replicated in a sample of 496 participants covering six international languages, with evidence of a RVF for each language studied.
-Not only does this task show validity, it seems reliable given Brederoo et al.'s (2020) observation that the task yields good split-half reliability when tested with 122 participants, *r* = 0.86 for error rates and *r* = 0.80 for response times.
+Since then this effect has been replicated in a sample of 496 participants covering six international languages, with evidence of a RVF for each language studied (Hausmann et al., 2019).
+Not only does this task show validity, it seems reliable given Brederoo et al.'s (2020) observation that the task yields good split-half reliability when tested with 122 participants: *r* = 0.86 for error rates and *r* = 0.80 for response times.
 
@@
 
-Simultaneous presentation of different auditory stimuli to left and right ears for immediate recall was a method originally used in studies of attention, but it became evident that there was usually an advantage for verbal stimuli presented to the right ear (Kimura, 1967).
+Simultaneous presentation of different auditory stimuli to left and right ears for immediate recall was a method originally used in studies of attention, but it became evident that there was usually better recall of verbal stimuli presented to the right ear (Kimura, 1967).
 Both ears project to both hemispheres, but the weaker ipsilateral pathways appear to be suppressed when the two sides compete, and so the right ear, with a contralateral projection to the left hemisphere, will have an advantage in tasks that involve language processing.
@@
 First, based on a review of the published literature, Westerhausen and Samuelsen outlined eight design features that should be used in an optimal dichotic listening paradigm and explained their justification for each.
-Their recommendations were: (1) stop consonant-vowel (CV) syllables as stimulus materials; (2) CV stimulus pairs from the same voicing category; (3) alternating trials of voiced and unvoiced stimulus pairs; (4) free-recall instruction; (5) one stimulus pair per trial; (6) a paradigm length of 120 dichotic trials; (7) all stimulus pairs presented at equal frequency to the two sides; and (8) inclusion of binaural (homonymic) trials.
+Their recommendations were: (1) stop consonant-vowel (CV) syllables as stimulus materials; (2) CV stimulus pairs from the same voicing category; (3) alternating trials of voiced and unvoiced stimulus pairs; (4) free-recall instruction; (5) one stimulus pair per trial; (6) a paradigm length of 120 dichotic trials; (7) all stimulus pairs presented at an equal frequency to the two sides; and (8) inclusion of binaural (homonymic) trials.
 Putting these criteria into practice, Westerhausen and Samuelsen administered a modified dichotic listening paradigm using both a verbal and manual response format to 50 right-handers.
@@
 (2013) developed a mobile device version of the consonant-vowel dichotic listening paradigm, with 36 simultaneous sound pairs.
-Under laboratory conditions, with a sample of 76 (33 Norwegian, 43 Austrialian) participants, they reported a test-retest reliability of *r*= 0.78 when using the mobile app.
+Under laboratory conditions, with a sample of 76 (33 Norwegian, 43 Australian) participants, they reported test-retest reliability of *r*= 0.78 when using the mobile app.
 Furthermore, both subgroups of participants showed clear right ear advantages at the population level.
@@
 Again Bless et al.
-(2015) reported a clear right ear advanatge at the population-level.
-They then used the large dataset to examine cultural/lingusitic effects on the degree on lateralisation; reporting that the degree of lateralization appears to be modulated by linguistic background.
+(2015) reported a clear right ear advantage at the population-level.
+They then used the large dataset to examine cultural/linguistic effects on the degree on lateralisation; reporting that the degree of lateralization appears to be modulated by linguistic background.
 Together these studies demonstrate the potential for online laterality research to generate large amounts of data while maintaining a tolerable degree of reliability.
@@
 We may note that Willemin et al.
-asked participants to respond via manual response, whereas other studies have involved an active speech production component; manual responding could possibly interact with handedness to influence the laterality index on this task.
+asked participants to respond via manual response, whereas other studies have involved an active speech production component; manual responding could interact with handedness to influence the laterality index on this task.
 
@@
 left hemisphere) advantage that is noted at the population level (see Hugdahl et al., 2009, and Westerhausen, 2019, for a review) is reduced in left-handers (Bethmann, Tempelmann, De Bleser, Scheich, & Brechmann, 2007; Bryden, 1970, 1988; Hugdahl, 2003).
-More recently, Karlsson, Johnstone, and Carey (2019) employed a consonant-vowel version of the dichotic listening task where participants reported which one of six syllables that they heard most clearly.
+Recently, Karlsson, Johnstone, and Carey (2019) employed a consonant-vowel version of the dichotic listening task where participants reported which one of six syllables that they heard most clearly.
 In study 1, 87% of right-handers and 78% of non-right-handers showed a right ear advantage.
@@
 When viewing a single face with a different emotion shown in either hemiface, 75% of right-handers and 60% of non-right-handers showed a leftwards bias.
-For the second chimeric faces task that involved two chimeric faces arranged vertically, a similar pattern emerged.
-73% of right-handers and 57% of non-right-handers showed a leftwards bias.
+For the second chimeric faces task that involved two chimeric faces arranged vertically, a similar pattern emerged: 73% of right-handers and 57% of non-right-handers showed a leftwards bias.
 
@@
 
-To date the focus has been in testing whether lateralisation shows  complementarity across tasks that are usually mediated by left and right hemispheres, especially whether there is a relationship between language vs visuospatial functioning.
+To date, the focus has been in testing whether lateralisation shows  complementarity across tasks that are usually mediated by the left and right hemispheres, especially whether there is a relationship between language vs visuospatial functioning.
 However, emerging lines of evidence have suggested that even within domains there may be dissociated patterns of lateralisation across similar tasks.
@@
 The first assumed that language laterality is a unidimensional trait and a single factor was used to model covariance.
-The second assumed that that two laterality factors underlie the pattern of covariance.
-The results supported a two factor model at the population level.
+The second assumed that two laterality factors underlie the pattern of covariance.
+The results supported a two-factor model at the population level.
 However, further examination of the data revealed that this was driven by seven left-handers.
-Woodhead, Thompson, Karlsson, and Bishop (2020) examined this factor structure in an extended sample of 74 participants with 43 right-handers and 31 left-handers.
-The data this time indicated that a two factor model was preferred for both right- and left-handers, with the preference for a two factor model being stronger for left-handers.
-Together, this indicated that even with the domain of language laterality indices may be dissociated.
+Furthermore, Woodhead, Thompson, Karlsson, and Bishop (2020) examined this factor structure in an extended sample of 74 participants (43 right-handers and 31 left-handers).
+The data this time indicated that a two-factor model was preferred for both right- and left-handers, with the preference for a two-factor model being stronger for left-handers.
+Together these data suggest that  laterality indices within the domain of language may be dissociated.
 
@@
 how feasible would it be to administer the task remotely via an online platform.
-We aimed to include at least two language measures, one tapping into receptive language and one expressive, in order to test the notion that language laterality can dissociate across different tasks.
+We aimed to include at least two language measures, one tapping into receptive language and one expressive, to test the notion that language laterality can dissociate across different tasks (i.e.
+Woodhead et al., 2019, 2020).
 We also aimed to include a task that involved face processing so that we could test for complementarity of lateralisation across tasks that usually lateralise to left and right hemispheres.
@@
 
-Given that numerous studies looking at the relationship lateralisation on language and non-language tasks have focused on the use of face stimuli (e.g.
+Given that numerous studies looking at the relationship between lateralisation on language and non-language tasks have focused on the use of face stimuli (e.g.
 Brederoo et al., 2020), we chose to include a face task to conduct similar comparisons.
-Regarding the chimeric faces, Levy et al.
+Regarding the chimeric faces task, Levy et al.
 (1983) reported good split-half reliability while Karlsson et al.
-had reported that, like the dichotic listening task, the the chimeric faces task is sensitive to the differences in perceptual asymmetry in right- and non-right-handers.
+had reported that, like the dichotic listening task, the chimeric faces task is sensitive to the differences in perceptual asymmetry in right- and non-right-handers.
 
@@
 The pegboard task is one of the most widely used performance tasks (see Hodgson & Hudson, 2016, for discussion), but the requirement for apparatus makes it unfeasible as an online test.
-We therefore took the opportunity to explore the reliability and validity of a finger-tapping task that required the participant to repeatedly tap a short sequence on the computer keyboard with the index finger of the preferred or nonpreferred hand, which capture one aspect of pegmoving, namely precise aiming of sequential movements.
+We therefore took the opportunity to explore the reliability and validity of a finger-tapping task that required the participant to repeatedly tap a short sequence on the computer keyboard with the index finger of the preferred or non-preferred hand, which captures one aspect of peg-moving, namely precise aiming of sequential movements.
 
@@
 However, their study was restricted to left-handers, who may be atypical in this regard, and they noted that lack of association was hard to interpret because it may be due to unreliability of the individual measures.
-In the pre-registration for our current study, we predicted that for right-handers, a model where the two language tasks are correlated with each other but not with the chimeric faces (model A) would be the best fit to the data, compared to either a model in which laterality indices are correlated between the language and chimeric face tasks (model B), or a base model where each task is independently lateralised (model C).
-Given our previous findings using fTCD to study dissociations in language laterality (Woodhead, Bradshaw, Wilson, Thompson, & Bishop, 2019), we further predicted that model fit may differ for right- and left-handers, with more dissociation between tasks in the latter group.
+In the pre-registration for our current study, we predicted that for right-handers, a two-factor model where the two language tasks are correlated with each other but not with the chimeric faces (model 2F) would be the best fit to the data, compared to either a single factor model in which laterality indices are correlated between the language and chimeric face tasks (model 1F), or a base model where each task is independently lateralised (model 3F).
+Given our previous findings using fTCD to study dissociations in language laterality (Woodhead et al., 2019, 2020), we further predicted that model fit may differ for right- and left-handers, with more dissociation between tasks in the latter group.
+While we report this analysis in the Supplementary Materials, we report an analysis of the whole sample in the main body as we found that the results from subgroup analyses were very unstable with the small sample of left-handers who were eligible for analysis, and could favour different models when quite minor changes were made to the sample.
 
@@
 In practice, our final sample exceeded 300, because we discarded participants from individual tasks for failing to meet prespecified criteria (see below).
-In total, 441 native English speakers were initially recruited through Prolific Academic (www.prolific.co) with all these participants completing both sessions,  Prolific is a platform for recruiting and screening paid research participants according to experimenter-specified criteria.
+
+In total, 441 native English speakers were initially recruited through Prolific Academic (www.prolific.co) with all these participants completing both sessions.
+Prolific is a platform for recruiting and screening paid research participants according to experimenter-specified criteria.
 Participants were 18- to 50-years-old, had normal or corrected-to-normal vision and reported no history of neurological disease (e.g.
@@
 For both questions, participants could answer "left", "right", or "no preference / Don't know".
+No specific predictions were made about footedness, but it was included to provide a comprehensive picture of behavioural laterality.
 
@@
 Each test classified participants as being either left or right eye dominant.
-We did not have specific predictions about ocular dominance, but gathered this information to further characterise behavioural laterality of the sample.
+We did not have specific predictions about ocular dominance but gathered this information to further characterise behavioural laterality of the sample.
 In practice, the two tests agreed perfectly, so we report here just the Miles test.
@@
 Hunter & Brysbaert, 2008) to make the task suitable for online presentation.
-This obviated the need for measuring latencies of oral responses, which would be challenging in the online context, both in terms of need for a high-quality microphone, and possible intrusion of background noises interfering with recording.
-Furthermore, the evaluation of oral responses would require additional scoring following data collection which would be a taxing procedure for researchers.
-It would also require additional ethical approvals with regards to the storing and processing of the data.
+This obviated the need for measuring latencies of oral responses, which would be challenging in the online context, both in terms of the need for a high-quality microphone, and possible intrusion of background noises interfering with recording.
+Furthermore, the evaluation of oral responses would require additional scoring following data collection which would be a taxing procedure for researchers and could introduce subjectivity into the measure.
+Voice recording would also require additional ethical approvals with regards to the storing and processing of the data.
 
@@
 #We will replace with read from OSF, but better stick with LIs to ensure final version is the one we use 
-allsum <- read.csv('./data/allsum.csv',stringsAsFactors = F)
-
+allsum <- read.csv('./data/allsum_out.csv',stringsAsFactors = F, na.strings = "NA")
+allsum <- allsum[!is.na(allsum$origID), ]
 ```
@@
 # reformat NA
+#identify missing cases for time 2
+day2cols<-c('Day2_CF_acc_LI','Day2_DL_acc_LI','Day2.RDT_RT_LI','Day2_finger_LI')
 allsum$day2.complete <- ifelse(is.na(allsum$day2.complete), 0, allsum$day2.complete)
@@
 Response times were log-transformed to reduce skew, as specified in our pre-registration.
-Outlier response times were identified using Hoaglin and Iglewicz’s (1987) procedure, whereby outliers are defined as response times that are 1.65 times the difference between the first and third quantiles (1.65 * (Q3-Q1)) below or above the first and third quantile values (e.g: lower limit = Q1 – 1.65 x (Q3-Q1); upper limit = Q3 + 1.65 x (Q3-Q1)).
+Outlier response times were identified using Hoaglin and Iglewicz’s (1987) procedure, whereby outliers are defined as response times that are 1.65 times the difference between the first and third quartiles (1.65 * (Q3-Q1)) below or above the first and third quartile values (e.g: lower limit = Q1 – 1.65 x (Q3-Q1); upper limit = Q3 + 1.65 x (Q3-Q1)).
 This criterion was applied to all responses, before subdivision into left- and right-sided responses.
@@
 For the chimeric faces task, we removed participants if they scored lower than 75% on trials in which an identical emotion was shown in each hemiface, and for the dichotic listening task, we removed those who scored less than 75% correct on trials where the same sound was played to each ear.
-These data trimming procedures left data for `r nrow(LIs)-nrow(LIs[LIs$exclude== 1 | LIs$exclude== 12,])` participants on the chimeric faces and `r nrow(LIs)-nrow(LIs[LIs$exclude== 2 | LIs$exclude== 12,])` participants on the dichotic listening task.
-An additional 7 cases had to be removed because of a technical error that meant they had less than 80 trials presented.
+These data trimming procedures left data for 304 participants on the dichotic listening task and 305 participants on the chimeric faces.
 
@@
 Furthermore, we removed participants who scored less than 75% accuracy on the catch trials (trials in which there was no rhyme).
-These trimming procedures left data for `r nrow(LIs[LIs$exRDT== 0,])` participants on this task.
+These trimming procedures left data for 270 participants on this task.
 We considered recruiting further participants to achieve our target sample size of 300 participants.
 This seemed uneconomic as we could already see that with close to 300 cases, the test-retest reliability of the rhyme detection task was unsatisfactory.
-With 241 participants, the test-retest reliability was *r* = .63, which is smaller than the *r* = 0.65 that we pre-registered as our criterion for a reliable task.
+With 270 participants, the test-retest reliability was *r* = .62, which is smaller than the *r* = 0.65 that we pre-registered as our criterion for a reliable task.
 We plan to evaluate modified versions of this task in future but decided to stop data collection at this point and so we present here data from those participants with usable data.
@@
 In fact, for RT data we specified a *t*-score, but *t* and *z* converge when sample size exceeds 30, and for simplicity, we refer to these as LI_z scores.
-These LI_z scores are very highly correlated with the traditional LI measures (correlations typically exceed .95) but can be used to provide a common metric for RT and accuracy measures when identifying significantly lateralised individuals.
+For accuracy data, these LI_z scores are very highly correlated with the traditional LI measures (correlations typically exceed .95) but can be used to provide a common metric for RT and accuracy measures when identifying significantly lateralised individuals.
 For example, using a conventional 2-sided alpha of .05, we can categorise individuals with an absolute LI_z greater than 1.96 as significantly lateralised.
@@
 To obtain LI_z we ran a *t*-test comparing accurate log RTs for left- and right-sided stimuli for each person.
+In this case, LI_z will depend on an individual's RT variability as well as the difference in means between the two sides, and so the correlation with the conventional LI, though positive, will be less strong than for LI_z scores based on accuracy.
 
@@
 
-We examined the relationship between LIs on the rhyme decision, dichotic listening, and chimeric faces tasks by modelling covariances using a simplified version of Structural Equation Modeling (SEM) that used the sum of LIs from the two sessions of each task to represent each of three factors, minimising the impact of measurement error.
+We examined the relationship between LIs on the rhyme decision, dichotic listening, and chimeric faces tasks by modeling covariances using a simplified version of Structural Equation Modeling (SEM) that used the sum of LIs from the two sessions of each task to represent each of three factors, minimising the impact of measurement error.
 In this approach, we constrain particular covariance patterns and report ‘best’ model fit according to AIC weights (Wagenmakers & Farrell, 2004).
 Akaike weights are transformations of the AIC values, which can be directly interpreted as conditional probabilities for each model, with the best-fitting model being selected.
-We  compared three models:
-
--	Model A, where the two language tasks are correlated with each other but not with the chimeric faces
-
--	Model B, where all LIs are correlated, and
-
--	Model C, a base model where the LIs are unrelated.
-
-This was completed separately for right- and left-handers, as well as for the whole combined group.
+We compared three models: Model 1F, where all LIs are correlated, Model 2F, where the two language tasks are correlated with each other but not with the chimeric faces, and Model 3F, a base model where the LIs are unrelated.
+This analysis was conducted using data for both left- and right-handers.
 
@@
 colnames(LItable)<-c('Task','N','Mean','SD','One sample v','One sample p','Normality p')
-LItable[,1]<-c('RD 1','RD 2','Dl 1','DL 2','CF 1','CF 2','FT 1','FT 2')
+LItable[,1]<-c('RD 1','RD 2','DL 1','DL 2','CF 1','CF 2','FT 1','FT 2')
 
@@
 If there were no population bias, we would expect around 2.5% of individuals to meet each of these criteria.
-It can be seen that for dichotic listening and chimeric faces, around 45% of individuals have hemispheric dominance in the predicted direction and around 9% have hemispheric dominance in the opposite direction.
-The proportions showing significant laterality are lower for the rhyme decision task (around 17% left-hemisphere dominant, and 5% right-hemisphere dominant), and for finger tapping, the proportions are only slightly greater than would be expected by chance, although there is a suggestion of increasing laterality on that task across the two sessions.
+It can be seen that for dichotic listening and chimeric faces, around 47% of individuals have hemispheric dominance in the predicted direction and around 7% have hemispheric dominance in the opposite direction.
+The proportions showing significant laterality are lower for the rhyme decision task (around 19% left-hemisphere dominant, and 4% right-hemisphere dominant), and for finger tapping, the proportions are only slightly greater than would be expected by chance, although there is a suggestion of increasing laterality on that task across the two sessions.
 Note, however, that we had oversampled left-handers for this study, and, as is evident from exploratory analyses presented below, finger-tapping was strongly influenced by handedness.
@@
 title= "(A)"
-N=241
+N=254
 mycors<-densplots(RD.dat, time1,time2,name1,name2,scattername, y1, y2, x1, x2, ax, ay, remove, title, N) 
@@
 title="(B)"
-N=290
+N=287
 mycors<-densplots(DL.dat, time1,time2,name1,name2,scattername,y1, y2, x1, x2, ax, ay, remove,title,N)
@@
 title="(C)"
-N=289
+N=287
 mycors<-densplots(CF.dat, time1,time2,name1,name2,scattername, y1, y2, x1, x2, ax, ay, remove,title,N)
@@
 title="(D)"
-N=288
+N=309
 mycors<-densplots(FT.dat, time1,time2,name1,name2,scattername, y1, y2, x1, x2, ax, ay, remove,title,N)
@@
 
-The rhyme decision task has few individuals who are significantly lateralised in either direction.
+The rhyme decision task has relatively few individuals who are significantly lateralised in either direction.
 We will therefore not consider it further, as our analyses are only justified if the task is an indicator of cerebral lateralisation.
@@
 Our final question was whether the laterality indices from the language and faces laterality tasks correlated.
-Specifically, we had predicted that in right-handers the two language tasks would correlate together, but the chimeric faces would form a separate factor.
-In left-handers, we predicted that the language tasks would fractionate so that none of the three tasks would correlate.
+While we had originally planned to examine the relationship between LI_z score from different measures separately for left- and right-handers, we report these analyses for a combined group of all participants who did not have missing data (69 left- and 146 right-handers).
+Our departure from pre-registered analysis followed reviewers concerns about limited variation and restricted ranges of LI_z scores when looking at intercorrelations within handedness subgroups.
+Furthermore, examining the relationship between LI_z scores for the sample as a whole protects against the instability of results from our planned modeling of covariances, which with small sample sizes could favour different models when quite minor changes were made to the sample.
 
@@
 right_dat <- subset(interRR, handedness== "Right")
-left_dat<-left_dat[,1:6]
-right_dat<-right_dat[,1:6]
+#left_dat<-left_dat[,1:6]
+#right_dat<-right_dat[,1:6]
 totcorr<-correlate(interRR[,1:6],method='spearman')
-Rcorr <- correlate(right_dat,method='spearman')
-Lcorr <- correlate(left_dat,method='spearman')
+#Rcorr <- correlate(right_dat,method='spearman')
+#Lcorr <- correlate(left_dat,method='spearman')
 
 #Change so we have one matrix with R handers above diagonal and L handers below
-Lcorr<-as.data.frame(Lcorr[,2:7])
-Rcorr <- as.data.frame(Rcorr[,2:7])
-allcorr<-Lcorr
-for (i in 1:6){
-  for (j in i:6){
-    allcorr[i,j]<-Rcorr[i,j]
-
-  }
-}
+#Lcorr<-as.data.frame(Lcorr[,2:7])
+#Rcorr <- as.data.frame(Rcorr[,2:7])
+#allcorr<-Lcorr
+#for (i in 1:6){
+#  for (j in i:6){
+#    allcorr[i,j]<-Rcorr[i,j]
+#  }
+#}
+allcorr <- totcorr[-c(1)]
 allcorr<- round(allcorr,3)
@@
 
-Table 6.
-Spearman correlations between the LI_z scores for the different tasks in two Sessions.
-The upper diagonal shows correlations for the `r NR` right-handers (black font), and the values of the lower diagonal show for the `r NL` left-handers (red font).
+Table 6. Spearman correlations between the LI_z scores for the different tasks in two Sessions.
 ```{r tab 5, echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
@@
 
-myft5<-color(myft5, i = c(2,3,4,5, 6), j = "RD 1", color= "red")
-myft5<-color(myft5, i = c(3,4,5, 6), j = "RD 2", color= "red")
-myft5<-color(myft5, i = c(4,5, 6), j = "DL 1", color= "red")
-myft5<-color(myft5, i = c(5, 6), j = "DL 2", color= "red")
-myft5<-color(myft5, i = c(6), j = "CF 1", color= "red")
+#myft5<-color(myft5, i = c(2,3,4,5, 6), j = "RD 1", color= "red")
+#myft5<-color(myft5, i = c(3,4,5, 6), j = "RD 2", color= "red")
+#myft5<-color(myft5, i = c(4,5, 6), j = "DL 1", color= "red")
+#myft5<-color(myft5, i = c(5, 6), j = "DL 2", color= "red")
+#myft5<-color(myft5, i = c(6), j = "CF 1", color= "red")
 
@@
 ```
-*Note*.
-For right-handers, an absolute value of .16 is significant at the .05 level, and a value of .22 is significant at the .01 level.
-For left-handers, a value of .25 is significant at the .05 level and a value of .32 is significant at the .01 level.
-
-In a more formal test, we compared three models: (a) where the two language tasks are correlated with each other but not with the chimeric faces (i.e.
-a language laterality factor); (b) all LIs are correlated (i.e.
-a single laterality factor); and (c) a base model where the LIs are unrelated (i.e.
+
+In a more formal test, we compared three models: (2F) where the two language tasks are correlated with each other but not with the chimeric faces (i.e.
+a language laterality factor); (1F) all LIs are correlated (i.e.
+a single laterality factor); and (3F) a base model where the LIs are unrelated (i.e.
 independent functions).
-We departed from our pregistered analysis in two respects on the basis of reviewer recommendations.
-First we show results for the full sample as well as for separate handedness groups; second we adopt an alternative approach to the Akaike weights analysis that takes measurement error into account by creating factor scores based on the summed scores for the two test sessions.
+We departed from our pre-registered analysis based on reviewer recommendations.
+First, we show results for the full sample rather than separately for left- and right-handers.
+Second, we modified the Akaike weights analysis to remove the effect of measurement error by using factor scores based on the summed scores computed separately for odd and even trials (i.e.
+split-half) from the two test sessions.
+The analysis is thus conducted on factor scores with four indicators for each factor.
 
 ```{r prepareKievit, echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
+#NB for logic of taking factor scores to remove measurement error, see: https://blog.efpsa.org/2015/12/14/structural-equation-modeling-what-is-it-what-does-it-have-in-common-with-hippie-music-and-why-does-it-eat-cake-to-get-rid-of-measurement-error/
+
+#Make a table from the saved split half values
+
+dat1<-read.csv("./data/LIzs1.csv",stringsAsFactors=F)
+dat2<-read.csv("./data/LIzs2.csv",stringsAsFactors=F)
+
+colnames(dat1)<-c('subject','handedness','faces11','faces21','dichotic11','dichotic21',
+                     'rhyme11','rhyme21')
+
+colnames(dat2)<-c('subject','handedness','faces12','faces22','dichotic12','dichotic22',
+                     'rhyme12','rhyme22')
+halfdat<-cbind(dat1,dat2[,3:8])
+halfdatall<-halfdat[complete.cases(halfdat),]
+
+#reorganise so cols for same variable are together
+halfdatall<-halfdatall[,c(2:4,9:10,5:6,11:12,7:8,13:14)]
+cor(halfdatall)
+
+#identify influential data points
+#take means for each measure
+mean1<-rowMeans(halfdatall[,2:5])
+mean2<-rowMeans(halfdatall[,6:9])
+mean3<-rowMeans(halfdatall[,10:13])
+allfac<-cbind(mean1,mean2,mean3)
+cor(allfac)
+
+seeplots<-0 #toggle to 1 to see pairwise plots
+if(seeplots==1){
+#visualise scattergraphs for outliers
+pairs(allfac[halfdatall$handedness==1,]) #plot of factors for R handers
+pairs(allfac[halfdatall$handedness==2,])#plot of factors for L handers
+pairs(allfac) #plot of allfac for all
+}
+
+#identify influential points with Cook's distance
+#lm12 <- lm(mean1~mean2)
+#cd12<-cooks.distance(lm12)
+#lm23 <- lm(mean2~mean3)
+#cd23<-cooks.distance(lm23)
+#lm13 <- lm(mean1~mean3)
+#cd13<-cooks.distance(lm13)
+#cdout<-c(which(cd12>.03),which(cd13>.03),which(cd23>.03))
+#cdout<-unique(cdout)
+#halfdatall$cookout<-0
+#halfdatall$cookout[cdout]<-1
+
+#remove cases with high Cook's distance
+#censorz<-halfdatall[halfdatall$cookout==0,]
+
+#nremoved<-nrow(halfdatall)-nrow(censorz)
+
+```
+
+```{r fitAikakefactor,echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
 library(lavaan)
@@
 library(qpcR)
-# created handedness data
+library(flextable)
+#New version of Akaike test combines 2 test scores for each into a factor to deal with measurement error
 # now create subset of data frame for analysis
-interRR <- dplyr::select(LIs, split.RDT.1.1, split.RDT.1.2, split.RDT.2.1, split.RDT.2.2,
-                              split.DL.1.1, split.DL.1.2, split.DL.2.1, split.DL.2.2,
-                              split.chim.1.1, split.chim.1.2, split.chim.2.1, split.chim.2.2, handedness)
+interRR <- halfdatall
 # rename, and reverse sign for CF
-interRR <- interRR %>% 
-  rename(
-    "rhyme11" = split.RDT.1.1,
-    "faces11" = split.chim.1.1,
-    "dichotic11" = split.DL.1.1,
-     "rhyme12" = split.RDT.1.2,
-    "faces12" = split.chim.1.2,
-    "dichotic12" = split.DL.1.2,
-      "rhyme21" = split.RDT.2.1,
-    "faces21" = split.chim.2.1,
-    "dichotic21" = split.DL.2.1,
-     "rhyme22" = split.RDT.2.2,
-    "faces22" = split.chim.2.2,
-    "dichotic22" = split.DL.2.2
-    )
-
-#remove cases with missing values
-interRR<-na.omit(interRR)
-interRR$faces11 <- (-interRR$faces11) 
+interRR$faces11 <- (-interRR$faces11) #reverse sign for faces
 interRR$faces12 <- (-interRR$faces12)
-interRR$faces21 <- (-interRR$faces21) 
+interRR$faces21 <- (-interRR$faces21) #reverse sign for faces
 interRR$faces22 <- (-interRR$faces22)
 # subset into right and left handers
-left_dat <- subset(interRR, handedness== "Left")
-right_dat <- subset(interRR, handedness== "Right")
-
-```
-
-As shown in Table 7, the Akaike weights favoured model B for both left-handers and right-handers, as well as for both groups combined, confirming that there is association for laterality indices for the two language tasks, but not between language latearlity and laterality for face processing of emotion.
-
-Table 7. Akaike weights for the language model, independent base model, and laterality model.
-```{r disattenuateR, ,echo=FALSE, comment=FALSE, warning=FALSE}
-# First check range of scores on all variables for left and right-handers to address point re restriction of range
-rangedf <- data.frame(matrix(NA,nrow=3,ncol=4))
-colnames(rangedf)<-c('Llower','Lupper','Rlower','Rupper')
-rownames(rangedf)<-c('Rhyme','Dichotic','Chimeric')
-thisrow<-0
-for (myc in c(1,3,5)){
-  thisrow<-thisrow+1
-  rangedf[thisrow,1:2]<-range(left_dat[myc:(myc+1)])
-  rangedf[thisrow,3:4]<-range(right_dat[myc:(myc+1)])
-}
-
-#Now for combined handedness group, compute disattenuated correlations
-#We already have computed totcorr as symmetric matrix
-totcorr<-as.data.frame(totcorr)
-r.rhyme<-totcorr[2,2]
-r.dich<-totcorr[4,4]
-r.chim<-totcorr[6,6]
-
-nutot<-totcorr
-for (c in 2:3){ #col 1 is name of task
-  for (r in 3:4){
-    nutot[r,c]<-totcorr[r,c]/sqrt(r.rhyme*r.dich)
-  }
-    for (r in 5:6){
-    nutot[r,c]<-totcorr[r,c]/sqrt(r.rhyme*r.chim)
-  }
-}
-for (c in 4:5){
-  for (r in 5:6){
-    nutot[r,c]<-totcorr[r,c]/sqrt(r.dich*r.chim)
-  }
-}
-
-#nutot shows disattenuated correlations below diagonal
-```
-
-```{r newAkaike_factor,echo=FALSE, comment=FALSE, warning=FALSE}
-#New version of Akaike test combines 2 test scores for each into a factor to deal with measurement error
-
+left_dat <- subset(interRR, handedness== 2)
+right_dat <- subset(interRR, handedness== 1)
+
+#NB in pre-reg we just had time 1 data, but here we use time 1 and 2
 #Model 1:all correlated
 model1F<-"
-    rhymeFACTOR =~ NA*rhyme11 + rhyme12 + rhyme21 + rhyme 22
-    dichoticFACTOR =~ NA*dichotic11 + dichotic12 + dichotic21 + dichotic22
-    facesFACTOR =~ NA*faces11 +faces12 + faces21 +faces22
+rhymeFACTOR =~ NA*rhyme11 + rhyme12+rhyme21 + rhyme22
+dichoticFACTOR =~ NA*dichotic11 + dichotic12+dichotic21 + dichotic22
+facesFACTOR =~ NA*faces11 +faces12+faces21 +faces22
 
@@
 model2F<-"
-    rhymeFACTOR =~ NA*rhyme11 + rhyme12 + rhyme21 + rhyme 22
-    dichoticFACTOR =~ NA*dichotic11 + dichotic12 + dichotic21 + dichotic22
-    facesFACTOR =~ NA*faces11 +faces12 + faces21 +faces22
+rhymeFACTOR =~ NA*rhyme11 + rhyme12+rhyme21 + rhyme22
+dichoticFACTOR =~ NA*dichotic11 + dichotic12+dichotic21 + dichotic22
+facesFACTOR =~ NA*faces11 +faces12+faces21 +faces22
 
@@
 
-    rhymeFACTOR =~ NA*rhyme11 + rhyme12 + rhyme21 + rhyme 22
-    dichoticFACTOR =~ NA*dichotic11 + dichotic12 + dichotic21 + dichotic22
-    facesFACTOR =~ NA*faces11 +faces12 + faces21 +faces22
+rhymeFACTOR =~ NA*rhyme11 + rhyme12+rhyme21 + rhyme22
+dichoticFACTOR =~ NA*dichotic11 + dichotic12+dichotic21 + dichotic22
+facesFACTOR =~ NA*faces11 +faces12+faces21 +faces22
 
@@
 }
+
 myft6 <- flextable(AICdf, col_keys = c("Handedness",
@@
 autofit(myft6, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
-```
+
+library(semPlot)
+layout(t(1:3))
+semPaths(fit1A,intercepts=FALSE,curvePivot=TRUE, 
+         nodeLabels= c("RD 1.1", "RD 1.2", "RD 2.1", "RD 2.2", "DL 1.1", "DL 1.2", "DL 2.1", "DL 2.2", 
+                       "CF 1.1", "CF 1.2", "CF 2.1", "CF 2.2", "RD", "DL", "CF"), edge.color="black", sizeMan=6)
+semPaths(fit2A,intercepts=FALSE,curvePivot=TRUE, 
+         nodeLabels= c("RD 1.1", "RD 1.2", "RD 2.1", "RD 2.2", "DL 1.1", "DL 1.2", "DL 2.1", "DL 2.2", 
+                       "CF 1.1", "CF 1.2", "CF 2.1", "CF 2.2", "RD", "DL", "CF"), edge.color="black", sizeMan=6)
+semPaths(fit3A,intercepts=FALSE,curvePivot=TRUE, 
+         nodeLabels= c("RD 1.1", "RD 1.2", "RD 2.1", "RD 2.2", "DL 1.1", "DL 1.2", "DL 2.1", "DL 2.2", 
+                       "CF 1.1", "CF 1.2", "CF 2.1", "CF 2.2", "RD", "DL", "CF"), edge.color="black", sizeMan=6)
+```
+
+The Akaike weight for the 1F, 2F, and 3F models were `r round(AICdf[1,2], 3)`, `r round(AICdf[1,3],3)`, and `r round(AICdf[1,4], 3)` respectively.
+Thus the Akaike weights favoured model 2F confirming that there is association for laterality indices for the two language tasks, but not between language laterality and laterality for face processing of emotion.
+
+For completeness, Supplementary Material Table S1 shows the original pre-registered Akaike weights analysis, which did not show a benefit for the two-factor model, but rather gave best fit to a model where there were distinct factors for each laterality test.
+We regard this model as misleading because it incorporates measurement error, which was particularly large for the rhyme detection task, and so attenuates the relationship between variables.
 
@@
 However, as the rhyme task was less reliable than other measures, and showed only small asymmetries, we chose to examine the relationship between the LexTALE and the other verbal laterality measure, dichotic listening, instead.
-For these exploratory analyses we included participants scoring less than 80 on the LexTALE in order to reduce the effect of a restricted range on the correlational analyses.
+For these exploratory analyses we included participants scoring less than 80 on the LexTALE to reduce the effect of a restricted range on the correlational analyses.
 There was no correlation between the LexTALE score and LIs on the dichotic listening at session 1, *r*= `r round(DL.lex.est$estimate, 2)`, *95% CI*= [`r round(DL.lex$conf.int[1], 2)`, `r round(DL.lex$conf.int[2], 2)`], *p*= `r round(p.adjust(DL.lex.est$p.value[1], n= 2, method= "bonferroni"), 3)`, or session 2, *r*= `r round(DL.lex.est1$estimate, 2)`, *95% CI*= [`r round(DL.lex1$conf.int[1], 2)`, `r round(DL.lex1$conf.int[2], 2)`], *p*= `r round(p.adjust(DL.lex.est1$p.value[1], n= 2, method= "bonferroni"), 3)`.
@@
 This pattern of results indicates that while very few participants showed statistically better performance for images in the right visual field (i.e.
-LI_z showed scores above |1.96|), many showed more subtle RVF advantages and had corresponding LI_z scores that were not statistically different from zero at the individual level.
-When comparing LI_z scores to zero at the population there will have been sufficient power to determine whether these small differences were different from zero.
+LI_z scores above |1.96|), many showed more subtle RVF advantages that did not exceed the threshold of statistical significance.
+When comparing LI_z scores to zero at the population level there will have been sufficient power to determine whether these small differences were reliably different from zero.
 This task also yielded the poorest test-retest reliability, just falling short of our target minimum cutoff of *r* = .65.
@@
 Knecht et al., 2000), and requiring speech production should yield a hemispheric advantage.
-Our task was designed to require covert speech, but an active speech production component might reveal stronger lateralisation.
-This would, however, be logistically difficult to implement online.
-The speech requirement may not, however, be a key factor, given that recent studies have reported robust right visual-field advantages on a lexical decision task that does not require active speech production (Brederoo et al., 2020; Hausmann et al., 2019; Willemin et al., 2019).
+Our task was designed to require covert speech, raising the question of whether an active speech production component might reveal stronger lateralisation.
+However, this explanation seems unlikely, given that recent studies have reported robust right visual-field advantages on a lexical decision task that does not require active speech production (Brederoo et al., 2020; Hausmann et al., 2019; Willemin et al., 2019).
 
@@
 Outside of the lab, participants can freely move their head which will influence the size of images on the retina, and it is difficult to ensure that images are shown exclusively in parafoveal vision.
-If on a subset of images are not equally displaced in parafoveal vision it could mean that processing is more efficient for the image in its corresponding visual field, reducing the validity of the task and adding considerable noise to response time estimates in each visual field.
-Without stict monitoring of eye-movements here it is not possible to control for this.
-In future work we plan to explore the use of images presented in peripheral vision, as it is unlikely that small head movements would result in the image shifting from peripheral vision.
+If a subset of images are not equally displaced in parafoveal vision it could mean that processing is more efficient for the image in its corresponding visual field, reducing the validity of the task and adding considerable noise to response time estimates in each visual field.
+Without strict monitoring of eye-movements here it is not possible to control for this.
+In future work, we plan to explore the use of images presented in peripheral vision, as it is unlikely that small head movements would result in the image shifting from peripheral vision.
 This would increase difficultly in the processing of information in this region, which may also result in stronger visual field advantages.
@@
 The task has previously been shown to have high split-half reliability (Levy et al., 1983), we show that it can meet the higher standard of achieving a high test-retest correlation.
-Our subsequent analysis showed that acceptable levels of reliability can be achieved with this task using very few trials: reliability did not suffer when only 35 trials were used, and test-retest reliability of *r*= .65 could be found with 25 trials.
+Our subsequent analysis showed that acceptable levels of reliability can be achieved with this task using very few trials: reliability did not suffer when only 36 trials were used, and test-retest reliability of *r*= .65 could be found with 25 trials.
 It seems likely that one reason for the success of the chimeric faces task online is the task’s simplistic nature.
@@
 Previous studies have reported correlations between hand preference and measures of relative hand skill, but different hand skill measures typically are weakly intercorrelated (Bishop, 1990), and it seems that asymmetries may vary depending on particular aspects of planning and execution of motor coordination that is involved.
-It would be of interest to see how far LIs resemble that from tasks such as the pegboard task (e.g.
+It would be of interest to compared LIs on this task with those from the pegboard task (e.g.
 Hodgson, Hirst, & Hudson, 2016), which is widely used in neuropsychological assessment.
@@
 However, several studies have reported no such association, or only a weak relationship (Bradzakova-Trajkov et al., 2010; Bryden et al.,1983; Rosch et al., 2012; Whitehouse & Bishop, 2009; van der Haegen & Brysbaert, 2018).
-Data from the current work supports a view in which laterality on each task is independent.
+Data from the current work supports a view in which laterality on verbal and nonverbal tasks is independent.
 Not only was our study well powered to examine this relationship, our sample included a large number of left-handers who showed increased variability in their LIs.
 Specifically, we found no association between the LI_z indices for the two language tasks and chimeric faces.
-There was a weak but measureable association between the two language tasks, rhyme decision and dichotic listening, found in both handedness groups.
-This is of particular interest given the different  demands of the two tasks - speech perception using dichotic presentation for the dichotic listening task, vs covert naming of a picture and comparison to the phonology of a written word in the rhyme judgement task.
+There was a weak but measurable association between the two language tasks, rhyme decision and dichotic listening.
+This is of particular interest given the different demands of the two tasks: speech perception using dichotic presentation for the dichotic listening task, vs covert naming of a picture and comparison to the phonology of a written word in the rhyme judgement task.
 The current result suggests there could be interest in replicating this comparison of the two tasks using a visual-half-field language task that has better reliability and validity.
-The lack of association between dichotic listening and chimeric faces is, however, much more convincing, given that both tasks show strong lateralisation at the population level and high test-retest reliability.
+We found no support for our prediction that dissociation between language tasks would be greater for left- than right-handers: the trend was for lower correlations between language tasks in the right-handers.
+
+The lack of association between dichotic listening and chimeric faces is convincing, given that both tasks show strong lateralisation at the population level and high test-retest reliability.
 This is consistent with the view that, at least for some functions, lateralisation is determined by independent biases: the left-hemisphere population bias for language is unrelated to the right-hemisphere bias for facial expressions of emotion.
@@
 (2020) confirms that lateralisation of different functions is associated far more than expected by chance, yet cases of dissociation are not uncommon.
-Research in future might explore which functions are associated, and to what extent, rather than making a simple contrast between models that postulate a general answer to explain all of laterality.
 
@@
 In their study, participants completed a variety of visual half-field tasks covering a range of domains such as word processing, face processing, and spatial frequency processing.
-At the population level Brederoo et al.
+At the population level, Brederoo et al.
 reported lateralisation in the expected direction for the majority of tasks.
-For left handers and those with right hemisphere dominant speech, there was evidence of a deviation from typical lateralisation.
+For left-handers and those with right hemisphere dominant speech, there was evidence of a deviation from typical lateralisation.
 Contrary to the current study, Brederoo et al.
 reported partial evidence to suggest that processes recruiting similar resources will lateralise to homologue areas to promote intra-hemispheric proximity to other cortical areas within their processing hierarchies.
-In one analysis they found the lateralisation of faces and word recognition to be correlated but when they looked at lateralisation in those who were right handed versus those with right dominant language processing there was considerably less evidence for causal complementarity.
+In one analysis they found the lateralisation of faces and word recognition to be correlated but when they looked at lateralisation in those who were right-handed versus those with right dominant language processing there was considerably less evidence for causal complementarity.
 However, Brederoo et al.
-reported consistent evidence within the visual domain, where lateralisation of low and high frequency information processing was associated, but independent of   lateralisation on other tasks.
+reported consistent evidence within the visual domain, where lateralisation of low- and high-frequency information processing was associated, but independent of lateralisation on other tasks.
 So it would appear by comparing tasks of a wider variety and considering laterality across domains, we may elucidate the issue of colateralisation and identify how far individual differences in lateralisation depend on task demands and shared processing resources.
+Complementarity may be specific to functions that engage homologous areas in the left and right hemispheres.
+As such, research in future might explore which functions are associated, and to what extent, rather than making a simple contrast between models that postulate a general answer to explain all of laterality.
 
@@
 It is important to note that this exclusion was not possible in the current study as we did not include a hearing test.
-With the current sample size, subtle hearing deficits in one ear will unlikely undermine the results.
+With the current sample size, subtle hearing deficits in one ear will be unlikely to undermine the results.
 However, it will be important for future studies to incorporate hearing accuracy tests in order to exclude participants with these hearing related difficulties.
@@
 In these Supplementary Materials we include the distribution of laterality indices for each behavioural laterality task across both testing sessions.
-We also include a series of t-tests examining the difference in LI_z scores between extreme left-handers and right-handers.
+We also include a series of non-parametric tests examining the difference in LI_z scores between extreme left-handers and right-handers and our pre-registered modeling of covariances between laterality indices.
 
@@
 ```{r hand.corrs,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
-comp_c <- ggplot(subset(LIs, handed3== "Right-hander" | handed3=="Extreme left-hander"),
-  aes(x= handed3, y= Day1_CF_acc_Z, fill= handed3)) +
+LIs$handed4 <- ifelse(LIs$handed3== "Right-hander", "RH", ifelse(LIs$handed3== "Extreme left-hander", "Extreme LH", "Mod"))
+comp_c <- ggplot(subset(LIs, handed4== "RH" | handed4=="Extreme LH"),
+  aes(x= handed4, y= Day1_CF_acc_Z, fill= handed4)) +
   geom_hline(yintercept=0) + 
@@
   geom_violin(alpha= .1, size= .5) +
-  geom_jitter(width= .1, aes(color= handed3), alpha= .25) +
+  geom_jitter(width= .1, aes(color= handed4), alpha= .25) +
   geom_boxplot(width=0.1, position=position_dodge(0.9), size= .5 , notch = T, notchwidth = 0.4, varwidth =F,
@@
   theme(legend.position= "none") + theme(plot.title = element_text(hjust = 0.5))
-comp_b <- ggplot(subset(LIs, handed3== "Right-hander" | handed3=="Extreme left-hander"),
-  aes(x= handed3, y= Day1_Dich_acc_Z, fill= handed3)) +
+comp_b <- ggplot(subset(LIs, handed4== "RH" | handed4=="Extreme LH"),
+  aes(x= handed4, y= Day1_Dich_acc_Z, fill= handed4)) +
   geom_hline(yintercept=0) + 
@@
   theme(legend.position= "none") + theme(plot.title = element_text(hjust = 0.5))
-comp_a <- ggplot(subset(LIs, handed3== "Right-hander" | handed3=="Extreme left-hander"),
-  aes(x= handed3, y= tlat_RDT_1, fill= handed3)) +
+comp_a <- ggplot(subset(LIs, handed4== "RH" | handed4=="Extreme LH"),
+  aes(x= handed4, y= tlat_RDT_1, fill= handed4)) +
   geom_hline(yintercept=0) + 
@@
   theme(legend.position= "none") + theme(plot.title = element_text(hjust = 0.5))
-t.plots <- ggpubr::ggarrange(comp_a, comp_b, comp_c, ncol = 1, nrow= 3)
-ggsave("./plots/t.plots.png", plot=t.plots,dpi = 300, width = 5, height = 15,)
+t.plots <- ggpubr::ggarrange(comp_a, comp_b, comp_c, ncol = 3, nrow= 1)
+ggsave("./plots/t.plots.png", plot=t.plots,dpi = 300, width = 12, height = 6,)
 # tests
@@
 This is not surprising given that very few participants were lateralised on this task.
-While LI_z scores did not differ between extreme left-handers and right-handers on the rhyme decision, they did on the dichotic listening, *U*= `r round(t2$statistic, 2)`, *95% CI*= [`r round(t2$conf.int[1], 2)`, `r round(t2$conf.int[2], 2)`], *p*= `r round(t2$p.value, 3)`, and chimeric faces, *U*= `r round(t3$statistic, 2)`, *95% CI*= [`r round(t3$conf.int[1], 2)`, `r round(t3$conf.int[2], 2)`], *p*= `r round(t3$p.value, 3)`, tasks.
+While LI_z scores did not differ between extreme left-handers and right-handers on the rhyme decision, they did on the dichotic listening, *U*= `r round(t2$statistic, 2)`, *95% CI*= [`r round(t2$conf.int[1], 2)`, `r round(t2$conf.int[2], 2)`], *p*= `r round(t2$p.value, 3)`, and chimeric faces, *U*= `r round(t3$statistic, 2)`, *95% CI*= [`r round(t3$conf.int[1], 2)`, `r round(t3$conf.int[2], 2)`], *p*< 0.001, tasks.
 These differences demonstrate that, as a population, extreme left-handers were less lateralised.
@@
 The red dashed lines are drawn at -1.96 and 1.96, indicating significant lateralisation to either the left or right hemisphere.
-Data points represent indiciaul participants' data.](./plots/t.plots.png){width=40%}
+Data points represent individual participants' data.](./plots/t.plots.png){width=60%}
+
+**Modeling of covariances**
+
+The modeling of covariances reported in our manuscript departed from our pre-registration.
+We originally specified that we would fit separate models for right- and left-handers, using a subset of SEM where we constrain particular covariance patterns and report ‘best’ model fit according to Akaike weights.
+Here we report these results.
+
+As shown in Table S1, the Akaike weights favoured model 1F for both left-handers and right-handers, indicating that lateralisation across tasks is largely independent.
+This reflects a potential lack of cross-task correlations for LI_z measures.
+However, for reasons stated in the manuscript, these results are somewhat misleading and should be approached with caution.
+
+```{r prepareKievitS1, echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
+# created handedness data
+# now create subset of data frame for analysis
+interRR <- dplyr::select(LIs, zlat_RDT_1, zlat_RDT_2, Day1_Dich_acc_Z,
+                         Day2_Dich_acc_Z, Day1_CF_acc_Z, Day2_CF_acc_Z,handedness)
+# rename, and reverse sign for CF
+interRR <- interRR %>% 
+  rename(
+    "rhyme1" = zlat_RDT_1,
+    "faces1" = Day1_CF_acc_Z,
+    "dichotic1" = Day1_Dich_acc_Z,
+     "rhyme2" = zlat_RDT_2,
+    "faces2" = Day2_CF_acc_Z,
+    "dichotic2" = Day2_Dich_acc_Z
+    )
+
+#remove cases with missing values
+interRR<-na.omit(interRR)
+interRR$faces1 <- (-interRR$faces1) #reverse sign for faces
+interRR$faces2 <- (-interRR$faces2)
+# subset into right and left handers
+left_dat <- subset(interRR, handedness== "Left")
+right_dat <- subset(interRR, handedness== "Right")
+
+#NB in pre-reg we just had time 1 data, but here we use time 1 and 2
+
+ #Model 1 all independent
+    model1<-"
+    rhyme1~~rhyme2
+    rhyme1~~0*dichotic1
+    rhyme1~~0*dichotic2
+     rhyme1~~0*faces1
+    rhyme1~~0*faces2
+    rhyme2~~0*dichotic1
+    rhyme2~~0*dichotic2
+     rhyme2~~0*faces1
+    rhyme2~~0*faces2
+    dichotic1~~dichotic2
+    dichotic1~~0*faces1
+    dichotic1~~0*faces2
+     dichotic2~~0*faces1
+    dichotic2~~0*faces2
+    faces1~~faces2
+    "
+#Model 2: language tasks are correlated
+    model2<-"
+  rhyme1~~rhyme2
+    rhyme1~~dichotic1
+    rhyme1~~dichotic2
+     rhyme1~~0*faces1
+    rhyme1~~0*faces2
+    rhyme2~~dichotic1
+    rhyme2~~dichotic2
+     rhyme2~~0*faces1
+    rhyme2~~0*faces2
+    dichotic1~~dichotic2
+    dichotic1~~0*faces1
+    dichotic1~~0*faces2
+     dichotic2~~0*faces1
+    dichotic2~~0*faces2
+    faces1~~faces2
+    "
+#Model 3:all correlated
+    model3<-"
+    rhyme1~~rhyme2
+    rhyme1~~dichotic1
+    rhyme1~~dichotic2
+     rhyme1~~faces1
+    rhyme1~~faces2
+    rhyme2~~dichotic1
+    rhyme2~~dichotic2
+     rhyme2~~faces1
+    rhyme2~~faces2
+    dichotic1~~dichotic2
+    dichotic1~~faces1
+    dichotic1~~faces2
+     dichotic2~~faces1
+    dichotic2~~faces2
+    faces1~~faces2
+    "
+#Make a little dataframe for AIC weights
+AICdf <- data.frame(matrix(NA,nrow=2,ncol=4))
+AICdf[,1]<-c('Right-handers','Left-handers')
+colnames(AICdf)<-c('Handedness','Model 1F','Model 2F','Model 3F')
+for (myhand in 1:2) {
+  mydat<-right_dat
+  if(myhand==2) {mydat<-left_dat}
+
+ #Using data from this handedness type, test all 3 models
+fit1A <- cfa(model1, data=mydat)
+fit2A <- cfa(model2, data=mydat)
+fit3A <- cfa(model3, data=mydat)
+    aic_vector1 <- c(fitMeasures(fit1A, "aic"),fitMeasures(fit2A, "aic"),fitMeasures(fit3A, "aic"))
+
+    #Confirms that for each model, lowest AIC obtained when the simulated data do match the model
+    # 
+    # The AIC weights make massive difference for preferred model!
+    AICweights<-akaike.weights(aic_vector1)$weights
+    # 
+ AICdf[myhand,2:4]<-round(AICweights,3)
+ }
+myftS1 <- flextable(AICdf, col_keys = c("Handedness",
+                                       "Model 1F",
+                                       "Model 2F", 
+                                       "Model 3F"))
+autofit(myftS1, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
+```
+
+Table S1. Akaike weights for the language model, independent base model, and laterality model when adopting our pre-registered analysis.
+```{r table S1, echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
+myftS1 <- flextable(AICdf, col_keys = c("Handedness",
+                                       "Model 1F",
+                                       "Model 2F", 
+                                       "Model 3F"))
+autofit(myftS1, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
+```
